### 2025-03-26 ms/response_letter.qmd
Initial commit for tree-allo

---
author: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  pdf:
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{xcolor}
        \usepackage{mdframed}
        % Define blockquote colors
        \definecolor{blockquote-border}{RGB}{221,221,221}
        \definecolor{blockquote-text}{RGB}{119,119,119}
        % Define custom blockquote environment
        \newmdenv[
          rightline=false,
          bottomline=false,
          topline=false,
          linewidth=3pt,
          linecolor=blockquote-border,
          skipabove=\parskip,
          innerleftmargin=10pt,
          innerrightmargin=0pt,
          frametitlebackgroundcolor=gray!20
        ]{customblockquote}
        \renewenvironment{quote}
          {\begin{customblockquote}\color{blockquote-text}\ignorespaces}
          {\end{customblockquote}}
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

Dear Editor,

We hereby resubmit the manuscript (JFR-D-25-00026), "Saturating allometric relationships reveal how wood density shapes global tree architecture".
We thank the handling editor and the reviewers for their careful attention to detail in our manuscript and for providing comments that have greatly improved the study.

Based on the comments, we have updated the manuscript to reflect the following:

1. **Improved content and clarity:**
We clarified the study’s main objective, added missing methodological details (e.g., wood density classification criteria and interpretation of $\Delta$LOOIC), and expanded the discussion to address potential confounding effects on allometric relationships.

2. **Enhanced presentation and formatting:**
We standardized the reporting of parameters, corrected figure labels, restructured supplementary tables for better readability, and revised the manuscript language to improve clarity and consistency.

We believe these revisions have strengthened the manuscript by clarifying our objectives, improving consistency and readability, and addressing key methodological concerns—thereby better highlighting the significance of our findings.

Below we provide detailed responses.
The editing history is recorded in a separate pdf file.

Sincerely,

Masatoshi Katabuchi

On behalf of the authors: Thi Duyen Nguyen

Corresponding author and contact information:
Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Mengla, Yunnan 666303, China\
Email: katabuchi@xtbg.ac.cn; mattocci27@gmail.com

\newpage

# Responses

Please note that line numbers refer to ms_diff.pdf that contains editing history.
The word file has slightly different line numbers.

> Reviewer #3: The manuscript presents a comprehensive study on developing allometric models for estimating forest carbon biomass using global datasets and Bayesian hierarchical models. The authors highlighted the importance of saturating functions and wood density to improve allometric models without using two-step fitting approaches in previous studies (which seems problematic). Overall, the study provides valuable contributions to ecological theory and practical forestry applications. I have reviewed the entire manuscript very carefully and prepared below a detailed list of corrections and suggestions that will certainly improve the quality and make clearer the text.

Thank you very much for your positive and constructive feedback.

> 1) The technical information presented in the paper was interesting and worth reading.

Thank you for your kind remark.

> 2) Abstract is informative and it reflects the body of the paper.

We appreciate your positive feedback on the abstract.

> 3) The objectives must be modified for a better understanding of the reader.

Thank you for your suggestion.
We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line XXX).

> 4) Line 366-369 & Table 2: In the sentence "In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent b is near the expected *2/3* ($b_{ang}$=0.625775 0.6168636,0.6345041, Table 2)" and the parameter estimates in Table 2, the values are not consistently presented. Check whole of the manuscript for a uniform presentation of the values.

We have standardized the presentation of all parameter estimates throughout the manuscript, including Table 2 and Table S6-8 and corresponding text.

> 5) Table 1: The table does not clearly explain how much of a difference in ΔLOOIC is considered significant between the two models.

We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line XXX).

> 6) Figure S1: According to the supplementary materials section, the figure is incorrectly labeled Figure 1. It should be Figure S1.

We have corrected the figure label.

> 7) The reference section was informative. However, incorporate some latest references and arrange them according to the journal guidelines.

We have updated the references with recent publications, including xxx, and formatted them according to the journal guidelines.

> 8) Supplementary Table S6 is very congested and hard to read. Please revise for better understanding.

We have reformatted Supplementary Tables S6, S7, and S8 for improved readability and included detailed captions to aid interpretation.

> 9) A revision of the English is needed as some sentences aren't properly structured, and it's often difficult to grasp what the authors are saying.

We have revised the manuscript for language, grammar, and sentence structure to enhance clarity and improve overall readability.
For example, we revised lines XXX–YYY in the Discussion section to ensure sentence brevity and clarity.
Additionally, we updated the figure legends and table captions for greater clarity and carefully cross-checked all technical terms to maintain consistency throughout the text.

> Reviewer #5: Congratulations on a well-written and clear manuscript. Your approach to exploring allometric relationships between DBH, height, wood density, biomass, and crown radius is highly interesting and contributes significantly to the field. While I primarily focus on the data analysis section and found no major concerns, I would like to suggest a few minor corrections and discussion points.

Thank you so much for your encouraging and constructive feedback.

> 1) In Figure 3, the distinction between species with light and dense wood density is presented. However, the criteria used for this classification are not mentioned in the Materials and Methods section. Clarifying this would improve the transparency of your methodology.

Thank you for pointing this out.
We have added a description in the Methods section clarifying the classification criteria (line XXX).

> 2) Figure 4 appears to have a formatting issue, as the letters 'a' and 'd' are not displayed.

We have corrected the formatting issue in Figure 4.

> 3) It may be valuable to discuss the potential confounding effects of environmental factors on observed allometric correlations when considering wood density. Given intraspecific variation, wood density may not have a strong impact on those allometric relationships, since Fajardo et al. (2022) suggest that wood density is more closely linked to vessel anatomy than to tree architecture within the same species across different environments.

We agree with your point and have expanded the Discussion section to address potential environmental confounding effects.
We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line XXX).

> 4) I found the "Consequences for AGB Estimates" section particularly insightful.

Thank you for your positive feedback.

> 5) Pages 132-140 are blank, likely due to a formatting issue with the final table. The table appears to have fixed columns that prevent the complete display of cell information.

We have resolved the formatting issue that caused the blank pages in the supplementary file.

### 2025-03-26 ms/response_letter.tex
Initial commit for tree-allo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{xcolor}
\usepackage{mdframed}
% Define blockquote colors
\definecolor{blockquote-border}{RGB}{221,221,221}
\definecolor{blockquote-text}{RGB}{119,119,119}
% Define custom blockquote environment
\newmdenv[
  rightline=false,
  bottomline=false,
  topline=false,
  linewidth=3pt,
  linecolor=blockquote-border,
  skipabove=\parskip,
  innerleftmargin=10pt,
  innerrightmargin=0pt,
  frametitlebackgroundcolor=gray!20
]{customblockquote}
\renewenvironment{quote}
  {\begin{customblockquote}\color{blockquote-text}\ignorespaces}
  {\end{customblockquote}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{2025-03-25}

\begin{document}

Dear Editor,

We hereby resubmit the manuscript (JFR-D-25-00026), ``Saturating
allometric relationships reveal how wood density shapes global tree
architecture''. We thank the handling editor and the reviewers for their
careful attention to detail in our manuscript and for providing comments
that have greatly improved the study.

Based on the comments, we have updated the manuscript to reflect the
following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Improved content and clarity:} We clarified the study's main
  objective, added missing methodological details (e.g., wood density
  classification criteria and interpretation of \(\Delta\)LOOIC), and
  expanded the discussion to address potential confounding effects on
  allometric relationships.
\item
  \textbf{Enhanced presentation and formatting:} We standardized the
  reporting of parameters, corrected figure labels, restructured
  supplementary tables for better readability, and revised the
  manuscript language to improve clarity and consistency.
\end{enumerate}

We believe these revisions have strengthened the manuscript by
clarifying our objectives, improving consistency and readability, and
addressing key methodological concerns---thereby better highlighting the
significance of our findings.

Below we provide detailed responses. The editing history is recorded in
a separate pdf file.

Sincerely,

Masatoshi Katabuchi

On behalf of the authors: Thi Duyen Nguyen

Corresponding author and contact information: Key Laboratory of Tropical
Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy
of Sciences, Mengla, Yunnan 666303, China\\
Email: katabuchi@xtbg.ac.cn; mattocci27@gmail.com

\newpage

\section{Responses}\label{responses}

Please note that line numbers refer to ms\_diff.pdf that contains
editing history. The word file has slightly different line numbers.

\begin{quote}
Reviewer \#3: The manuscript presents a comprehensive study on
developing allometric models for estimating forest carbon biomass using
global datasets and Bayesian hierarchical models. The authors
highlighted the importance of saturating functions and wood density to
improve allometric models without using two-step fitting approaches in
previous studies (which seems problematic). Overall, the study provides
valuable contributions to ecological theory and practical forestry
applications. I have reviewed the entire manuscript very carefully and
prepared below a detailed list of corrections and suggestions that will
certainly improve the quality and make clearer the text.
\end{quote}

Thank you very much for your positive and constructive feedback.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  The technical information presented in the paper was interesting and
  worth reading.
\end{enumerate}
\end{quote}

Thank you for your kind remark.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  Abstract is informative and it reflects the body of the paper.
\end{enumerate}
\end{quote}

We appreciate your positive feedback on the abstract.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  The objectives must be modified for a better understanding of the
  reader.
\end{enumerate}
\end{quote}

Thank you for your suggestion. We have revised the Introduction to
clearly state the main objective of the study before presenting the
specific questions (line XXX).

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{3}
\tightlist
\item
  Line 366-369 \& Table 2: In the sentence ``In addition, the estimated
  exponent aligns closely with MST predictions for crown radius,
  supporting the selection of the power-law model for angiosperms, where
  the community-level exponent b is near the expected \emph{2/3}
  (\(b_{ang}\)=0.625775 0.6168636,0.6345041, Table 2)'' and the
  parameter estimates in Table 2, the values are not consistently
  presented. Check whole of the manuscript for a uniform presentation of
  the values.
\end{enumerate}
\end{quote}

We have standardized the presentation of all parameter estimates
throughout the manuscript, including Table 2 and Table S6-8 and
corresponding text.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{4}
\tightlist
\item
  Table 1: The table does not clearly explain how much of a difference
  in ΔLOOIC is considered significant between the two models.
\end{enumerate}
\end{quote}

We have revised the caption of Table 1 to clarify that a \(\Delta\)LOOIC
greater than 4 suggests a meaningful difference in predictive
performance (line XXX).

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{5}
\tightlist
\item
  Figure S1: According to the supplementary materials section, the
  figure is incorrectly labeled Figure 1. It should be Figure S1.
\end{enumerate}
\end{quote}

We have corrected the figure label.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{6}
\tightlist
\item
  The reference section was informative. However, incorporate some
  latest references and arrange them according to the journal
  guidelines.
\end{enumerate}
\end{quote}

We have updated the references with recent publications, including xxx,
and formatted them according to the journal guidelines.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{7}
\tightlist
\item
  Supplementary Table S6 is very congested and hard to read. Please
  revise for better understanding.
\end{enumerate}
\end{quote}

We have reformatted Supplementary Tables S6, S7, and S8 for improved
readability and included detailed captions to aid interpretation.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{8}
\tightlist
\item
  A revision of the English is needed as some sentences aren't properly
  structured, and it's often difficult to grasp what the authors are
  saying.
\end{enumerate}
\end{quote}

We have revised the manuscript for language, grammar, and sentence
structure to enhance clarity and improve overall readability.

\begin{quote}
Reviewer \#5: Congratulations on a well-written and clear manuscript.
Your approach to exploring allometric relationships between DBH, height,
wood density, biomass, and crown radius is highly interesting and
contributes significantly to the field. While I primarily focus on the
data analysis section and found no major concerns, I would like to
suggest a few minor corrections and discussion points.
\end{quote}

Thank you so much for your encouraging and constructive feedback.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  In Figure 3, the distinction between species with light and dense wood
  density is presented. However, the criteria used for this
  classification are not mentioned in the Materials and Methods section.
  Clarifying this would improve the transparency of your methodology.
\end{enumerate}
\end{quote}

Thank you for pointing this out. We have added a description in the
Methods section clarifying the classification criteria (line XXX).

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  Figure 4 appears to have a formatting issue, as the letters `a' and
  `d' are not displayed.
\end{enumerate}
\end{quote}

We have corrected the formatting issue in Figure 4.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  It may be valuable to discuss the potential confounding effects of
  environmental factors on observed allometric correlations when
  considering wood density. Given intraspecific variation, wood density
  may not have a strong impact on those allometric relationships, since
  Fajardo et al.~(2022) suggest that wood density is more closely linked
  to vessel anatomy than to tree architecture within the same species
  across different environments.
\end{enumerate}
\end{quote}

We agree with your point and have expanded the Discussion section to
address potential environmental confounding effects. We also cited
Fajardo et al.~(2022) to highlight less clear linkages between tree
architecture and intraspecific variation in wood density (line XXX).

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{3}
\tightlist
\item
  I found the ``Consequences for AGB Estimates'' section particularly
  insightful.
\end{enumerate}
\end{quote}

Thank you for your positive feedback.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{4}
\tightlist
\item
  Pages 132-140 are blank, likely due to a formatting issue with the
  final table. The table appears to have fixed columns that prevent the
  complete display of cell information.
\end{enumerate}
\end{quote}

We have resolved the formatting issue that caused the blank pages in the
supplementary file.

\end{document}

### 2025-03-26 docs/R2.qmd
Initial commit for tree-allo

---
title: "Caculating Bayesian R2"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r echo=FALSE}
library(targets)
library(tidyverse)
source("R/functions.R")
```

# Read in data

# Power-law

## H
```{r}
stan_data_lr_h <- tar_read(stan_data_lr_h)
str(stan_data_lr_h)

# Without wood density
pl_nou_h <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl_nou_h

gamma1_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[1]"])
gamma2_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[2]"])
sigma_draws_pl_nou_h <- as.matrix(pl_nou_h[, "sigma"])

num_sam_pl_nou_h <- nrow(gamma1_draws_pl_nou_h)
num_obs_pl_nou_h <- nrow(stan_data_lr_h$log_x)

log_y_pred_draws_pl_nou_h <- matrix(NA, nrow = num_sam_pl_nou_h, ncol = num_obs_pl_nou_h)

for (i in 1:num_sam_pl_nou_h) {
  log_y_pred_draws_pl_nou_h[i, ] <- gamma1_draws_pl_nou_h[i] + gamma2_draws_pl_nou_h[i] * stan_data_lr_h$log_x[, 2]
}

var_log_y_pred_pl_nou_h <- apply(log_y_pred_draws_pl_nou_h, 1, var)

var_res_pl_nou_h <- sigma_draws_pl_nou_h^2

bayes_R2_draws_pl_nou_h <- var_log_y_pred_pl_nou_h / (var_log_y_pred_pl_nou_h + var_res_pl_nou_h)

summary_bayes_R2_pl_nou_h <- summary(bayes_R2_draws_pl_nou_h)

summary_bayes_R2_pl_nou_h

# With wood density
pl_h <- tar_read(fit_lr2_draws_power_law1_h)
pl_h

gamma1_draws_pl_h <- as.matrix(pl_h[, "gamma[1,1]"])
gamma2_draws_pl_h <- as.matrix(pl_h[, "gamma[2,1]"])
sigma_draws_pl_h <- as.matrix(pl_h[, "sigma"])

num_sam_pl_h <- nrow(gamma1_draws_pl_h)
num_obs_pl_h <- nrow(stan_data_lr_h$log_x)

log_y_pred_draws_pl_h <- matrix(NA, nrow = num_sam_pl_h, ncol = num_obs_pl_h)

for (i in 1:num_sam_pl_h) {
  log_y_pred_draws_pl_h[i, ] <- gamma1_draws_pl_h[i] + gamma2_draws_pl_h[i] * stan_data_lr_h$log_x[, 2]
}

var_log_y_pred_pl_h <- apply(log_y_pred_draws_pl_h, 1, var)

var_res_pl_h <- sigma_draws_pl_h^2

bayes_R2_draws_pl_h <- var_log_y_pred_pl_h / (var_log_y_pred_pl_h + var_res_pl_h)

summary_bayes_R2_pl_h <- summary(bayes_R2_draws_pl_h)
summary_bayes_R2_pl_h
```

## CR

```{r}
stan_data_lr_cr <- tar_read(stan_data_lr_cr)

# Without wood density
pl_nou_cr <- tar_read(fit_lr2_draws_power_law1_nou_cr)

gamma1_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "gamma[1]"])
gamma2_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "gamma[2]"])
sigma_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "sigma"])

num_sam_pl_nou_cr <- nrow(gamma1_draws_pl_nou_cr)
num_obs_pl_nou_cr <- nrow(stan_data_lr_cr$log_x)

log_y_pred_draws_pl_nou_cr <- matrix(NA, nrow = num_sam_pl_nou_cr, ncol = num_obs_pl_nou_cr)

for (i in 1:num_sam_pl_nou_cr) {
  log_y_pred_draws_pl_nou_cr[i, ] <- gamma1_draws_pl_nou_cr[i] + 
                                     gamma2_draws_pl_nou_cr[i] * stan_data_lr_cr$log_x[, 2]
}

var_log_y_pred_pl_nou_cr <- apply(log_y_pred_draws_pl_nou_cr, 1, var)

var_res_pl_nou_cr <- sigma_draws_pl_nou_cr^2

bayes_R2_draws_pl_nou_cr <- var_log_y_pred_pl_nou_cr / (var_log_y_pred_pl_nou_cr + var_res_pl_nou_cr)

summary_bayes_R2_pl_nou_cr <- summary(bayes_R2_draws_pl_nou_cr)

summary_bayes_R2_pl_nou_cr

# With wood density
stan_data_lr_cr <- tar_read(stan_data_lr_cr)  
pl_cr <- tar_read(fit_lr2_draws_power_law1_cr)

gamma1_draws_pl_cr <- as.matrix(pl_cr[, "gamma[1,1]"])
gamma2_draws_pl_cr <- as.matrix(pl_cr[, "gamma[2,1]"])
sigma_draws_pl_cr <- as.matrix(pl_cr[, "sigma"])

num_sam_pl_cr <- nrow(gamma1_draws_cr)
num_obs_pl_cr <- nrow(stan_data_lr_cr$log_x)

log_y_pred_draws_pl_cr <- matrix(NA, nrow = num_sam_pl_cr, ncol = num_obs_pl_cr)

for (i in 1:num_sam_cr) {
  log_y_pred_draws_pl_cr[i, ] <- gamma1_draws_pl_cr[i] + gamma2_draws_pl_cr[i] * stan_data_lr_cr$log_x[, 2]
}

var_log_y_pred_pl_cr <- apply(log_y_pred_draws_pl_cr, 1, var)

var_res_pl_cr <- sigma_draws_pl_cr^2

bayes_R2_draws_pl_cr <- var_log_y_pred_pl_cr / (var_log_y_pred_pl_cr + var_res_pl_cr)

summary_bayes_R2_pl_cr <- summary(bayes_R2_draws_pl_cr)
summary_bayes_R2__pl_cr
```

# DBH

## DBH0

```{r}
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

# Without wood density
pl_nou_dbh <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl_nou_dbh

gamma1_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[1]"])
gamma2_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[2]"])
gamma3_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[3]"])
sigma_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "sigma"])

num_sam_nou_dbh <- nrow(gamma1_draws_nou_dbh)
num_obs_nou_dbh <- nrow(stan_data_lr_dbh$log_x)

log_y_pred_draws_nou_dbh <- matrix(NA, nrow = num_sam_nou_dbh, ncol = num_obs_nou_dbh)

for (i in 1:num_sam_nou_dbh) {
  log_y_pred_draws_nou_dbh[i, ] <- gamma1_draws_nou_dbh[i] + 
                               gamma2_draws_nou_dbh[i] * stan_data_lr_dbh$log_x[, 2] + 
                               gamma3_draws_nou_dbh[i] * stan_data_lr_dbh$log_x[, 3]
}

var_log_y_pred_nou_dbh <- apply(log_y_pred_draws_nou_dbh, 1, var)

var_res_nou_dbh <- sigma_draws_nou_dbh^2

bayes_R2_draws_nou_dbh <- var_log_y_pred_nou_dbh / (var_log_y_pred_nou_dbh + var_res_nou_dbh)

summary_bayes_R2_nou_dbh <- summary(bayes_R2_draws_nou_dbh)

summary_bayes_R2_nou_dbh

# With wood density
pl_dbh <- tar_read(fit_lr2_draws_power_law1_dbh)  
pl_dbh

gamma1_draws_dbh <- as.matrix(pl_dbh[, "gamma[1,1]"])
gamma2_draws_dbh <- as.matrix(pl_dbh[, "gamma[2,1]"])
gamma3_draws_dbh <- as.matrix(pl_dbh[, "gamma[3,1]"])
sigma_draws_dbh <- as.matrix(pl_dbh[, "sigma"])

num_sam_dbh <- nrow(gamma1_draws_dbh)
num_obs_dbh <- nrow(stan_data_lr_dbh$log_x)

log_y_pred_draws_dbh <- matrix(NA, nrow = num_sam_dbh, ncol = num_obs_dbh)

for (i in 1:num_sam_dbh) {
  log_y_pred_draws_dbh[i, ] <- gamma1_draws_dbh[i] + 
                               gamma2_draws_dbh[i] * stan_data_lr_dbh$log_x[, 2] + 
                               gamma3_draws_dbh[i] * stan_data_lr_dbh$log_x[, 3]
}

var_log_y_pred_dbh <- apply(log_y_pred_draws_dbh, 1, var)

var_res_dbh <- sigma_draws_dbh^2

bayes_R2_draws_dbh <- var_log_y_pred_dbh / (var_log_y_pred_dbh + var_res_dbh)

summary_bayes_R2_dbh <- summary(bayes_R2_draws_dbh)

summary_bayes_R2_dbh
```

## DBH1

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Without wood density
pl_nou_dbh1 <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl_nou_dbh1

gamma1_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "gamma[1]"])
gamma2_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "gamma[2]"])
sigma_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "sigma"])

num_sam_nou_dbh1 <- nrow(gamma1_draws_nou_dbh1)
num_obs_nou_dbh1 <- nrow(stan_data_lr_dbh1$log_x)

log_y_pred_draws_nou_dbh1 <- matrix(NA, nrow = num_sam_nou_dbh1, ncol = num_obs_nou_dbh1)

for (i in 1:num_sam_nou_dbh1) {
  log_y_pred_draws_nou_dbh1[i, ] <- gamma1_draws_nou_dbh1[i] + 
                                    gamma2_draws_nou_dbh1[i] * stan_data_lr_dbh1$log_x[, 2]
}

var_log_y_pred_nou_dbh1 <- apply(log_y_pred_draws_nou_dbh1, 1, var)

var_res_nou_dbh1 <- sigma_draws_nou_dbh1^2

bayes_R2_draws_nou_dbh1 <- var_log_y_pred_nou_dbh1 / (var_log_y_pred_nou_dbh1 + var_res_nou_dbh1)

summary_bayes_R2_nou_dbh1 <- summary(bayes_R2_draws_nou_dbh1)

summary_bayes_R2_nou_dbh1

# With wood density
pl_dbh1 <- tar_read(fit_lr2_draws_power_law1_dbh1)  
pl_dbh1

gamma1_draws_dbh1 <- as.matrix(pl_dbh1[, "gamma[1,1]"])
gamma2_draws_dbh1 <- as.matrix(pl_dbh1[, "gamma[2,1]"])
sigma_draws_dbh1 <- as.matrix(pl_dbh1[, "sigma"])

# Number of posterior samples and observations
num_sam_dbh1 <- nrow(gamma1_draws_dbh1)
num_obs_dbh1 <- nrow(stan_data_lr_dbh1$log_x)

# Calculate predicted log values across all posterior samples using two predictors
log_y_pred_draws_dbh1 <- matrix(NA, nrow = num_sam_dbh1, ncol = num_obs_dbh1)

for (i in 1:num_sam_dbh1) {
  log_y_pred_draws_dbh1[i, ] <- gamma1_draws_dbh1[i] + 
                                gamma2_draws_dbh1[i] * stan_data_lr_dbh1$log_x[, 2]
}

var_log_y_pred_dbh1 <- apply(log_y_pred_draws_dbh1, 1, var)

var_res_dbh1 <- sigma_draws_dbh1^2

bayes_R2_draws_dbh1 <- var_log_y_pred_dbh1 / (var_log_y_pred_dbh1 + var_res_dbh1)

summary_bayes_R2_dbh1 <- summary(bayes_R2_draws_dbh1)

summary_bayes_R2_dbh1
```

## DBH2

```{r}
stan_data_lr_dbh2 <- tar_read(stan_data_lr_dbh2)
str(stan_data_lr_dbh2)

# Without wood desnity
pl_nou_dbh2 <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl_nou_dbh2

gamma1_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "gamma[1]"])
gamma2_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "gamma[2]"])
sigma_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "sigma"])

num_sam_nou_dbh2 <- nrow(gamma1_draws_nou_dbh2)
num_obs_nou_dbh2 <- nrow(stan_data_lr_dbh2$log_x)

log_y_pred_draws_nou_dbh2 <- matrix(NA, nrow = num_sam_nou_dbh2, ncol = num_obs_nou_dbh2)

for (i in 1:num_sam_nou_dbh2) {
  log_y_pred_draws_nou_dbh2[i, ] <- gamma1_draws_nou_dbh2[i] + 
                                    gamma2_draws_nou_dbh2[i] * stan_data_lr_dbh2$log_x[, 2]
}

var_log_y_pred_nou_dbh2 <- apply(log_y_pred_draws_nou_dbh2, 1, var)

var_res_nou_dbh2 <- sigma_draws_nou_dbh2^2

bayes_R2_draws_nou_dbh2 <- var_log_y_pred_nou_dbh2 / (var_log_y_pred_nou_dbh2 + var_res_nou_dbh2)

summary_bayes_R2_nou_dbh2 <- summary(bayes_R2_draws_nou_dbh2)

summary_bayes_R2_nou_dbh2

# With wood density
pl_dbh2 <- tar_read(fit_lr2_draws_power_law1_dbh2)  
pl_dbh2

gamma1_draws_dbh2 <- as.matrix(pl_dbh2[, "gamma[1,1]"])
gamma2_draws_dbh2 <- as.matrix(pl_dbh2[, "gamma[2,1]"])
sigma_draws_dbh2 <- as.matrix(pl_dbh2[, "sigma"])

num_sam_dbh2 <- nrow(gamma1_draws_dbh2)
num_obs_dbh2 <- nrow(stan_data_lr_dbh2$log_x)

log_y_pred_draws_dbh2 <- matrix(NA, nrow = num_sam_dbh2, ncol = num_obs_dbh2)

for (i in 1:num_sam_dbh2) {
  log_y_pred_draws_dbh2[i, ] <- gamma1_draws_dbh2[i] + 
                                gamma2_draws_dbh2[i] * stan_data_lr_dbh2$log_x[, 2]
}

var_log_y_pred_dbh2 <- apply(log_y_pred_draws_dbh2, 1, var)

var_res_dbh2 <- sigma_draws_dbh2^2

bayes_R2_draws_dbh2 <- var_log_y_pred_dbh2 / (var_log_y_pred_dbh2 + var_res_dbh2)

summary_bayes_R2_dbh2 <- summary(bayes_R2_draws_dbh2)

summary_bayes_R2_dbh2
```

# DBH3

```{r}
stan_data_lr_dbh3 <- tar_read(stan_data_lr_dbh3)
str(stan_data_lr_dbh3)

# Without wood density
pl_nou_dbh3 <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl_nou_dbh3

gamma1_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "gamma[1]"])
gamma2_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "gamma[2]"])
sigma_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "sigma"])

num_sam_nou_dbh3 <- nrow(gamma1_draws_nou_dbh3)
num_obs_nou_dbh3 <- nrow(stan_data_lr_dbh3$log_x)

log_y_pred_draws_nou_dbh3 <- matrix(NA, nrow = num_sam_nou_dbh3, ncol = num_obs_nou_dbh3)

for (i in 1:num_sam_nou_dbh3) {
  log_y_pred_draws_nou_dbh3[i, ] <- gamma1_draws_nou_dbh3[i] + 
                                    gamma2_draws_nou_dbh3[i] * stan_data_lr_dbh3$log_x[, 2]
}

var_log_y_pred_nou_dbh3 <- apply(log_y_pred_draws_nou_dbh3, 1, var)

var_res_nou_dbh3 <- sigma_draws_nou_dbh3^2

bayes_R2_draws_nou_dbh3 <- var_log_y_pred_nou_dbh3 / (var_log_y_pred_nou_dbh3 + var_res_nou_dbh3)

summary_bayes_R2_nou_dbh3 <- summary(bayes_R2_draws_nou_dbh3)

summary_bayes_R2_nou_dbh3

# With wood density
pl_dbh3 <- tar_read(fit_lr2_draws_power_law1_dbh3)  
pl_dbh3

gamma1_draws_dbh3 <- as.matrix(pl_dbh3[, "gamma[1,1]"])
gamma2_draws_dbh3 <- as.matrix(pl_dbh3[, "gamma[2,1]"])
sigma_draws_dbh3 <- as.matrix(pl_dbh3[, "sigma"])

num_sam_dbh3 <- nrow(gamma1_draws_dbh3)
num_obs_dbh3 <- nrow(stan_data_lr_dbh3$log_x)

log_y_pred_draws_dbh3 <- matrix(NA, nrow = num_sam_dbh3, ncol = num_obs_dbh3)

for (i in 1:num_sam_dbh3) {
  log_y_pred_draws_dbh3[i, ] <- gamma1_draws_dbh3[i] + 
                                gamma2_draws_dbh3[i] * stan_data_lr_dbh3$log_x[, 2]
}

var_log_y_pred_dbh3 <- apply(log_y_pred_draws_dbh3, 1, var)

var_res_dbh3 <- sigma_draws_dbh3^2

bayes_R2_draws_dbh3 <- var_log_y_pred_dbh3 / (var_log_y_pred_dbh3 + var_res_dbh3)

summary_bayes_R2_dbh3 <- summary(bayes_R2_draws_dbh3)

summary_bayes_R2_dbh3
```

# gMM

## H
```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Without wood density
gmm_nou_h <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm_nou_h

gamma1_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[1]"])
gamma2_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[2]"])
gamma3_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[3]"])

sigma_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "sigma"])

num_sam_gmm_nou_h <- nrow(gamma1_draws_gmm_nou_h)
num_obs_gmm_nou_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_gmm_nou_h <- matrix(NA, nrow = num_sam_gmm_nou_h, ncol = num_obs_gmm_nou_h)

for (i in 1:num_sam_gmm_nou_h) {
  log_y_pred_draws_gmm_nou_h[i, ] <- exp(gamma1_draws_gmm_nou_h[i]) + 
                                     exp(gamma2_draws_gmm_nou_h[i]) * log(stan_data_nlr_h$x) - 
                                     log(exp(gamma3_draws_gmm_nou_h[i]) + (stan_data_nlr_h$x)^(exp(gamma2_draws_gmm_nou_h[i])))
}

var_log_y_pred_gmm_nou_h <- apply(log_y_pred_draws_gmm_nou_h, 1, var)

var_res_gmm_nou_h <- sigma_draws_gmm_nou_h^2

bayes_R2_draws_gmm_nou_h <- var_log_y_pred_gmm_nou_h / (var_log_y_pred_gmm_nou_h + var_res_gmm_nou_h)

summary_bayes_R2_gmm_nou_h <- summary(bayes_R2_draws_gmm_nou_h)
summary_bayes_R2_gmm_nou_h

# With wood density
gmm_h <- tar_read(fit_nlr_draws_gmm1_h)
gmm_h

gamma1_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,1]"])
gamma2_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,2]"])
gamma3_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,3]"])
sigma_draws_gmm_h <- as.matrix(gmm_h[, "sigma"])

num_sam_gmm_h <- nrow(gamma1_draws_gmm_h)
num_obs_gmm_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_gmm_h <- matrix(NA, nrow = num_sam_gmm_h, ncol = num_obs_gmm_h)

for (i in 1:num_sam_gmm_h) {
  log_y_pred_draws_gmm_h[i, ] <- exp(gamma1_draws_gmm_h[i]) + 
                                 exp(gamma2_draws_gmm_h[i]) * log(stan_data_nlr_h$x) - 
                                 log(exp(gamma3_draws_gmm_h[i]) + (stan_data_nlr_h$x)^(exp(gamma2_draws_gmm_h[i])))
}

var_log_y_pred_gmm_h <- apply(log_y_pred_draws_gmm_h, 1, var)

var_res_gmm_h <- sigma_draws_gmm_h^2

bayes_R2_draws_gmm_h <- var_log_y_pred_gmm_h / (var_log_y_pred_gmm_h + var_res_gmm_h)

summary_bayes_R2_gmm_h <- summary(bayes_R2_draws_gmm_h)
summary_bayes_R2_gmm_h
```

## CR

```{r}
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Without wood density
gmm_nou_cr <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm_nou_cr

gamma1_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[1]"])
gamma2_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[2]"])
gamma3_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[3]"])

sigma_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "sigma"])

num_sam_gmm_nou_cr <- nrow(gamma1_draws_gmm_nou_cr)
num_obs_gmm_nou_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_gmm_nou_cr <- matrix(NA, nrow = num_sam_gmm_nou_cr, ncol = num_obs_gmm_nou_cr)

for (i in 1:num_sam_gmm_nou_cr) {
  log_y_pred_draws_gmm_nou_cr[i, ] <- exp(gamma1_draws_gmm_nou_cr[i]) + 
                                      exp(gamma2_draws_gmm_nou_cr[i]) * log(stan_data_nlr_cr$x) - 
                                      log(exp(gamma3_draws_gmm_nou_cr[i]) + (stan_data_nlr_cr$x)^(exp(gamma2_draws_gmm_nou_cr[i])))
}

var_log_y_pred_gmm_nou_cr <- apply(log_y_pred_draws_gmm_nou_cr, 1, var)

var_res_gmm_nou_cr <- sigma_draws_gmm_nou_cr^2

bayes_R2_draws_gmm_nou_cr <- var_log_y_pred_gmm_nou_cr / (var_log_y_pred_gmm_nou_cr + var_res_gmm_nou_cr)

summary_bayes_R2_gmm_nou_cr <- summary(bayes_R2_draws_gmm_nou_cr)
summary_bayes_R2_gmm_nou_cr

# With wood density
gmm_cr <- tar_read(fit_nlr_draws_gmm1_cr)
gmm_cr

gamma1_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,1]"])
gamma2_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,2]"])
gamma3_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,3]"])
sigma_draws_gmm_cr <- as.matrix(gmm_cr[, "sigma"])

num_sam_gmm_cr <- nrow(gamma1_draws_gmm_cr)
num_obs_gmm_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_gmm_cr <- matrix(NA, nrow = num_sam_gmm_cr, ncol = num_obs_gmm_cr)

for (i in 1:num_sam_gmm_cr) {
  log_y_pred_draws_gmm_cr[i, ] <- exp(gamma1_draws_gmm_cr[i]) + 
                                  exp(gamma2_draws_gmm_cr[i]) * log(stan_data_nlr_cr$x) - 
                                  log(exp(gamma3_draws_gmm_cr[i]) + (stan_data_nlr_cr$x)^(exp(gamma2_draws_gmm_cr[i])))
}

var_log_y_pred_gmm_cr <- apply(log_y_pred_draws_gmm_cr, 1, var)

var_res_gmm_cr <- sigma_draws_gmm_cr^2

bayes_R2_draws_gmm_cr <- var_log_y_pred_gmm_cr / (var_log_y_pred_gmm_cr + var_res_gmm_cr)

summary_bayes_R2_gmm_cr <- summary(bayes_R2_draws_gmm_cr)
summary_bayes_R2_gmm_cr
```

# Weibull

## H

```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Without wood density
wb_nou_h <- tar_read(fit_nlr_draws_weibull1_nou_h)
wb_nou_h

gamma1_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[1]"])
gamma2_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[2]"])
gamma3_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[3]"])

sigma_draws_wb_nou_h <- as.matrix(wb_nou_h[, "sigma"])

num_sam_wb_nou_h <- nrow(gamma1_draws_wb_nou_h)
num_obs_wb_nou_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_wb_nou_h <- matrix(NA, nrow = num_sam_wb_nou_h, ncol = num_obs_wb_nou_h)

for (i in 1:num_sam_wb_nou_h) {
  log_y_pred_draws_wb_nou_h[i, ] <- exp(gamma1_draws_wb_nou_h[i]) + 
                                    log(1- exp(-(exp(gamma2_draws_wb_nou_h[i])) * (stan_data_nlr_h$x)^(exp(gamma3_draws_wb_nou_h[i]))))
}

var_log_y_pred_wb_nou_h <- apply(log_y_pred_draws_wb_nou_h, 1, var)

var_res_wb_nou_h <- sigma_draws_wb_nou_h^2

bayes_R2_draws_wb_nou_h <- var_log_y_pred_wb_nou_h / (var_log_y_pred_wb_nou_h + var_res_wb_nou_h)

summary_bayes_R2_wb_nou_h <- summary(bayes_R2_draws_wb_nou_h)
summary_bayes_R2_wb_nou_h

# With wood density
wb_h <- tar_read(fit_nlr_draws_weibull1_h)
wb_h

# Extract the gamma and sigma draws
gamma1_draws_wb_h <- as.matrix(wb_h[, "gamma[1,1]"])
gamma2_draws_wb_h <- as.matrix(wb_h[, "gamma[1,2]"])
gamma3_draws_wb_h <- as.matrix(wb_h[, "gamma[1,3]"])
sigma_draws_wb_h <- as.matrix(wb_h[, "sigma"])

num_sam_wb_h <- nrow(gamma1_draws_wb_h)
num_obs_wb_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_wb_h <- matrix(NA, nrow = num_sam_wb_h, ncol = num_obs_wb_h)

for (i in 1:num_sam_wb_h) {
  log_y_pred_draws_wb_h[i, ] <- exp(gamma1_draws_wb_h[i]) + 
                                log(1 - exp(-(exp(gamma2_draws_wb_h[i])) * 
                                (stan_data_nlr_h$x)^(exp(gamma3_draws_wb_h[i]))))
}

var_log_y_pred_wb_h <- apply(log_y_pred_draws_wb_h, 1, var)

var_res_wb_h <- sigma_draws_wb_h^2

bayes_R2_draws_wb_h <- var_log_y_pred_wb_h / (var_log_y_pred_wb_h + var_res_wb_h)

summary_bayes_R2_wb_h <- summary(bayes_R2_draws_wb_h)

summary_bayes_R2_wb_h
```

## CR

```{r}
wb_nou_cr <- tar_read(fit_nlr_draws_weibull1_nou_cr)
wb_nou_cr

gamma1_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[1]"])
gamma2_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[2]"])
gamma3_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[3]"])
sigma_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "sigma"])

num_sam_wb_nou_cr <- nrow(gamma1_draws_wb_nou_cr)
num_obs_wb_nou_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_wb_nou_cr <- matrix(NA, nrow = num_sam_wb_nou_cr, ncol = num_obs_wb_nou_cr)

for (i in 1:num_sam_wb_nou_cr) {
  log_y_pred_draws_wb_nou_cr[i, ] <- exp(gamma1_draws_wb_nou_cr[i]) + 
                                     log(1 - exp(-(exp(gamma2_draws_wb_nou_cr[i])) * 
                                     (stan_data_nlr_cr$x)^(exp(gamma3_draws_wb_nou_cr[i]))))
}

var_log_y_pred_wb_nou_cr <- apply(log_y_pred_draws_wb_nou_cr, 1, var)

var_res_wb_nou_cr <- sigma_draws_wb_nou_cr^2

bayes_R2_draws_wb_nou_cr <- var_log_y_pred_wb_nou_cr / (var_log_y_pred_wb_nou_cr + var_res_wb_nou_cr)

summary_bayes_R2_wb_nou_cr <- summary(bayes_R2_draws_wb_nou_cr)

summary_bayes_R2_wb_nou_cr

# With wood density 
wb_cr <- tar_read(fit_nlr_draws_weibull1_cr)
wb_cr

gamma1_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,1]"])
gamma2_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,2]"])
gamma3_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,3]"])
sigma_draws_wb_cr <- as.matrix(wb_cr[, "sigma"])

num_sam_wb_cr <- nrow(gamma1_draws_wb_cr)
num_obs_wb_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_wb_cr <- matrix(NA, nrow = num_sam_wb_cr, ncol = num_obs_wb_cr)

for (i in 1:num_sam_wb_cr) {
  log_y_pred_draws_wb_cr[i, ] <- exp(gamma1_draws_wb_cr[i]) + 
                                 log(1 - exp(-(exp(gamma2_draws_wb_cr[i])) * 
                                 (stan_data_nlr_cr$x)^(exp(gamma3_draws_wb_cr[i]))))
}

var_log_y_pred_wb_cr <- apply(log_y_pred_draws_wb_cr, 1, var)

var_res_wb_cr <- sigma_draws_wb_cr^2

bayes_R2_draws_wb_cr <- var_log_y_pred_wb_cr / (var_log_y_pred_wb_cr + var_res_wb_cr)

summary_bayes_R2_wb_cr <- summary(bayes_R2_draws_wb_cr)

summary_bayes_R2_wb_cr
```

```{r}
# H (Height) data
stan_data_lr_h <- tar_read(stan_data_lr_h)
pl_nou_h <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl_h <- tar_read(fit_lr2_draws_power_law1_h)

# CR (Crown Radius) data
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
pl_nou_cr <- tar_read(fit_lr2_draws_power_law1_nou_cr)
pl_cr <- tar_read(fit_lr2_draws_power_law1_cr)

# DBH0 (Diameter at Breast Height) data
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
pl_nou_dbh <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl_dbh <- tar_read(fit_lr2_draws_power_law1_dbh)

# DBH1 data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
pl_nou_dbh1 <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl_dbh1 <- tar_read(fit_lr2_draws_power_law1_dbh1)

# DBH2 data
stan_data_lr_dbh2 <- tar_read(stan_data_lr_dbh2)
pl_nou_dbh2 <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl_dbh2 <- tar_read(fit_lr2_draws_power_law1_dbh2)

# DBH3 data
stan_data_lr_dbh3 <- tar_read(stan_data_lr_dbh3)
pl_nou_dbh3 <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl_dbh3 <- tar_read(fit_lr2_draws_power_law1_dbh3)

# gMM models for H
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
gmm_nou_h <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm_h <- tar_read(fit_nlr_draws_gmm1_h)

# gMM models for CR
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
gmm_nou_cr <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm_cr <- tar_read(fit_nlr_draws_gmm1_cr)

# Weibull models for H
wb_nou_h <- tar_read(fit_nlr_draws_weibull1_nou_h)
wb_h <- tar_read(fit_nlr_draws_weibull1_h)

# Weibull models for CR
wb_nou_cr <- tar_read(fit_nlr_draws_weibull1_nou_cr)
wb_cr <- tar_read(fit_nlr_draws_weibull1_cr)
```

# Usage

```{r}
source("R/functions.R")

# lr
bayes_R2_pl_nou_h <- calculate_bayes_R2_lr(stan_data_lr_h, pl_nou_h, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_h <- calculate_bayes_R2_lr(stan_data_lr_h, pl_h, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_cr <- calculate_bayes_R2_lr(stan_data_lr_cr, pl_nou_cr, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_cr <- calculate_bayes_R2_lr(stan_data_lr_cr, pl_cr, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh <- calculate_bayes_R2_lr(stan_data_lr_dbh, pl_nou_dbh, c("gamma[1]", "gamma[2]", "gamma[3]"))
bayes_R2_pl_dbh <- calculate_bayes_R2_lr(stan_data_lr_dbh, pl_dbh, c("gamma[1,1]", "gamma[2,1]", "gamma[3,1]"))

bayes_R2_pl_nou_dbh1 <- calculate_bayes_R2_lr(stan_data_lr_dbh1, pl_nou_dbh1, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh1 <- calculate_bayes_R2_lr(stan_data_lr_dbh1, pl_dbh1, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh2 <- calculate_bayes_R2_lr(stan_data_lr_dbh2, pl_nou_dbh2, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh2 <- calculate_bayes_R2_lr(stan_data_lr_dbh2, pl_dbh2, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh3 <- calculate_bayes_R2_lr(stan_data_lr_dbh3, pl_nou_dbh3, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh3 <- calculate_bayes_R2_lr(stan_data_lr_dbh3, pl_dbh3, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_gmm_nou_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, gmm_nou_h, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_gmm_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, gmm_h, c("gamma[11,]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_gmm_nou_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, gmm_nou_cr, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_gmm_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, gmm_cr, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_wb_nou_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, wb_nou_h, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_wb_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, wb_h, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_wb_nou_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, wb_nou_cr, c("gamma[1]", "gamma[2]", "gamma[3]"), "weibull")
bayes_R2_wb_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, wb_cr, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "weibull")

```

```{r}
source("R/functions.R")
r2_df <- tar_read(R2_df)
r2_df
r2_tbl <- generate_r2_table(r2_df)
r2_tbl

loo_tbl <- tar_read(loo_tbl)
loo_tbl
com_tbl <- generate_comparison_table(loo_tbl)
com_tbl

loo_r2_tbl <- loo_r2_table(r2_tbl, com_tbl)
loo_r2_tbl
```

### 2025-03-26 docs/R2_new.qmd
Initial commit for tree-allo

--- 2025-03-26 docs/R2.qmd
@@
 
-```{r setup, include=FALSE}
-knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
-```
-
 ```{r global_options, include=FALSE}
-library(knitr)
+knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
+knitr::opts_knit$set(root.dir = here::here())
 knitr::opts_chunk$set(
@@
 )
+
 ```
@@
 library(tidyverse)
-source("R/functions.R")
+# source("R/functions.R")
 ```
@@
 ```{r}
-stan_data_lr_h <- tar_read(stan_data_lr_h)
-str(stan_data_lr_h)
+pl_nou_ang_h_sum <- tar_read(fit_lr_nou_summary_pl_ang_h)
+pl_nou_ang_h_sum
+
+beta_pl_nou_ang_h_sum <- pl_nou_ang_h_sum |> filter(grepl("beta", variable))
+beta_pl_nou_ang_h_sum
+
+### ANG
+stan_data_lr_ang_h <- tar_read(stan_data_lr_ang_h)
+str(stan_data_lr_ang_h)
 
 # Without wood density
-pl_nou_h <- tar_read(fit_lr2_draws_power_law1_nou_h)
-pl_nou_h
+pl_nou_ang_h <- tar_read(fit_lr_nou_draws_pl_ang_h)
+pl_nou_ang_h
 
-gamma1_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[1]"])
-gamma2_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[2]"])
-sigma_draws_pl_nou_h <- as.matrix(pl_nou_h[, "sigma"])
+# Get column names that match beta[1, ] and beta[2, ]
+beta1_cols <- grep("^beta\\[1,\\d+\\]", colnames(pl_nou_ang_h), value = TRUE)
+beta2_cols <- grep("^beta\\[2,\\d+\\]", colnames(pl_nou_ang_h), value = TRUE)
 
-num_sam_pl_nou_h <- nrow(gamma1_draws_pl_nou_h)
-num_obs_pl_nou_h <- nrow(stan_data_lr_h$log_x)
+# Extract the beta[1, ] and beta[2, ] columns
+beta1 <- as.matrix(pl_nou_ang_h[, beta1_cols])
+beta2 <- as.matrix(pl_nou_ang_h[, beta2_cols])
 
-log_y_pred_draws_pl_nou_h <- matrix(NA, nrow = num_sam_pl_nou_h, ncol = num_obs_pl_nou_h)
+dim(beta1)
+dim(beta2)
 
-for (i in 1:num_sam_pl_nou_h) {
-  log_y_pred_draws_pl_nou_h[i, ] <- gamma1_draws_pl_nou_h[i] + gamma2_draws_pl_nou_h[i] * stan_data_lr_h$log_x[, 2]
+# Assuming beta1 and beta2 are already extracted as matrices with dimensions (3000, 1214)
+y_pred <- matrix(NA, nrow = nrow(stan_data_lr_ang_h$log_x), ncol = nrow(beta1))  # Empty matrix for predictions
+
+# Loop over posterior draws and compute predicted log_y for each tree
+for (i in 1:nrow(beta1)) {  # Loop over iterations (posterior draws)
+  for (n in 1:stan_data_lr_ang_h$N) {       # Loop over trees
+    # Extract the corresponding beta[1, jj[n]] and beta[2, jj[n]]
+    beta_n <- c(beta1[i, stan_data_lr_ang_h$jj[n]], beta2[i, stan_data_lr_ang_h$jj[n]])
+
+    # Multiply the log_x (tree-level predictors) by the corresponding beta values
+    y_pred[n, i] <- stan_data_lr_ang_h$log_x[n, ] %*% beta_n  # Predicted log_y for each tree
+  }
 }
 
-var_log_y_pred_pl_nou_h <- apply(log_y_pred_draws_pl_nou_h, 1, var)
+dim(y_pred)
 
-var_res_pl_nou_h <- sigma_draws_pl_nou_h^2
-
-bayes_R2_draws_pl_nou_h <- var_log_y_pred_pl_nou_h / (var_log_y_pred_pl_nou_h + var_res_pl_nou_h)
-
-summary_bayes_R2_pl_nou_h <- summary(bayes_R2_draws_pl_nou_h)
-
-summary_bayes_R2_pl_nou_h
-
-# With wood density
-pl_h <- tar_read(fit_lr2_draws_power_law1_h)
-pl_h
-
-gamma1_draws_pl_h <- as.matrix(pl_h[, "gamma[1,1]"])
-gamma2_draws_pl_h <- as.matrix(pl_h[, "gamma[2,1]"])
-sigma_draws_pl_h <- as.matrix(pl_h[, "sigma"])
-
-num_sam_pl_h <- nrow(gamma1_draws_pl_h)
-num_obs_pl_h <- nrow(stan_data_lr_h$log_x)
-
-log_y_pred_draws_pl_h <- matrix(NA, nrow = num_sam_pl_h, ncol = num_obs_pl_h)
-
-for (i in 1:num_sam_pl_h) {
-  log_y_pred_draws_pl_h[i, ] <- gamma1_draws_pl_h[i] + gamma2_draws_pl_h[i] * stan_data_lr_h$log_x[, 2]
-}
-
-var_log_y_pred_pl_h <- apply(log_y_pred_draws_pl_h, 1, var)
-
-var_res_pl_h <- sigma_draws_pl_h^2
-
-bayes_R2_draws_pl_h <- var_log_y_pred_pl_h / (var_log_y_pred_pl_h + var_res_pl_h)
-
-summary_bayes_R2_pl_h <- summary(bayes_R2_draws_pl_h)
-summary_bayes_R2_pl_h
 ```
@@
 ```
+
+# H-DBH
+
+## ANG
+
+### Power-law
+```{r}
+stan_data_lr_ang_h <- tar_read(stan_data_lr_ang_h)
+str(stan_data_lr_ang_h)
+
+fit_lr_nou_draws_pl_ang_h <- tar_read(fit_lr_nou_draws_pl_ang_h)
+fit_lr_nou_draws_pl_ang_h
+
+r2_pl_nou_h_ang <- calculate_bayes_R2(
+  stan_data_lr_ang_h,
+  fit_lr_nou_draws_pl_ang_h,
+  c("gamma[1]", "gamma[2]"),
+  "lr"
+)
+
+summary(r2_pl_nou_h_ang)
+```
+
+### gMM
+```{r}
+stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
+str(stan_data_nlr_ang_h)
+
+fit_nlr_nou_draws_gmm_ang_h <- tar_read(fit_nlr_nou_draws_gmm_ang_h)
+fit_nlr_nou_draws_gmm_ang_h
+
+r2_gmm_nou_h_ang <- calculate_bayes_R2(
+  stan_data_nlr_ang_h,
+  fit_nlr_nou_draws_gmm_ang_h,
+  c("gamma[1]", "gamma[2]", "gamma[3]"),
+  "gmm"
+)
+
+summary(r2_gmm_nou_h_ang)
+```
+
+## Weibull
+
+```{r}
+stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
+str(stan_data_nlr_ang_h)
+
+fit_nlr_nou_draws_weibull_ang_h <- tar_read(fit_nlr_nou_draws_weibull_ang_h)
+fit_nlr_nou_draws_weibull_ang_h
+
+r2_weibull_nou_h_ang <- calculate_bayes_R2(
+  stan_data_nlr_ang_h,
+  fit_nlr_nou_draws_weibull_ang_h,
+  c("gamma[1]", "gamma[2]", "gamma[3]"),
+  "weibull"
+)
+
+summary(r2_weibull_nou_h_ang)
+```
+
+```{r}
+
+```{r}
+
+```
+
+```{r}

### 2025-03-26 docs/agb.qmd
Initial commit for tree-allo

---
title: "AGB estimation"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r include=FALSE}
library(tidyverse)
library(here)
library(targets)
library(knitr)
library(scales)
library(patchwork)
library(cowplot)
```

# Parameter extraction

```{r}
options(width=150)
tallo_wd <- tar_read(tallo_wd) |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
         log_h = log(h),
         log_cr = log(cr),
         log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  select(-division,-family, -genus, - latitude, - longitude, -reference_id, -height_outlier, -crown_radius_outlier, -mean_wd, -unit, -wd_s) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd
summary(tallo_wd)

sp_posterior_df <- tar_read(sp_posterior_df)
head(sp_posterior_df)

# H models 2a, 2b
h_wb_df <- sp_posterior_df |>
  filter(Dependent_variable == "Tree Height") |>
  dplyr::select(sp, a, b, k)

head(h_wb_df)

h_pl_df <- sp_posterior_df |>
  filter(Dependent_variable == "Tree Height", is.na(k)) |>
   dplyr::select(sp, a, b)

head(h_pl_df)

# DBH models 3a, 3b, 3c
dbh_df <- sp_posterior_df |>
  filter(Dependent_variable == "DBH") |>
  dplyr::select(sp, a, b, c)

head(dbh_df)

dbh2_df <- sp_posterior_df |>
  filter(Dependent_variable == "DBH2") |>
  dplyr::select(sp, a, b, c)

head(dbh2_df)

dbh3_df <- sp_posterior_df |>
  filter(Dependent_variable == "DBH3") |>
  dplyr::select(sp, a, b, c)

head(dbh3_df)
```

# Compute AGB for various scenarios

```{r}
h_wb_df <- h_wb_df |> distinct(sp, .keep_all = TRUE)
h_pl_df <- h_pl_df |> distinct(sp, .keep_all = TRUE)
dbh_df <- dbh_df |> distinct(sp, .keep_all = TRUE)
dbh2_df <- dbh2_df |> distinct(sp, .keep_all = TRUE)
dbh3_df <- dbh3_df |> distinct(sp, .keep_all = TRUE)

agb_df <- tallo_wd |>
  left_join(h_wb_df |> rename(a_h_wb = a, b_h_wb = b, k_h_wb = k), by = "sp") |>    # Joining h_df with renamed columns
  left_join(h_pl_df |> rename(a_h_pl = a, b_h_pl = b), by = "sp") |>    # Joining h_pl_df with renamed columns
  left_join(dbh_df |> rename(a_dbh = a, b_dbh = b, c_dbh = c), by = "sp") |>  # Joining dbh_df with renamed columns
  left_join(dbh2_df |> rename(a_dbh2 = a, b_dbh2 = b), by = "sp") |> # Joining dbh2_df with renamed columns
  left_join(dbh3_df |> rename(a_dbh3 = a, b_dbh3 = b), by = "sp")

agb_df <- agb_df |>
  mutate(
    a_h_wb = as.numeric(gsub(" .*", "", a_h_wb)),
    b_h_wb = as.numeric(gsub(" .*", "", b_h_wb)),
    k_h_wb = as.numeric(gsub(" .*", "", k_h_wb)),

    a_h_pl = as.numeric(gsub(" .*", "", a_h_pl)),
    b_h_pl = as.numeric(gsub(" .*", "", b_h_pl)),

    a_dbh = as.numeric(gsub(" .*", "", a_dbh)),
    b_dbh = as.numeric(gsub(" .*", "", b_dbh)),
    c_dbh = as.numeric(gsub(" .*", "", c_dbh)),

    a_dbh2 = as.numeric(gsub(" .*", "", a_dbh2)),
    b_dbh2 = as.numeric(gsub(" .*", "", b_dbh2)),

    a_dbh3 = as.numeric(gsub(" .*", "", a_dbh3)),
    b_dbh3 = as.numeric(gsub(" .*", "", b_dbh3))
  )

agb_df
```

```{r}
calculate_agb <- function(agb_df) {
  agb_df |>
    mutate(
      # Baseline biomass calculation
      log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

      # Weibull-based height estimation and biomass calculation
      height_wb = as.numeric(a_h_wb) * (1 - exp(-as.numeric(b_h_wb) * dbh^as.numeric(k_h_wb))),
      log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),

      # Power-law based height estimation and biomass calculation
      height_pl = as.numeric(a_h_pl) * dbh^as.numeric(b_h_pl),
      log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),

      # Power-law DBH estimation using CR and H, and biomass calculation
      dbh_pl = as.numeric(a_dbh) * (cr^as.numeric(b_dbh)) * (h^as.numeric(c_dbh)),
      log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,

      # Power-law DBH estimation using CR only, and biomass calculation
      dbh_pl2 = as.numeric(a_dbh2) * (cr^as.numeric(b_dbh2)),
      log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,

      # Power-law DBH estimation using H only, and biomass calculation
      dbh_pl3 = as.numeric(a_dbh3) * (h^as.numeric(b_dbh3)),
      log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h
    ) |>
    dplyr::select(
      tree_id, sp, log_dbh, log_h, log_wd, 
      log_AGB_bl, log_AGB_wb_h, log_AGB_pl_h, 
      log_AGB_pl_dbh, log_AGB_pl_dbh2, log_AGB_pl_dbh3
    )
}

# Apply the function to your dataset
agb_df <- calculate_agb(agb_df)
agb_df

```

# Box plot

```{r}
agb_long <- agb_df |>
  tidyr::pivot_longer(cols = starts_with("log_AGB"), names_to = "Scenario", values_to = "AGB")

ggplot(agb_long, aes(x = Scenario, y = AGB)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Senarios",
       y = "log predicted AGB (kg)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Performance metrics

```{r}
metrics_df <- agb_df |>
  summarize(
    # Weibull Height Model 2a
    mae_wb_h = mean(abs(log_AGB_bl - log_AGB_wb_h), na.rm = TRUE),
    rmse_wb_h = sqrt(mean((log_AGB_bl - log_AGB_wb_h)^2, na.rm = TRUE)),
    r2_wb_h = 1 - (sum((log_AGB_bl - log_AGB_wb_h)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_wb_h = mean(log_AGB_wb_h - log_AGB_bl, na.rm = TRUE),

    # Power-Law Height Model 2b
    mae_pl_h = mean(abs(log_AGB_bl - log_AGB_pl_h), na.rm = TRUE),
    rmse_pl_h = sqrt(mean((log_AGB_bl - log_AGB_pl_h)^2, na.rm = TRUE)),
    r2_pl_h = 1 - (sum((log_AGB_bl - log_AGB_pl_h)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_pl_h = mean(log_AGB_pl_h - log_AGB_bl, na.rm = TRUE),

    # Power-Law Scenario 3a
    mae_pl_dbh = mean(abs(log_AGB_bl - log_AGB_pl_dbh), na.rm = TRUE),
    rmse_pl_dbh = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh)^2, na.rm = TRUE)),
    r2_pl_dbh = 1 - (sum((log_AGB_bl - log_AGB_pl_dbh)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_pl_dbh = mean(log_AGB_pl_dbh - log_AGB_bl, na.rm = TRUE),

    # Power-Law Scenario 3b
    mae_pl_dbh2 = mean(abs(log_AGB_bl - log_AGB_pl_dbh2), na.rm = TRUE),
    rmse_pl_dbh2 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh2)^2, na.rm = TRUE)),
    r2_pl_dbh2 = 1 - (sum((log_AGB_bl - log_AGB_pl_dbh2)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_pl_dbh2 = mean(log_AGB_pl_dbh2 - log_AGB_bl, na.rm = TRUE),

    # Power-Law Scenario 3c
    mae_pl_dbh3 = mean(abs(log_AGB_bl - log_AGB_pl_dbh3), na.rm = TRUE),
    rmse_pl_dbh3 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh3)^2, na.rm = TRUE)),
    r2_pl_dbh3 = 1 - (sum((log_AGB_bl - log_AGB_pl_dbh3)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_pl_dbh3 = mean(log_AGB_pl_dbh3 - log_AGB_bl, na.rm = TRUE),

    # CR*H
    mae_cr_h = mean(abs(log_AGB_bl - log_AGB_obs_CR_H), na.rm = TRUE),
    rmse_cr_h = sqrt(mean((log_AGB_bl - log_AGB_obs_CR_H)^2, na.rm = TRUE)),
    r2_pl_cr_h = 1 - (sum((log_AGB_bl - log_AGB_obs_CR_H)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)),
    bias_cr_h = mean(log_AGB_obs_CR_H - log_AGB_bl, na.rm = TRUE),
  )

metrics_df

metrics_long <- metrics_df |>
  pivot_longer(cols = everything(), names_to = "Metric_Model", values_to = "Value") |>
  separate(Metric_Model, into = c("Metric", "Model"), sep = "_", extra = "merge") |>
  filter(Metric != "bias")  # Exclude the bias metric

ggplot(metrics_long, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Senarios",
       y = "Metric Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set2")
```

# Predicted vs. baseline AGB plots

```{r}
model_colors <- c("h_wb" = "#d55e00",    # Magenta/pinkish color
                  "h_pl" = "#0072b2",    # Blue
                  "dbh_pl" = "#e69f00",  # Orange
                  "dbh2_pl" = "#009e73", # Green
                  "dbh3_pl" = "#56b4e9", # Cyan/Light blue
                  "cr_h" = "#56b8e9")

# Refactor the pred_bl function to include all models
pred_bl <- function(data, baseline, predicted, rmse, bias, model_name) {
  ggplot(data, aes(x = exp(baseline), y = exp(predicted))) +
    geom_point(alpha = 0.3, color = "grey") +
    geom_smooth(method = "lm", se = FALSE, color = model_colors[model_name], linetype = "solid") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
    scale_x_log10(labels = scales::comma) +
    scale_y_log10(labels = scales::comma) +
    labs(x = "Baseline AGB (kg)", y = paste("Predicted biomass (kg) -", model_name)) +
    annotate("text", x = 1, y = 1e5, label = paste("RMSE =", format(round(rmse, 2), nsmall = 2), "Mg"), hjust = 0) +
    annotate("text", x = 1, y = 5e4, label = paste("Bias =", format(round(bias * 100, 2), nsmall = 2), "%"), hjust = 0) +
    theme_minimal()
}

p1 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_wb_h, 
              mean(metrics_df$rmse_wb_h), mean(metrics_df$bias_wb_h), "h_wb")

p2 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_h, 
              mean(metrics_df$rmse_pl_h), mean(metrics_df$bias_pl_h), "h_pl")

p3 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh, 
              mean(metrics_df$rmse_pl_dbh), mean(metrics_df$bias_pl_dbh), "dbh_pl")

p4 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh2, 
              mean(metrics_df$rmse_pl_dbh2), mean(metrics_df$bias_pl_dbh2), "dbh2_pl")

p5 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh3, 
              mean(metrics_df$rmse_pl_dbh3), mean(metrics_df$bias_pl_dbh3), "dbh3_pl")

p6 <- pred_bl(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_obs_CR_H, 
              mean(metrics_df$rmse_cr_h), mean(metrics_df$bias_cr_h), "cr_h")

```

```{r}
# agb_df <- agb_df |>
#   mutate(
#     abs_error_wb_h = abs(exp(log_AGB_wb_h) - exp(log_AGB_bl)),
#     abs_error_pl_h = abs(exp(log_AGB_pl_h) - exp(log_AGB_bl)),
#     abs_error_pl_dbh = abs(exp(log_AGB_pl_dbh) - exp(log_AGB_bl)),
#     abs_error_pl_dbh2 = abs(exp(log_AGB_pl_dbh2) - exp(log_AGB_bl)),
#     abs_error_pl_dbh3 = abs(exp(log_AGB_pl_dbh3) - exp(log_AGB_bl))
#   )

# errors_long <- agb_df |>
#   select(tree_id, abs_error_wb_h, abs_error_pl_h, abs_error_pl_dbh, abs_error_pl_dbh2, abs_error_pl_dbh3) |>
#   pivot_longer(cols = starts_with("abs_error"), names_to = "Model", values_to = "Abs_Error") |>
#   mutate(Model = case_when(
#     Model == "abs_error_wb_h" ~ "h_wb",
#     Model == "abs_error_pl_h" ~ "h_pl",
#     Model == "abs_error_pl_dbh" ~ "dbh_pl",
#     Model == "abs_error_pl_dbh2" ~ "dbh2_pl",
#     Model == "abs_error_pl_dbh3" ~ "dbh3_pl"
#   ))

# p6 <- ggplot(errors_long, aes(x = Abs_Error, color = Model)) +
#   geom_density(fill = NA) + 
#   scale_x_log10(labels = scales::comma) + 
#   scale_y_continuous(position = "right") + 
#   scale_color_manual(values = model_colors) + 
#   labs(x = "Error (kg)", y = "Density") +
#   theme_minimal() +
#   theme(legend.position = "none", plot.margin = margin(5.5, 100, 5.5, 5.5))

# p6 <- ggdraw(p6) +
#   draw_line(x = c(0.88, 0.9), y = c(0.85, 0.85), color = "#d55e00", size = 1) +  # h_wb line
#   draw_text("h_wb", x = 0.91, y = 0.85, hjust = 0, size = 12) +  # h_wb label
#   draw_line(x = c(0.88, 0.9), y = c(0.8, 0.8), color = "#0072b2", size = 1) +   # h_pl line
#   draw_text("h_pl", x = 0.91, y = 0.8, hjust = 0, size = 12) +  # h_pl label
#   draw_line(x = c(0.88, 0.9), y = c(0.75, 0.75), color = "#e69f00", size = 1) +  # dbh_pl line
#   draw_text("dbh_pl", x = 0.91, y = 0.75, hjust = 0, size = 12) +  # dbh_pl label
#   draw_line(x = c(0.88, 0.9), y = c(0.7, 0.7), color = "#009e73", size = 1) +   # dbh2_pl line
#   draw_text("dbh2_pl", x = 0.91, y = 0.7, hjust = 0, size = 12) +  # dbh2_pl label
#   draw_line(x = c(0.88, 0.9), y = c(0.65, 0.65), color = "#56b4e9", size = 1) +  # dbh3_pl line
#   draw_text("dbh3_pl", x = 0.91, y = 0.65, hjust = 0, size = 12)  # dbh3_pl label

```

```{r}
p1
p2
p3
p4
p5
p6
```

# Combine in 1 plot

```{r}
# p <- (
#   p1 + labs(tag = "(a)", x = NULL, y = "Predicted AGB_h_wb (kg)") + 
#     theme(
#       axis.title = element_text(size = 14, margin = margin(t = 10, r = 10, b = 10, l = 10)),  # Adjust title margin
#       axis.text = element_text(size = 12),   # Adjust axis text size
#       plot.tag = element_text(face = "bold", size = 14)  # Bold and resize plot tag
#     ) |

#   p2 + labs(tag = "(b)", x = NULL, y = "Predicted AGB_h_pl (kg)") + 
#     theme(
#       axis.title = element_text(size = 14, margin = margin(t = 10, r = 10, b = 10, l = 10)),
#       axis.text = element_text(size = 12),
#       plot.tag = element_text(face = "bold", size = 14)
#     ) |

#   p3 + labs(tag = "(c)", x = NULL, y = "Predicted AGB_dbh_pl (kg)") + 
#     theme(
#       axis.title = element_text(size = 16, margin = margin(t = 10, r = 10, b = 10, l = 10)),
#       axis.text = element_text(size = 12),
#       plot.tag = element_text(face = "bold", size = 14)
#     ) |

#   p4 + labs(tag = "(d)", x = "AGB_obs (kg)", y = "Predicted AGB_dbh2_pl (kg)") + 
#     theme(
#       axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 10, b = 20, l = 10)),  # Increased bottom margin for distance
#       axis.text = element_text(size = 12),
#       plot.tag = element_text(face = "bold", size = 14)
#     ) |

#   p5 + labs(tag = "(e)", x = "AGB_obs (kg)", y = "Predicted AGB_dbh3_pl (kg)") + 
#     theme(
#       axis.title.x = element_text(size = 16, margin = margin(t = 10, r = 10, b = 20, l = 10)),  # Increased bottom margin for distance
#       axis.text = element_text(size = 12),
#       plot.tag = element_text(face = "bold", size = 14)
#     ) |

#   p6 + labs(tag = "(f)", x = "Error (kg)", y = "Density") + 
#     theme(
#       axis.title = element_text(size = 14, margin = margin(t = 10, r = 10, b = 10, l = 10)),
#       axis.text = element_text(size = 12),
#       plot.tag = element_text(face = "bold", size = 14)
#     )
# ) + 
#   plot_layout(ncol = 2, nrow = 3) + 
#   plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"))) & 
#   theme(
#     plot.tag = element_text(face = "bold", size = 14),  # Set tag text to bold and larger
#     axis.title = element_text(size = 14),  # Set axis title text size
#     axis.text = element_text(size = 12),   # Set axis label text size
#     plot.margin = margin(5, 10, 5, 5)       # Set plot margins
#   )

# ggsave(
#   filename = "agb2.png",  # Specify the filename and format
#   plot = p,                             # Pass the combined plot object
#   width = 13,                           # Set the width of the plot in inches
#   height = 16,                          # Set the height of the plot in inches
#   dpi = 300                             # Set the resolution (300 for high quality)
# )
```

```{r}
p <- (
  p1 + labs(tag = "(a)", x = NULL, y = "Predicted AGB_h_wb (kg)") + 
    theme(
      axis.title = element_text(size = 14, margin = margin(t = 10, r = 10, b = 10, l = 10)),  # Adjust title margin
      axis.text = element_text(size = 12),   # Adjust axis text size
      plot.tag = element_text(face = "bold", size = 14)  # Bold and resize plot tag
    ) |

  p2 + labs(tag = "(b)", x = NULL, y = "Predicted AGB_h_pl (kg)") + 
    theme(
      axis.title = element_text(size = 14, margin = margin(t = 10, r = 10, b = 10, l = 10)),
      axis.text = element_text(size = 12),
      plot.tag = element_text(face = "bold", size = 14)
    ) |

  p3 + labs(tag = "(c)", x = NULL, y = "Predicted AGB_dbh_pl (kg)") + 
    theme(
      axis.title = element_text(size = 16, margin = margin(t = 10, r = 10, b = 10, l = 10)),
      axis.text = element_text(size = 12),
      plot.tag = element_text(face = "bold", size = 14)
    ) |

  p4 + labs(tag = "(d)", x = "AGB_obs (kg)", y = "Predicted AGB_dbh2_pl (kg)") + 
    theme(
      axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 10, b = 20, l = 10)),  # Increased bottom margin for distance
      axis.text = element_text(size = 12),
      plot.tag = element_text(face = "bold", size = 14)
    ) |

  p5 + labs(tag = "(e)", x = "AGB_obs (kg)", y = "Predicted AGB_dbh3_pl (kg)") + 
    theme(
      axis.title.x = element_text(size = 16, margin = margin(t = 10, r = 10, b = 20, l = 10)),  # Increased bottom margin for distance
      axis.text = element_text(size = 12),
      plot.tag = element_text(face = "bold", size = 14)
    )
) + 
  plot_layout(ncol = 2, nrow = 3) + 
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)"))) & 
  theme(
    plot.tag = element_text(face = "bold", size = 14),  # Set tag text to bold and larger
    axis.title = element_text(size = 14),  # Set axis title text size
    axis.text = element_text(size = 12),   # Set axis label text size
    plot.margin = margin(5, 10, 5, 5)       # Increased right margin to 50
  )

ggsave(
  filename = "agb2.png",  # Specify the filename and format
  plot = p,                             # Pass the combined plot object
  width = 13,                           # Set the width of the plot in inches
  height = 16,                          # Set the height of the plot in inches
  dpi = 300                             # Set the resolution (300 for high quality)
)
```

```{r}
options(width = 140)

calculate_agb <- function(agb_df) {
  agb_df |>
    mutate(
      # Existing biomass calculations:
      # Baseline biomass calculation
      log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

      # Weibull-based height estimation and biomass calculation
      height_wb = as.numeric(a_h_wb) * (1 - exp(-as.numeric(b_h_wb) * dbh^as.numeric(k_h_wb))),
      log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),

      # Power-law based height estimation and biomass calculation
      height_pl = as.numeric(a_h_pl) * dbh^as.numeric(b_h_pl),
      log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),

      # Power-law DBH estimation using CR and H, and biomass calculation
      dbh_pl = as.numeric(a_dbh) * (cr^as.numeric(b_dbh)) * (h^as.numeric(c_dbh)),
      log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,

      # Power-law DBH estimation using CR only, and biomass calculation
      dbh_pl2 = as.numeric(a_dbh2) * (cr^as.numeric(b_dbh2)),
      log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,

      # Power-law DBH estimation using H only, and biomass calculation
      dbh_pl3 = as.numeric(a_dbh3) * (h^as.numeric(b_dbh3)),
      log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h,

      # New biomass calculation using observed CR and H
      log_AGB_obs_CR_H = log(0.0559) + log_wd + 2 * log((2*cr) * h) + log(h)
    ) |>
    dplyr::select(
      tree_id, sp, log_dbh, log_h, log_wd, 
      log_AGB_bl, log_AGB_wb_h, log_AGB_pl_h, 
      log_AGB_pl_dbh, log_AGB_pl_dbh2, log_AGB_pl_dbh3,
      log_AGB_obs_CR_H # Include the new AGB calculation
    )
}

agb_df
```

```{r}
fit_lr_nou_summary_pl_ang_h <- tar_read(fit_lr_nou_summary_pl_ang_h)
fit_lr_nou_summary_pl_ang_h

fit_lr_nou_summary_pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
fit_lr_nou_summary_pl_gym_h

```

### 2025-03-26 docs/agb_dat_check.qmd
Initial commit for tree-allo

---
title: "Angiosperms' AGB data checking
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r include=FALSE}
library(tidyverse)
library(here)
library(targets)
options(width=200)
library(ggplot2)
```

```{r}
tallo_wd_df0 <- tar_read(tallo_wd_df0)
tallo_wd_df0
colnames(tallo_wd_df0)

sp_posterior_agb_df <- tar_read(sp_posterior_agb_df)

tallo_wd_df <- tallo_wd_df0 |>
    filter(!is.na(wd)) |>
    # filter( dbh >= 5) |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr),
           log_wd = log(wd)) |>
    filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
    dplyr::select(tree_id, division, biome, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
    filter( dbh >= 5) |>
    group_by(sp) |>
    filter(n() >= 20) |>
    ungroup()

tallo_wd_ang <- tallo_wd_df |>
  filter(division == "Angiosperm")

# Posterior data into Angiosperms
sp_posterior_ang <- sp_posterior_agb_df |> filter(Division == "Angiosperm")

filter_common_species <- function(tallo_wd_df, sp_posterior_agb_df) {
  h_wb_df <- sp_posterior_agb_df |>
  filter(Dependent_variable == "Tree Height", !is.na(a), !is.na(b),!is.na(k)) |>
    dplyr::select(sp, a, b, k) |>
    distinct()

  h_pl_df <- sp_posterior_agb_df |>
    filter(Dependent_variable == "Tree Height", is.na(k)) |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh_df <- sp_posterior_agb_df |>
    filter(Dependent_variable == "DBH", !is.na(a), !is.na(b), !is.na(c)) |>
    dplyr::select(sp, a, b, c) |>
    distinct()

  dbh1_df <- sp_posterior_agb_df |>
    filter(Dependent_variable == "DBH1") |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh2_df <- sp_posterior_agb_df |>
    filter(Dependent_variable == "DBH2") |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh3_df <- sp_posterior_agb_df |>
    filter(Dependent_variable == "DBH3") |>
    dplyr::select(sp, a, b) |>
    distinct()

  common_species <- Reduce(intersect, list(
    tallo_wd_df$sp,
    h_wb_df$sp,
    h_pl_df$sp,
    dbh_df$sp,
    dbh1_df$sp,
    dbh2_df$sp,
    dbh3_df$sp
  ))

  list(
    filtered_tallo_wd_df = tallo_wd_df |> filter(sp %in% common_species),
    filtered_h_wb_df = h_wb_df |> filter(sp %in% common_species),
    filtered_h_pl_df = h_pl_df |> filter(sp %in% common_species),
    filtered_dbh_df = dbh_df |> filter(sp %in% common_species),
    filtered_dbh1_df = dbh1_df |> filter(sp %in% common_species),
    filtered_dbh2_df = dbh2_df |> filter(sp %in% common_species),
    filtered_dbh3_df = dbh3_df |> filter(sp %in% common_species),
    num_species = length(common_species),
    num_trees = tallo_wd_df |> filter(sp %in% common_species) |> nrow()

  )
}

# Filter common species for Angiosperms and Gymnosperms
filtered_ang <- filter_common_species(tallo_wd_ang, sp_posterior_ang)

calculate_agb <- function(filtered_data) {
    tallo_wd_df <- filtered_data$filtered_tallo_wd_df
    agb_df <- tallo_wd_df |>
      left_join(
        filtered_data$filtered_h_wb_df |> rename(a_h_wb = a, b_h_wb = b, k_h_wb = k),
        by = "sp"
      ) |>
      left_join(
        filtered_data$filtered_h_pl_df |> rename(a_h_pl = a, b_h_pl = b),
        by = "sp"
      ) |>
      left_join(
        filtered_data$filtered_dbh_df |> rename(a_dbh = a, b_dbh = b, c_dbh = c),
        by = "sp"
      ) |>
      left_join(
        filtered_data$filtered_dbh1_df |> rename(a_dbh1 = a, b_dbh1 = b),
        by = "sp"
      ) |>
      left_join(
        filtered_data$filtered_dbh2_df |> rename(a_dbh2 = a, b_dbh2 = b),
        by = "sp"
      ) |>
      left_join(
        filtered_data$filtered_dbh3_df |> rename(a_dbh3 = a, b_dbh3 = b),
        by = "sp"
      )

    # Calculate the log_AGB based on various models
    agb_df <- agb_df |>
      mutate(
        a_h_wb = as.numeric(gsub(" \\(.*\\)", "", a_h_wb)),
        b_h_wb = as.numeric(gsub(" \\(.*\\)", "", b_h_wb)),
        k_h_wb = as.numeric(gsub(" \\(.*\\)", "", k_h_wb)),
        a_h_pl = as.numeric(gsub(" \\(.*\\)", "", a_h_pl)),
        b_h_pl = as.numeric(gsub(" \\(.*\\)", "", b_h_pl)),
        a_dbh = as.numeric(gsub(" \\(.*\\)", "", a_dbh)),
        b_dbh = as.numeric(gsub(" \\(.*\\)", "", b_dbh)),
        c_dbh = as.numeric(gsub(" \\(.*\\)", "", c_dbh)),
        a_dbh1 = as.numeric(gsub(" \\(.*\\)", "", a_dbh1)),
        b_dbh1 = as.numeric(gsub(" \\(.*\\)", "", b_dbh1)),
        a_dbh2 = as.numeric(gsub(" \\(.*\\)", "", a_dbh2)),
        b_dbh2 = as.numeric(gsub(" \\(.*\\)", "", b_dbh2)),
        a_dbh3 = as.numeric(gsub(" \\(.*\\)", "", a_dbh3)),
        b_dbh3 = as.numeric(gsub(" \\(.*\\)", "", b_dbh3))
      ) |>
      mutate(
        # Updated baseline AGB calculation with new equation
        log_AGB_bl = log(0.0673) + 0.976 * (log_wd + 2 * log_dbh + log_h),

        height_wb = a_h_wb * (1 - exp(-b_h_wb * dbh^k_h_wb)),    # Weibull-based height estimation
        log_AGB_wb_h = log(0.0673) + 0.976 * (log_wd + 2 * log_dbh + log(height_wb)),  # Biomass from Weibull height

        height_pl = a_h_pl * dbh^b_h_pl,  # Power-law height estimation
        log_AGB_pl_h = log(0.0673) + 0.976 * (log_wd + 2 * log_dbh + log(height_pl)),  # Biomass from Power-law height

        dbh_pl = a_dbh * (cr^b_dbh) * (h^c_dbh),  # Power-law DBH estimation using CR and H
        log_AGB_pl_dbh = log(0.0673) + 0.976 * (log_wd + 2 * log(dbh_pl) + log_h),  # Biomass from Power-law DBH

        dbh_pl1 = a_dbh1 * (cr * h)^b_dbh1,  # Power-law DBH estimation using CR*H
        log_AGB_pl_dbh1 = log(0.0673) + 0.976 * (log_wd + 2 * log(dbh_pl1) + log_h),  # Biomass from Power-law DBH1

        dbh_pl2 = a_dbh2 * (cr^b_dbh2),  # Power-law DBH estimation using CR only
        log_AGB_pl_dbh2 = log(0.0673) + 0.976 * (log_wd + 2 * log(dbh_pl2) + log_h),  # Biomass from Power-law DBH2

        dbh_pl3 = a_dbh3 * (h^b_dbh3),  # Power-law DBH estimation using H only
        log_AGB_pl_dbh3 = log(0.0673) + 0.976 * (log_wd + 2 * log(dbh_pl3) + log_h)  # Biomass from Power-law DBH3
      )

    return(agb_df)
  }

# Calculate AGB for Angiosperms
agb_ang <- calculate_agb(filtered_ang)

```

```{r}
agb_ang
summary(agb_ang)

# agb_ang <- agb_ang |>
#   mutate(basal_area = pi * (dbh / 2)^2)

large_tree <- agb_ang |>
  filter(dbh >=  70)

num_trees_large_dbh <- nrow(large_tree)

percent_large_dbh <- (num_trees_large_dbh / nrow(agb_ang)) * 100

total_biomass <- sum(agb_ang$log_AGB_bl, na.rm = TRUE)

large_tree_biomass <- sum(large_tree$log_AGB_bl, na.rm = TRUE)

percent_biomass_large_trees <- (large_tree_biomass / total_biomass) * 100

num_trees_large_dbh
percent_large_dbh
large_tree_biomass
total_biomass
percent_biomass_large_trees

```

# Collinearity checking

```{r}
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)

cor(tallo_reduced_lr_df_ang_dbh$h, tallo_reduced_lr_df_ang_dbh$cr, use = "complete.obs")

ggplot(tallo_reduced_lr_df_ang_dbh, aes(x = h, y = cr)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Scatterplot of Height vs Crown Radius",
       x = "Tree Height (h)",
       y = "Crown Radius (cr)")

```

### 2025-03-26 docs/agb_new.qmd
Initial commit for tree-allo

---
title: "AGB estimation"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r include=FALSE}
library(tidyverse)
library(here)
library(targets)
library(knitr)
library(scales)
library(patchwork)
library(cowplot)
options(width=200)

```

# Loading data

```{r}
# sp_posterior_df <- tar_read(sp_posterior_agb_df)
# head(sp_posterior_df, n = 6)

sp_posterior_df <- tar_read(test_agb_post)
head(sp_posterior_df, n = 6)

tallo_wd_df0 <- tar_read(tallo_wd_df0)

tallo_wd_df0
summary(tallo_wd_df0)

# Remove sp < 20 obs
tallo_wd_df <- tallo_wd_df0 |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
         log_h = log(h),
         log_cr = log(cr),
         log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  select(tree_id, division, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd_df

# common_sp <- intersect(tallo_wd_df$sp, sp_posterior_df$sp)
# tallo_wd_df <- tallo_wd_df |> filter(sp %in% common_sp)
# sp_posterior_df <- sp_posterior_df |> filter(sp %in% common_sp)

tallo_wd_ang <- tallo_wd_df |>
  filter(division == "Angiosperm")
tallo_wd_ang

tallo_wd_gym <- tallo_wd_df |>
  filter(division == "Gymnosperm")
tallo_wd_gym

sp_posterior_ang <- sp_posterior_df |>
  filter(Division == "Angiosperm")

sp_posterior_gym <- sp_posterior_df |>
  filter(Division == "Gymnosperm")

```

# Parameter extraction

## Angiosperm

```{r}
calculate_agb <- function(tallo_wd_df, sp_posterior_df) {
  h_wb_df <- sp_posterior_df |>
    filter(Dependent_variable == "Tree Height") |>
    dplyr::select(sp, a, b, k, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  h_pl_df <- sp_posterior_df |>
    filter(Dependent_variable == "Tree Height", is.na(k)) |>
    dplyr::select(sp, a, b, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  dbh_df <- sp_posterior_df |>
    filter(Dependent_variable == "DBH") |>
    dplyr::select(sp, a, b, c, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  dbh1_df <- sp_posterior_df |>
    filter(Dependent_variable == "DBH1") |>
    dplyr::select(sp, a, b, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  dbh2_df <- sp_posterior_df |>
    filter(Dependent_variable == "DBH2") |>
    dplyr::select(sp, a, b, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  dbh3_df <- sp_posterior_df |>
    filter(Dependent_variable == "DBH3") |>
    dplyr::select(sp, a, b, Division) |>
    distinct(sp, .keep_all = TRUE) |>
    arrange(sp)

  # Join the posterior data with tallo_wd_df data and calculate AGB
  agb_df <- tallo_wd_df |>
    left_join(h_wb_df |> rename(a_h_wb = a, b_h_wb = b, k_h_wb = k), by = "sp") |> 
    left_join(h_pl_df |> rename(a_h_pl = a, b_h_pl = b), by = "sp") |> 
    left_join(dbh_df |> rename(a_dbh = a, b_dbh = b, c_dbh = c), by = "sp") |>
    left_join(dbh1_df |> rename(a_dbh1 = a, b_dbh1 = b), by = "sp") |> 
    left_join(dbh2_df |> rename(a_dbh2 = a, b_dbh2 = b), by = "sp") |> 
    left_join(dbh3_df |> rename(a_dbh3 = a, b_dbh3 = b), by = "sp")

  # Calculate the log_AGB based on various models
  agb_df <- agb_df |>
    mutate(
      a_h_wb = as.numeric(gsub(" .*", "", a_h_wb)),
      b_h_wb = as.numeric(gsub(" .*", "", b_h_wb)),
      k_h_wb = as.numeric(gsub(" .*", "", k_h_wb)),
      a_h_pl = as.numeric(gsub(" .*", "", a_h_pl)),
      b_h_pl = as.numeric(gsub(" .*", "", b_h_pl)),
      a_dbh = as.numeric(gsub(" .*", "", a_dbh)),
      b_dbh = as.numeric(gsub(" .*", "", b_dbh)),
      c_dbh = as.numeric(gsub(" .*", "", c_dbh)),
      a_dbh1 = as.numeric(gsub(" .*", "", a_dbh1)),
      b_dbh1 = as.numeric(gsub(" .*", "", b_dbh1)),
      a_dbh2 = as.numeric(gsub(" .*", "", a_dbh2)),
      b_dbh2 = as.numeric(gsub(" .*", "", b_dbh2)),
      a_dbh3 = as.numeric(gsub(" .*", "", a_dbh3)),
      b_dbh3 = as.numeric(gsub(" .*", "", b_dbh3))
    ) |>
    mutate(
      log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

      height_wb = a_h_wb * (1 - exp(-b_h_wb * dbh^k_h_wb)),    # Weibull-based height estimation
      log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),  # Biomass from Weibull height

      height_pl = a_h_pl * dbh^b_h_pl,  # Power-law height estimation
      log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),  # Biomass from Power-law height

      dbh_pl = a_dbh * (cr^b_dbh) * (h^c_dbh),  # Power-law DBH estimation using CR and H
      log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,  # Biomass from Power-law DBH

      dbh_pl1 = a_dbh1 * (cr * h)^b_dbh1,  # Power-law DBH estimation using CR*H
      log_AGB_pl_dbh1 = log(0.0559) + log_wd + 2 * log(dbh_pl1) + log_h,  # Biomass from Power-law DBH1

      dbh_pl2 = a_dbh2 * (cr^b_dbh2),  # Power-law DBH estimation using CR only
      log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,  # Biomass from Power-law DBH2

      dbh_pl3 = a_dbh3 * (h^b_dbh3),  # Power-law DBH estimation using H only
      log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h  # Biomass from Power-law DBH3
    )
}

agb_ang <- calculate_agb(tallo_wd_ang, sp_posterior_ang)

agb_gym <- calculate_agb(tallo_wd_gym, sp_posterior_gym)

```

# Box plot

```{r}
#ANG
agb_ang_long <- agb_ang |>
  tidyr::pivot_longer(cols = starts_with("log_AGB"), names_to = "Scenario", values_to = "AGB")

ggplot(agb_ang_long, aes(x = Scenario, y = AGB)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Angiosperm",
    x = "Senarios",
    y = "log predicted AGB (kg)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# GYM
agb_gym_long <- agb_gym |>
  tidyr::pivot_longer(cols = starts_with("log_AGB"), names_to = "Scenario", values_to = "AGB")

ggplot(agb_gym_long, aes(x = Scenario, y = AGB)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Gymnosperm",
    x = "Senarios",
    y = "log predicted AGB (kg)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Performance metrics

```{r}
# combine_metrics <- function(agb_df) {
#   agb_df |>
#     summarize(
#       rmse_wb_h = sqrt(mean((log_AGB_bl - log_AGB_wb_h)^2, na.rm = TRUE)),
#       bias_wb_h = mean(log_AGB_wb_h - log_AGB_bl, na.rm = TRUE),

#       rmse_pl_h = sqrt(mean((log_AGB_bl - log_AGB_pl_h)^2, na.rm = TRUE)),
#       bias_pl_h = mean(log_AGB_pl_h - log_AGB_bl, na.rm = TRUE),

#       rmse_pl_dbh = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh)^2, na.rm = TRUE)),
#       bias_pl_dbh = mean(log_AGB_pl_dbh - log_AGB_bl, na.rm = TRUE),

#       rmse_pl_dbh1 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh1)^2, na.rm = TRUE)),
#       bias_pl_dbh1 = mean(log_AGB_pl_dbh1 - log_AGB_bl, na.rm = TRUE),

#       rmse_pl_dbh2 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh2)^2, na.rm = TRUE)),
#       bias_pl_dbh2 = mean(log_AGB_pl_dbh2 - log_AGB_bl, na.rm = TRUE),

#       rmse_pl_dbh3 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh3)^2, na.rm = TRUE)),
#       bias_pl_dbh3 = mean(log_AGB_pl_dbh3 - log_AGB_bl, na.rm = TRUE)
#     )
# }

combine_metrics <- function(agb_df) {
  agb_df |>
    summarize(
      # Weibull-based height metrics
      mae_wb_h = mean(abs(log_AGB_bl - log_AGB_wb_h), na.rm = TRUE),
      rmse_wb_h = sqrt(mean((log_AGB_bl - log_AGB_wb_h)^2, na.rm = TRUE)),
      bias_wb_h = mean(log_AGB_wb_h - log_AGB_bl, na.rm = TRUE),
      r2_wb_h = 1 - sum((log_AGB_wb_h - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law height metrics
      mae_pl_h = mean(abs(log_AGB_bl - log_AGB_pl_h), na.rm = TRUE),
      rmse_pl_h = sqrt(mean((log_AGB_bl - log_AGB_pl_h)^2, na.rm = TRUE)),
      bias_pl_h = mean(log_AGB_pl_h - log_AGB_bl, na.rm = TRUE),
      r2_pl_h = 1 - sum((log_AGB_pl_h - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH metrics
      mae_pl_dbh = mean(abs(log_AGB_bl - log_AGB_pl_dbh), na.rm = TRUE),
      rmse_pl_dbh = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh)^2, na.rm = TRUE)),
      bias_pl_dbh = mean(log_AGB_pl_dbh - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh = 1 - sum((log_AGB_pl_dbh - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH1 metrics
      mae_pl_dbh1 = mean(abs(log_AGB_bl - log_AGB_pl_dbh1), na.rm = TRUE),
      rmse_pl_dbh1 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh1)^2, na.rm = TRUE)),
      bias_pl_dbh1 = mean(log_AGB_pl_dbh1 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh1 = 1 - sum((log_AGB_pl_dbh1 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH2 metrics
      mae_pl_dbh2 = mean(abs(log_AGB_bl - log_AGB_pl_dbh2), na.rm = TRUE),
      rmse_pl_dbh2 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh2)^2, na.rm = TRUE)),
      bias_pl_dbh2 = mean(log_AGB_pl_dbh2 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh2 = 1 - sum((log_AGB_pl_dbh2 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH3 metrics
      mae_pl_dbh3 = mean(abs(log_AGB_bl - log_AGB_pl_dbh3), na.rm = TRUE),
      rmse_pl_dbh3 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh3)^2, na.rm = TRUE)),
      bias_pl_dbh3 = mean(log_AGB_pl_dbh3 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh3 = 1 - sum((log_AGB_pl_dbh3 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)
    )
}

# ANG 
metrics_ang <- combine_metrics(agb_ang)
head(metrics_ang)

metrics_ang_long <- metrics_ang |>
  pivot_longer(cols = everything(), names_to = "Metric_Model", values_to = "Value") |>
  separate(Metric_Model, into = c("Metric", "Model"), sep = "_", extra = "merge") |>
  filter(Metric != "bias")  # Exclude the bias metric

ggplot(metrics_ang_long, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Angiosperm",
    x = "Senarios",
    y = "Metric Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set2")

# GYM

metrics_gym <- combine_metrics(agb_gym)
head(metrics_gym)

metrics_gym_long <- metrics_gym |>
  pivot_longer(cols = everything(), names_to = "Metric_Model", values_to = "Value") |>
  separate(Metric_Model, into = c("Metric", "Model"), sep = "_", extra = "merge") |>
  filter(Metric != "bias")  # Exclude the bias metric

ggplot(metrics_gym_long, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Gymnosperm",
    x = "Senarios",
    y = "Metric Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set2")

```

# Predicted vs. baseline AGB plots

```{r}
plot_agb <- function(agb_df, metrics_df, model_colors, division_label) {
  create_plot <- function(data, baseline, predicted, rmse, bias, model_name, tag_label, x_axis_label = NULL) {
      ggplot(data, aes(x = exp(baseline), y = exp(predicted))) +
        geom_point(alpha = 0.5, color = "grey", size = 0.3) +
        geom_smooth(method = "lm", se = FALSE, color = model_colors[model_name], size = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black", size = 0.5) +
        scale_x_log10(
          breaks = c(0.1, 10, 1000, 100000),
          labels = function(x) ifelse(x == 0.1, "0.1", scales::label_number(accuracy = 1)(x))
        ) +
        scale_y_log10(
          breaks = c(0.1, 10, 1000, 100000),
          limits = c(NA, 100000), 
          labels = function(x) ifelse(x == 0.1, "0.1", scales::label_number(accuracy = 1)(x))
        ) +
        labs(tag = tag_label, 
            x = x_axis_label, 
            y = switch(model_name,
                        "h_wb" = expression(Predicted~AGB[H*"\u00A0"*"-"*"\u00A0"*WB]~"(kg)"),
                        "h_pl" = expression(Predicted~AGB[H*"\u00A0"*"-"*"\u00A0"*PL]~"(kg)"),
                        "dbh_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR*"\u002B"*H]~"(kg)"),
                        "dbh1_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR*"\u00D7"*H]~"(kg)"),
                        "dbh2_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR]~"(kg)"),
                        "dbh3_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*H]~"(kg)"))) +
        # Reduce size of RMSE and Bias annotations
        annotate("text", x = 0.9, y = 1e5, label = paste("RMSE =", format(round(rmse, 2), nsmall = 2), "Mg"), hjust = 0, size = 2.5) +
        annotate("text", x = 0.9, y = 3e4, label = paste("Bias =", format(round(bias * 100, 2), nsmall = 2), "%"), hjust = 0, size = 2.5) +
        theme_minimal() +
        theme(plot.margin = margin(5, 15, 5, 5)) +
        theme(axis.title = element_text(size = 9), 
              plot.tag = element_text(face = "bold", size = 10), 
              axis.text = element_text(size = 8))  # Reduce the axis text size
  }
    model_colors <- c(
      "h_wb" = "#D55E00",
      "h_pl" = "#0072B2",
      "dbh_pl" = "#E69F00",
      "dbh1_pl" = "#F0E442",
      "dbh2_pl" = "#009E73",
      "dbh3_pl" = "#56B4E9"
    )

    p1 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_wb_h, 
                      metrics_df$rmse_wb_h, metrics_df$bias_wb_h, "h_wb", "(a)")

    p2 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_h, 
                      metrics_df$rmse_pl_h, metrics_df$bias_pl_h, "h_pl", "(b)")

    p3 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh, 
                      metrics_df$rmse_pl_dbh, metrics_df$bias_pl_dbh, "dbh_pl", "(c)")

    p4 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh1, 
                      metrics_df$rmse_pl_dbh1, metrics_df$bias_pl_dbh1, "dbh1_pl", "(d)")

    p5 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh2, 
                      metrics_df$rmse_pl_dbh2, metrics_df$bias_pl_dbh2, "dbh2_pl", "(e)",
                      x_axis_label = expression(AGB[obs]~"(kg)"))

    p6 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh3, 
                      metrics_df$rmse_pl_dbh3, metrics_df$bias_pl_dbh3, "dbh3_pl", "(f)",
                      x_axis_label = expression(AGB[obs]~"(kg)"))

    # Combine plots
    plot <- (p1 | p2 | p3 | p4 | p5 | p6) + 
      plot_layout(ncol = 2, nrow = 3) + 
      plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)")))

    plot_with_label <- ggdraw() +
    draw_plot(plot, x = 0, y = 0, width = 1, height = 0.9) +  
    draw_label(division_label, x = 0.5, y = 0.92, fontface = 'bold', size = 10)

  return(plot_with_label)
}
```

```{r}
# Plot for Angiosperms
plot_ang <- plot_agb(agb_ang, metrics_ang, model_colors, "Angiosperm")
plot_ang
```

```{r, height = 20, width = 10}
# Plot for Gymnosperms
plot_gym <- plot_agb(agb_gym, metrics_gym, model_colors, "Gymnosperm")
plot_gym
```

# CHECKING FUNCTION

```{r}
tallo_wd_df0 <- tar_read(tallo_wd_df0)

tallo_wd_df <- tallo_wd_df0 |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
          log_h = log(h),
          log_cr = log(cr),
          log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  dplyr::select(tree_id, division, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd_df

# Wood density for Angiosperms
wd_df_ang <- tallo_wd_df |>
  filter(division == "Angiosperm") |>
  group_by(sp) |>
  summarize(wd = mean(wd, na.rm = TRUE)) |>
  mutate(sp_id = row_number())
wd_df_ang

# Wood density for Gymnosperms
wd_df_gym <- tallo_wd_df |>
  filter(division == "Gymnosperm") |>
  group_by(sp) |>
  summarize(wd = mean(wd, na.rm = TRUE)) |>
  mutate(sp_id = row_number())

wd_df_gym

```

```{r}
# Read the datasets
tallo_wd_df0 <- tar_read(tallo_wd_df0)

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
# tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
tallo_reduced_lr_df_ang_dbh1 <- tar_read(tallo_reduced_lr_df_ang_dbh1)
tallo_reduced_lr_df_ang_dbh2 <- tar_read(tallo_reduced_lr_df_ang_dbh2)
tallo_reduced_lr_df_ang_dbh3 <- tar_read(tallo_reduced_lr_df_ang_dbh3)

tallo_reduced_nlr_df_gym_h <- tar_read(tallo_reduced_nlr_df_gym_h)
tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
# tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)
tallo_reduced_lr_df_gym_dbh <- tar_read(tallo_reduced_lr_df_gym_dbh)
tallo_reduced_lr_df_gym_dbh1 <- tar_read(tallo_reduced_lr_df_gym_dbh1)
tallo_reduced_lr_df_gym_dbh2 <- tar_read(tallo_reduced_lr_df_gym_dbh2)
tallo_reduced_lr_df_gym_dbh3 <- tar_read(tallo_reduced_lr_df_gym_dbh3)

# ANG
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_lr_nou_summary_pl_ang_h <- tar_read(fit_lr_nou_summary_pl_ang_h)
# fit_lr_nou_summary_pl_ang_cr <- tar_read(fit_lr_nou_summary_pl_ang_cr)
fit_lr_nou_summary_pl_ang_dbh <- tar_read(fit_lr_nou_summary_pl_ang_dbh)
fit_lr_nou_summary_pl_ang_dbh1 <- tar_read(fit_lr_nou_summary_pl_ang_dbh1)
fit_lr_nou_summary_pl_ang_dbh2 <- tar_read(fit_lr_nou_summary_pl_ang_dbh2)
fit_lr_nou_summary_pl_ang_dbh3 <- tar_read(fit_lr_nou_summary_pl_ang_dbh3)

stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
stan_data_lr_ang_h <- tar_read(stan_data_lr_ang_h)
# stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
stan_data_lr_ang_dbh1 <- tar_read(stan_data_lr_ang_dbh1)
stan_data_lr_ang_dbh2 <- tar_read(stan_data_lr_ang_dbh2)
stan_data_lr_ang_dbh3 <- tar_read(stan_data_lr_ang_dbh3)

str(stan_data_nlr_ang_h)
str(stan_data_lr_ang_h)

str(stan_data_lr_ang_dbh)
str(stan_data_lr_ang_dbh1)
str(stan_data_lr_ang_dbh2)
str(stan_data_lr_ang_dbh3)

# GYM
fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_lr_nou_summary_pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
# fit_nlr_nou_summary_gmm_gym_cr <- tar_read(fit_nlr_nou_summary_gmm_gym_cr)
fit_lr_nou_summary_pl_gym_dbh <- tar_read(fit_lr_nou_summary_pl_gym_dbh)
fit_lr_nou_summary_pl_gym_dbh1 <- tar_read(fit_lr_nou_summary_pl_gym_dbh1)
fit_lr_nou_summary_pl_gym_dbh2 <- tar_read(fit_lr_nou_summary_pl_gym_dbh2)
fit_lr_nou_summary_pl_gym_dbh3 <- tar_read(fit_lr_nou_summary_pl_gym_dbh3)

stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_gym_h <- tar_read(stan_data_lr_gym_h)
# stan_data_nlr_gym_cr <- tar_read(stan_data_nlr_gym_cr)
stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
stan_data_lr_gym_dbh1 <- tar_read(stan_data_lr_gym_dbh1)
stan_data_lr_gym_dbh2 <- tar_read(stan_data_lr_gym_dbh2)
stan_data_lr_gym_dbh3 <- tar_read(stan_data_lr_gym_dbh3)

str(stan_data_nlr_gym_h)
str(stan_data_lr_gym_h)
str(stan_data_lr_gym_dbh)
str(stan_data_lr_gym_dbh1)
str(stan_data_lr_gym_dbh2)
str(stan_data_lr_gym_dbh3)

```

```{r}
filter_species <- function(dataset, division) {
  dataset |> 
    filter(division == division, !is.na(dbh), !is.na(cr), !is.na(h)) |> 
    group_by(sp) |> 
    summarise(count = n()) |> 
    filter(count >= 20) |> 
    pull(sp)
}
create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
  if (functional_form == "Weibull" || functional_form == "gMM") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

    k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
    k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
    k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  } else if (functional_form == "Power-Law") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

    c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
    c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
    c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  }

  return(ci_95)
}
# Function to process datasets and align species IDs with Stan data
process_dataset <- function(dataset, stan_data, division_filter) {
  # Filter the dataset by division
  dataset <- dataset |> filter(division == division_filter)

  # Prepare species data
  sp_data <- dataset |>
    group_by(sp) |>
    summarise(wd = mean(wd, na.rm = TRUE)) |>
    arrange(sp) |>
    mutate(sp_id = row_number()) |>
    mutate(wd_s = scale(wd) |> as.numeric()) |>
    select(sp_id, sp, wd, wd_s)

  # Verify Stan data structure
  if (!all(c("jj", "u") %in% names(stan_data))) {
    stop("Stan data must contain 'jj' and 'u'.")
  }

  # Handle `nlr` structure: `u` has rows indexed by `jj`
  if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
    stan_processed <- data.frame(
      sp_id = as.factor(stan_data$jj),
      wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
    )
  } 
  # Handle `lr` structure: `u` has two rows (variables) and columns for species
  else if (nrow(stan_data$u) == 2) {
    stan_processed <- data.frame(
      sp_id = 1:ncol(stan_data$u),
      wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
    )
  } 
  # Error handling for unexpected structures
  else {
    stop("Unexpected structure for 'u' in Stan data.")
  }

  # Add species names and calculate tree counts
  stan_processed <- stan_processed |>
    group_by(sp_id, wd_s) |>
    summarise(tree_count = n(), .groups = "drop") |>
    mutate(sp_id = as.integer(as.character(sp_id))) |>
    left_join(sp_data, by = "sp_id")

  return(list(sp_data = sp_data, stan_processed = stan_processed))
}
```

```{r}
ang_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Angiosperm")
gym_dbh_species <- filter_species(tallo_reduced_lr_df_gym_dbh, "Gymnosperm")

ang_dbh1_species <- filter_species(tallo_reduced_lr_df_ang_dbh1, "Angiosperm")
gym_dbh1_species <- filter_species(tallo_reduced_lr_df_gym_dbh1, "Gymnosperm")

ang_dbh2_species <- filter_species(tallo_reduced_lr_df_ang_dbh2, "Angiosperm")
gym_dbh2_species <- filter_species(tallo_reduced_lr_df_gym_dbh2, "Gymnosperm")

ang_dbh3_species <- filter_species(tallo_reduced_lr_df_ang_dbh3, "Angiosperm")
gym_dbh3_species <- filter_species(tallo_reduced_lr_df_gym_dbh3, "Gymnosperm")

h_ang_wb <- process_dataset(tallo_reduced_nlr_df_ang_h, stan_data_nlr_ang_h, "Angiosperm")
h_ang_wb

h_ang_pl <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_lr_ang_h, "Angiosperm")
h_ang_pl

cr_ang_pl <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_lr_ang_cr, "Angiosperm")
cr_ang_pl

dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% ang_dbh_species), stan_data_lr_ang_dbh, "Angiosperm")
dbh_ang

dbh1_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh1 |> filter(sp %in% ang_dbh1_species), stan_data_lr_ang_dbh1, "Angiosperm")
dbh1_ang

dbh2_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh2 |> filter(sp %in% ang_dbh2_species), stan_data_lr_ang_dbh2, "Angiosperm")
dbh2_ang

dbh3_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh3 |> filter(sp %in% ang_dbh3_species), stan_data_lr_ang_dbh3, "Angiosperm")
dbh3_ang

###

h_gym_wb <- process_dataset(tallo_reduced_nlr_df_gym_h, stan_data_nlr_gym_h, "Gymnosperm")
h_gym_wb

h_gym_pl <- process_dataset(tallo_reduced_lr_df_gym_h, stan_data_lr_ang_h, "Gymnosperm")
h_gym_pl

cr_gym_gmm <- process_dataset(tallo_reduced_nlr_df_gym_cr, stan_data_nlr_gym_cr, "Gymnosperm")
cr_gym_gmm

dbh_gym <- process_dataset(tallo_reduced_lr_df_gym_dbh |> filter(sp %in% gym_dbh_species), stan_data_lr_gym_dbh, "Gymnosperm")
dbh_gym

dbh1_gym <- process_dataset(tallo_reduced_lr_df_gym_dbh1 |> filter(sp %in% gym_dbh1_species), stan_data_lr_gym_dbh1, "Gymnosperm")
dbh1_gym

dbh2_gym <- process_dataset(tallo_reduced_lr_df_gym_dbh2 |> filter(sp %in% gym_dbh2_species), stan_data_lr_gym_dbh2, "Gymnosperm")
dbh2_gym

dbh3_gym <- process_dataset(tallo_reduced_lr_df_gym_dbh3 |> filter(sp %in% gym_dbh3_species), stan_data_lr_gym_dbh3, "Gymnosperm")
dbh3_gym

```

```{r}
# ANG
h_ang_wb_df <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang_wb$sp)
head(h_ang_wb_df)
summary(h_ang_wb_df)

h_ang_pl_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_h, "Tree Height", "Power-Law", h_ang_pl$sp)
head(h_ang_pl_df)
summary(h_ang_pl_df)

cr_ang_pl_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_cr, "Crown Radius", "Power-Law", cr_ang_pl$sp)
head(cr_ang_pl_df)
summary(cr_ang_pl_df)

dbh_ang_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh, "DBH", "Power-Law", dbh_ang$sp)
head(dbh_ang_df)
summary(dbh_ang_df)

dbh1_ang_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh1, "DBH1", "Power-Law", dbh1_ang$sp)
head(dbh1_ang_df)
summary(dbh1_ang_df)

dbh2_ang_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh2, "DBH2", "Power-Law", dbh2_ang$sp)
head(dbh2_ang_df)
summary(dbh2_ang_df)

dbh3_ang_df <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh3, "DBH3", "Power-Law", dbh3_ang$sp)
head(dbh3_ang_df)
summary(dbh3_ang_df)

# GYM

h_gym_wb_df <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_gym_wb$sp)
head(h_gym_wb_df)
summary(h_gym_wb_df)

h_gym_pl_df <- create_species_model_df(fit_lr_nou_summary_pl_gym_h, "Tree Height", "Power-Law", h_gym_pl$sp)
head(h_df_gym)
summary(h_df_gym)

cr_gym_gmm_df <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_cr, "Crown Radius", "gMM", cr_gym_gmm$sp)
head(cr_gym_gmm_df)
summary(cr_gym_gmm_df)

dbh_gym_df <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh, "DBH", "Power-Law", dbh_gym$sp)
head(dbh_gym_df)
summary(dbh_gym_df)

dbh1_gym_df <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh1, "DBH1", "Power-Law", dbh1_gym$sp)
head(dbh1_gym_df)
summary(dbh1_gym_df)

dbh2_gym_df <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh2, "DBH2", "Power-Law", dbh2_gym$sp)
head(dbh2_gym_df)
summary(dbh2_gym_df)

dbh3_gym_df <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh3, "DBH3", "Power-Law", dbh3_gym$sp)
head(dbh3_gym_df)
summary(dbh3_gym_df)

```

```{r}
# sp_posterior_agb_ang <- bind_rows(
#   h_ang_wb_df |> mutate(Division = "Angiosperm"),
#   h_ang_pl_df |> mutate(Division = "Angiosperm"),
#   cr_ang_pl_df |> mutate(Division = "Angiosperm"),
#   dbh_ang_df |> mutate(Division = "Angiosperm"),
#   dbh1_ang_df |> mutate(Division = "Angiosperm"),
#   dbh2_ang_df |> mutate(Division = "Angiosperm"),
#   dbh3_ang_df |> mutate(Division = "Angiosperm")
# )

# sp_posterior_agb_gym <- bind_rows(
#   h_gym_wb_df |> mutate(Division = "Gymnosperm"),
#   h_gym_pl_df |> mutate(Division = "Gymnosperm"),
#   cr_gym_gmm_df |> mutate(Division = "Gymnosperm"),
#   dbh_gym_df |> mutate(Division = "Gymnosperm"),
#   dbh1_gym_df |> mutate(Division = "Gymnosperm"),
#   dbh2_gym_df |> mutate(Division = "Gymnosperm"),
#   dbh3_gym_df |> mutate(Division = "Gymnosperm")
# )

# head(sp_posterior_agb_ang)
# head(sp_posterior_agb_gym)

```

## ANGIOSPERM
```{r}
head(sp_posterior_agb_ang)
summary(sp_posterior_agb_ang)
```

## CHECK PARAMETER IN DF AND CACULATE AGB

```{r}
h_ang_wb_df <- h_ang_wb_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(h_ang_wb_df)

h_ang_pl_df <- h_ang_pl_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(h_ang_pl_df)

dbh_ang_df <- dbh_ang_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh_ang_df)

dbh1_ang_df <- dbh1_ang_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh1_ang_df)

dbh2_ang_df <- dbh2_ang_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh2_ang_df)

dbh3_ang_df <- dbh3_ang_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh3_ang_df)

colnames(h_ang_wb_df)
colnames(h_ang_pl_df)
colnames(dbh_ang_df)
colnames(dbh1_ang_df)
colnames(dbh2_ang_df)
colnames(dbh3_ang_df)

tallo_wd_df0 <- tar_read(tallo_wd_df0)

tallo_wd_df0
summary(tallo_wd_df0)

# Remove sp < 20 obs
tallo_wd_df <- tallo_wd_df0 |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
         log_h = log(h),
         log_cr = log(cr),
         log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  select(tree_id, division, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd_ang <- tallo_wd_df |>
  filter(division == "Angiosperm")
tallo_wd_ang

common_sp_ang <- Reduce(intersect, list(
  tallo_wd_df$sp,
  h_ang_wb_df$sp,
  h_ang_pl_df$sp,
  dbh_ang_df$sp,
  dbh1_ang_df$sp,
  dbh2_ang_df$sp,
  dbh3_ang_df$sp
))

filtered_tallo_wd_df <- tallo_wd_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_h_ang_wb_df <- h_ang_wb_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_h_ang_pl_df <- h_ang_pl_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_dbh_ang_df <- dbh_ang_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_dbh1_ang_df <- dbh1_ang_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_dbh2_ang_df <- dbh2_ang_df |> dplyr::filter(sp %in% common_sp_ang)
filtered_dbh3_ang_df <- dbh3_ang_df |> dplyr::filter(sp %in% common_sp_ang)

ang_agb_df <- filtered_tallo_wd_df |>
  dplyr::left_join(
    filtered_h_ang_wb_df |>
      dplyr::rename(a_h_wb_ang = a, b_h_wb_ang = b, k_h_wb_ang = k),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_h_ang_pl_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_h_pl_ang = a, b_h_pl_ang = b, c_h_pl_ang = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh_ang_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh_ang = a, b_dbh_ang = b, c_dbh_ang = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh1_ang_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh1_ang = a, b_dbh1_ang = b, c_dbh1_ang = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh2_ang_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh2_ang = a, b_dbh2_ang = b, c_dbh2_ang = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh3_ang_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh3_ang = a, b_dbh3_ang = b, c_dbh3_ang = c),
    by = "sp"
  )

head(ang_agb_df)

ang_agb_df <- ang_agb_df |>
  mutate(
    # Extract numeric values for all coefficient columns
    a_h_wb_ang = as.numeric(gsub(" .*", "", a_h_wb_ang)),
    b_h_wb_ang = as.numeric(gsub(" .*", "", b_h_wb_ang)),
    k_h_wb_ang = as.numeric(gsub(" .*", "", k_h_wb_ang)),
    a_h_pl_ang = as.numeric(gsub(" .*", "", a_h_pl_ang)),
    b_h_pl_ang = as.numeric(gsub(" .*", "", b_h_pl_ang)),
    a_dbh_ang = as.numeric(gsub(" .*", "", a_dbh_ang)),
    b_dbh_ang = as.numeric(gsub(" .*", "", b_dbh_ang)),
    c_dbh_ang = as.numeric(gsub(" .*", "", c_dbh_ang)),
    a_dbh1_ang = as.numeric(gsub(" .*", "", a_dbh1_ang)),
    b_dbh1_ang = as.numeric(gsub(" .*", "", b_dbh1_ang)),
    a_dbh2_ang = as.numeric(gsub(" .*", "", a_dbh2_ang)),
    b_dbh2_ang = as.numeric(gsub(" .*", "", b_dbh2_ang)),
    a_dbh3_ang = as.numeric(gsub(" .*", "", a_dbh3_ang)),
    b_dbh3_ang = as.numeric(gsub(" .*", "", b_dbh3_ang))
  ) |>
  mutate(
    # Calculate biomass and related variables
    log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

    # Weibull-based height and biomass
    height_wb = a_h_wb_ang * (1 - exp(-b_h_wb_ang * dbh^k_h_wb_ang)),
    log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),

    # Power-law height and biomass
    height_pl = a_h_pl_ang * dbh^b_h_pl_ang,
    log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),

    # Power-law DBH estimations and biomass
    dbh_pl = a_dbh_ang * (cr^b_dbh_ang) * (h^c_dbh_ang),
    log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,

    dbh_pl1 = a_dbh1_ang * (cr * h)^b_dbh1_ang,
    log_AGB_pl_dbh1 = log(0.0559) + log_wd + 2 * log(dbh_pl1) + log_h,

    dbh_pl2 = a_dbh2_ang * (cr^b_dbh2_ang),
    log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,

    dbh_pl3 = a_dbh3_ang * (h^b_dbh3_ang),
    log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h
  )

agb_ang_long <- ang_agb_df |>
  tidyr::pivot_longer(cols = starts_with("log_AGB"), names_to = "Scenario", values_to = "AGB")

ggplot(agb_ang_long, aes(x = Scenario, y = AGB)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Angiosperm",
    x = "Senarios",
    y = "log predicted AGB (kg)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

combine_metrics <- function(agb_df) {
  agb_df |>
    summarize(
      # Weibull-based height metrics
      mae_wb_h = mean(abs(log_AGB_bl - log_AGB_wb_h), na.rm = TRUE),
      rmse_wb_h = sqrt(mean((log_AGB_bl - log_AGB_wb_h)^2, na.rm = TRUE)),
      bias_wb_h = mean(log_AGB_wb_h - log_AGB_bl, na.rm = TRUE),
      r2_wb_h = 1 - sum((log_AGB_wb_h - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law height metrics
      mae_pl_h = mean(abs(log_AGB_bl - log_AGB_pl_h), na.rm = TRUE),
      rmse_pl_h = sqrt(mean((log_AGB_bl - log_AGB_pl_h)^2, na.rm = TRUE)),
      bias_pl_h = mean(log_AGB_pl_h - log_AGB_bl, na.rm = TRUE),
      r2_pl_h = 1 - sum((log_AGB_pl_h - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH metrics
      mae_pl_dbh = mean(abs(log_AGB_bl - log_AGB_pl_dbh), na.rm = TRUE),
      rmse_pl_dbh = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh)^2, na.rm = TRUE)),
      bias_pl_dbh = mean(log_AGB_pl_dbh - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh = 1 - sum((log_AGB_pl_dbh - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH1 metrics
      mae_pl_dbh1 = mean(abs(log_AGB_bl - log_AGB_pl_dbh1), na.rm = TRUE),
      rmse_pl_dbh1 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh1)^2, na.rm = TRUE)),
      bias_pl_dbh1 = mean(log_AGB_pl_dbh1 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh1 = 1 - sum((log_AGB_pl_dbh1 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH2 metrics
      mae_pl_dbh2 = mean(abs(log_AGB_bl - log_AGB_pl_dbh2), na.rm = TRUE),
      rmse_pl_dbh2 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh2)^2, na.rm = TRUE)),
      bias_pl_dbh2 = mean(log_AGB_pl_dbh2 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh2 = 1 - sum((log_AGB_pl_dbh2 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE),

      # Power-law DBH3 metrics
      mae_pl_dbh3 = mean(abs(log_AGB_bl - log_AGB_pl_dbh3), na.rm = TRUE),
      rmse_pl_dbh3 = sqrt(mean((log_AGB_bl - log_AGB_pl_dbh3)^2, na.rm = TRUE)),
      bias_pl_dbh3 = mean(log_AGB_pl_dbh3 - log_AGB_bl, na.rm = TRUE),
      r2_pl_dbh3 = 1 - sum((log_AGB_pl_dbh3 - log_AGB_bl)^2, na.rm = TRUE) / sum((log_AGB_bl - mean(log_AGB_bl, na.rm = TRUE))^2, na.rm = TRUE)
    )
}

metrics_ang1 <- combine_metrics(ang_agb_df)
head(metrics_ang)

metrics_ang_long <- metrics_ang |>
  pivot_longer(cols = everything(), names_to = "Metric_Model", values_to = "Value") |>
  separate(Metric_Model, into = c("Metric", "Model"), sep = "_", extra = "merge") |>
  filter(Metric != "bias")  # Exclude the bias metric

ggplot(metrics_ang_long, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Angiosperm",
    x = "Senarios",
    y = "Metric Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set2")

plot_agb <- function(agb_df, metrics_df, model_colors, division_label) {
  create_plot <- function(data, baseline, predicted, rmse, bias, model_name, tag_label, x_axis_label = NULL) {
      ggplot(data, aes(x = exp(baseline), y = exp(predicted))) +
        geom_point(alpha = 0.5, color = "grey", size = 0.3) +
        geom_smooth(method = "lm", se = FALSE, color = model_colors[model_name], size = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black", size = 0.5) +
        scale_x_log10(
          breaks = c(0.1, 10, 1000, 100000),
          labels = function(x) ifelse(x == 0.1, "0.1", scales::label_number(accuracy = 1)(x))
        ) +
        scale_y_log10(
          breaks = c(0.1, 10, 1000, 100000),
          limits = c(NA, 100000), 
          labels = function(x) ifelse(x == 0.1, "0.1", scales::label_number(accuracy = 1)(x))
        ) +
        labs(tag = tag_label, 
            x = x_axis_label, 
            y = switch(model_name,
                        "h_wb" = expression(Predicted~AGB[H*"\u00A0"*"-"*"\u00A0"*WB]~"(kg)"),
                        "h_pl" = expression(Predicted~AGB[H*"\u00A0"*"-"*"\u00A0"*PL]~"(kg)"),
                        "dbh_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR*"\u002B"*H]~"(kg)"),
                        "dbh1_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR*"\u00D7"*H]~"(kg)"),
                        "dbh2_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*CR]~"(kg)"),
                        "dbh3_pl" = expression(Predicted~AGB[DBH*phantom()*"\u223C"*phantom()*H]~"(kg)"))) +
        # Reduce size of RMSE and Bias annotations
        annotate("text", x = 0.9, y = 1e5, label = paste("RMSE =", format(round(rmse, 2), nsmall = 2), "Mg"), hjust = 0, size = 2.5) +
        annotate("text", x = 0.9, y = 3e4, label = paste("Bias =", format(round(bias * 100, 2), nsmall = 2), "%"), hjust = 0, size = 2.5) +
        theme_minimal() +
        theme(plot.margin = margin(5, 15, 5, 5)) +
        theme(axis.title = element_text(size = 9), 
              plot.tag = element_text(face = "bold", size = 10), 
              axis.text = element_text(size = 8))  # Reduce the axis text size
  }
    model_colors <- c(
      "h_wb" = "#D55E00",
      "h_pl" = "#0072B2",
      "dbh_pl" = "#E69F00",
      "dbh1_pl" = "#F0E442",
      "dbh2_pl" = "#009E73",
      "dbh3_pl" = "#56B4E9"
    )

    p1 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_wb_h, 
                      metrics_df$rmse_wb_h, metrics_df$bias_wb_h, "h_wb", "(a)")

    p2 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_h, 
                      metrics_df$rmse_pl_h, metrics_df$bias_pl_h, "h_pl", "(b)")

    p3 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh, 
                      metrics_df$rmse_pl_dbh, metrics_df$bias_pl_dbh, "dbh_pl", "(c)")

    p4 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh1, 
                      metrics_df$rmse_pl_dbh1, metrics_df$bias_pl_dbh1, "dbh1_pl", "(d)")

    p5 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh2, 
                      metrics_df$rmse_pl_dbh2, metrics_df$bias_pl_dbh2, "dbh2_pl", "(e)",
                      x_axis_label = expression(AGB[obs]~"(kg)"))

    p6 <- create_plot(agb_df, agb_df$log_AGB_bl, agb_df$log_AGB_pl_dbh3, 
                      metrics_df$rmse_pl_dbh3, metrics_df$bias_pl_dbh3, "dbh3_pl", "(f)",
                      x_axis_label = expression(AGB[obs]~"(kg)"))

    # Combine plots
    plot <- (p1 | p2 | p3 | p4 | p5 | p6) + 
      plot_layout(ncol = 2, nrow = 3) + 
      plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)")))

    plot_with_label <- ggdraw() +
    draw_plot(plot, x = 0, y = 0, width = 1, height = 0.9) +  
    draw_label(division_label, x = 0.5, y = 0.92, fontface = 'bold', size = 10)

  return(plot_with_label)
}

plot_ang <- plot_agb(ang_agb_df, metrics_ang, model_colors, "Angiosperm")
plot_ang

ggsave("test_ang.png", plot = plot_ang, width = 120, height = 130, units = "mm")
```

## GYMNOSPERM

```{r}
## GYMNOSPERM DATA FRAME PROCESSING

# Process the Gymnosperm height dataset (h_gym_wb_df)
h_gym_wb_df <- h_gym_wb_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(h_gym_wb_df)

# Process the Gymnosperm height power-law dataset (h_gym_pl_df)
h_gym_pl_df <- h_gym_pl_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(h_gym_pl_df)

# Process the Gymnosperm DBH dataset (dbh_gym_df)
dbh_gym_df <- dbh_gym_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh_gym_df)

# Process the Gymnosperm DBH1 dataset (dbh1_gym_df)
dbh1_gym_df <- dbh1_gym_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh1_gym_df)

# Process the Gymnosperm DBH2 dataset (dbh2_gym_df)
dbh2_gym_df <- dbh2_gym_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh2_gym_df)

# Process the Gymnosperm DBH3 dataset (dbh3_gym_df)
dbh3_gym_df <- dbh3_gym_df |>
  dplyr::select(-sp.sp_id, -sp.wd_s, -sp.wd) |>
  dplyr::rename(sp = sp.sp)
head(dbh3_gym_df)

# Filter the dataset for gymnosperms
tallo_wd_gym <- tallo_wd_df |>
  filter(division == "Gymnosperm")

common_sp_gym <- Reduce(intersect, list(
  tallo_wd_df$sp,
  h_gym_wb_df$sp,
  h_gym_pl_df$sp,
  dbh_gym_df$sp,
  dbh1_gym_df$sp,
  dbh2_gym_df$sp,
  dbh3_gym_df$sp
))
common_sp_gym

filtered_tallo_wd_df_gym <- tallo_wd_df |>
  filter(sp %in% common_sp_gym)
filtered_h_gym_wb_df <- h_gym_wb_df |>
  filter(sp %in% common_sp_gym)
filtered_h_gym_pl_df <- h_gym_pl_df |>
  filter(sp %in% common_sp_gym)
filtered_dbh_gym_df <- dbh_gym_df |>
  filter(sp %in% common_sp_gym)
filtered_dbh1_gym_df <- dbh1_gym_df |>
  filter(sp %in% common_sp_gym)
filtered_dbh2_gym_df <- dbh2_gym_df |>
  filter(sp %in% common_sp_gym)
filtered_dbh3_gym_df <- dbh3_gym_df |>
  filter(sp %in% common_sp_gym)

gym_agb_df <- filtered_tallo_wd_df_gym |>
  dplyr::left_join(
    filtered_h_gym_wb_df |>
      dplyr::rename(a_h_wb_gym = a, b_h_wb_gym = b, k_h_wb_gym = k),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_h_gym_pl_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_h_pl_gym = a, b_h_pl_gym = b, c_h_pl_gym = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh_gym_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh_gym = a, b_dbh_gym = b, c_dbh_gym = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh1_gym_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh1_gym = a, b_dbh1_gym = b, c_dbh1_gym = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh2_gym_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh2_gym = a, b_dbh2_gym = b, c_dbh2_gym = c),
    by = "sp"
  ) |>
  dplyr::left_join(
    filtered_dbh3_gym_df |>
      dplyr::select(sp, a, b, c) |>
      dplyr::rename(a_dbh3_gym = a, b_dbh3_gym = b, c_dbh3_gym = c),
    by = "sp"
  )

# View the joined dataset
head(gym_agb_df)

# Convert coefficients to numeric values
gym_agb_df <- gym_agb_df |>
  mutate(
    a_h_wb_gym = as.numeric(gsub(" .*", "", a_h_wb_gym)),
    b_h_wb_gym = as.numeric(gsub(" .*", "", b_h_wb_gym)),
    k_h_wb_gym = as.numeric(gsub(" .*", "", k_h_wb_gym)),
    a_h_pl_gym = as.numeric(gsub(" .*", "", a_h_pl_gym)),
    b_h_pl_gym = as.numeric(gsub(" .*", "", b_h_pl_gym)),
    a_dbh_gym = as.numeric(gsub(" .*", "", a_dbh_gym)),
    b_dbh_gym = as.numeric(gsub(" .*", "", b_dbh_gym)),
    c_dbh_gym = as.numeric(gsub(" .*", "", c_dbh_gym)),
    a_dbh1_gym = as.numeric(gsub(" .*", "", a_dbh1_gym)),
    b_dbh1_gym = as.numeric(gsub(" .*", "", b_dbh1_gym)),
    a_dbh2_gym = as.numeric(gsub(" .*", "", a_dbh2_gym)),
    b_dbh2_gym = as.numeric(gsub(" .*", "", b_dbh2_gym)),
    a_dbh3_gym = as.numeric(gsub(" .*", "", a_dbh3_gym)),
    b_dbh3_gym = as.numeric(gsub(" .*", "", b_dbh3_gym))
  ) |>
  mutate(
    log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

    # Weibull-based height and biomass
    height_wb = a_h_wb_gym * (1 - exp(-b_h_wb_gym * dbh^k_h_wb_gym)),
    log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),

    # Power-law height and biomass
    height_pl = a_h_pl_gym * dbh^b_h_pl_gym,
    log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),

    # Power-law DBH estimations and biomass
    dbh_pl = a_dbh_gym * (cr^b_dbh_gym) * (h^c_dbh_gym),
    log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,

    dbh_pl1 = a_dbh1_gym * (cr * h)^b_dbh1_gym,
    log_AGB_pl_dbh1 = log(0.0559) + log_wd + 2 * log(dbh_pl1) + log_h,

    dbh_pl2 = a_dbh2_gym * (cr^b_dbh2_gym),
    log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,

    dbh_pl3 = a_dbh3_gym * (h^b_dbh3_gym),
    log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h
  )

# Reshape data for visualization
agb_gym_long <- gym_agb_df |>
  tidyr::pivot_longer(cols = starts_with("log_AGB"), names_to = "Scenario", values_to = "AGB")

# Visualize the data using boxplots
ggplot(agb_gym_long, aes(x = Scenario, y = AGB)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Gymnosperm",
    x = "Scenarios",
    y = "log predicted AGB (kg)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

metrics_gym <- combine_metrics(gym_agb_df)
head(metrics_gym)

plot_gym <- plot_agb(gym_agb_df, metrics_gym, model_colors, "Gymnosperm")

ggsave("test_gym.png", plot = plot_ang, width = 120, height = 130, units = "mm")

```

# CHECKING GYM WB

```{r}
tallo_wd_df <- tallo_wd_df0 |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
          log_h = log(h),
          log_cr = log(cr),
          log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  dplyr::select(tree_id, division, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd_ang <- tallo_wd_df |>
  filter(division == "Angiosperm")

tallo_wd_gym <- tallo_wd_df |>
  filter(division == "Gymnosperm")

sp_posterior_agb_df1 <- tar_read(sp_posterior_agb_df1)
head(sp_posterior_agb_df1)

sp_posterior_ang <- sp_posterior_agb_df1 |>
  filter(Division == "Angiosperm")

sp_posterior_gym <- sp_posterior_agb_df1 |>
  filter(Division == "Gymnosperm")
head(sp_posterior_gym)

filter_common_species <- function(tallo_wd_df, posterior_df) {
  common_species <- intersect(tallo_wd_df$sp, posterior_df$sp)
  tallo_wd_df_filtered <- tallo_wd_df |>
    filter(sp %in% common_species)
  posterior_df_filtered <- posterior_df |>
    filter(sp %in% common_species)
  return(list(tallo_wd_df_filtered, posterior_df_filtered))
}
# Filter common species for Angiosperms and Gymnosperms
filtered_ang <- filter_common_species(tallo_wd_ang, sp_posterior_ang)
filtered_gym <- filter_common_species(tallo_wd_gym, sp_posterior_gym)

tallo_wd_ang <- filtered_ang[[1]]
sp_posterior_ang <- filtered_ang[[2]]

tallo_wd_gym <- filtered_gym[[1]]
sp_posterior_gym <- filtered_gym[[2]]

```

#  CHECKING FUNCTION "generate_agb_estimation_com1"

```{r}
options(with = 200)
tallo_wd_df0 <- tar_read(tallo_wd_df0)
tallo_wd_df0

tallo_wd_df <- tallo_wd_df0 |>
  filter(!is.na(wd)) |>
  mutate(log_dbh = log(dbh),
          log_h = log(h),
          log_cr = log(cr),
          log_wd = log(wd)) |>
  filter(!is.na(log_dbh), !is.na(log_cr), !is.na(log_h), !is.na(log_wd)) |>
  dplyr::select(tree_id, division, sp, dbh, h, cr, wd, log_dbh, log_h, log_cr, log_wd) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

tallo_wd_df
# Split the tallo_wd_df data by Division
tallo_wd_ang <- tallo_wd_df |>
  filter(division == "Angiosperm")
tallo_wd_ang

tallo_wd_gym <- tallo_wd_df |>
  filter(division == "Gymnosperm")
tallo_wd_gym

sp_posterior_agb_df1 <- tar_read(sp_posterior_agb_df1)
head(sp_posterior_agb_df1)

# Splitting the posterior data into Angiosperms and Gymnosperms
sp_posterior_ang <- sp_posterior_agb_df1 |> filter(Division == "Angiosperm")
head(sp_posterior_ang)

sp_posterior_gym <- sp_posterior_agb_df1 |> filter(Division == "Gymnosperm")
head(sp_posterior_gym)

```

```{r}
filter_common_species <- function(tallo_wd_df, sp_posterior_agb_df1) {
  h_wb_df <- sp_posterior_agb_df1 |>
  filter(Dependent_variable == "Tree Height", !is.na(a), !is.na(b),!is.na(k)) |>
    dplyr::select(sp, a, b, k) |>
    distinct()

  h_pl_df <- sp_posterior_agb_df1 |>
    filter(Dependent_variable == "Tree Height", is.na(k)) |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh_df <- sp_posterior_agb_df1 |>
    filter(Dependent_variable == "DBH", !is.na(a), !is.na(b), !is.na(c)) |>
    dplyr::select(sp, a, b, c) |>
    distinct()

  dbh1_df <- sp_posterior_agb_df1 |>
    filter(Dependent_variable == "DBH1") |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh2_df <- sp_posterior_agb_df1 |>
    filter(Dependent_variable == "DBH2") |>
    dplyr::select(sp, a, b) |>
    distinct()

  dbh3_df <- sp_posterior_agb_df1 |>
    filter(Dependent_variable == "DBH3") |>
    dplyr::select(sp, a, b) |>
    distinct()

  common_species <- Reduce(intersect, list(
    tallo_wd_df$sp,
    h_wb_df$sp,
    h_pl_df$sp,
    dbh_df$sp,
    dbh1_df$sp,
    dbh2_df$sp,
    dbh3_df$sp
  ))

  list(
    filtered_tallo_wd_df = tallo_wd_df |> filter(sp %in% common_species),
    filtered_h_wb_df = h_wb_df |> filter(sp %in% common_species),
    filtered_h_pl_df = h_pl_df |> filter(sp %in% common_species),
    filtered_dbh_df = dbh_df |> filter(sp %in% common_species),
    filtered_dbh1_df = dbh1_df |> filter(sp %in% common_species),
    filtered_dbh2_df = dbh2_df |> filter(sp %in% common_species),
    filtered_dbh3_df = dbh3_df |> filter(sp %in% common_species)
  )
}

# Filter common species for Angiosperms and Gymnosperms
filtered_ang <- filter_common_species(tallo_wd_ang, sp_posterior_ang)
filtered_gym <- filter_common_species(tallo_wd_gym, sp_posterior_gym)

head(filtered_gym$filtered_h_pl_df)

# Function to join posterior data and estimate AGB
calculate_agb <- function(filtered_data) {
  tallo_wd_df <- filtered_data$filtered_tallo_wd_df
  agb_df <- tallo_wd_df |>
    left_join(
      filtered_data$filtered_h_wb_df |> rename(a_h_wb = a, b_h_wb = b, k_h_wb = k),
      by = "sp"
    ) |>
    left_join(
      filtered_data$filtered_h_pl_df |> rename(a_h_pl = a, b_h_pl = b),
      by = "sp"
    ) |>
    left_join(
      filtered_data$filtered_dbh_df |> rename(a_dbh = a, b_dbh = b, c_dbh = c),
      by = "sp"
    ) |>
    left_join(
      filtered_data$filtered_dbh1_df |> rename(a_dbh1 = a, b_dbh1 = b),
      by = "sp"
    ) |>
    left_join(
      filtered_data$filtered_dbh2_df |> rename(a_dbh2 = a, b_dbh2 = b),
      by = "sp"
    ) |>
    left_join(
      filtered_data$filtered_dbh3_df |> rename(a_dbh3 = a, b_dbh3 = b),
      by = "sp"
    )

  # Calculate the log_AGB based on various models
  agb_df <- agb_df |>
    mutate(
    a_h_wb = as.numeric(gsub(" \\(.*\\)", "", a_h_wb)),
    b_h_wb = as.numeric(gsub(" \\(.*\\)", "", b_h_wb)),
    k_h_wb = as.numeric(gsub(" \\(.*\\)", "", k_h_wb)),
    a_h_pl = as.numeric(gsub(" \\(.*\\)", "", a_h_pl)),
    b_h_pl = as.numeric(gsub(" \\(.*\\)", "", b_h_pl)),
    a_dbh = as.numeric(gsub(" \\(.*\\)", "", a_dbh)),
    b_dbh = as.numeric(gsub(" \\(.*\\)", "", b_dbh)),
    c_dbh = as.numeric(gsub(" \\(.*\\)", "", c_dbh)),
    a_dbh1 = as.numeric(gsub(" \\(.*\\)", "", a_dbh1)),
    b_dbh1 = as.numeric(gsub(" \\(.*\\)", "", b_dbh1)),
    a_dbh2 = as.numeric(gsub(" \\(.*\\)", "", a_dbh2)),
    b_dbh2 = as.numeric(gsub(" \\(.*\\)", "", b_dbh2)),
    a_dbh3 = as.numeric(gsub(" \\(.*\\)", "", a_dbh3)),
    b_dbh3 = as.numeric(gsub(" \\(.*\\)", "", b_dbh3))
    ) |>
    mutate(
      log_AGB_bl = log(0.0559) + log_wd + 2 * log_dbh + log_h,

      height_wb = a_h_wb * (1 - exp(-b_h_wb * dbh^k_h_wb)),    # Weibull-based height estimation
      log_AGB_wb_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_wb),  # Biomass from Weibull height

      height_pl = a_h_pl * dbh^b_h_pl,  # Power-law height estimation
      log_AGB_pl_h = log(0.0559) + log_wd + 2 * log_dbh + log(height_pl),  # Biomass from Power-law height

      dbh_pl = a_dbh * (cr^b_dbh) * (h^c_dbh),  # Power-law DBH estimation using CR and H
      log_AGB_pl_dbh = log(0.0559) + log_wd + 2 * log(dbh_pl) + log_h,  # Biomass from Power-law DBH

      dbh_pl1 = a_dbh1 * (cr * h)^b_dbh1,  # Power-law DBH estimation using CR*H
      log_AGB_pl_dbh1 = log(0.0559) + log_wd + 2 * log(dbh_pl1) + log_h,  # Biomass from Power-law DBH1

      dbh_pl2 = a_dbh2 * (cr^b_dbh2),  # Power-law DBH estimation using CR only
      log_AGB_pl_dbh2 = log(0.0559) + log_wd + 2 * log(dbh_pl2) + log_h,  # Biomass from Power-law DBH2

      dbh_pl3 = a_dbh3 * (h^b_dbh3),  # Power-law DBH estimation using H only
      log_AGB_pl_dbh3 = log(0.0559) + log_wd + 2 * log(dbh_pl3) + log_h  # Biomass from Power-law DBH3
    )

  return(agb_df)
}

# Calculate AGB for Angiosperms and Gymnosperms
agb_ang <- calculate_agb(filtered_ang)
agb_ang
agb_gym <- calculate_agb(filtered_gym)
agb_gym

# Find columns that have all NA values
na_columns <- colSums(is.na(agb_ang)) == nrow(agb_ang)
na_column_names <- names(na_columns[na_columns])

# Display the column names with all NA values
print(na_column_names)

```

# Checking AGB data

```{r}

```

### 2025-03-26 docs/allo_name.qmd
Initial commit for tree-allo

---
title: "Allometry name matching"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r}
library(tidyverse)
library(here)
```

- species list with no c_name

```{r}
sp_list <- read_csv(here("data-raw/bb_sp_list.csv")) |>
  janitor::clean_names()

sp_list
```

- species list with c_name

```{r}
abund <- read_csv(here("data-raw/abund_xtbg.csv")) |>
  janitor::clean_names()

abund
```
- We just keep `cname`, `sp`, and `life` columns.

```{r}
abund <- abund |> dplyr::select(cname, sp, life)
```

- allometry with c_name

```{r}
allo <- read_csv(here("data-raw/bb_allometry.csv")) |>
  janitor::clean_names()

allo
```

- Join allometry data with abundance data on common species names

```{r}
allo_re <- left_join(allo, abund, by = c("sp_cn" = "cname"))

allo_re
```
-  Filter the joined dataset to find entries with missing species names

```{r}
allo_re  |>
  filter(is.na(sp))
```

```{r}
allo_re  |>
  filter(is.na(sp)) |>
  pull(sp_cn)
```

```{r}
abund |>
  filter(str_detect(cname, "橄榄"))
```

橄榄 in `allo` seems to be 橄榄 sp. (`CANATO`) in `abund`.
So, we add `CANATO` to `sp` if `sp_cn` is 橄榄.

```{r}
allo_re2  <- allo_re  |>
  mutate(sp = ifelse(sp_cn == "橄榄", "CANATO", sp))

allo_re2 |>
  filter(is.na(sp))
```

Now, only 33 data points are missing species names.
We can just ignore these data when we do the analysis.

- Write the data to a csv file.
```{r}
allo_re2 |>
  write_csv(here("data/bb_allometry_re.csv"))
```

I thought we had to edit these data manually, but it seems that we don't have to.

### 2025-03-26 docs/allo_relationships.qmd
Initial commit for tree-allo

```{r}
library(readr)
library(tidyverse)
library(here)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(patchwork)
```

# data preparation
 ```{r}
allo <- read_csv("data/overlap.csv")
head(allo, 6)
allo2 <- head(allo, 300000)
```

- checking data
```{r}
hist_h <- ggplot(allo2, aes(x = h)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Tree Height", x = "Height (m)", y = "Frequency")

hist_dbh <- ggplot(allo2, aes(x = dbh)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of DBH", x = "DBH (cm)", y = "Frequency")

hist_cr <- ggplot(allo2, aes(x = cr)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of CR", x = "CR", y = "Frequency")

combined_plot <- grid.arrange(hist_h, hist_dbh, hist_cr, ncol = 3)
ggsave("figs/combined_histograms.png", plot = combined_plot, width = 12, height = 6)
```

- tranformation of data
```{r}
allo2$dbh.ln <- log(allo2$dbh)
allo2$h.ln <- log(allo2$h)
allo2$cr.ln <- log(allo2$cr)
```

- checking again
```{r}
hist_h_ln <- ggplot(allo2, aes(x = h.ln)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Log(Height)", x = "Log(Height)", y = "Frequency")

hist_dbh_ln <- ggplot(allo2, aes(x = dbh.ln)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Log(DBH)", x = "Log(DBH)", y = "Frequency")

hist_cr_ln <- ggplot(allo2, aes(x = cr.ln)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of Log(CR)", x = "Log(CR)", y = "Frequency")

combined_plot_ln <- grid.arrange(hist_h_ln, hist_dbh_ln, hist_cr_ln, ncol = 3)
ggsave("figs/combined_histograms_ln.png", plot = combined_plot_ln, width = 12, height = 6)
```

- negative values in cr => transfer
```{r}
allo2$cr.ln <- log(allo2$cr + 1)
allo3 <- allo2
summary(allo3)
```

```{r}
hist_h_ln <- ggplot(allo2, aes(x = h.ln)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Log(Height)", x = "Log(Height)", y = "Frequency")

hist_dbh_ln <- ggplot(allo2, aes(x = dbh.ln)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Log(DBH)", x = "Log(DBH)", y = "Frequency")

hist_cr_ln <- ggplot(allo2, aes(x = cr.ln)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of Log(CR)", x = "Log(CR)", y = "Frequency")

combined_plot_ln <- grid.arrange(hist_h_ln, hist_dbh_ln, hist_cr_ln, ncol = 3)
ggsave("figs/combined_histograms_ln_plus_1.png", plot = combined_plot_ln, width = 12, height = 6)
```

# fitting power law

```{r}
dhmod1 <- lm((dbh.ln ~ h.ln), data = allo3) 
dcmod1 <- lm((dbh.ln ~ cr.ln), data = allo3)
hdmod1 <- lm((h.ln ~ dbh.ln), data = allo3)
cdmod1 <- lm((cr.ln ~ dbh.ln), data = allo3)
dhcmod1 <- lm((dbh.ln ~ h.ln + cr.ln), data = allo3)
```

 - plotting models

```{r}
plot_dhmod1 <- ggplot(allo2, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

plot_dcmod1 <- ggplot(allo2, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln", x = "cr.ln", y = "dbh.ln")

plot_hdmod1 <- ggplot(allo2, aes(x = dbh.ln, y = h.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: h.ln vs. dbh.ln", x = "dbh.ln", y = "h.ln")

plot_cdmod1 <- ggplot(allo2, aes(x = dbh.ln, y = cr.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: cr.ln vs. dbh.ln", x = "dbh.ln", y = "cr.ln")

plot_dhcmod1 <- ggplot(allo2, aes(x = h.ln + cr.ln, y = dbh.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln + cr.ln", x = "h.ln + cr.ln", y = "dbh.ln")

# Arrange plots in one figure
combined_plots <- grid.arrange(plot_dhmod1, plot_dcmod1, plot_hdmod1, plot_cdmod1, plot_dhcmod1,
                               nrow = 2, ncol = 3)

# Save the combined plot
ggsave("figs/combined_plots.png", plot = combined_plots, width = 12, height = 8)             
```

```{r}
summary(dhmod1)
summary(dcmod1)
summary(hdmod1)
summary(cdmod1)
summary(dhcmod1)
```

```{r}
line_plot_allo2 <- ggplot(allo2) +
  geom_line(aes(x = h.ln, y = dbh.ln, group = sp)) +
  labs(title = "Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")
# Save the plot
ggsave("figs/line_plot_allo2.png", plot = line_plot_allo2, width = 12, height = 8)
```

```{r}
# Define the plot with points and linear regression lines
line_plot_allo2 <- ggplot(allo2, aes(x = h.ln, y = dbh.ln, group = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") +  # Add linear regression line
  labs(title = "Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/_scatter_plot_allo2_with_lm.png", plot = line_plot_allo2, width = 12, height = 8)
```

```{r}
plot <- ggplot(data = allo2, aes(y = h.ln, x = dbh.ln, group = sp, group = sp, shape = sp)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1) +
  theme_minimal() +
  labs(title = "Relationship between DBH and Height by Species",
       subtitle = "Regression lines for each species, colored and shaped by species",
       group = "sp", shape = "sp")
ggsave("figs/_scatter_plot_allo2_with_lm.png", plot = plot, width = 12, height = 8)
```

work
```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "green") +  # Add linear regression line
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height)",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height.png", plot = scatter_plot, width = 12, height = 8)
```

- does not work 
#```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE) +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)

```

- does not work
```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)
```

```{r}
# Scatter plot with separate linear regression lines for each species
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp), color = "black") +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)
```

#```{r}
plot_allo3 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, group = sp, color = sp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression lines
  labs(title = "Relationship between Log(DBH) and Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/relationship_dbh_height_by_species.png", plot = plot_allo3, width = 12, height = 8)
#```

```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_lines.png", plot = plot_dhmod1, width = 12, height = 8)
```

#```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp, colour = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_color.png", plot = plot_dhmod1, width = 12, height = 8)
```

```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(colour = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_colour.png", plot = plot_dhmod1, width = 12, height = 8)
```

```{r}
scatter_plot <- ggplot( allo3, aes(x = h.ln, y = dbh.ln))+
 geom_point()+
 geom_smooth(method = "lm", se = FALSE, color ="blue") +
 labs( title = "Plot of log(dbh) vs. log(h)"; x = "log(h)", y = "log(dbh)")
ggsave("figs/scatter_plot_dbh_vs_h.png", plot = scatter_plot, width = 12, height = 8)
```

```{r}
plot_dcmod1 <- ggplot(allo3, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln by Species", x = "cr.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/dcmod1_scatter_plot_species_lm_lines.png", plot = plot_dcmod1, width = 12, height = 8)
```

```{r}
# Scatter plot for dhmod1
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

# Scatter plot for dcmod1
plot_dcmod1 <- ggplot(allo3, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln", x = "cr.ln", y = "dbh.ln")

# Scatter plot for hdmod1
plot_hdmod1 <- ggplot(allo3, aes(x = dbh.ln, y = h.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: h.ln vs. dbh.ln", x = "dbh.ln", y = "h.ln")

# Scatter plot for cdmod1
plot_cdmod1 <- ggplot(allo3, aes(x = dbh.ln, y = cr.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: cr.ln vs. dbh.ln", x = "dbh.ln", y = "cr.ln")

# Scatter plot for dhcmod1
plot_dhcmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(color = cr.ln)) +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

# Combine all plots into one figure
combined_plots <- grid.arrange(plot_dhmod1, plot_dcmod1, plot_hdmod1, plot_cdmod1, plot_dhcmod1,
                               nrow = 2, ncol = 3, 
                               top = textGrob("Scatter Plots with Fitted Regression Lines", 
                                              gp = gpar(fontsize = 16, fontface = "bold")))

# Save figure 
ggsave("figs/power_combined_plots.png", combined_plots, width = 15, height = 10)
```

# fitting gMM (drm())
```{r}
library(drc)
library(minpack.lm)

```

- Remove NaNs

```{r}
allo4 <- allo3 %>% 
  mutate(hcr = h.ln + cr.ln)
head(allo4)
```

```{r}
sum(is.na(allo4))
allo4 <- na.omit(allo4)
summary(allo4)
```

```{r}
michaelis_menten <- function(x, a, b, k) {
  a * x^b / (k + x^b)
}
dhmod2 <- nlsLM(dbh.ln ~ michaelis_menten(h.ln, a, b, k), data = allo4,
                 start = list(a = 1, b = 5, k = 0.1), 
                 control = nls.lm.control(maxiter = 100, maxfev = 500))

dcmod2 <- nlsLM(dbh.ln ~ michaelis_menten(cr.ln, a, b, k), data = allo4,
                start = list(a = 1, b = 1, k = 1))

hdmod2 <- nlsLM(h.ln ~ michaelis_menten(dbh.ln, a, b, k), data = allo4,
                start = list(a = 1, b = 1, k = 1))

cdmod2 <- nlsLM(cr.ln ~ michaelis_menten(dbh.ln, a, b, k), 
                data = allo4,
                start = list(a = 1, b = 1, k = 1),
                control = nls.lm.control(maxiter = 500, maxfev = 1000))

dhcmod2 <- nlsLM(dbh.ln ~ michaelis_menten(hcr, a, b, k), data = allo4,
                 start = list(a = 1, b = 1, k = 1),
                 control = nls.lm.control(maxiter = 500, maxfev = 500))

summary(dhmod2)
summary(dcmod2)
summary(hdmod2)
summary(cdmod2)
summary(dhcmod2)
```

### 2025-03-26 docs/bays_check1.qmd
Initial commit for tree-allo

---
title: "Diagnostics for MCMC"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
     collapsed: TRUE
     smooth_scroll: TRUE
     toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(dplyr)
```

# Models

```{r}
loo_tbl <- tar_read(loo_tbl)
loo_tbl
print(loo_tbl , n = Inf)

#gMM1, PL1, and Weibull1 models
loo_tbl <- loo_tbl |>
  filter(grepl("gmm1|power_law1|weibull1", model))
print(loo_tbl)
```

8 models to predict DBH:

- DBH: DBH = a.(CR^b).(H^c)

- DBH1: DBH = a.(CR.H)^b

- DBH2: DBH = a.CR^b

- DBH3: DBH = a.H^b

6 models to predic CR:

- CR ~ DBH (pl1, gMM1, Weibull1)

6 models to predic H:

- H ~ DBH (pl1, gMM1, Weibull1)

Species-level Predictor: With and Without Wood density

# Convergence

I consider the ESS < 300 | Rhat >1.1 to indicate the convergence issue.

```{r}
options(width = 150)
pl1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
pl1_h <- tar_read(fit_lr2_summary_power_law1_h)
pl1_dbh <- tar_read(fit_lr2_summary_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_summary_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_summary_power_law1_dbh3)
pl1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
pl1_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)
pl1_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_summary_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_summary_power_law1_nou_dbh3)

gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_summary_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_summary_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_summary_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
```
```{r}
pl1_cr
pl1_h
pl1_dbh
pl1_dbh1
pl1_dbh2
pl1_dbh3
pl1_nou_cr
pl1_nou_h
pl1_nou_dbh
pl1_nou_dbh1
pl1_nou_dbh2
pl1_nou_dbh3
gmm1_cr
gmm1_h
gmm1_nou_cr
gmm1_nou_h
weibull1_cr
weibull1_h
weibull1_nou_cr
weibull1_nou_h
```

```{r}
check_div <- function(model_data) {
  model_data |>
    filter((grepl("^beta", variable) | grepl("^gamma", variable)) & (rhat > 1.1 | ess_bulk < 300))
}

div_pl1_cr <- check_div(pl1_cr)
div_pl1_h <- check_div(pl1_h)
div_pl1_dbh <- check_div(pl1_dbh)
div_pl1_dbh1 <- check_div(pl1_dbh1)
div_pl1_dbh2 <- check_div(pl1_dbh2)
div_pl1_dbh3 <- check_div(pl1_dbh3)
div_pl1_nou_cr <- check_div(pl1_nou_cr)
div_pl1_nou_h <- check_div(pl1_nou_h)
div_pl1_nou_dbh <- check_div(pl1_nou_dbh)
div_pl1_nou_dbh1 <- check_div(pl1_nou_dbh1)
div_pl1_nou_dbh2 <- check_div(pl1_nou_dbh2)
div_pl1_nou_dbh3 <- check_div(pl1_nou_dbh3)
div_gmm1_cr <- check_div(gmm1_cr)
div_gmm1_h <- check_div(gmm1_h)
div_gmm1_nou_cr <- check_div(gmm1_nou_cr)
div_gmm1_nou_h <- check_div(gmm1_nou_h)
div_weibull1_cr <- check_div(weibull1_cr)
div_weibull1_h <- check_div(weibull1_h)
div_weibull1_nou_cr <- check_div(weibull1_nou_cr)
div_weibull1_nou_h <- check_div(weibull1_nou_h)
```

```{r}
div_pl1_cr
div_pl1_h
div_pl1_dbh
div_pl1_dbh1
div_pl1_dbh2
div_pl1_dbh3
div_pl1_nou_cr
div_pl1_nou_h
div_pl1_nou_dbh
div_pl1_nou_dbh1
div_pl1_nou_dbh2
div_pl1_nou_dbh3
div_gmm1_cr
div_gmm1_h
div_gmm1_nou_cr
div_gmm1_nou_h
div_weibull1_cr
div_weibull1_h
div_weibull1_nou_cr
div_weibull1_nou_h
```

## Checking the problematic species
```{r}
sp_ids <- function(div_data) {
  div_data |>
    filter(grepl("^beta_hat\\[", variable)) |>
    mutate(sp_id = as.numeric(sub("beta_hat\\[([0-9]+),.*\\]", "\\1", variable))) |>
    pull(sp_id) |>
    unique()
}

sp_ids_gmm1_cr <- sp_ids(div_gmm1_cr)
sp_ids_gmm1_h <- sp_ids(div_gmm1_h)
sp_ids_gmm1_nou_cr <- sp_ids(div_gmm1_nou_cr)
sp_ids_gmm1_nou_h <- sp_ids(div_gmm1_nou_h)
sp_ids_weibull1_cr <- sp_ids(div_weibull1_cr)
sp_ids_weibull1_h <- sp_ids(div_weibull1_h)
sp_ids_weibull1_nou_cr <- sp_ids(div_weibull1_nou_cr)
sp_ids_weibull1_nou_h <- sp_ids(div_weibull1_nou_h)

## Print the unique sp_ids for each model
print(sp_ids_gmm1_cr)
print(sp_ids_gmm1_h)
print(sp_ids_gmm1_nou_cr)
print(sp_ids_gmm1_nou_h)
print(sp_ids_weibull1_cr)
print(sp_ids_weibull1_h)
print(sp_ids_weibull1_nou_cr)
print(sp_ids_weibull1_nou_h)

```

Rhat are good, mostly close to 1 (but sp_id 593 is another case).
Model 1-12: mostly have convergent problems in gamma parameter with low ESS.
Non-linear models, in general, have many problematic species with low ESS.
We should increase MCMC iterations, adapt_delta.

# Comparing models
```{r}
loo_tbl

# CR models
cr_models <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  arrange(looic)

print(cr_models)

# H models
h_models <- loo_tbl |>
  filter(grepl("_h", model)) |>
  arrange(looic)

print(h_models)

# DBH models
dbh_models <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  arrange(looic)

# Print the filtered and sorted models
print(dbh_models)
```

- CR models: Weibull incorporating WD is the best model 
- H models: Weibull incorporating WD is the best model 
- DBH: Incorporating CR and H improved DBH estimation accuracy.
- WD slightly effects tree architecture differentiation.

<!-- 
# Table
```{r}
calculate_delta_looic <- function(df) {
  df <- df |>
    arrange(looic) |>
    mutate(ΔLOOIC = looic - min(looic))
  return(df)
}

# Calculate ΔLOOIC for CR models
cr_models_tbl <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the CR models table with ΔLOOIC
print(cr_models_tbl)

# Calculate ΔLOOIC for H models
h_models_tbl <- loo_tbl |>
  filter(grepl("_h", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the H models table with ΔLOOIC
print(h_models_tbl)

# Calculate ΔLOOIC for DBH models
dbh_models_tbl <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the DBH models table with ΔLOOIC
print(dbh_models_tbl)
```

```{r}
# Define the custom format function for tables
format_table <- function(df, title) {
  formatted_df <- df |>
    rename(
      "Model" = model,
      "Wood Density" = Wood_Density,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    select(Model, `Wood Density`, ELPD, pLOO, LOOIC, `ΔLOOIC`) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(formatted_df)

  # Return the formatted dataframe
  return(formatted_df)
}

# Applying the function to CR, H, and DBH models
cr_models_tbl_formatted <- format_table(cr_models_tbl, "Crown Radius Models")
h_models_tbl_formatted <- format_table(h_models_tbl, "Height Models")
dbh_models_tbl_formatted <- format_table(dbh_models_tbl, "DBH Models")
```
 -->

### 2025-03-26 docs/bays_plot.qmd
Initial commit for tree-allo

---
title: "Bayesian results"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the required packages

```{r}
# Load necessary libraries
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(bayesplot)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(rstantools)
library(kableExtra)
library(ggpubr)
```

# Compare Model Performance using PSIS-LOO
```{r}
# Load the models
pl_dbh <- tar_read(fit_lr_mcmc_power_law_dbh)
pl_dbh1 <- tar_read(fit_lr_mcmc_power_law_dbh1)
```

```{r}
# Extract log-likelihood for both models
log_lik_dbh <- pl_dbh$draws(variables = "log_lik", format = "draws_matrix")
log_lik_dbh_matrix <- as_draws_matrix(log_lik_dbh)

log_lik_dbh1 <- pl_dbh1$draws(variables = "log_lik", format = "draws_matrix")
log_lik_dbh1_matrix <- as_draws_matrix(log_lik_dbh1)

# Calculate LOO for both models
loo_dbh <- loo(log_lik_dbh_matrix, reloo = TRUE, save_psis = TRUE)
loo_dbh1 <- loo(log_lik_dbh1_matrix, reloo = TRUE, save_psis = TRUE)

# Print LOO results
print(loo_dbh)
print(loo_dbh1)

# Compare models
model_comparison <- loo_compare(loo_dbh1, loo_dbh)
print(model_comparison)

```

```{r}
# Extract relevant metrics
metrics <- data.frame(
  Dependent_variable = c("DBH"),
  Functional_form = c("Power_law", "gMM", "Weibull")
  Model = c("pl_dbh", "pl_dbh1", "pl_dbh_", "pl_dbh1_"),
  group_level_predictor = c("Wood density", "-", "Wood density", "-"),
  elpd_loo = c(loo_dbh$estimates[1, 1], loo_dbh1$estimates[1, 1]),
  se_elpd_loo = c(loo_dbh$estimates[1, 2], loo_dbh1$estimates[1, 2]),
  p_loo = c(loo_dbh$estimates[2, 1], loo_dbh1$estimates[2, 1]),
  se_p_loo = c(loo_dbh$estimates[2, 2], loo_dbh1$estimates[2, 2]),
  looic = c(loo_dbh$estimates[3, 1], loo_dbh1$estimates[3, 1]),
  se_looic = c(loo_dbh$estimates[3, 2], loo_dbh1$estimates[3, 2])
)

# Create the table as a plot
table_plot <- ggtexttable(metrics, rows = NULL, theme = ttheme("mBlue"))

# Save the table as an image
ggsave("figs/model_comparison_table.png", table_plot, width = 10, height = 6)

# Display the table in the document
print(table_plot)
```

## Visualization
```{r}
# Extract the posterior draws for pl_dbh
posterior_draws_dbh <- as_draws_matrix(pl_dbh$draws())
posterior_draws_dbh1 <- as_draws_matrix(pl_dbh1$draws())
```

## pl_dbh
```{r}
# Trace plots for parameters
trace_plot_dbh <- mcmc_trace(posterior_draws_dbh[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]) +
  ggtitle("Trace Plots for Parameters (Model pl_dbh)")

# Density plots for parameters
density_plot_dbh <- mcmc_areas(posterior_draws_dbh[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]) +
  ggtitle("Density Plots for Parameters (pl_dbh)")

# Pareto k diagnostics
pareto_k_plot_dbh <- plot(loo_dbh) +
  title("Pareto k Diagnostics (pl_dbh)")

# Combine and display plots for pl_dbh
grid.arrange(trace_plot_dbh, density_plot_dbh, ncol = 2)
grid.draw(pareto_k_plot_dbh)

```

## pl_dbh1
```{r}
# Trace plots for parameters
trace_plot_dbh1 <- mcmc_trace(posterior_draws[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]) +
  ggtitle("Trace Plots for Parameters (pl_dbh1)")

# Density plots for parameters
density_plot_dbh1 <- mcmc_areas(posterior_draws[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]) +
  ggtitle("Density Plots for Parameters (pl_dbh1)")

# Pareto k diagnostics
pareto_k_plot_dbh1 <- plot(loo_dbh1) +
  title("Pareto k Diagnostics")

# Combine and display plots
gridExtra::grid.arrange(trace_plot, density_plot, pareto_k_plot, ncol = 2)

```

# Marginal posterior predictive checks
```{r}
# stan_data_lr_dbh
# stan_data_lr_dbh1
stan_lr_data_dbh <- tar_read(stan_data_lr_dbh)
stan_lr_data_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_lr_data_dbh)
str(stan_lr_data_dbh1)

str(stan_lr_data_dbh)
str(stan_lr_data_dbh1)
```

```{r}
predictive_function <- function(draws, data) {
  # Extract parameters from draws
  gamma <- as.matrix(draws[, grepl("gamma", colnames(draws))])

  # Compute posterior predictions
  predictions <- apply(gamma, 1, function(params) {
    # Example formula: y_pred = exp(params[1]) * data$log_x[, 1] ^ params[2]
    y_pred <- exp(params[1]) * data$log_x[, 1] ^ params[2]
    return(y_pred)
  })

  return(predictions)
}
```

## pl_dbh
```{r}
# Compute posterior predictions
y_rep_dbh <- predictive_function(posterior_draws_dbh, stan_lr_data_dbh)

# Assuming `log_y` is observed data
y_obs_dbh <- stan_lr_data_dbh$log_y

str(y_rep_dbh)
str(y_obs_dbh)

# Ensure dimensions match
y_rep_dbh <- t(y_rep_dbh) 
str(y_rep_dbh)

# Marginal posterior predictive checks
ppc_dbh <- ppc_loo_pit_overlay(
  y = y_obs_dbh,
  yrep = y_rep_dbh, 
  lw = weights(loo_dbh$psis_object)
) + ggtitle("Marginal Posterior Predictive Check (pl_dbh)")

print(ppc_dbh)
```

## pl_dbh1
```{r}
# Compute posterior predictions
y_rep_dbh1 <- predictive_function(posterior_draws_dbh1, stan_lr_data_dbh1)

# Assuming `log_y` is observed data
y_obs_dbh1 <- stan_lr_data_dbh1$log_y

# Ensure dimensions match
y_rep_dbh1 <- t(y_rep_dbh1)  # Transpose to match dimensions

# Marginal posterior predictive checks
ppc_dbh1 <- ppc_loo_pit_overlay(
  y = y_obs_dbh1,
  yrep = y_rep_dbh1, 
  lw = weights(loo_dbh1$psis_object)
) + ggtitle("Marginal Posterior Predictive Check (pl_dbh1)")

print(ppc_dbh1)
```

# Comparison of H model
```{r}
pl_h <-  tar_read(fit_lr_mcmc_power_law_h)
pl_h
gmm_h <-  tar_read(fit_nlr1_mcmc_gmm_h)
# gmm_h <- summary$gmm_h(variable = "gamma")
gmm_h
```

```{r}
# Extract log-likelihood for both models
log_lik_pl_h <- pl_h$draws(variables = "log_lik")
log_lik_pl_h_matrix <- as_draws_matrix(log_lik_pl_h)

log_lik_gmm_h <- gmm_h$draws(variables = "log_lik", format = "draws_matrix")
log_lik_gmm_h_matrix <- as_draws_matrix(log_lik_gmm_h)

# Calculate LOO for both models
loo_pl_h <- loo(log_lik_pl_h_matrix, reloo = TRUE, save_psis = TRUE)
loo_gmm_h <- loo(log_lik_gmm_h_matrix, reloo = TRUE, save_psis = TRUE)

# Print LOO results
print(loo_pl_h)
print(loo_gmm_h)

# Compare models
h_model_comparison <- loo_compare(loo_pl_h, loo_gmm_h)
print(h_model_comparison)
```

# Comparison of CR model 
```{r}
pl_cr <- tar_read(fit_lr_mcmc_power_law_cr)
pl_cr
gmm_cr <-  tar_read(fit_nlr1_mcmc_gmm_cr)
# gmm_cr <- summary$gmm_cr(variable = "gamma")
```

```{r}
# Extract log-likelihood for both models
log_lik_pl_cr <- pl_cr$draws(variables = "log_lik")
log_lik_pl_cr_matrix <- as_draws_matrix(log_lik_pl_cr)

log_lik_gmm_cr <- gmm_cr$draws(variables = "log_lik", format = "draws_matrix")
log_lik_gmm_cr_matrix <- as_draws_matrix(log_lik_gmm_cr)

# Calculate LOO for both models
loo_pl_cr <- loo(log_lik_pl_cr_matrix, reloo = TRUE, save_psis = TRUE)
loo_gmm_cr <- loo(log_lik_gmm_cr_matrix, reloo = TRUE, save_psis = TRUE)

# Print LOO results
print(loo_pl_cr)
print(loo_gmm_cr)

# Compare models
cr_model_comparison <- loo_compare(loo_pl_cr, loo_gmm_cr)
print(cr_model_comparison)
```

# Try to make plots
```{r}
pl_cr <- tar_read(fit_lr_mcmc_power_law_cr)
pl_cr
stan_lr_data_cr <- tar_read(stan_data_lr_cr)
str(stan_lr_data_cr)
```

```{r}
# Extract summaries
summary_pl_cr <- pl_cr$summary()

# Extract mean values for log(a) and b
log_a_mean <- summary_pl_cr[summary_pl_cr$variable == "gamma[1,1]", "mean"]
b_mean <- summary_pl_cr[summary_pl_cr$variable == "gamma[2,1]", "mean"]

# Convert log(a) to a
a_mean <- exp(log_a_mean)

# Define a function to make predictions using the power law model
make_predictions_pl <- function(a, b, x_seq) {
  a * x_seq^b
}

# Generate sequence for trunk diameter
dbh_seq <- seq(1, 100, length.out = 100)

# Generate predictions
cr_pred <- make_predictions_pl(a_mean, b_mean, dbh_seq)

df_pred_pl_cr <- data.frame(
  dbh = dbh_seq,
  crown_radius = cr_pred
)

```

```{r}
str(stan_lr_data_h)
head(stan_lr_data_h)

hist(stan_lr_data_h$log_y)

dbh <- stan_lr_data_h$log_x
cr <- stan_lr_data_h$log_y
sp <- factor(stan_lr_data_h$jj)

# Create the data frame
df <- data.frame(
  dbh = dbh,
  cr = cr,
  sp = sp
)

head(df)
str(df)

pl_cr_plot <- ggplot(df, aes(x= dbh.2, y = cr, color = sp))
pl_cr_plot
```

```{r}
library(cmdstanr)
library(posterior)
library(tidyverse)

pl_cr <- tar_read(fit_lr_mcmc_power_law_cr)
pl_cr

stan_lr_data_cr <- tar_read(stan_data_lr_cr)
str(stan_lr_data_cr)

# Assuming `stan_lr_data_cr` is your data list
log_y <- stan_lr_data_cr$log_y
log_x <- stan_lr_data_cr$log_x[, 2]  
data$sp <- as.factor(stan_lr_data_cr$jj)

# Creating a data frame for ggplot
data <- data.frame(CR = log_y, DBH = log_x, sp = sp)
head(data)
summary(data)

# Scatter plot of CR vs DBH
ggplot(data, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5) +
  labs(title = "Scatter Plot of CR vs DBH",
       x = "DBH (cm)",
       y = "CR (m)") +
  theme_minimal()

# Extract summaries
summary_pl_cr <- pl_cr$summary()

# Extract mean values for log(a) and b
log_a_mean <- summary_pl_cr |> filter(variable == "gamma[1,1]") |> pull(mean)
b_mean <- summary_pl_cr |> filter(variable == "gamma[1,2]") |> pull(mean)

# Compute the fitted values for CR using the formula log_y = log_a + b * log_x
fitted_log_y <- log_a_mean + b_mean * log_x
str(fitted_log_y)

# Create a data frame for ggplot
data <- data.frame(CR = log_y, DBH = log_x, fitted_CR = fitted_log_y, sp = sp)
head(data)
str(data)

# Plot the scatter plot of observed CR vs DBH and add the fitted line
ggplot(data, aes(x = exp(DBH), y = exp(CR))) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = fitted_CR), color = "red") +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  theme_minimal()

ggplot(data, aes(x = exp(DBH), y = exp(CR))) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = exp(DBH), y = exp(fitted_CR)), color = "red") +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Plot the scatter plot of observed CR vs DBH and add the fitted line for each species
p <- ggplot(data, aes(x = exp(DBH), y = exp(CR))) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = exp(DBH), y = exp(fitted_CR), color = sp, group = sp), show.legend = FALSE) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Set background color to white
par(bg = "white")
bg = "white"
# Save the plot
ggsave("figs/scatter_plot_CR_vs_DBH.png", plot = p, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
```

```{r}
library(cmdstanr)
library(posterior)
library(tidyverse)

gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
gmm_h

stan_nlr_data_h <- tar_read(stan_data_nlr_h)
str(stan_nlr_data_h)

# Assuming `stan_lr_data_cr` is your data list
log_y <- stan_nlr_data_h$log_y
x <- stan_nlr_data_h$x
data$sp <- as.factor(stan_nlr_data_h$jj)

# Extract summaries
summary_gmm_h <- gmm_h$summary()

# Extract mean values for log(a) and b
log_a2 <- summary_gmm_h |> filter(variable == "gamma[1,1]") |> pull(mean)
b2 <- summary_gmm_h |> filter(variable == "gamma[1,2]") |> pull(mean)
k1 <- summary_gmm_h |> filter(variable == "gamma[1,3]") |> pull(mean)

# Compute the fitted values for H using the formula gMM
fitted_log_y_h <- log_a2 + b2 * log(x) - log(k1 + x^b2)

# Create a data frame for ggplot
data_h <- data.frame(H = exp(log_y), DBH = x, fitted_H = exp(fitted_log_y_h), sp = sp)

# Check the structure of the data frame
head(data_h)
str(data_h)

# Plot the scatter plot of observed H vs DBH and add the fitted line for each species

h <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = DBH, y = fitted_H, color = sp, group = sp), show.legend = FALSE) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

h
# Save the plot
ggsave("figs/scatter_plot_H_vs_DBH.png", plot = h, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
```

```{r}
library(targets)
library(bayesplot)
gmm_cr <- tar_read(fit_nlr4_mcmc_gmm_cr)
gmm_cr
gamma_trace <- gmm_cr$draws("gamma")
gamma_trace
mcmc_trace(gamma_trace)

gamma_trace <- gmm_nou_h$draws("gamma")
mcmc_trace(gamma_trace)

gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)
gamma_trace <- gmm_nou_h$draws("gamma")
mcmc_trace(gamma_trace)
```

```{r}
library(dplyr)
library(readr)
library(targets)
library(stringr)

tallo_wd_200_df <- read_csv("data/tallo_wd_200.csv")
tallo_wd_200_df
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)

beta <- gmm_nou_h$summary(variable = "beta")
str(beta)
beta

# Filter beta parameters with small ess_bulk
small_ess_beta <- beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

# Print the filtered beta parameters
print(small_ess_beta)

# Count the number of individuals per species
species_count <- tallo_wd_200_df |>
  group_by(sp) |>
  summarise(num_individuals = n())

# Print the species count
print(species_count)

# Create a mapping from species_id to species name and number of individuals
species_map <- tallo_wd_200_df |>
  mutate(species_id = as.integer(as.factor(sp))) |>
  group_by(species_id, sp) |>
  summarise(num_individuals = n(), .groups = 'drop')

# Join the filtered beta parameters with the species count
small_ess_species <- small_ess_beta |>
  left_join(species_map, by = "species_id")

# Print the result
options(width = 200)
print(small_ess_species)
summary(small_ess_species)

# Extract the unique species and their number of individuals
unique_species_count <- small_ess_species |>
  select(sp, num_individuals) |>
  distinct()

# Print the unique species and their number of individuals
print(unique_species_count, n = Inf)

```

# Covariance between beta
```{r}
library(cmdstanr)
library(posterior)
library(dplyr)
library(tidyr)

gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)

# Extract beta samples
beta_samples <- as_draws_df(gmm_nou_h, variable = "beta")

# Check the structure of the extracted samples
str(beta_samples)

# Reshape the data frame to long format
beta_long <- beta_samples |>
  pivot_longer(cols = starts_with("beta"), names_to = "parameter", values_to = "value")

# Extract species and parameter indices from the names
beta_long <- beta_long |>
  mutate(species = as.integer(str_extract(parameter, "(?<=\\[)\\d+")),
         param = as.integer(str_extract(parameter, "(?<=,)\\d+(?=\\])"))) |>
  select(-parameter)

# Filter for the three beta parameters of interest (e.g., beta[,1], beta[,2], beta[,3])
beta_filtered <- beta_long %>%
  filter(param %in% c(1, 2, 3))

# Reshape to wide format for covariance calculation
beta_wide <- beta_filtered %>%
  pivot_wider(names_from = param, values_from = value, names_prefix = "beta_")

# Print the reshaped data frame
print(beta_wide)

# Compute the covariance matrix for beta parameters 1, 2, and 3
cov_matrix <- cov(beta_wide[, c("beta_1", "beta_2", "beta_3")])

# Print the covariance matrix
print(cov_matrix)

```

### 2025-03-26 docs/biome.qmd
Initial commit for tree-allo

---
title: "Biomes' allometry plotting"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r include=FALSE}
library(tidyverse)
library(targets)
library(viridis)
library(ggplot2)
options(width=150)
```

# Read in data

```{r}
tallo_wd_df0 <- tar_read(tallo_wd_df0)
unique(tallo_wd_df0$biome)
```

```{r}
# Boreal/montane forest
boreal_ang <- tallo_wd_df0 |> filter(biome == "Boreal/montane forest" & division == "Angiosperm")
boreal_ang

boreal_gym <- tallo_wd_df0 |> filter(biome == "Boreal/montane forest" & division == "Gymnosperm")
boreal_gym

# Tropical rain forest
trop_rain_ang <- tallo_wd_df0 |> filter(biome == "Tropical rain forest" & division == "Angiosperm")
trop_rain_ang

trop_rain_gym <- tallo_wd_df0 |> filter(biome == "Tropical rain forest" & division == "Gymnosperm")
trop_rain_gym

# Temperate broadleaf forest
tem_broad_ang <- tallo_wd_df0 |> filter(biome == "Temperate broadleaf forest" & division == "Angiosperm")
tem_broad_ang

tem_broad_gym <- tallo_wd_df0 |> filter(biome == "Temperate broadleaf forest" & division == "Gymnosperm")
tem_broad_gym

# Tropical savanna
trop_sav_ang <- tallo_wd_df0 |> filter(biome == "Tropical savanna" & division == "Angiosperm")
trop_sav_ang

trop_sav_gym <- tallo_wd_df0 |> filter(biome == "Tropical savanna" & division == "Gymnosperm")
trop_sav_gym

# Temperate grassland
tem_grass_ang <- tallo_wd_df0 |> filter(biome == "Temperate grassland" & division == "Angiosperm")
tem_grass_ang

tem_grass_gym <- tallo_wd_df0 |> filter(biome == "Temperate grassland" & division == "Gymnosperm")
tem_grass_gym

# Temperate conifer forest
tem_coni_ang <- tallo_wd_df0 |> filter(biome == "Temperate conifer forest" & division == "Angiosperm")
tem_coni_ang

tem_coni_gym <- tallo_wd_df0 |> filter(biome == "Temperate conifer forest" & division == "Gymnosperm")
tem_coni_gym

# Mediterranean woodland
medi_ang <- tallo_wd_df0 |> filter(biome == "Mediterranean woodland" & division == "Angiosperm")
medi_ang

medi_gym <- tallo_wd_df0 |> filter(biome == "Mediterranean woodland" & division == "Gymnosperm")
medi_gym

# Tropical dry forest
trop_dry_ang <- tallo_wd_df0 |> filter(biome == "Tropical dry forest" & division == "Angiosperm")
trop_dry_ang

trop_dry_gym <- tallo_wd_df0 |> filter(biome == "Tropical dry forest" & division == "Gymnosperm")
trop_dry_gym

# Dryland
dryland_ang <- tallo_wd_df0 |> filter(biome == "Dryland" & division == "Angiosperm")
dryland_ang

dryland_gym <- tallo_wd_df0 |> filter(biome == "Dryland" & division == "Gymnosperm")
dryland_gym

# Mangrove
mangrove_ang <- tallo_wd_df0 |> filter(biome == "Mangrove" & division == "Angiosperm")
mangrove_ang

mangrove_gym <- tallo_wd_df0 |> filter(biome == "Mangrove" & division == "Gymnosperm")
mangrove_gym

```

# Plotting

```{r}
plot_biome <- function(df, biome_name, division_type) {

  biome_df <- df |> filter(biome == biome_name & division == division_type)

  cr_dbh_plot <- ggplot() +
    geom_bin2d(data = biome_df, aes(x = dbh, y = cr), bins = 100) +
    scale_fill_viridis_c(option = "D") +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "DBH (cm)", y = "CR (m)", title = paste("CR-DBH:", division_type, "-", biome_name)) +
    theme_minimal()

  h_dbh_plot <- ggplot() +
    geom_bin2d(data = biome_df, aes(x = dbh, y = h), bins = 100) +
    scale_fill_viridis_c(option = "D") +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "DBH (cm)", y = "H (m)", title = paste("H-DBH:", division_type, "-", biome_name)) +
    theme_minimal()

  list(cr_dbh_plot = cr_dbh_plot, h_dbh_plot = h_dbh_plot)
}

tallo_wd_df0 <- tar_read(tallo_wd_df0)

biomes <- c("Boreal/montane forest", "Tropical rain forest", "Temperate broadleaf forest", 
            "Tropical savanna", "Temperate grassland", "Temperate conifer forest", 
            "Mediterranean woodland", "Tropical dry forest", "Dryland", "Mangrove")

for (biome in biomes) {

  # Ang
  angiosperm_plots <- plot_biome(tallo_wd_df0, biome, "Angiosperm")

  print(angiosperm_plots$cr_dbh_plot)

  print(angiosperm_plots$h_dbh_plot)

  # Gym
  gymnosperm_plots <- plot_biome(tallo_wd_df0, biome, "Gymnosperm")

  print(gymnosperm_plots$cr_dbh_plot)

  print(gymnosperm_plots$h_dbh_plot)
}
```

### 2025-03-26 docs/check_H _new.qmd
Initial commit for tree-allo

---
title: "Plotting H-DBH allometry"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# Read in data
```{r}
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h
h_data <- tallo_reduced_lr_df_ang_h |> 
  filter(!is.na(h))
h_data

sp_posterior_h_df <- tar_read(sp_posterior_h_df)
head(sp_posterior_h_df)
dim(sp_posterior_h_df)

large_tree_species <- h_data |>
  group_by(sp) |>
  filter(any(dbh > 150)) |>
  ungroup()

large_tree_summary <- large_tree_species |>
  group_by(sp) |>
  summarize(
    min_dbh = min(dbh, na.rm = TRUE),
    max_dbh = max(dbh, na.rm = TRUE),
    mean_dbh = mean(dbh, na.rm = TRUE),
    count = n()
  ) |>
  arrange(desc(max_dbh))

top_30_species <- large_tree_summary |>
  slice(1:30)

top_30_species

top_30_data <- h_data |>
  filter(sp %in% top_30_species$sp) |>
  select(tree_id, sp, dbh, h)

sub_datasets <- top_30_data |>
  group_split(sp)

names(sub_datasets) <- unique(top_30_data$sp)
```

# Plot function
```{r}
# generate_plot <- function(sub_datasets, sp_posterior_h_df) {
#   combined_data <- data.frame()

#   for (i in seq_along(sub_datasets)) {
#       species_data <- sub_datasets[[i]]
#       species_name <- names(sub_datasets)[i]

#       species_post <- sp_posterior_h_df |>
#         filter(Species == species_name)

#       if (nrow(species_post) == 0) {
#         warning(paste("No posterior data for species:", species_name))
#         next
#       }

#       # Extract parameters
#       a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
#       log_a <- log(a)
#       b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
#       k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

#       dbh <- species_data$dbh

#       # Compute the fitted height values
#       fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
#       fitted_height <- exp(fitted_log_y)

#       # Create a dataframe for the fitted curve
#       fitted_curve <- data.frame(
#         DBH = dbh,
#         H = fitted_height,
#         sp = species_name,
#         Source = "Fitted"
#       )

#       # Combine observed and fitted data
#       combined_species_data <- species_data |>
#         mutate(Species = species_name, Source = "Observed") |>
#         dplyr::select(DBH = dbh, H = h, sp, Source) |>
#         bind_rows(fitted_curve)

#       # Append to the combined dataset
#       combined_data <- bind_rows(combined_data, combined_species_data)
#   }

#   p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
#         geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
#         geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#0f92e9", linewidth = 0.6) +
#         facet_wrap(~ sp, scales = "free", ncol = 4) +
#         labs(
#           x = "DBH (cm)",
#           y = "Tree Height (m)"
#         ) +
#         my_theme()
#   return(p)
# }

# generate_plot_log <- function(sub_datasets, sp_posterior_h_df) {
#   combined_data <- data.frame()

#   for (i in seq_along(sub_datasets)) {
#       species_data <- sub_datasets[[i]]
#       species_name <- names(sub_datasets)[i]

#       species_post <- sp_posterior_h_df |>
#         filter(Species == species_name)

#       if (nrow(species_post) == 0) {
#         warning(paste("No posterior data for species:", species_name))
#         next
#       }

#       # Extract parameters
#       a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
#       log_a <- log(a)
#       b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
#       k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

#       dbh <- species_data$dbh

#       # Compute the fitted height values
#       fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
#       fitted_height <- exp(fitted_log_y)

#       # Create a dataframe for the fitted curve
#       fitted_curve <- data.frame(
#         DBH = dbh,
#         H = fitted_height,
#         sp = species_name,
#         Source = "Fitted"
#       )

#       # Combine observed and fitted data
#       combined_species_data <- species_data |>
#         mutate(Species = species_name, Source = "Observed") |>
#         dplyr::select(DBH = dbh, H = h, sp, Source) |>
#         bind_rows(fitted_curve)

#       # Append to the combined dataset
#       combined_data <- bind_rows(combined_data, combined_species_data)
#   }

#   p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
#         geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
#         geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#0f92e9", linewidth = 0.6) +
#         facet_wrap(~ sp, scales = "free", ncol = 4) +
#         labs(
#           x = "DBH (cm)",
#           y = "Tree Height (m)"
#         ) +
#     scale_x_log10() +
#     scale_y_log10() +
#     my_theme()

#   return(p)
# }

generate_plot <- function(sub_datasets, sp_posterior_h_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
      species_data <- sub_datasets[[i]]
      species_name <- names(sub_datasets)[i]

      species_post <- sp_posterior_h_df |>
        filter(Species == species_name)

      if (nrow(species_post) == 0) {
        warning(paste("No posterior data for species:", species_name))
        next
      }

      # Extract parameters
      a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
      log_a <- log(a)
      b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
      k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

      # Extend DBH range
      dbh_min <- min(species_data$dbh, na.rm = TRUE)
      dbh_max <- max(species_data$dbh, na.rm = TRUE)
      extended_dbh <- seq(dbh_min * 0.1, dbh_max * 2, length.out = 100)

      # Compute the fitted height values
      fitted_log_y <- log_a + log(1 - exp(-b * (extended_dbh ^ k)))
      fitted_height <- exp(fitted_log_y)

      # Create a dataframe for the fitted curve
      fitted_curve <- data.frame(
        DBH = extended_dbh,
        H = fitted_height,
        sp = species_name,
        Source = "Fitted"
      )

      # Combine observed and fitted data
      combined_species_data <- species_data |>
        mutate(Species = species_name, Source = "Observed") |>
        dplyr::select(DBH = dbh, H = h, sp, Source) |>
        bind_rows(fitted_curve)

      # Append to the combined dataset
      combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
        geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
        geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#0f92e9", linewidth = 0.6) +
        facet_wrap(~ sp, scales = "free", ncol = 4) +
        labs(
          x = "DBH (cm)",
          y = "Tree Height (m)"
        ) +
        my_theme()
  return(p)
}

generate_plot_log <- function(sub_datasets, sp_posterior_h_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
      species_data <- sub_datasets[[i]]
      species_name <- names(sub_datasets)[i]

      species_post <- sp_posterior_h_df |>
        filter(Species == species_name)

      if (nrow(species_post) == 0) {
        warning(paste("No posterior data for species:", species_name))
        next
      }

      # Extract parameters
      a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
      log_a <- log(a)
      b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
      k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

      # Extend DBH range
      dbh_min <- min(species_data$dbh, na.rm = TRUE)
      dbh_max <- max(species_data$dbh, na.rm = TRUE)
      extended_dbh <- seq(dbh_min * 0.1, dbh_max * 2, length.out = 100)

      # Compute the fitted height values
      fitted_log_y <- log_a + log(1 - exp(-b * (extended_dbh ^ k)))
      fitted_height <- exp(fitted_log_y)

      # Create a dataframe for the fitted curve
      fitted_curve <- data.frame(
        DBH = extended_dbh,
        H = fitted_height,
        sp = species_name,
        Source = "Fitted"
      )

      # Combine observed and fitted data
      combined_species_data <- species_data |>
        mutate(Species = species_name, Source = "Observed") |>
        dplyr::select(DBH = dbh, H = h, sp, Source) |>
        bind_rows(fitted_curve)

      # Append to the combined dataset
      combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
        geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
        geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#0f92e9", linewidth = 0.6) +
        facet_wrap(~ sp, scales = "free", ncol = 4) +
        labs(
          x = "DBH (cm)",
          y = "Tree Height (m)"
        ) +
        scale_x_log10() +
        scale_y_log10() +
        my_theme()

  return(p)
}

```

```{r, fig.width=14, fig.height=20}
p <- generate_plot(sub_datasets, sp_posterior_h_df)
p
```

```{r, fig.width=14, fig.height=20}
p_log <- generate_plot_log(sub_datasets, sp_posterior_h_df)
p_log
```

### 2025-03-26 docs/check_H.qmd
Initial commit for tree-allo

---
title: "Checking H-DBH allometry for angiosperms"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# Species ID

```{r}
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h
ang_data_h <- tallo_reduced_lr_df_ang_h |> 
  filter(division == "Angiosperm") |>
  filter(!is.na(h)) |>

ang_data_h

wd_df <- ang_data_h |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(sp_id = 1:n()) |>
  mutate(wd_s = scale(wd) |> as.numeric()) |>
  dplyr::select(sp_id, sp, wd, wd_s)
head(wd_df)
str(wd_df)
summary(wd_df)

stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
str(stan_data_nlr_ang_h)

h_stan_ang <- data.frame(
  sp_id = as.factor(stan_data_nlr_ang_h$jj),
  # dbh = stan_data_nlr_ang_h$x,
  wd_s = stan_data_nlr_ang_h$u[stan_data_nlr_ang_h$jj, 2])

head(h_stan_ang)
# id_66 <- wd_df |> filter(wd_df$sp_id == "66")
# id_66

h_stan_ang <- h_stan_ang |>
  group_by(sp_id, wd_s) |>
  summarise(tree_count = n(), .groups = "drop")

head(h_stan_ang)
str(h_stan_ang)
# id_66_stan <- h_stan_ang |> filter(h_stan_ang$sp_id == "66")
# id_66_stan

h_ang <- h_stan_ang |>
  mutate(sp_id = as.integer(as.character(sp_id))) |>
  left_join(wd_df |> mutate(sp_id = as.integer(sp_id)) |> select(sp_id, sp), by = "sp_id")
head(h_ang)
str(h_ang)

```

# Extracting sp-specific parameter

```{r}
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_nlr_nou_summary_weibull_ang_h

create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
  if (functional_form == "Weibull" || functional_form == "gMM") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

    k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
    k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
    k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  } else if (functional_form == "Power-Law") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

    c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
    c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
    c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  }

  return(ci_95)
}

h_posterior_df <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang)

head(h_posterior_df)

h_posterior_df <- h_posterior_df |>
  rename(sp = sp.sp)

head(h_posterior_df)

head(h_posterior_df)
ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")
ang_data_h
summary(ang_data_h)

large_tree_species <- ang_data_h |>
  group_by(sp) |>
  filter(any(dbh > 150)) |>
  ungroup()

large_tree_summary <- large_tree_species |>
  group_by(sp) |>
  summarize(
    min_dbh = min(dbh, na.rm = TRUE),
    max_dbh = max(dbh, na.rm = TRUE),
    mean_dbh = mean(dbh, na.rm = TRUE),
    count = n()
  ) |>
  arrange(desc(max_dbh))

top_30_species <- large_tree_summary |>
  slice(1:30)

top_30_species

top_30_data <- ang_data_h |>
  filter(sp %in% top_30_species$sp) |>
  select(tree_id, sp, dbh, h)

sub_datasets <- top_30_data |>
  group_split(sp)

names(sub_datasets) <- unique(top_30_data$sp)
```

# Function
```{r}
generate_plot <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    my_theme()

  return(p)
}

generate_plot_log <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
       dplyr::select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    scale_x_log10() +
    scale_y_log10() +
    my_theme()

  return(p)
}
```

# Plotting

```{r, fig.width=14, fig.height=20}
p <- generate_plot(sub_datasets, h_posterior_df)
p

p_log <- generate_plot_log(sub_datasets, h_posterior_df)
p_log
```

# SPECIES POSTERIOR DATAFRAME

```{r}
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_lr_nou_summary_pl_ang_cr <- tar_read(fit_lr_nou_summary_pl_ang_cr)
fit_nlr_nou_summary_gmm_gym_cr <- tar_read(fit_nlr_nou_summary_gmm_gym_cr)
fit_lr_nou_summary_pl_ang_dbh <- tar_read(fit_lr_nou_summary_pl_ang_dbh)
fit_lr_nou_summary_pl_gym_dbh <- tar_read(fit_lr_nou_summary_pl_gym_dbh)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
stan_data_nlr_gym_cr <- tar_read(stan_data_nlr_gym_cr)
stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
```

```{r}
str(stan_data_nlr_ang_h)
str(stan_data_nlr_gym_h)
str(stan_data_lr_ang_cr)
str(stan_data_nlr_gym_cr)
str(stan_data_lr_ang_dbh)
str(stan_data_lr_gym_dbh)

  filter_species <- function(dataset, division) {
    dataset |> 
      filter(division == division, !is.na(dbh), !is.na(cr), !is.na(h)) |> 
      group_by(sp) |> 
      summarise(count = n()) |> 
      filter(count >= 20) |> 
      pull(sp)
  }
  create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
    if (functional_form == "Weibull" || functional_form == "gMM") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

      k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
      k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
      k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    } else if (functional_form == "Power-Law") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

      c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
      c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
      c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    }

    return(ci_95)
  }
  # Function to process datasets and align species IDs with Stan data
 process_dataset <- function(dataset, stan_data, division_filter) {
    # Filter the dataset by division
    dataset <- dataset |> filter(division == division_filter)

    # Prepare species data
    sp_data <- dataset |>
      group_by(sp) |>
      summarise(wd = mean(wd, na.rm = TRUE)) |>
      arrange(sp) |>
      mutate(sp_id = row_number()) |>
      mutate(wd_s = scale(wd) |> as.numeric()) |>
      select(sp_id, sp, wd, wd_s)

    # Verify Stan data structure
    if (!all(c("jj", "u") %in% names(stan_data))) {
      stop("Stan data must contain 'jj' and 'u'.")
    }

    # Handle `nlr` structure: `u` has rows indexed by `jj`
    if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
      stan_processed <- data.frame(
        sp_id = as.factor(stan_data$jj),
        wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
      )
    } 
    # Handle `lr` structure: `u` has two rows (variables) and columns for species
    else if (nrow(stan_data$u) == 2) {
      stan_processed <- data.frame(
        sp_id = 1:ncol(stan_data$u),
        wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
      )
    } 
    # Error handling for unexpected structures
    else {
      stop("Unexpected structure for 'u' in Stan data.")
    }

    # Add species names and calculate tree counts
    stan_processed <- stan_processed |>
      group_by(sp_id, wd_s) |>
      summarise(tree_count = n(), .groups = "drop") |>
      mutate(sp_id = as.integer(as.character(sp_id))) |>
      left_join(sp_data, by = "sp_id")

    return(list(sp_data = sp_data, stan_processed = stan_processed))
  }

```

```{r}
h_ang <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_ang_h, "Angiosperm")

h_ang

h_gym <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_gym_h, "Gymnosperm")
h_gym

cr_ang <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_lr_ang_cr, "Angiosperm")

cr_ang

cr_gym <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_nlr_gym_cr, "Gymnosperm")
cr_gym

# dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh, stan_data_lr_ang_dbh, "Angiosperm")
# dbh_ang

# dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh, stan_data_lr_gym_dbh, "Gymnosperm")
# dbh_gym
ang_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Angiosperm")
gym_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Gymnosperm")

dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% ang_dbh_species), stan_data_lr_ang_dbh, "Angiosperm")

dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% gym_dbh_species), stan_data_lr_gym_dbh, "Gymnosperm")

```

```{r}
h_df_ang <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang$sp)
head(h_df_ang)

h_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_gym$sp)
head(h_df_gym)

cr_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_cr, "Crown Radius", "Power-Law", cr_ang$sp)
head(cr_df_ang)

cr_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_cr, "Crown Radius", "gMM", cr_gym$sp)
head(cr_df_gym)

dbh_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh, "DBH", "Power-Law", dbh_ang$sp)
head(dbh_df_ang)

dbh_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh, "DBH", "Power-Law", dbh_gym$sp)
head(dbh_df_gym)

```

```{r}
sp_posterior_df <- bind_rows(
  h_df_ang |> mutate(Division = "Angiosperm"),
  h_df_gym |> mutate(Division = "Gymnosperm"),
  cr_df_ang |> mutate(Division = "Angiosperm"),
  cr_df_gym |> mutate(Division = "Gymnosperm"),
  dbh_df_ang |> mutate(Division = "Angiosperm"),
  dbh_df_gym |> mutate(Division = "Gymnosperm")
)

head(sp_posterior_df)
```

```{r}
  filter_species <- function(dataset, division) {
    dataset |> 
      filter(division == division, !is.na(dbh), !is.na(cr), !is.na(h)) |> 
      group_by(sp) |> 
      summarise(count = n()) |> 
      filter(count >= 20) |> 
      pull(sp)
  }

  # Helper function to extract posterior summaries and prepare species-specific output
  create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
    if (functional_form == "Weibull" || functional_form == "gMM") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

      k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
      k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
      k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

      # Ensure species list and parameters match
      if (length(sp_list) != length(a_median)) {
        stop("Mismatch in species list and parameter lengths.")
      }

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    } else if (functional_form == "Power-Law") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

      c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
      c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
      c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

      # Ensure species list and parameters match
      if (length(sp_list) != length(a_median)) {
        stop("Mismatch in species list and parameter lengths.")
      }

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    }

    return(ci_95)
  }

  # Function to process datasets and align species IDs with Stan data
  process_dataset <- function(dataset, stan_data, division_filter) {
    # Filter the dataset by division
    dataset <- dataset |> filter(division == division_filter)

    # Prepare species data
    sp_data <- dataset |>
      group_by(sp) |>
      summarise(wd = mean(wd, na.rm = TRUE)) |>
      arrange(sp) |>
      mutate(sp_id = row_number()) |>
      mutate(wd_s = scale(wd) |> as.numeric()) |>
      select(sp_id, sp, wd, wd_s)

    # Verify Stan data structure
    if (!all(c("jj", "u") %in% names(stan_data))) {
      stop("Stan data must contain 'jj' and 'u'.")
    }

    # Handle `nlr` structure: `u` has rows indexed by `jj`
    if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
      stan_processed <- data.frame(
        sp_id = as.factor(stan_data$jj),
        wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
      )
    } else if (nrow(stan_data$u) == 2) {
      stan_processed <- data.frame(
        sp_id = 1:ncol(stan_data$u),
        wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
      )
    } else {
      stop("Unexpected structure for 'u' in Stan data.")
    }

    # Add species names and calculate tree counts
    stan_processed <- stan_processed |>
      group_by(sp_id, wd_s) |>
      summarise(tree_count = n(), .groups = "drop") |>
      mutate(sp_id = as.integer(as.character(sp_id))) |>
      left_join(sp_data, by = "sp_id")

    return(list(sp_data = sp_data, stan_processed = stan_processed))
  }

  # Filter species
  ang_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Angiosperm")
  gym_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Gymnosperm")

  # Process datasets for Angiosperms and Gymnosperms
  h_ang <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_ang_h, "Angiosperm")
  h_gym <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_gym_h, "Gymnosperm")
  cr_ang <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_lr_ang_cr, "Angiosperm")
  cr_gym <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_nlr_gym_cr, "Gymnosperm")
  dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% ang_dbh_species), stan_data_lr_ang_dbh, "Angiosperm")
  dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% gym_dbh_species), stan_data_lr_gym_dbh, "Gymnosperm")

  # Generate posterior dataframes
  h_df_ang <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang$sp)
  h_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_gym$sp)
  cr_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_cr, "Crown Radius", "Power-Law", cr_ang$sp)
  cr_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_cr, "Crown Radius", "gMM", cr_gym$sp)
  dbh_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh, "DBH", "Power-Law", dbh_ang$sp)
  dbh_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh, "DBH", "Power-Law", dbh_gym$sp)

  # # Combine results
  # sp_posterior_df <- bind_rows(
  #   h_df_ang |> mutate(Division = "Angiosperm"),
  #   h_df_gym |> mutate(Division = "Gymnosperm"),
  #   cr_df_ang |> mutate(Division = "Angiosperm"),
  #   cr_df_gym |> mutate(Division = "Gymnosperm"),
  #   dbh_df_ang |> mutate(Division = "Angiosperm"),
  #   dbh_df_gym |> mutate(Division = "Gymnosperm")
  # )

```

```{r}

```

### 2025-03-26 docs/check_H_gym.qmd
Initial commit for tree-allo

---
title: "Plotting H-DBH allometry"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# Extract parameters of gymnosperms H-DBH from 3 functional forms

```{r}
create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
  if (functional_form == "Weibull" || functional_form == "gMM") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

    k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
    k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
    k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(format(a_median, scientific = FALSE), " (", format(a_lower, scientific = FALSE), ", ", format(a_upper, scientific = FALSE), ")"),
      b = paste0(format(b_median, scientific = FALSE), " (", format(b_lower, scientific = FALSE), ", ", format(b_upper, scientific = FALSE), ")"),
      k = paste0(format(k_median, scientific = FALSE), " (", format(k_lower, scientific = FALSE), ", ", format(k_upper, scientific = FALSE), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  } else if (functional_form == "Power-Law") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

    c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
    c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
    c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

  ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(format(a_median, scientific = FALSE), " (", format(a_lower, scientific = FALSE), ", ", format(a_upper, scientific = FALSE), ")"),
        b = paste0(format(b_median, scientific = FALSE), " (", format(b_lower, scientific = FALSE), ", ", format(b_upper, scientific = FALSE), ")"),
        c = paste0(format(c_median, scientific = FALSE), " (", format(c_lower, scientific = FALSE), ", ", format(c_upper, scientific = FALSE), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
    )
  }

  return(ci_95)
}
# Function to process datasets and align species IDs with Stan data
process_dataset <- function(dataset, stan_data, division_filter) {
# Filter the dataset by division
dataset <- dataset |> filter(division == division_filter)

# Prepare species data
sp_data <- dataset |>
  group_by(sp) |>
  summarise(wd = mean(wd, na.rm = TRUE)) |>
  arrange(sp) |>
  mutate(sp_id = row_number()) |>
  mutate(wd_s = scale(wd) |> as.numeric()) |>
  dplyr::select(sp_id, sp, wd, wd_s)

if (!all(c("jj", "u") %in% names(stan_data))) {
  stop("Stan data must contain 'jj' and 'u'.")
}

# Handle `nlr` structure: `u` has rows indexed by `jj`
if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
  stan_processed <- data.frame(
    sp_id = as.factor(stan_data$jj),
    wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
  )
} 
# Handle `lr` structure: `u` has two rows (variables) and columns for species
else if (nrow(stan_data$u) == 2) {
  stan_processed <- data.frame(
    sp_id = 1:ncol(stan_data$u),
    wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
  )
} 
# Error handling for unexpected structures
else {
  stop("Unexpected structure for 'u' in Stan data.")
}

# Add species names and calculate tree counts
stan_processed <- stan_processed |>
  group_by(sp_id, wd_s) |>
  summarise(tree_count = n(), .groups = "drop") |>
  mutate(sp_id = as.integer(as.character(sp_id))) |>
  left_join(sp_data, by = "sp_id")

return(list(sp_data = sp_data, stan_processed = stan_processed))
}
```

```{r}
stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_gym_h <- tar_read(stan_data_lr_gym_h)

tallo_reduced_nlr_df_gym_h <- tar_read(tallo_reduced_nlr_df_gym_h)
tallo_reduced_nlr_df_gym_h

tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
tallo_reduced_lr_df_gym_h

fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_nlr_nou_summary_weibull_gym_h

fit_lr_nou_summary_pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
fit_lr_nou_summary_pl_gym_h

fit_nlr_nou_summary_gmm_gym_h <- tar_read(fit_nlr_nou_summary_gmm_gym_h)
fit_nlr_nou_summary_gmm_gym_h

h_wb_gym <- process_dataset(tallo_reduced_nlr_df_gym_h, stan_data_nlr_gym_h, "Gymnosperm")
head(h_wb_gym)

h_pl_gym <- process_dataset(tallo_reduced_lr_df_gym_h, stan_data_lr_gym_h, "Gymnosperm")
head(h_pl_gym)

h_gmm_gym <- process_dataset(tallo_reduced_nlr_df_gym_h, stan_data_nlr_gym_h, "Gymnosperm")
head(h_gmm_gym)

h_wb_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_wb_gym$sp_data$sp)
head(h_wb_df_gym)

h_pl_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_h, "Tree Height", "Power-Law", h_pl_gym$sp_data$sp)
head(h_pl_df_gym)

h_gmm_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_h, "Tree Height", "gMM", h_gmm_gym$sp_data$sp)
head(h_gmm_df_gym)

```

```{r}
tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
tallo_reduced_lr_df_gym_h

h_data <- tallo_reduced_lr_df_gym_h |> 
    filter(division == "Gymnosperm") |>
    filter(!is.na(h)) |>
    group_by(sp) |>
    ungroup()

h_data
num_species = 10

species_names <- sample(unique(h_data$sp), min(num_species, n_distinct(h_data$sp)))

# Split the dataset by species
sub_datasets <- h_data |>
  filter(sp %in% species_names) |>
  group_split(sp)

# Assign species names as dataset names
names(sub_datasets) <- species_names

```

```{r}
generate_plot <- function(sub_datasets, gmm_df, weibull_df, pl_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Filter data for each model
    gmm_post <- gmm_df |> filter(sp == species_name)
    weibull_post <- weibull_df |> filter(sp == species_name)
    pl_post <- pl_df |> filter(sp == species_name)

    if (nrow(gmm_post) == 0 || nrow(weibull_post) == 0 || nrow(pl_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Use only observed DBH range for predictions
    dbh_values <- species_data$dbh

    # Power-Law model
    a_pl <- as.numeric(sub(" \\(.*\\)", "", pl_post$a))
    b_pl <- as.numeric(sub(" \\(.*\\)", "", pl_post$b))
    pl_pred <- a_pl * (dbh_values^b_pl)

    # gMM model
    a_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$a))
    b_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$b))
    k_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$k))
    gmm_pred <- a_gmm * (dbh_values^b_gmm) / (k_gmm + dbh_values^b_gmm)

    # Weibull model
    a_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$a))
    b_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$b))
    k_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$k))
    wb_pred <- a_wb * (1 - exp(-b_wb * dbh_values^k_wb))

    # Create fitted curves for each model
    model_predictions <- list()
    model_predictions[["Power-Law"]] <- data.frame(DBH = dbh_values, H = pl_pred, sp = species_name, Model = "Power-Law")
    model_predictions[["gMM"]] <- data.frame(DBH = dbh_values, H = gmm_pred, sp = species_name, Model = "gMM")
    model_predictions[["Weibull"]] <- data.frame(DBH = dbh_values, H = wb_pred, sp = species_name, Model = "Weibull")

    # Combine observed and fitted data
    fitted_data <- bind_rows(model_predictions)
    observed_data <- species_data |>
      mutate(DBH = dbh, H = h, Model = "Observed", sp = species_name) |>
      dplyr::select(DBH, H, sp, Model)

    combined_species_data <- bind_rows(fitted_data, observed_data)
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  # Create the plot
  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Model)) +
    geom_point(data = combined_data |> filter(Model == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Model != "Observed"), aes(linetype = Model), linewidth = 0.6, linetype = "solid") +
    facet_wrap(~ sp, scales = "free", ncol = 3) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    scale_color_manual(values = c("gMM" = "#1b9e77", "Weibull" = "#0f92e9", "Power-Law" = "#FF5733", "Observed" = "gray")) +
    my_theme()

  return(p)
}
```

```{r, fig.width=30, fig.height=60, out.width='100%', out.height='100%'}
p <- generate_plot(sub_datasets, h_gmm_df_gym, h_wb_df_gym, h_pl_df_gym)
p
```

### 2025-03-26 docs/check_dat.qmd
Initial commit for tree-allo

---
title: "Data checking"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r include=FALSE}
library(tidyverse)
library(here)
library(targets)
library(knitr)
options(width=200)
```

# Read in data

```{r}
tmp <- tar_manifest()

tmp |>
  filter(str_detect(name, "summary")) |>
  filter(str_detect(name, "wd"))

tallo_wd_df <- tar_read(tallo_wd_df)
tallo_wd_df
summary(tallo_wd_df)
count_sp <- tallo_wd_df |>
  group_by(sp) |>
  summarise(count = n()
)
count_sp

```

# Overlapping data
```{r}
tallo_wd_df0 <- tar_read(tallo_wd_df0)
tallo_wd_df0
count_sp0 <- tallo_wd_df0 |>
  group_by(sp) |>
  summarise(count = n()
)
count_sp0

```

- 355,881 trees
- 2,961 sp

# H data

```{r}
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm") |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)

ang_h_sp

gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm") |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
gym_h_sp

stan_data_lr_ang_h <- tar_read(stan_data_lr_ang_h)
str(stan_data_lr_ang_h)

stan_data_lr_gym_h <- tar_read(stan_data_lr_gym_h)
str(stan_data_lr_gym_h)

```

# CR data

```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
cr_sp <- tallo_reduced_lr_df_ang_cr |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
cr_sp

ang_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
ang_data_cr
ang_cr_sp <- ang_data_cr |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
ang_cr_sp

gym_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Gymnosperm")
gym_data_cr
gym_cr_sp <- gym_data_cr |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
gym_cr_sp

stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
str(stan_data_lr_ang_cr)

stan_data_lr_gym_cr <- tar_read(stan_data_lr_gym_cr)
str(stan_data_lr_gym_cr)

```

# DBH data

```{r}
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
tallo_reduced_lr_df_ang_dbh
dbh_sp <- tallo_reduced_lr_df_ang_dbh |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
dbh_sp

ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |> filter(division == "Angiosperm")
ang_data_dbh

ang_dbh_data <- ang_data_dbh |> filter(!is.na(dbh), !is.na(cr), !is.na(h)) |> group_by(sp)

ang_dbh_sp <- ang_dbh_data |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
ang_dbh_sp

# 744

gym_data_dbh <- tallo_reduced_lr_df_ang_dbh |> filter(division == "Gymnosperm")
gym_data_dbh

gym_dbh_data <- gym_data_dbh |> filter(!is.na(dbh), !is.na(cr), !is.na(h)) |> group_by(sp)

gym_dbh_sp
gym_dbh_sp <- gym_dbh_data |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
gym_dbh_sp

#56
#now it ok

stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
str(stan_data_lr_ang_dbh)

stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
str(stan_data_lr_gym_dbh)

# tallo_wd_df0 <- tar_read(tallo_wd_df0)
# tallo_wd_df0

# dbh_data <- tallo_wd_df0 |> filter(!is.na(dbh), !is.na(cr), !is.na(h)) |> group_by(sp) |> filter(n() >= 20) |> ungroup()
# dbh_data

# dbh_sp <- dbh_data |>
#   group_by(sp) |>
#   summarise(count = n())

# dbh_sp
# Load the tallo_reduced_lr_df for DBH1 data processing
tallo_reduced_lr_df_ang_dbh1 <- tar_read(tallo_reduced_lr_df_ang_dbh1)
tallo_reduced_lr_df_ang_dbh1

# Summarise species count for Angiosperms with DBH1 data
dbh1_sp <- tallo_reduced_lr_df_ang_dbh1 |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
dbh1_sp

# Filter Angiosperm data with DBH1
ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |> filter(division == "Angiosperm")
ang_data_dbh1

# Filter and clean the DBH1 data for Angiosperms
ang_dbh1_data <- ang_data_dbh1 |> filter(!is.na(dbh), !is.na(cr), !is.na(h)) |> group_by(sp)

# Summarise species with enough samples (>= 20) for Angiosperms
ang_dbh1_sp <- ang_dbh1_data |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
ang_dbh1_sp

# Gymnosperm data processing for DBH1
gym_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |> filter(division == "Gymnosperm")
gym_data_dbh1

# Filter and clean the DBH1 data for Gymnosperms
gym_dbh1_data <- gym_data_dbh1 |> filter(!is.na(dbh), !is.na(cr), !is.na(h)) |> group_by(sp)

# Summarise species with enough samples (>= 20) for Gymnosperms
gym_dbh1_sp <- gym_dbh1_data |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
gym_dbh1_sp

# Load Stan data for DBH1 for Angiosperms and Gymnosperms
stan_data_lr_ang_dbh1 <- tar_read(stan_data_lr_ang_dbh1)
str(stan_data_lr_ang_dbh1)

ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
    filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
    group_by(sp) |> filter(n() >= 20) |>
    ungroup()

stan_data_lr_gym_dbh1 <- tar_read(stan_data_lr_gym_dbh1)
str(stan_data_lr_gym_dbh1)

stan_data_lr_ang_dbh2 <- tar_read(stan_data_lr_ang_dbh2)
str(stan_data_lr_ang_dbh2)

stan_data_lr_gym_dbh2 <- tar_read(stan_data_lr_gym_dbh2)
str(stan_data_lr_gym_dbh2)

  ang_data_dbh2 <- tallo_reduced_lr_df_ang_dbh2 |>
    filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
    group_by(sp) |> filter(n() >= 20) |>
    ungroup()

stan_data_lr_ang_dbh3 <- tar_read(stan_data_lr_ang_dbh3)
str(stan_data_lr_ang_dbh3)

stan_data_lr_gym_dbh3 <- tar_read(stan_data_lr_gym_dbh3)
str(stan_data_lr_gym_dbh3)

stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
str(stan_data_lr_ang_dbh)

stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
str(stan_data_lr_gym_dbh)

```

# Characteristics of dataset and sub-datasets

```{r}
# Dataset
# d0 <- tar_read(tallo_wd_df0)
# ori_sample_size <- nrow(d0)
# ori_species_number <- d0 |>
#   distinct(sp) |>
#   nrow()

# H allometry sub dataset
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")

ang_h_obs <- nrow(ang_data_h)
ang_h_sp <- ang_data_h |>
  distinct(sp) |>
  nrow()

ang_h_trait_ranges <- ang_data_h |>
    dplyr::select(dbh, h, wd) |>
    apply(2, range)

gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm")

gym_h_obs <- nrow(gym_data_h)
gym_h_sp <- gym_data_h |>
  distinct(sp) |>
  nrow()

gym_h_trait_ranges <- gym_data_h |>
    dplyr::select(dbh, h, wd) |>
    apply(2, range)

# CR allometry sub dataset
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

ang_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
ang_cr_obs <- nrow(ang_data_cr)
ang_cr_sp <- ang_data_cr |>
  distinct(sp) |>
  nrow()
ang_cr_trait_ranges <- ang_data_cr |>
  dplyr::select(dbh, cr, wd) |>
  apply(2, range)

gym_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Gymnosperm")
gym_cr_obs <- nrow(gym_data_cr)
gym_cr_sp <- gym_data_cr |>
  distinct(sp) |>
  nrow()
gym_cr_trait_ranges <- gym_data_cr |>
  dplyr::select(dbh, cr, wd) |>
  apply(2, range)

# DBH allometry sub dataset
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)

# Angiosperm data
ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |> 
  filter(division == "Angiosperm", 
         !is.na(dbh), 
         !is.na(cr), 
         !is.na(h)) |> 
  group_by(sp) |> 
  filter(n() >= 20) |>
  ungroup()

ang_dbh_obs <- nrow(ang_data_dbh)
ang_dbh_sp <- ang_data_dbh |> 
  distinct(sp) |> 
  nrow()
ang_dbh_trait_ranges <- ang_data_dbh |> 
  dplyr::select(dbh, cr, h, wd) |> 
  apply(2, range)

# Gymnosperm data
gym_data_dbh <- tallo_reduced_lr_df_ang_dbh |> 
  filter(division == "Gymnosperm", 
         !is.na(dbh), 
         !is.na(cr), 
         !is.na(h)) |> 
  group_by(sp) |> 
  filter(n() >= 20) |>
  ungroup()

gym_dbh_obs <- nrow(gym_data_dbh)
gym_dbh_sp <- gym_data_dbh |> 
  distinct(sp) |> 
  nrow()
gym_dbh_trait_ranges <- gym_data_dbh |> 
  dplyr::select(dbh, cr, h, wd) |> 
  apply(2, range)

```

## TableS1
```{r}
library(dplyr)
library(tidyr)

combined_data <- tibble(
  `Dependent variables` = rep(c("H", "CR", "DBH"), each = 2),
  Division = rep(c("Angiosperm", "Gymnosperm"), times = 3),
  `Number of trees` = c(
    nrow(ang_data_h), nrow(gym_data_h),
    nrow(ang_data_cr), nrow(gym_data_cr),
    nrow(ang_data_dbh), nrow(gym_data_dbh)
  ),
  `Number of species` = c(
    ang_data_h |> distinct(sp) |> nrow(),
    gym_data_h |> distinct(sp) |> nrow(),
    ang_data_cr |> distinct(sp) |> nrow(),
    gym_data_cr |> distinct(sp) |> nrow(),
    ang_data_dbh |> distinct(sp) |> nrow(),
    gym_data_dbh |> distinct(sp) |> nrow()
  ),
  `DBH range` = c(
    paste0(round(min(ang_data_h$dbh), 2), " - ", round(max(ang_data_h$dbh), 2)),
    paste0(round(min(gym_data_h$dbh), 2), " - ", round(max(gym_data_h$dbh), 2)),
    paste0(round(min(ang_data_cr$dbh), 2), " - ", round(max(ang_data_cr$dbh), 2)),
    paste0(round(min(gym_data_cr$dbh), 2), " - ", round(max(gym_data_cr$dbh), 2)),
    paste0(round(min(ang_data_dbh$dbh), 2), " - ", round(max(ang_data_dbh$dbh), 2)),
    paste0(round(min(gym_data_dbh$dbh), 2), " - ", round(max(gym_data_dbh$dbh), 2))
  ),
  `Dependent variable range` = c(
    paste0(round(min(ang_data_h$h), 2), " - ", round(max(ang_data_h$h), 2)),
    paste0(round(min(gym_data_h$h), 2), " - ", round(max(gym_data_h$h), 2)),
    paste0(round(min(ang_data_cr$cr), 2), " - ", round(max(ang_data_cr$cr), 2)),
    paste0(round(min(gym_data_cr$cr), 2), " - ", round(max(gym_data_cr$cr), 2)),
    paste0(round(min(ang_data_dbh$cr), 2), " - ", round(max(ang_data_dbh$cr), 2)),
    paste0(round(min(gym_data_dbh$cr), 2), " - ", round(max(gym_data_dbh$cr), 2))
  ),
  `WD range` = c(
    paste0(round(min(ang_data_h$wd), 3), " - ", round(max(ang_data_h$wd), 3)),
    paste0(round(min(gym_data_h$wd), 3), " - ", round(max(gym_data_h$wd), 3)),
    paste0(round(min(ang_data_cr$wd), 3), " - ", round(max(ang_data_cr$wd), 3)),
    paste0(round(min(gym_data_cr$wd), 3), " - ", round(max(gym_data_cr$wd), 3)),
    paste0(round(min(ang_data_dbh$wd), 3), " - ", round(max(ang_data_dbh$wd), 3)),
    paste0(round(min(gym_data_dbh$wd), 3), " - ", round(max(gym_data_dbh$wd), 3))
  ),
  Relationship = c(
    "H-DBH", "H-DBH",
    "CR-DBH", "CR-DBH",
    "DBH-CR,H", "DBH-CR,H"
  )
)

combined_data

combined_data_fixed <- combined_data |>
  mutate(across(c(`Number of trees`, `Number of species`, `DBH range`, `Dependent variable range`, `WD range`), as.character))

formatted_table <- combined_data_fixed |>
  pivot_longer(
    cols = c(`Number of trees`, `Number of species`, `DBH range`, `Dependent variable range`, `WD range`),
    names_to = "Characteristics",
    values_to = "Values"
  ) |>
  pivot_wider(
    names_from = Division,
    values_from = Values
  ) |>
  dplyr::select(`Dependent variables`, Characteristics, Relationship, Angiosperm, Gymnosperm)

formatted_table

```

```{r}
library(targets)
library(dplyr)
library(tidyr)

# Read and process H allometry sub-dataset
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")
gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm")

ang_h_obs <- nrow(ang_data_h)
ang_h_sp <- ang_data_h |> distinct(sp) |> nrow()
ang_h_trait_ranges <- ang_data_h |> dplyr::select(dbh, h, wd) |> apply(2, range)

gym_h_obs <- nrow(gym_data_h)
gym_h_sp <- gym_data_h |> distinct(sp) |> nrow()
gym_h_trait_ranges <- gym_data_h |> dplyr::select(dbh, h, wd) |> apply(2, range)

# Read and process CR allometry sub-dataset
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
gym_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Gymnosperm")

ang_cr_obs <- nrow(ang_data_cr)
ang_cr_sp <- ang_data_cr |> distinct(sp) |> nrow()
ang_cr_trait_ranges <- ang_data_cr |> dplyr::select(dbh, cr, wd) |> apply(2, range)

gym_cr_obs <- nrow(gym_data_cr)
gym_cr_sp <- gym_data_cr |> distinct(sp) |> nrow()
gym_cr_trait_ranges <- gym_data_cr |> dplyr::select(dbh, cr, wd) |> apply(2, range)

# Read and process DBH allometry sub-dataset
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)

ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

gym_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

ang_dbh_obs <- nrow(ang_data_dbh)
ang_dbh_sp <- ang_data_dbh |> distinct(sp) |> nrow()
ang_dbh_trait_ranges <- ang_data_dbh |> dplyr::select(dbh, cr, h, wd) |> apply(2, range)

gym_dbh_obs <- nrow(gym_data_dbh)
gym_dbh_sp <- gym_data_dbh |> distinct(sp) |> nrow()
gym_dbh_trait_ranges <- gym_data_dbh |> dplyr::select(dbh, cr, h, wd) |> apply(2, range)

# Combine the data
combined_data <- tibble(
  `Dependent variables` = rep(c("H", "CR", "DBH"), each = 2),
  Division = rep(c("Angiosperm", "Gymnosperm"), times = 3),
  `Number of trees` = as.character(c(
    nrow(ang_data_h), nrow(gym_data_h),
    nrow(ang_data_cr), nrow(gym_data_cr),
    nrow(ang_data_dbh), nrow(gym_data_dbh)
  )),
  `Number of species` = as.character(c(
    ang_data_h |> distinct(sp) |> nrow(),
    gym_data_h |> distinct(sp) |> nrow(),
    ang_data_cr |> distinct(sp) |> nrow(),
    gym_data_cr |> distinct(sp) |> nrow(),
    ang_data_dbh |> distinct(sp) |> nrow(),
    gym_data_dbh |> distinct(sp) |> nrow()
  )),
  `DBH range` = c(
    range(ang_data_h$dbh) |> paste(collapse = " - "), range(gym_data_h$dbh) |> paste(collapse = " - "),
    range(ang_data_cr$dbh) |> paste(collapse = " - "), range(gym_data_cr$dbh) |> paste(collapse = " - "),
    range(ang_data_dbh$dbh) |> paste(collapse = " - "), range(gym_data_dbh$dbh) |> paste(collapse = " - ")
  ),
  `Dependent variable range` = c(
    range(ang_data_h$h) |> paste(collapse = " - "), range(gym_data_h$h) |> paste(collapse = " - "),
    range(ang_data_cr$cr) |> paste(collapse = " - "), range(gym_data_cr$cr) |> paste(collapse = " - "),
    range(ang_data_dbh$cr) |> paste(collapse = " - "), range(gym_data_dbh$cr) |> paste(collapse = " - ")
  ),
  `WD range` = c(
    range(ang_data_h$wd) |> paste(collapse = " - "), range(gym_data_h$wd) |> paste(collapse = " - "),
    range(ang_data_cr$wd) |> paste(collapse = " - "), range(gym_data_cr$wd) |> paste(collapse = " - "),
    range(ang_data_dbh$wd) |> paste(collapse = " - "), range(gym_data_dbh$wd) |> paste(collapse = " - ")
  ),
  Relationship = rep(c("H-DBH", "CR-DBH", "DBH-CR,H"), each = 2)
)

formatted_table <- combined_data |>
  pivot_longer(
    cols = c(`Number of trees`, `Number of species`, `DBH range`, `Dependent variable range`, `WD range`),
    names_to = "Characteristics",
    values_to = "Values"
  ) |>
  pivot_wider(
    names_from = Division,
    values_from = Values
  ) |>
  dplyr::select(`Dependent variables`, Characteristics, Relationship, Angiosperm, Gymnosperm)

formatted_table
```

# Take 3 digits for WD

```{r}
library(dplyr)
library(tidyr)
library(targets)

# Read and process H allometry sub-dataset
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")
gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm")

ang_h_obs <- nrow(ang_data_h)
ang_h_sp <- ang_data_h |> distinct(sp) |> nrow()
ang_h_trait_ranges <- ang_data_h |> dplyr::select(dbh, h, wd) |> apply(2, range)

gym_h_obs <- nrow(gym_data_h)
gym_h_sp <- gym_data_h |> distinct(sp) |> nrow()
gym_h_trait_ranges <- gym_data_h |> dplyr::select(dbh, h, wd) |> apply(2, range)

# Read and process CR allometry sub-dataset
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
gym_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Gymnosperm")

ang_cr_obs <- nrow(ang_data_cr)
ang_cr_sp <- ang_data_cr |> distinct(sp) |> nrow()
ang_cr_trait_ranges <- ang_data_cr |> dplyr::select(dbh, cr, wd) |> apply(2, range)

gym_cr_obs <- nrow(gym_data_cr)
gym_cr_sp <- gym_data_cr |> distinct(sp) |> nrow()
gym_cr_trait_ranges <- gym_data_cr |> dplyr::select(dbh, cr, wd) |> apply(2, range)

# Read and process DBH allometry sub-dataset
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)

ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

gym_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |>
  filter(n() >= 20) |>
  ungroup()

ang_dbh_obs <- nrow(ang_data_dbh)
ang_dbh_sp <- ang_data_dbh |> distinct(sp) |> nrow()
ang_dbh_trait_ranges <- ang_data_dbh |> dplyr::select(dbh, cr, h, wd) |> apply(2, range)

gym_dbh_obs <- nrow(gym_data_dbh)
gym_dbh_sp <- gym_data_dbh |> distinct(sp) |> nrow()
gym_dbh_trait_ranges <- gym_data_dbh |> dplyr::select(dbh, cr, h, wd) |> apply(2, range)

# Combine the data
combined_data <- tibble(
  `Dependent variables` = rep(c("H", "CR", "DBH"), each = 2),
  Division = rep(c("Angiosperm", "Gymnosperm"), times = 3),
  `Number of trees` = as.character(c(
    nrow(ang_data_h), nrow(gym_data_h),
    nrow(ang_data_cr), nrow(gym_data_cr),
    nrow(ang_data_dbh), nrow(gym_data_dbh)
  )),
  `Number of species` = as.character(c(
    ang_data_h |> distinct(sp) |> nrow(),
    gym_data_h |> distinct(sp) |> nrow(),
    ang_data_cr |> distinct(sp) |> nrow(),
    gym_data_cr |> distinct(sp) |> nrow(),
    ang_data_dbh |> distinct(sp) |> nrow(),
    gym_data_dbh |> distinct(sp) |> nrow()
  )),
  `DBH range` = c(
    range(ang_data_h$dbh) |> paste(collapse = " - "), range(gym_data_h$dbh) |> paste(collapse = " - "),
    range(ang_data_cr$dbh) |> paste(collapse = " - "), range(gym_data_cr$dbh) |> paste(collapse = " - "),
    range(ang_data_dbh$dbh) |> paste(collapse = " - "), range(gym_data_dbh$dbh) |> paste(collapse = " - ")
  ),
  `Dependent variable range` = c(
    range(ang_data_h$h) |> paste(collapse = " - "), range(gym_data_h$h) |> paste(collapse = " - "),
    range(ang_data_cr$cr) |> paste(collapse = " - "), range(gym_data_cr$cr) |> paste(collapse = " - "),
    range(ang_data_dbh$cr) |> paste(collapse = " - "), range(gym_data_dbh$cr) |> paste(collapse = " - ")
  ),
  `WD range` = c(
    ang_data_h$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_h$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_cr$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_cr$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_dbh$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_dbh$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - ")
  ),
  Relationship = rep(c("H-DBH", "CR-DBH", "DBH-CR,H"), each = 2)
)

formatted_table <- combined_data |>
  pivot_longer(
    cols = c(`Number of trees`, `Number of species`, `DBH range`, `Dependent variable range`, `WD range`),
    names_to = "Characteristics",
    values_to = "Values"
  ) |>
  pivot_wider(
    names_from = Division,
    values_from = Values
  ) |>
  dplyr::select(`Dependent variables`, Characteristics, Relationship, Angiosperm, Gymnosperm)

formatted_table
```

> formatted_table
# A tibble: 15 × 5
   `Dependent variables` Characteristics          Relationship Angiosperm   Gymnosperm  
   <chr>                 <chr>                    <chr>        <chr>        <chr>       
 1 H                     Number of trees          H-DBH        79332        6306        
 2 H                     Number of species        H-DBH        1214         76          
 3 H                     DBH range                H-DBH        1 - 560      1 - 648     
 4 H                     Dependent variable range H-DBH        1.3 - 94     1.3 - 115.8 
 5 H                     WD range                 H-DBH        0.131 - 1.17 0.32 - 0.689
 6 CR                    Number of trees          CR-DBH       45350        4632        
 7 CR                    Number of species        CR-DBH       762          59          
 8 CR                    DBH range                CR-DBH       1 - 385      1 - 266     
 9 CR                    Dependent variable range CR-DBH       0.05 - 24.25 0.1 - 8.75  
10 CR                    WD range                 CR-DBH       0.131 - 1.15 0.32 - 0.689
11 DBH                   Number of trees          DBH-CR,H     41713        3685        
12 DBH                   Number of species        DBH-CR,H     744          56          
13 DBH                   DBH range                DBH-CR,H     1 - 385      1 - 266     
14 DBH                   Dependent variable range DBH-CR,H     0.1 - 24.25  0.2 - 10.65 
15 DBH                   WD range                 DBH-CR,H     0.131 - 1.15 0.32 - 0.651

```{r}
library(dplyr)
library(tidyr)
library(targets)

options(width = 150)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
tallo_reduced_lr_df_ang_dbh1 <- tar_read(tallo_reduced_lr_df_ang_dbh1)
tallo_reduced_lr_df_ang_dbh2 <- tar_read(tallo_reduced_lr_df_ang_dbh2)
tallo_reduced_lr_df_ang_dbh3 <- tar_read(tallo_reduced_lr_df_ang_dbh3)

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")
ang_data_h
summary(ang_data_h)

gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm")
summary(gym_data_h)

# Process CR allometry dataset
ang_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
summary(ang_data_cr)

gym_data_cr <- tallo_reduced_lr_df_ang_cr |> filter(division == "Gymnosperm")
summary(gym_data_cr)

# Process DBH allometry datasets with specific filters
ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(ang_data_dbh)

gym_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(gym_data_dbh)

ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(ang_data_dbh1)

gym_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(gym_data_dbh1)

ang_data_dbh2 <- tallo_reduced_lr_df_ang_dbh2 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(ang_data_dbh2)

gym_data_dbh2 <- tallo_reduced_lr_df_ang_dbh2 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(gym_data_dbh2)

ang_data_dbh3 <- tallo_reduced_lr_df_ang_dbh3 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary(ang_data_dbh3)

gym_data_dbh3 <- tallo_reduced_lr_df_ang_dbh3 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

summary()

# Recreate `combined_data` with corrected Predictor variable range logic
combined_data <- tibble(
  `Dependent variables` = rep(c("H", "CR", "DBH", "DBH", "DBH", "DBH"), each = 2),
  `Predictor variable` = c(
    "DBH", "DBH", 
    "DBH", "DBH",
    "CR, H", "CR, H", 
    "CR * H", "CR * H", 
    "CR", "CR", 
    "H", "H"
  ),
  Division = rep(c("Angiosperm", "Gymnosperm"), times = 6),
  `Number of trees` = as.character(c(
    nrow(ang_data_h), nrow(gym_data_h),
    nrow(ang_data_cr), nrow(gym_data_cr),
    nrow(ang_data_dbh), nrow(gym_data_dbh),
    nrow(ang_data_dbh1), nrow(gym_data_dbh1),
    nrow(ang_data_dbh2), nrow(gym_data_dbh2),
    nrow(ang_data_dbh3), nrow(gym_data_dbh3)
  )),
  `Number of species` = as.character(c(
    ang_data_h |> distinct(sp) |> nrow(),
    gym_data_h |> distinct(sp) |> nrow(),
    ang_data_cr |> distinct(sp) |> nrow(),
    gym_data_cr |> distinct(sp) |> nrow(),
    ang_data_dbh |> distinct(sp) |> nrow(),
    gym_data_dbh |> distinct(sp) |> nrow(),
    ang_data_dbh1 |> distinct(sp) |> nrow(),
    gym_data_dbh1 |> distinct(sp) |> nrow(),
    ang_data_dbh2 |> distinct(sp) |> nrow(),
    gym_data_dbh2 |> distinct(sp) |> nrow(),
    ang_data_dbh3 |> distinct(sp) |> nrow(),
    gym_data_dbh3 |> distinct(sp) |> nrow()
  )),
  `Predictor variable range` = c(
    range(ang_data_h$dbh) |> paste(collapse = " - "), range(gym_data_h$dbh) |> paste(collapse = " - "),
    range(ang_data_cr$dbh) |> paste(collapse = " - "), range(gym_data_cr$dbh) |> paste(collapse = " - "),
    paste("CR:", range(ang_data_dbh$cr) |> paste(collapse = " - "), 
          "H:", range(ang_data_dbh$h) |> paste(collapse = " - "), sep = " "),
    paste("CR:", range(gym_data_dbh$cr) |> paste(collapse = " - "), 
          "H:", range(gym_data_dbh$h) |> paste(collapse = " - "), sep = " "),
    paste("CR:", range(ang_data_dbh1$cr) |> paste(collapse = " - "),
          "H:", range(ang_data_dbh1$h) |> paste(collapse = " - "), sep = " "),
    paste("CR:", range(gym_data_dbh1$cr) |> paste(collapse = " - "),
          "H:", range(gym_data_dbh1$h) |> paste(collapse = " - "), sep = " "),
    paste("CR:", range(ang_data_dbh2$cr) |> paste(collapse = " - ")),
    paste("CR:", range(gym_data_dbh2$cr) |> paste(collapse = " - ")),
    paste("H:", range(ang_data_dbh3$h) |> paste(collapse = " - ")),
    paste("H:", range(gym_data_dbh3$h) |> paste(collapse = " - "))
  ),
  `Dependent variable range` = c(
    range(ang_data_h$h) |> paste(collapse = " - "), range(gym_data_h$h) |> paste(collapse = " - "),
    range(ang_data_cr$cr) |> paste(collapse = " - "), range(gym_data_cr$cr) |> paste(collapse = " - "),
    range(ang_data_dbh$cr) |> paste(collapse = " - "), range(gym_data_dbh$cr) |> paste(collapse = " - "),
    range(ang_data_dbh1$cr) |> paste(collapse = " - "), range(gym_data_dbh1$cr) |> paste(collapse = " - "),
    range(ang_data_dbh2$cr) |> paste(collapse = " - "), range(gym_data_dbh2$cr) |> paste(collapse = " - "),
    range(ang_data_dbh3$h) |> paste(collapse = " - "), range(gym_data_dbh3$h) |> paste(collapse = " - ")
  ),
  `WD range` = c(
    ang_data_h$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_h$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_cr$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_cr$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_dbh$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_dbh$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_dbh1$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_dbh1$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_dbh2$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_dbh2$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    ang_data_dbh3$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - "),
    gym_data_dbh3$wd |> range(na.rm = TRUE) |> round(3) |> paste(collapse = " - ")
  )
)
combined_data

formatted_table <- combined_data |>
  pivot_longer(
    cols = c(`Number of trees`, `Number of species`, `Predictor variable range`, `Dependent variable range`, `WD range`),
    names_to = "Characteristics",
    values_to = "Values"
  ) |>
  pivot_wider(
    names_from = Division,
    values_from = Values
  ) |>
  dplyr::select(`Dependent variables`, `Predictor variable`, Characteristics, Angiosperm, Gymnosperm)

print(formatted_table, n = Inf)
```

# Checking species-specific posterior

```{r}
options(width = 200)
sp_posterior_df_rounded <- tar_read(sp_posterior_df_rounded)
head(sp_posterior_df_rounded)
summary(sp_posterior_df_rounded)

# h_posterior_df <- sp_posterior_df_rounded |>
#   filter(Dependent_variable == "Tree Height") |>
#   dplyr::select(sp, Division, Dependent_variable, a, b, k)

h_posterior_df <- sp_posterior_df_rounded |>
  dplyr::select(sp, Division, Dependent_variable, a, b, k) |>
  filter(
    Dependent_variable == "Tree Height",
    !is.na(a),
    !is.na(b),
    !is.na(k)
  )

head(h_posterior_df)
summary(h_posterior_df)
length(unique(h_posterior_df$sp))

h_ang <- h_posterior_df |>
  filter(Division == "Angiosperm")
head(h_ang)

summary(h_ang)
length(unique(h_ang$sp))

h_gym <- h_posterior_df |>
  filter(Division == "Gymnosperm")
h_gym

```

```{r}
options(width = 200)
sp_posterior_df <- tar_read(sp_posterior_df)
head(sp_posterior_df)
summary(sp_posterior_df)

# h_posterior_df <- sp_posterior_df_rounded |>
#   filter(Dependent_variable == "Tree Height") |>
#   dplyr::select(sp, Division, Dependent_variable, a, b, k)

h_posterior_df <- sp_posterior_df |>
  dplyr::select(sp, Division, Dependent_variable, a, b, k) |>
  filter(
    Dependent_variable == "Tree Height",
    !is.na(a),
    !is.na(b),
    !is.na(k)
  )

head(h_posterior_df)
summary(h_posterior_df)
length(unique(h_posterior_df$sp))

```

# Check H, CR data

```{r}
options(width = 150)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_lr <- tallo_reduced_lr_df_ang_h |>
   filter(division == "Angiosperm")
ang_lr

gym_lr <- tallo_reduced_lr_df_ang_h |>
   filter(division == "Gymnosperm")
gym_lr

tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_nlr <- tallo_reduced_lr_df_ang_h |>
   filter(division == "Angiosperm")
ang_nlr

gym_nlr <- tallo_reduced_lr_df_ang_h |>
   filter(division == "Gymnosperm")

```

## CHECKING DATA

```{r}
# Angiosperm datasets
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_lr_nou_summary_pl_ang_h <- tar_read(fit_lr_nou_summary_pl_ang_h)
fit_lr_nou_summary_pl_ang_dbh <- tar_read(fit_lr_nou_summary_pl_ang_dbh)
fit_lr_nou_summary_pl_ang_dbh1 <- tar_read(fit_lr_nou_summary_pl_ang_dbh1)
fit_lr_nou_summary_pl_ang_dbh2 <- tar_read(fit_lr_nou_summary_pl_ang_dbh2)
fit_lr_nou_summary_pl_ang_dbh3 <- tar_read(fit_lr_nou_summary_pl_ang_dbh3)
tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
tallo_reduced_lr_df_ang_dbh1 <- tar_read(tallo_reduced_lr_df_ang_dbh1)
tallo_reduced_lr_df_ang_dbh2 <- tar_read(tallo_reduced_lr_df_ang_dbh2)
tallo_reduced_lr_df_ang_dbh3 <- tar_read(tallo_reduced_lr_df_ang_dbh3)
stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
stan_data_lr_ang_h <- tar_read(stan_data_lr_ang_h)
stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
stan_data_lr_ang_dbh1 <- tar_read(stan_data_lr_ang_dbh1)
stan_data_lr_ang_dbh2 <- tar_read(stan_data_lr_ang_dbh2)
stan_data_lr_ang_dbh3 <- tar_read(stan_data_lr_ang_dbh3)

# Gymnosperm datasets
fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_lr_nou_summary_pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
fit_lr_nou_summary_pl_gym_dbh <- tar_read(fit_lr_nou_summary_pl_gym_dbh)
fit_lr_nou_summary_pl_gym_dbh1 <- tar_read(fit_lr_nou_summary_pl_gym_dbh1)
fit_lr_nou_summary_pl_gym_dbh2 <- tar_read(fit_lr_nou_summary_pl_gym_dbh2)
fit_lr_nou_summary_pl_gym_dbh3 <- tar_read(fit_lr_nou_summary_pl_gym_dbh3)
tallo_reduced_nlr_df_gym_h <- tar_read(tallo_reduced_nlr_df_gym_h)
tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
tallo_reduced_lr_df_gym_dbh <- tar_read(tallo_reduced_lr_df_gym_dbh)
tallo_reduced_lr_df_gym_dbh1 <- tar_read(tallo_reduced_lr_df_gym_dbh1)
tallo_reduced_lr_df_gym_dbh2 <- tar_read(tallo_reduced_lr_df_gym_dbh2)
tallo_reduced_lr_df_gym_dbh3 <- tar_read(tallo_reduced_lr_df_gym_dbh3)
stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_gym_h <- tar_read(stan_data_lr_gym_h)
stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
stan_data_lr_gym_dbh1 <- tar_read(stan_data_lr_gym_dbh1)
stan_data_lr_gym_dbh2 <- tar_read(stan_data_lr_gym_dbh2)
stan_data_lr_gym_dbh3 <- tar_read(stan_data_lr_gym_dbh3)

tallo_reduced_lr_df_ang_dbh
tallo_reduced_lr_df_gym_dbh

ang_data_dbh <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

ang_data_dbh
length(unique(ang_data_dbh$sp))
str(stan_data_lr_ang_dbh)

gym_data_dbh <- tallo_reduced_lr_df_gym_dbh |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()
gym_data_dbh
length(unique(gym_data_dbh$sp))
str(stan_data_lr_gym_dbh)

tallo_reduced_lr_df_ang_dbh1
tallo_reduced_lr_df_gym_dbh1

ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()
ang_data_dbh1
length(unique(ang_data_dbh1$sp))
str(stan_data_lr_ang_dbh1)

gym_data_dbh1 <- tallo_reduced_lr_df_gym_dbh1 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

gym_data_dbh1
length(unique(gym_data_dbh1$sp))
str(stan_data_lr_gym_dbh1)

ang_data_dbh2 <- tallo_reduced_lr_df_ang_dbh2 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

ang_data_dbh2
length(unique(ang_data_dbh2$sp))
str(stan_data_lr_ang_dbh2)

gym_data_dbh2 <- tallo_reduced_lr_df_gym_dbh2 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

gym_data_dbh2
length(unique(gym_data_dbh2$sp))
str(stan_data_lr_gym_dbh2)

ang_data_dbh3 <- tallo_reduced_lr_df_ang_dbh3 |>
  filter(division == "Angiosperm", !is.na(dbh), !is.na(h), !is.na(cr)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()
ang_data_dbh3
length(unique(ang_data_dbh3$sp))

str(stan_data_lr_ang_dbh3)

gym_data_dbh3 <- tallo_reduced_lr_df_gym_dbh3 |>
  filter(division == "Gymnosperm", !is.na(dbh), !is.na(h), !is.na(cr)) |>
  group_by(sp) |> filter(n() >= 20) |>
  ungroup()

gym_data_dbh3
length(unique(gym_data_dbh3$sp))

str(stan_data_lr_gym_dbh3)

```

```{r}
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm") |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)

ang_h_sp

gym_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm") |>
  group_by(sp) |>
  summarise(count = n()) |>
  filter(count >= 20)
gym_data_h

```

### 2025-03-26 docs/check_sp_estimates.qmd
Initial commit for tree-allo

---
title: "Checking species-level estimates"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# H-DBH allometry

## Read in data

```{r}
tallo_reduced_nlr_df_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_h

summary(tallo_reduced_nlr_df_h)

species_data <- tallo_reduced_nlr_df_h |>
  select(tree_id, sp, dbh, h)

sub_datasets <- species_data |> 
  group_split(sp)

names(sub_datasets) <- unique(species_data$sp)

subset_species <- names(sub_datasets)[1:100]  # Select the first 100 species
subset_data <- sub_datasets[names(sub_datasets) %in% subset_species]
```

## Extracting sp-specific parameter

```{r}
sp_posterior_h_df <- tar_read(sp_posterior_h_df)
head(sp_posterior_h_df)
sp_posterior_h_df <- sp_posterior_h_df |> 
  rename(sp = Species)

```

# Function
```{r}
generate_plot <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    my_theme()

  return(p)
}

generate_plot_log <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    scale_x_log10() +
    scale_y_log10() +
    my_theme()

  return(p)
}
```

# Plotting

```{r, fig.width=14, fig.height=20}
# Split species into 13 groups
species_groups <- split(sub_datasets, cut(seq_along(sub_datasets), breaks = 15, labels = FALSE))

# Loop through each group and generate plots
for (i in seq_along(species_groups)) {
  # Get the current group of species
  group_data <- species_groups[[i]]

  # Generate the plot
  p <- generate_plot(group_data, sp_posterior_h_df)

  # Save the plot
  ggsave(
    filename = paste0("h_ang_sp_group_", i, ".png"),
    plot = p,
    width = 14,
    height = 45,
    units = "in",
    dpi = 300
  )

  # Generate the log-log plot
  p_log <- generate_plot_log(group_data, sp_posterior_h_df)

  # Save the log-log plot
  ggsave(
    filename = paste0("h_ang_sp_group_", i, "_log.png"),
    plot = p_log,
    width = 14,
    height = 40,
    units = "in",
    dpi = 300
  )
}
```

# CR-DBH

## ANG
```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
tallo_reduced_lr_df_ang_cr <- tallo_reduced_lr_df_ang_cr |>
   filter(division == "Angiosperm")
tallo_reduced_lr_df_ang_cr
length(unique(tallo_reduced_lr_df_ang_cr$sp))

ang_cr_pl <- tallo_reduced_lr_df_ang_cr |> 
  group_split(sp)

names(ang_cr_pl) <- unique(tallo_reduced_lr_df_ang_cr$sp)
ang_cr_pl_sp <- names(ang_cr_pl)

sp_posterior_cr_df <- tar_read(sp_posterior_cr_df)

sp_posterior_cr_df <- sp_posterior_cr_df |>
  rename(sp = Species)
head(sp_posterior_cr_df)

sp_posterior_cr_ang <- sp_posterior_cr_df |>
  filter(Division =="Angiosperm")
head(sp_posterior_cr_ang)
summary(sp_posterior_cr_ang)
```

```{r}
### POWER-LAW FUNCTION
generate_plot_cr <- function(sub_datasets, sp_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- sp_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))

    dbh <- species_data$dbh

    # Compute the fitted CR values
    fitted_log_y <- log_a + b*log(dbh)
    fitted_cr <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      CR = fitted_cr,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, CR = cr, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  # Plot the combined data
  p <- ggplot(combined_data, aes(x = DBH, y = CR, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Crown Radius (m)"
    ) +
    my_theme()

  return(p)
}

generate_plot_cr_log <- function(sub_datasets, sp_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- sp_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))

    dbh <- species_data$dbh

    # Compute the fitted CR values
    fitted_log_y <- log_a + b*log(dbh)
    fitted_cr <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      CR = fitted_cr,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, CR = cr, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  # Plot the combined data in log-log scale
  p <- ggplot(combined_data, aes(x = DBH, y = CR, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Crown Radius (m)"
    ) +
    scale_x_log10() +
    scale_y_log10() +
    my_theme()

  return(p)
}
```

```{r}
subset_species <- ang_cr_pl_sp[1:100]  # Select the first 100 species
subset_data <- ang_cr_pl[names(ang_cr_pl) %in% subset_species]

# Generate the plots
p <- generate_plot_cr(subset_data, sp_posterior_cr_ang)

# Save the plot
ggsave(
  filename = "cr_ang_pl.png",
  plot = p,
  width = 14,
  height = 45,
  units = "in",
  dpi = 300
)

# Generate and save log-log plot
p_log <- generate_plot_cr_log(subset_data, sp_posterior_cr_ang)

ggsave(
  filename = "cr_ang_pl_log.png",
  plot = p_log,
  width = 14,
  height = 45,
  units = "in",
  dpi = 300
)

```

## GYM

```{r}
tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)
tallo_reduced_nlr_df_gym_cr
tallo_reduced_nlr_df_gym_cr <- tallo_reduced_nlr_df_gym_cr |>
   filter(division == "Gymnosperm")

tallo_reduced_nlr_df_gym_cr
length(unique(tallo_reduced_nlr_df_gym_cr$sp))

gym_cr_gmm <- tallo_reduced_nlr_df_gym_cr |> 
  group_split(sp)

names(gym_cr_gmm) <- unique(tallo_reduced_nlr_df_gym_cr$sp)
gym_cr_gmm_sp <- names(gym_cr_gmm)

sp_posterior_cr_df <- tar_read(sp_posterior_cr_df)

sp_posterior_cr_df <- sp_posterior_cr_df |>
  rename(sp = Species)
head(sp_posterior_cr_df)

sp_posterior_cr_gym <- sp_posterior_cr_df |>
  filter(Division =="Gymnosperm")
head(sp_posterior_cr_gym)
summary(sp_posterior_cr_gym)

```

```{r}
### gMM function
generate_plot_cr_gmm <- function(sub_datasets, sp_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- sp_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))    
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted CR values
    fitted_log_y <- log_a + b*log(dbh) + log(k + dbh^b)
    fitted_cr <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      CR = fitted_cr,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, CR = cr, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  # Plot the combined data
  p <- ggplot(combined_data, aes(x = DBH, y = CR, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Crown Radius (m)"
    ) +
    my_theme()

  return(p)
}

generate_plot_cr_log_gmm <- function(sub_datasets, sp_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- sp_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))    
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted CR values
    fitted_log_y <- log_a + b*log(dbh) + log(k + dbh^b)
    fitted_cr <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      CR = fitted_cr,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, CR = cr, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }
  # Plot the combined data in log-log scale
  p <- ggplot(combined_data, aes(x = DBH, y = CR, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Crown Radius (m)"
    ) +
    scale_x_log10() +
    scale_y_log10() +
    my_theme()

  return(p)
}
```

```{r}
subset_species <- gym_cr_gmm_sp[1:59]  # Select the first 100 species
subset_data <- gym_cr_gmm[names(gym_cr_gmm) %in% subset_species]

# Generate the plots
p <- generate_plot_cr_gmm(subset_data, sp_posterior_cr_gym)

# Save the plot
ggsave(
  filename = "cr_gym_gmm.png",
  plot = p,
  width = 14,
  height = 45,
  units = "in",
  dpi = 300
)

# Generate and save log-log plot
p_log <- generate_plot_cr_log_gmm(subset_data, sp_posterior_cr_ang)

ggsave(
  filename = "cr_gym_gmm_log.png",
  plot = p_log,
  width = 14,
  height = 45,
  units = "in",
  dpi = 300
)

```

### 2025-03-26 docs/com.qmd
Initial commit for tree-allo

---
title: "Model comparison figure"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
# Required Libraries
library(targets)
library(tidyverse)
library(scales)
library(MASS)
library(viridis)
library(patchwork)
library(ggplot2)
source("R/figs.R")
```

# Read in data

```{r}
# Load Data for H-DBH
stan_data_lr_h <- tar_read(stan_data_lr_h)
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
gmm_summary_h <- tar_read(fit_nlr_summary_gmm1_h)
wb_summary_h <- tar_read(fit_nlr_summary_weibull1_h)
pl_summary_h <- tar_read(fit_lr2_summary_power_law1_h)

# Load Data for CR-DBH
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
gmm_summary_cr <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary_cr <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary_cr <- tar_read(fit_lr2_summary_power_law1_cr)
```

# Extract scaling parameters
```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
  filter(!is.na(dbh), !is.na(h), !is.na(cr), !is.na(wd_s)) |>
  mutate(log_dbh = log(dbh),
         log_h = log(h),
         log_cr = log(cr)) |>
  mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
         log_h_s = scale(log_h) |> as.numeric(),
         log_cr_s = scale(log_cr) |> as.numeric()) |>
  mutate(dbh_s = scale(dbh) |> as.numeric())

# Extract scaling parameters for unscaling
log_dbh_sd <- sd(data2$log_dbh)
log_dbh_mean <- mean(data2$log_dbh)
log_h_sd <- sd(data2$log_h)
log_h_mean <- mean(data2$log_h)
log_cr_sd <- sd(data2$log_cr)
log_cr_mean <- mean(data2$log_cr)
```

# Function to Compute Fitted Values

```{r}
compute_fitted_values <- function(model_summary, x_seq, log_y_mean, log_y_sd, log_x_sd, log_x_mean, model_type, log_x_seq = NULL) {
  gamma <- model_summary |> filter(str_detect(variable, "gamma"))

  if (model_type == "pl") {
    # Power-Law model: use log_x_seq for predictions
    log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[2,1]") |> pull(q50)
    log_y_pred_scaled <- log_a + b * log_x_seq

    # Unscale the predicted log-transformed y
    log_y_pred <- log_y_pred_scaled * log_y_sd + log_y_mean

    # Unscale the x variable and back-transform to the original scale
    log_x_unscaled <- (log_x_seq * log_x_sd) + log_x_mean
    fitted_values <- tibble(y = exp(log_y_pred), x = exp(log_x_unscaled))

  } else if (model_type == "gmm") {
    # Generalized Michaelis-Menten model
    log_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
    k <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

    log_a <- exp(log_log_a)
    b <- exp(b)
    k <- exp(k)

    log_y_pred <- log_a + b * log(x_seq) - log(k + x_seq^b)
    fitted_values <- tibble(y = exp(log_y_pred), x = x_seq)

  } else if (model_type == "weibull") {
    # Weibull model
    log_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
    k <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

    log_a <- exp(log_log_a)
    b <- exp(b)
    k <- exp(k)

    log_y_pred <- log_a + log(1 - exp(-b * (x_seq ^ k)))
    fitted_values <- tibble(y = exp(log_y_pred), x = x_seq)

  } else {
    stop("Invalid model type specified.")
  }

  return(fitted_values)
}
```

# Extract fitted values
```{r}
# Define sequences for Power-Law x variables (scaled log x)
scaled_x_range_h <- stan_data_lr_h$log_x[, 2] |> range()
log_x_seq_h <- seq(scaled_x_range_h[1], scaled_x_range_h[2], length.out = 100)

scaled_x_range_cr <- stan_data_lr_cr$log_x[, 2] |> range()
log_x_seq_cr <- seq(scaled_x_range_cr[1], scaled_x_range_cr[2], length.out = 100)

# Define sequences for gMM and Weibull x variables (original x)
x_nlr_h <- stan_data_nlr_h$x
x_nlr_range_h <- range(x_nlr_h)
x_nlr_seq_h <- seq(x_nlr_range_h[1], x_nlr_range_h[2], length.out = 100)

x_nlr_cr <- stan_data_nlr_cr$x
x_nlr_range_cr <- range(x_nlr_cr)
x_nlr_seq_cr <- seq(x_nlr_range_cr[1], x_nlr_range_cr[2], length.out = 100)

# Compute fitted values for Power-Law model (unscaled)
df_pl_h <- compute_fitted_values(pl_summary_h, NULL, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "pl", log_x_seq = log_x_seq_h)
df_pl_cr <- compute_fitted_values(pl_summary_cr, NULL, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "pl", log_x_seq = log_x_seq_cr)

# Compute fitted values for gMM model (original x)
df_gmm_h <- compute_fitted_values(gmm_summary_h, x_nlr_seq_h, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "gmm")
df_gmm_cr <- compute_fitted_values(gmm_summary_cr, x_nlr_seq_cr, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "gmm")

# Compute fitted values for Weibull model (original x)
df_wb_h <- compute_fitted_values(wb_summary_h, x_nlr_seq_h, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "weibull")
df_wb_cr <- compute_fitted_values(wb_summary_cr, x_nlr_seq_cr, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "weibull")
```

# Plotting

## Plotting function
```{r}
# Plotting function
plot_model_comparison <- function(data, df_pl, df_gmm, df_wb, x_var, y_var, x_label, y_label, log_scale = TRUE) {
  # Compute density for the scatter plot
  if (log_scale) {
    # Log-scale density calculation
    dens <- with(data, kde2d(log10(data[[x_var]]), log10(data[[y_var]]), n = 300))
    ix <- findInterval(log10(data[[x_var]]), dens$x)
    iy <- findInterval(log10(data[[y_var]]), dens$y)
  } else {
    # Linear-scale density calculation
    dens <- with(data, kde2d(data[[x_var]], data[[y_var]], n = 300))
    ix <- findInterval(data[[x_var]], dens$x)
    iy <- findInterval(data[[y_var]], dens$y)
  }

  data$density <- dens$z[cbind(ix, iy)]

  # Base plot
  p <- ggplot(data, aes_string(x = x_var, y = y_var, color = "density")) +
    geom_point(size = 0.5, alpha = 0.8) +
    scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
    labs(x = x_label, y = y_label) +
    theme_minimal()

  # Add model lines
  p <- p +
    geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1) +
    geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1) +
    geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1)

  # Handle log scale if requested
  if (log_scale) {
    p <- p + scale_x_log10() + scale_y_log10()
  }

  return(p)
}
```

## Plot
```{r}
# Plot for H-DBH using Power-Law model type
combined_plot_h <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_h,
  df_gmm = df_gmm_h,
  df_wb = df_wb_h,
  x_var = "dbh",
  y_var = "h",
  x_label = "DBH (cm)",
  y_label = "Height (m)",
  log_scale = TRUE    # or FALSE, depending on your preference
)

# Plot for CR-DBH using Power-Law model type
combined_plot_cr <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_cr,
  df_gmm = df_gmm_cr,
  df_wb = df_wb_cr,
  x_var = "dbh",
  y_var = "cr",
  x_label = "DBH (cm)",
  y_label = "Crown Radius (m)",
  log_scale = TRUE    # or FALSE
)

```

# Combining plots

```{r}
# Increase the size of axis titles
combined_plot_h <- combined_plot_h + 
  theme(
    plot.title = element_text(hjust = 0, face = "bold"),  
    axis.title.x = element_blank(),                       # Remove x-axis title from the top plot
    axis.text.x = element_blank(),                        # Remove x-axis ticks from the top plot
    axis.ticks.x = element_blank(),                       # Remove x-axis ticks from the top plot
    axis.title.y = element_text(size = 16),               # Increase y-axis title size
    plot.margin = margin(5, 5, 5, 5),                     # Adjust margins if needed
    legend.position = "none"                              # Remove the legend from the top plot
  ) + ggtitle("(a)")

combined_plot_cr <- combined_plot_cr + 
  theme(
    plot.title = element_text(hjust = 0, face = "bold"),  # Left-align the title and make it bold
    axis.title.x = element_text(size = 16),               # Increase x-axis title size
    axis.title.y = element_text(size = 16),               # Increase y-axis title size
    plot.margin = margin(5, 5, 5, 5)                      # Adjust margins if needed
  ) + ggtitle("(b)")

combined_plots <- combined_plot_h / combined_plot_cr + 
  plot_layout(ncol = 1, heights = c(1, 1))

ggsave("combined_plots_with_larger_labels.png", plot = combined_plots, width = 10, height = 12, dpi = 300)
```

```{r}
# Plot for H-DBH using Power-Law model type
combined_plot_h_non <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_h,
  df_gmm = df_gmm_h,
  df_wb = df_wb_h,
  x_var = "dbh",
  y_var = "h",
  x_label = "DBH (cm)",
  y_label = "H (m)",
  log_scale = FALSE    
)

# Plot for CR-DBH using Power-Law model type
combined_plot_cr_non <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_cr,
  df_gmm = df_gmm_cr,
  df_wb = df_wb_cr,
  x_var = "dbh",
  y_var = "cr",
  x_label = "DBH (cm)",
  y_label = "CR (m)",
  log_scale = FALSE
)

print(combined_plot_h_non)
print(combined_plot_cr_non)
```

### 2025-03-26 docs/cov_beta.qmd
Initial commit for tree-allo

---
title: "Checking the covariance between beta parameter of gmm_nou_h"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load required packages 
```{r}
library(cmdstanr)
library(posterior)
library(targets)
```

# Data preparation
```{r}
gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)
```

# Extract correlations among each parameter 
```{r}
# Convert posterior samples to an array
posterior_samples <- as_draws_array(gmm_nou_h$draws())

# Number of posterior samples
num_samples <- dim(posterior_samples)[1]
print(num_samples)

# Number of predictors (K)
K <- 3

# Extract L_Sigma_beta samples
# Adjust the regex pattern if the variable names differ
L_Sigma_beta_samples <- posterior_samples[, , grep("^L_Sigma_beta", dimnames(posterior_samples)$variable)]

# Initialize array to store covariance matrices
cov_matrices <- array(NA, dim = c(K, K, num_samples))

# Calculate covariance matrix for each sample
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrices[, , i] <- L_Sigma_beta %*% t(L_Sigma_beta)
  }

```

```{r}
# Calculate the mean covariance matrix
mean_cov_matrix <- apply(cov_matrices, c(1, 2), mean)

# Print the mean covariance matrix
print(mean_cov_matrix)

# Calculate the mean correlation matrix
mean_cor_matrix <- cov2cor(mean_cov_matrix)

# Print the mean correlation matrix
print(mean_cor_matrix)
```

# Interpretation:

- Diagonal Element (Variance):
Variance of beta1: 853.0958
Variance of beta 2: 112.9586
Variance of beta 3: 478.6792
- Off-Diagonal Elements (Covariance):
Covariance between beta1 and beta2: 306.1926
Covariance between beta1 and beta3: 636.5709
Covariance between beta2 and beta3: 230.3332
- The Mean Correlation Matrix
Correlation Coefficients:
Correlation between beta1 and beta2: 0.9863616
Correlation between beta1 and beta3: 0.9961519
Correlation between beta2 and beta3: 0.9905461

# Summary
- All correlations are very high (close to 1), indicating strong positive relationship between beta parameters.
- Betas move together strongly in the same direction (in gmm_nou_h model)

# Checking 95% CI of the correlations coefficients
```{r}
# Calculate correlation matrices for each sample
cor_matrices <- array(NA, dim = c(K, K, num_samples))
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrix <- L_Sigma_beta %*% t(L_Sigma_beta)
  cor_matrices[, , i] <- cov2cor(cov_matrix)
}

# Initialize matrices to store the lower and upper bounds of the 95% CI
lower_ci <- matrix(NA, nrow = K, ncol = K)
upper_ci <- matrix(NA, nrow = K, ncol = K)

# Calculate the 95% CI for each correlation coefficient
for (row in 1:K) {
  for (col in 1:K) {
    cor_values <- cor_matrices[row, col, ]
    lower_ci[row, col] <- quantile(cor_values, probs = 0.025)
    upper_ci[row, col] <- quantile(cor_values, probs = 0.975)
  }
}

# Print the 95% CI for each correlation coefficient
print(lower_ci)
print(upper_ci)

```

```{r}
# Assuming you have the posterior samples already extracted
posterior_samples <- as_draws_array(gmm_h$draws())

# Number of posterior samples
num_samples <- dim(posterior_samples)[1]

# Define the number of tree-level predictors (K)
K <- 2

# Extract L_Sigma_beta samples
L_Sigma_beta_samples <- posterior_samples[, , grep("^L_Sigma_beta", dimnames(posterior_samples)$variable)]

# Initialize array to store covariance matrices
cov_matrices <- array(NA, dim = c(K, K, num_samples))

# Calculate covariance matrix for each sample
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrices[,,i] <- L_Sigma_beta %*% t(L_Sigma_beta)
}

# Calculate the mean covariance matrix
mean_cov_matrix <- apply(cov_matrices, c(1, 2), mean)

# Print the mean covariance matrix
print(mean_cov_matrix)

# Calculate the correlations
mean_cor_matrix <- cov2cor(mean_cov_matrix)

# Print the mean correlation matrix
print(mean_cor_matrix)

# Calculate 95% credible intervals for the correlations
cor_samples <- array(NA, dim = c(K, K, num_samples))
for (i in 1:num_samples) {
  cor_samples[,,i] <- cov2cor(cov_matrices[,,i])
}

# Calculate the 95% credible intervals
lower_ci <- apply(cor_samples, c(1, 2), function(x) quantile(x, 0.025))
upper_ci <- apply(cor_samples, c(1, 2), function(x) quantile(x, 0.975))

# Print the 95% credible intervals
print(lower_ci)
print(upper_ci)
```

### 2025-03-26 docs/cr.qmd
Initial commit for tree-allo

---
title: "Plot the best model to predict CR from DBH: Weibull"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
```

# Read in data

```{r}
# Read in the data (wb, gMM)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data 
log_y <- stan_data_nlr_cr$log_y
x <- stan_data_nlr_cr$x
sp <- as.factor(stan_data_nlr_cr$jj)

# Read in the model summary
options(width = 120)
wb1_cr_summary <- tar_read(fit_nlr_summary_weibull1_cr)
wb1_cr_summary

# Extract the actual beta parameters
beta <- wb1_cr_summary |> filter(grepl("beta\\[", variable))
beta

log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)

# Compute fitted values for CR using the Weibull formula
fitted_log_y_cr <- sapply(1:length(log_a), function(i) {
  log_a[i] + log(1 - exp(-b[i] * (x ^ k[i])))
})

fitted_cr <- exp(fitted_log_y_cr)

# Extract community-level parameters directly
gamma <- wb1_cr_summary |> filter(grepl("gamma", variable))

# Correctly map community parameters
community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

# Apply exponential transformation to community-level parameters
community_log_a <- exp(community_log_a_hat)
community_b <- exp(community_b_hat) 
community_k <- exp(community_k_hat)

community_fitted_log_y_cr <- community_log_a + log(1 - exp(-community_b * (x^community_k)))
community_fitted_CR <- exp(community_fitted_log_y_cr)

# Create a data frame for plotting
data_cr <- data.frame(CR = exp(log_y), DBH = x, sp = sp)
data_cr$community_fitted_CR <- community_fitted_CR
head(data_cr)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_cr[[paste0("fitted_CR_sp_", i)]] <- fitted_cr[, i]
}

# Convert data to long format
data_long <- reshape2::melt(data_cr, id.vars = c("CR", "DBH", "sp", "community_fitted_CR"), measure.vars = grep("fitted_CR_sp_", names(data_cr), value = TRUE))

head(data_long)

# Function to create DBH sequence for each species based on actual range
create_dbh_sequence <- function(species, x, length_out = 100) {
  min_dbh <- min(x[sp == species])
  max_dbh <- max(x[sp == species])
  seq(min_dbh, max_dbh, length.out = length_out)
}

# Calculate fitted CR values for each species using gamma parameters
fitted_cr_list <- lapply(1:length(log_a), \(i) {
  dbh_seq <- create_dbh_sequence(i, x)
  fitted_log_cr <- log_a[i] + log(1 - exp(-b[i] * (dbh_seq ^ k[i])))
  fitted_cr <- exp(fitted_log_cr)

  data.frame(
    DBH = dbh_seq,
    value = fitted_cr,
    variable = paste0("fitted_CR_sp_", i)
  )
})

# Combine all species data frames into one
data_long <- do.call(rbind, fitted_cr_list)
head(data_long)

# Original data for points
data_cr <- data.frame(CR = exp(log_y), DBH = x, sp = sp)
head(data_cr)
```

# Plot
```{r}
# # Plotting the CR against DBH
# wb1_cr_plot <- ggplot(data_cr, aes(x = DBH, y = CR)) +
#   geom_point(alpha = 0.5, color = "grey") +
#   geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) +
#   geom_line(aes(y = community_fitted_CR), color = "red", linewidth = 1.2) +
#   labs(title = "CR = a(1 - exp(-b * DBH^k)",
#        x = "DBH (cm)",
#        y = "CR (m)") +
#   scale_x_log10(labels = scales::label_number()) +
#   scale_y_log10(labels = scales::label_number()) +
#   theme_minimal()

# # Save the plot
# # ggsave(filename = "figs/cr/DBH_CR_wb1_corrected.png", plot = wb1_cr_plot, width = 12, height = 8, dpi = 300, bg = "white")

# wb1_cr_plot
```

```{r}
# Plotting the CR against DBH
wb1_cr_plot <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  # geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#72b6e3", alpha = 0.3) +
  geom_line(aes(y = community_fitted_CR), color = "#0f92e9", linewidth = 1.2) +
  labs(
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

wb1_cr_plot
```

### 2025-03-26 docs/cr_3functions.qmd
Initial commit for tree-allo

---
title: "CR models - plotting 3 community-lines (pl, wd, gMM in a single panel)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)

# Define and run the targets pipeline
# tar_make()
```

# Unscale power-law's parameters

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
scaling_params <- data2 |>
    summarise(mean_log_dbh = mean(log_dbh), sd_log_dbh = sd(log_dbh),
              mean_log_cr = mean(log_cr), sd_log_cr = sd(log_cr))

scaling_params
```

# Plotting 3 community-lines (pl, wd, gMM in a single panel)

```{r}
# Read in the data (pl)
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
str(stan_data_lr_cr)

# Extract relevant data from stan_data_lr_cr (log_y, x, jj)
log_y_lr <- stan_data_lr_cr$log_y
log_x_lr <- stan_data_lr_cr$log_x[, 2]
sp_lr <- as.factor(stan_data_lr_cr$jj)

unscaled_dbh_pl <- exp((log_x_lr * scaling_params$sd_log_dbh) + scaling_params$mean_log_dbh)
unscaled_cr_pl <- exp((log_y_lr * scaling_params$sd_log_cr) + scaling_params$mean_log_cr)

# Read in the data (gMM, wb)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data from stan_data_nlr_cr (log_y, x, jj)
log_y_nlr <- stan_data_nlr_cr$log_y
x_nlr <- stan_data_nlr_cr$x
sp_nlr <- as.factor(stan_data_nlr_cr$jj)

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary <- tar_read(fit_lr2_summary_power_law1_cr)

gmm_summary
wb_summary
pl_summary

# Extract beta parameters for power-law model
beta_pl <- pl_summary |>
  filter(grepl("beta\\[", variable))

beta_pl

# Extract log_a (beta[1, ]) and b (beta[2, ])
log_a_pl <- beta_pl |>
  filter(str_detect(variable, "beta\\[1,\\d+\\]")) |>
  pull(mean)

b_pl <- beta_pl |>
  filter(str_detect(variable, "beta\\[2,\\d+\\]")) |>
  pull(mean)

# Print extracted values to verify
head(log_a_pl)
head(b_pl)

# Extract beta parameters for gMM model
beta_gmm <- gmm_summary |>
  filter(grepl("beta\\[", variable))

beta_gmm

# Extract log_a, b, k from beta_gmm
log_a_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,1\\]")) |>
  pull(mean)

b_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,2\\]")) |>
  pull(mean)

k_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,3\\]")) |>
  pull(mean)

head(log_a_gmm)
head(b_gmm)
head(k_gmm)

# Extract beta parameters for Weibull model
beta_wb <- wb_summary |>
  filter(grepl("beta\\[", variable))

beta_wb

# Extract log_a, b, k from beta_wb
log_a_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,1\\]")) |>
  pull(mean)

b_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,2\\]")) |>
  pull(mean)

k_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,3\\]")) |>
  pull(mean)

head(log_a_wb)
head(b_wb)
head(k_wb)

```

# Extract community-level parameters of each function
(I will change by median later)
```{r}
##pl
pl_summary
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_hat_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(mean)

community_log_a_pl <- exp(community_log_a_hat_pl)
community_b_pl <- exp(community_b_hat_pl)

community_log_a_pl
community_b_pl

## gMM
gmm_summary
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
gamma_gmm
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm) 
community_k_gmm <- exp(community_k_hat_gmm)

community_log_a_gmm
community_b_gmm
community_k_gmm

# Weibull
wb_summary
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
gamma_wb
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb) 
community_k_wb <- exp(community_k_hat_wb)

community_log_a_wb
community_b_wb
community_b_wb

# Computing Fitted Values for Community-level Parameters
#pl
community_fitted_log_y_cr_pl <- community_log_a_pl + community_b_pl * log_x_lr
community_fitted_cr_pl <- exp(community_fitted_log_y_cr_pl)

head(community_fitted_cr_pl)

#gMM

community_fitted_log_y_cr_gmm <- community_log_a_gmm + community_b_gmm * log(x_nlr) - log(community_k_gmm + x_nlr^community_b_gmm)
community_fitted_cr_gmm <- exp(community_fitted_log_y_cr_gmm)

head(community_fitted_cr_gmm)

# wb

community_fitted_log_y_cr_wb <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr ^ community_k_wb)))
community_fitted_cr_wb <- exp(community_fitted_log_y_cr_wb)

head(community_fitted_cr_wb)

# Create a data frame for plotting
data_cr_pl_unscaled <- data.frame(CR = unscaled_cr_pl, DBH = unscaled_dbh_pl, sp = as.factor(stan_data_lr_cr$jj), community_fitted_CR_pl = community_fitted_cr_pl)

head(data_cr_pl_unscaled)

data_cr_gb <- data.frame(CR = exp(log_y_nlr), DBH = x_nlr, sp = sp_nlr)
head(data_cr_gb)

data_cr_gb$community_fitted_CR_gmm <- community_fitted_cr_gmm
data_cr_gb$community_fitted_CR_wb <- community_fitted_cr_wb

head(data_cr_pl_unscaled)
head(data_cr_gb)

# Combine the data frames
data_combined <- merge(data_cr_pl_unscaled, data_cr_gb, by = "sp", all = TRUE)
head(data_combined)

# Separate the community-fitted values for plotting
data_community_pl <- data_combined |> select(DBH.x, community_fitted_CR_pl) |> distinct()
data_community_gmm <- data_combined |> select(DBH.y, community_fitted_CR_gmm) |> distinct()
data_community_wb <- data_combined |> select(DBH.y, community_fitted_CR_wb) |> distinct()

head(data_community_pl)
head(data_community_gmm)
head(data_community_wb)
```

# Plot the results
```{r}
combined_plot <- ggplot(data_combined) +
  geom_point(aes(x = DBH.x, y = CR.x), alpha = 0.5, color = "grey") +
  geom_point(aes(x = DBH.y, y = CR.y), alpha = 0.5, color = "grey") +
  geom_line(data = data_community_pl, aes(x = DBH.x, y = community_fitted_CR_pl, color = "Power-law"), size = 1) +
  geom_line(data = data_community_gmm, aes(x = DBH.y, y = community_fitted_CR_gmm, color = "gMM"), size = 1) +
  geom_line(data = data_community_wb, aes(x = DBH.y, y = community_fitted_CR_wb, color = "Weibull"), size = 1) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "blue", "gMM" = "red", "Weibull" = "green"))

# Save the plot
ggsave(filename = "figs/cr/DBH_CR_community_combined.png", plot = combined_plot, width = 12, height = 8, dpi = 300, bg = "white")

print(combined_plot)
```

### 2025-03-26 docs/cr_conv.qmd
Initial commit for tree-allo

---
title: "CR-DBH allometry convergence's checking"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
```

# Power-law

## Without wood density

```{r}

fit_lr2_summary_power_law1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
fit_lr2_summary_power_law1_nou_cr
beta_pl_nou_cr <- fit_lr2_summary_power_law1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_pl_nou_cr)
```

- No problem

# With wood density
```{r}
fit_lr2_summary_power_law1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
fit_lr2_summary_power_law1_cr
beta_pl_cr <- fit_lr2_summary_power_law1_cr |> filter(grepl("beta\\[", variable))
beta_pl_cr
summary(beta_pl_cr)
```

- No problem

# gMM

## Without wood density

```{r}
fit_nlr_summary_gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
fit_nlr_summary_gmm1_nou_cr
beta_gmm_nou_cr <- fit_nlr_summary_gmm1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_gmm_nou_cr)

low_ess_gmm_nou_cr_beta <- beta_gmm_nou_cr |>
  filter(ess_bulk < 300)

low_ess_gmm_nou_cr_beta
summary(low_ess_gmm_nou_cr_beta)
```

- 5 species has convergent issue but ESS >= 250 so I think it OK.

## With wood density
```{r}
fit_nlr_summary_gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
fit_nlr_summary_gmm1_cr
beta_gmm_cr <- fit_nlr_summary_gmm1_cr |> filter(grepl("beta\\[", variable))
summary(beta_gmm_cr)

low_ess_gmm_cr_beta <- beta_gmm_cr |>
  filter(ess_bulk < 300)

print(low_ess_gmm_cr_beta, n = Inf)

low_ess_gmm_cr_beta <- low_ess_gmm_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable),
         dimension = sub("beta\\[[0-9]+,([0-9]+)\\]", "\\1", variable))

species_counts <- low_ess_gmm_cr_beta |>
  group_by(sp_id) |>
  summarise(dimension_count = n_distinct(dimension))

species_counts

num_species <- low_ess_gmm_cr_beta |>
  distinct(sp_id) |>
  nrow()
```

- 22 species has convergent issue but ESS >= 250

# Weibull

## Without wood density

```{r}
fit_nlr_summary_weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
fit_nlr_summary_weibull1_nou_cr
beta_weibull_nou_cr <- fit_nlr_summary_weibull1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_weibull_nou_cr)

low_ess_weibull_nou_cr_beta <- beta_weibull_nou_cr |>
  filter(ess_bulk < 300)

low_ess_weibull_nou_cr_beta
summary(low_ess_weibull_nou_cr_beta)

low_ess_weibull_nou_cr_beta <- low_ess_weibull_nou_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable))

num_species_low_ess <- low_ess_weibull_nou_cr_beta |>
  distinct(sp_id) |>
  nrow()

print(num_species_low_ess)
```

# With wood density

```{r}
fit_nlr_summary_weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
fit_nlr_summary_weibull1_cr
beta_weibull_cr <- fit_nlr_summary_weibull1_cr |> filter(grepl("beta\\[", variable))
summary(beta_weibull_cr)

low_ess_weibull_cr_beta <- beta_weibull_cr |>
  filter(ess_bulk < 300)

low_ess_weibull_cr_beta

summary(low_ess_weibull_cr_beta)
print(low_ess_weibull_cr_beta, n = Inf)

low_ess_weibull_cr_beta <- low_ess_weibull_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable))

num_species_low_ess <- low_ess_weibull_cr_beta |>
  distinct(sp_id) |>
  nrow()

print(num_species_low_ess)
```

- 400 sp have convergent issue.

### 2025-03-26 docs/cr_new.qmd
Initial commit for tree-allo

---
title: "Comparing CR models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
```

# Read in data

```{r}
options(width = 200)
tallo_wd_df_cr <- tar_read(tallo_wd_df_cr)
tallo_wd_df_cr

ang_cr_df <- tallo_wd_df_cr |> filter(division == "Angiosperm")
ang_cr_df

gym_cr_df <- tallo_wd_df_cr |> filter(division == "Gymnosperm")
gym_cr_df

```

# Power-law

## Ang
```{r}
fit_lr_nou_data_ang_cr <- tar_read(fit_lr_nou_data_ang_cr)
str(fit_lr_nou_data_ang_cr)

pl_ang_cr <- tar_read(fit_lr_nou_summary_pl_ang_cr)
pl_ang_cr

gamma_ang_pl <- pl_ang_cr |> filter(grepl("gamma", variable))

community_log_a_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_ang <- fit_lr_nou_data_ang_cr$log_x[,2] |> range()
log_x_seq_ang <- seq(log_x_range_ang[1], log_x_range_ang[2], length.out = 100)
fitted_log_y_seq_ang <- community_log_a_ang_pl + community_b_ang_pl * log_x_seq_ang
y_pred_ang <- exp(fitted_log_y_seq_ang)
df_ang_pl_cr <- tibble(y = y_pred_ang, x = exp(log_x_seq_ang))

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_pl_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")

```

## Gym

```{r}
fit_lr_nou_data_gym_cr <- tar_read(fit_lr_nou_data_gym_cr)
str(fit_lr_nou_data_gym_cr)

# pl_gym_cr <- tar_read(fit_lr_nou_summary_pl_biome_gym_cr)
# pl_gym_cr

pl_gym_cr <- tar_read(fit_lr_nou_summary_pl_gym_cr)
pl_gym_cr

gamma_gym_pl <- pl_gym_cr |> filter(grepl("gamma", variable))

community_log_a_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_gym_cr <- fit_lr_nou_data_gym_cr$log_x[,2] |> range()
log_x_seq_gym_cr <- seq(log_x_range_gym_cr[1], log_x_range_gym_cr[2], length.out = 100)

fitted_log_y_seq_gym <- community_log_a_gym_pl + community_b_gym_pl * log_x_seq_gym_cr
y_pred_gym_cr <- exp(fitted_log_y_seq_gym)

df_gym_pl_cr <- tibble(y = y_pred_gym_cr, x = exp(log_x_seq_gym_cr))
df_gym_pl_cr

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y))

```

# gMM

## Ang
```{r}
fit_nlr_nou_data_ang_cr <- tar_read(fit_nlr_nou_data_ang_cr)
str(fit_nlr_nou_data_ang_cr)

x_nlr_range_ang <- range(fit_nlr_nou_data_ang_cr$x)
x_nlr_seq_ang <- seq(x_nlr_range_ang[1], x_nlr_range_ang[2], length.out = 100)

gmm_ang_cr <- tar_read(fit_nlr_nou_summary_gmm_ang_cr)
gamma_ang_gmm <- gmm_ang_cr |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
community_log_a_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[2]") |> pull(q50)
community_k_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_gmm_seq <- community_log_a_ang_gmm + community_b_ang_gmm * log(x_nlr_seq_ang) - log(community_k_ang_gmm + x_nlr_seq_ang^community_b_ang_gmm)
y_ang_gmm_pred <- exp(community_fitted_log_y_cr_gmm_seq)
df_ang_gmm_cr <- tibble(y = y_ang_gmm_pred, x = x_nlr_seq_ang)

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")

```

## Gym

```{r}
fit_nlr_nou_data_gym_cr <- tar_read(fit_nlr_nou_data_gym_cr)
str(fit_nlr_nou_data_gym_cr)

x_nlr_range_gym_cr <- range(fit_nlr_nou_data_gym_cr$x)
x_nlr_seq_gym_cr  <- seq(x_nlr_range_gym_cr[1], x_nlr_range_gym_cr[2], length.out = 100)

gmm_gym_cr <- tar_read(fit_nlr_nou_summary_gmm_gym_cr)
gmm_gym_cr

gamma_gym_gmm <- gmm_gym_cr |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
gamma_gym_gmm

community_log_a_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_gym_gmm_seq <- community_log_a_gym_gmm + community_b_gym_gmm * log(x_nlr_seq_gym_cr) - log(community_k_gym_gmm + x_nlr_seq_gym_cr^community_b_gym_gmm)

y_gym_gmm_pred_cr <- exp(community_fitted_log_y_cr_gym_gmm_seq)

df_gym_gmm_cr <- tibble(y = y_gym_gmm_pred_cr, x = x_nlr_seq_gym_cr)
df_gym_gmm_cr

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y))
```

# Weibull

## Ang
```{r}
fit_nlr_nou_data_ang_cr <- tar_read(fit_nlr_nou_data_ang_cr)
x_nlr_range_ang_cr <- range(fit_nlr_nou_data_ang_cr$x)
x_nlr_seq_ang_cr <- seq(x_nlr_range_ang_cr[1], x_nlr_range_ang_cr[2], length.out = 100)

weibull_ang_cr <- tar_read(fit_nlr_nou_summary_weibull_ang_cr)
weibull_ang_cr

gamma_ang_wb <- weibull_ang_cr |> filter(grepl("gamma", variable))
community_log_a_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[2]") |> pull(q50)
community_k_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_wb_ang_seq <- community_log_a_ang_wb + log(1 - exp(-community_b_ang_wb * (x_nlr_seq_ang ^ community_k_ang_wb)))

y_ang_wb_pred_cr <- exp(community_fitted_log_y_cr_wb_ang_seq)

df_ang_wb_cr <- tibble(y = y_ang_wb_pred_cr, x = x_nlr_seq_ang_cr)

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")
```

## Gym

```{r}
fit_nlr_nou_data_gym_cr <- tar_read(fit_nlr_nou_data_gym_cr)
str(fit_nlr_nou_data_gym_cr)

x_nlr_range_gym_cr <- range(fit_nlr_nou_data_gym_cr$x)
x_nlr_seq_gym_cr <- seq(x_nlr_range_gym_cr[1], x_nlr_range_gym_cr[2], length.out = 100)

weibull_gym_cr <- tar_read(fit_nlr_nou_summary_weibull_gym_cr)
gamma_gym_wb <- weibull_gym_cr |> filter(grepl("gamma", variable))

community_log_a_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_wb_gym_seq <- community_log_a_gym_wb + log(1 - exp(-community_b_gym_wb * (x_nlr_seq_gym_cr ^ community_k_gym_wb)))

y_gym_wb_pred_cr <- exp(community_fitted_log_y_cr_wb_gym_seq)

df_gym_wb_cr <- tibble(y = y_gym_wb_pred_cr, x = x_nlr_seq_gym_cr)

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y))
```

# Plotting with density-based coloring log-scale

## Ang

```{r}
dens <- with(ang_cr_df, kde2d(log10(dbh), log10(cr), n = 300))
ix <- findInterval(log10(ang_cr_df$dbh), dens$x)
iy <- findInterval(log10(ang_cr_df$cr), dens$y)
ang_cr_df$density <- dens$z[cbind(ix, iy)]

p_ang_cr_den <- ggplot(data = ang_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_ang (cm)", y = "CR_ang (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

p_ang_cr_den

ggsave("figs/p_cr_ang_den.png", plot = p_ang_cr_den, width = 8, height = 6, dpi = 300, bg = "white")
```

## Gym

```{r}
dens <- with(gym_cr_df, kde2d(log10(dbh), log10(cr), n = 300))
ix <- findInterval(log10(gym_cr_df$dbh), dens$x)
iy <- findInterval(log10(gym_cr_df$cr), dens$y)
gym_cr_df$density <- dens$z[cbind(ix, iy)]

# Create the plot with log scale
p_gym_cr_den_log <- ggplot(data = gym_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "CR_gym (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

p_gym_cr_den_log

ggsave("figs/p_gym_cr_log_den.png", plot = p_gym_cr_den_log, width = 8, height = 6, dpi = 300, bg = "white")
```

# Plotting with density-based coloring non-log scale

## Ang

```{r}
dens <- with(ang_cr_df, kde2d(dbh, cr, n = 300))
ix <- findInterval(ang_cr_df$dbh, dens$x)
iy <- findInterval(ang_cr_df$cr, dens$y)
ang_cr_df$density <- dens$z[cbind(ix, iy)]

p_ang_cr_den_non <- ggplot(data = ang_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_ang_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_ang (cm)", y = "CR_ang (m)") +
  theme_minimal()

p_ang_cr_den_non

ggsave("figs/p_ang_cr_non_den_non.png", plot = p_ang_cr_den_non, width = 8, height = 6, dpi = 300, bg = "white")
```

## Gym

```{r}
dens <- with(gym_cr_df, kde2d(dbh, cr, n = 300))
ix <- findInterval(gym_cr_df$dbh, dens$x)
iy <- findInterval(gym_cr_df$cr, dens$y)
gym_cr_df$density <- dens$z[cbind(ix, iy)]

p_gym_cr_den_non <- ggplot(data = gym_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "CR_gym (m)") +
  theme_minimal()

p_gym_cr_den_non

ggsave("figs/p_gym_cr_non_den.png", plot = p_gym_cr_den_non, width = 8, height = 6, dpi = 300, bg = "white")
```

### 2025-03-26 docs/cr_up.qmd
Initial commit for tree-allo

---
title: "Comparing CR models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")

library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
```

# Read in data

```{r}
# Read in the data (pl)
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
str(stan_data_lr_cr)

# Read in the data (gMM, wb)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data from stan_data_nlr_cr (log_y, x, jj)
x_nlr <- stan_data_nlr_cr$x

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary <- tar_read(fit_lr2_summary_power_law1_cr)

gmm_summary
wb_summary
pl_summary
```

# Extract community-level parameters of each function

## Power-law

### Extract scaling parameters

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_cr = log(cr),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
log_dbh_sd <- sd(data2$log_dbh)
log_dbh_mean <- mean(data2$log_dbh)

log_cr_sd <- sd(data2$log_cr)
log_cr_mean <- mean(data2$log_cr)
```

### Computing Fitted Values

```{r}
pl_summary
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(q50)
community_log_a_pl
community_b_pl

## Computing Fitted Values
scaled_x_range <- stan_data_lr_cr$log_x[, 2] |> range()
log_x_seq <- seq(scaled_x_range[1], scaled_x_range[2], length.out = 100)
log_xx <- (log_x_seq * log_dbh_sd) + log_dbh_mean

# use scaled values to predict scaled values
log_y_pred_scaled <- community_log_a_pl + community_b_pl * log_x_seq
# unscaled
log_y_pred <- log_y_pred_scaled * log_cr_sd + log_cr_mean
community_fitted_cr_pl <- exp(log_y_pred)

df_pl <- tibble(y = community_fitted_cr_pl, x = exp(log_xx))
head(df_pl)
```

## gMM
```{r}
gmm_summary
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
gamma_gmm
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(q50)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm) 
community_k_gmm <- exp(community_k_hat_gmm)

community_log_a_gmm
community_b_gmm
community_k_gmm

## Computing Fitted Values
x_nlr_range <- range(x_nlr)
x_nlr_seq <- seq(x_nlr_range[1], x_nlr_range[2], length.out = 100)

# Compute the fitted values using the sequence of x-values
community_fitted_log_y_cr_gmm_seq <- community_log_a_gmm + community_b_gmm * log(x_nlr_seq) - log(community_k_gmm + x_nlr_seq^community_b_gmm)
community_fitted_cr_gmm_seq <- exp(community_fitted_log_y_cr_gmm_seq)

df_gmm <- tibble(y = community_fitted_cr_gmm_seq, x = x_nlr_seq)
head(df_gmm)
```

## Weibull
```{r}
wb_summary
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
gamma_wb
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(q50)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb) 
community_k_wb <- exp(community_k_hat_wb)

community_log_a_wb
community_b_wb
community_b_wb

# Compute the fitted values using the sequence of x-values
community_fitted_log_y_cr_wb_seq <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr_seq ^ community_k_wb)))
community_fitted_cr_wb_seq <- exp(community_fitted_log_y_cr_wb_seq)

df_wb <- tibble(y = community_fitted_cr_wb_seq, x = x_nlr_seq)
head(df_wb)
```

# Plot
## Power-law

```{r}
head(data2)
head(df_pl)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = cr), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "blue")

```

## gMM
```{r}
head(df_gmm)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = cr), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "green")

```

## Weibull
```{r}
head(df_wb)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = cr), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red")

```

## 3 functions in 1 panel: Log-scale
```{r}
combined_plot <- ggplot(data2) +
  geom_bin2d(data = data2, aes(x = dbh, y = cr), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y, color = "Power-law"), size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y, color = "gMM"), size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y, color = "Weibull"), size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "blue", "gMM" = "red", "Weibull" = "green"))

combined_plot

```

## 3 functions in 1 panel: Non-log scale

```{r}
combined_plot_non <- ggplot(data2) +
  geom_bin2d(data = data2, aes(x = dbh, y = cr), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y, color = "Power-law"), size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y, color = "gMM"), size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y, color = "Weibull"), size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "blue", "gMM" = "green", "Weibull" = "red"))

combined_plot_non
```

# Plotting with density-based coloring

## Log scale
```{r}
dens <- with(data2, kde2d(log10(dbh), log10(cr), n = 300))
ix <- findInterval(log10(data2$dbh), dens$x)
iy <- findInterval(log10(data2$cr), dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

combined_plot <- ggplot(data = data2, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

combined_plot
ggsave("cr_plot.png", plot = combined_plot, width = 8, height = 6, dpi = 300, bg= "white")
```

```{r}
combined_plot <- ggplot(data = data2, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +  # Power-Law
  geom_line(data = df_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +  # GMM
  geom_line(data = df_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +  # Weibull
  labs(
    x = "DBH (cm)",
    y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16),
        legend.position = "none") +  # Remove default legend

  annotate("segment", x = 100, xend = 130, y = 0.3, yend = 0.3, color = "#FF5733", size = 1) +  # Power-Law line
    annotate("text", x = 135, y = 0.3, label = "Power-Law", size = 5, hjust = 0) +  # Power-Law text

    annotate("segment", x = 100, xend = 130, y = 0.2, yend = 0.2, color = "#2ECC71", size = 1) +  # GMM line
    annotate("text", x = 135, y = 0.2, label = "gMM", size = 5, hjust = 0) +  # GMM text

    annotate("segment", x = 100, xend = 130, y = 0.125, yend = 0.125, color = "#0f92e9", size = 1) +  # Weibull line
    annotate("text", x = 135, y = 0.125, label = "Weibull", size = 5, hjust = 0)  # Weibull text
# Save the plot with annotations
ggsave("cr_plot_anno.png", plot = combined_plot, width = 8, height = 6, dpi = 300, bg = "white")
```

## Non-log scale
```{r}
dens <- with(data2, kde2d(dbh, cr, n = 300))
ix <- findInterval(data2$dbh, dens$x)
iy <- findInterval(data2$cr, dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

# Create the plot without log scales
combined_plot_non <- ggplot(data = data2, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "blue", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "green", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  theme_minimal()

combined_plot_non
```

### 2025-03-26 docs/dbh.qmd
Initial commit for tree-allo

---
title: "Plotting DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
# Define and run the targets pipeline
# tar_make()
```

# Plot DBH = aCR^b * H^c model (1, 5) (no WD)

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Compute CR^b and H^c for each observation
data_dbh$CR_b <- data_dbh$CR^community_b_cr
data_dbh$H_c <- data_dbh$H^community_b_h
data_dbh$CR_H <- data_dbh$CR_b*data_dbh$H_c

head(data_dbh)

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dbh_plot <- ggplot(data_long, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*CR^b * H^c",
       x = "CR_H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot
```

- Note: Plot looks weird and I am struggling with this TT_TT

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h

# Create a data frame for plotting with log-transformed values
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
data_dbh$community_fitted_log_DBH <- community_fitted_log_y_dbh

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_log_DBH_sp_", i)]] <- fitted_log_y_dbh[, i]
}

# Compute CR^b and H^c for each observation in the log scale
data_dbh$log_CR_b <- community_b_cr * data_dbh$log_CR
data_dbh$log_H_c <- community_b_h * data_dbh$log_H
data_dbh$log_CR_H <- data_dbh$log_CR_b + data_dbh$log_H_c

# Reshape the data for plotting
data_long <- reshape2::melt(data_dbh, id.vars = c("log_DBH", "log_CR_H", "sp", "community_fitted_log_DBH"), measure.vars = grep("fitted_log_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

# Plot with log-transformed values
dbh_plot <- ggplot(data_long, aes(x = log_CR_H, y = log_DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = log_CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) +  # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_log_DBH), color = "white", linewidth = 1.2) +
  labs(title = "log(DBH) = log(a) + b*log(CR) + c*log(H)",
       x = "log(CR^b * H^c)",
       y = "log(DBH)") +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1_log.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot

```

# Plot DBH = a(CR*H)^b model (6, 2) (with WD)

```{r}
# Read the data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x[, 2]
sp <- as.factor(stan_data_lr_dbh1$jj)

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract community-level parameters directly
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[2,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b * log_x

community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh1 <- data.frame(DBH = exp(log_y), CRxH = exp(log_x), sp = sp)
data_dbh1$community_fitted_DBH <- community_fitted_DBH

head(data_dbh1)
summary(data_dbh1)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh1[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

data_long <- melt(data_dbh1, id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh1), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)

# Plotting
dbh_plot1 <- ggplot(data_dbh1, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot1
```

# Fix the range of the CR*H axis

```{r}
# Function to create CR*H sequence for each species based on actual range
create_crxh_sequence <- function(species, log_x, length_out = 100) {
  crxh_values <- exp(log_x[stan_data_lr_dbh1$jj == species])
  min_crxh <- min(crxh_values)
  max_crxh <- max(crxh_values)
  seq(min_crxh, max_crxh, length.out = length_out)
}

# Calculate fitted DBH values for each species using gamma parameters
fitted_dbh_list <- lapply(1:length(log_a), function(i) {
  crxh_seq <- create_crxh_sequence(i, log_x)
  fitted_log_dbh <- log_a[i] + b[i] * log(crxh_seq)
  fitted_dbh <- exp(fitted_log_dbh)

  data.frame(
    CRxH = crxh_seq,
    DBH = fitted_dbh,
    variable = paste0("fitted_DBH_sp_", i)
  )
})

# Combine all species data frames into one
data_long <- do.call(rbind, fitted_dbh_list)
head(data_long)

# Original data for points
data_dbh1 <- data.frame(DBH = exp(log_y), CRxH = exp(log_x), sp = sp)
head(data_dbh1)

# Plotting the DBH against CR*H
dbh_plot1 <- ggplot(data_dbh1, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = DBH, group = variable), color = "#4682B4", alpha = 0.3) +
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CRxH_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot1
```

### 2025-03-26 docs/dbh_new.qmd
Initial commit for tree-allo

---
title: "Comparing DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
options(width = 150)
```

# Read in data

-

```{r}
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_dbh

ang_df <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Angiosperm")

gym_df <- tallo_reduced_lr_df_ang_dbh |>
  filter(division == "Gymnosperm")

angiosperms_df
gymnosperms_df

# tallo_reduced_lr_df_gym_dbh <- tar_read(tallo_reduced_lr_df_gym_dbh)
# tallo_reduced_lr_df_gym_dbh

# angiosperms_df1 <- tallo_reduced_lr_df_gym_dbh |>
#   filter(division == "Angiosperm")

# gymnosperms_df1 <- tallo_reduced_lr_df_gym_dbh |>
#   filter(division == "Gymnosperm")

# angiosperms_df1
# gymnosperms_df1

```

# DBH

## Ang
```{r}
fit_lr_nou_data_ang_dbh <- tar_read(fit_lr_nou_data_ang_dbh)
str(fit_lr_nou_data_ang_dbh)

dbh_ang <- tar_read(fit_lr_nou_summary_pl_ang_dbh)
dbh_ang

dbh_gamma_ang <- dbh_ang |> filter(grepl("gamma", variable))
dbh_gamma_ang

community_log_a_dbh_ang <- dbh_gamma_ang |> filter(variable == "gamma[1]") |> pull(q50)
community_b_dbh_ang <- dbh_gamma_ang |> filter(variable == "gamma[2]") |> pull(q50)

community_c_dbh_ang <- dbh_gamma_ang |> filter(variable == "gamma[3]") |> pull(q50)

head(fit_lr_nou_data_ang_dbh$log_x)

fitted_log_y_dbh_ang <- community_log_a_dbh_ang + community_b_dbh_ang * fit_lr_nou_data_ang_dbh$log_x[,2] + community_c_dbh_ang * fit_lr_nou_data_ang_dbh$log_x[,3]

y_pred_dbh_ang <- exp(fitted_log_y_dbh_ang)

df_dbh_ang <- tibble(y = y_pred_dbh_ang, x1 = exp(fit_lr_nou_data_ang_dbh$log_x[,2]), x2 = exp(fit_lr_nou_data_ang_dbh$log_x[,3]))

head(df_dbh_ang)

ggplot(df_dbh_ang, aes(x = x1, y = y)) +
  geom_point() +
  theme_minimal()

```

```{r}

```

```{r}

```

### 2025-03-26 docs/dbh_up.qmd
Initial commit for tree-allo

---
title: "Comparing DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(ggplot2)
options(width = 120)
```

# Read in data

```{r}
loo_tbl <- tar_read(loo_tbl)
print(loo_tbl, n = Inf)

dbh_models <- loo_tbl |> 
  filter(grepl("dbh($|1|2|3)", model)) |> 
  arrange(looic)
dbh_models

```

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(q50)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(q50)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Compute CR^b and H^c for each observation
data_dbh$CR_b <- data_dbh$CR^community_b_cr
data_dbh$H_c <- data_dbh$H^community_b_h
data_dbh$CR_H <- data_dbh$CR_b*data_dbh$H_c

head(data_dbh)

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

data_long <- data_long |> 
  mutate(partial_residuals = DBH - value)

data_summary <- data_long |> 
  group_by(sp, CR_H) |> 
  summarize(mean_residuals = mean(partial_residuals),
            sd_residuals = sd(partial_residuals),
            .groups = 'drop')me(legend.position = "none")

head(data_summary)

ggplot(data_summary, aes(x = CR_H, y = mean_residuals)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_residuals - sd_residuals, ymax = mean_residuals + sd_residuals), width = 0.01) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "CR^b * H^c", y = "Partial Residuals") +
  theme_minimal()

```

```{r}
# Extract relevant data for model (stan_data_lr_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)
# Ensure the correct indexing for variables
log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]  
log_x_h <- stan_data_lr_dbh$log_x[, 3]   
sp <- as.factor(stan_data_lr_dbh$jj)

# Read in the model summary
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)

# Filter rows where the variable is "beta" and extract parameters
beta <- dbh_nou_summary |> filter(grepl("beta", variable))

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)

# Compute fitted values for DBH in log scale using the multiple regression formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

# Extract gamma values for community-level effects
gamma <- dbh_nou_summary |> filter(grepl("gamma", variable))
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(q50)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(q50)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(q50)

# Calculate community-level fitted line in log scale
community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h

# Create a data frame for plotting (log scale)
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
data_dbh$community_fitted_log_DBH <- community_fitted_log_y_dbh

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_log_DBH_sp_", i)]] <- fitted_log_y_dbh[, i]
}

# Reshape data for plotting
data_long <- reshape2::melt(data_dbh, id.vars = c("log_DBH", "log_CR", "log_H", "sp", "community_fitted_log_DBH"),
                            measure.vars = grep("fitted_log_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dim(data_long)
```

# Visualization

## Scatter Plot of Observed vs. Fitted Values
```{r}
> ggplot(data_long, aes(x = value, y = log_DBH, color = sp)) +
+   geom_point() +
+   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+   labs(x = "Fitted Log(DBH)", y = "Observed Log(DBH)") +
+   theme_minimal() +
+   theme(legend.position = "none")
```

## Partial Residuals Plot
```{r}
data_long <- data_long |> 
  mutate(partial_residuals = log_DBH - value)

head(data_long)

ggplot(data_long, aes(x = sp, y = partial_residuals, color = sp)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Species", y = "Partial Residuals (Log Scale)") +
  theme_minimal() +
  theme(legend.position = "none")

```

## Residuals vs. Fitted Plot

```{r}
ggplot(data_long, aes(x = value, y = partial_residuals, color = sp)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Fitted Log(DBH)", y = "Partial Residuals") +
  theme_minimal() +
  theme(legend.position = "none")

```

### 2025-03-26 docs/den_plot.qmd
Initial commit for tree-allo

---
title: "Crown radius/Height vs DBH: size range of sample"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(targets)
library(ggplot2)
library(MASS) # For kde2d function
library(scales)
library(dplyr)
library(stringr)
library(reshape2)
# Define and run the targets pipeline
# tar_make()
```

# H-DBH

## Data preparation

```{r}
# Read in the data
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Extract relevant data
log_y <- stan_data_nlr_h$log_y
x <- stan_data_nlr_h$x
sp <- as.factor(stan_data_nlr_h$jj)

# Read in the model summary
wb1_h_summary <- tar_read(fit_nlr_summary_weibull1_h)
wb1_h_summary

# Extract the actual beta parameters
beta <- wb1_h_summary |> filter(grepl("beta\\[", variable))
beta

log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(mean)
k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(mean)

# Compute fitted values for H using the Weibull formula
fitted_log_y_h <- sapply(1:length(log_a), function(i) {
  log_a[i] + log(1 - exp(-b[i] * (x ^ k[i])))
})

fitted_h <- exp(fitted_log_y_h)

# Extract community-level parameters directly
gamma <- wb1_h_summary |> filter(grepl("gamma", variable))

# Correctly map community parameters
community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(mean)

# Apply exponential transformation to community-level parameters
community_log_a <- exp(community_log_a_hat)
community_b <- exp(community_b_hat) 
community_k <- exp(community_k_hat)

community_fitted_log_y_h <- community_log_a + log(1 - exp(-community_b * (x^community_k)))
community_fitted_H <- exp(community_fitted_log_y_h)

# Create a data frame for plotting
data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)
data_h$community_fitted_H <- community_fitted_H
head(data_h)
```

## Density plotting
```{r}
# Compute density using kde2d function
dens <- with(data_h, kde2d(log(DBH), log(H), n = 150))
ix <- findInterval(log(data_h$DBH), dens$x)
iy <- findInterval(log(data_h$H), dens$y)
data_h$density <- dens$z[cbind(ix, iy)]

# Plot with points and density-based coloring
plot_h_density_points <- ggplot(data_h, aes(x = DBH, y = H, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  labs(x = " DBH (cm)", y = " H (m)") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/h/H_DBH_density.png", plot = plot_h_density_points, width = 12, height = 8, dpi = 300, bg = "white")

print(plot_h_density_points)
```

## Plot 3 function in 1 panel with density based coloring

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(h)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
scaling_params <- data2 |>
    summarise(mean_log_dbh = mean(log_dbh), sd_log_dbh = sd(log_dbh),
              mean_log_h = mean(log_h), sd_log_h = sd(log_h))

scaling_params

# Functions for scaling and unscaling based on log-transformed values
scale_log_dbh <- function(log_dbh) (log_dbh - scaling_params$mean_log_dbh) / scaling_params$sd_log_dbh
unscale_log_h <- function(log_h_scaled) (log_h_scaled * scaling_params$sd_log_h) + scaling_params$mean_log_h

# Read in the data (power-law)
stan_data_lr_h <- tar_read(stan_data_lr_h)
log_y_lr <- stan_data_lr_h$log_y
log_x_lr <- stan_data_lr_h$log_x[, 2]
sp_lr <- as.factor(stan_data_lr_h$jj)

head(log_y_lr)
head(log_x_lr)

# Unscale log-transformed DBH and height for the power-law model
unscaled_log_dbh_pl <- (log_x_lr * scaling_params$sd_log_dbh) + scaling_params$mean_log_dbh
unscaled_log_h_pl <- (log_y_lr * scaling_params$sd_log_h) + scaling_params$mean_log_h

head(unscaled_log_dbh_pl)
head(unscaled_log_h_pl)

# Read in the data (gMM, Weibull)
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
log_y_nlr <- stan_data_nlr_h$log_y
x_nlr <- stan_data_nlr_h$x
sp_nlr <- as.factor(stan_data_nlr_h$jj)

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_h)
wb_summary <- tar_read(fit_nlr_summary_weibull1_h)
pl_summary <- tar_read(fit_lr2_summary_power_law1_h)

gmm_summary
wb_summary
pl_summary

# Extract community-level parameters for power-law
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
community_log_a_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(mean)

# Compute fitted values for power-law model using scaled log(DBH)
community_fitted_log_y_h_pl <- community_log_a_pl + community_b_pl * scale_log_dbh(unscaled_log_dbh_pl)
community_fitted_h_pl <- exp(unscale_log_h(community_fitted_log_y_h_pl))

# Extract community-level parameters for gMM
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm)
community_k_gmm <- exp(community_k_hat_gmm)

community_fitted_log_y_h_gmm <- community_log_a_gmm + community_b_gmm * log(x_nlr) - log(community_k_gmm + x_nlr^community_b_gmm)
community_fitted_h_gmm <- exp(community_fitted_log_y_h_gmm)

head(community_fitted_h_gmm)

# Extract community-level parameters for Weibull
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb)
community_k_wb <- exp(community_k_hat_wb)

community_fitted_log_y_h_wb <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr ^ community_k_wb)))
community_fitted_h_wb <- exp(community_fitted_log_y_h_wb)

head(community_fitted_h_wb)

# Create a data frame for plotting
data_h_pl_unscaled <- data.frame(H = exp(unscaled_log_h_pl), DBH = exp(unscaled_log_dbh_pl), sp = as.factor(stan_data_lr_h$jj), community_fitted_H_pl = community_fitted_h_pl)

head(data_h_pl_unscaled)

data_h_gb <- data.frame(H = exp(log_y_nlr), DBH = x_nlr, sp = sp_nlr)
head(data_h_gb)

data_h_gb$community_fitted_H_gmm <- community_fitted_h_gmm
data_h_gb$community_fitted_H_wb <- community_fitted_h_wb

head(data_h_pl_unscaled)
head(data_h_gb)

# Combine the data frames
data_combined <- merge(data_h_pl_unscaled, data_h_gb, by = "sp", all = TRUE)
head(data_combined)

# Separate the community-fitted values for plotting
data_community_pl <- data_combined |> select(DBH.x, community_fitted_H_pl) |> distinct()
data_community_gmm <- data_combined |> select(DBH.y, community_fitted_H_gmm) |> distinct()
data_community_wb <- data_combined |> select(DBH.y, community_fitted_H_wb) |> distinct()

head(data_community_pl)
head(data_community_gmm)
```

## Log scale plot with density based coloring (non-log scale)

```{r}
library(MASS)
library(dplyr)

head(data_combined)

dens <- with(data_combined, kde2d(log(DBH.y), log(H.y), n = 150))
ix <- findInterval(log(data_combined$DBH.y), dens$x)
iy <- findInterval(log(data_combined$H.y), dens$y)
data_combined$density <- dens$z[cbind(ix, iy)]

head(data_combined)

# Plotting the three functions in one panel with density-based coloring for points
combined_plot <- ggplot(data_combined) +
  geom_point(aes(x = DBH.y, y = H.y, color = density), size = 0.5, alpha = 0.8) +
  geom_line(aes(x = DBH.x, y = community_fitted_H_pl), color = "blue", size = 1, data = data_community_pl) +
  geom_line(aes(x = DBH.y, y = community_fitted_H_gmm), color = "red", size = 1, data = data_community_gmm) +
  geom_line(aes(x = DBH.y, y = community_fitted_H_wb), color = "green", size = 1, data = data_community_wb) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/h/DBH_H_community_combined_with_density.png", plot = combined_plot, width = 12, height = 8, dpi = 300, bg = "white")
```

## Non-log scale plot with density based coloring

```{r}
# Plotting with a smoother density representation
combined_plot2 <- ggplot(data_combined, aes(x = DBH.y, y = H.y, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  geom_line(aes(x = DBH.x, y = community_fitted_H_pl), color = "blue", size = 1, data = data_community_pl) +
  geom_line(aes(x = DBH.y, y = community_fitted_H_gmm), color = "red", size = 1, data = data_community_gmm) +
  geom_line(aes(x = DBH.y, y = community_fitted_H_wb), color = "green", size = 1, data = data_community_wb) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/h/DBH_H_community_combined_non_log.png", plot = combined_plot2, width = 12, height = 8, dpi = 300, bg = "white")

```

# CR-DBH

## Data preparation
```{r}
# Read in the data (wb, gMM)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data 
log_y <- stan_data_nlr_cr$log_y
x <- stan_data_nlr_cr$x
sp <- as.factor(stan_data_nlr_cr$jj)

# Read in the model summary
options(width = 120)
wb1_cr_summary <- tar_read(fit_nlr_summary_weibull1_cr)
wb1_cr_summary

# Extract the actual beta parameters
beta <- wb1_cr_summary |> filter(grepl("beta\\[", variable))
beta

log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(mean)
k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(mean)

# Compute fitted values for CR using the Weibull formula
fitted_log_y_cr <- sapply(1:length(log_a), function(i) {
  log_a[i] + log(1 - exp(-b[i] * (x ^ k[i])))
})

fitted_cr <- exp(fitted_log_y_cr)

# Extract community-level parameters directly
gamma <- wb1_cr_summary |> filter(grepl("gamma", variable))

# Correctly map community parameters
community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(mean)

# Apply exponential transformation to community-level parameters
community_log_a <- exp(community_log_a_hat)
community_b <- exp(community_b_hat) 
community_k <- exp(community_k_hat)

community_fitted_log_y_cr <- community_log_a + log(1 - exp(-community_b * (x^community_k)))
community_fitted_CR <- exp(community_fitted_log_y_cr)

# Create a data frame for plotting
data_cr <- data.frame(CR = exp(log_y), DBH = x, sp = sp)
data_cr$community_fitted_CR <- community_fitted_CR
head(data_cr)

```

## Density plotting

```{r}
# Compute density using kde2d function
dens <- with(data_cr, kde2d(log(DBH), log(CR), n = 100))
ix <- findInterval(log(data_cr$DBH), dens$x)
iy <- findInterval(log(data_cr$CR), dens$y)
data_cr$density <- dens$z[cbind(ix, iy)]

# Plot with points and density-based coloring
plot_cr_density_points <- ggplot(data_cr, aes(x = DBH, y = CR, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  labs(x = " DBH (cm)", y = " CR (m)") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/cr/CR_DBH_density.png", plot = plot_cr_density_points, width = 12, height = 8, dpi = 300, bg = "white")

print(plot_cr_density_points)
```

## Plot 3 functions in 1 panel with density-based coloring (log scale)

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
scaling_params <- data2 |>
    summarise(mean_log_dbh = mean(log_dbh), sd_log_dbh = sd(log_dbh),
              mean_log_cr = mean(log_cr), sd_log_cr = sd(log_cr))

scaling_params

scale_log_dbh <- function(log_dbh) (log_dbh - scaling_params$mean_log_dbh) / scaling_params$sd_log_dbh
unscale_log_cr <- function(log_cr_scaled) (log_cr_scaled * scaling_params$sd_log_cr) + scaling_params$mean_log_cr

# Read in the data (pl)
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
str(stan_data_lr_cr)

# Extract relevant data from stan_data_lr_cr (log_y, x, jj)
log_y_lr <- stan_data_lr_cr$log_y
log_x_lr <- stan_data_lr_cr$log_x[, 2]
sp_lr <- as.factor(stan_data_lr_cr$jj)

str(log_y_lr)
str(log_x_lr)
str(sp_lr)

# Unscale log-transformed DBH and height for the power-law model
unscaled_log_dbh_pl <- (log_x_lr * scaling_params$sd_log_dbh) + scaling_params$mean_log_dbh
unscaled_log_cr_pl <- (log_y_lr * scaling_params$sd_log_cr) + scaling_params$mean_log_cr

head(unscaled_log_dbh_pl)
head(unscaled_log_cr_pl)

# Read in the data (gMM, wb)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data from stan_data_nlr_cr (log_y, x, jj)
log_y_nlr <- stan_data_nlr_cr$log_y
x_nlr <- stan_data_nlr_cr$x
sp_nlr <- as.factor(stan_data_nlr_cr$jj)

str(log_y_nlr)
str(x_nlr)

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary <- tar_read(fit_lr2_summary_power_law1_cr)

pl_summary
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(mean)

community_log_a_pl
community_b_pl

## gMM
gmm_summary
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
gamma_gmm
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm) 
community_k_gmm <- exp(community_k_hat_gmm)

community_log_a_gmm
community_b_gmm
community_k_gmm

# Weibull
wb_summary
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
gamma_wb
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb) 
community_k_wb <- exp(community_k_hat_wb)

community_log_a_wb
community_b_wb
community_b_wb

# Computing Fitted Values for Community-level Parameters
#pl
community_fitted_log_y_cr_pl <- community_log_a_pl + community_b_pl * scale_log_dbh(unscaled_log_dbh_pl)
community_fitted_cr_pl <- exp(unscale_log_cr(community_fitted_log_y_cr_pl))

head(community_fitted_cr_pl)

# head(data2$log_dbh)
# head(unscaled_log_dbh_pl)
# head(scale_log_dbh(unscaled_log_dbh_pl))

#gMM

community_fitted_log_y_cr_gmm <- community_log_a_gmm + community_b_gmm * log(x_nlr) - log(community_k_gmm + x_nlr^community_b_gmm)
community_fitted_cr_gmm <- exp(community_fitted_log_y_cr_gmm)

head(community_fitted_cr_gmm)

# wb

community_fitted_log_y_cr_wb <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr ^ community_k_wb)))
community_fitted_cr_wb <- exp(community_fitted_log_y_cr_wb)

head(community_fitted_cr_wb)

# Create a data frame for plotting
data_cr_pl_unscaled <- data.frame(CR = exp(unscaled_log_cr_pl), DBH = exp(unscaled_log_dbh_pl), sp = as.factor(stan_data_lr_cr$jj), community_fitted_CR_pl = community_fitted_cr_pl)

head(data_cr_pl_unscaled)

data_cr_gb <- data.frame(CR = exp(log_y_nlr), DBH = x_nlr, sp = sp_nlr)
head(data_cr_gb)

data_cr_gb$community_fitted_CR_gmm <- community_fitted_cr_gmm
data_cr_gb$community_fitted_CR_wb <- community_fitted_cr_wb

head(data_cr_pl_unscaled)
head(data_cr_gb)

# Combine the data frames
data_combined <- merge(data_cr_pl_unscaled, data_cr_gb, by = "sp", all = TRUE)
head(data_combined)

# Separate the community-fitted values for plotting
data_community_pl <- data_combined |> select(DBH.x, community_fitted_CR_pl) |> distinct()
data_community_gmm <- data_combined |> select(DBH.y, community_fitted_CR_gmm) |> distinct()
data_community_wb <- data_combined |> select(DBH.y, community_fitted_CR_wb) |> distinct()

head(data_community_pl)
head(data_community_gmm)
head(data_community_wb)

```

## Log scale plot with density based coloring (non-log scale)

```{r}
library(MASS)
library(dplyr)

head(data_combined)

# Calculate density
dens <- with(data_combined, kde2d(log(DBH.y), log(CR.y), n = 200))
ix <- findInterval(log(data_combined$DBH.y), dens$x)
iy <- findInterval(log(data_combined$CR.y), dens$y)
data_combined$density <- dens$z[cbind(ix, iy)]

head(data_combined)

combined_plot <- ggplot(data_combined) +
  geom_point(aes(x = DBH.y, y = CR.y, color = density), size = 0.5, alpha = 0.8) +
  geom_line(data = data_community_pl, aes(x = DBH.x, y = community_fitted_CR_pl), color = "blue", size = 1) +
  geom_line(data = data_community_gmm, aes(x = DBH.y, y = community_fitted_CR_gmm), color = "red", size = 1) +
  geom_line(data = data_community_wb, aes(x = DBH.y, y = community_fitted_CR_wb), color = "green", size = 1) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/cr/DBH_CR_community_combined_with_density.png", plot = combined_plot, width = 12, height = 8, dpi = 300, bg = "white")
```

## Non-log scale plot with density based coloring

```{r}
# # Calculate density using kde2d and convert it to a data frame for plotting
# dens <- with(data_combined, kde2d(log(DBH.y), log(CR.y), n = 150))
# ix <- findInterval(log(data_combined$DBH.y), dens$x)
# iy <- findInterval(log(data_combined$CR.y), dens$y)
# data_combined$density <- dens$z[cbind(ix, iy)]

# Plotting with a smoother density representation
combined_plot2 <- ggplot(data_combined, aes(x = DBH.y, y = CR.y, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  geom_line(aes(x = DBH.x, y = community_fitted_CR_pl), color = "blue", size = 1, data = data_community_pl) +
  geom_line(aes(x = DBH.y, y = community_fitted_CR_gmm), color = "red", size = 1, data = data_community_gmm) +
  geom_line(aes(x = DBH.y, y = community_fitted_CR_wb), color = "green", size = 1, data = data_community_wb) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 12),
    panel.grid = element_blank(),
    legend.position = "none"
  )

# Save the plot
ggsave(filename = "figs/cr/DBH_CR_community_combined_non_log.png", plot = combined_plot2, width = 12, height = 8, dpi = 300, bg = "white")
```

# DBH - CR*H

```{r}
# Load necessary libraries
library(ggplot2)
library(MASS)
library(reshape2)

# Data preparation
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x[,2]
sp <- as.factor(stan_data_lr_dbh1$jj)

# Read in the model summary
dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)

# Extract community-level parameters
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[2,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b * log_x
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create data frame for plotting
data_dbh1 <- data.frame(DBH = exp(log_y), CRxH = exp(log_x), sp = sp)
data_dbh1$community_fitted_DBH <- community_fitted_DBH

# Add species-specific fitted values
for (i in 1:length(log_a)) {
  data_dbh1[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Reshape data for plotting
data_long <- melt(data_dbh1, id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), 
                  measure.vars = grep("fitted_DBH_sp_", names(data_dbh1), value = TRUE))

# Density Calculation
dens <- with(data_dbh1, kde2d(log(DBH), log(CRxH), n = 50))
ix <- findInterval(log(data_dbh1$CRxH), dens$x, all.inside = TRUE)
iy <- findInterval(log(data_dbh1$DBH), dens$y, all.inside = TRUE)
data_dbh1$density <- dens$z[cbind(ix, iy)]

# Plotting
dbh_plot1 <- ggplot(data_dbh1, aes(x = CRxH, y = DBH)) +
  geom_point(aes(color = density), alpha = 0.5) +
  geom_line(aes(x = CRxH, y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR_CR*H_density.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")
```

### 2025-03-26 docs/diag.qmd
Initial commit for tree-allo

---
title: "Diagnostic checking"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
# tar_make()
```

# Diagnostic-checking

- Note: List of all potential diagnostic file names (lr2, nlr)
  "fit_lr2_diagnostics_power_law1_cr", "fit_lr2_diagnostics_power_law1_dbh",
  "fit_lr2_diagnostics_power_law1_dbh1", "fit_lr2_diagnostics_power_law1_dbh2",
  "fit_lr2_diagnostics_power_law1_dbh3", "fit_lr2_diagnostics_power_law1_h",
  "fit_lr2_diagnostics_power_law1_nou_cr", "fit_lr2_diagnostics_power_law1_nou_dbh",
  "fit_lr2_diagnostics_power_law1_nou_dbh1", "fit_lr2_diagnostics_power_law1_nou_dbh2",
  "fit_lr2_diagnostics_power_law1_nou_dbh3", "fit_lr2_diagnostics_power_law1_nou_h",
  "fit_nlr_diagnostics_gmm1_cr", "fit_nlr_diagnostics_gmm1_h",
  "fit_nlr_diagnostics_gmm1_nou_cr", "fit_nlr_diagnostics_gmm1_nou_h",
  "fit_nlr_diagnostics_weibull1_cr", "fit_nlr_diagnostics_weibull1_h",
  "fit_nlr_diagnostics_weibull1_nou_cr", "fit_nlr_diagnostics_weibull1_nou_h"

```{r}
options(width = 200)
# Load diagnostics data for specific fit_lr2 and fit_nlr models
pl1_cr <- tar_read(fit_lr2_diagnostics_power_law1_cr)
pl1_dbh <- tar_read(fit_lr2_diagnostics_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_diagnostics_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_diagnostics_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_diagnostics_power_law1_dbh3)
pl1_h <- tar_read(fit_lr2_diagnostics_power_law1_h)
pl1_nou_cr <- tar_read(fit_lr2_diagnostics_power_law1_nou_cr)
pl1_nou_dbh <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh3)
pl1_nou_h <- tar_read(fit_lr2_diagnostics_power_law1_nou_h)
gmm1_cr <- tar_read(fit_nlr_diagnostics_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_diagnostics_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_diagnostics_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_diagnostics_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_diagnostics_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_diagnostics_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_diagnostics_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_diagnostics_weibull1_nou_h)
```

```{r}
summary(pl1_cr)
summary(pl1_dbh)
summary(pl1_dbh1)
summary(pl1_dbh2)
summary(pl1_dbh3)
summary(pl1_h)
summary(pl1_nou_cr)
summary(pl1_nou_dbh)
summary(pl1_nou_dbh1)
summary(pl1_nou_dbh2)
summary(pl1_nou_dbh3)
summary(pl1_nou_h)

# fit_nlr models summaries
summary(gmm1_cr)
summary(gmm1_h)
summary(gmm1_nou_cr)
summary(gmm1_nou_h)
summary(weibull1_cr)
summary(weibull1_h)
summary(weibull1_nou_cr)
summary(weibull1_nou_h)

```

# Checking ESS_bulk and R_hat
```{r}
pl1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
pl1_dbh <- tar_read(fit_lr2_summary_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_summary_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_summary_power_law1_dbh3)
pl1_h <- tar_read(fit_lr2_summary_power_law1_h)
pl1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
pl1_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_summary_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_summary_power_law1_nou_dbh3)
pl1_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)

gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_summary_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_summary_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_summary_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
```

```{r}
pl1_cr
pl1_dbh
pl1_dbh1
pl1_dbh2
pl1_dbh3
pl1_h
pl1_nou_cr
pl1_nou_dbh
pl1_nou_dbh1
pl1_nou_dbh2
pl1_nou_dbh3
pl1_nou_h
gmm1_cr
gmm1_h
gmm1_nou_cr
gmm1_nou_h
weibull1_cr
weibull1_h
weibull1_nou_cr
weibull1_nou_h
```

```{r}
summary(pl1_cr)
summary(pl1_dbh)
summary(pl1_dbh1)
summary(pl1_dbh2)
summary(pl1_dbh3)
summary(pl1_h)
summary(pl1_nou_cr)
summary(pl1_nou_dbh)
summary(pl1_nou_dbh1)
summary(pl1_nou_dbh2)
summary(pl1_nou_dbh3)
summary(pl1_nou_h)

# fit_nlr models summaries
summary(gmm1_cr)
summary(gmm1_h)
summary(gmm1_nou_cr)
summary(gmm1_nou_h)
summary(weibull1_cr)
summary(weibull1_h)
summary(weibull1_nou_cr)
summary(weibull1_nou_h)
```

# Divergent transition checking
```{r}
library(targets)
library(posterior)
library(bayesplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cmdstanr)
library(stringr)

```

```{r}
divergent_tbl <- tar_read(divergent_tbl)
print(divergent_tbl, n = Inf)
```

# Parallel coordinates plot
```{r}
pl1_h_sam <- tar_read(fit_lr2_draws_power_law1_h)
pl1_h_sam_df <- as_draws_df(pl1_h_sam)
pl1_h_sam_plot <- pl1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]

pl1_h_pairs <- mcmc_pairs(pl1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_h_pairs
```

```{r}
pl1_cr_sam <- tar_read(fit_lr2_draws_power_law1_cr)
pl1_cr_sam_df <- as_draws_df(pl1_cr_sam)
pl1_cr_sam_df
pl1_cr_sam_plot <- pl1_cr_sam_df[,c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_cr_pairs <- mcmc_pairs(pl1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_cr_pairs

```

```{r}
pl1_dbh_sam <- tar_read(fit_lr2_draws_power_law1_dbh)
pl1_dbh_sam_df <- as_draws_df(pl1_dbh_sam)
pl1_dbh_sam_df
pl1_dbh_sam_plot <- pl1_dbh_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[3,1]",  "gamma[1,2]", "gamma[2,2]", "gamma[3,2]")]
pl1_dbh_pairs <- mcmc_pairs(pl1_dbh_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh_pairs
```

```{r}
pl1_dbh1_sam <- tar_read(fit_lr2_draws_power_law1_dbh1)
pl1_dbh1_sam_df <- as_draws_df(pl1_dbh1_sam)
pl1_dbh1_sam_df
pl1_dbh1_sam_plot <- pl1_dbh1_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh1_pairs <- mcmc_pairs(pl1_dbh1_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh1_pairs

```

```{r}
pl1_dbh2_sam <- tar_read(fit_lr2_draws_power_law1_dbh2)
pl1_dbh2_sam_df <- as_draws_df(pl1_dbh2_sam)
pl1_dbh2_sam_df
pl1_dbh2_sam_plot <- pl1_dbh2_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh2_pairs <- mcmc_pairs(pl1_dbh2_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh2_pairs
```

```{r}
pl1_dbh3_sam <- tar_read(fit_lr2_draws_power_law1_dbh3)
pl1_dbh3_sam_df <- as_draws_df(pl1_dbh3_sam)
pl1_dbh3_sam_df
pl1_dbh3_sam_plot <- pl1_dbh3_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh3_pairs <- mcmc_pairs(pl1_dbh3_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh3_pairs
```

```{r}
pl1_nou_h_sam <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl1_nou_h_sam_df <- as_draws_df(pl1_nou_h_sam)
pl1_nou_h_sam_df
pl1_nou_h_sam_plot <- pl1_nou_h_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_h_pairs <- mcmc_pairs(pl1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_h_pairs
```

```{r}
pl1_nou_cr_sam <- tar_read(fit_lr2_draws_power_law1_nou_cr)
pl1_nou_cr_sam_df <- as_draws_df(pl1_nou_cr_sam)
pl1_nou_cr_sam_df
pl1_nou_cr_sam_plot <- pl1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_cr_pairs <- mcmc_pairs(pl1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_cr_pairs
```

```{r}
pl1_nou_dbh_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl1_nou_dbh_sam_df <- as_draws_df(pl1_nou_dbh_sam)
pl1_nou_dbh_sam_df
pl1_nou_dbh_sam_plot <- pl1_nou_dbh_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
pl1_nou_dbh_pairs <- mcmc_pairs(pl1_nou_dbh_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh_pairs
```

```{r}
pl1_nou_dbh1_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl1_nou_dbh1_sam_df <- as_draws_df(pl1_nou_dbh1_sam)
pl1_nou_dbh1_sam_df 
pl1_nou_dbh1_sam_plot <- pl1_nou_dbh1_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh1_pairs <- mcmc_pairs(pl1_nou_dbh1_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh1_pairs
```

```{r}
pl1_nou_dbh2_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl1_nou_dbh2_sam_df <- as_draws_df(pl1_nou_dbh2_sam)
pl1_nou_dbh2_sam_df
pl1_nou_dbh2_sam_plot <- pl1_nou_dbh2_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh2_pairs <- mcmc_pairs(pl1_nou_dbh2_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh2_pairs
```

```{r}
pl1_nou_dbh3_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl1_nou_dbh3_sam_df <- as_draws_df(pl1_nou_dbh3_sam)
pl1_nou_dbh3_sam_df
pl1_nou_dbh3_sam_plot <- pl1_nou_dbh3_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh3_pairs <- mcmc_pairs(pl1_nou_dbh3_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh3_pairs
```

```{r}
gmm1_h_sam <- tar_read(fit_nlr_draws_gmm1_h)
gmm1_h_sam_df <- as_draws_df(gmm1_h_sam)
gmm1_h_sam_df
gmm1_h_sam_plot <- gmm1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
gmm1_h_pairs <- mcmc_pairs(gmm1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_h_pairs
```

```{r}
gmm1_cr_sam <- tar_read(fit_nlr_draws_gmm1_cr)
gmm1_cr_sam_df <- as_draws_df(gmm1_cr_sam)
gmm1_cr_sam_df
gmm1_cr_sam_plot <- gmm1_cr_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
gmm1_cr_pairs <- mcmc_pairs(gmm1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_cr_pairs
```

```{r}
gmm1_nou_h_sam <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm1_nou_h_sam_df <- as_draws_df(gmm1_nou_h_sam)
gmm1_nou_h_sam_df
gmm1_nou_h_sam_plot <- gmm1_nou_h_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
gmm1_nou_h_pairs <- mcmc_pairs(gmm1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_nou_h_pairs
```

```{r}
gmm1_nou_cr_sam <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm1_nou_cr_sam_df <- as_draws_df(gmm1_nou_cr_sam)
gmm1_nou_cr_sam_df
gmm1_nou_cr_sam_plot <- gmm1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
gmm1_nou_cr_pairs <- mcmc_pairs(gmm1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_nou_cr_pairs
```

```{r}
weibull1_h_sam <- tar_read(fit_nlr_draws_weibull1_h)
weibull1_h_sam_df <- as_draws_df(weibull1_h_sam)
weibull1_h_sam_df
weibull1_h_sam_plot <- weibull1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
weibull1_h_pairs <- mcmc_pairs(weibull1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_h_pairs
```

```{r}
weibull1_cr_sam <- tar_read(fit_nlr_draws_weibull1_cr)
weibull1_cr_sam_df <- as_draws_df(weibull1_cr_sam)
weibull1_cr_sam_df
weibull1_cr_sam_plot <- weibull1_cr_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
weibull1_cr_pairs <- mcmc_pairs(weibull1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_cr_pairs
```

```{r}
weibull1_nou_h_sam <- tar_read(fit_nlr_draws_weibull1_nou_h)
weibull1_nou_h_sam_df <- as_draws_df(weibull1_nou_h_sam)
weibull1_nou_h_sam
weibull1_nou_h_sam_plot <- weibull1_nou_h_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
weibull1_nou_h_pairs <- mcmc_pairs(weibull1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_nou_h_pairs
```

```{r}
weibull1_nou_cr_sam <- tar_read(fit_nlr_draws_weibull1_nou_cr)
weibull1_nou_cr_sam_df <- as_draws_df(weibull1_nou_cr_sam)
weibull1_nou_cr_sam_df
weibull1_nou_cr_sam_plot <- weibull1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
weibull1_nou_cr_pairs <- mcmc_pairs(weibull1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_nou_cr_pairs
```

### 2025-03-26 docs/div_trans.qmd
Initial commit for tree-allo

---
title: "Divergent transitions (gmm1_nou_h)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(posterior)
library(bayesplot)
library(ggplot2)
library(cmdstanr)
library(targets)
```

# Data loading

```{r}
gmm1_nou_h_diag <- tar_read(fit_nlr_diagnostics_gmm1_nou_h) 
gmm1_nou_h_diag
summary(gmm1_nou_h_diag)
```

# Divergent Iterations

```{r}
# Identify divergent iterations
divergent_iterations <- gmm1_nou_h_diag[gmm1_nou_h_diag$divergent__ == 1, ]
divergent_iterations

```

# Plotting

```{r}
fit <- tar_read(fit_nlr_mcmc_gmm1_nou_h) 
mcmc_samples <- as_draws_array(fit$draws())
mcmc_samples

trace_plot <- mcmc_trace(mcmc_samples, pars = c("gamma[1]", "gamma[2]", "gamma[3]"))

pair_plot <- mcmc_pairs(mcmc_samples, pars = c("gamma[1]", "gamma[2]", "gamma[3]"))

trace_plot
pair_plot
```

### 2025-03-26 docs/examples.qmd
Initial commit for tree-allo

---
title: "Crown width vs. DBH"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

I will explain a MLE example for crown width vs. DBH using a power-law model.

# Load the packages
```{r}
library(tidyverse)
library(here)
```

# Data preparation

I prepared `tallo_cleaned.csv` withouth wood density but with other iformation like laitiude and family.
We may need these information later.
For example, we can use faimly or biomes as random effects if needed.

```{r}
tallo_df <- read_csv(here("data/tallo_cleaned.csv"))
tallo_df
```

I also prepared wood density data for each species (`wd.csv`)
It's better to have species-based data (i.e., number of species = number of rows) for the analysis.

```{r}
wd_df <- read_csv(here("data/wd.csv"))
wd_df <- wd_df |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_df
```

You can easily join these two dataframes by `sp` (species name) before the analysis.

```{r}
tallo_wd <- left_join(tallo_df, wd_df, by = "sp")
tallo_wd
```

# Example for *C* vs. *D*

For tree *i* and species *j*, we assume that crown width ($C$) and DBH ($D$) has a power-law relationship :
$$
\text{ln}C_{i} \sim N(\alpha_j + \beta_j \text{ln}D_{i})
$$ {#eq-cd}

and species-level varaition can be explaiend by wood density ($\rho$):
$$
\begin{split}
\text{ln}C_{i} &= \alpha_{j} + \beta_{j}\text{ln}D_{i} + \epsilon_{i} \\
\alpha_j &= \gamma_{0}^{(\alpha)} + \gamma_{1}^{(\alpha)}\rho_{j} + \eta_{j}^{(\alpha)} \\
\beta_j &= \gamma_{0}^{(\beta)} + \gamma_{1}^{(\beta)}\rho_{j} +\eta_{j}^{(\beta)}.
\end{split}
$$ {#eq-wd_lmm}

To use `lme4::lmer`, we need to re-express the model (@eq-wd_lmm) as:

$$
\text{ln}C_{i} =
(\gamma_{0}^{(\alpha)} + \gamma_{1}^{(\alpha)}\rho_{j} + \eta_{j}^{(\alpha)})
 + (\gamma_{0}^{(\beta)} + \gamma_{1}^{(\beta)}\rho_{j} +\eta_{j}^{(\beta)})\text{ln}D_j + \epsilon_{i}.
$$ {#eq-wd_lmm_re}

We define new variables (
$a_j = \gamma_{0}^{(\alpha)} + \eta_{j}^{(\alpha)}$,
$b_j = \gamma_{0}^{(\beta)} + \eta_{j}^{(\beta)}$,
$c = \gamma_{1}^{(\alpha)}$, and
$d = \gamma_{1}^{(\beta)}$),
and we can further re-express the above model (@eq-wd_lmm_re) as:

$$
\text{ln}C_{i}= a_j + b_j \text{ln}D_{i} + c \rho_{j} + d \text{ln}D_{i} \times \rho_{j} + \epsilon_{i}.
$$ {#eq-lmer}

The correspoding R code looks like the following:

```{r, eval=FALSE}
lme4::lmer(log(cr) ~ log(dbh) + wd_s + log(dbh):wd_s + (1 + log(dbh) | sp), ...)
```

In this code, `(1 + log(dbh) | sp)` indiates varying intercepts ($a_j$) and slopes ($b_j$) for each species.
If you use `(1 | sp)`, this indicats only varying intercepts ($a_j$).

Note that:

- $a_j$ and $b_j$ are vectors of lenght $J$ (number of species).
- $c$ and $d$ are sclars (single values).

```{r}
fit_cd_power <- lme4::lmer(log(cr) ~ log(dbh) + wd_s + log(dbh):wd_s + (1 + log(dbh) | sp), data = tallo_wd)
summary(fit_cd_power)
```

The t-value for wd_s is relatively large (> 3), which indicates $c = \gamma_{1}^{(\alpha)}$ is (probably) significant.
This pattern was not that clear in Martínez Cano's paper.
On the other hand, $d = \gamma_{1}^{(\beta)}$ doesn't seem to be significant.

# Next

Power-law models with `lme4` are just preliminary analysis.
We also want to use gMM and Weibull, which requrie other R packages.
It will be more straightforward to use `stan` for all the models.

- Write **equations** for stan codes for:
  - power-law
  - gMM
  - Weibull

- Write stan codes for:
  - power-law
  - gMM
  - Weibull

- Write simple stan codes for simple data.

- Create dummy data for your stan codes.

#  Liner models with dummy data example

## Dummy data
$$
y_i \sim N(\beta_0 + \beta_1 x_i, \sigma)
$$ {#eq-lm}

where $\beta_0$ = 1.2, $\beta_1$ = -0.8, $\sigma$ = 0.5, and the number of samples (n) is 100.

```{r}
set.seed(123)
n <- 100
beta0 <- 1.2
beta1 <- -0.8
sigma <- 0.5

x <- rnorm(n)
mu <- beta0 + beta1 * x
y <- rnorm(n, mu, sigma)
plot(x, y)
```

### Matrix Operations

In practice, we use matrix operations to handle computations:

```{r}
x_mat <- cbind(1, x)
mu2 <- x_mat %*% c(beta0, beta1)

# mu and mu2 are expected to be equal
testthat::expect_equal(mu, as.numeric(mu2))
```

## Linear Regression (MLE Approach)

```{r}
fit <- lm(y ~ x)
summary(fit)
```

## Bayesian Analysis with Stan

### Stan model

The Stan model code is stored in `stan/lm_example.stan`:

```{stan, size="small", file="../stan/lm_example.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

As you can see, the stan code looks like how the dummy data was generated.
This is why it is a good practice to write both stan code and code to generate dummy data.
It helps you to understand the model and to debug the code.

Summary of posteriors:

<!-- ```{r}
# targets::tar_read(fit0_summary_lm_example)
withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit0_summary_lm_example))
```
 -->

### 2025-03-26 docs/gmm_cr.qmd
Initial commit for tree-allo

---
title: "Checking gmm model"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the required packages

```{r}
library(dplyr)
library(readr)
library(targets)
library(stringr)
library(cmdstanr)
library(tidyverse)
library(reshape2)
library(bayesplot)
library(cmdstanr) 
library(posterior) 
```

# Data preparation
```{r}
tallo_wd_200_df <- read_csv("data/tallo_wd_200.csv")
tallo_wd_200_df
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

options(width = 100)
gmm_cr <- tar_read(fit_nlr_mcmc_gmm_cr)
gmm_cr

gmm2_cr <- tar_read(fit_nlr_mcmc_gmm2_cr)
gmm2_cr

gmm_beta <- gmm_cr$summary(variable = "beta")
gmm2_beta <- gmm2_cr$summary(variable = "beta")

gmm_beta
gmm2_beta

summary(gmm_beta)
summary(gmm2_beta)

# Filter beta parameters with small ESS
small_ess_gmm_beta <- gmm_beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

small_ess_gmm1_beta <- gmm1_beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk)
small_ess_gmm1_beta

high_rhat_gmm1_beta <- gmm1_beta |>
   filter(rhat > 1.1) |> 
   arrange(rhat)

high_rhat_gmm1_beta
```

# Species with low ESS of gmm_cr and gmm2_cr model

```{r}
beta1 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]"))
beta1
summary(beta1)

small_ess_beta1 <- beta1 |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

small_ess_beta1

beta2 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]"))

beta2
summary(beta2)

beta3 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]"))

beta3
summary(beta3)
```

```{r}
gamma_trace <- gmm_cr$draws("gamma")
mcmc_trace(gamma_trace)

mcmc_trace(gamma_trace) + ggtitle("Trace Plot for gmm_cr Gamma")

gamma2_trace <- gmm2_cr$draws("gamma")
mcmc_trace(gamma2_trace) + ggtitle("Trace Plot for gmm2_cr Gamma")

mcmc_trace(gamma_trace) + ggtitle("Trace plot of gmm1_cr")
```

### 2025-03-26 docs/gmm_d_h.qmd
Initial commit for tree-allo

---
title: "Example for gMM DBH-H"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the required packages

```{r}
library(cmdstanr)
library(posterior)
library(tidyverse)
library(targets)
```

# Load the results and data
```{r}
gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
gmm_h

stan_nlr_data_h <- tar_read(stan_data_nlr_h)
str(stan_nlr_data_h)

log_y <- stan_nlr_data_h$log_y
x <- stan_nlr_data_h$x
sp <- as.factor(stan_nlr_data_h$jj)

# Extract summaries
summary_gmm_h <- gmm_h$summary()
head(summary_gmm_h)

beta <- summary_gmm_h |> filter(variable == "beta[,1]")

# Extract mean values for log(a), b and k
log_a2 <- summary_gmm_h |> filter(variable == "beta[,1]") |> pull(mean)
b2 <- summary_gmm_h |> filter(variable == "beta[,2]") |> pull(mean)
k1 <- summary_gmm_h |> filter(variable == "beta[,3]") |> pull(mean)

# Compute the fitted values for H using the formula gMM
fitted_log_y_h <- log_a2 + b2 * log(x) - log(k1 + x^b2)

# Create a data frame for ggplot
data_h <- data.frame(H = exp(log_y), DBH = x, fitted_H = exp(fitted_log_y_h), sp = sp)

# Check the structure of the data frame
head(data_h)
str(data_h)

# Plot the scatter plot of observed H vs DBH and add the fitted line for each species

h <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = DBH, y = fitted_H, color = sp, group = sp), show.legend = FALSE) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

h
# Save the plot
ggsave("figs/scatter_plot_H_vs_DBH.png", plot = h, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
```

```{r}
gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
gmm_h

stan_nlr_data_h <- tar_read(stan_data_nlr_h)
str(stan_nlr_data_h)

log_y <- stan_nlr_data_h$log_y
x <- stan_nlr_data_h$x
sp <- as.factor(stan_nlr_data_h$jj)
# sp <- as.factor(stan_nlr_data_h$u)

# Extract beta from result
beta <- gmm_h$summary(variable = "beta")
beta

log_a2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(mean)
k1 <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(mean)

# Verify the lengths of the extracted parameters
print(length(log_a2))
print(length(b2))
print(length(k1))
length(sp)

# Compute the fitted values for H using the formula gMM
fitted_log_y_h <- log_a2 + b2 * log(x) - log(k1 + x^b2)

# Create a data frame for ggplot
data_h <- data.frame(H = exp(log_y), DBH = x, fitted_H = exp(fitted_log_y_h), sp = sp)

# Check the structure of the data frame
head(data_h)
str(data_h)
summary(data_h)

# Plot the scatter plot of observed H vs DBH and add the fitted line for each species

gmm_h_plot <- ggplot(data_h, aes(x = DBH, y = H), group = sp, color = sp) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = DBH, y = fitted_H, color = sp, group = sp), show.legend = FALSE) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

gmm_h_plot

# Save the plot
ggsave("figs/plot_H_vs_DBH.png", plot = gmm_h_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
```

---
title: "Example for gMM DBH-H"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(ggplot2)
library(dplyr)
library(stringr)
```

# Data preparation
```{r}
# Load MCMC results and Stan data
gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
stan_nlr_data_h <- tar_read(stan_data_nlr_h)

# Extract relevant data
log_y <- stan_nlr_data_h$log_y
x <- stan_nlr_data_h$x
sp <- as.factor(stan_nlr_data_h$jj)

# Extract beta parameters
beta <- gmm_h$summary(variable = "beta")

# Ensure dplyr and stringr are loaded and then extract the mean values of beta parameters
log_a2 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,1\\]")) %>% pull(mean)
b2 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,2\\]")) %>% pull(mean)
k1 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,3\\]")) %>% pull(mean)

# Verify the lengths of the extracted parameters
print(length(log_a2))
print(length(b2))
print(length(k1))
print(length(sp))

# Compute fitted values for H using the gMM formula
fitted_log_y_h <- sapply(1:length(log_a2), function(i) {
  log_a2[i] + b2[i] * log(x) - log(k1[i] + x^b2[i])
})

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a2 <- mean(log_a2)
community_b2 <- mean(b2)
community_k1 <- mean(k1)
community_fitted_log_y_h <- community_log_a2 + community_b2 * log(x) - log(community_k1 + x^community_b2)
community_fitted_H <- exp(community_fitted_log_y_h)

# Create a data frame for plotting
data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a2)) {
  data_h[[paste0("fitted_H_sp_", i)]] <- exp(fitted_log_y_h[, i])
}

# Plot observed H vs DBH with fitted lines for each species and community-level line
gmm_h_plot <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(aes(y = community_fitted_H), color = "white", linewidth = 1.2) +
  geom_line(aes(y = rowMeans(fitted_log_y_h)), color = "blue", alpha = 0.3) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "black"))

# Save the plot
ggsave("figs/beta_plot_H_vs_DBH.png", plot = gmm_h_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

gmm_h_plot

# Plot observed H vs DBH with fitted lines for each species and community-level line
gmm_h_plot1 <- ggplot(data_h, aes(x = DBH, y = H, group = sp)) +
  geom_point(alpha = 0.5, color = "grey") +
  # geom_line(aes(y = community_fitted_H), color = "#4682B4", linewidth = 0.3) +
  geom_line(aes(y = rowMeans(fitted_log_y_h), group = sp), color = "#4682B4", alpha = 0.3) + # SteelBlue color
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))

# Save the plot
ggsave("figs/beta1_plot_H_vs_DBH.png", plot = gmm_h_plot1, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(gmm_h_plot1)

# Melt data_h to long format for easier ggplot handling
library(reshape2)
data_long <- melt(data_h, id.vars = c("H", "DBH", "sp"), measure.vars = grep("fitted_H_sp_", names(data_h), value = TRUE))

# Plot observed H vs DBH with fitted lines for each species and community-level line
gmm_h_plot2 <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_H), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "black"))

# Save the plot
ggsave("figs/beta2_plot_H_vs_DBH.png", plot = gmm_h_plot2, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(gmm_h_plot2)

# Plot observed H vs DBH with fitted lines for each species and community-level line
gmm_h_plot3 <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_H), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Save the plot
ggsave("figs/beta3_plot_H_vs_DBH.png", plot = gmm_h_plot3, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(gmm_h_plot3)

### 2025-03-26 docs/gmm_h.qmd
Initial commit for tree-allo

---
title: "Example for gMM DBH-H"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(targets)
library(ggplot2)
library(dplyr)
library(stringr)
library(reshape2)
```

# Data preparation
```{r}
# Load MCMC results and Stan data
gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
stan_nlr_data_h <- tar_read(stan_data_nlr_h)
str(stan_nlr_data_h)
```

# Extract relevant data
```{r}
log_y <- stan_nlr_data_h$log_y
x <- stan_nlr_data_h$x
sp <- as.factor(stan_nlr_data_h$jj)

# Extract beta parameters
beta <- gmm_h$summary(variable = "beta")

log_a2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(mean)
k1 <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(mean)

# Verify the lengths of the extracted parameters
print(length(log_a2))
print(length(b2))
print(length(k1))
print(length(sp))

# Compute fitted values for H using the gMM formula
fitted_log_y_h <- sapply(1:length(log_a2), function(i) {
  log_a2[i] + b2[i] * log(x) - log(k1[i] + x^b2[i])
})

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a2 <- mean(log_a2)
community_b2 <- mean(b2)
community_k1 <- mean(k1)
community_fitted_log_y_h <- community_log_a2 + community_b2 * log(x) - log(community_k1 + x^community_b2)
community_fitted_H <- exp(community_fitted_log_y_h)

# Create a data frame for plotting
data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)
head(daha_h)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a2)) {
  data_h[[paste0("fitted_H_sp_", i)]] <- exp(fitted_log_y_h[, i])
}

data_long <- melt(data_h, id.vars = c("H", "DBH", "sp"), measure.vars = grep("fitted_H_sp_", names(data_h), value = TRUE))

```

# Plot observed H vs DBH with fitted lines for each species and community-level line
```{r}
gmm_h_plot4 <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_H), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of H vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Save the plot
ggsave("figs/beta4_plot_H_vs_DBH.png", plot = gmm_h_plot4, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(gmm_h_plot4)

```

```{r}
library(dplyr)
library(readr)
library(targets)
library(stringr)
library(cmdstanr)
library(bayesplot)
```

```{r}
gmm_h <- tar_read(fit_nlr7_mcmc_gmm_h)
gamma_trace <- gmm_h$draws("gamma")
mcmc_trace(gamma_trace)

tallo_wd_200_df <- read_csv("data/tallo_wd_200.csv")
tallo_wd_200_df
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

beta <- gmm_h$summary(variable = "beta")
str(beta)
beta
```

# Covariance between beta parameters
```{r}
# Extract beta samples
beta_samples <- gmm_h$draws(variables = "beta", format = "draws_matrix")
str(beta_samples)
beta_samples

# Extract beta1, beta2, and beta3 separately
beta1_samples <- beta_samples[, grep("beta\\[\\d+,1\\]", colnames(beta_samples))]
beta2_samples <- beta_samples[, grep("beta\\[\\d+,2\\]", colnames(beta_samples))]
beta3_samples <- beta_samples[, grep("beta\\[\\d+,3\\]", colnames(beta_samples))]

str(beta1_samples)
str(beta2_samples)
str(beta3_samples)

# Combine the beta samples into a single data frame
combined_beta_samples <- cbind(beta1_samples, beta2_samples, beta3_samples)
str(combined_beta_samples)

# Convert the combined samples to a matrix
combined_beta_matrix <- as.matrix(combined_beta_samples)

# Compute the covariance matrix for the combined beta parameters
cov_matrix <- cov(combined_beta_matrix)
head(cov_matrix)

# Print the covariance matrix
str(cov_matrix)

# Calculate summary statistics for the covariance matrix (excluding diagonal)
cov_values <- cov_matrix[lower.tri(cov_matrix)]
summary(cov_values)

# Calculate the correlation matrix
cor_matrix <- cov2cor(cov_matrix)

# Summary statistics for the correlation values (excluding diagonal)
cor_values <- cor_matrix[lower.tri(cor_matrix, diag = FALSE)]
summary(cor_values)

threshold <- 0.7
strong_correlations <- sum(abs(cor_values) > threshold)

# Calculate the percentage of strong correlations
total_pairs <- length(cor_values)
percentage_strong <- (strong_correlations / total_pairs) * 100
percentage_strong

```

```{r}
summary(cor_values)
hist(cov_values)

hist(cov_values, breaks=50, main="Histogram of Covariance Values", xlab="Covariance Values", col="blue")
```

### 2025-03-26 docs/gmm_nou_h.qmd
Initial commit for tree-allo

---
title: "Checking gmm_nou_h model"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the required packages

```{r}
library(dplyr)
library(readr)
library(targets)
library(stringr)
library(cmdstanr)
library(tidyverse)
library(reshape2)
```

# Data preparation
```{r}
tallo_wd_200_df <- read_csv("data/tallo_wd_200.csv")
tallo_wd_200_df
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)
gmm_nou_h
# beta <- gmm_nou_h$summary(variable = "beta")
# str(beta)
# beta

```

# Species with low ESS of gmm_nou_h model
```{r}
# # Filter beta parameters with small ess_bulk
# small_ess_beta <- beta |>
#   filter(ess_bulk < 300) |>
#   arrange(ess_bulk) |>
#   mutate(species_id = as.integer(str_extract(variable, "\\d+")))

# # Print the filtered beta parameters
# print(small_ess_beta)

# # Count the number of individuals per species
# species_count <- tallo_wd_200_df |>
#  filter(!is.na(wd)) |>
#   group_by(sp) |>
#   summarise(num_individuals = n())

# # Print the species count
# print(species_count)

# # Create a mapping from species_id to species name and number of individuals
# species_map <- tallo_wd_200_df |>
#   filter(!is.na(wd)) |>
#   mutate(species_id = as.integer(as.factor(sp))) |>
#   group_by(species_id, sp) |>
#   summarise(num_individuals = n(), .groups = 'drop')

# species_map

# # Join the filtered beta parameters with the species count
# small_ess_species <- small_ess_beta |>
#   left_join(species_map, by = "species_id")

# # Print the result
# options(width = 200)
# print(small_ess_species)

# # Extract the unique species and their number of individuals
# unique_species_count <- small_ess_species |>
#   select(sp, num_individuals, species_id) |>
#   distinct()

# # Print the unique species and their number of individuals
# print(unique_species_count, n = Inf)
```

# Covariance between beta parameters
```{r}
# Extract beta samples
beta_samples <- gmm_nou_h$draws(variables = "beta", format = "draws_matrix")
str(beta_samples)
beta_samples

# Extract beta1, beta2, and beta3 separately
beta1_samples <- beta_samples[, grep("beta\\[\\d+,1\\]", colnames(beta_samples))]
beta2_samples <- beta_samples[, grep("beta\\[\\d+,2\\]", colnames(beta_samples))]
beta3_samples <- beta_samples[, grep("beta\\[\\d+,3\\]", colnames(beta_samples))]

str(beta1_samples)
str(beta2_samples)
str(beta3_samples)

# Combine the beta samples into a single data frame
combined_beta_samples <- cbind(beta1_samples, beta2_samples, beta3_samples)
str(combined_beta_samples)

# Convert the combined samples to a matrix
combined_beta_matrix <- as.matrix(combined_beta_samples)

# Compute the covariance matrix for the combined beta parameters
cov_matrix <- cov(combined_beta_matrix)
head(cov_matrix)

# Print the covariance matrix
str(cov_matrix)

# Calculate summary statistics for the covariance matrix (excluding diagonal)
cov_values <- cov_matrix[lower.tri(cov_matrix)]
summary(cov_values)

# Calculate the correlation matrix
cor_matrix <- cov2cor(cov_matrix)

# Summary statistics for the correlation values (excluding diagonal)
cor_values <- cor_matrix[lower.tri(cor_matrix, diag = FALSE)]
summary(cor_values)
head(cor_values)

threshold <- 0.7
strong_correlations <- sum(abs(cor_values) > threshold)

# Calculate the percentage of strong correlations
total_pairs <- length(cor_values)
percentage_strong <- (strong_correlations / total_pairs) * 100
percentage_strong

```

```{r}
summary(cor_values)
hist(cov_values)

hist(cov_values, breaks=50, main="Histogram of Covariance Values", xlab="Covariance Values", col="blue")
```

Key findings: 
The summary of covariance values shows a wide range, with a minimum of -1142.68 and a maximum of 1070.34. The mean is close to zero (0.5014), indicating that while there are some extreme values, the average covariance is relatively low. The median is also close to zero (0.0002), suggesting that the majority of covariances are small.

# Species with low ESS of gmm_nou_h model

```{r}
log_y <- stan_data_nlr_h$log_y
x <- stan_data_nlr_h$x
sp <- as.factor(stan_data_nlr_h$jj)

beta <- gmm_nou_h$summary(variable = "beta")

log_a2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(mean)
b2 <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(mean)
k1 <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(mean)

```

```{r}
# Filter beta parameters with small ESS
small_ess_beta <- beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

# Print the filtered beta parameters
print(small_ess_beta)

# Count the number of individuals per species
species_count <- tallo_wd_200_df |>
 filter(!is.na(wd)) |>
  group_by(sp) |>
  summarise(num_individuals = n())

# Print the species count
print(species_count)

# Create a mapping from species_id to species name and number of individuals
species_map <- tallo_wd_200_df |>
  filter(!is.na(wd)) |>
  mutate(species_id = as.integer(as.factor(sp))) |>
  group_by(species_id, sp) |>
  summarise(num_individuals = n(), .groups = 'drop')

species_map

# Join the filtered beta parameters with the species count
small_ess_species <- small_ess_beta |>
  left_join(species_map, by = "species_id")

# Print the result
options(width = 200)
print(small_ess_species)

# Extract the unique species and their number of individuals
unique_species_count <- small_ess_species |>
  select(sp, num_individuals, species_id) |>
  distinct()

# Print the unique species and their number of individuals
print(unique_species_count, n = Inf)
unique_species_count <- as.data.frame(unique_species_count)
# Save as a CSV file
write.csv(unique_species_count, "data/gmm_nou_h_lowESS.csv", row.names = FALSE)

small_ess_species

# Extract parameters only for species with low ESS
low_ess_species_ids <- small_ess_species$species_id
log_a2_low_ess <- log_a2[low_ess_species_ids]
b2_low_ess <- b2[low_ess_species_ids]
k1_low_ess <- k1[low_ess_species_ids]

# Compute fitted values for H using the gMM formula for species with low ESS
fitted_log_y_h_low_ess <- sapply(1:length(log_a2_low_ess), function(i) {
  log_a2_low_ess[i] + b2_low_ess[i] * log(x) - log(k1_low_ess[i] + x^b2_low_ess[i])
})

# Create a data frame for plotting
data_h_low_ess <- data.frame(H = exp(log_y), DBH = x, sp = sp)
head(data_h_low_ess)
summary(data_h_low_ess)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a2_low_ess)) {
  data_h_low_ess[[paste0("fitted_H_sp_", low_ess_species_ids[i])]] <- exp(fitted_log_y_h_low_ess[, i])
}

# Transform data for plotting
data_long_low_ess <- melt(data_h_low_ess, id.vars = c("H", "DBH", "sp"), measure.vars = grep("fitted_H_sp_", names(data_h_low_ess), value = TRUE))

head(data_long_low_ess)
summary(data_long_low_ess)

# write.csv(as.data.frame(data_long_low_ess), "data/data_long_low_ess.csv", row.names = FALSE)
```

# Make the separate plots for each species

```{r}
# data_long_low_ess <- read_csv("data/data_long_low_ess.csv")
# data_long_low_ess

# Update the variable names to include species names
data_long_low_ess$variable <- sub("fitted_H_sp_", "", data_long_low_ess$variable)

head(data_long_low_ess)
summary(data_long_low_ess)
```

```{r}
# Function to plot and save each species
plot_species <- function(species_name) {
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_plot <- ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = data_h_low_ess %>% filter(sp == species_name), aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "H") +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank())

  # Save the plot
  ggsave(paste0("figs/gmm_nou_h", species_name, "_H_vs_DBH.png"), plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

  # Display the plot
  print(species_plot)
}

library(ggplot2)
library(patchwork)
library(dplyr)

# Generate plots for each species and store in a list
plots <- lapply(unique(data_long_low_ess$variable), plot_species)

# Combine all plots into a single plot layout
combined_plot <- wrap_plots(plots, ncol = 4)  # 4 columns, 6 rows

# Save the combined plot to a PDF file
ggsave("figs/gmm_nou_h/combined_figures.pdf", combined_plot, width = 16, height = 24, units = "in", dpi = 300)

```

```{r}
# Define the plotting function
plot_species <- function(species_name) {
  # Filter data for the selected species
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_dbh <- species_data$DBH

  # Determine the range of DBH for the selected species
  dbh_min <- min(species_dbh)
  dbh_max <- min(max(species_dbh), 50)  # Cap the max at 100 for better visualization

  # Adjust the x-axis limits
  x_limits <- c(dbh_min, dbh_max)

  species_plot <- ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = data_h_low_ess %>% filter(sp == as.numeric(species_name)), aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "H") +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank()) +
    scale_x_continuous(limits = x_limits)

  # Save the plot
  ggsave(paste0("figs/gmm_nou_h_", species_name, "_H_vs_DBH.png"), plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

  # Display the plot
  print(species_plot)
}

# Example usage for all species
unique_species <- unique(data_long_low_ess$variable)
plots <- lapply(unique_species, plot_species)

# Combine all plots into a single plot layout
combined_plot <- wrap_plots(plots, ncol = 4)  # 4 columns

# Save the combined plot to a PDF file
ggsave("figs/gmm_nou_h/combined_figures.pdf", combined_plot, width = 16, height = 24, units = "in", dpi = 300)
```

```{r}
# Add species-specific fitted values to the data frame
for (i in 1:length(log_a2_low_ess)) {
  species_name <- small_ess_species %>% filter(species_id == low_ess_species_ids[i]) %>% pull(sp) %>% unique()
  species_name <- species_name[1]  # Ensure we only get one unique species name
  data_h_low_ess[[paste0("fitted_H_sp_", species_name)]] <- exp(fitted_log_y_h_low_ess[, i])
}

# Transform data for plotting
data_long_low_ess <- data_h_low_ess %>%
  pivot_longer(cols = starts_with("fitted_H_sp_"), names_to = "variable", values_to = "value")

# Update the variable names to include species names
data_long_low_ess$variable <- sub("fitted_H_sp_", "", data_long_low_ess$variable)

head(data_long_low_ess)

plot_species <- function(species_name) {
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_plot <- ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = data_h_low_ess %>% filter(sp == species_name), aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "H") +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank())

  # Save the plot
  ggsave(paste0("figs/gmm_nou_h", species_name, "_H_vs_DBH.png"), plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

  # Display the plot
  print(species_plot)
}
# Generate plots for each species and store in a list
plots <- lapply(unique(data_long_low_ess$variable), plot_species)

# Combine all plots into a single plot layout
combined_plot <- wrap_plots(plots, ncol = 4)  # 4 columns, 6 rows

# Save the combined plot to a PDF file
ggsave("figs/gmm_nou_h/combined_figures.pdf", combined_plot, width = 16, height = 24, units = "in", dpi = 300)

```

```{r}
# Function to plot and save each species
plot_species <- function(species_name) {
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_points <- data_h_low_ess %>% filter(sp == species_name)
  species_plot <- ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = data_h_low_ess %>% filter(sp == species_name), aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "Fitted H") +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank())

  # Save the plot
  ggsave(paste0("figs/gmm/", gsub(" ", "_", species_name), "2_H_vs_DBH.png"), plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

  # Display the plot
  print(species_plot)
}

# Loop through each species and plot
unique_species_names <- unique(data_long_low_ess$variable)
for (species in unique_species_names) {
  plot_species(species)
}
```

```{r}
# Function to plot and save each species
plot_species <- function(species_name) {
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_points <- data_h_low_ess %>% filter(sp == species_name)
  species_plot <- ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = species_points, aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "Fitted H") +
    theme_minimal() +
    theme(legend.position = "bottom", legend.title = element_blank())

  # Save the plot
  ggsave(paste0("figs/", gsub(" ", "_", species_name), "1_H_vs_DBH.png"), plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

  # Display the plot
  print(species_plot)
}

# Loop through each species and plot
unique_species_names <- unique(data_long_low_ess$variable)
for (species in unique_species_names) {
  plot_species(species)
}
```

```{r}
library(cowplot)

# Function to create each species plot
create_species_plot <- function(species_name) {
  species_data <- data_long_low_ess %>% filter(variable == species_name)
  species_points <- data_h_low_ess %>% filter(sp == species_name)
  ggplot(species_data, aes(x = DBH, y = value, color = variable)) +
    geom_line() +
    geom_point(data = species_points, aes(x = DBH, y = H), alpha = 0.5, color = "grey") +
    labs(title = paste("Fitted Values for", species_name, "with ESS < 300"), x = "DBH", y = "Fitted H") +
    theme_minimal() +
    theme(legend.position = "none", legend.title = element_blank())
}

# Create a list of plots
unique_species_names <- unique(data_long_low_ess$variable)
plots <- lapply(unique_species_names, create_species_plot)

# Define the number of plots per page and total pages
plots_per_page <- 3 * 8

# Open a PDF device to save the plots
pdf("figs/species_fitted_plots.pdf", width = 11, height = 8.5)

# Loop through the pages and plot
for (i in seq(1, length(plots), by = plots_per_page)) {
  page_plots <- plots[i:min(i + plots_per_page - 1, length(plots))]
  plot_grid(plotlist = page_plots, ncol = 3, nrow = 8)
}

# Close the PDF device
dev.off()
```

# Arrange figures

```{r}
library(ggplot2)
library(patchwork)
library(grid)
library(png)
library(jpeg)

# Directory containing the images
image_dir <- "figs/gmm1"
image_files <- list.files(image_dir, pattern = "\\.(png|PNG|jpg|jpeg|JPG|JPEG|gif|GIF|svg|SVG)$", full.names = TRUE)

# Check if any image files are found
if (length(image_files) == 0) {
  stop("No image files found in the directory.")
}

# Create a list to store the plots
plots <- list()

# Create a ggplot object for each image and add it to the list
for (image_file in image_files) {
  img <- NULL
  if (grepl("\\.png$|\\.PNG$", image_file)) {
    img <- try(readPNG(image_file), silent = TRUE)
  } else if (grepl("\\.jpg$|\\.jpeg$|\\.JPG$|\\.JPEG$", image_file)) {
    img <- try(readJPEG(image_file), silent = TRUE)
  } else {
    next # Skip unsupported file types
  }

  # Check if the image was successfully read
  if (inherits(img, "try-error")) {
    message("Failed to read image: ", image_file)
    next
  }

  plot <- ggplot() +
    annotation_custom(rasterGrob(img, width = unit(1, "npc"), height = unit(1, "npc")), -Inf, Inf, -Inf, Inf) +
    ggtitle(basename(image_file)) +
    theme_void() +
    theme(plot.title = element_text(size = 10))

  plots <- append(plots, list(plot))
}

# Check if any plots were created
if (length(plots) == 0) {
  stop("No plots created. Please check your image files.")
}

# Combine all plots into a single plot layout
combined_plot <- wrap_plots(plots, ncol = 4) # 4 columns

# Save the combined plot to a PDF file
ggsave("figs/combined_figures.pdf", combined_plot, width = 16, height = 24, dpi = 300 ) # Adjust width and height as needed
```

# Check the covariance

```{r}
# Load the required library
library(cmdstanr)
library(posterior)

gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)

str(gmm_nou_h)

# Extract the posterior samples directly as a draws_array
posterior_samples <- as_draws_array(gmm_nou_h$draws())

# Number of posterior samples
num_samples <- dim(posterior_samples)[1]
head(num_samples)

# Assuming K is known
K <- 3  # Change this to the actual value of K

# Extract L_Sigma_beta samples
# The variable names for L_Sigma_beta may be different; adapt as necessary
L_Sigma_beta_samples <- posterior_samples[, , grep("^L_Sigma_beta", dimnames(posterior_samples)$variable)]

# Initialize array to store covariance matrices
cov_matrices <- array(NA, dim = c(K, K, num_samples))

# Calculate covariance matrix for each sample
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrices[,,i] <- L_Sigma_beta %*% t(L_Sigma_beta)
}

# Calculate the mean covariance matrix
mean_cov_matrix <- apply(cov_matrices, c(1, 2), mean)

# Print the mean covariance matrix
print(mean_cov_matrix)

# Calculate the correlations
mean_cor_matrix <- cov2cor(mean_cov_matrix)

# Print the mean correlation matrix
print(mean_cor_matrix)
```

### 2025-03-26 docs/h.qmd
Initial commit for tree-allo

---
title: "Plot the best model to predict H from DBH: Weibull"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
```

# Read in data
```{r}
# Read in the data
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Extract relevant data
log_y <- stan_data_nlr_h$log_y
x <- stan_data_nlr_h$x
sp <- as.factor(stan_data_nlr_h$jj)

# Read in the model summary
wb1_h_summary <- tar_read(fit_nlr_summary_weibull1_h)
wb1_h_summary

# Extract the actual beta parameters
beta <- wb1_h_summary |> filter(grepl("beta\\[", variable))
beta

log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50)
b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)

# Compute fitted values for H using the Weibull formula
fitted_log_y_h <- sapply(1:length(log_a), function(i) {
  log_a[i] + log(1 - exp(-b[i] * (x ^ k[i])))
})

fitted_h <- exp(fitted_log_y_h)

# Extract community-level parameters directly
gamma <- wb1_h_summary |> filter(grepl("gamma", variable))

# Correctly map community parameters
community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

# Apply exponential transformation to community-level parameters
community_log_a <- exp(community_log_a_hat)
community_b <- exp(community_b_hat) 
community_k <- exp(community_k_hat)

community_fitted_log_y_h <- community_log_a + log(1 - exp(-community_b * (x^community_k)))
community_fitted_H <- exp(community_fitted_log_y_h)

# Create a data frame for plotting
data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)
data_h$community_fitted_H <- community_fitted_H
head(data_h)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_h[[paste0("fitted_H_sp_", i)]] <- fitted_h[, i]
}

# Convert data to long format
data_long <- reshape2::melt(data_h, id.vars = c("H", "DBH", "sp", "community_fitted_H"), measure.vars = grep("fitted_H_sp_", names(data_h), value = TRUE))

head(data_long)

# Function to create DBH sequence for each species based on actual range
create_dbh_sequence <- function(species, x, length_out = 100) {
  min_dbh <- min(x[sp == species])
  max_dbh <- max(x[sp == species])
  seq(min_dbh, max_dbh, length.out = length_out)
}

# Calculate fitted H values for each species
fitted_h_list <- lapply(1:length(log_a), \(i) {
  dbh_seq <- create_dbh_sequence(i, x)
  fitted_log_y <- log_a[i] + log(1 - exp(-b[i] * (dbh_seq ^ k[i])))
  fitted_h <- exp(fitted_log_y)

  data.frame(
    DBH = dbh_seq,
    value = fitted_h,
    variable = paste0("fitted_H_sp_", i)
  )
})

# Combine all species data frames into one
data_long <- do.call(rbind, fitted_h_list)
head(data_long)

# Original data for points
data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)
head(data_h)
```

# Plot
```{r}
wb1_h_plot1 <- ggplot(data_h, aes(x = DBH, y = H)) +
  # geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#72b6e3", alpha = 0.5) +
  geom_line(aes(y = community_fitted_H), color = "#0f92e9", linewidth = 1.5) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

wb1_h_plot1
```

### 2025-03-26 docs/h_cr_bb.qmd
Initial commit for tree-allo

---
title: "Overlaps species"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
```

# Read the CSV file
```{r}
hcr <- read_csv("data/bb_allometry_re.csv")
summary(hcr)
head(hcr, n=100)
summary(hcr)
```

# Convert sp to a factor
```{r}
hcr$sp <- factor(hcr$sp)
hcr$life <- factor(hcr$life)
#Construct the histogram 
par(mfrow=c(2,1))
hist(hcr$tree_height)
hist(hcr$crown_width_we) # positive skewed
```

#log transform
```{r}
hcr$crown_width_we.ln <- log(hcr$crown_width_we)
hcr$tree_height.ln <- log(hcr$tree_height)
summary(hcr)
```

#check
```{r}
par(mfrow=c(2,1))
hist(hcr$tree_height.ln) #look better
hist(hcr$crown_width_we.ln) 
```

#plot the relationship of interest
```{r}
plot(crown_width_we.ln ~ tree_height.ln, data = hcr)
plot(hcr$tree_height.ln ,hcr$crown_width_we.ln ,col= c("red", "blue", "green")[as.numeric(hcr$sp)])
```

# Remove rows with -Inf and NA in crown_width_we.ln, tree_height.ln
```{r}
hcr <- hcr[is.finite(hcr$crown_width_we.ln) & !is.na(hcr$crown_width_we.ln), ]
hcr <- hcr[is.finite(hcr$tree_height.ln) & !is.na(hcr$tree_height.ln), ]

# Check the summary again
summary(hcr$crown_width_we.ln)
summary(hcr$tree_height.ln)
```

# Fit a linear model
```{r}
hcr.rg <- lm(crown_width_we.ln ~ tree_height.ln, data = hcr)
summary(hcr.rg)
plot(hcr.rg)
# Create a line graph using ggplot2 with regression lines for each species
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = sp, colour = sp, shape = sp)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = sp)) +
  theme(legend.position = "none")  # Optional: To remove legend for shapes due to the large number of levels
```

# Relationship between Tree Height and Crown Width by Life
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life, shape = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = life)) +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored and shaped by life",
       color = "Life", shape = "Life")
```

#separate plots for each level of the life variable
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = life)) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored by life",
       color = "Life")
```

# 
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = sp)) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored by life",
       color = "Life")
```

# facetting the plot by the life variable and showing regression lines for each life category, colored by the life variable.
```{r}
bb_plot <- ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = interaction(sp, life), colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = interaction(sp, life))) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each species within different life categories",
       color = "Life")

# Save the plot
ggsave("figs/bb.png", plot = bb_plot, width = 10, height = 6)

```

```{r}

bb <- ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = interaction(sp, life), colour = life)) +
  geom_point(size = 7, alpha = 0.7) +  # Increase point size and add transparency
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 0.7, aes(group = interaction(sp, life))) +  # Reduce line size
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each species within different life categories",
       color = "Life")

# Save the plot
ggsave("figs/bb_2.png", plot = bb, width = 10, height = 6)
```

### 2025-03-26 docs/h_new.qmd
Initial commit for tree-allo

---
title: "Comparing H models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
options(width = 200)
```

# Read in data

```{r}
tallo_wd_df0 <- tar_read(tallo_wd_df0)
tallo_wd_df0
tallo_wd_df0 <- tallo_wd_df0 |> 
                dplyr::filter(!is.na(h)) |>
                dplyr::group_by(sp) |>
                dplyr::filter(n() > 20) |>
                dplyr::ungroup()

ang_df <- tallo_wd_df0 |> filter(division == "Angiosperm")
ang_df

gym_df <- tallo_wd_df0 |> filter(division == "Gymnosperm")
gym_df

```

```{r}
tallo_wd_df_h <- tar_read(tallo_wd_df_h)
tallo_wd_df_h

ang_df <- tallo_wd_df_h |> filter(division == "Angiosperm")
ang_df

gym_df <- tallo_wd_df_h |> filter(division == "Gymnosperm")
gym_df

tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h

ang_data <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")

gym_data <- tallo_reduced_lr_df_ang_h |> filter(division == "Gymnosperm")

tallo_reduced_lr_df_gym_h

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

tallo_reduced_nlr_df_gym_h

```

# Power-law

## Ang

```{r}
# Data
fit_lr_nou_data_ang_h <- tar_read(fit_lr_nou_data_ang_h)
str(fit_lr_nou_data_ang_h)

pl_ang_h <- tar_read(fit_lr_nou_summary_pl_ang_h)
pl_ang_h

gamma_ang_pl <- pl_ang_h |> filter(grepl("gamma", variable))
gamma_ang_pl

community_log_a_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_ang <- fit_lr_nou_data_ang_h$log_x[,2] |> range()

log_x_seq_ang <- seq(log_x_range_ang[1], log_x_range_ang[2], length.out = 100)

fitted_log_y_seq_ang <- community_log_a_ang_pl + community_b_ang_pl * log_x_seq_ang

y_pred_ang <- exp(fitted_log_y_seq_ang)

df_ang_pl <- tibble(y = y_pred_ang, x = exp(log_x_seq_ang))

df_ang_pl

ggplot() +
  geom_bin2d(data = ang_df, aes(x = dbh, y = h), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_pl, aes(x = x, y = y))
```

## Gym

```{r}
# Data
fit_lr_nou_data_gym_h <- tar_read(fit_lr_nou_data_gym_h)
str(fit_lr_nou_data_gym_h)

pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
pl_gym_h

gamma_gym_pl <- pl_gym_h |> filter(grepl("gamma", variable))
gamma_gym_pl

community_log_a_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_gym <- fit_lr_nou_data_gym_h$log_x[,2] |> range()

log_x_seq_gym <- seq(log_x_range_gym[1], log_x_range_gym[2], length.out = 100)

fitted_log_y_seq_gym <- community_log_a_gym_pl + community_b_gym_pl * log_x_seq_gym

y_pred_gym <- exp(fitted_log_y_seq_gym)

df_gym_pl <- tibble(y = y_pred_gym, x = exp(log_x_seq_gym))

df_gym_pl

ggplot() +
  geom_bin2d(data = gym_df, aes(x = dbh, y = h), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_pl, aes(x = x, y = y))
```

# gMM

## Ang

```{r}
# Data
fit_nlr_nou_data_ang_h <- tar_read(fit_nlr_nou_data_ang_h)
str(fit_nlr_nou_data_ang_h)

x_nlr_range_ang <- range(fit_nlr_nou_data_ang_h$x)
x_nlr_seq_ang <- seq(x_nlr_range_ang[1], x_nlr_range_ang[2], length.out = 100)

gmm_ang_h <- tar_read(fit_nlr_nou_summary_gmm_ang_h)
gmm_ang_h

gamma_ang_gmm <- gmm_ang_h |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
gamma_ang_gmm

community_log_a_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[1]") |> pull(q50)

community_b_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[2]") |> pull(q50)

community_k_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_h_gmm_seq <- community_log_a_ang_gmm + community_b_ang_gmm * log(x_nlr_seq_ang) - log(community_k_ang_gmm + x_nlr_seq_ang^community_b_ang_gmm)

y_ang_gmm_pred <- exp(community_fitted_log_y_h_gmm_seq)

df_ang_gmm <- tibble(y = y_ang_gmm_pred, x = x_nlr_seq_ang)
df_ang_gmm

ggplot() +
  geom_bin2d(data = ang_df, aes(x = dbh, y = h), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_gmm, aes(x = x, y = y))
```

## Gym

```{r}
# Data
fit_nlr_nou_data_gym_h <- tar_read(fit_nlr_nou_data_gym_h)
str(fit_nlr_nou_data_gym_h)

x_nlr_range_gym <- range(fit_nlr_nou_data_gym_h$x)
x_nlr_seq_gym <- seq(x_nlr_range_gym[1], x_nlr_range_gym[2], length.out = 100)

gmm_gym_h <- tar_read(fit_nlr_nou_summary_gmm_gym_h)
gmm_gym_h

gamma_gym_gmm <- gmm_gym_h |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
gamma_gym_gmm

community_log_a_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_h_gym_gmm_seq <- community_log_a_gym_gmm + community_b_gym_gmm * log(x_nlr_seq_gym) - log(community_k_gym_gmm + x_nlr_seq_gym^community_b_gym_gmm)

y_gym_gmm_pred <- exp(community_fitted_log_y_h_gym_gmm_seq)

df_gym_gmm <- tibble(y = y_gym_gmm_pred, x = x_nlr_seq_gym)
df_gym_gmm

ggplot() +
  geom_bin2d(data = gym_df, aes(x = dbh, y = h), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_gmm, aes(x = x, y = y))

```

# Weibull

## Ang

```{r}
# Data
fit_nlr_nou_data_ang_h <- tar_read(fit_nlr_nou_data_ang_h)
str(fit_nlr_nou_data_ang_h)

x_nlr_range_ang <- range(fit_nlr_nou_data_ang_h$x)
x_nlr_seq_ang <- seq(x_nlr_range_ang[1], x_nlr_range_ang[2], length.out = 100)

# weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_biome_ang_h)
# weibull_ang_h

weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
weibull_ang_h

gamma_ang_wb <- weibull_ang_h |> filter(grepl("gamma", variable))
gamma_ang_wb

community_log_a_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[1]") |> pull(q50)

community_b_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[2]") |> pull(q50)

community_k_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_h_wb_ang_seq <- community_log_a_ang_wb + log(1 - exp(-community_b_ang_wb * (x_nlr_seq_ang ^ community_k_ang_wb)))

y_ang_wb_pred <- exp(community_fitted_log_y_h_wb_ang_seq)

df_ang_wb <- tibble(y = y_ang_wb_pred, x = x_nlr_seq_ang)

df_ang_wb

ggplot() +
  geom_bin2d(data = ang_df, aes(x = dbh, y = h), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_wb, aes(x = x, y = y))

```

## Gym

```{r}
# Data
fit_nlr_nou_data_gym_h <- tar_read(fit_nlr_nou_data_gym_h)
str(fit_nlr_nou_data_gym_h)

x_nlr_range_gym <- range(fit_nlr_nou_data_gym_h$x)
x_nlr_seq_gym <- seq(x_nlr_range_gym[1], x_nlr_range_gym[2], length.out = 100)

# Weibull model summary
weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
weibull_gym_h

# Extract community-level parameters
gamma_gym_wb <- weibull_gym_h |> filter(grepl("gamma", variable))
gamma_gym_wb

community_log_a_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[3]") |> pull(q50)

# Calculate fitted values for Weibull model
community_fitted_log_y_h_wb_gym_seq <- community_log_a_gym_wb + log(1 - exp(-community_b_gym_wb * (x_nlr_seq_gym ^ community_k_gym_wb)))

y_gym_wb_pred <- exp(community_fitted_log_y_h_wb_gym_seq)

df_gym_wb <- tibble(y = y_gym_wb_pred, x = x_nlr_seq_gym)

df_gym_wb

# Plot with gym data
ggplot() +
  geom_bin2d(data = gym_df, aes(x = dbh, y = h), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_wb, aes(x = x, y = y))

```

# Plotting with density-based coloring

## Ang

```{r}
dens <- with(ang_df, kde2d(log10(dbh), log10(h), n = 300))
ix <- findInterval(log10(ang_df$dbh), dens$x)
iy <- findInterval(log10(ang_df$h), dens$y)
ang_df$density <- dens$z[cbind(ix, iy)]
head(ang_df)

p_ang_den <- ggplot(data = ang_df, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_ang_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(
       x = "DBH_ang (cm)",
       y = "H_ang (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

p_ang_den

ggsave("figs/p_ang_den.png", plot = p_ang_den, width = 8, height = 6, dpi = 300, bg = "white")

```

## Gym

```{r}
dens <- with(gym_df, kde2d(log10(dbh), log10(h), n = 300))
ix <- findInterval(log10(gym_df$dbh), dens$x)
iy <- findInterval(log10(gym_df$h), dens$y)
gym_df$density <- dens$z[cbind(ix, iy)]

p_gym_den <- ggplot(data = gym_df, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "H_gym (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

p_gym_den

ggsave("figs/p_gym_den.png", plot = p_gym_den, width = 8, height = 6, dpi = 300, bg = "white")

```

# Plotting with density-based coloring non-log scale

## Ang

```{r}
dens <- with(ang_df, kde2d(dbh, h, n = 300))
ix <- findInterval(ang_df$dbh, dens$x)
iy <- findInterval(ang_df$h, dens$y)
ang_df$density <- dens$z[cbind(ix, iy)]
head(ang_df)

p_ang_den_non <- ggplot(data = ang_df, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_ang_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(
       x = "DBH_ang (cm)",
       y = "H_ang (m)") +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

p_ang_den_non

ggsave("figs/p_ang_non_den.png", plot = p_ang_den_non, width = 8, height = 6, dpi = 300, bg = "white")

```

## Gym

```{r}
dens <- with(gym_df, kde2d(dbh, h, n = 300))
ix <- findInterval(gym_df$dbh, dens$x)
iy <- findInterval(gym_df$h, dens$y)
gym_df$density <- dens$z[cbind(ix, iy)]

p_gym_den_non <- ggplot(data = gym_df, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "H_gym (m)") +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

p_gym_den_non

ggsave("figs/p_gym_non_den.png", plot = p_gym_den_non, width = 8, height = 6, dpi = 300, bg = "white")
```

### 2025-03-26 docs/h_re.qmd
Initial commit for tree-allo

---
title: "Comparing H models (updated)"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
```

```{r}
pl_summary <- tar_read(fit_lr2_summary_power_law1_h)
```

```{r}
##pl
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(q50)
```

We need sd and mean to unscale DBH and H.
```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
# tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
  filter(!is.na(dbh)) |>
  filter(!is.na(h)) |>
  filter(!is.na(cr)) |>
  filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

# we need sd and mean to unscale the values
log_dbh_sd <- sd(data2$log_dbh)
log_dbh_mean <- mean(data2$log_dbh)

log_h_sd <- sd(data2$log_h)
log_h_mean <- mean(data2$log_h)

stan_data_lr_h <- tar_read(stan_data_lr_h)
str(stan_data_lr_h)
```

We only need several data points for x to plot the line.
We can generate a sequence of x-values from the range of the scaled x-values.
We also need to unscale them to get the actual values.

```{r}
scaled_x_range <- stan_data_lr_h$log_x[, 2] |> range()
log_x_seq <- seq(scaled_x_range[1], scaled_x_range[2], length.out = 100)
log_xx <- (log_x_seq * log_dbh_sd) + log_dbh_mean
```

Because the predicted y-values are also scaled (according to `generate_stan_data`), we need to unscale them to get the actual values.
You didn't have to scale y-values for the model in the first place, but it's okay.

```{r}
# use scaled values to predict scaled values
log_y_pred_scaled <- community_log_a_pl + community_b_pl * log_x_seq
# unscaled
log_y_pred <- log_y_pred_scaled * log_h_sd + log_h_mean
y_pred <- exp(log_y_pred)

# use unscaled values to draw lines.
df <- tibble(y = y_pred, x = exp(log_xx))
```

For the desnity plot, you can use other functions but here is a simple example.
```{r}
ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df, aes(x = x, y = y))
```

### 2025-03-26 docs/h_up.qmd
Initial commit for tree-allo

---
title: "Comparing H models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
```

# Read in data

```{r}
# Read in the data (pl)
stan_data_lr_h <- tar_read(stan_data_lr_h)
str(stan_data_lr_h)

# Read in the data (gMM, wb)
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Extract relevant data from stan_data_nlr_h (log_y, x, jj)
x_nlr <- stan_data_nlr_h$x

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_h)
wb_summary <- tar_read(fit_nlr_summary_weibull1_h)
pl_summary <- tar_read(fit_lr2_summary_power_law1_h)

gmm_summary
wb_summary
pl_summary
```

# Extract community-level parameters of each function

## Power-law

### Extract scaling parameters
```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(h)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
log_dbh_sd <- sd(data2$log_dbh)
log_dbh_mean <- mean(data2$log_dbh)

log_h_sd <- sd(data2$log_h)
log_h_mean <- mean(data2$log_h)
```

### Computing Fitted Values
```{r}
pl_summary
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(q50)
community_log_a_pl
community_b_pl

## Computing Fitted Values
scaled_x_range <- stan_data_lr_h$log_x[, 2] |> range()
log_x_seq <- seq(scaled_x_range[1], scaled_x_range[2], length.out = 100)
log_xx <- (log_x_seq * log_dbh_sd) + log_dbh_mean

# use scaled values to predict scaled values
log_y_pred_scaled <- community_log_a_pl + community_b_pl * log_x_seq
# unscaled
log_y_pred <- log_y_pred_scaled * log_h_sd + log_h_mean
community_fitted_h_pl <- exp(log_y_pred)

df_pl <- tibble(y = community_fitted_h_pl, x = exp(log_xx))
head(df_pl)
```

## gMM
```{r}
gmm_summary
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
gamma_gmm
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(q50)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm) 
community_k_gmm <- exp(community_k_hat_gmm)

community_log_a_gmm
community_b_gmm
community_k_gmm

## Computing Fitted Values
x_nlr_range <- range(x_nlr)
x_nlr_seq <- seq(x_nlr_range[1], x_nlr_range[2], length.out = 100)

# Compute the fitted values using the sequence of x-values
community_fitted_log_y_h_gmm_seq <- community_log_a_gmm + community_b_gmm * log(x_nlr_seq) - log(community_k_gmm + x_nlr_seq^community_b_gmm)
community_fitted_h_gmm_seq <- exp(community_fitted_log_y_h_gmm_seq)

df_gmm <- tibble(y = community_fitted_h_gmm_seq, x = x_nlr_seq)
head(df_gmm)
```

## Weibull
```{r}
wb_summary
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
gamma_wb
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(q50)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(q50)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(q50)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb) 
community_k_wb <- exp(community_k_hat_wb)

community_log_a_wb
community_b_wb
community_b_wb

# Compute the fitted values using the sequence of x-values
community_fitted_log_y_h_wb_seq <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr_seq ^ community_k_wb)))
community_fitted_h_wb_seq <- exp(community_fitted_log_y_h_wb_seq)

df_wb <- tibble(y = community_fitted_h_wb_seq, x = x_nlr_seq)
head(df_wb)
```

# Plot

## Power-law
```{r}
head(data2)
head(df_pl)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "#fc8d62" )

```

## gMM
```{r}
head(df_gmm)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "#80b1d3" )

```

## Weibull

```{r}
head(df_wb)

ggplot() +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_wb, aes(x = x, y = y), color = "#1b9e77" )

```

## 3 functions in 1 panel: Log-scale
```{r}
combined_plot <- ggplot(data2) +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y, color = "Power-law"), size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y, color = "gMM"), size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y, color = "Weibull"), size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "#fc8d62", "gMM" = "#80b1d3", "Weibull" = "#1b9e77"))

combined_plot
```

## 3 functions in 1 panel: Non-log scale
```{r}
combined_plot_non <- ggplot(data2) +
  geom_bin2d(data = data2, aes(x = dbh, y = h), bins = 200)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_pl, aes(x = x, y = y, color = "Power-law"), size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y, color = "gMM"), size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y, color = "Weibull"), size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "blue", "gMM" = "green", "Weibull" = "red"))

combined_plot_non
```

# Plotting with density-based coloring

## Log scale
```{r}
dens <- with(data2, kde2d(log10(dbh), log10(h), n = 300))
ix <- findInterval(log10(data2$dbh), dens$x)
iy <- findInterval(log10(data2$h), dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

combined_plot <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16))

combined_plot
ggsave("h_plot.png", plot = combined_plot, width = 8, height = 6, dpi = 300, bg= "white")

combined_plot <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 20),   
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16),    
        axis.text.y = element_text(size = 16),
        legend.position = "none") +  # Remove default legend

    # Manually add lines and descriptions for the three models in the upper-right corner
    annotate("segment", x = 100, xend = 130, y = 1.0, yend = 1.0, color = "#FF5733", size = 1) +  # Power-Law line
    annotate("text", x = 135, y = 1.0, label = "Power-Law", size = 5, hjust = 0) +  # Power-Law text

    annotate("segment", x = 100, xend = 130, y = 0.8, yend = 0.8, color = "#2ECC71", size = 1) +  # GMM line
    annotate("text", x = 135, y = 0.8, label = "gMM", size = 5, hjust = 0) +  # GMM text

    annotate("segment", x = 100, xend = 130, y = 0.6, yend = 0.6, color = "#0f92e9", size = 1) +  # Weibull line
    annotate("text", x = 135, y = 0.6, label = "Weibull", size = 5, hjust = 0)  # Weibull text

ggsave("h_plot_anno.png", plot = combined_plot, width = 8, height = 6, dpi = 300, bg = "white")

```

```{r}
library(viridis)

combined_plot1 <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.6) +  # Reduced alpha for points
  scale_color_viridis(option = "viridis", name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1.2) +  # Increased line width
  geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1.2) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1.2) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

combined_plot1
```

## Non-log scale

```{r}
dens <- with(data2, kde2d(dbh, h, n = 300))
ix <- findInterval(data2$dbh, dens$x)
iy <- findInterval(data2$h, dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

# Create the plot without log scales
combined_plot_non <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "blue", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "green", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  theme_minimal()

combined_plot_non
```

```{r}
library(viridis)

combined_plot_non1 <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.6) +  # Reduced alpha for points
  scale_color_viridis(option = "viridis", name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1.2) +  # Increased line width
  geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1.2) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1.2) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

combined_plot_non1
```

# Plotting with density-based coloring2

## Log scale
```{r}
dens <- with(data2, kde2d(log10(dbh), log10(h), n = 300))
ix <- findInterval(log10(data2$dbh), dens$x)
iy <- findInterval(log10(data2$h), dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

combined_plot1 <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.6) +  # Reduced alpha for points
  scale_color_viridis(option = "viridis", name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1.2) +  # Increased line width
  geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1.2) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1.2) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

combined_plot1
```

## Non-log scale

```{r}
dens <- with(data2, kde2d(dbh, h, n = 300))
ix <- findInterval(data2$dbh, dens$x)
iy <- findInterval(data2$h, dens$y)
data2$density <- dens$z[cbind(ix, iy)]
head(data2)

# Create the plot without log scales
combined_plot_non <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "blue", size = 0.8) +
  geom_line(data = df_gmm, aes(x = x, y = y), color = "green", size = 0.8) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 0.8) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "H (m)") +
  theme_minimal()

combined_plot_non
```

```{r}
library(viridis)

combined_plot_non1 <- ggplot(data = data2, aes(x = dbh, y = h, color = density)) +
  geom_point(size = 0.5, alpha = 0.6) +  # Reduced alpha for points
  scale_color_viridis(option = "viridis", name = "Density") +
  geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1.2) +  # Increased line width
  geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1.2) +
  geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1.2) +
  labs(
       x = "DBH (cm)",
       y = "H (m)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

combined_plot_non1
```

### 2025-03-26 docs/lm_theta_wd.qmd
Initial commit for tree-allo

---
title: "The relationships between tree architecture and wood density"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the packages
```{r}
library(targets)
library(targets)
library(tidyverse)
library(stringr)
```

# log(h) ~ log(dbh)
```{r}
p1 <- tar_read(fit_lm_summary_lm)
p1
p1$variable <- as.factor(p1$variable)
ln_a1 <- p1 %>%
  filter(str_detect(variable, "theta\\[.*,1\\]")) %>%
  pull(mean)

# Filter rows corresponding to slopes
b1 <- p1 %>%
  filter(str_detect(variable, "theta\\[.*,2\\]")) %>%
  pull(mean)

summary(ln_a1)
summary(b1)

```
```{r}
par(mfrow = c(1, 2))
hist(ln_a1)
hist(b1)
```

```{r}
alpha1_1 <- p1[p1$variable == "alpha[1]", ]
alpha2_1 <- p1[p1$variable == "alpha[2]", ]
beta1_1 <- p1[p1$variable == "beta[1]", ]
beta2_1 <- p1[p1$variable == "beta[2]", ]
alpha1_1
alpha2_1
beta1_1
beta2_1
```

r$> beta1_1
# A tibble: 1 × 15
  variable     mean     sd    mad    q2.5      q5     q25      q50     q75    q95  q97.5  rhat ess_bulk ess_tail .join_data
  <fct>       <dbl>  <dbl>  <dbl>   <dbl>   <dbl>   <dbl>    <dbl>   <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta[1]  -0.00135 0.0137 0.0139 -0.0267 -0.0233 -0.0109 -0.00185 0.00789 0.0211 0.0274  1.00     381.     342.         NA

r$> beta2_1
# A tibble: 1 × 15
  variable    mean     sd    mad    q2.5      q5     q25     q50    q75    q95  q97.5  rhat ess_bulk ess_tail .join_data
  <fct>      <dbl>  <dbl>  <dbl>   <dbl>   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta[2]  0.00899 0.0122 0.0125 -0.0166 -0.0109 0.00117 0.00857 0.0178 0.0283 0.0329  1.01     533.     411.         NA

- Key findings:

## log(cr) ~ log(dbh)
```{r}
p2 <- tar_read(fit_lm_summary_lm)
p2
p2$variable <- as.factor(p2$variable)
ln_a2 <- p2 %>%
  filter(str_detect(variable, "theta\\[.*,1\\]")) %>%
  pull(mean)
beta <- p2 %>% filter(str_detect(variable, "beta\\[.*,1\\")) %>% pull(mean)
# Filter rows corresponding to slopes
b2 <- p2 %>%
  filter(str_detect(variable, "theta\\[.*,2\\]")) %>%
  pull(mean)

summary(ln_a2)
summary(b2)

```
```{r}
par(mfrow = c(1, 2))
hist(ln_a2)
hist(b2)
```

```{r}
alpha1_2 <- p2[p2$variable == "alpha[1]", ]
alpha2_2 <- p2[p2$variable == "alpha[2]", ]
beta1_2 <- p2[p2$variable == "beta[1]", ]
beta2_2 <- p2[p2$variable == "beta[2]", ]
alpha1_2
alpha2_2
beta1_2
beta2_2
```

- beta1_2

# A tibble: 1 × 15
  variable    mean     sd    mad    q2.5      q5        q25     q50    q75    q95  q97.5  rhat ess_bulk ess_tail
  <fct>      <dbl>  <dbl>  <dbl>   <dbl>   <dbl>      <dbl>   <dbl>  <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>
1 beta[1]  0.00868 0.0116 0.0129 -0.0147 -0.0113 -0.0000615 0.00900 0.0174 0.0269 0.0293  1.02     355.     299.
# ℹ 1 more variable: .join_data <dbl>

- beta2_2
# A tibble: 1 × 15
  variable     mean      sd     mad    q2.5      q5      q25      q50     q75     q95  q97.5  rhat ess_bulk
  <fct>       <dbl>   <dbl>   <dbl>   <dbl>   <dbl>    <dbl>    <dbl>   <dbl>   <dbl>  <dbl> <dbl>    <dbl>
1 beta[2]  -0.00322 0.00771 0.00831 -0.0166 -0.0151 -0.00891 -0.00350 0.00233 0.00931 0.0120  1.01     374.
# ℹ 2 more variables: ess_tail <dbl>, .join_data <dbl>

## log(dbh) ~ log(h) + log(cr)

```{r}
p3 <- tar_read(fit_lm_summary_lm)
p3
p3$variable <- as.factor(p3$variable)
ln_a3 <- p3 %>%
  filter(str_detect(variable, "theta\\[.*,1\\]")) %>%
  pull(mean)

# Filter rows corresponding to slopes
b3 <- p3 %>%
  filter(str_detect(variable, "theta\\[.*,2\\]")) %>%
  pull(mean)

summary(ln_a3)
summary(b3)

```
```{r}
par(mfrow = c(1, 2))
hist(ln_a3)
hist(b3)
```

```{r}
alpha1_3 <- p3[p3$variable == "alpha[1]", ]
alpha2_3 <- p3[p3$variable == "alpha[2]", ]
beta1_3 <- p3[p3$variable == "beta[1]", ]
beta2_3 <- p3[p3$variable == "beta[2]", ]
alpha1_3
alpha2_3
beta1_3
beta2_3
```

r$> beta1_3
# A tibble: 1 × 15
  variable     mean     sd    mad    q2.5      q5     q25      q50     q75    q95  q97.5  rhat ess_bulk ess_tail .join_data
  <chr>       <dbl>  <dbl>  <dbl>   <dbl>   <dbl>   <dbl>    <dbl>   <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta[1]  -0.00338 0.0124 0.0122 -0.0279 -0.0241 -0.0114 -0.00350 0.00498 0.0171 0.0194  1.00     499.     374.         NA

r$> beta2_3
# A tibble: 1 × 15
  variable   mean     sd    mad    q2.5       q5     q25     q50    q75    q95  q97.5  rhat ess_bulk ess_tail .join_data
  <chr>     <dbl>  <dbl>  <dbl>   <dbl>    <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta[2]  0.0101 0.0109 0.0110 -0.0113 -0.00809 0.00289 0.00999 0.0176 0.0283 0.0322  1.00     434.     375.         NA

```{r}
p <- tar_read(fit_lm_summary_lm)
p
p$variable <- as.factor(p$variable)
library(dplyr)

# Extract theta1[i,1]
theta1_1 <- p %>% 
  filter(grepl("^theta1\\[.*,1\\]$", variable))

# Extract theta1[i,2]
theta1_2 <- p %>% 
  filter(grepl("^theta1\\[.*,2\\]$", variable))

# Extract alpha1[i]
alpha1 <- p %>% 
  filter(grepl("^alpha1\\[.*\\]$", variable))

# Extract alpha2[i]
alpha2 <- p %>% 
  filter(grepl("^alpha2\\[.*\\]$", variable))

# Extract alpha3[i]
alpha3 <- p %>% 
  filter(grepl("^alpha3\\[.*\\]$", variable))

# Extract beta1[i]
beta1 <- p %>% 
  filter(grepl("^beta1\\[.*\\]$", variable))

# Extract beta2[i]
beta2 <- p %>% 
  filter(grepl("^beta2\\[.*\\]$", variable))

# Extract beta3[i]
beta3 <- p %>% 
  filter(grepl("^beta3\\[.*\\]$", variable))

# Display the extracted data
theta1_1
theta1_2
alpha1
alpha2
alpha3
beta1
beta2
beta3
```

r$> beta1
# A tibble: 2 × 15
  variable     mean      sd     mad    q2.5      q5     q25      q50      q75     q95  q97.5  rhat ess_bulk ess_tail .join_data
  <fct>       <dbl>   <dbl>   <dbl>   <dbl>   <dbl>   <dbl>    <dbl>    <dbl>   <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta1[1] -0.00199 0.0122  0.0126  -0.0266 -0.0212 -0.0106 -0.00209 0.00637  0.0189  0.0230  1.00     447.     392.         NA
2 beta1[2] -0.00537 0.00813 0.00789 -0.0204 -0.0173 -0.0107 -0.00598 0.000108 0.00846 0.0109  1.02     382.     313.         NA

r$> beta2
# A tibble: 2 × 15
  variable     mean      sd     mad    q2.5      q5       q25      q50     q75     q95  q97.5  rhat ess_bulk ess_tail .join_data
  <fct>       <dbl>   <dbl>   <dbl>   <dbl>   <dbl>     <dbl>    <dbl>   <dbl>   <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta2[1]  0.00871 0.0116  0.0125  -0.0141 -0.0100 -0.000392  0.00850 0.0166  0.0276  0.0321  1.01     382.     341.         NA
2 beta2[2] -0.00340 0.00804 0.00762 -0.0194 -0.0168 -0.00841  -0.00323 0.00188 0.00911 0.0119  1.00     418.     372.         NA

r$> beta3
# A tibble: 2 × 15
  variable     mean      sd     mad     q2.5       q5      q25      q50     q75    q95  q97.5  rhat ess_bulk ess_tail .join_data
  <fct>       <dbl>   <dbl>   <dbl>    <dbl>    <dbl>    <dbl>    <dbl>   <dbl>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>      <dbl>
1 beta3[1] -0.00293 0.0125  0.0119  -0.0269  -0.0237  -0.0108  -0.00309 0.00505 0.0174 0.0209  1.01     383.     335.         NA
2 beta3[2]  0.00925 0.00844 0.00868 -0.00628 -0.00471  0.00330  0.00951 0.0148  0.0226 0.0251  1.00     430.     304.         NA

# Using multivariate normal distribution

## log(cr) ~ log(dbh)
```{r}
p2 <- tar_read(fit_lm_summary_lm)
```

r$> p2
# A tibble: 230,753 × 15
   variable         mean       sd      mad       q2.5         q5        q25        q50        q75        q95      q97.5  rhat ess_bulk ess_tail .join_data
   <fct>           <dbl>    <dbl>    <dbl>      <dbl>      <dbl>      <dbl>      <dbl>      <dbl>      <dbl>      <dbl> <dbl>    <dbl>    <dbl>      <dbl>
 1 lp__       72673.     35.6     35.9     72602.     72612.     72649.     72673.     72698.     72731.     72738.      1.01     128.     274.         NA
 2 Omega[1,1]     1       0        0           1          1          1          1          1          1          1      NA         NA       NA          NA
 3 tau[1]         0.249   0.00386  0.00377     0.241      0.242      0.246      0.249      0.251      0.255      0.256   1.00     428.     304.         NA
 4 gamma[1,1]    -0.198   0.00689  0.00744    -0.211     -0.210     -0.203     -0.199     -0.193     -0.187     -0.185   1.01     390.     266.         NA
 5 beta[1,1]      0.368   0.0393   0.0422      0.292      0.307      0.340      0.368      0.396      0.433      0.445   1.00     377.     282.         NA
 6 beta[2,1]      0.236   0.00428  0.00462     0.228      0.229      0.233      0.236      0.239      0.243      0.245   1.02     437.     124.         NA
 7 beta[3,1]      0.268   0.0303   0.0298      0.209      0.223      0.247      0.265      0.287      0.318      0.324   1.03     517.     215.         NA
 8 beta[4,1]      0.0544  0.00357  0.00379     0.0486     0.0489     0.0519     0.0545     0.0570     0.0598     0.0609  1.04     345.     313.         NA
 9 beta[5,1]      0.251   0.0185   0.0193      0.217      0.221      0.239      0.252      0.265      0.281      0.285   1.01     465.     229.         NA
10 beta[6,1]      0.262   0.0120   0.0124      0.239      0.243      0.253      0.262      0.269      0.281      0.286   1.01     413.     268.         NA
# ℹ 230,743 more rows
# ℹ Use `print(n = ...)` to see more rows

- Key findings:
For the effect of wood density on log(cr) ~ log(dbh) relationship:
The mean value of gamma is approximately -0.198 with a standard deviation of around 0.00689. The 95% credible interval ranges from approximately -0.211 to -0.185.
Since the 95% credible interval does not include zero (-0.198 ± 0.013), it suggests that gamma is significantly different from zero.
Suggesting that the effect of wood density (gamma) on the relationship between log(cr) and log(dbh) is statistically significant.

```{r}
beta <- p2 %>%
  filter(str_detect(variable, "beta\\[.*,1\\]")) %>%
  pull(mean)
hist(beta)
```

```{r}
# Calculate the number of beta values where the 95% CI includes zero
num_beta_ci_include_zero <- sum(p2$q2.5 <= 0 & p2$q97.5 >= 0 & str_detect(p2$variable, "beta\\[.*,1\\]"))

# Print the result
print(num_beta_ci_include_zero)
```

[1] 514
- Key findings:
For the interaction between log(dbh) and wood density.
There are 514 beta values where the 95% CI includes zero.
This suggests that the interaction between log(dbh) and wood density (beta) does not significantly impact the relationship between log(cr) and log(dbh).

### 2025-03-26 docs/lmer.qmd
Initial commit for tree-allo

---
title: "Relationships of tree architecture and wood density"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the packages
```{r}
library(readr)
library(tidyverse)
library(here)
library(lme4)
library(e1071)
library(car)
library(sjPlot)
library(gridExtra)
```

# Data preparation
```{r}
ov <- read_csv(here("data/overlap_20.csv"))
ov <- na.omit(ov)
ov$sp <- as.factor(ov$sp)
summary(ov)
```

- Checking the data
```{r}
# Histograms

hist_dbh <- ggplot(ov, aes(x = dbh)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of DBH", x = "DBH", y = "Frequency")

hist_h <- ggplot(ov, aes(x = h)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Height", x = "Height", y = "Frequency")

hist_cr <- ggplot(ov, aes(x = cr)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of CR", x = "CR", y = "Frequency")

# Q-Q plots
qq_dbh <- ggplot(ov, aes(sample = dbh)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for DBH", x = "Theoretical Quantiles", y = "Sample Quantiles")

qq_h <- ggplot(ov, aes(sample = h)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for Height", x = "Theoretical Quantiles", y = "Sample Quantiles")

qq_cr <- ggplot(ov, aes(sample = cr)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for CR", x = "Theoretical Quantiles", y = "Sample Quantiles")

# # Combine plots
# combined_plots <- gridExtra::grid.arrange(hist_dbh, hist_h, hist_cr, qq_dbh, qq_h, qq_cr, ncol = 3)

# # Save the combined plots
# ggsave(
#   filename = here("figs/lmer/distribution_plots.png"),
#   plot = combined_plots,
#   width = 12,
#   height = 8
# )
```

- Transforming data
```{r}
ov$dbh_ln <- log(ov$dbh)
ov$h_ln <- log(ov$h)
ov$cr_ln <- log(ov$cr)
summary(ov)
```

- Checking data
```{r}
# Calculate mean values
mean_h <- mean(ov$h.ln)
mean_dbh <- mean(ov$dbh.ln)
mean_cr <- mean(ov$cr.ln)
mean_wd <- mean(ov$wd)

hist_h <- ggplot(ov, aes(x = h.ln)) +
  geom_histogram(fill = "lightblue", color = "black") +
  geom_vline(xintercept = mean_h, color = "red", linetype = "dashed", size = 1) +  # Add mean line
  labs(title = "Histogram of Tree Height (ln)", x = "Height (ln)", y = "Frequency")

hist_dbh <- ggplot(ov, aes(x = dbh.ln)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  geom_vline(xintercept = mean_dbh, color = "red", linetype = "dashed", size = 1) +  # Add mean line
  labs(title = "Histogram of DBH (ln)", x = "DBH (ln)", y = "Frequency")

hist_cr <- ggplot(ov, aes(x = cr.ln)) +
  geom_histogram(fill = "lightpink", color = "black") +
  geom_vline(xintercept = mean_cr, color = "red", linetype = "dashed", size = 1) +  # Add mean line
  labs(title = "Histogram of CR (ln)", x = "CR (ln)", y = "Frequency")

hist_wd <- ggplot(ov, aes(x = wd)) +
  geom_histogram(fill = "orange", color = "black") +
  geom_vline(xintercept = mean_wd, color = "red", linetype = "dashed", size = 1) +  # Add mean line
  labs(title = "Histogram of WD", x = "WD", y = "Frequency")

combined_plot <- gridExtra::grid.arrange(hist_h, hist_dbh, hist_cr, hist_wd, ncol = 4)

ggsave(
  filename = here("figs/lmer/combined_histograms_ln_with_mean.png"),
  plot = combined_plot,
  width = 12,
  height = 6
)
```

- Normality qq-plot

```{r}
# Create Q-Q plots
qq_h.ln <- ggplot(ov, aes(sample = h.ln)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for Tree Height (ln)", x = "Theoretical Quantiles", y = "Sample Quantiles")

qq_dbh.ln <- ggplot(ov, aes(sample = dbh.ln)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for DBH (ln)", x = "Theoretical Quantiles", y = "Sample Quantiles")

qq_cr.ln <- ggplot(ov, aes(sample = cr.ln)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for CR (ln)", x = "Theoretical Quantiles", y = "Sample Quantiles")

qq_wd <- ggplot(ov, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot for WD", x = "Theoretical Quantiles", y = "Sample Quantiles")

# Combine Q-Q plots into one grid
combined_qq_plots <- gridExtra::grid.arrange(qq_h.ln, qq_dbh.ln, qq_cr.ln, qq_wd, ncol = 2)

# Save the combined Q-Q plots
ggsave(
  filename = here("figs/lmer/combined_qq_ln_plots.png"),
  plot = combined_qq_plots,
  width = 12,
  height = 8
)
```

# Fitting lme4::lmer

```{r}
dhcmod1 <- lme4::lmer(dbh_ln ~ (h_ln + cr_ln) + (1 + (h_ln + cr_ln) | sp), data = ov2)
hdmod1 <- lme4::lmer(h_ln ~ dbh_ln + (1 + dbh_ln | sp), data = ov2)
cdmod1 <- lme4::lmer(cr_ln ~ dbh_ln + (1 + dbh_ln | sp), data = ov2)
```

- Issues:
> dhcmod1 <- lme4::lmer(dbh.ln ~ h.ln + cr.ln + (1 + h.ln + cr.ln | sp), data = ov)
Warning message:
In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  Model failed to converge with max|grad| = 0.00437759 (tol = 0.002, component 1)

```{r}
print(dhmod1)
print(dcmod1)
print(hdmod1)
print(cdmod1)
print(dhcmod1)
```

- Comparison of models (estimating DBH)

```{r}
# Compare AIC and BIC
aic_bic <- data.frame(
  Model = c("dhmod1", "dcmod1", "dhcmod1"),
  AIC = c(AIC(dhmod1), AIC(dcmod1), AIC(dhcmod1)),
  BIC = c(BIC(dhmod1), BIC(dcmod1), BIC(dhcmod1))
)
print(aic_bic)
```

- Key findings:
> print(aic_bic)
    Model       AIC      BIC
1  dhmod1 219829.56 219891.3
2  dcmod1 191989.70 192051.5
3 dhcmod1  88852.92  88955.9

The model with the lowest AIC and BIC values is generally considered the best-fit.
In this comparison, dhcmod1 has the lowest AIC and BIC values.
Therefore, dhcmod1 is the better-fitting model (estimation of DBH) according to these criteria.
These results imply that tree crown size might play a crucial role in accurately estimating DBH.

## Check the residual distribution

- Extract the residuals
```{r}
residuals_dhmod1 <- resid(dhmod1)
residuals_dcmod1 <- resid(dcmod1)
residuals_hdmod1 <- resid(hdmod1)
residuals_cdmod1 <- resid(cdmod1)
residuals_dhcmod1 <- resid(dhcmod1)
```

- Extract the  fitted values
```{r}
fitted_dhmod1 <- fitted(dhmod1)
fitted_dcmod1 <- fitted(dcmod1)
fitted_hdmod1 <- fitted(hdmod1)
fitted_cdmod1 <- fitted(cdmod1)
fitted_dhcmod1 <- fitted(dhcmod1)
```

- Histograming residuals
```{r}
hist_dhmod1_plot <- ggplot(data.frame(residuals = residuals_dhmod1), aes(x = residuals)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Residuals for dhmod1") +
  theme_minimal()

hist_dcmod1_plot <- ggplot(data.frame(residuals = residuals_dcmod1), aes(x = residuals)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Residuals for dcmod1") +
  theme_minimal()

hist_hdmod1_plot <- ggplot(data.frame(residuals = residuals_hdmod1), aes(x = residuals)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Residuals for hdmod1") +
  theme_minimal()

hist_cdmod1_plot <- ggplot(data.frame(residuals = residuals_cdmod1), aes(x = residuals)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Residuals for cdmod1") +
  theme_minimal()

hist_dhcmod1_plot <- ggplot(data.frame(residuals = residuals_dhcmod1), aes(x = residuals)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Residuals for dhcmod1") +
  theme_minimal()

# Combine all plots with a common layout
combined_plots_re <- gridExtra::grid.arrange(
  hist_dhmod1_plot, hist_dcmod1_plot, hist_hdmod1_plot, hist_cdmod1_plot, hist_dhcmod1_plot,
  nrow = 2
)
# Save the combined plots
ggsave(
  filenam = here("figs/lmer/combined_residual_ln_histograms.png"),
  plot = combined_plots_re,
  width = 20,
  height = 8)
```

- Redidual dependent plot

```{r}
# Function to create residual dependent plots
create_residual_dependent_plot <- function(residuals, model_name) {
  ggplot(data.frame(Residuals = residuals), aes(x = seq_along(residuals), y = residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    labs(title = paste("Residuals for", model_name), x = "Observation", y = "Residuals")
}

# List of model names and their corresponding residuals
model_residuals <- list(
  "dhmod1" = residuals_dhmod1,
  "dcmod1" = residuals_dcmod1,
  "hdmod1" = residuals_hdmod1,
  "cdmod1" = residuals_cdmod1,
  "dhcmod1" = residuals_dhcmod1
)

# Create individual residual dependent plots
plots <- lapply(names(model_residuals), function(model_name) {
  create_residual_dependent_plot(model_residuals[[model_name]], model_name)
})

# Arrange all plots in a grid
combined_plots_rd <- gridExtra::grid.arrange(grobs = plots, ncol = 3)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/combined_residual_dependent_plots.png"),
  plot = combined_plots_rd,
  width = 16,
  height = 12
)
```

- Scale - Location plot
```{r}
# Function to create Scale-Location (S-L) Plot
create_sl_plot <- function(residuals, fitted_values, model_name) {
  # Calculate square root of absolute residuals
  sqrt_abs_residuals <- sqrt(abs(residuals))

  # Create data frame
  df <- data.frame(Fitted_Values = fitted_values, Sqrt_Abs_Residuals = sqrt_abs_residuals)

  # Plot
  ggplot(df, aes(x = Fitted_Values, y = Sqrt_Abs_Residuals)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(title = paste("Scale-Location Plot for", model_name),
         x = "Fitted Values",
         y = "Square Root of Absolute Residuals") +
    theme_minimal()
}

# List of model names and their corresponding residuals and fitted values
model_data <- list(
  "dhmod1" = list(residuals = residuals_dhmod1, fitted_values = fitted_dhmod1),
  "dcmod1" = list(residuals = residuals_dcmod1, fitted_values = fitted_dcmod1),
  "hdmod1" = list(residuals = residuals_hdmod1, fitted_values = fitted_hdmod1),
  "cdmod1" = list(residuals = residuals_cdmod1, fitted_values = fitted_cdmod1),
  "dhcmod1" = list(residuals = residuals_dhcmod1, fitted_values = fitted_dhcmod1)
)

# Create individual S-L plots
plots <- lapply(names(model_data), function(model_name) {
  create_sl_plot(model_data[[model_name]]$residuals, model_data[[model_name]]$fitted_values, model_name)
})

# Arrange all plots in a grid
combined_plots_sl <- gridExtra::grid.arrange(grobs = plots, ncol = 3)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/combined_sl_plots.png"),
  plot = combined_plots_sl,
  width = 16,
  height = 12
)
```

```{r}
# Create diagnostic plots for each model
diag_plots_dcmod1 <- plot_model(dcmod1, type = "diag")
diag_plots_hdmod1 <- plot_model(hdmod1, type = "diag")
diag_plots_cdmod1 <- plot_model(cdmod1, type = "diag")
diag_plots_dhcmod1 <- plot_model(dhcmod1, type = "diag")

# Combine plots into a single grid
combined_plots <- gridExtra::grid.arrange(
  diag_plots_dcmod1[[1]],
  diag_plots_hdmod1[[1]],
  diag_plots_cdmod1[[1]],
  diag_plots_dhcmod1[[1]],
  nrow = 2
)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/diagnostic_plots_combined.png"),
  plot = combined_plots,
  width = 12,
  height = 8
)
```

- Normality qq-plot of residuals
```{r}
# Create Q-Q plots for the residuals of each model
qq_plots <- lapply(list(residuals_dcmod1, residuals_hdmod1, residuals_cdmod1, residuals_dhcmod1), function(residuals) {
  ggplot(data.frame(Residuals = residuals), aes(sample = residuals)) +
    geom_qq() +
    geom_qq_line() +
    labs(title = "Q-Q Plot of Residuals")
})

# Combine Q-Q plots into a single grid
combined_qq_plots <- grid.arrange(
  grobs = qq_plots,
  nrow = 2
)

# Save the combined Q-Q plots
ggsave(
  filename = here("figs/lmer/residuals_qq_plots_combined.png"),
  plot = combined_qq_plots,
  width = 12,
  height = 8
)
```

- Issues: Not so good, I am not sure if the residuals follow normal distribution (with positive kurtosis in histogram and outliers in QQ-plot).

## Extract group-level intercepts and slopes

- Using lme4::ranef and lme4::fixef functions
```{r}
# Combine all models into a list
models <- list(dhmod1 = dhmod1, dcmod1 = dcmod1, hdmod1 = hdmod1, cdmod1 = cdmod1, dhcmod1 = dhcmod1)

# Create an empty data frame to store the results
coefficients_df <- data.frame(
  Model = character(),
  sp = character(),
  Intercept = numeric(),
  Slope = numeric(),
  stringsAsFactors = FALSE
)

# Function to extract both random and fixed effects
extract_effects <- function(model, model_name) {
  random_effects <- ranef(model)$sp
  fixed_effects <- fixef(model)

  random_df <- data.frame(
    Model = model_name,
    sp = rownames(random_effects),
    Intercept = random_effects[, 1],
    Slope = random_effects[, 2]
  )

  fixed_df <- data.frame(
    Model = model_name,
    sp = "Fixed",
    Intercept = as.numeric(fixed_effects[1]),
    Slope = as.numeric(fixed_effects[2])
  )

  return(rbind(random_df, fixed_df))
}

# Iterate over each model to extract effects and add to coefficients_df
for (model_name in names(models)) {
  coefficients_df <- rbind(coefficients_df, extract_effects(models[[model_name]], model_name))
}
```

```{r}
head(coefficients_df)
head(ov)
```

- Join parameters and overlap_20.csv data
```{r}
ov2 <- merge(ov, coefficients_df, by = 'sp', all.x = TRUE)
head(ov2)
summary(ov2)
```

# Fitting linear model for group  level predictor (wood density) and parameters of first- level models
- Issues:
I am not sure if the variables of my sub dataset are following a normal distribution.
I checked skewness for all and tried several transformation methods but QQ-plot and Histagram do not show that my transformed data follow normal distribution.
So, now, I am trying to fit the linear model for log wood density (wd.ln) and the parameters of linear mixed models first. Then I will make tar_script for them and I am finding a suitable transform method as well.

## dhmod1

- Subset data for dhmod1
```{r}
dhmod1_data <- subset(ov2, Model == "dhmod1")
head(dhmod1_data)
```

- Checking distribution of parameters
```{r}
# Histogram for Intercept
intercept_hist_dh <- ggplot(dhmod1_data, aes(x = Intercept)) +
  geom_histogram(binwidth = 0.1, fill = "gray", color = "black") +
  labs(x = "Intercept", y = "Frequency", title = "Distribution of Intercept for dhmod1")

# Histogram for Slope
slope_hist_dh <- ggplot(dhmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "orange", color = "black") +
  labs(x = "Slope", y = "Frequency", title = "Distribution of Slope for dhmod1")

# Histogram for Wood Density
wd_hist_dh <- ggplot(dhmod1_data, aes(x = wd)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +  # Adjust color and fill as desired
  labs(x = "Wood Density (g m⁻³)", y = "Frequency", title = "Distribution of Wood Density for dhmod1") +
  theme_minimal()

# Arrange all histograms into a grid
combined_hist_dh <- gridExtra::grid.arrange(intercept_hist_dh, slope_hist_dh, wd_hist_dh, nrow = 1)

# Save the combined histogram
ggsave(
  filename = here("figs/lmer/dhmod1_parameter_distribution.png"),
  plot = combined_hist_dh,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Normality qq-plot
```{r}
# Create QQ plot for Intercept
qq_plot_dh_i <- ggplot(data = dhmod1_data, aes(sample = Intercept)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Slope
qq_plot_dh_s <- ggplot(data = dhmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Wood Density
qq_plot_dh_wd <- ggplot(data = dhmod1_data, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line()

# Arrange all QQ plots into a grid
combine_qq_dh <- gridExtra::grid.arrange(qq_plot_dh_i, qq_plot_dh_s, qq_plot_dh_wd, nrow = 1)

# Save the combined QQ plot
ggsave(
  filename = here("figs/lmer/dhmod1_para_qq_plot.png"),
  plot = combine_qq_dh,
  width = 16,
  height = 6,
  dpi = 120
)
```

- Checking skewness
```{r}
# Calculate skewness for Intercept
skewness_intercept <- skewness(dhmod1_data$Intercept)
print(paste("Skewness of Intercept:", skewness_intercept))

# Calculate skewness for Slope
skewness_slope <- skewness(dhmod1_data$Slope)
print(paste("Skewness of Slope:", skewness_slope))
# Calculate skewness for Wood Density
skewness_wd <- skewness(dhmod1_data$wd)
print(paste("Skewness of Wood Density:", skewness_wd))
```

- Issues:
Intercept and wd of dhmod1_data do not follow normal distribution
[1] "Skewness of Intercept: -0.308043701635711"
[1] "Skewness of Slope: -0.0219872668781178" (it's ok i think)
[1] "Skewness of Wood Density: 0.512113500065508" (log)

- Transforming data

```{r}
dhmod1_data$Intercept.yj <- yjPower(dhmod1_data$Intercept, lambda = 1.39, jacobian.adjusted = FALSE)
dhmod1_data$Slope <- yjPower(dhmod1_data$Slope,lambda = 1.05,  jacobian.adjusted = FALSE)
dhmod1_data$wd.ln <- log(dhmod1_data$wd)
```

```{r}
# Check skewness of 'Intercept.yj' and 'wd.ln'
skewness_intercept_yj <- skewness(dhmod1_data$Intercept.yj)
skewness_wd_ln <- skewness(dhmod1_data$wd.ln)
# Print skewness values
print(paste("Skewness of Transformed Intercept (Yeo-Johnson):", skewness_intercept_yj))
print(paste("Skewness of Transformed wd (Natural Log):", skewness_wd_ln))
```

[1] "Skewness of Transformed Intercept (Yeo-Johnson): -0.00567449414355722"
[1] "Skewness of Transformed wd (Natural Log): -0.0021015611785854"

- Normality qq-plot
```{r}
# Create QQ plot for Intercept
qq_plot_dh_i.yj <- ggplot(data = dhmod1_data, aes(sample = Intercept.yj)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Slope
qq_plot_dh_s <- ggplot(data = dhmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Wood Density
qq_plot_dh_wd.ln <- ggplot(data = dhmod1_data, aes(sample = wd.ln)) +
  stat_qq() +
  stat_qq_line()

# Arrange plots into a grid
combined_qq_plots_yj <- gridExtra::grid.arrange(qq_plot_dh_i.yj, qq_plot_dh_s, qq_plot_dh_wd.ln, nrow = 1)

# Save the combined QQ plots
ggsave(
  filename = "figs/lmer/combined_qq_yj_plots.png",
  plot = combined_qq_plots_yj,
  width = 16,
  height = 6,
  dpi = 120
)
```

- Histogram
```{r}
# Histogram for Intercept.yj
histogram_intercept_yj <- ggplot(dhmod1_data, aes(x = Intercept.yj)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Intercept.yj")

# Histogram for Slope.yj
histogram_slope <- ggplot(dhmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Slope")

# Histogram for wd.ln
histogram_wd_ln <- ggplot(dhmod1_data, aes(x = wd.ln)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of wd.ln")

# Arrange histograms into a grid
combined_histograms <- gridExtra::grid.arrange(histogram_intercept_yj, histogram_slope, histogram_wd_ln, nrow = 1)

# Save the combined histograms
ggsave(
  filename = "figs/lmer/dhmod1_yj_combined_histograms.png",
  plot = combined_histograms,
  width = 16,
  height = 6,
  dpi = 120
)
```

- Fitting second- level model
```{r}
i_dhmod1 <- lm(Intercept ~ wd.ln, data = dhmod1_data)
s_dhmod1 <- lm(Slope ~ wd.ln, data = dhmod1_data)
summary(i_dhmod1)
summary(s_dhmod1)
```

- Plot models
```{r}
# Scatter plot for Intercept vs. Wood Density
intercept_plot_dh <- ggplot(dhmod1_data, aes(x = wd.ln, y = Intercept)) +
  geom_point(color = "gray", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Intercept", title = "Intercept vs. Wood Density for dhmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Scatter plot for Slope vs. Wood Density
slope_plot_dh <- ggplot(dhmod1_data, aes(x = wd.ln, y = Slope)) +
  geom_point(color = "orange", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Slope", title = "Slope vs. Wood Density for dhmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Arrange plots into a grid
combined_plot_dh <- gridExtra::grid.arrange(intercept_plot_dh, slope_plot_dh, nrow = 1)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/dhmod1_parameter_plots.png"),
  plot = combined_plot_dh,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Key findings:
There is a negative relationship between wood density and intercept but positive with slope of dhmod1 model.
Wood Density has a significant effect on both Intercept and Slope of dhmod1 model.

## dcmod1

- Subset data for dcmod1
```{r}
dcmod1_data <- subset(ov2, Model == "dcmod1")
head(dcmod1_data)
```

- Checking distribution of parameters
```{r}
# Histogram for Intercept
intercept_hist_dc <- ggplot(dcmod1_data, aes(x = Intercept)) +
  geom_histogram(binwidth = 0.1, fill = "gray", color = "black") +
  labs(x = "Intercept", y = "Frequency", title = "Distribution of Intercept for dhmod1")

# Histogram for Slope
slope_hist_dc <- ggplot(dcmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "orange", color = "black") +
  labs(x = "Slope", y = "Frequency", title = "Distribution of Slope for dhmod1")

# Histogram for Wood Density (wd)
wd_hist_dc <- ggplot(dcmod1_data, aes(x = wd)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Wood Density", y = "Frequency", title = "Distribution of Wood Density for dcmod1")

# Arrange plots into a grid
combined_hist_dc <- gridExtra::grid.arrange(intercept_hist_dc, slope_hist_dc, wd_hist_dc, nrow = 1)

# Save the combined histogram
ggsave(
  filename = here("figs/lmer/dcmod1_parameter_distribution.png"),
  plot = combined_hist_dc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Normality qq-plot
```{r}
# Create QQ plot for Intercept
qq_plot_dc_i <- ggplot(data = dcmod1_data, aes(sample = Intercept)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Slope
qq_plot_dc_s <- ggplot(data = dcmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for wd
qq_plot_dc_wd <- ggplot(data = dcmod1_data, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line()

# Arrange plots into a grid
combine_qq_dc <- gridExtra::grid.arrange(qq_plot_dc_i, qq_plot_dc_s, qq_plot_dc_wd, nrow = 1)

# Save the combined QQ plot
ggsave(
  filename = here("figs/lmer/dcmod1_para_qq_plot.png"),
  plot = combine_qq_dc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Issues:
Parameters of dcmod1_data do not follow normal distribution?

[1] "Skewness of Intercept: -1.92651634512216"
[1] "Skewness of Slope: -0.503247269693186"
[1] "Skewness of Wood Density: 0.512113500065508" (log)

- Transforming data
```{r}
dcmod1_data$Intercept.yj <- yjPower(dcmod1_data$Intercept, lambda = 2.3, jacobian.adjusted = FALSE)
dcmod1_data$Slope.yj <- yjPower(dcmod1_data$Slope,lambda = 1.45,  jacobian.adjusted = FALSE)
dcmod1_data$wd.ln <- log(dcmod1_data$wd)
```

```{r}
# Check skewness of 'Intercept.yj' and 'wd.ln'
skewness_intercept_yj <- skewness(dcmod1_data$Intercept.yj)
skewness_wd_ln <- skewness(dcmod1_data$wd.ln)
skewness_slope_yj <- skewness(dcmod1_data$Slope.yj)
# Print skewness values
print(paste("Skewness of Transformed Intercept (Yeo-Johnson):", skewness_intercept_yj))
print(paste("Skewness of Transformed Slope:", skewness_slope_yj))
print(paste("Skewness of Transformed wd (Natural Log):", skewness_wd_ln))
```

[1] "Skewness of Transformed Intercept (Yeo-Johnson): -0.0344020665645285"
[1] "Skewness of Transformed Slope: 0.0656546901573137"
[1] "Skewness of Transformed wd (Natural Log): -0.0021015611785854"

- Fitting the second-level model

```{r}
i_dcmod1 <- lm(Intercept ~ wd.ln, data = dcmod1_data)
s_dcmod1 <- lm(Slope ~ wd.ln, data = dcmod1_data)
summary(i_dcmod1)
summary(s_dcmod1)
```

- Plot models
```{r}
# Scatter plot for Intercept vs. Wood Density
intercept_plot_dc <- ggplot(dcmod1_data, aes(x = wd.ln, y = Intercept)) +
  geom_point(color = "gray", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Intercept", title = "Intercept vs. Wood Density for dcmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Scatter plot for Slope vs. Wood Density
slope_plot_dc <- ggplot(dcmod1_data, aes(x = wd.ln, y = Slope)) +
  geom_point(color = "orange", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Slope", title = "Slope vs. Wood Density for dcmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Arrange plots into a grid
combined_plot_dc <- gridExtra::grid.arrange(intercept_plot_dc, slope_plot_dc, nrow = 1)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/dcmod1_parameter_plots.png"),
  plot = combined_plot_dc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Key findings:
There is a negative relationship between wood density and the intercept but a positive relationship with the slope of the dcmod1 model.
Wood density has a significant effect on both the intercept and slope of the model.

## hdmod1

- Subset data for hdmod1
```{r}
hdmod1_data <- subset(ov2, Model == "hdmod1")
summary(hdmod1_data)
```

- Checking distribution of parameters
```{r}
# Histogram for Intercept
intercept_hist_hd <- ggplot(hdmod1_data, aes(x = Intercept)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Intercept", y = "Frequency", title = "Distribution of Intercept for hdmod1")

# Histogram for Slope
slope_hist_hd <- ggplot(hdmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "black") +
  labs(x = "Slope", y = "Frequency", title = "Distribution of Slope for hdmod1")

# Histogram for Wood Density
wd_hist_hd <- ggplot(hdmod1_data, aes(x = wd)) +
  geom_histogram(binwidth = 0.1, fill = "lightcoral", color = "black") +
  labs(x = "Wood Density", y = "Frequency", title = "Distribution of Wood Density for hdmod1")

# Arrange plots into a grid
combined_hist_hd <- gridExtra::grid.arrange(intercept_hist_hd, slope_hist_hd, wd_hist_hd, nrow = 1)

# Save the combined histogram
ggsave(
  filename = here("figs/lmer/hdmod1_parameter_distribution.png"),
  plot = combined_hist_hd,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Normality qq-plot
```{r}
# Create QQ plot for Intercept
qq_plot_hd_i <- ggplot(data = hdmod1_data, aes(sample = Intercept)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Slope
qq_plot_hd_s <- ggplot(data = hdmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line()

# Create QQ plot for Wood Density
qq_plot_hd_wd <- ggplot(data = hdmod1_data, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line()

# Arrange plots into a grid
combine_qq_hd <- gridExtra::grid.arrange(qq_plot_hd_i, qq_plot_hd_s, qq_plot_hd_wd, nrow = 1)

# Save the combined QQ plot
ggsave(
  filename = here("figs/lmer/hdmod1_para_qq_plot.png"),
  plot = combine_qq_hd,
  width = 16,
  height = 6,
  dpi = 120
)
```

- Checking skewness
```{r}
skewness_intercept <- skewness(hdmod1_data$Intercept)
print(paste("Skewness of Intercept:", skewness_intercept))

skewness_slope <- skewness(hdmod1_data$Slope)
print(paste("Skewness of Slope:", skewness_slope))

skewness_wd <- skewness(hdmod1_data$wd)
print(paste("Skewness of Wood Density:", skewness_wd))
```

- Issues:
The parameters of hdmod1 model do not follow normal distribution
[1] "Skewness of Intercept: 0.0118844338848712"
[1] "Skewness of Slope: -0.146462294861349"
[1] "Skewness of Wood Density: 0.512113500065508"

- Transforming data
```{r}
hdmod1_data$Slope.yj <- yjPower(hdmod1_data$Slope,lambda = 1.5,  jacobian.adjusted = FALSE)
hdmod1_data$wd.ln <- log(dcmod1_data$wd)
```

```{r}
skewness_wd_ln <- skewness(hdmod1_data$wd.ln)
skewness_slope_yj <- skewness(hdmod1_data$Slope.yj)
# Print skewness values
print(paste("Skewness of Transformed Intercept (Yeo-Johnson):", skewness_intercept_yj))
print(paste("Skewness of Transformed Slope:", skewness_slope_yj))
print(paste("Skewness of Transformed wd (Natural Log):", skewness_wd_ln))
```

[1] "Skewness of Transformed Intercept (Yeo-Johnson): -0.0344020665645285"
[1] "Skewness of Transformed Slope: 0.00458622401929467")
[1] "Skewness of Transformed wd (Natural Log): -0.0021015611785854"

- Fitting the second-level model
```{r}
i_hdmod1 <- lm(Intercept ~ wd.ln, data = hdmod1_data)
s_hdmod1 <- lm(Slope ~ wd.ln, data = hdmod1_data)
summary(i_hdmod1)
summary(s_hdmod1)
```

- Plot models
```{r}
# Scatter plot for Intercept vs. Wood Density
intercept_plot_hd <- ggplot(hdmod1_data, aes(x = wd.ln, y = Intercept)) +
  geom_point(color = "gray", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Intercept", title = "Intercept vs. Wood Density for hdmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Scatter plot for Slope vs. Wood Density
slope_plot_hd <- ggplot(hdmod1_data, aes(x = wd.ln, y = Slope)) +
  geom_point(color = "orange", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Slope", title = "Slope vs. Wood Density for hdmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Arrange plots into a grid
combined_plot_hd <- gridExtra::grid.arrange(intercept_plot_hd, slope_plot_hd, nrow = 1)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/hdmod1_parameter_plots.png"),
  plot = combined_plot_hd,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Key findings:
There is positive relationship between wood density and intercept but negative with slope of hdmod1 model.
Wood density has a significant impact on both the intercept and slope of the model.

## cdmod1

- Subset data for cdmod1
```{r}
cdmod1_data <- subset(ov2, Model == "cdmod1")
```

- Checking distribution of parameters
```{r}
# Histogram for Intercept
intercept_hist_cd <- ggplot(cdmod1_data, aes(x = Intercept)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Intercept", y = "Frequency", title = "Distribution of Intercept for cdmod1")

# Histogram for Slope
slope_hist_cd <- ggplot(cdmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "black") +
  labs(x = "Slope", y = "Frequency", title = "Distribution of Slope for cdmod1")

# Histogram for Wood Density
wd_hist_cd <- ggplot(cdmod1_data, aes(x = wd)) +
  geom_histogram(binwidth = 0.1, fill = "lightyellow", color = "black") +
  labs(x = "Wood Density", y = "Frequency", title = "Distribution of Wood Density for cdmod1")

# Arrange plots into a grid
combined_hist_cd <- gridExtra::grid.arrange(intercept_hist_cd, slope_hist_cd, wd_hist_cd, nrow = 1)

# Save the combined histogram
ggsave(
  filename = here("figs/lmer/cdmod1_parameter_distribution.png"),
  plot = combined_hist_cd,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Normality qq-plot
```{r}
# Normality QQ plot for Intercept
qq_plot_cd_intercept <- ggplot(cdmod1_data, aes(sample = Intercept)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Normal Q-Q Plot for Intercept")

# Normality QQ plot for Slope
qq_plot_cd_slope <- ggplot(cdmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Normal Q-Q Plot for Slope")

# Normality QQ plot for Wood Density
qq_plot_cd_wd <- ggplot(cdmod1_data, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Normal Q-Q Plot for Wood Density")

# Arrange plots into a grid
combined_qq_cd <- gridExtra::grid.arrange(qq_plot_cd_intercept, qq_plot_cd_slope, qq_plot_cd_wd, nrow = 1)

# Save the combined QQ plot
ggsave(
  filename = here("figs/lmer/cdmod1_parameter_qq_plot.png"),
  plot = combined_qq_cd,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Issues: Parameters of cdmod1_data do not follow normal distribution?

- Checking skewness
```{r}
# Calculate skewness for Intercept
skewness_intercept <- skewness(cdmod1_data$Intercept)
print(paste("Skewness of Intercept:", skewness_intercept))

# Calculate skewness for Slope
skewness_slope <- skewness(cdmod1_data$Slope)
print(paste("Skewness of Slope:", skewness_slope))

# Calculate skewness for Wood Density (wd)
skewness_wd <- skewness(cdmod1_data$wd)
print(paste("Skewness of Wood Density:", skewness_wd))
```

[1] "Skewness of Intercept: 0.537678950391731"
[1] "Skewness of Slope: -0.763657379988937"
[1] "Skewness of Wood Density: 0.512113500065508"

- Transforming data
```{r}
library(car)
cdmod1_data$Intercept.yj <- yjPower(cdmod1_data$Intercept, lambda = 0.42, jacobian.adjusted = FALSE)
cdmod1_data$Slope.yj <- yjPower(cdmod1_data$Slope,lambda = 2.4,  jacobian.adjusted = FALSE)
cdmod1_data$wd.ln <- log(cdmod1_data$wd)
```

```{r}
# Check skewness of 'Intercept.yj' and 'wd.ln'
skewness_intercept_yj <- skewness(cdmod1_data$Intercept.yj)
skewness_wd_ln <- skewness(cdmod1_data$wd.ln)
skewness_slope_yj <- skewness(cdmod1_data$Slope.yj)
# Print skewness values
print(paste("Skewness of Transformed Intercept (Yeo-Johnson):", skewness_intercept_yj))
print(paste("Skewness of Transformed Slope:", skewness_slope_yj))
print(paste("Skewness of Transformed wd (Natural Log):", skewness_wd_ln))
```

[1] "Skewness of Transformed Intercept (Yeo-Johnson): -0.000305219223815829"
[1] "Skewness of Transformed Slope: 0.00916408756428457"
[1] "Skewness of Transformed wd (Natural Log): -0.0021015611785854"

- Fitting second-level model
```{r}
i_cdmod1 <- lm(Intercept ~ wd.ln, data = cdmod1_data)
s_cdmod1 <- lm(Slope ~ wd.ln, data = cdmod1_data)
summary(i_cdmod1)
summary(s_cdmod1)
```

- Plot models
```{r}
# Scatter plot for Intercept vs. Wood Density
intercept_plot_cd <- ggplot(cdmod1_data, aes(x = wd.ln, y = Intercept)) +
  geom_point(color = "gray", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Intercept", title = "Intercept vs. Wood Density for cdmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Scatter plot for Slope vs. Wood Density
slope_plot_cd <- ggplot(cdmod1_data, aes(x = wd.ln, y = Slope)) +
  geom_point(color = "orange", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Slope", title = "Slope vs. Wood Density for cdmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Arrange plots into a grid
combined_plot_cd <- gridExtra::grid.arrange(intercept_plot_cd, slope_plot_cd, nrow = 1)

# Save the combined plot
ggsave(
  filename = here("figs/lmer/cdmod1_parameter_plots.png"),
  plot = combined_plot_cd,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Key findings:
There is positive relationship between wood density and intercept but negative with slope of cdmod1 model.
Wood density has a significant impact on both the intercept and slope of the model.

## dhcmod1

- Subset data for dhcmod1
```{r}
dhcmod1_data <- subset(ov2, Model == "dhcmod1")
```

- Checking distribution of parameters
```{r}
# Histogram for Intercept
intercept_hist_dhc <- ggplot(dhcmod1_data, aes(x = Intercept)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Intercept", y = "Frequency", title = "Distribution of Intercept for dhcmod1")

# Histogram for Slope
slope_hist_dhc <- ggplot(dhcmod1_data, aes(x = Slope)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Slope", y = "Frequency", title = "Distribution of Slope for dhcmod1")

# Histogram for Wood Density
wd_hist_dhc <- ggplot(dhcmod1_data, aes(x = wd)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(x = "Wood Density", y = "Frequency", title = "Distribution of Wood Density for dhcmod1")

# Arrange plots into a grid
combined_hist_dhc <- gridExtra::grid.arrange(intercept_hist_dhc, slope_hist_dhc, wd_hist_dhc, nrow = 1)

# Save the combined histogram
ggsave(
  filename = here("figs/lmer/dhcmod1_parameter_distribution.png"),
  plot = combined_hist_dhc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Normality qq-plot

```{r}
# QQ Plot for Intercept
qq_plot_dhc_intercept <- ggplot(dhcmod1_data, aes(sample = Intercept)) +
  stat_qq() +
  stat_qq_line() +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles", title = "QQ Plot for Intercept of dhcmod1")

# QQ Plot for Slope
qq_plot_dhc_slope <- ggplot(dhcmod1_data, aes(sample = Slope)) +
  stat_qq() +
  stat_qq_line() +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles", title = "QQ Plot for Slope of dhcmod1")

# QQ Plot for Wood Density
qq_plot_dhc_wd <- ggplot(dhcmod1_data, aes(sample = wd)) +
  stat_qq() +
  stat_qq_line() +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles", title = "QQ Plot for Wood Density of dhcmod1")

# Arrange plots into a grid
combined_qq_dhc <- gridExtra::grid.arrange(qq_plot_dhc_intercept, qq_plot_dhc_slope, qq_plot_dhc_wd, nrow = 1)

# Save the combined QQ plot
ggsave(
  filename = here("figs/lmer/dhcmod1_parameter_qq_plots.png"),
  plot = combined_qq_dhc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Checking skewness
```{r}
# Check skewness of Intercept
skewness_intercept <- skewness(dhcmod1_data$Intercept)
print(paste("Skewness of Intercept:", skewness_intercept))

# Check skewness of Slope
skewness_slope <- skewness(dhcmod1_data$Slope)
print(paste("Skewness of Slope:", skewness_slope))

# Check skewness of Wood Density
skewness_wd <- skewness(dhcmod1_data$wd)
print(paste("Skewness of Wood Density:", skewness_wd))
```

Issues:
The varibles of dhcmod1_data do not follow normal distribution

[1] "Skewness of Intercept: -0.48944085463939"
[1] "Skewness of Slope: 0.392645198693742"
[1] "Skewness of Wood Density: 0.512113500065508"

- Transforming data

```{r}
dhcmod1_data$Intercept.yj <- yjPower(dhcmod1_data$Intercept, lambda = 1.85, jacobian.adjusted = FALSE)
dhcmod1_data$Slope.yj <- yjPower(dhcmod1_data$Slope,lambda = 0.000001,  jacobian.adjusted = FALSE)
dhcmod1_data$wd.ln <- log(dhcmod1_data$wd)
```

```{r}
# Check skewness of 'Intercept.yj' and 'wd.ln'
skewness_intercept_yj <- skewness(dhcmod1_data$Intercept.yj)
skewness_wd_ln <- skewness(dhcmod1_data$wd.ln)
skewness_slope_yj <- skewness(dhcmod1_data$Slope.yj)
# Print skewness values
print(paste("Skewness of Transformed Intercept (Yeo-Johnson):", skewness_intercept_yj))
print(paste("Skewness of Transformed Slope:", skewness_slope_yj))
print(paste("Skewness of Transformed wd (Natural Log):", skewness_wd_ln))
```

- Fitting second-level model
```{r}
i_dhcmod1 <- lm(Intercept ~ wd.ln, data = dhcmod1_data)
s_dhcmod1 <- lm(Slope ~ wd.ln, data = dhcmod1_data)
summary(i_dhcmod1)
summary(s_dhcmod1)
```

- Plot models
```{r}
# Scatter plot for Intercept vs. Wood Density
intercept_plot_dhc <- ggplot(dhcmod1_data, aes(x = wd.ln, y = Intercept)) +
  geom_point(color = "gray", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Intercept", title = "Intercept vs. Wood Density for dhcmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Scatter plot for Slope vs. Wood Density
slope_plot_dhc <- ggplot(dhcmod1_data, aes(x = wd.ln, y = Slope)) +
  geom_point(color = "orange", size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1) +
  labs(x = "wd.ln", y = "Slope", title = "Slope vs. Wood Density for dhcmod1") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size = 12))

# Arrange plots into a grid
combined_plot_dhc <- gridExtra::grid.arrange(intercept_plot_dhc, slope_plot_dhc, nrow = 1)

# Save the combined plot
ggsave(
  filenam = here("figs/lmer/dhcmod1_parameter_plots.png"),
  plot = combined_plot_dhc,
  width = 12,
  height = 6,
  dpi = 120
)
```

- Key findings:
There is a negative relationship between wood density with both Intercept and  Slope of dhcmod1 model.
Wood density has a notable impact on both the intercept and slope of the dhcmod1 model.

### 2025-03-26 docs/loo_new.qmd
Initial commit for tree-allo

---
title: "Comparing models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(targets)
library(dplyr)
# library(officer)
library(stringr)
library(here)
# tar_make()
```

# Loading data

```{r}
# Read the loo_tbl
loo_tbl <- tar_read(loo_tbl)

print(loo_tbl , n = Inf)

# H models
h_models <- loo_tbl |>
  filter(str_detect(tmp, "_h"), !str_detect(sp_biome, "biome")) |>
  arrange(looic)

print(h_models)

# CR models
cr_models <- loo_tbl |>
  filter(str_detect(tmp, "_cr"), !str_detect(sp_biome, "biome")) |>
  arrange(looic)

print(cr_models)

# DBH models
dbh_models <- loo_tbl |>
  filter(str_detect(tmp, "dbh|dbh1|dbh2|dbh3"), !str_detect(sp_biome, "biome")) |>
  arrange(looic)

print(dbh_models)
```

# Calculating ΔLOOIC

```{r}
calculate_delta_looic <- function(df) {
  df <- df |>
    group_by(div) |>
    arrange(looic, .by_group = TRUE) |>
    mutate(ΔLOOIC = looic - min(looic)) |>
    ungroup()
  return(df)
}

# H models
h_models_tbl <- h_models |>
  filter(str_detect(tmp, "_h")) |>
  mutate(
    Wood_Density = ifelse(wd == "no", "Without", "With")
  ) |>
  select(tmp, model, div, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

print(h_models_tbl)

# CR models
cr_models_tbl <- cr_models |>
  filter(str_detect(tmp, "_cr")) |>
  mutate(
    Wood_Density = ifelse(wd == "no", "Without", "With")
  ) |>
  select(tmp, model, div, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

print(cr_models_tbl)

# DBH models
dbh_models_tbl <- dbh_models |>
  filter(str_detect(tmp, "dbh|dbh1|dbh2|dbh3")) |>
  mutate(
    Wood_Density = ifelse(wd == "no", "Without", "With")
  ) |>
  select(tmp, model, div, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

print(dbh_models_tbl)

```

# Formating table

```{r}
format_table <- function(df, title) {
  formatted_df <- df |>
    mutate(
      # Define Dependent Variable based on the model type
      `Dependent variable` = case_when(
        grepl("dbh", tmp) ~ "DBH",
        grepl("_h", tmp) ~ "Tree Height",
        grepl("_cr", tmp) ~ "Crown Radius",
        TRUE ~ NA_character_
      ),
      # Define Predictor Variable for each model type and specific cases
      `Predictor variable` = case_when(
        grepl("dbh_mcmc_pl_wd", tmp) ~ "CR, H",
        grepl("dbh1", tmp) ~ "CR × H",
        grepl("dbh2", tmp) ~ "CR",
        grepl("dbh3", tmp) ~ "H",
        grepl("dbh$", tmp) ~ "CR, H",
        grepl("_h", tmp) ~ "DBH",
        grepl("_cr", tmp) ~ "DBH",
        TRUE ~ NA_character_
      ),
      # Map model types to simplified names
      `Functional form` = case_when(
        grepl("pl", model, ignore.case = TRUE) ~ "Power-law",
        grepl("gmm", model, ignore.case = TRUE) ~ "gMM",
        grepl("weibull", model, ignore.case = TRUE) ~ "Weibull",
        TRUE ~ model  # Preserve original name if no match found
      )
    ) |>
    rename(
      "Division" = div,
      "Wood Density" = Wood_Density,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    select(`Dependent variable`, `Predictor variable`, `Functional form`, Division, `Wood Density`, ELPD, pLOO, LOOIC, `ΔLOOIC`) |>
    as.data.frame()

  return(formatted_df)
}

h_models_tbl_fm <- format_table(h_models_tbl, "Tree Height Models")
cr_models_tbl_fm <- format_table(cr_models_tbl, "Crown Radius Models")
dbh_models_tbl_fm <- format_table(dbh_models_tbl, "DBH Models")

options(width = 200)
print(h_models_tbl_fm)
print(cr_models_tbl_fm)
print(dbh_models_tbl_fm)

```

### 2025-03-26 docs/looic.qmd
Initial commit for tree-allo

---
title: "Comparing models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(dplyr)
library(ggplot2)
library(flextable)
library(officer)
# Define and run the targets pipeline
# tar_make()
```

# Compare models using loo_tbl

```{r}
loo_tbl <- tar_read(loo_tbl)
loo_tbl
print(loo_tbl , n = Inf)

# Extract gMM1, PL1, and Weibull1 models
loo_tbl <- loo_tbl |>
  filter(grepl("gmm1|power_law1|weibull1", model))
print(loo_tbl)

# CR models
cr_models <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  arrange(looic)

print(cr_models)

# H models
h_models <- loo_tbl |>
  filter(grepl("_h", model)) |>
  arrange(looic)

print(h_models)

# DBH models
dbh_models <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  arrange(looic)

# Print the filtered and sorted models
print(dbh_models)
```

# Function to calculate ΔLOOIC for a given table
```{r}
calculate_delta_looic <- function(df) {
  df <- df |>
    arrange(looic) |>
    mutate(ΔLOOIC = looic - min(looic))
  return(df)
}

# Calculate ΔLOOIC for CR models
cr_models_tbl <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the CR models table with ΔLOOIC
print(cr_models_tbl)

# Calculate ΔLOOIC for H models
h_models_tbl <- loo_tbl |>
  filter(grepl("_h", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the H models table with ΔLOOIC
print(h_models_tbl)

# Calculate ΔLOOIC for DBH models
dbh_models_tbl <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the DBH models table with ΔLOOIC
print(dbh_models_tbl)

```

```{r}
# Define the custom format function for tables
format_table <- function(df, title) {
  formatted_df <- df |>
    rename(
      "Model" = model,
      "Wood Density" = Wood_Density,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    select(Model, `Wood Density`, ELPD, pLOO, LOOIC, `ΔLOOIC`) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(formatted_df)

  # Return the formatted dataframe
  return(formatted_df)
}

# Applying the function to CR, H, and DBH models
cr_models_tbl_formatted <- format_table(cr_models_tbl, "Crown Radius Models")
h_models_tbl_formatted <- format_table(h_models_tbl, "Height Models")
dbh_models_tbl_formatted <- format_table(dbh_models_tbl, "DBH Models")
```

# Power-law1, gMM1, Weibull1
```{r}
options(width = 100)
# Define the custom format function for tables, filtering specific models
format_table <- function(df, title) {
  # Filter the models based on specific types
  filtered_df <- df |>
    filter(grepl("power_law1|gmm1|weibull1", model, ignore.case = TRUE)) |>
    rename(
      "Model" = model,
      "Wood Density" = Wood_Density,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    select(Model, `Wood Density`, ELPD, pLOO, LOOIC, `ΔLOOIC`) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(filtered_df)

  # Return the formatted and filtered dataframe
  return(filtered_df)
}

# Applying the function to CR, H, and DBH models, filtering only the specified model types
cr_models_tbl_formatted <- format_table(cr_models_tbl, "Crown Radius Models")
h_models_tbl_formatted <- format_table(h_models_tbl, "Height Models")
dbh_models_tbl_formatted <- format_table(dbh_models_tbl, "DBH Models")
```

```{r}
# Combine all formatted tables into one dataframe
combined_tbl <- bind_rows(
  cr_models_tbl_formatted %>% mutate(Category = "Crown Radius"),
  h_models_tbl_formatted %>% mutate(Category = "Height"),
  dbh_models_tbl_formatted %>% mutate(Category = "DBH")
)

# Create a flextable from the combined dataframe
ft_combined <- flextable(combined_tbl)

# Customize the flextable appearance
ft_combined <- ft_combined |>
  merge_v(j = c("Category")) |>
  fix_border_issues() |>
  set_header_labels(
    Category = "Category",
    Model = "Model",
    Wood_Density = "Wood Density",
    ELPD = "ELPD",
    pLOO = "pLOO",
    LOOIC = "LOOIC",
    `ΔLOOIC` = "ΔLOOIC"
  )

# Print the combined flextable
ft_combined
```

# Changing the format (Model name)
```{r}
# Define the custom format function for tables with simplified model names
format_table <- function(df, title) {
  # Mapping of model types to simplified names
  model_map <- c(
    "power_law1" = "Power-law",
    "gmm1" = "gMM",
    "weibull1" = "Weibull"
  )

  # Replace model names with simplified names
  filtered_df <- df |>
    filter(grepl("power_law1|gmm1|weibull1", model, ignore.case = TRUE)) |>
    mutate(Model_Type = case_when(
      grepl("power_law1", model, ignore.case = TRUE) ~ "Power-law",
      grepl("gmm1", model, ignore.case = TRUE) ~ "gMM",
      grepl("weibull1", model, ignore.case = TRUE) ~ "Weibull"
    )) |>
    select(Model_Type, Wood_Density, elpd, p_loo, looic, ΔLOOIC) |>
    rename(
      "Model" = Model_Type,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(filtered_df)

  # Return the formatted and filtered dataframe
  return(filtered_df)
}

# Applying the function to CR, H, and DBH models with simplified names
cr_models_tbl_formatted <- format_table(cr_models_tbl, "Crown Radius Models")
h_models_tbl_formatted <- format_table(h_models_tbl, "Height Models")
dbh_models_tbl_formatted <- format_table(dbh_models_tbl, "DBH Models")
```

# Changing format of DBH model 
```{r}
library(dplyr)

# Define the custom format function for DBH models with specific functional forms
format_dbh_table <- function(df, title) {
  # Mapping of model types to detailed functional form descriptions
  model_description_map <- c(
    "dbh" = "DBH = a * CR^b * H^c",
    "dbh1" = "DBH = a * (CR * H)^b",
    "dbh2" = "DBH = a * CR^b",
    "dbh3" = "DBH = a * H^b"
  )

  # Extract the variable type from the model name and map to descriptions
  filtered_df <- df |>
    filter(grepl("power_law1", model, ignore.case = TRUE)) |>
    mutate(Model_Type = case_when(
      grepl("dbh1", model, ignore.case = TRUE) ~ model_description_map["dbh1"],
      grepl("dbh2", model, ignore.case = TRUE) ~ model_description_map["dbh2"],
      grepl("dbh3", model, ignore.case = TRUE) ~ model_description_map["dbh3"],
      TRUE ~ model_description_map["dbh"]
    )) |>
    select(Model_Type, Wood_Density, elpd, p_loo, looic, ΔLOOIC) |>
    rename(
      "Model" = Model_Type,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(filtered_df)

  # Return the formatted and filtered dataframe
  return(filtered_df)
}

# Applying the function to the DBH models with detailed descriptions
dbh_models_tbl_formatted <- format_dbh_table(dbh_models_tbl, "DBH Models")

options(width = 150)
dbh_models_tbl_formatted
```

# All models in 1 table 
```{r}
options(width = 500)
# Combining all formatted tables into one
all_models_combined <- bind_rows(
  cr_models_tbl_formatted |> mutate(Dependent_variable = "Crown Radius"),
  h_models_tbl_formatted |> mutate(Dependent_variable = "Height"),
  dbh_models_tbl_formatted |> mutate(Dependent_variable = "DBH")
)
# Print the combined table
print(all_models_combined)

```

# Assuming all_models_combined is already created and contains the combined data

```{r}

# Define a mapping of old model names to new simplified names
model_map <- c(
  "weibull1" = "Weibull",
  "gmm1" = "gMM",
  "power_law1" = "Power-law"
)

# Modify the 'Model' column to replace the names based on the mapping
all_models_combined$Model <- sapply(all_models_combined$Model, function(model) {
  for (pattern in names(model_map)) {
    if (grepl(pattern, model, ignore.case = TRUE)) {
      return(model_map[[pattern]])
    }
  }
  return(model)
})

# Reorder columns to move 'Dependent_variable' to the first column
all_models_combined <- all_models_combined |>
  select(Dependent_variable, everything())

# Print the modified dataframe
print(all_models_combined)

```

```{r}
# Define the function to create the combined table with formulas and sort from best to worst within each Dependent_Variable
create_combined_table <- function(cr_df, h_df, dbh_df) {
  # Define mappings for dependent variables, functional forms, and their formulas
  formula_map <- list(
    "Diameter at Breast Height" = list(
      "power_law1_dbh" = "DBH = a * CR^b * H^c",
      "power_law1_dbh1" = "DBH = a * (CR * H)^b",
      "power_law1_dbh2" = "DBH = a * CR^b",
      "power_law1_dbh3" = "DBH = a * H^b"
    )
  )

  # Combine all models into one dataframe and filter for specific model types
  combined_df <- bind_rows(
    cr_df |> mutate(Dependent_Variable = "Crown Radius"),
    h_df |> mutate(Dependent_Variable = "Height"),
    dbh_df |> mutate(Dependent_Variable = "Diameter at Breast Height")
  ) |>
    filter(grepl("power_law1|gmm1|weibull1", model, ignore.case = TRUE)) |>
    mutate(
      Functional_Form = case_when(
        grepl("gmm1", model, ignore.case = TRUE) ~ "gMM",
        grepl("weibull1", model, ignore.case = TRUE) ~ "Weibull",
        grepl("power_law1", model, ignore.case = TRUE) ~ "Power-law"
      ),
      Formula = case_when(
        Dependent_Variable == "Diameter at Breast Height" & grepl("power_law1_dbh", model, ignore.case = TRUE) ~ formula_map[["Diameter at Breast Height"]][["power_law1_dbh"]],
        Dependent_Variable == "Diameter at Breast Height" & grepl("power_law1_dbh1", model, ignore.case = TRUE) ~ formula_map[["Diameter at Breast Height"]][["power_law1_dbh1"]],
        Dependent_Variable == "Diameter at Breast Height" & grepl("power_law1_dbh2", model, ignore.case = TRUE) ~ formula_map[["Diameter at Breast Height"]][["power_law1_dbh2"]],
        Dependent_Variable == "Diameter at Breast Height" & grepl("power_law1_dbh3", model, ignore.case = TRUE) ~ formula_map[["Diameter at Breast Height"]][["power_law1_dbh3"]],
        TRUE ~ NA_character_
      ),
      Species_Level_Predictor = ifelse(grepl("nou", model, ignore.case = TRUE), "Without", "With")
    ) |>
    rename(
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    select(Dependent_Variable, Formula, Functional_Form, Species_Level_Predictor, ELPD, pLOO, LOOIC, ΔLOOIC) %>%
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    arrange(Dependent_Variable, LOOIC)

  # Move DBH to the end
  dbh_rows <- combined_df[combined_df$Dependent_Variable == "Diameter at Breast Height", ]
  other_rows <- combined_df[combined_df$Dependent_Variable != "Diameter at Breast Height", ]
  combined_df <- bind_rows(other_rows, dbh_rows)

  return(combined_df)
}

# Create the combined table
combined_models_tbl <- create_combined_table(cr_models_tbl, h_models_tbl, dbh_models_tbl)

# Create a flextable
ft <- flextable(combined_models_tbl)

# Merge cells for common Dependent_Variable
ft <- ft |>
  merge_v(j = c("Dependent_Variable")) |>
  fix_border_issues() |>
  set_header_labels(
    Dependent_Variable = "Dependent Variable",
    Formula = "Formula",
    Functional_Form = "Functional Form",
    Species_Level_Predictor = "Species-Level Predictor",
    ELPD = "ELPD",
    pLOO = "pLOO",
    LOOIC = "LOOIC",
    `ΔLOOIC` = "ΔLOOIC"
  )

# Print the flextable
ft
```

### 2025-03-26 docs/mle.qmd
Initial commit for tree-allo

---
title: "The relationships between tree architecture and wood density"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the packages
```{r}
library(readr)
library(tidyverse)
library(here)
library(lme4)
library(e1071)
library(car)
library(sjPlot)
library(gridExtra)
```

# Data preparation
```{r}
tallo_df <- read_csv(here("data/tallo_cleaned.csv"))
tallo_df
```

```{r}
wd_df <- read_csv(here("data/wd.csv"))
wd_df <- wd_df |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_df
```

## Join tallo and wood density dataframe
```{r}
tallo_wd <- left_join(tallo_df, wd_df, by = "sp")
tallo_wd
```

## Data transformation
```{r}
tallo_wd$dbh_ln <- log(tallo_wd$dbh)
tallo_wd$cr_ln <- log(tallo_wd$cr)
tallo_wd$h_ln <- log(tallo_wd$h)
tallo_wd
```

```{r}
wd_s_numeric <- as.numeric(tallo_wd$wd_s)
dbh_ln_numeric <- as.numeric(tallo_wd$dbh_ln)
h_ln_numeric <- as.numeric(tallo_wd$h_ln)
cr_ln_numeric <- as.numeric(tallo_wd$cr_ln)
str(tallo_wd)
```

# Fitting lme4::lmer

## Scatter plot
```{r}
scatter_h_dbh <- ggplot(tallo_wd, aes(x = dbh_ln, y = h_ln)) +
  geom_point(color = "skyblue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Log(dbh)", y = "Log(h)", title = "Scatter Plot of Log(dbh) vs Log(h)")

scatter_cr_dbh <- ggplot(tallo_wd, aes(x = dbh_ln, y = cr_ln)) +
  geom_point(color = "skyblue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Log(dbh)", y = "Log(cr)", title = "Scatter Plot of Log(dbh) vs Log(cr)")

scatter_dbh_hcr <- ggplot(tallo_wd, aes(x = h_ln + cr_ln, y = dbh_ln)) +
  geom_point(color = "skyblue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Log(h) + Log(cr)", y = "Log(dbh)", title = "Scatter Plot of Log(h) + Log(cr) vs Log(dbh)")

scatter_h_dbh 
scatter_cr_dbh
scatter_dbh_hcr
```

## Fitting models (lmer)
```{r}
dhcmod1 <- lme4::lmer(dbh_ln ~ (h_ln + cr_ln) + wd_s + (h_ln + cr_ln):wd_s + (1 + (h_ln + cr_ln) | sp), data = tallo_wd)
hdmod1 <- lme4::lmer(h_ln ~ dbh_ln + wd_s + dbh_ln:wd_s + (1 + dbh_ln | sp), data = tallo_wd)
cdmod1 <- lme4::lmer(cr_ln ~ dbh_ln + wd_s + dbh_ln:wd_s + (1 + dbh_ln | sp), data = tallo_wd)
```

```{r}
summary(hdmod1)
summary(cdmod1)
summary(dhcmod1)
```

```{r}
plot(dhcmod1)
plot(hdmod1)
plot(cdmod1)
```

## Key findings
- hdmod1: 
The t-value of wd_s is 2.110. This t-value indicates the significant positive effect of coefficiect estimate for wd_s on h_ln.
The t-values of dbh_ln:wd_s is -5.182, which indicates that the interaction between dbh_ln and wd_s has a significant negative effect on h_ln.
- cdmod1: The t-value associated with wd_s is 3.819. This t-value indicates the significance of wd_s.
On the other hand, the interaction between dbh_ln and wd_s does not have a significant effect in cr_ln in this model.
- dhcmod1: The t-value associated with wd_s is -1.982, suggesting that wd_s has a negative effect on dbh_ln (but it is not statistically significant at conventional levels with p > 0.05).
Hence, in this model, neither wd_s nor its interations with h_ln or cr_ln have statistically significant effects on dbh_ln.

### 2025-03-26 docs/overlap_sp.qmd
Initial commit for tree-allo

---
title: "Overlaps species"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r}
library(readr)
library(tidyverse)
library(tidyr)
library(here)
library(dplyr)
```

# Preparation of data

- tallo with allometric variables (h, dbh, cr)

```{r}
allo <- read_csv(here("data-raw/Tallo.csv"))
```

- "try" with wood density for each species

```{r}
wood <- read_csv(here("data/try.csv"))
wood[1:5]
```

# Cleaning data
- tallo

```{r}
selected_data <- allo %>%
  select(sp = species, dbh = stem_diameter_cm, h = height_m, cr = crown_radius_m)
write_csv(selected_data, "data/allo.csv")
```

- try

```{r}
selected_data <- wood %>%
  select(sp = AccSpeciesName, wd = OrigValueStr, unit = OrigUnitStr)
write_csv(selected_data, "data/wd.csv")
wd <- separate(wd, sp.wd., into = c("sp", "wd", "unit"), sep = ";")
wd1 <- read_csv("data/wd.csv")
str(wd1)
head(wd1, n=5) 
```

- filter
```{r}
allo2 <- read_csv("data/allo.csv")
wood2 <- read_csv("data/wd.csv")
allo3 <- allo2 %>%
  filter(!is.na(sp) & sp != "unknown" & !is.na(dbh) & !is.na(h) & !is.na(cr))

wood3 <- wood2 %>%
  filter(!is.na(sp) & sp != "unknown" & !is.na(wd) & !is.na(unit))

write_csv(allo3, "data/allo_cleaned.csv")
write_csv(wood3, "data/wd_cleaned.csv")

```

- checking for the overlap (2209 species)
```{r}
allo_cleaned <- read_csv("data/allo_cleaned.csv")
wd_cleaned <- read_csv("data/wd_cleaned.csv")
allo_species <- unique(allo_cleaned$sp)
wd_species <- unique(wd_cleaned$sp)
overlap_sp <- intersect(allo_species, wd_species)
cat("Overlapping species:\n")
print(overlap_sp)
overlap_sp
```

- cleaning wd_cleaned.csv data
```{r}
wd2<- read_csv("data/wd_cleaned.csv")
table(wd2$unit)
```

- chosing units
```{r}
print(wd2)
```

```{r}
selected_units <- c("(g/cm3)*100", "g /(cm3*10)", "g*cm-3", "g/dm3", "gr/cm3", "kg/m3", "mg/mm^3", 
                    "t/m3", "cm3/mg", "g / cm3", "g cm-3", "g/cm^3", "g\xb7cm-3", "mg mm-3", 
                    "mg/mm3", "g / ml", "g_cm3-1", "g/cm3", "mg / mm3", "mg/cm3", "t m3", "kg m-3")
```

```{r}
wd3 <- wd2[wd2$unit %in% selected_units, ]
write_csv(wd3, "data/wd_unit.csv")
head(wd3, 6)
summary(wd3)
wd3$wd <- as.numeric(wd3$wd)
```

- convert units to g/cm^3
```{r}
conversion_factors <- c(`(g/cm3)*100` = 1 / 100, 
                        `g /(cm3*10)` = 1 / 10, 
                        `g*cm-3` = 1 , 
                        `g/dm3` = 1 / 1000, 
                        `gr/cm3` = 1, 
                        `kg/m3` = 1 / 1000, 
                        `mg/mm^3` = 1, 
                        `t/m3` = 1 , 
                        `cm3/mg` = 1 / (1 * 1000), 
                        `g / cm3` = 1, 
                        `g cm-3` = 1, 
                        `g/cm^3` = 1, 
                        `g\xb7cm-3` = 1, 
                        `mg mm-3` = 1 , 
                        `mg/mm3` = 1 , 
                        `g / ml` = 1 , 
                        `g_cm3-1` = 1, 
                        `g/cm3` = 1, 
                        `mg / mm3` = 1 , 
                        `mg/cm3` = 1 / (1 * 1000), 
                        `t m3` = 1 ,
                        `kg m-3` = 1 / 1000)
```

```{r}
wd3$wd_convert <- wd3$wd * conversion_factors[as.character(wd3$unit)]
```

```{r}
wd4 <- wd3[, c("sp", "wd_convert")]
colnames(wd4) <- c("sp", "wd_converted")  # Rename the columns
write.csv(wd4, "data/wd_convert.csv", row.names = FALSE)
summary(wd4)
```

# Checking after converting
```{r}
wd_filtered <- wd4 %>% 
  filter(wd_convert > 0.2 & wd_convert <= 1.2)
write.csv(wd_filtered, "data/wd_converted.csv", row.names = FALSE)
wd5 <- read_csv("data/wd_converted.csv")
head(wd5, 6)
```

- checking overlap species (2208)
```{r}
allo4 <- read_csv("data/allo_cleaned.csv")
wd6 <- read_csv("data/wd_converted.csv")
species_allo4 <- unique(allo4$sp)
species_wd6 <- unique(wd6$sp)
overlapping_species <- intersect(species_allo4, species_wd6)
overlapping_species
```

# Join wd data with tallo allometry data (my macbook can not torelate)

```{r}
allo_wd <- left_join(allo4, wd6, by = "sp")
head(allo_wd, 3)
```

- write the data to a csv file
```{r}
write_csv(allo_wd, "data/overlap.csv")
```

- checking
```{r}
ov1 <- read_csv("data/overlap.csv")
summary(ov1)
head(ov1, 20)
```

### 2025-03-26 docs/pl_cr.qmd
Initial commit for tree-allo

---
title: "Example for Power-law DBH-CR"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(targets)
library(ggplot2)
library(dplyr)
library(stringr)
library(reshape2)
library(gridExtra)
library(readr)
```

# Data preparation
```{r}
# Load MCMC results and Stan data
pl_cr <- tar_read(fit_lr_mcmc_power_law_cr)
stan_lr_data_cr <- tar_read(stan_data_lr_cr)
str(stan_lr_data_cr)

```

# Extract relevant data
```{r}
log_y <- stan_lr_data_cr$log_y
log_x <- stan_lr_data_cr$log_x
sp <- as.factor(stan_lr_data_cr$jj)

str(log_x)
head(log_x) # Use log_x[, 2] for the fitted values
summary(exp(log_y))
summary(exp(log_x))

# Extract beta parameters
beta <- pl_cr$summary(variable = "beta")
beta

# Extract the mean values of beta parameters for each species
log_a1 <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b1 <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

# Verify the lengths of the extracted parameters
print(length(log_a1))
print(length(b1))
print(length(sp))

# Compute fitted values for CR using the power law formula
fitted_log_y_cr <- sapply(1:length(log_a1), function(i) {
  log_a1[i] + b1[i] * log_x[, 2]  
})

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a1 <- mean(log_a1)
community_b1 <- mean(b1)
community_fitted_log_y_cr <- community_log_a1 + community_b1 * log_x[, 2]
community_fitted_CR <- exp(community_fitted_log_y_cr)

# Create a data frame for plotting
data_cr <- data.frame(CR = exp(log_y), DBH = exp(log_x[, 2]), sp = sp)
head(data_cr)
summary(data_cr)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a1)) {
  data_cr[[paste0("fitted_CR_sp_", i)]] <- exp(fitted_log_y_cr[, i])
}

data_long <- melt(data_cr, id.vars = c("CR", "DBH", "sp"), measure.vars = grep("fitted_CR_sp_", names(data_cr), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)
str(data_cr)

```

# Plot observed CR vs DBH with fitted lines for each species and community-level line
```{r}
pl_cr_plot <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  theme_minimal()

# Save the plot
ggsave("figs/plot_CR_vs_DBH.png", plot = pl_cr_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot)

```

```{r}
pl_cr_plot1 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_continuous(trans = "log10", breaks = scales::trans_breaks("log10", function(x) 10^x),
                     labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_y_continuous(trans = "log10", breaks = scales::trans_breaks("log10", function(x) 10^x),
                     labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  theme_minimal()

# Save the plot
ggsave("figs/plot1_CR_vs_DBH.png", plot = pl_cr_plot1, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot1)
```

```{r}
library(scales)

# Plot observed CR vs DBH with fitted lines for each species and community-level line
pl_cr_plot5 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  theme_minimal()

# Save the plot
ggsave("figs/plot5_CR_vs_DBH.png", plot = pl_cr_plot5, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

```

# Plot for some sp

```{r}
# Identify the top 5 species with the most observations
top_5_species <- data_cr |>
  count(sp, sort = TRUE) |>
  top_n(5, n) |>
  pull(sp)
top_5_species
# Create and save plots for the top 5 species
output_dir <- "figs/top_5_species_plots"

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

for (species in top_5_species) {
  species_data_cr <- data_cr %>% filter(sp == species)
  species_data_long <- data_long %>% filter(sp == species)

  species_plot <- ggplot(species_data_cr, aes(x = DBH, y = CR)) +
    geom_point(alpha = 0.5, color = "grey") +
    geom_line(data = species_data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) +
    geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
    labs(title = paste("Scatter Plot of CR vs DBH for Species", species),
         x = "DBH (cm)",
         y = "CR (m)") +
    scale_x_log10(labels = scales::comma, limits = c(1, 150)) +
    scale_y_log10(labels = scales::comma) +
    theme_minimal()

  plot_filename <- file.path(output_dir, paste0("plot_CR_vs_DBH_species_", species, ".png"))
  ggsave(plot_filename, plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
    print(species_plot)
}
```

# How wood density effect tree architecture differentiation

```{r}
library(dplyr)
str(stan_lr_data_cr)
u <- stan_lr_data_cr$u[2, ]

# # Load the tallo_wf_df_200 dataset
# tallo_wf_df_200 <- read_csv("data/tallo_wd_200.csv")

# # Calculate mean wd for each species and filter species with at least 5 observations
# wd_df <- tallo_wf_df_200 |>
#   group_by(sp) |>
#   filter(n() >= 5) |>
#   select(sp, mean_wd)

# Create a data frame for plotting
data_wd_plot <- data.frame(
  a1 = exp(log_a1),
  b1 = b1,
  wd = u
)

head(data_wd_plot)
summary(data_wd_plot)

# Plot for intercepts (a)
plot_a <- ggplot(data_wd_plot, aes(x = wd, y = a1)) +
  geom_point(color = "grey", alpha = 0.5) +
  geom_smooth(method = "lm", color = "black", fill = "grey") +
  labs(title = "(a)", x = "WD_s", y = "Intercept a") +
  theme_minimal()

plot_a

# Plot for exponents (b)
plot_b <- ggplot(data_wd_plot, aes(x = wd, y = b1)) +
  geom_point(color = "orange", alpha = 0.5) +
  geom_smooth(method = "lm", color = "black", fill = "orange") +
  labs(title = "(b)", x = "WD_s", y = "Exponent b") +
  theme_minimal()

plot_b

# Arrange the plots vertically
combined_plot <- grid.arrange(plot_a, plot_b, ncol = 1)

# Save the plot
ggsave("figs/intercept_exponent_plot.png", plot = combined_plot, width = 8, height = 10, units = "in", dpi = 300, bg = "white")

# Display the plot
print(combined_plot)
```

```{r}
pl_cr_plot2 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Save the plot
ggsave("figs/plot2_CR_vs_DBH.png", plot = pl_cr_plot2, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot2)
```

```{r}
# Plot observed CR vs DBH with fitted lines for each species and community-level line
pl_cr_plot3 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of Crown Radius vs Diameter with Fitted Lines for Each Species and Community-Level Line",
       x = "Diameter at Breast Height (cm)",
       y = "Crown Radius (m)") +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  annotation_logticks(sides = "bl") +  # Add logarithmic ticks on both axes
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

# Save the plot
ggsave("figs/plot3_CR_vs_DBH.png", plot = pl_cr_plot3, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

pl_cr_plot3
```

# Comparison of pl_cr and pl_nou_cr

```{r}
# Load necessary libraries
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
```
```{r}
pl_cr <- tar_read(fit_lr2_mcmc_power_law_cr)
pl_nou_cr <- tar_read(fit_lr2_mcmc_power_law_nou_cr)
```

```{r}
# Extract log-likelihood for both models
log_lik_cr <- pl_cr$draws(variables = "log_lik", format = "draws_matrix")
log_lik_cr_matrix <- as_draws_matrix(log_lik_cr)

log_lik_nou_cr <- pl_nou_cr$draws(variables = "log_lik", format = "draws_matrix")
log_lik_nou_cr_matrix <- as_draws_matrix(log_lik_nou_cr)

# Calculate LOO for both models
loo_cr <- loo(log_lik_cr_matrix, reloo = TRUE, save_psis = TRUE)
loo_nou_cr <- loo(log_lik_nou_cr_matrix, reloo = TRUE, save_psis = TRUE)

# Print LOO results
print(loo_cr)
print(loo_nou_cr)

# Compare models
model_comparison <- loo_compare(loo_cr, loo_nou_cr)
print(model_comparison)
```

### 2025-03-26 docs/pl_dbh.qmd
Initial commit for tree-allo

---
title: "Comparing DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
# Define and run the targets pipeline
# tar_make()
```

- Note:

MCMC setting:
  chains = 3,
  parallel_chains = 3,
  iter_warmup = 1000,
  iter_sampling = 1000,
  adapt_delta = 0.9,
  max_treedepth = 15,

# Diagnostics checking

```{r}
options(width = 100)
# Assuming you have CmdStanMCMC or CmdStanFit objects for each model
# Load diagnostics for each model
diagnostics_dbh <- tar_read(fit_lr1_diagnostics_power_law_dbh)
diagnostics_dbh1 <- tar_read(fit_lr1_diagnostics_power_law_dbh1)
diagnostics_dbh2 <- tar_read(fit_lr1_diagnostics_power_law_dbh2)
diagnostics_dbh3 <- tar_read(fit_lr1_diagnostics_power_law_dbh3)
diagnostics_nou_dbh <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh)
diagnostics_nou_dbh1 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh1)
diagnostics_nou_dbh2 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh2)
diagnostics_nou_dbh3 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh3)

summary(diagnostics_dbh)
summary(diagnostics_dbh1)
summary(diagnostics_dbh2)
summary(diagnostics_dbh3)
summary(diagnostics_nou_dbh)
summary(diagnostics_nou_dbh1)
summary(diagnostics_nou_dbh2)
summary(diagnostics_nou_dbh3)
```

- All models exhibit stable treedepth, no divergences, and consistent behavior across chains and iterations. The mean energy levels, acceptance statistics, and step sizes vary slightly among the models but remain within acceptable ranges, indicating well-mixed chains and reliable diagnostics.

# Extract log-likelihood values

```{r}
draws_dbh <- tar_read(fit_lr1_draws_power_law_dbh)
draws_dbh1 <- tar_read(fit_lr1_draws_power_law_dbh1)
draws_dbh2 <- tar_read(fit_lr1_draws_power_law_dbh2)
draws_dbh3 <- tar_read(fit_lr1_draws_power_law_dbh3)
draws_nou_dbh <- tar_read(fit_lr1_draws_power_law_nou_dbh)
draws_nou_dbh1 <- tar_read(fit_lr1_draws_power_law_nou_dbh1)
draws_nou_dbh2 <- tar_read(fit_lr1_draws_power_law_nou_dbh2)
draws_nou_dbh3 <- tar_read(fit_lr1_draws_power_law_nou_dbh3)
```

```{r}
log_lik_dbh <- as.matrix(draws_dbh |> select(starts_with("log_lik")))
log_lik_dbh1 <- as.matrix(draws_dbh1 |> select(starts_with("log_lik")))
log_lik_dbh2 <- as.matrix(draws_dbh2 |> select(starts_with("log_lik")))
log_lik_dbh3 <- as.matrix(draws_dbh3 |> select(starts_with("log_lik")))
log_lik_nou_dbh <- as.matrix(draws_nou_dbh |> select(starts_with("log_lik")))
log_lik_nou_dbh1 <- as.matrix(draws_nou_dbh1 |> select(starts_with("log_lik")))
log_lik_nou_dbh2 <- as.matrix(draws_nou_dbh2 |> select(starts_with("log_lik")))
log_lik_nou_dbh3 <- as.matrix(draws_nou_dbh3 |> select(starts_with("log_lik")))
```

# Compute LOOIC for each model

```{r}
loo_dbh <- loo(log_lik_dbh)
loo_dbh1 <- loo(log_lik_dbh1)
loo_dbh2 <- loo(log_lik_dbh2)
loo_dbh3 <- loo(log_lik_dbh3)
loo_nou_dbh <- loo(log_lik_nou_dbh)
loo_nou_dbh1 <- loo(log_lik_nou_dbh1)
loo_nou_dbh2 <- loo(log_lik_nou_dbh2)
loo_nou_dbh3 <- loo(log_lik_nou_dbh3)

# loo_dbh
# loo_dbh1
# loo_dbh2
# loo_dbh3
# loo_nou_dbh
# loo_nou_dbh1
# loo_nou_dbh2
# loo_nou_dbh3
```

```{r}
# Function to save LOO plots
save_loo_plot <- function(loo_object, filename) {
  png(filename, width = 8, height = 6, units = "in", res = 300)
  plot(loo_object)
  dev.off()
}

# Create and save the plots with titles
save_loo_plot(loo_dbh, "figs/dbh/dbh_loo_plot.png")
save_loo_plot(loo_dbh1, "figs/dbh/dbh1_loo_plot.png")
save_loo_plot(loo_dbh2, "figs/dbh/dbh2_loo_plot.png")
save_loo_plot(loo_dbh3, "figs/dbh/dbh3_loo_plot.png")
save_loo_plot(loo_nou_dbh, "figs/dbh/nou_dbh_loo_plot.png")
save_loo_plot(loo_nou_dbh1, "figs/dbh/nou_dbh1_loo_plot.png")
save_loo_plot(loo_nou_dbh2, "figs/dbh/nou_dbh2_loo_plot.png")
save_loo_plot(loo_nou_dbh3, "figs/dbh/nou_dbh3_loo_plot.png")
```

# Compare models using LOOIC
```{r}
comparison <- loo_compare(loo_dbh, loo_dbh1, loo_dbh2, loo_dbh3, loo_nou_dbh, loo_nou_dbh1, loo_nou_dbh2, loo_nou_dbh3)

# Print comparison results
print(comparison)
```

- Note:

print(comparison)
       elpd_diff se_diff 
model1      0.0       0.0
model5     -0.9       2.1
model6  -2705.2      93.2
model2  -2707.0      93.2
model8 -10973.9     169.4
model4 -10975.6     169.4
model7 -17706.9     213.3
model3 -17707.4     213.3

```{r}
# Create a data frame for comparison
model_names <- c("loo_dbh", "loo_dbh1", "loo_dbh2", "loo_dbh3", "loo_nou_dbh", "loo_nou_dbh1", "loo_nou_dbh2", "loo_nou_dbh3")
elpd_diff <- comparison[, "elpd_diff"]
se_diff <- comparison[, "se_diff"]

# Map the results back to the model names based on loo_compare output
comparison_df <- data.frame(
  Model = model_names[c(1, 5, 6, 2, 8, 4, 7, 3)],  # Reorder based on loo_compare output order
  ELPD_Difference = round(elpd_diff, 1),
  SE_Difference = round(se_diff, 1),
  Formula = c(
    "DBH = a * (CR^b * H^c)",
    "DBH = a * (CR^b * H^c)",
    "DBH = a * (CR * H)^b",
    "DBH = a * (CR * H)^b",
    "DBH = a * H^b",
    "DBH = a * H^b",
    "DBH = a * CR^b",
    "DBH = a * CR^b"
  )
)
comparison_df
# Print the comparison table
library(knitr)
kable(comparison_df, caption = "LOOIC Comparison of DBH Models")
```

- Best Model:
Model 1 (loo_dbh) with the formula DBH = a * (CR^b * H^c) is the best-performing model.

The best-performing models (1, 5, 6, 2) use the formula DBH = a * (CR^b * H^c), DBH = a * (CR * H)^b, indicating that including both CR and H is crucial for accurate DBH estimation.

Models without wood density (indicated by the prefix “nou” in  table) do not show significantly worse performance compared to their counterparts with wood density. For example, model 1 (loo_dbh) and model 5 (loo_nou_dbh) have very similar ELPD differences, suggesting that simpler models without wood density are nearly as effective.

While simpler models without wood density perform well, the model incorporating wood density as a species-level predictor (model 1: DBH = a * (CR^b * H^c)) shows the best overall performance. This indicates that, despite the minimal difference in performance, including wood density may still provide some benefit.

- Key findings:

Accurate DBH Estimation: Incorporating both tree height (H) and crown radius (CR) is essential for accurate DBH estimation.

Wood Density: While the inclusion of wood density does not significantly improve the model’s performance, it provides a slight benefit, making the model with wood density the best performer overall.

# Plot DBH = aCR^b * H^c model (1, 5)

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_summary <- tar_read(fit_lr1_summary_power_law_dbh)
colnames(dbh_summary)

# Filter rows where the variable is "beta"
beta <- dbh_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Combine CR and H for plotting
data_dbh$CR_H <- data_dbh$CR^community_b_cr * data_dbh$H^community_b_h

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dbh_plot <- ggplot(data_dbh, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*CR^b * H^c",
       x = "CR_H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")
```

- Note: Plot looks weird

# Fixxing code
```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)
head(stan_data_lr_dbh$log_x)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

head(log_x_cr)
head(log_x_h)

# Read in the model summary for model 
dbh_summary <- tar_read(fit_lr1_summary_power_law_dbh)

# Filter rows where the variable is "beta"
beta <- dbh_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Debug: Check the lengths of log_a, b_cr, and b_h
cat("Lengths of log_a, b_cr, b_h:\n", length(log_a), length(b_cr), length(b_h), "\n")

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)
gamma <- dbh_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Combine CR and H for plotting
data_dbh$CR_H <- data_dbh$CR^community_b_cr * data_dbh$H^community_b_h

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

# Plotting with no background
dbh_plot <- ggplot(data_dbh, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(x = CR_H, y = community_fitted_DBH), color = "red", linewidth = 1.2) + # Community fit line in red
  labs(title = "DBH = a * CR^b * H^c",
       x = "CR^b * H^c",
       y = "DBH") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal() +
  theme(panel.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.background = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_blank())

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1_2.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "transparent")

# Print the plot to visualize in R
print(dbh_plot)
```

- Note: still weird

```{r}
# Create a data frame for plotting
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
head(data_dbh)

# Fit the overall model
fit <- lm(log_DBH ~ log_CR + log_H, data = data_dbh)
summary(fit)

# Generate overall fitted values
data_dbh$fitted_log_DBH <- predict(fit, newdata = data_dbh)
head(data_dbh)

species_fitted_values <- data_dbh |>
  group_by(sp) |>
  do({
    fit_sp <- lm(log_DBH ~ log_CR + log_H, data = .)
    data.frame(
      log_DBH = .$log_DBH,
      fitted_log_DBH_sp = predict(fit_sp, newdata = .),
      sp = .$sp
    )
  })

# Merge the species fitted values back to the main data frame
data_dbh <- merge(data_dbh, species_fitted_values, by = c("log_DBH", "sp"))
head(data_dbh)

plot_observed_vs_predicted <- ggplot(data_dbh, aes(x = fitted_log_DBH, y = log_DBH)) +
  geom_point(alpha = 0.3, color = "grey") +
  geom_abline(intercept = 0, slope = 1, color = "#4682B4", linetype = "dashed") + # Perfect fit line
  geom_smooth(aes(group = sp, color = ), method = "lm", se = FALSE, alpha = 0.5) + # Species-specific fitted lines
  labs(title = "DBH = aCR^b*H^c",
       x = "log(CR) + log(H)",
       y = "log(DBH)") +
  theme_minimal() +
  guides(color = "none") # Remove the legend
plot_observed_vs_predicted
```

# Plot DBH = a(CR*H)^b model (6, 2)

```{r}
comparison1 <- loo_compare(loo_dbh1, loo_dbh2, loo_dbh3, loo_nou_dbh1, loo_nou_dbh2, loo_nou_dbh3)

# Print comparison results
print(comparison1)
```

# Read the data

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)
# u <- stan_data_lr_dbh1$u

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x[, 2]
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract the gamma values
gamma <- dbh1 |> filter(grepl("gamma", variable))

# Compute the community-level intercept and slope using gamma
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)

# Compute fitted values for the community-level line
community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), sp = sp)

data_dbh$community_fitted_DBH <- community_fitted_DBH

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Transform data for plotting
data_long <- data_dbh |> melt(id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)
```

```{r}
# Plotting
dbh_plot1 <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

```

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)
# u <- stan_data_lr_dbh1$u

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Community parameters
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[1,2]") |> pull(mean)
community_log_a
community_b

community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), community_fitted_DBH, sp = sp)

data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)
str(data_dbh)

dbh <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = CRxH, y = community_fitted_DBH, color = sp, group = sp), show.legend = FALSE) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CRxH ",
       y = "DBH (cm)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

dbh
```

# Plot DBH = a(CR*H)^b model (6, 2) (Using gamma as community-level parameters)

```{r}
# Read the data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x[, 2]
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract community-level parameters directly
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[2,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]

community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH

head(data_dbh)
summary(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

data_long <- melt(data_dbh, id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)

# Plotting
dbh_plot1 <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

```

### 2025-03-26 docs/sp_cm_plot.qmd
Initial commit for tree-allo

---
title: "Weibull-based allometric models for Tree Height and Crown Radius with DBH at community and species level"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r include=FALSE}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
library(patchwork)
```

# Read in data

```{r}
# Load H data
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Load CR data
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

```

# Ranging DBH

```{r}
dbh_data <- data.frame(sp = as.factor(stan_data_nlr_h$jj), DBH = stan_data_nlr_h$x)

# Calculate the range of DBH for each species
dbh_range <- dbh_data |>
  group_by(sp) |>
  summarize(min_dbh = min(DBH), max_dbh = max(DBH))

```

# Extract parameters

## H parameters

```{r}
# Function to extract beta parameters and compute fitted values using species-specific DBH range
extract_and_compute_fitted_h <- function(data, summary, dbh_range) {
  log_y <- data$log_y
  x <- data$x
  sp <- as.factor(data$jj)

  # Extract beta parameters
  beta <- summary |> filter(grepl("beta\\[", variable))

  log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50)
  b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
  k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)

  # Create DBH sequence for each species based on actual range
  create_dbh_sequence <- function(species, dbh_range, length_out = 100) {
    min_dbh <- dbh_range$min_dbh[dbh_range$sp == species]
    max_dbh <- dbh_range$max_dbh[dbh_range$sp == species]
    seq(min_dbh, max_dbh, length.out = length_out)
  }

  # Calculate fitted H values for each species using species-specific DBH range
  fitted_h_list <- lapply(unique(sp), function(s) {
    dbh_seq <- create_dbh_sequence(s, dbh_range)
    i <- which(unique(sp) == s)  # Get the index of the species
    fitted_log_y <- log_a[i] + log(1 - exp(-b[i] * (dbh_seq ^ k[i])))
    fitted_h <- exp(fitted_log_y)

    data.frame(
      DBH = dbh_seq,
      value = fitted_h,
      variable = paste0("fitted_H_sp_", i)
    )
  })

  # Combine all species data frames into one
  data_long_h <- do.call(rbind, fitted_h_list)

  # Extract community-level parameters
  gamma <- summary |> filter(grepl("gamma", variable))

  community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
  community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
  community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

  community_log_a <- exp(community_log_a_hat)
  community_b <- exp(community_b_hat)
  community_k <- exp(community_k_hat)

  community_fitted_log_y_h <- community_log_a + log(1 - exp(-community_b * (x ^ community_k)))
  community_fitted_H <- exp(community_fitted_log_y_h)

  list(fitted_h = fitted_h_list, community_fitted_H = community_fitted_H, log_y = log_y, x = x, sp = sp, data_long_h = data_long_h)
}

# Apply function to H data
h_params <- extract_and_compute_fitted_h(stan_data_nlr_h, tar_read(fit_nlr_summary_weibull1_h), dbh_range)

# Check the extracted and computed values
str(h_params)
```

# CR paramter

```{r}
# Function to extract beta parameters and compute fitted values using species-specific DBH range
extract_and_compute_fitted_cr <- function(data, summary, dbh_range) {
  log_y <- data$log_y
  x <- data$x
  sp <- as.factor(data$jj)

  # Extract beta parameters
  beta <- summary |> filter(grepl("beta\\[", variable))

  log_a <- beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50)
  b <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
  k <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)

  # Create DBH sequence for each species based on actual range
  create_dbh_sequence <- function(species, dbh_range, length_out = 100) {
    min_dbh <- dbh_range$min_dbh[dbh_range$sp == species]
    max_dbh <- dbh_range$max_dbh[dbh_range$sp == species]
    seq(min_dbh, max_dbh, length.out = length_out)
  }

  # Calculate fitted CR values for each species using species-specific DBH range
  fitted_cr_list <- lapply(unique(sp), function(s) {
    dbh_seq <- create_dbh_sequence(s, dbh_range)
    i <- which(unique(sp) == s)  # Get the index of the species
    fitted_log_y <- log_a[i] + log(1 - exp(-b[i] * (dbh_seq ^ k[i])))
    fitted_cr <- exp(fitted_log_y)

    data.frame(
      DBH = dbh_seq,
      value = fitted_cr,
      variable = paste0("fitted_CR_sp_", i)
    )
  })

  # Combine all species data frames into one
  data_long_cr <- do.call(rbind, fitted_cr_list)

  # Extract community-level parameters
  gamma <- summary |> filter(grepl("gamma", variable))

  community_log_a_hat <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
  community_b_hat <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
  community_k_hat <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

  community_log_a <- exp(community_log_a_hat)
  community_b <- exp(community_b_hat)
  community_k <- exp(community_k_hat)

  community_fitted_log_y_cr <- community_log_a + log(1 - exp(-community_b * (x ^ community_k)))
  community_fitted_CR <- exp(community_fitted_log_y_cr)

  list(fitted_cr = fitted_cr_list, community_fitted_CR = community_fitted_CR, log_y = log_y, x = x, sp = sp, data_long_cr = data_long_cr)
}

# Apply function to CR data
cr_params <- extract_and_compute_fitted_cr(stan_data_nlr_cr, tar_read(fit_nlr_summary_weibull1_cr), dbh_range)

# Check the extracted and computed values
str(cr_params)
```

# Plot data

```{r}
# Data frame for plotting H
data_h <- data.frame(H = exp(h_params$log_y), DBH = h_params$x, sp = h_params$sp)
data_h$community_fitted_H <- h_params$community_fitted_H

# Check the data for plotting
head(h_params$data_long_h)

# Data frame for plotting CR
data_cr <- data.frame(CR = exp(cr_params$log_y), DBH = cr_params$x, sp = cr_params$sp)
data_cr$community_fitted_CR <- cr_params$community_fitted_CR

# Check the data for plotting
head(cr_params$data_long_cr)
```

# Plot
```{r}
# Plot H
p1 <- ggplot(data_h, aes(x = DBH, y = H)) +
  geom_line(data = h_params$data_long_h, aes(x = DBH, y = value, group = variable), color = "#72b6e3", alpha = 0.3, linewidth = 0.15) +
  geom_line(aes(y = community_fitted_H), color = "#0f92e9", linewidth = 0.6) +
  labs(x = "DBH (cm)", y = "Tree Height (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

# Plot CR
p2 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_line(data = cr_params$data_long_cr, aes(x = DBH, y = value, group = variable), color = "#72b6e3", alpha = 0.3, linewidth = 0.15) +
  geom_line(aes(y = community_fitted_CR), color = "#0f92e9", linewidth = 0.6) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

# Combine and save plots
p1 <- p1 +
  theme(
    axis.title.x = element_blank(),   # Remove the x-axis title
    axis.text.x = element_blank()     # Remove the x-axis text labels
  )

combined_plot <- p1 / p2 + 
  plot_layout(ncol = 1) &  
  theme(
    plot.margin = unit(c(10, 10, 10, 10), "pt"),  
    text = element_text(size = 9),               
    legend.title = element_text(size = 10)       
  )

ggsave(
  filename = "p.png",
  plot = combined_plot,
  width = 110,    
  height = 180,
  units = "mm",  
  dpi = 600     
)
```

```{r}

```

### 2025-03-26 docs/sp_para.qmd
Initial commit for tree-allo

---
title: "Extracting species-specific paramters"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
```

# Species id

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

wd_df <- data2 |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

wd_df

wd_df <- wd_df |>
  mutate(sp_id = 1:n()) |>
  select(-wd_s)
wd_df

```

# Extracting species-specific parameters

```{r}
options(width = 200)
source("R/functions.R")

# Best H model
wb_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
gamma_nou_h <- wb_nou_h |> filter(str_detect(variable, "^gamma\\["))
beta_nou_h <- wb_nou_h |> filter(str_detect(variable, "^beta\\["))

# Pl H model

pl_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)
gamma_nou_pl_h <- pl_nou_h |> filter(str_detect(variable, "^gamma\\["))
beta_nou_pl_h <- pl_nou_h |> filter(str_detect(variable, "^beta\\["))
gamma_nou_pl_h
beta_nou_pl_h

# Best CR model
wb_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
gamma_nou_cr <- wb_nou_cr |> filter(str_detect(variable, "^gamma\\["))
beta_nou_cr <- wb_nou_cr |> filter(str_detect(variable, "^beta\\["))

# Best DBH model
pl_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
gamma_dbh <- pl_nou_dbh |> filter(str_detect(variable, "^gamma\\["))
beta_dbh <- pl_nou_dbh |> filter(str_detect(variable, "^beta\\["))

species_specific_df <- generate_species_posterior_df(wb_nou_h, wb_nou_cr, pl_nou_dbh)
head(species_specific_df)
dim(species_specific_df)

cr_df <- species_specific_df |> filter(Dependent_variable == "Crown Radius")
h_df <- species_specific_df |> filter(Dependent_variable == "Tree Height")
dbh_df <- species_specific_df |> filter(Dependent_variable == "DBH")

# Crown Radius
cr_df_a <- cr_df |> filter(Parameter == "a")
cr_df_b <- cr_df |> filter(Parameter == "b")
cr_df_k <- cr_df |> filter(Parameter == "k")
cr_df <- data.frame(
  sp_id = seq_along(cr_df_a$CIs),  
  a = cr_df_a$CIs,
  b = cr_df_b$CIs,
  k = cr_df_k$CIs
)

# Tree Height
h_df_a <- h_df |> filter(Parameter == "a")
h_df_b <- h_df |> filter(Parameter == "b")
h_df_k <- h_df |> filter(Parameter == "k")
h_df <- data.frame(
  sp_id = seq_along(h_df_a$CIs),  
  a = h_df_a$CIs,
  b = h_df_b$CIs,
  k = h_df_k$CIs
)

# DBH
dbh_df_a <- dbh_df |> filter(Parameter == "a")
dbh_df_b <- dbh_df |> filter(Parameter == "b")
dbh_df_c <- dbh_df |> filter(Parameter == "c")
dbh_df <- data.frame(
  sp_id = seq_along(dbh_df_a$CIs),  
  a = dbh_df_a$CIs,
  b = dbh_df_b$CIs,
  c = dbh_df_c$CIs
)

posterior_df <- bind_rows(
  cr_df |> mutate(Dependent_variable = "Crown Radius", c = "-", k = k), 
  h_df |> mutate(Dependent_variable = "Tree Height", c = "-", k = k),
  dbh_df |> mutate(Dependent_variable = "DBH", k = "-", c = c)
)

posterior_df <- sp_df |> select(sp, Dependent_variable, everything())

head(posterior_df)
```

# Posterior df
```{r}
options(width = 150)
posterior_df <- posterior_df |> 
  mutate(sp_id = as.numeric(sp_id))

head(posterior_df)
wd_df

sp_posterior_df <- posterior_df |>
  left_join(wd_df, by = "sp_id") |>
  select(sp_id, sp, Dependent_variable, a, b, k, c, wd)

head(sp_posterior_df)
```

# Cheking the df of each variables

```{r}
# Split into three data frames
cr_df <- sp_posterior_df |> filter(Dependent_variable == "Crown Radius")
h_df <- sp_posterior_df |> filter(Dependent_variable == "Tree Height")
dbh_df <- sp_posterior_df |> filter(Dependent_variable == "DBH")
h_pl_df <- sp_posterior_df |> filter(Dependent_variable == "Tree Height")

head(cr_df)
head(h_df)
head(dbh_df)

options(width = 150)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
wb_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
pl_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)
wb_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
pl_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)

pl_nou_dbh2 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl_nou_dbh3 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl_nou_dbh2
pl_nou_dbh3

# source("R/functions.R")
sp_posterior_df <- generate_sp_posterior_df(wb_nou_h, wb_nou_cr, pl_nou_dbh, pl_nou_h, tallo_wd_df_200)

head(sp_posterior_df)
dim(sp_posterior_df)

sp_posterior_df <- tar_read(sp_posterior_df)
head(sp_posterior_df)
summary(sp_posterior_df)

# Split into 4 data frames
cr_df <- sp_posterior_df |> filter(Dependent_variable == "Crown Radius")
h_df <- sp_posterior_df |> filter(Dependent_variable == "Tree Height")
dbh_df <- sp_posterior_df |> filter(Dependent_variable == "DBH")
h_pl_df <- sp_posterior_df |> 
  filter(Dependent_variable == "Tree Height", is.na(k))

head(cr_df)
head(h_df)
head(dbh_df)
head(h_pl_df)

dim(h_pl_df)

```

### 2025-03-26 docs/wd_ef_CIs.qmd
Initial commit for tree-allo

---
title: "Wood density's effect"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
source("R/figs.R")
options(width = 150)
library(cowplot)

```

# H-DBH angiosperms

```{r}
tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

wd_h_ang_dat <- tallo_reduced_nlr_df_ang_h |>
  filter(division == "Angiosperm") |>
  dplyr::select(sp, dbh, h, cr, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_h_ang_dat

light_h_ang_wd_s <- quantile(wd_h_ang_dat$wd_s, 0.1, na.rm = TRUE)
dense_h_ang_wd_s <- quantile(wd_h_ang_dat$wd_s, 0.9, na.rm = TRUE)
light_h_ang_wd_s
dense_h_ang_wd_s

fit_wd_ang_h_draws_weibull_wd <- tar_read(fit_wd_ang_h_draws_weibull_wd)
fit_wd_ang_h_draws_weibull_wd

gamma_params_h <- fit_wd_ang_h_draws_weibull_wd |>
  dplyr::select(starts_with("gamma_int"), starts_with("gamma_slope"))

gamma_params_h

median_a <- median(exp(gamma_params_h$`gamma_int[1]`), na.rm = TRUE)
median_b <- median(gamma_params_h$`gamma_int[2]`, na.rm = TRUE)

gamma_params_h <- gamma_params_h |>
  mutate(
    k_light = `gamma_int[3]` + `gamma_slope[3]` * light_h_ang_wd_s,
    k_dense = `gamma_int[3]` + `gamma_slope[3]` * dense_h_ang_wd_s
  )
gamma_params_h

dbh_seq <- seq(1, 150, length.out = 100)

pred_h <- gamma_params_h |>
  rowwise() |>
  mutate(
    h_light = list(tibble(
      dbh = dbh_seq,
      h_light = median_a * (1 - exp(-median_b * dbh^k_light))
    )),
    h_dense = list(tibble(
      dbh = dbh_seq,
      h_dense = median_a * (1 - exp(-median_b * dbh^k_dense))
    ))
  ) |>
  select(h_light, h_dense) |>
  unnest(cols = c(h_light, h_dense), names_sep = "_") |>
  rename(
    dbh = h_light_dbh,
    h_light = h_light_h_light,
    h_dense_dbh = h_dense_dbh,
    h_dense = h_dense_h_dense
  )

pred_h
summary(pred_h)
str(pred_h)

h_ang_light <- pred_h |>
  group_by(dbh) |>
  summarize(
    h_median_light = median(h_light),
    h_lower_light = quantile(h_light, 0.025),
    h_upper_light = quantile(h_light, 0.975),
    .groups = "drop"
  )

h_ang_light

h_ang_dense <- pred_h |>
  group_by(dbh) |>
  summarize(
    h_median_dense = median(h_dense),
    h_lower_dense = quantile(h_dense, 0.025),
    h_upper_dense = quantile(h_dense, 0.975),
    .groups = "drop"
  )

h_ang_dense

h_ang_dat <- left_join(
  h_ang_light,
  h_ang_dense,
  by = "dbh"
)
h_ang_dat

p1 <- ggplot(h_ang_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = h_lower_light, ymax = h_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y = h_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = h_lower_dense, ymax = h_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = h_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Height (m)") +
  my_theme()

p1

```

# CR-DBH angiosperms

```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

wd_cr_ang_dat <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  dplyr::select(sp, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_cr_ang_dat

light_cr_ang_wd_s <- quantile(wd_cr_ang_dat$wd_s, 0.1, na.rm = TRUE)
dense_cr_ang_wd_s <- quantile(wd_cr_ang_dat$wd_s, 0.9, na.rm = TRUE)
light_cr_ang_wd_s
dense_cr_ang_wd_s

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

gamma_params_cr_ang <- fit_wd_ang_cr_draws_pl_wd |>
  dplyr::select(starts_with("gamma"))

gamma_params_cr_ang

gamma_params_cr_ang <- gamma_params_cr_ang |>
  mutate(
    a_light = exp(`gamma[1,1]` + `gamma[1,2]` * light_cr_ang_wd_s),
    a_dense = exp(`gamma[1,1]` + `gamma[1,2]` * dense_cr_ang_wd_s),
    b_light = `gamma[2,1]` + `gamma[2,2]` * light_cr_ang_wd_s,
    b_dense = `gamma[2,1]` + `gamma[2,2]` * dense_cr_ang_wd_s
  )

gamma_params_cr_ang

dbh_cr_ang_seq <- seq(1, 150, length.out = 100)

pred_ang_cr <- gamma_params_cr_ang |>
  rowwise() |>
  mutate(
    cr_light = list(tibble(
      dbh = dbh_cr_ang_seq,
      cr_light = a_light * dbh^b_light
    )),
    cr_dense = list(tibble(
      dbh = dbh_cr_ang_seq,
      cr_dense = a_dense * dbh^b_dense
    ))
  ) |>
  select(cr_light, cr_dense) |>
  unnest(cols = c(cr_light, cr_dense), names_sep = "_") |>
  rename(
    dbh = cr_light_dbh,
    cr_light = cr_light_cr_light,
    cr_dense_dbh = cr_dense_dbh,
    cr_dense = cr_dense_cr_dense
  )

pred_ang_cr
summary(pred_ang_cr)
str(pred_ang_cr)

cr_ang_light <- pred_ang_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_light = median(cr_light),
    cr_lower_light = quantile(cr_light, 0.025),
    cr_upper_light = quantile(cr_light, 0.975),
    .groups = "drop"
  )

cr_ang_light

cr_ang_dense <- pred_ang_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_dense = median(cr_dense),
    cr_lower_dense = quantile(cr_dense, 0.025),
    cr_upper_dense = quantile(cr_dense, 0.975),
    .groups = "drop"
  )

cr_ang_dense

cr_ang_dat <- left_join(
  cr_ang_light,
  cr_ang_dense,
  by = "dbh"
)

cr_ang_dat

p2 <- ggplot(cr_ang_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = cr_lower_light, ymax = cr_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y =  cr_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = cr_lower_dense, ymax = cr_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = cr_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)") +
  my_theme()

p2
```

# CR-DBH gymnosperms

```{r}
tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)
tallo_reduced_nlr_df_gym_cr

wd_cr_gym_dat <- tallo_reduced_nlr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  dplyr::select(sp, dbh, h, cr, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_cr_gym_dat

light_cr_gym_wd_s <- quantile(wd_cr_gym_dat$wd_s, 0.1, na.rm = TRUE)
dense_cr_gym_wd_s <- quantile(wd_cr_gym_dat$wd_s, 0.9, na.rm = TRUE)
light_cr_gym_wd_s
dense_cr_gym_wd_s

fit_wd_gym_cr_draws_gmm_wd <- tar_read(fit_wd_gym_cr_draws_gmm_wd)
fit_wd_gym_cr_draws_gmm_wd

gamma_params_cr_gym <- fit_wd_gym_cr_draws_gmm_wd |>
  dplyr::select(starts_with("gamma_int"), starts_with("gamma_slope"))

gamma_params_cr_gym

median_cr_gym_a <- median(exp(gamma_params_cr_gym$`gamma_int[1]`), na.rm = TRUE)
median_cr_gym_k <- median(gamma_params_cr_gym$`gamma_int[3]`, na.rm = TRUE)

gamma_params_cr_gym <- gamma_params_cr_gym |>
  mutate(
    b_light = `gamma_int[2]` + `gamma_slope[2]` * light_cr_gym_wd_s,
    b_dense = `gamma_int[2]` + `gamma_slope[2]` * dense_cr_gym_wd_s
  )
gamma_params_cr_gym

dbh_cr_gym_seq <- seq(1, 150, length.out = 100)

pred_gym_cr <- gamma_params_cr_gym |>
  rowwise() |>
  mutate(
    cr_light = list(tibble(
      dbh = dbh_cr_gym_seq,
      cr_light = median_cr_gym_a * dbh^b_light / (median_cr_gym_k + dbh^b_light)
    )),
    cr_dense = list(tibble(
      dbh = dbh_cr_gym_seq,
      cr_dense = median_cr_gym_a * dbh^b_dense / (median_cr_gym_k + dbh^b_dense)
    ))
  ) |>
  select(cr_light, cr_dense) |>
  unnest(cols = c(cr_light, cr_dense), names_sep = "_") |>
  rename(
    dbh = cr_light_dbh,
    cr_light = cr_light_cr_light,
    cr_dense_dbh = cr_dense_dbh,
    cr_dense = cr_dense_cr_dense
  )

pred_gym_cr
summary(pred_gym_cr)
str(pred_gym_cr)

cr_gym_light <- pred_gym_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_light = median(cr_light),
    cr_lower_light = quantile(cr_light, 0.025),
    cr_upper_light = quantile(cr_light, 0.975),
    .groups = "drop"
  )

cr_gym_light

cr_gym_dense <- pred_gym_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_dense = median(cr_dense),
    cr_lower_dense = quantile(cr_dense, 0.025),
    cr_upper_dense = quantile(cr_dense, 0.975),
    .groups = "drop"
  )

pred_gym_cr

cr_gym_dat <- left_join(
  cr_gym_light,
  cr_gym_dense,
  by = "dbh"
)

cr_gym_dat

p3 <- ggplot(cr_gym_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = cr_lower_light, ymax = cr_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y =  cr_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = cr_lower_dense, ymax = cr_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = cr_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)") +
  my_theme()

p3
```

# Combine in 1 plot

```{r}
p <- p1 + p2 + p3 + plot_layout(nrow = 1, guides = "collect") +
      plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)"))) &
      theme(
      plot.tag = element_text(size = 8),
      axis.title = element_text(size = 9),
      text = element_text(size = 8),
      legend.title = element_blank(),
      plot.margin = margin(t = 2, r = 2, b = 2, l = 2, unit = "pt")
    )

p <- ggdraw() +
draw_plot(p, x = 0, y = 0, width = 1, height = 0.95) +  
draw_label("Angiosperm", x = 0.172, y = 0.94, fontface = 'bold', size = 10) +  
draw_label("Angiosperm", x = 0.429, y = 0.94, fontface = 'bold', size = 10) +  
draw_label("Gymnosperm", x = 0.689, y = 0.94, fontface = 'bold', size = 10)

```

# Writing function

```{r}
process_wd_data <- function(df, division, light_quantile = 0.1, dense_quantile = 0.9) {
  df_filtered <- df |>
    filter(division == division) |>
    dplyr::select(sp, dbh, h, cr, wd) |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

  light_wd_s <- quantile(df_filtered$wd_s, light_quantile, na.rm = TRUE)
  dense_wd_s <- quantile(df_filtered$wd_s, dense_quantile, na.rm = TRUE)

  return(list(df = df_filtered, light_wd_s = light_wd_s, dense_wd_s = dense_wd_s))
}

```

```{r}

wd_h_ang_dat <- process_wd_data(tallo_reduced_nlr_df_ang_h)
wd_h_ang_dat

wd_cr_ang_dat <- process_wd_data(tallo_reduced_lr_df_ang_cr)
wd_cr_ang_dat

wd_cr_gym_dat <- process_wd_data(tallo_reduced_nlr_df_gym_cr)
wd_cr_gym_dat

```

### 2025-03-26 docs/wd_ef_check.qmd
Initial commit for tree-allo

---
title: "Checking Wood density's effect plot"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
source("R/figs.R")
options(width = 150)

```

# WD vs. CR-DBH

## Angiosperm

### No CIs
```{r}
sp_posterior_cr_ang_df_wd <- tar_read(sp_posterior_cr_ang_df_wd)
head(sp_posterior_cr_ang_df_wd)

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_cr_post <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_cr_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b) |>
  mutate(
    a = as.numeric(sub(" .*", "", a)),
    b = as.numeric(sub(" .*", "", b))
  )

ang_cr_post
summary(ang_cr_post)

wd_quantiles <- quantile(ang_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)

light_wood <- ang_cr_post |>
  filter(wd < wd_quantiles[1])

dense_wood <- ang_cr_post |>
  filter(wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)

mean_light_wood <- light_wood |>
  summarise(across(c(a, b), mean, na.rm = TRUE))

mean_dense_wood <- dense_wood |>
  summarise(across(c(a, b), mean, na.rm = TRUE))

predicted_cr_light <- mean_light_wood$a * dbh_seq^mean_light_wood$b
predicted_cr_dense <- mean_dense_wood$a * dbh_seq^mean_dense_wood$b

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_cr, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
sp_posterior_cr_ang_df_wd <- tar_read(sp_posterior_cr_ang_df_wd)
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_cr_post <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_cr_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))"))
  ) |>
  dplyr::select(-a, - b)

ang_cr_post
summary(ang_cr_post)

light_wood <- ang_cr_post |>
  filter(wd < quantile(ang_cr_post$wd, probs = 0.1, na.rm = TRUE))

dense_wood <- ang_cr_post |>
  filter(wd > quantile(ang_cr_post$wd, probs = 0.9, na.rm = TRUE))

mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE)
  )

dbh_seq <- seq(1, 100, length.out = 100)

predicted_cr_light <- mean_light_wood$a_median * dbh_seq^mean_light_wood$b_median
predicted_cr_dense <- mean_dense_wood$a_median * dbh_seq^mean_dense_wood$b_median

# Calculate CI bounds for both wood types using lower and upper values of a and b
predicted_cr_light_lower <- mean_light_wood$a_lower * dbh_seq^mean_light_wood$b_lower
predicted_cr_light_upper <- mean_light_wood$a_upper * dbh_seq^mean_light_wood$b_upper

predicted_cr_dense_lower <- mean_dense_wood$a_lower * dbh_seq^mean_dense_wood$b_lower
predicted_cr_dense_upper <- mean_dense_wood$a_upper * dbh_seq^mean_dense_wood$b_upper

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  predicted_cr_lower = c(predicted_cr_light_lower, predicted_cr_dense_lower),
  predicted_cr_upper = c(predicted_cr_light_upper, predicted_cr_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

head(prediction_df_mean)

ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_cr_lower, ymax = predicted_cr_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_cr, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()

```

## Gymnosperms

### No CIs

```{r}
sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
head(sp_posterior_cr_gym_df_wd)

tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_df <- tallo_reduced_lr_df_gym_cr |> filter(division == "Gymnosperm")
gym_cr_df

gym_cr_post <- gym_cr_df |>
  left_join(sp_posterior_cr_gym_df_wd, by = c("sp" = "sp")) |>
  dplyr::select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(
    a = as.numeric(sub(" .*", "", a)),
    b = as.numeric(sub(" .*", "", b)),
    k = as.numeric(sub(" .*", "", k))
  )

gym_cr_post
summary(gym_cr_post)

sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_post <- tallo_reduced_lr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  left_join(sp_posterior_cr_gym_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(across(c(a, b, k), ~ as.numeric(sub(" .*", "", .))))

wd_quantiles <- quantile(gym_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)
light_wood <- gym_cr_post |> filter(wd < wd_quantiles[1])
dense_wood <- gym_cr_post |> filter(wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)

mean_light_wood <- light_wood |> summarise(across(c(a, b, k), mean, na.rm = TRUE))
mean_dense_wood <- dense_wood |> summarise(across(c(a, b, k), mean, na.rm = TRUE))

predicted_cr_light <- (mean_light_wood$a * dbh_seq^mean_light_wood$b) / (mean_light_wood$k + dbh_seq^mean_light_wood$b)
predicted_cr_dense <- (mean_dense_wood$a * dbh_seq^mean_dense_wood$b) / (mean_dense_wood$k + dbh_seq^mean_dense_wood$b)

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_cr, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

# tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_post <- tallo_reduced_lr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  left_join(sp_posterior_cr_gym_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),
    k_median = as.numeric(str_extract(k, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),
    k_lower = as.numeric(str_extract(k, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))")),
    k_upper = as.numeric(str_extract(k, "(?<=,).*?(?=\\))"))
  ) |>
  select(-a, -b, -k)

gym_cr_post
wd_quantiles <- quantile(gym_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)
light_wood <- gym_cr_post |> filter(wd < wd_quantiles[1])
dense_wood <- gym_cr_post |> filter(wd > wd_quantiles[2])

mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

dbh_seq <- seq(1, 150, length.out = 100)

predicted_cr_light <- (mean_light_wood$a_median * dbh_seq^mean_light_wood$b_median) / 
  (mean_light_wood$k_median + dbh_seq^mean_light_wood$b_median)

predicted_cr_dense <- (mean_dense_wood$a_median * dbh_seq^mean_dense_wood$b_median) / 
  (mean_dense_wood$k_median + dbh_seq^mean_dense_wood$b_median)

predicted_cr_light_lower <- (mean_light_wood$a_lower * dbh_seq^mean_light_wood$b_lower) / 
  (mean_light_wood$k_lower + dbh_seq^mean_light_wood$b_lower)

predicted_cr_light_upper <- (mean_light_wood$a_upper * dbh_seq^mean_light_wood$b_upper) / 
  (mean_light_wood$k_upper + dbh_seq^mean_light_wood$b_upper)

predicted_cr_dense_lower <- (mean_dense_wood$a_lower * dbh_seq^mean_dense_wood$b_lower) / 
  (mean_dense_wood$k_lower + dbh_seq^mean_dense_wood$b_lower)

predicted_cr_dense_upper <- (mean_dense_wood$a_upper * dbh_seq^mean_dense_wood$b_upper) / 
  (mean_dense_wood$k_upper + dbh_seq^mean_dense_wood$b_upper)

# Combine into a single data frame
prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  predicted_cr_lower = c(predicted_cr_light_lower, predicted_cr_dense_lower),
  predicted_cr_upper = c(predicted_cr_light_upper, predicted_cr_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_cr_lower, ymax = predicted_cr_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_cr, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()
```

# WD vs. H-DBH

## Angiosperm

### No CIs
```{r}
sp_posterior_h_ang_df_wd <- tar_read(sp_posterior_h_ang_df_wd)
# tallo_reduced_lr_df_h_ang <- tar_read(tallo_reduced_nlr_df_ang_h)

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)

ang_h_post <- tallo_reduced_lr_df_h_ang |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_h_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, wd, a, b, k) |>
  mutate(across(c(a, b, k), ~ as.numeric(sub(" .*", "", .))))

wd_quantiles <- quantile(ang_h_post$wd, probs = c(0.25, 0.75), na.rm = TRUE)
light_wood <- filter(ang_h_post, wd < wd_quantiles[1])
dense_wood <- filter(ang_h_post, wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)
mean_light_wood <- summarise(light_wood, across(c(a, b, k), mean))
mean_dense_wood <- summarise(dense_wood, across(c(a, b, k), mean))

predicted_h_light <- mean_light_wood$a * (1 - exp(-mean_light_wood$b * dbh_seq^mean_light_wood$k))
predicted_h_dense <- mean_dense_wood$a * (1 - exp(-mean_dense_wood$b * dbh_seq^mean_dense_wood$k))

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_h = c(predicted_h_light, predicted_h_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_h, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Tree Height (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
# Load data
sp_posterior_h_ang_df_wd <- tar_read(sp_posterior_h_ang_df_wd)
tallo_reduced_lr_df_h_ang <- tar_read(tallo_reduced_nlr_df_ang_h)

# Process data and parse CIs
ang_h_post <- tallo_reduced_lr_df_h_ang |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_h_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, wd, a, b, k) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),
    k_median = as.numeric(str_extract(k, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),
    k_lower = as.numeric(str_extract(k, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))")),
    k_upper = as.numeric(str_extract(k, "(?<=,).*?(?=\\))"))
  ) |>
  select(-a, -b, -k)

# Quantiles for Wood Density
wd_quantiles <- quantile(ang_h_post$wd, probs = c(0.25, 0.75), na.rm = TRUE)
light_wood <- filter(ang_h_post, wd < wd_quantiles[1])
dense_wood <- filter(ang_h_post, wd > wd_quantiles[2])

# Summary statistics
mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

# Generate predictions
dbh_seq <- seq(1, 150, length.out = 100)

# Median predictions
predicted_h_light <- mean_light_wood$a_median * (1 - exp(-mean_light_wood$b_median * dbh_seq^mean_light_wood$k_median))
predicted_h_dense <- mean_dense_wood$a_median * (1 - exp(-mean_dense_wood$b_median * dbh_seq^mean_dense_wood$k_median))

# Lower and upper bounds
predicted_h_light_lower <- mean_light_wood$a_lower * (1 - exp(-mean_light_wood$b_lower * dbh_seq^mean_light_wood$k_lower))
predicted_h_light_upper <- mean_light_wood$a_upper * (1 - exp(-mean_light_wood$b_upper * dbh_seq^mean_light_wood$k_upper))

predicted_h_dense_lower <- mean_dense_wood$a_lower * (1 - exp(-mean_dense_wood$b_lower * dbh_seq^mean_dense_wood$k_lower))
predicted_h_dense_upper <- mean_dense_wood$a_upper * (1 - exp(-mean_dense_wood$b_upper * dbh_seq^mean_dense_wood$k_upper))

# Combine into a single data frame
prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_h = c(predicted_h_light, predicted_h_dense),
  predicted_h_lower = c(predicted_h_light_lower, predicted_h_dense_lower),
  predicted_h_upper = c(predicted_h_light_upper, predicted_h_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

# Plot with CIs
ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_h_lower, ymax = predicted_h_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_h, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Tree Height (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()
```

### 2025-03-26 docs/wd_para.qmd
Initial commit for tree-allo

---
title: "Relationship between species-specific H/CR allometry parameters with wood density in the best-fit hierarchical model (Weibull function)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)
```

# Wood density data preparation
```{r}
options(width = 150)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

wd_df <- data2 |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

wd_df
dim(wd_df)
wd_df <- wd_df |>
  mutate(sp_id = 1:n())

head(wd_df)
dim(wd_df)
```

# H-DBH

## Extract a, b, k

```{r}
# Load the necessary data
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
wd_s <- stan_data_nlr_h$u[, 2]
wb1_h_summary <- tar_read(fit_nlr_summary_weibull1_h)
wb1_h_summary

# Extract the beta parameters (species-specific) with 95% CI
beta <- wb1_h_summary |> filter(grepl("beta\\[", variable))

# Extract the beta parameters (species-specific) with 95% CI
beta <- wb1_h_summary |> filter(grepl("beta\\[", variable))

# Calculate mean and 90% CI for a, b, k
a_mean <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
a_lower <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))
a_upper <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))

b_mean <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
b_lower <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)
b_upper <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)

k_mean <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
k_lower <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)
k_upper <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)

# Extract community-level parameters and calculate fitted values
gamma <- wb1_h_summary |> filter(grepl("gamma", variable))
gamma

# Correct variable names
gamma_log_a_intercept_median <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
gamma_log_a_wd_median <- gamma |> filter(variable == "gamma[2,1]") |> pull(q50)
gamma_b_intercept_median <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
gamma_b_wd_median <- gamma |> filter(variable == "gamma[2,2]") |> pull(q50)
gamma_k_intercept_median <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)
gamma_k_wd_median <- gamma |> filter(variable == "gamma[2,3]") |> pull(q50)

# Compute community fitted values using median (q50)
log_community_fitted_a_median <- gamma_log_a_intercept_median + gamma_log_a_wd_median * wd_s
community_fitted_a_median <- exp(log_community_fitted_a_median)

log_community_fitted_b_median <- gamma_b_intercept_median + gamma_b_wd_median * wd_s
community_fitted_b_median <- exp(log_community_fitted_b_median)

log_community_fitted_k_median <- gamma_k_intercept_median + gamma_k_wd_median * wd_s
community_fitted_k_median <- exp(log_community_fitted_k_median)

# Calculate 90% credible intervals for community fitted values
gamma_log_a_intercept_lower <- gamma |> filter(variable == "gamma[1,1]") |> pull(q5)
gamma_log_a_wd_lower <- gamma |> filter(variable == "gamma[2,1]") |> pull(q5)

gamma_log_a_intercept_upper <- gamma |> filter(variable == "gamma[1,1]") |> pull(q95)
gamma_log_a_wd_upper <- gamma |> filter(variable == "gamma[2,1]") |> pull(q95)

gamma_b_intercept_lower <- gamma |> filter(variable == "gamma[1,2]") |> pull(q5)
gamma_b_wd_lower <- gamma |> filter(variable == "gamma[2,2]") |> pull(q5)

gamma_b_intercept_upper <- gamma |> filter(variable == "gamma[1,2]") |> pull(q95)
gamma_b_wd_upper <- gamma |> filter(variable == "gamma[2,2]") |> pull(q95)

gamma_k_intercept_lower <- gamma |> filter(variable == "gamma[1,3]") |> pull(q5)
gamma_k_wd_lower <- gamma |> filter(variable == "gamma[2,3]") |> pull(q5)

gamma_k_intercept_upper <- gamma |> filter(variable == "gamma[1,3]") |> pull(q95)
gamma_k_wd_upper <- gamma |> filter(variable == "gamma[2,3]") |> pull(q95)

# Compute lower and upper bounds for community fitted values
log_log_community_fitted_a_lower <- gamma_log_a_intercept_lower + gamma_log_a_wd_lower * wd_s
log_log_community_fitted_a_upper <- gamma_log_a_intercept_upper + gamma_log_a_wd_upper * wd_s
community_fitted_a_lower <- exp(log_log_community_fitted_a_lower)
community_fitted_a_upper <- exp(log_log_community_fitted_a_upper)
log_community_fitted_b_lower <- gamma_b_intercept_lower + gamma_b_wd_lower * wd_s
log_community_fitted_b_upper <- gamma_b_intercept_upper + gamma_b_wd_upper * wd_s
community_fitted_b_lower <- exp(log_community_fitted_b_lower)
community_fitted_b_upper <- exp(log_community_fitted_b_upper)
log_community_fitted_k_lower <- gamma_k_intercept_lower + gamma_k_wd_lower * wd_s
log_community_fitted_k_upper <- gamma_k_intercept_upper + gamma_k_wd_upper * wd_s
community_fitted_k_lower <- exp(log_community_fitted_k_lower)
community_fitted_k_upper <- exp(log_community_fitted_k_upper)

# Prepare data for plotting
sp_id <- as.factor(unique(stan_data_nlr_h$jj))
h_wd_data <- data.frame(
  sp_id = sp_id,
  wd_s = wd_s,
  a_mean = a_mean,
  a_lower = a_lower,
  a_upper = a_upper,
  b_mean = b_mean,
  b_lower = b_lower,
  b_upper = b_upper,
  k_mean = k_mean,
  k_lower = k_lower,
  k_upper = k_upper,
  community_fitted_a_median = community_fitted_a_median,
  community_fitted_b_median = community_fitted_b_median,
  community_fitted_k_median = community_fitted_k_median,
  community_fitted_a_lower = community_fitted_a_lower,
  community_fitted_a_upper = community_fitted_a_upper,
  community_fitted_b_lower = community_fitted_b_lower,
  community_fitted_b_upper = community_fitted_b_upper,
  community_fitted_k_lower = community_fitted_k_lower,
  community_fitted_k_upper = community_fitted_k_upper
)

head(h_wd_data)

h_wd_data$sp_id <- as.integer(as.character(h_wd_data$sp_id))
h_wd_data <- inner_join(h_wd_data, wd_df, by = "sp_id")
head(h_wd_data)
```

## Plotting

```{r}
plot_h_a <- ggplot(h_wd_data, aes(x = wd, y = a_mean)) +
  geom_errorbar(aes(ymin = a_lower, ymax = a_upper), color = "lightgrey", width = 0.0001) +
  geom_point(color = "darkgrey", size = 1) +
  geom_ribbon(aes(ymin = community_fitted_a_lower, ymax = community_fitted_a_upper), fill = "grey80", alpha = 0.5) +
  geom_line(aes(y = community_fitted_a_median), color = "black", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Intercept, a") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

plot_h_b <- ggplot(h_wd_data, aes(x = wd, y = b_mean)) +
  geom_errorbar(aes(ymin = b_lower, ymax = b_upper), color = "orange", width = 0.0001) +
  geom_point(color = "orange", size = 1) +
  geom_ribbon(aes(ymin = community_fitted_b_lower, ymax = community_fitted_b_upper), fill = "orange", alpha = 0.5) +
  geom_line(aes(y = community_fitted_b_median), color = "darkorange", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Exponent, b") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

plot_h_k <- ggplot(h_wd_data, aes(x = wd, y = k_mean)) +
  geom_errorbar(aes(ymin = k_lower, ymax = k_upper), color = "blue", width = 0.0001) +
  geom_point(color = "blue", size = 1) +
  # geom_ribbon(aes(ymin = community_fitted_k_lower, ymax = community_fitted_k_upper), fill = "lightblue", alpha = 0.5) +
  geom_line(aes(y = community_fitted_k_median), color = "darkblue", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Inflection parameter, k") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

# Print plots
plot_h_a
plot_h_b
plot_h_k

```

# CR-DBH

## Extract a, b, k
```{r}
# Load the necessary data
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

wd_s <- stan_data_nlr_cr$u[, 2]
wb1_cr_summary <- tar_read(fit_nlr_summary_weibull1_cr)
wb1_cr_summary

# Extract the beta parameters (species-specific) with 90% CI
beta <- wb1_cr_summary |> filter(grepl("beta\\[", variable))
beta

# Calculate mean and 90% CI for a, b, k
a_mean <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
a_lower <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q5))
a_upper <- exp(beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q95))

b_mean <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
b_lower <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q5)
b_upper <- beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q95)

k_mean <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
k_lower <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q5)
k_upper <- beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q95)

# Extract community-level parameters and calculate fitted values
gamma <- wb1_cr_summary |> filter(grepl("gamma", variable))

gamma

gamma_log_a_intercept <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
gamma_log_a_wd <- gamma |> filter(variable == "gamma[2,1]") |> pull(q50)
gamma_b_intercept <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
gamma_b_wd <- gamma |> filter(variable == "gamma[2,2]") |> pull(q50)
gamma_k_intercept <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)
gamma_k_wd <- gamma |> filter(variable == "gamma[2,3]") |> pull(q50)

# Compute community fitted values
log_log_community_fitted_a <- gamma_log_a_intercept + gamma_log_a_wd * wd_s
community_fitted_a <- exp(exp(log_log_community_fitted_a))

log_community_fitted_b <- gamma_b_intercept + gamma_b_wd * wd_s
community_fitted_b <- exp(log_community_fitted_b)
log_community_fitted_k <- gamma_k_intercept + gamma_k_wd * wd_s
community_fitted_k <- exp(log_community_fitted_k)

# Calculate 90% credible intervals for community fitted values
gamma_log_a_intercept_lower <- gamma |> filter(variable == "gamma[1,1]") |> pull(q5)
gamma_log_a_wd_lower <- gamma |> filter(variable == "gamma[2,1]") |> pull(q5)

gamma_log_a_intercept_upper <- gamma |> filter(variable == "gamma[1,1]") |> pull(q95)
gamma_log_a_wd_upper <- gamma |> filter(variable == "gamma[2,1]") |> pull(q95)

gamma_b_intercept_lower <- gamma |> filter(variable == "gamma[1,2]") |> pull(q5)
gamma_b_wd_lower <- gamma |> filter(variable == "gamma[2,2]") |> pull(q5)

gamma_b_intercept_upper <- gamma |> filter(variable == "gamma[1,2]") |> pull(q95)
gamma_b_wd_upper <- gamma |> filter(variable == "gamma[2,2]") |> pull(q95)

gamma_k_intercept_lower <- gamma |> filter(variable == "gamma[1,3]") |> pull(q5)
gamma_k_wd_lower <- gamma |> filter(variable == "gamma[2,3]") |> pull(q5)

gamma_k_intercept_upper <- gamma |> filter(variable == "gamma[1,3]") |> pull(q95)
gamma_k_wd_upper <- gamma |> filter(variable == "gamma[2,3]") |> pull(q95)

# Compute lower and upper bounds for community fitted values
log_log_community_fitted_a_lower <- gamma_log_a_intercept_lower + gamma_log_a_wd_lower * wd_s
log_log_community_fitted_a_upper <- gamma_log_a_intercept_upper + gamma_log_a_wd_upper * wd_s
community_fitted_a_lower <- exp(exp(log_log_community_fitted_a_lower))
community_fitted_a_upper <- exp(exp(log_log_community_fitted_a_upper))
log_community_fitted_b_lower <- gamma_b_intercept_lower + gamma_b_wd_lower * wd_s
log_community_fitted_b_upper <- gamma_b_intercept_upper + gamma_b_wd_upper * wd_s
community_fitted_b_lower <- exp(log_community_fitted_b_lower)
community_fitted_b_upper <- exp(log_community_fitted_b_upper)
log_community_fitted_k_lower <- gamma_k_intercept_lower + gamma_k_wd_lower * wd_s
log_community_fitted_k_upper <- gamma_k_intercept_upper + gamma_k_wd_upper * wd_s
community_fitted_k_lower <- exp(log_community_fitted_k_lower)
community_fitted_k_upper <- exp(log_community_fitted_k_upper)

# Prepare data for plotting
sp_id <- as.factor(unique(stan_data_nlr_cr$jj))
cr_wd_data <- data.frame(
  sp_id = sp_id,
  wd_s = wd_s,
  a_mean = a_mean,
  a_lower = a_lower,
  a_upper = a_upper,
  b_mean = b_mean,
  b_lower = b_lower,
  b_upper = b_upper,
  k_mean = k_mean,
  k_lower = k_lower,
  k_upper = k_upper,
  community_fitted_a = community_fitted_a,
  community_fitted_b = community_fitted_b,
  community_fitted_k = community_fitted_k,
  community_fitted_a_lower = community_fitted_a_lower,
  community_fitted_a_upper = community_fitted_a_upper,
  community_fitted_b_lower = community_fitted_b_lower,
  community_fitted_b_upper = community_fitted_b_upper,
  community_fitted_k_lower = community_fitted_k_lower,
  community_fitted_k_upper = community_fitted_k_upper
)
head(cr_wd_data)

# Join with wood density data
cr_wd_data$sp_id <- as.integer(as.character(cr_wd_data$sp_id))
cr_wd_data <- inner_join(cr_wd_data, wd_df, by = "sp_id")
head(cr_wd_data)

```

## Plotting

```{r}
plot_cr_a <- ggplot(cr_wd_data, aes(x = wd, y = a_mean)) +
  geom_errorbar(aes(ymin = a_lower, ymax = a_upper), color = "lightgrey", width = 0.0001) +
  geom_point(color = "darkgrey", size = 1) +
  # geom_ribbon(aes(ymin = community_fitted_a_lower, ymax = community_fitted_a_upper), fill = "grey80", alpha = 0.5) +
  geom_line(aes(y = community_fitted_a), color = "black", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Intercept, a") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

plot_cr_b <- ggplot(cr_wd_data, aes(x = wd, y = b_mean)) +
  geom_errorbar(aes(ymin = b_lower, ymax = b_upper), color = "orange", width = 0.0001) +
  geom_point(color = "orange", size = 1) +
  # geom_ribbon(aes(ymin = community_fitted_b_lower, ymax = community_fitted_b_upper), fill = "orange", alpha = 0.5) +
  geom_point(color = "orange", size = 2) +
  geom_line(aes(y = community_fitted_b), color = "darkorange", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Exponent, b") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

plot_cr_k <- ggplot(cr_wd_data, aes(x = wd, y = k_mean)) +
  geom_errorbar(aes(ymin = k_lower, ymax = k_upper), color = "blue", width = 0.0001) +
  geom_point(color = "blue", size = 1) +
  # geom_ribbon(aes(ymin = community_fitted_k_lower, ymax = community_fitted_k_upper), fill = "lightblue", alpha = 0.5) +
  geom_line(aes(y = community_fitted_k), color = "darkblue", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Inflection parameter, k") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

# Print plots
plot_cr_a
plot_cr_b
plot_cr_k
```

## Combine the figures

```{r}

wd_plots <- (plot_h_a + labs(x = NULL, tag = "(a)") |
                   plot_h_b + labs(x = NULL, tag = "(b)") |
                   plot_h_k + labs(x = NULL, tag = "(c)")) / 
                  (plot_cr_a + labs(tag = "(d)") |
                   plot_cr_b + labs(tag = "(e)") |
                   plot_cr_k + labs(tag = "(f)")) +
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"))) &
  theme(plot.tag = element_text(face = "bold"))

wd_plots

ggsave("wd_para_plot.png", plot = wd_plots, width = 8, height = 6, dpi = 300, bg = "white")

wd_plots <- (plot_h_a + labs(x = NULL, tag = "(a)") |
                   plot_h_b + labs(x = NULL, tag = "(b)") |
                   plot_h_k + labs(x = NULL, tag = "(c)")) / 
                  (plot_cr_a + labs(tag = "(d)") |
                   plot_cr_b + labs(tag = "(e)") |
                   plot_cr_k + labs(tag = "(f)")) +
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"))) &
  theme(
    plot.tag = element_text(face = "bold", size = 28),  # Increase tag size
    axis.title.x = element_text(size = 24),             # Increase x-axis title size
    axis.title.y = element_text(size = 24),             # Increase y-axis title size
    axis.text.x = element_text(size = 24),              # Increase x-axis number size
    axis.text.y = element_text(size = 24)               # Increase y-axis number size
  )

wd_plots

# Save the plot with the updated theme and larger dimensions
ggsave("wd_para_plot_large_axes.png", plot = wd_plots, width = 28, height = 18, dpi = 300, bg = "white")

# Combine the top row of plots without removing x-axis labels
wd_plots_abc <- (plot_h_a + labs(tag = "(a)") |
                 plot_h_b + labs(tag = "(b)") |
                 plot_h_k + labs(tag = "(c)")) +
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)"))) &
  theme(
    plot.tag = element_text(face = "bold", size = 28), 
    axis.title.x = element_text(size = 28),           
    axis.title.y = element_text(size = 28),           
    axis.text.x = element_text(size = 24),            
    axis.text.y = element_text(size = 24)             
  )

# Save the top row as a separate figure
ggsave("wd_para_plot_abc.png", plot = wd_plots_abc, width = 26, height = 9, dpi = 300, bg = "white")

# Combine the bottom row of plots
wd_plots_def <- (plot_cr_a + labs(tag = "(d)") |
                 plot_cr_b + labs(tag = "(e)") |
                 plot_cr_k + labs(tag = "(f)")) +
  plot_annotation(tag_levels = list(c("(d)", "(e)", "(f)"))) &
  theme(
    plot.tag = element_text(face = "bold", size = 28), 
    axis.title.x = element_text(size = 28),           
    axis.title.y = element_text(size = 28),           
    axis.text.x = element_text(size = 24),            
    axis.text.y = element_text(size = 24)             
  )

# Save the bottom row as a separate figure
ggsave("wd_para_plot_def.png", plot = wd_plots_def, width = 26, height = 9, dpi = 300, bg = "white")
```

# CR-DBH (Power-law)

# Load necessary libraries

```{r}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)
```

# Wood density data preparation

```{r}
options(width = 150)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

wd_df <- data2 |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

wd_df
dim(wd_df)
wd_df <- wd_df |>
  mutate(sp_id = 1:n())

head(wd_df)
dim(wd_df)
```

## Extract a, b

```{r}
# Load the necessary data
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
str(stan_data_lr_cr)

wd_s <- stan_data_lr_cr$u[, 2]

pl_cr_summary <- tar_read(fit_lr2_summary_power_law1_cr)
pl_cr_summary

# Extract the beta parameters (species-specific) with 95% CI
beta <- pl_cr_summary |> filter(grepl("beta\\[", variable))
beta

# Calculate mean and 90% CI for a, b, k
a_median <- beta |> filter(str_detect(variable, "beta\\[1,")) |> pull(q50)
a_lower <- beta |> filter(str_detect(variable, "beta\\[1,")) |> pull(q2.5)
a_upper <- beta |> filter(str_detect(variable, "beta\\[1,")) |> pull(q97.5)

b_median <- beta |> filter(str_detect(variable, "beta\\[2,")) |> pull(q50)
b_lower <- beta |> filter(str_detect(variable, "beta\\[2,")) |> pull(q2.5)
b_upper <- beta |> filter(str_detect(variable, "beta\\[2,")) |> pull(q97.5)

# Extract community-level parameters and calculate fitted values
gamma <- pl_cr_summary |> filter(grepl("gamma", variable))

gamma_log_a_intercept <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
gamma_a_slope_wd <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)

gamma_b_intercept <- gamma |> filter(variable == "gamma[2,1]") |> pull(q50)
gamma_b_wd <- gamma |> filter(variable == "gamma[2,2]") |> pull(q50)

# Compute community fitted values
community_fitted_log_a <- gamma_log_a_intercept + gamma_a_slope_wd * wd_s
community_fitted_a <- exp(community_fitted_log_a)

community_fitted_b <- gamma_b_intercept + gamma_b_wd * wd_s

# Calculate 95% credible intervals for community fitted values
gamma_log_a_intercept_lower <- gamma |> filter(variable == "gamma[1,1]") |> pull(q2.5)
gamma_log_a_wd_lower <- gamma |> filter(variable == "gamma[1,2]") |> pull(q2.5)

gamma_log_a_intercept_upper <- gamma |> filter(variable == "gamma[1,1]") |> pull(q97.5)
gamma_log_a_wd_upper <- gamma |> filter(variable == "gamma[1,2]") |> pull(q97.5)

gamma_b_intercept_lower <- gamma |> filter(variable == "gamma[2,1]") |> pull(q2.5)
gamma_b_wd_lower <- gamma |> filter(variable == "gamma[2,2]") |> pull(q97.5)

gamma_b_intercept_upper <- gamma |> filter(variable == "gamma[1,2]") |> pull(q97.5)
gamma_b_wd_upper <- gamma |> filter(variable == "gamma[2,2]") |> pull(q97.5)

# Compute lower and upper bounds for community fitted values
# log_community_fitted_a_lower <- gamma_log_a_intercept_lower + gamma_log_a_wd_lower * wd_s
# log_community_fitted_a_upper <- gamma_log_a_intercept_upper + gamma_log_a_wd_upper * wd_s

# community_fitted_a_lower <- exp(log_community_fitted_a_lower)
# community_fitted_a_upper <- exp(log_community_fitted_a_upper)

community_fitted_a_lower <- gamma_log_a_intercept_lower + gamma_log_a_wd_lower * wd_s
community_fitted_a_upper <- gamma_log_a_intercept_upper + gamma_log_a_wd_upper * wd_s

# community_fitted_a_lower <- exp(log_community_fitted_a_lower)
# community_fitted_a_upper <- exp(log_community_fitted_a_upper)

community_fitted_b_lower <- gamma_b_intercept_lower + gamma_b_wd_lower * wd_s
community_fitted_b_upper <- gamma_b_intercept_upper + gamma_b_wd_upper * wd_s

# Prepare data for plotting
sp_id <- as.factor(unique(stan_data_lr_cr$jj))
cr_wd_data <- data.frame(
  sp_id = sp_id,
  wd_s = wd_s,
  a_median = a_median,
  a_lower = a_lower,
  a_upper = a_upper,
  b_median = b_median,
  b_lower = b_lower,
  b_upper = b_upper,
  community_fitted_a = community_fitted_a,
  community_fitted_b = community_fitted_b,
  community_fitted_a_lower = community_fitted_a_lower,
  community_fitted_a_upper = community_fitted_a_upper,
  community_fitted_b_lower = community_fitted_b_lower,
  community_fitted_b_upper = community_fitted_b_upper
)
head(cr_wd_data)

# Join with wood density data
cr_wd_data$sp_id <- as.integer(as.character(cr_wd_data$sp_id))
cr_wd_data <- inner_join(cr_wd_data, wd_df, by = "sp_id")
head(cr_wd_data)

```

## Plotting

```{r}
plot_cr_a <- ggplot(cr_wd_data, aes(x = wd, y = a_median)) +
  geom_errorbar(aes(ymin = a_lower, ymax = a_upper), color = "lightgrey", width = 0.0001) +
  geom_point(color = "darkgrey", size = 1) +
  geom_ribbon(aes(ymin = community_fitted_a_lower, ymax = community_fitted_a_upper), fill = "grey80", alpha = 0.5) +
  geom_line(aes(y = community_fitted_a), color = "black", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Amplitude parameter, a") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

plot_cr_b <- ggplot(cr_wd_data, aes(x = wd, y = b_median)) +
  geom_errorbar(aes(ymin = b_lower, ymax = b_upper), color = "orange", width = 0.0001) +
  geom_point(color = "orange", size = 1) +
  # geom_ribbon(aes(ymin = community_fitted_b_lower, ymax = community_fitted_b_upper), fill = "orange", alpha = 0.5) +
  geom_point(color = "orange", size = 2) +
  geom_line(aes(y = community_fitted_b), color = "darkorange", linewidth = 1.2) +
  labs(x = "Wood Density (g cm⁻³)", y = "Exponent, b") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

# Print plots
plot_cr_a
plot_cr_b
```

# Updating the code

```{r}
library(targets)
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)
library(Rmpfr)
```

# Wood density data preparation
```{r}
options(width = 150)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

wd_df <- data2 |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

wd_df
dim(wd_df)
wd_df <- wd_df |>
  mutate(sp_id = 1:n())

head(wd_df)
dim(wd_df)
```

```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)
wd_s <- stan_data_nlr_h$u[, 2]
wb1_h_summary <- tar_read(fit_nlr_summary_weibull1_h)
wb1_h_summary

wb1_h <- tar_read(fit_nlr_draws_weibull1_h)
wb1_h

# Extract the necessary variables from the wb1_h_summary
gamma_1_1 <- wb1_h_summary |> filter(variable == "gamma[1,1]") |> pull(mean)
gamma_2_1 <- wb1_h_summary |> filter(variable == "gamma[2,1]") |> pull(mean)
gamma_1_2 <- wb1_h_summary |> filter(variable == "gamma[1,2]") |> pull(mean)
gamma_2_2 <- wb1_h_summary |> filter(variable == "gamma[2,2]") |> pull(mean)
gamma_1_3 <- wb1_h_summary |> filter(variable == "gamma[1,3]") |> pull(mean)
gamma_2_3 <- wb1_h_summary |> filter(variable == "gamma[2,3]") |> pull(mean)

# Compute the predicted values using these gammas
pred_a <- exp(gamma_1_1 + gamma_2_1 * wd_s)
pred_b <- gamma_1_2 + gamma_2_2 * wd_s
pred_k <- gamma_1_3 + gamma_2_3 * wd_s

# Combine into a matrix for easier manipulation
pred_matrix <- cbind(pred_a, pred_b, pred_k)

head(pred_matrix)

for (n in 1:length(predicted_log_y)) {
  predicted_log_y[n] <- pred_a[stan_data_nlr_h$jj[n]] + 
                        log1mexp(-pred_b[stan_data_nlr_h$jj[n]] * 
                        (stan_data_nlr_h$x[n] ^ pred_k[stan_data_nlr_h$jj[n]]))
}

# Prepare data for plotting
h_wd_data <- data.frame(
  sp_id = stan_data_nlr_h$jj,  # species ID for each observation
  wd_s = wd_s[stan_data_nlr_h$jj],  # wood density corresponding to each observation's species
  predicted_log_y = predicted_log_y,
  observed_log_y = stan_data_nlr_h$log_y
)

# Check the data
head(h_wd_data)
```

# REVISING

```{r}
library(targets)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
library(patchwork)
library(cowplot)
```

# Wood density data preparation

```{r}
options(width = 150)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
  filter(!is.na(dbh)) |>
  filter(!is.na(cr)) |>
  filter(!is.na(h)) |>
  filter(!is.na(wd_s))

wd_df <- data2 |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())

wd_df <- wd_df |>
  mutate(sp_id = 1:n())

head(wd_df)
```

```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)
wd_s <- stan_data_nlr_h$u[, 2]
length(wd_s)

wb1_h <- tar_read(fit_nlr_draws_weibull1_h)
wb1_h

gamma_params <- wb1_h |>
  select(starts_with("gamma"))
gamma_params

beta_params <- wb1_h |>
  select(starts_with("beta["))
beta_params
```

```{r}
# Step 1: Reshape the beta_params tibble into long format to separate the parameters
beta_long <- beta_params |>
  pivot_longer(
    cols = everything(),
    names_to = c("species", "parameter"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(species = as.integer, parameter = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",
    parameter == 2 ~ "b",
    parameter == 3 ~ "k"
  ),
  value = log(value))  # Log-transform the beta values to match the scale of gamma

# Step 2: Join with wood density data
# We join the reshaped beta parameters with the wood density (wd_s) information
plot_data <- beta_long |>
  left_join(wd_df, by = c("species" = "sp_id"))
plot_data

# Step 3: Summarize the posterior distributions for plotting
summary_stats <- plot_data |>
  group_by(species, parameter, wd_s) |>
  summarize(
    median = median(value),
    lower = quantile(value, 0.025),  # 2.5th percentile for lower bound of 95% CI
    upper = quantile(value, 0.975),  # 97.5th percentile for upper bound of 95% CI
    .groups = "drop"
  )

# Step 4: Generate the "pile of lines" plot using gamma parameters
gamma_log_a <- gamma_params |>
  select(`gamma[1,1]`, `gamma[2,1]`)

wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)

lines_data <- do.call(rbind, lapply(1:nrow(gamma_log_a), function(i) {
  data.frame(
    wd_s = wd_seq,
    log_log_a = gamma_log_a$`gamma[1,1]`[i] + gamma_log_a$`gamma[2,1]`[i] * wd_seq,
    draw = i
  )
}))

median_intercept <- median(gamma_log_a$`gamma[1,1]`)
median_slope <- median(gamma_log_a$`gamma[2,1]`)

fit_line <- data.frame(
  wd_s = wd_seq,
  log_log_a = median_intercept + median_slope * wd_seq
)

# Step 5: Plot the "pile of lines", median fit line, and species-specific points with CIs
ggplot() +
  geom_line(data = lines_data, aes(x = wd_s, y = log_log_a, group = draw), alpha = 0.1, color = "blue") +  # Pile of lines
  geom_line(data = fit_line, aes(x = wd_s, y = log_log_a), color = "red", size = 1) +  # Median fit line
  geom_point(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd_s, y = median), color = "green") +  # Species-specific points
  geom_errorbar(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd_s, ymin = lower, ymax = upper), width = 0.02, color = "green") +  # 95% CIs
  theme_minimal() +
  labs(
    x = "Scaled Wood Density (WD)",
    y = "Log(log(a))",
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )
```

### 2025-03-26 docs/wd_para_cr_new.qmd
Initial commit for tree-allo

---
title: "Wood density vs. CR-DBH allometry parameters (Weibull)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(patchwork)
library(posterior)
options(width = 150)
```

# Summary result

## Ang

```{r}

pl_cr_wd_ang <- tar_read(fit_wd_ang_cr_summary_pl_wd)
pl_cr_wd_ang

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

```

- Key finding: Wood density is signifcant effect the horizontal expansion of *Angiosperm* tree.

## Gym

```{r}
gmm_cr_wd_gym <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
gmm_cr_wd_gym

gmm_cr_wd_gym <- gmm_cr_wd_gym |>
  filter(str_detect(variable, "gamma"))

gmm_cr_wd_gym

gmm_cr_wd_gym <- gmm_cr_wd_gym |>
  filter(str_detect(variable, "gamma_int") | str_detect(variable, "gamma_slope"))
gmm_cr_wd_gym

```

- Key finding: Wood density significantly affects parameter b, but not a and k of CR-DBH allometry of *Gymnosperm*.

# Plotting

## Data preparation

### Ang

```{r}
tallo_wd_df_cr <- tar_read(tallo_wd_df_cr)
tallo_wd_df_cr

ang_cr_df <- tallo_wd_df_cr |> filter(division == "Angiosperm")
ang_cr_df

wd_df_cr_ang <- ang_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_cr_ang

```

- 762 species

```{r}
pl_cr_wd_ang_draws <- tar_read(fit_wd_ang_cr_draws_pl_wd)
pl_cr_wd_ang_draws

gamma_ang_cr <- pl_cr_wd_ang_draws |>
  select(starts_with("gamma"))

gamma_ang_cr

beta_ang_cr <- pl_cr_wd_ang_draws |>
  select(starts_with("beta["))

beta_ang_cr

beta_ang_cr_long <- beta_ang_cr |>
  pivot_longer(
    cols = everything(),
    names_to = c("parameter", "species"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(parameter = as.integer, species = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",  # Assign 'log_a' for [1,]
    parameter == 2 ~ "b"       # Assign 'b' for [2,]
  ),
  value = log(value))

beta_ang_cr_long

plot_ang_cr_data <- beta_ang_cr_long |>
  left_join(wd_df_cr_ang, by = c("species" = "sp_id"))

plot_ang_cr_data

```

### Gym

#### Species ID

```{r}

gym_cr_df <- tallo_wd_df_cr |> filter(division == "Gymnosperm")
gym_cr_df

wd_df_cr_gym <- gym_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_cr_gym

wd_seq_cr_gym <- seq(min(wd_df_cr_gym$wd_s), max(wd_df_cr_gym$wd_s), length.out = 100)
```

#### Extracting parameters

```{r}
gmm_cr_wd_gym_draws <- tar_read(fit_wd_gym_cr_draws_gmm_wd)
gmm_cr_wd_gym_draws

fit_wd_gym_cr_data <- tar_read(fit_wd_gym_cr_data)

gamma_gym_cr <- gmm_cr_wd_gym_draws |>
  select(starts_with("gamma"))

gamma_gym_cr

beta_gym_cr <- gmm_cr_wd_gym_draws |>
  select(starts_with("beta["))

beta_gym_cr

beta_gym_cr_long <- beta_gym_cr |>
  pivot_longer(
    cols = everything(),
    names_to = c("species", "parameter"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(species = as.integer, parameter = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",
    parameter == 2 ~ "b",
    parameter == 3 ~ "k"
  ),
  value = log(value))

beta_gym_cr_long

plot_gym_cr_data <- beta_gym_cr_long |>
  left_join(wd_df_cr_gym, by = c("species" = "sp_id"))

plot_gym_cr_data
```

# Plotting function

```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

fit_wd_ang_cr_data <- tar_read(fit_wd_ang_cr_data)
str(fit_wd_ang_cr_data)

stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
str(stan_data_lr_ang_cr)

# same as fit_wd_ang_cr_data

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
str(tallo_reduced_lr_df_ang_cr)

ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

ang_cr_df

wd_df_ang <- ang_cr_df |>
    group_by(sp) |>
    summarize(wd = mean(wd, na.rm = TRUE)) |>
    mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_seq_ang <- seq(min(wd_df_ang$wd_s), max(wd_df_ang$wd_s), length.out = 100)

wd_seq_df_ang <- data.frame(
    wd_s = wd_seq_ang,
    wd = approx(wd_df_ang$wd_s, wd_df_ang$wd, xout = wd_seq_ang)$y
  )

```

```{r}
fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

gamma_ang_cr <- fit_wd_ang_cr_draws_pl_wd |>
  select(starts_with("gamma"))

gamma_ang_cr

beta_ang_cr <- fit_wd_ang_cr_draws_pl_wd |>
  select(starts_with("beta["))

beta_ang_cr

beta_ang_cr_long <- beta_ang_cr |>
  pivot_longer(
    cols = everything(),  # Pivot all columns
    names_to = c("parameter", "species"),  # Split column names into 'parameter' and 'species'
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",  # Extract the two parts from column names like beta[1,1]
    names_transform = list(parameter = as.integer, species = as.integer),  # Convert to integers
    values_to = "value"  # Store values in 'value'
  ) |>
  mutate(
    parameter = case_when(
      parameter == 1 ~ "log_a",  # Map '1' to 'log_a'
      parameter == 2 ~ "b"       # Map '2' to 'b'
    )
  )

beta_ang_cr_long

plot_ang_cr_data <- beta_ang_cr_long |>
  left_join(wd_df_ang, by = c("species" = "sp_id"))

plot_ang_cr_data

summary_stats <- plot_ang_cr_data |>
  group_by(species, parameter, wd, wd_s) |>
  summarize(
    median = median(value),
    lower = quantile(value, 0.025),
    upper = quantile(value, 0.975),
    .groups = "drop"
  )

# Join summary_stats with wd_df_ang to get the `wd` values
# summary_stats <- summary_stats |>
#   left_join(wd_df_ang, by = c("wd_s"))

summary_stats

gamma_log_a <- gamma_ang_cr |>
  select(`gamma[1,1]`, `gamma[1,2]`)
gamma_b <- gamma_ang_cr |>
  select(`gamma[1,2]`, `gamma[2,2]`)

lines_data <- do.call(rbind, lapply(1:nrow(gamma_log_a), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    log_a = gamma_log_a$`gamma[1,1]`[i] + gamma_log_a$`gamma[1,2]`[i] * wd_seq_df_ang$wd_s,
    draw = i
  )
}))

median_intercept <- median(gamma_log_a$`gamma[1,1]`)
median_slope <- median(gamma_log_a$`gamma[1,2]`)

# Update `lines_data` to compute `a` from `log_a`
lines_data <- do.call(rbind, lapply(1:nrow(gamma_log_a), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    a = exp(gamma_log_a$`gamma[1,1]`[i] + gamma_log_a$`gamma[1,2]`[i] * wd_seq_df_ang$wd_s),  # Now using `a` instead of `log_a`
    draw = i
  )
}))

fit_line_a <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    a = exp(median_intercept + median_slope * wd_seq_df_ang$wd_s)
)

# Updated plot code
ggplot() +
  geom_line(data = lines_data, aes(x = wd, y = a, group = draw), alpha = 0.1, color = "blue") +  # Pile of lines using `a` for y
  geom_line(data = fit_line_a, aes(x = wd, y = a), color = "red", size = 1) +  # Median fit line using `a`
  geom_point(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd, y = exp(median)), color = "green") +  # Species-specific points using `a`
  geom_errorbar(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd, ymin = exp(lower), ymax = exp(upper)), width = 0.02, color = "green") +  # 95% CIs using `a`
  theme_minimal() +
  labs(
    x = "WD",  # Label for x-axis
    y = "a"  # Label for y-axis (now using `a`)
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

```

### for b

```{r}
# Select gamma terms for `b` from gamma_ang_cr
gamma_b <- gamma_ang_cr |>
  select(`gamma[2,1]`, `gamma[2,2]`)

# Prepare data for lines corresponding to `b`
lines_data_b <- do.call(rbind, lapply(1:nrow(gamma_b), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    b = gamma_b$`gamma[2,1]`[i] + gamma_b$`gamma[2,2]`[i] * wd_seq_df_ang$wd_s,
    draw = i
  )
}))

# Median fit line for `b`
median_intercept_b <- median(gamma_b$`gamma[2,1]`)
median_slope_b <- median(gamma_b$`gamma[2,2]`)

fit_line_b <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    b = median_intercept_b + median_slope_b * wd_seq_df_ang$wd_s
)

# Updated plot for `b`
ggplot() +
  geom_line(data = lines_data_b, aes(x = wd, y = b, group = draw), alpha = 0.1, color = "blue") +  # Pile of lines using `b` for y
  geom_line(data = fit_line_b, aes(x = wd, y = b), color = "red", size = 1) +  # Median fit line using `b`
  geom_point(data = summary_stats |> filter(parameter == "b"), aes(x = wd, y = median), color = "green") +  # Species-specific points using `b`
  geom_errorbar(data = summary_stats |> filter(parameter == "b"), aes(x = wd, ymin = lower, ymax = upper), width = 0.02, color = "green") +  # 95% CIs using `b`
  theme_minimal() +
  labs(
    x = "WD",  # Label for x-axis
    y = "b"  # Label for y-axis (now using `b`)
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )
```

## The scale of points is weird? WHYYY?

```{r}
fit_wd_ang_cr_data <- tar_read(fit_wd_ang_cr_data)

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
str(tallo_reduced_lr_df_ang_cr)

# Filter Angiosperms
ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

# Group by species and calculate mean wood density for Angiosperms
wd_df_ang <- ang_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd, na.rm = TRUE)) |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

# Create the sequence for wood density for Angiosperms
wd_seq_ang <- seq(min(wd_df_ang$wd_s), max(wd_df_ang$wd_s), length.out = 100)

wd_seq_ang

wd_seq_df_ang <- data.frame(
  wd_s = wd_seq_ang,
  wd = approx(wd_df_ang$wd_s, wd_df_ang$wd, xout = wd_seq_ang)$y
)

# Prepare function for plotting
prepare_and_plot_parameter <- function(fit_data, fit_draws, wd_df, param_name, color_point, color_ribbon, line_color, show_ribbon = TRUE) {
  # Extract gamma parameters for the selected parameter
  gamma_params <- fit_draws |>
    dplyr::select(starts_with("gamma"))

  # Extract beta (species-specific) parameters
  beta_params <- fit_draws |>
    dplyr::select(starts_with("beta[")) |>
    pivot_longer(
      cols = everything(),
      names_to = c("parameter", "species"),
      names_pattern = "beta\\[(\\d+),(\\d+)\\]",
      names_transform = list(parameter = as.integer, species = as.integer),
      values_to = "value"
    ) |>
    mutate(parameter = case_when(
      parameter == 1 ~ "log_a",
      parameter == 2 ~ "b"
    ))

  # Combine species-specific beta with wd_df
  plot_data <- beta_params |>
    left_join(wd_df, by = c("species" = "sp_id"))

  # Summarize the posterior distributions for plotting
  summary_stats <- plot_data |>
    group_by(species, parameter, wd, wd_s) |>
    summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  # Prepare data for lines using gamma parameters
  gamma_param <- gamma_params |>
    dplyr::select(paste0("gamma[1,", which(c("log_a", "b") == param_name), "]"), 
            paste0("gamma[2,", which(c("log_a", "b") == param_name), "]"))

  lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
    data.frame(
      wd_s = wd_seq_df_ang$wd_s,
      wd = wd_seq_df_ang$wd,
      value = if (param_name == "log_a") {
        exp(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df_ang$wd_s)
      } else {
        gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df_ang$wd_s
      },
      draw = i
    )
  }))

  # Calculate the fit line and 95% CIs
  median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
  median_slope <- median(gamma_param[[2]], na.rm = TRUE)

  fit_line <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    value = if (param_name == "log_a") {
      exp(median_intercept + median_slope * wd_seq_df_ang$wd_s)
    } else {
      median_intercept + median_slope * wd_seq_df_ang$wd_s
    }
  )

  ci_ribbon <- lines_data |>
    group_by(wd) |>
    summarize(
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

  if (param_name == "log_a") {
    summary_stats <- summary_stats |>
      mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
  }

  # Create plot with the selected colors
  plot <- ggplot() +
    geom_errorbar(data = summary_stats |> filter(parameter == param_name),
                  aes(x = wd, ymin = lower, ymax = upper), 
                  width = 0, linewidth = 0.2, color = color_ribbon) +
    geom_point(data = summary_stats |> filter(parameter == param_name), 
                aes(x = wd, y = median), color = color_point, size = 0.5) +
    geom_line(data = fit_line, aes(x = wd, y = value), 
              color = line_color, linewidth = 0.6) +
    theme_minimal() +
    labs(x = "Wood Density (g cm⁻³)", y = ifelse(param_name == "log_a", "Amplitude (a)", "Scale (b)")) +
    theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9))

  if (show_ribbon) {
    plot <- plot + geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), alpha = 0.5, fill = color_ribbon)
  }

  return(plot)
}

```

```{r}
# Generate plots using the custom colors for log_a and b
p1 <- prepare_and_plot_parameter(fit_wd_ang_cr_data, fit_wd_ang_cr_draws_pl_wd, wd_df_ang, "log_a", "#9e9ac8", "#cbc9e2", "#6a51a3", show_ribbon = TRUE)
p2 <- prepare_and_plot_parameter(fit_wd_ang_cr_data, fit_wd_ang_cr_draws_pl_wd, wd_df_ang, "b", "#74c476", "#bae4b3", "#238b45", show_ribbon = TRUE)

# Combine the plots into one layout
p <- (p1 + labs(tag = "(a)", x = NULL) + theme(axis.text.x = element_blank()) |
      p2 + labs(tag = "(b)", x = expression("Wood Density (g cm"^"-3"~")"))) +
  plot_layout(ncol = 1, nrow = 2) +
  plot_annotation(tag_levels = list(c("(a)", "(b)"))) &
  theme(
    plot.tag = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    plot.margin = margin(5, 5, 5, 5)
  )

# Add a label for Angiosperm
p <- ggdraw() +
  draw_plot(p, x = 0, y = 0, width = 1, height = 0.9) +  
  draw_label("Angiosperm", x = 0.75, y = 0.91, fontface = 'bold', size = 10)

```

### 2025-03-26 docs/wd_para_h_new.qmd
Initial commit for tree-allo

---
title: "Wood density vs. H-DBH allometry parameters (Weibull)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(patchwork)
library(posterior)
options(width = 150)
```

# Read in data

- What is the difference between fit_wd_data_ang_h (768 sp) and fit_wd_ang_h_data (1214 sp), same for Gym case (62 vs 76 sp)?

- What is S parameter in h_data?

- I do not see the S parameter in weibull_wd.stan model.

- Best predictive model for Ang and Gym: Weibull

# Summary result

## Ang

```{r}
wb_h_wd_ang <- tar_read(fit_wd_ang_h_summary_weibull_wd)
wb_h_wd_ang
```

- Key finding: gamma_slope[1] and gamma_slope[2]: The 95% CIs overlap 0, so the effect of wood density to a and b are insignificant. gamma_slope[3]: The 95% CI does not overlap 0 (from -0.0254 to -0.00591), indicating that this effect to k is significant for *Angiosperm*.

## Gym

```{r}
wb_h_wd_gym <- tar_read(fit_wd_gym_h_summary_weibull_wd)
wb_h_wd_gym
```

- Key finding: The 95% CIs of gamma_slope parameters all do overlap 0. Hence, the effect of wood density on the H-DBH allometry of *Gymnosperms* is statistically insignificant.

# Plotting

## Data preparation

### Ang

```{r}
tallo_wd_df_h <- tar_read(tallo_wd_df_h)
tallo_wd_df_h

ang_h_df <- tallo_wd_df_h |> filter(division == "Angiosperm")
ang_h_df

wd_df_h_ang <- ang_h_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_h_ang

```

- 1214 species

```{r}
wb_h_wd_ang_draws <- tar_read(fit_wd_ang_h_draws_weibull_wd)

gamma_ang_h <- wb_h_wd_ang_draws |>
  select(starts_with("gamma"))

beta_ang_h <- wb_h_wd_ang_draws |>
  select(starts_with("beta["))

beta_ang_h_long <- beta_ang_h |>
  pivot_longer(
    cols = everything(),
    names_to = c("species", "parameter"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(species = as.integer, parameter = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",
    parameter == 2 ~ "b",
    parameter == 3 ~ "k"
  ),
  value = log(value))

plot_ang_h_data <- beta_ang_h_long |>
  left_join(wd_df_h_ang, by = c("species" = "sp_id"))

plot_ang_h_data
```

### Gym

#### Species ID

```{r}
gym_h_df <- tallo_wd_df_h |> filter(division == "Gymnosperm")
gym_h_df

unique_sp_h_gym <- gym_h_df |>
  distinct(sp)

unique_sp_h_gym

wd_df_h_gym <- gym_h_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_h_gym

wd_seq_h_gym <- seq(min(wd_df_h_gym$wd_s), max(wd_df_h_gym$wd_s), length.out = 100)

wd_df_h_gym_seq_df <- data.frame(
  wd_s = wd_seq_h_gym,
  wd = approx(wd_df_h_gym$wd_s, wd_df_h_gym$wd, xout = wd_seq_h_gym)$y  
)
```

- 76 species

#### Extracting parameters

- WD vs a/b/k: insignificant

```{r}
wb_h_wd_gym_draws <- tar_read(fit_wd_gym_h_draws_weibull_wd)
wb_h_wd_gym_draws

fit_wd_gym_h_data <- tar_read(fit_wd_gym_h_data)

wd_s <- fit_wd_gym_h_data$u[, 2]
length(wd_s)

gamma_gym_h <- wb_h_wd_gym_draws |>
  select(starts_with("gamma"))

gamma_gym_h

beta_gym_h <- wb_h_wd_gym_draws |>
  select(starts_with("beta["))

beta_gym_h

beta_gym_h_long <- beta_gym_h |>
  pivot_longer(
    cols = everything(),
    names_to = c("species", "parameter"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(species = as.integer, parameter = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",
    parameter == 2 ~ "b",
    parameter == 3 ~ "k"
  ),
  value = log(value))

beta_gym_h_long

plot_gym_h_data <- beta_gym_h_long |>
  left_join(wd_df_h_gym, by = c("species" = "sp_id"))

plot_gym_h_data
```

# Plotting function

```{r}
prepare_and_plot_parameter <- function(wd_df, fit_draws_df, param_name, color_point, color_ribbon, line_color, line_type = "solid", show_ribbon = TRUE) {

  param_index <- which(c("log_a", "b", "k") == param_name)

  gamma_param <- fit_draws_df |>
    select(paste0("gamma_int[", param_index, "]"), paste0("gamma_slope[", param_index, "]"))

  transform_value <- if (param_name == "log_a") {
    function(x) exp(x)  # Apply exp() only for log_a
  } else {
    function(x) x  # No transformation for b and k
  }

  wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)
  wd_seq_df <- data.frame(
    wd_s = wd_seq,
    wd = approx(wd_df$wd_s, wd_df$wd, xout = wd_seq)$y
  )

  lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
    data.frame(
      wd_s = wd_seq_df$wd_s,
      wd = wd_seq_df$wd,
      value = transform_value(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df$wd_s),
      draw = i
    )
  }))

  median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
  median_slope <- median(gamma_param[[2]], na.rm = TRUE)

  fit_line <- data.frame(
    wd_s = wd_seq_df$wd_s,
    wd = wd_seq_df$wd,
    value = transform_value(median_intercept + median_slope * wd_seq_df$wd_s)
  )

  ci_ribbon <- lines_data |>
    group_by(wd) |>
    summarize(
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

  summary_stats <- fit_draws_df |>
    select(starts_with("beta[")) |>
    pivot_longer(
      cols = everything(),
      names_to = c("species", "parameter"),
      names_pattern = "beta\\[(\\d+),(\\d+)\\]",
      names_transform = list(species = as.integer, parameter = as.integer),
      values_to = "value"
    ) |>
    mutate(parameter = case_when(
      parameter == 1 ~ "log_a",
      parameter == 2 ~ "b",
      parameter == 3 ~ "k"
    )) |>
    filter(parameter == param_name) |>
    left_join(wd_df, by = c("species" = "sp_id")) |>
    group_by(species, parameter, wd, wd_s) |>
    summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  if (param_name == "log_a") {
    summary_stats <- summary_stats |>
      mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
  }

  y_label <- case_when(
    param_name == "log_a" ~ "Amplitude parameter, a",
    param_name == "b" ~ "Scale parameter, b",
    param_name == "k" ~ "Shape parameter, k"
  )

  plot <- ggplot() +
    geom_errorbar(data = summary_stats, 
                  aes(x = wd, ymin = lower, ymax = upper), 
                  width = 0, linewidth = 0.2, color = color_ribbon) +
    geom_point(data = summary_stats, 
               aes(x = wd, y = median), color = color_point, size = 0.5) +
    geom_line(data = fit_line, aes(x = wd, y = value), 
              color = line_color, linewidth = 1, linetype = line_type) +
    theme_minimal() +
    labs(
      x = "Wood Density (g cm⁻³)",
      y = y_label
    ) +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12)
    )

  if (show_ribbon) {
    plot <- plot +
      geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), 
                  alpha = 0.5, fill = color_ribbon)
  }

  return(plot)
}

```

```{r}
p1 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_ang, 
  fit_draws_df = wb_h_wd_ang_draws, 
  param_name = "log_a", 
  color_point = "#9e9ac8", 
  color_ribbon = "#cbc9e2", 
  line_color = "#6a51a3",
  line_type = "dashed", 
  show_ribbon = FALSE
)

p2 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_ang, 
  fit_draws_df = wb_h_wd_ang_draws, 
  param_name = "b", 
  color_point = "#74c476", 
  color_ribbon = "#bae4b3", 
  line_color = "#238b45",
  line_type = "dashed", 
  show_ribbon = FALSE
)

p3 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_ang, 
  fit_draws_df = wb_h_wd_ang_draws, 
  param_name = "k",
  color_point = "#6baed6", 
  color_ribbon = "#8fcbee", 
  line_color = "#2171b5", 
  show_ribbon = TRUE
)

p4 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_gym, 
  fit_draws_df = wb_h_wd_gym_draws, 
  param_name = "log_a", 
  color_point = "#9e9ac8", 
  color_ribbon = "#cbc9e2", 
  line_color = "#6a51a3",
  line_type = "dashed", 
  show_ribbon = FALSE
)

p5 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_gym, 
  fit_draws_df = wb_h_wd_gym_draws, 
  param_name = "b", 
  color_point = "#74c476", 
  color_ribbon = "#bae4b3", 
  line_color = "#238b45",
  line_type = "dashed", 
  show_ribbon = FALSE
)

p6 <- prepare_and_plot_parameter(
  wd_df = wd_df_h_gym, 
  fit_draws_df = wb_h_wd_gym_draws, 
  param_name = "k",
  color_point = "#6baed6", 
  color_ribbon = "#8fcbee", 
  line_color = "#2171b5", 
  line_type = "dashed", 
  show_ribbon = FALSE
)

p1
p2
p3
p4
p5
p6
```

# Making figure

```{r}
prepare_and_plot_parameter <- function(wd_df, fit_draws_df, param_name, color_point, color_ribbon, line_color, line_type = "solid", show_ribbon = TRUE) {
    # Prepare wood density data
    wd_df <- wd_df |>
      filter(!is.na(wd_s)) |>
      group_by(sp) |>
      summarize(wd = mean(wd, na.rm = TRUE)) |>
      mutate(wd_s = scale(wd) |> as.numeric(), sp_id = 1:n())

    # Extract gamma and beta parameters from fit_draws
    gamma_params <- fit_draws |>
      dplyr::select(starts_with("gamma"))

    # Reshape the beta parameters for analysis
    beta_params <- fit_draws |>
      dplyr::select(starts_with("beta[")) |>
      pivot_longer(
        cols = everything(),
        names_to = c("species", "parameter"),
        names_pattern = "beta\\[(\\d+),(\\d+)\\]",
        names_transform = list(species = as.integer, parameter = as.integer),
        values_to = "value"
      ) |>
      mutate(parameter = case_when(
        parameter == 1 ~ "log_a",
        parameter == 2 ~ "b",
        parameter == 3 ~ "k"
      ))

    # Join wood density data with beta parameters
    plot_data <- beta_params |>
      left_join(wd_df, by = c("species" = "sp_id"))

    # Summarize the posterior distributions for plotting
    summary_stats <- plot_data |>
      group_by(species, parameter, wd, wd_s) |>
      summarize(
        median = median(value, na.rm = TRUE),
        lower = quantile(value, 0.025, na.rm = TRUE),
        upper = quantile(value, 0.975, na.rm = TRUE),
        .groups = "drop"
      )

    # Create a sequence for wd_s and merge with corresponding wd values
    wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)
    wd_seq_df <- data.frame(
      wd_s = wd_seq,
      wd = approx(wd_df$wd_s, wd_df$wd, xout = wd_seq)$y
    )

    # Extract relevant gamma parameters for the chosen model
    gamma_param <- gamma_params |>
      dplyr::select(paste0("gamma[1,", which(c("log_a", "b", "k") == param_name), "]"), 
             paste0("gamma[2,", which(c("log_a", "b", "k") == param_name), "]"))

    # Generate lines and fit data for the selected parameter
    transform_value <- if (param_name == "log_a") {
      function(x) exp(exp(x))
    } else {
      function(x) exp(x)
    }

    lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
      data.frame(
        wd_s = wd_seq_df$wd_s,
        wd = wd_seq_df$wd,
        value = transform_value(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df$wd_s),
        draw = i
      )
    }))

    # Compute the fit line and 95% credible intervals
    median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
    median_slope <- median(gamma_param[[2]], na.rm = TRUE)

    fit_line <- data.frame(
      wd_s = wd_seq_df$wd_s,
      wd = wd_seq_df$wd,
      value = transform_value(median_intercept + median_slope * wd_seq_df$wd_s)
    )

    ci_ribbon <- lines_data |>
      group_by(wd) |>
      summarize(
        lower = quantile(value, 0.025, na.rm = TRUE),
        upper = quantile(value, 0.975, na.rm = TRUE),
        .groups = "drop"
      )

    fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

    # Apply transformation for log_a in summary_stats
    if (param_name == "log_a") {
      summary_stats <- summary_stats |>
        mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
    }

    # Set y-axis label based on parameter
    y_label <- case_when(
      param_name == "log_a" ~ "Amplitude parameter, a",
      param_name == "b" ~ "Scale parameter, b",
      param_name == "k" ~ "Shape parameter, k"
    )

    # Create plot
    plot <- ggplot() +
      geom_errorbar(data = summary_stats |> filter(parameter == param_name), 
                    aes(x = wd, ymin = lower, ymax = upper), 
                    width = 0, linewidth = 0.2, color = color_ribbon) +
      geom_point(data = summary_stats |> filter(parameter == param_name), 
                 aes(x = wd, y = median), color = color_point, size = 0.5) +
      geom_line(data = fit_line, aes(x = wd, y = value), 
                color = line_color, linewidth = 0.6, linetype = line_type) +
      theme_minimal() +
      labs(x = "Wood Density (g cm⁻³)", y = y_label) +
      theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9))

    if (show_ribbon) {
      plot <- plot + geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), alpha = 0.5, fill = color_ribbon)
    }

    return(plot)
  }

```

```{r}
p1 <- prepare_and_plot_parameter(
  fit_wd_gym_h_data,
  wb_h_wd_gym_draws,
  wd_df,
  "log_a", "#9e9ac8", "#cbc9e2", "#6a51a3",
  show_ribbon = TRUE)

```

```{r}
tallo_wd_df_h <- tar_read(tallo_wd_df_h)
tallo_wd_df_h

ang_h_df <- tallo_wd_df_h |> filter(division == "Angiosperm")
ang_h_df

gym_h_df <- tallo_wd_df_h |> filter(division == "Gymnosperm")
gym_h_df

fit_wd_ang_h_draws_weibull_wd <- tar_read(fit_wd_ang_h_draws_weibull_wd)

fit_wd_gym_h_draws_weibull_wd <- tar_read(fit_wd_gym_h_draws_weibull_wd)

fit_wd_ang_h_data <- tar_read(fit_wd_ang_h_data)

fit_wd_gym_h_data <- tar_read(fit_wd_gym_h_data)

```

```{r}
generate_wd_para_h <- function(tallo_wd_df_h, fit_wd_ang_h_draws_weibull_wd, fit_wd_gym_h_draws_weibull_wd, fit_wd_ang_h_data, fit_wd_gym_h_data) {

# Filter Angiosperms and Gymnosperms
  ang_h_df <- tallo_wd_df_h |> filter(division == "Angiosperm")
  gym_h_df <- tallo_wd_df_h |> filter(division == "Gymnosperm")

  # Group by species and calculate mean wood density for Angiosperms
  wd_df_ang <- ang_h_df |>
    group_by(sp) |>
    summarize(wd = mean(wd, na.rm = TRUE)) |>
    mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

  # Group by species and calculate mean wood density for Gymnosperms
  wd_df_gym <- gym_h_df |>
    group_by(sp) |>
    summarize(wd = mean(wd, na.rm = TRUE)) |>
    mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

  # Create the sequence for wood density for Angiosperms and Gymnosperms separately
  wd_seq_ang <- seq(min(wd_df_ang$wd_s), max(wd_df_ang$wd_s), length.out = 100)
  wd_seq_gym <- seq(min(wd_df_gym$wd_s), max(wd_df_gym$wd_s), length.out = 100)

  wd_seq_df_ang <- data.frame(
    wd_s = wd_seq_ang,
    wd = approx(wd_df_ang$wd_s, wd_df_ang$wd, xout = wd_seq_ang)$y
  )

  wd_seq_df_gym <- data.frame(
    wd_s = wd_seq_gym,
    wd = approx(wd_df_gym$wd_s, wd_df_gym$wd, xout = wd_seq_gym)$y
  )

  # Function to prepare and plot parameters (log_a, b, k) for H-DBH of Ang and Gym
  prepare_and_plot_parameter <- function(stan_data, fit_draws, wd_df, param_name, color_point, color_ribbon, line_color, line_type = "solid", show_ribbon = TRUE) {
    # Prepare wood density data
    wd_df <- wd_df |>
      filter(!is.na(wd_s)) |>
      group_by(sp) |>
      summarize(wd = mean(wd, na.rm = TRUE)) |>
      mutate(wd_s = scale(wd) |> as.numeric(), sp_id = 1:n())

    # Extract gamma and beta parameters from fit_draws
    gamma_params <- fit_draws |>
      dplyr::select(starts_with("gamma"))

    # Reshape the beta parameters for analysis
    beta_params <- fit_draws |>
      dplyr::select(starts_with("beta[")) |>
      pivot_longer(
        cols = everything(),
        names_to = c("species", "parameter"),
        names_pattern = "beta\\[(\\d+),(\\d+)\\]",
        names_transform = list(species = as.integer, parameter = as.integer),
        values_to = "value"
      ) |>
      mutate(parameter = case_when(
        parameter == 1 ~ "log_a",
        parameter == 2 ~ "b",
        parameter == 3 ~ "k"
      ))

    # Join wood density data with beta parameters
    plot_data <- beta_params |>
      left_join(wd_df, by = c("species" = "sp_id"))

    # Summarize the posterior distributions for plotting
    summary_stats <- plot_data |>
      group_by(species, parameter, wd, wd_s) |>
      summarize(
        median = median(value, na.rm = TRUE),
        lower = quantile(value, 0.025, na.rm = TRUE),
        upper = quantile(value, 0.975, na.rm = TRUE),
        .groups = "drop"
      )

    # Create a sequence for wd_s and merge with corresponding wd values
    wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)
    wd_seq_df <- data.frame(
      wd_s = wd_seq,
      wd = approx(wd_df$wd_s, wd_df$wd, xout = wd_seq)$y
    )

    # Extract relevant gamma parameters for the chosen model
    gamma_param <- gamma_params |>
      dplyr::select(paste0("gamma[1,", which(c("log_a", "b", "k") == param_name), "]"), 
             paste0("gamma[2,", which(c("log_a", "b", "k") == param_name), "]"))

    # Generate lines and fit data for the selected parameter
    transform_value <- if (param_name == "log_a") {
      function(x) exp(exp(x))
    } else {
      function(x) exp(x)
    }

    lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
      data.frame(
        wd_s = wd_seq_df$wd_s,
        wd = wd_seq_df$wd,
        value = transform_value(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df$wd_s),
        draw = i
      )
    }))

    # Compute the fit line and 95% credible intervals
    median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
    median_slope <- median(gamma_param[[2]], na.rm = TRUE)

    fit_line <- data.frame(
      wd_s = wd_seq_df$wd_s,
      wd = wd_seq_df$wd,
      value = transform_value(median_intercept + median_slope * wd_seq_df$wd_s)
    )

    ci_ribbon <- lines_data |>
      group_by(wd) |>
      summarize(
        lower = quantile(value, 0.025, na.rm = TRUE),
        upper = quantile(value, 0.975, na.rm = TRUE),
        .groups = "drop"
      )

    fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

    # Apply transformation for log_a in summary_stats
    if (param_name == "log_a") {
      summary_stats <- summary_stats |>
        mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
    }

    # Set y-axis label based on parameter
    y_label <- case_when(
      param_name == "log_a" ~ "Amplitude parameter, a",
      param_name == "b" ~ "Scale parameter, b",
      param_name == "k" ~ "Shape parameter, k"
    )

    # Create plot
    plot <- ggplot() +
      geom_errorbar(data = summary_stats |> filter(parameter == param_name), 
                    aes(x = wd, ymin = lower, ymax = upper), 
                    width = 0, linewidth = 0.2, color = color_ribbon) +
      geom_point(data = summary_stats |> filter(parameter == param_name), 
                 aes(x = wd, y = median), color = color_point, size = 0.5) +
      geom_line(data = fit_line, aes(x = wd, y = value), 
                color = line_color, linewidth = 0.6, linetype = line_type) +
      theme_minimal() +
      labs(x = "Wood Density (g cm⁻³)", y = y_label) +
      theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9))

    if (show_ribbon) {
      plot <- plot + geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), alpha = 0.5, fill = color_ribbon)
    }

    return(plot)
  }

  # Generate individual plots for both H-DBH of Ang and Gym
  p1 <- prepare_and_plot_parameter(fit_wd_ang_h_data, fit_wd_ang_h_draws_weibull_wd, wd_df, "log_a", "#9e9ac8", "#cbc9e2", "#6a51a3", "dashed", show_ribbon = FALSE)
  p2 <- prepare_and_plot_parameter(fit_wd_ang_h_data, fit_wd_ang_h_draws_weibull_wd, wd_df, "b", "#74c476", "#bae4b3", "#238b45", "dashed", show_ribbon = FALSE)
  p3 <- prepare_and_plot_parameter(fit_wd_ang_h_data, fit_wd_ang_h_draws_weibull_wd, wd_df, "k", "#6baed6", "#8fcbee", "#2171b5", show_ribbon = TRUE)
  p4 <- prepare_and_plot_parameter(fit_wd_gym_h_data, fit_wd_gym_h_draws_weibull_wd, wd_df, "log_a", "#9e9ac8", "#cbc9e2", "#6a51a3", "dashed", show_ribbon = FALSE)
  p5 <- prepare_and_plot_parameter(fit_wd_gym_h_data, fit_wd_gym_h_draws_weibull_wd, wd_df, "b", "#74c476", "#bae4b3", "#238b45", "dashed", show_ribbon = FALSE)
  p6 <- prepare_and_plot_parameter(fit_wd_gym_h_data, fit_wd_gym_h_draws_weibull_wd, wd_df, "k", "#6baed6", "#8fcbee", "#2171b5", "dashed", show_ribbon = FALSE)

  # Combine all plots
  p <- (p1 + labs(tag = "(a)", x = NULL) + theme(axis.text.x = element_blank()) |
                    p4 + labs(tag = "(b)", x = NULL) + theme(axis.title.y = element_blank(), axis.text.x = element_blank()) |
                    p2 + labs(tag = "(c)", x = NULL) + theme(axis.text.x = element_blank()) |
                    p5 + labs(tag = "(d)", x = NULL) + theme(axis.title.y = element_blank(), axis.text.x = element_blank()) |
                    p3 + labs(tag = "(e)", x = expression("Wood Density (g cm"^"-3"~")")) |
                    p6 + labs(tag = "(f)", x = expression("Wood Density (g cm"^"-3"~")")) + theme(axis.title.y = element_blank())) +
    plot_layout(ncol = 2, nrow = 3) +
    plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"))) &
    theme(
      plot.tag = element_text(face = "bold", size = 10), 
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9),
      plot.margin = margin(5, 5, 5, 5))

  # Add labels
  p <- ggdraw() +
    draw_plot(p, x = 0, y = 0, width = 1, height = 0.9) +  
    draw_label("Angiosperm", x = 0.25, y = 0.91, fontface = 'bold', size = 10) +  
    draw_label("Gymnosperm", x = 0.75, y = 0.91, fontface = 'bold', size = 10)

  return(p)
}
```

# Plotting p
```{r}

```

```{r}

```

```{r}

```

### 2025-03-26 docs/wd_para_new.qmd
Initial commit for tree-allo

---
title: "WD's effect"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
options(width = 150)
```

# WD vs. CR-DBH

## Angiosperm

### All species

```{r}
fit_wd_ang_cr_summary_pl_wd <- tar_read(fit_wd_ang_cr_summary_pl_wd)
fit_wd_ang_cr_summary_pl_wd
beta_ang_cr_pl  <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
beta_ang_cr_pl

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
ang_cr_df
unique_species_cr_df <- unique(ang_cr_df$sp)
sp_id_mapping <- ang_cr_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

angiosperm_parameters <- beta_ang_cr_pl |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("1,", variable) ~ "log_a",
    grepl("2,", variable) ~ "b"
  ))

angiosperm_parameters <- angiosperm_parameters |>
  mutate(sp_id = rep(1:(nrow(angiosperm_parameters) / 2), each = 2))

angiosperm_params_wide <- angiosperm_parameters |>
  pivot_wider(
    names_from = param_type,
    values_from = mean,
    id_cols = sp_id
  )

angiosperm_params_wide

angiosperm_params_with_sp <- angiosperm_params_wide |>
  left_join(sp_id_mapping, by = "sp_id")

angiosperm_params_with_sp

angiosperm_data_full <- ang_cr_df |>
  left_join(angiosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, )

angiosperm_data_full

species_summary <- angiosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avr_cr = mean(cr, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )

species_summary
summary(species_summary)

dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

prediction_df <- data.frame()

for(i in 1:nrow(species_summary)) {
  sp <- species_summary$sp[i]
  avg_log_a <- species_summary$avg_log_a[i]
  avg_b <- species_summary$avg_b[i]
  avg_wd <- species_summary$avg_wd[i]

  predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df <- rbind(prediction_df, temp_df)
}

# ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = sp, color = avg_wd)) +
#   geom_line(alpha = 0.7) +
#   scale_color_viridis_c() +
#   labs(x = "DBH (cm)", y = "Crown Radius (m)") +
#   theme_minimal() +
#   theme(legend.title = element_text(size = 12),
#         plot.title = element_text(size = 14)) +
#   guides(color = guide_colorbar(title = "Wood Density"))
```

### Light and Dense

```{r}
# Calculate the 25% and 75% quantiles of WD
wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
high_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))

# Combine low and high WD data
wd_filtered <- bind_rows(low_wd_species, high_wd_species)

# Create DBH sequence
dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

# Generate predictions for low and high WD
prediction_df <- data.frame()

for (i in 1:nrow(wd_filtered)) {
  sp <- wd_filtered$sp[i]
  avg_log_a <- wd_filtered$avg_log_a[i]
  avg_b <- wd_filtered$avg_b[i]
  avg_wd <- wd_filtered$avg_wd[i]

  predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df <- rbind(prediction_df, temp_df)
}

ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)", color = "WD Quantiles") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  )

p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p1

```

## Gymnosperm

### All species

```{r}
fit_wd_gym_cr_summary_gmm_wd <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
fit_wd_gym_cr_summary_gmm_wd
beta_gym_cr_gmm  <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
beta_gym_cr_gmm

tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")
gym_cr_df

unique_species_cr_df_gym <- unique(gym_cr_df$sp)
sp_id_mapping_gym <- gym_cr_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

gymnosperm_parameters <- beta_gym_cr_gmm |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("\\[.*,1\\]", variable) ~ "log_a",
    grepl("\\[.*,2\\]", variable) ~ "b",
    grepl("\\[.*,3\\]", variable) ~ "k"
  ))

gymnosperm_parameters <- gymnosperm_parameters |>
  mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

gymnosperm_params_wide <- gymnosperm_parameters |>
  select(sp_id, param_type, mean) |>
  pivot_wider(names_from = param_type, values_from = mean)

gymnosperm_params_with_sp <- gymnosperm_params_wide |>
  left_join(sp_id_mapping_gym, by = "sp_id")

gymnosperm_data_full <- gym_cr_df |>
  left_join(gymnosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)

species_summary_gym <- gymnosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avr_cr = mean(cr, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_k = mean(k, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )

species_summary_gym

summary(species_summary_gym)

dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)

prediction_df_gym <- data.frame()

for (i in 1:nrow(species_summary_gym)) {
  sp <- species_summary_gym$sp[i]
  avg_log_a <- species_summary_gym$avg_log_a[i]
  avg_b <- species_summary_gym$avg_b[i]
  avg_k <- species_summary_gym$avg_k[i]
  avg_wd <- species_summary_gym$avg_wd[i]

  predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df_gym <- rbind(prediction_df_gym, temp_df)
}

ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = sp, color = avg_wd)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  labs(x = "DBH (cm)", y = "Crown Radius (CR)") +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        plot.title = element_text(size = 14)) +
  guides(color = guide_colorbar(title = "Wood Density"))

```

### Light and Dense

```{r}
# Calculate the 25% and 75% quantiles of WD for Gymnosperms
wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
high_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))

# Combine low and high WD data
wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)

# Create predictions for Gymnosperms based on selected species
prediction_df_gym <- data.frame()

for (i in 1:nrow(wd_filtered_gym)) {
  sp <- wd_filtered_gym$sp[i]
  avg_log_a <- wd_filtered_gym$avg_log_a[i]
  avg_b <- wd_filtered_gym$avg_b[i]
  avg_k <- wd_filtered_gym$avg_k[i]
  avg_wd <- wd_filtered_gym$avg_wd[i]

  predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df_gym <- rbind(prediction_df_gym, temp_df)
}

# ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
#   geom_line(alpha = 0.7) +
#   scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
#   labs(x = "DBH (cm)", y = "Crown Radius (CR)", color = "WD Quantiles") +
#   theme_minimal() +
#   theme(
#     legend.title = element_text(size = 12),
#     plot.title = element_text(size = 14)
#   )

p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p2
```

# WD vs. H-DBH

## Angiosperm

### All species
```{r}
fit_wd_ang_h_summary_weibull_wd <- tar_read(fit_wd_ang_h_summary_weibull_wd)
fit_wd_ang_h_summary_weibull_wd

beta_ang_h_wb  <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
beta_ang_h_wb

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")
summary(ang_h_df)

unique_species_h_df_ang <- unique(ang_h_df$sp)
sp_id_mapping_ang <- ang_h_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

beta_ang_h_weibull <- fit_wd_ang_h_summary_weibull_wd |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("\\[.*,1\\]", variable) ~ "log_a",
    grepl("\\[.*,2\\]", variable) ~ "b",
    grepl("\\[.*,3\\]", variable) ~ "k"
  ))

beta_ang_h_weibull <- beta_ang_h_weibull |>
  mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

angiosperm_params_wide <- beta_ang_h_weibull |>
  select(sp_id, param_type, mean) |>
  pivot_wider(names_from = param_type, values_from = mean)

angiosperm_params_with_sp <- angiosperm_params_wide |>
  left_join(sp_id_mapping_ang, by = "sp_id")

angiosperm_data_full <- ang_h_df |>
  left_join(angiosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)

species_summary_ang <- angiosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avg_h = mean(h, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_k = mean(k, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )
species_summary_ang

summary(species_summary_ang)

dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)

prediction_df_ang <- data.frame()

for (i in 1:nrow(species_summary_ang)) {
  sp <- species_summary_ang$sp[i]
  avg_log_a <- species_summary_ang$avg_log_a[i]
  avg_b <- species_summary_ang$avg_b[i]
  avg_k <- species_summary_ang$avg_k[i]
  avg_wd <- species_summary_ang$avg_wd[i]

  predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
  prediction_df_ang <- rbind(prediction_df_ang, temp_df)
}

ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = sp, color = avg_wd)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  labs(x = "DBH (cm)", y = "Height (H)") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  ) +
  guides(color = guide_colorbar(title = "Wood Density"))
```

### Light and Dense

```{r}
# Using Weibull model, calculate the 25% and 75% quantiles for WD in Angiosperms
wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species_ang <- species_summary_ang |> filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
high_wd_species_ang <- species_summary_ang |> filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))

# Combine low and high WD data
wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)

# Generate height predictions for low and high WD
prediction_df_ang <- data.frame()

for (i in 1:nrow(wd_filtered_ang)) {
  sp <- wd_filtered_ang$sp[i]
  avg_log_a <- wd_filtered_ang$avg_log_a[i]
  avg_b <- wd_filtered_ang$avg_b[i]
  avg_k <- wd_filtered_ang$avg_k[i]
  avg_wd <- wd_filtered_ang$avg_wd[i]

  predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
  prediction_df_ang <- rbind(prediction_df_ang, temp_df)
}

# # Plot DBH vs. Height (H) for Angiosperms with low and high WD
# ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
#   geom_line(alpha = 0.7) +
#   scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
#   labs(x = "DBH (cm)", y = "Height (m)", color = "WD Quantiles") +
#   theme_minimal() +
#   theme(
#     legend.title = element_text(size = 12),
#     plot.title = element_text(size = 14)
#   )

# Plot DBH vs. Height (H) for Angiosperms with low and high WD
p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Height", color = "WD Quantiles") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p3

```

# Combine in 1 plot

```{r}
p1 <- p1 + theme(legend.position = "none")
p2 <- p2 + theme(legend.position = "none")
p3 <- p3 + theme(legend.position = "right")

p <- p1 + p2 + p3 + plot_layout(ncol = 3)
p

```

# Writing function

## p1
```{r}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
```

```{r}
fit_wd_ang_cr_summary_pl_wd <- tar_read(fit_wd_ang_cr_summary_pl_wd)
fit_wd_ang_cr_summary_pl_wd
beta_ang_cr_pl  <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
beta_ang_cr_pl

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

```

```{r}
generate_p1 <- function(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr) {
  # Filter the beta parameters from the fit data
  beta_ang_cr_pl <- fit_wd_ang_cr_summary_pl_wd |> 
    filter(grepl("beta", variable))

  # Filter Angiosperm data from the tallo data
  ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

  # Create species ID mapping
  sp_id_mapping <- ang_cr_df |> 
    select(sp) |> 
    distinct() |> 
    mutate(sp_id = row_number())

  # Process beta parameters into log_a and b
  angiosperm_parameters <- beta_ang_cr_pl |>
    mutate(param_type = case_when(
      grepl("1,", variable) ~ "log_a",
      grepl("2,", variable) ~ "b"
    )) |>
    mutate(sp_id = rep(1:(nrow(beta_ang_cr_pl) / 2), each = 2))

  # Pivot the parameters into wide format
  angiosperm_params_wide <- angiosperm_parameters |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Angiosperm data
  angiosperm_params_with_sp <- angiosperm_params_wide |>
    left_join(sp_id_mapping, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  angiosperm_data_full <- ang_cr_df |>
    left_join(angiosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b)

  # Summarize species data to calculate avg values
  species_summary <- angiosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species <- species_summary |>
    filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
  high_wd_species <- species_summary |>
    filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))

  # Combine low and high WD data
  wd_filtered <- bind_rows(low_wd_species, high_wd_species)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

  # Generate predictions for low and high WD
  prediction_df <- data.frame()

  for (i in 1:nrow(wd_filtered)) {
    sp <- wd_filtered$sp[i]
    avg_log_a <- wd_filtered$avg_log_a[i]
    avg_b <- wd_filtered$avg_b[i]
    avg_wd <- wd_filtered$avg_wd[i]

    predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
    prediction_df <- rbind(prediction_df, temp_df)
  }

  # Plot
  p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p1)
}

# Now call the function using the fit data and tallo data
p1 <- generate_p1(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr)
print(p1)
```

## p2

```{r}
fit_wd_gym_cr_summary_gmm_wd <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
fit_wd_gym_cr_summary_gmm_wd
beta_gym_cr_gmm  <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
beta_gym_cr_gmm

tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

generate_p2 <- function(fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr) {
  # Filter the beta parameters from the fit data
  beta_gym_cr_gmm <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))

  # Filter Gymnosperm data from the tallo data
  gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")

  # Create species ID mapping
  sp_id_mapping_gym <- gym_cr_df |> 
    select(sp) |> 
    distinct() |> 
    mutate(sp_id = row_number())

  # Process beta parameters into log_a, b, and k
  gymnosperm_parameters <- beta_gym_cr_gmm |>
    mutate(param_type = case_when(
      grepl("\\[.*,1\\]", variable) ~ "log_a",
      grepl("\\[.*,2\\]", variable) ~ "b",
      grepl("\\[.*,3\\]", variable) ~ "k"
    )) |>
    mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

  # Pivot the parameters into wide format
  gymnosperm_params_wide <- gymnosperm_parameters |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Gymnosperm data
  gymnosperm_params_with_sp <- gymnosperm_params_wide |>
    left_join(sp_id_mapping_gym, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  gymnosperm_data_full <- gym_cr_df |>
    left_join(gymnosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)

  # Summarize species data to calculate avg values
  species_summary_gym <- gymnosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_k = mean(k, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species_gym <- species_summary_gym |>
    filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
  high_wd_species_gym <- species_summary_gym |>
    filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))

  # Combine low and high WD data
  wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)

  # Generate predictions for low and high WD
  prediction_df_gym <- data.frame()

  for (i in 1:nrow(wd_filtered_gym)) {
    sp <- wd_filtered_gym$sp[i]
    avg_log_a <- wd_filtered_gym$avg_log_a[i]
    avg_b <- wd_filtered_gym$avg_b[i]
    avg_k <- wd_filtered_gym$avg_k[i]
    avg_wd <- wd_filtered_gym$avg_wd[i]

    predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
    prediction_df_gym <- rbind(prediction_df_gym, temp_df)
  }

  # Plot
  p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p2)
}

# Now call the function using the fit data and tallo data
p2 <- generate_p2(fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr)
print(p2)
```

## p3
```{r}
fit_wd_ang_h_summary_weibull_wd <- tar_read(fit_wd_ang_h_summary_weibull_wd)
fit_wd_ang_h_summary_weibull_wd

beta_ang_h_wb  <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
beta_ang_h_wb

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

generate_p3 <- function(fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h) {
  # Filter the beta parameters from the fit data
  beta_ang_h_wb <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))

  # Filter Angiosperm data from the tallo data
  ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")

  # Create species ID mapping
  sp_id_mapping_ang <- ang_h_df |>
    select(sp) |>
    distinct() |>
    mutate(sp_id = row_number())

  # Process beta parameters into log_a, b, and k
  beta_ang_h_weibull <- beta_ang_h_wb |>
    mutate(param_type = case_when(
      grepl("\\[.*,1\\]", variable) ~ "log_a",
      grepl("\\[.*,2\\]", variable) ~ "b",
      grepl("\\[.*,3\\]", variable) ~ "k"
    )) |>
    mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

  # Pivot the parameters into wide format
  angiosperm_params_wide <- beta_ang_h_weibull |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Angiosperm data
  angiosperm_params_with_sp <- angiosperm_params_wide |>
    left_join(sp_id_mapping_ang, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  angiosperm_data_full <- ang_h_df |>
    left_join(angiosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)

  # Summarize species data to calculate avg values
  species_summary_ang <- angiosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_h = mean(h, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_k = mean(k, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species_ang <- species_summary_ang |>
    filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
  high_wd_species_ang <- species_summary_ang |>
    filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))

  # Combine low and high WD data
  wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)

  # Generate height predictions for low and high WD
  prediction_df_ang <- data.frame()

  for (i in 1:nrow(wd_filtered_ang)) {
    sp <- wd_filtered_ang$sp[i]
    avg_log_a <- wd_filtered_ang$avg_log_a[i]
    avg_b <- wd_filtered_ang$avg_b[i]
    avg_k <- wd_filtered_ang$avg_k[i]
    avg_wd <- wd_filtered_ang$avg_wd[i]

    predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
    prediction_df_ang <- rbind(prediction_df_ang, temp_df)
  }

  # Plot
  p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Height", color = "WD Quantiles") +
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p3)
}

p3 <- generate_p3(fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h)
print(p3)
```

# 3 function in 1
```{r}

generate_wd_ef <- function(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr,
                           fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr,
                           fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h) {

  generate_p1 <- function() {
    beta_ang_cr_pl <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
    ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
    sp_id_mapping <- ang_cr_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    angiosperm_parameters <- beta_ang_cr_pl |>
      mutate(param_type = case_when(
        grepl("1,", variable) ~ "log_a",
        grepl("2,", variable) ~ "b"
      )) |>
      mutate(sp_id = rep(1:(nrow(beta_ang_cr_pl) / 2), each = 2))
    angiosperm_params_wide <- angiosperm_parameters |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    angiosperm_params_with_sp <- angiosperm_params_wide |>
      left_join(sp_id_mapping, by = "sp_id")
    angiosperm_data_full <- ang_cr_df |>
      left_join(angiosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b)
    species_summary <- angiosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_log_a = mean(log_a, na.rm = TRUE),
                avg_b = mean(b, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
    high_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))
    wd_filtered <- bind_rows(low_wd_species, high_wd_species)
    dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)
    prediction_df <- data.frame()
    for (i in 1:nrow(wd_filtered)) {
      sp <- wd_filtered$sp[i]
      avg_log_a <- wd_filtered$avg_log_a[i]
      avg_b <- wd_filtered$avg_b[i]
      avg_wd <- wd_filtered$avg_wd[i]
      predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
      prediction_df <- rbind(prediction_df, temp_df)
    }
    p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p1)
  }

  generate_p2 <- function() {
    beta_gym_cr_gmm <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
    gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")
    sp_id_mapping_gym <- gym_cr_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    gymnosperm_parameters <- beta_gym_cr_gmm |>
      mutate(param_type = case_when(
        grepl("\\[.*,1\\]", variable) ~ "log_a",
        grepl("\\[.*,2\\]", variable) ~ "b",
        grepl("\\[.*,3\\]", variable) ~ "k"
      )) |>
      mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))
    gymnosperm_params_wide <- gymnosperm_parameters |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    gymnosperm_params_with_sp <- gymnosperm_params_wide |>
      left_join(sp_id_mapping_gym, by = "sp_id")
    gymnosperm_data_full <- gym_cr_df |>
      left_join(gymnosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)
    species_summary_gym <- gymnosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_log_a = mean(log_a, na.rm = TRUE),
                avg_b = mean(b, na.rm = TRUE), avg_k = mean(k, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
    high_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))
    wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)
    dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)
    prediction_df_gym <- data.frame()
    for (i in 1:nrow(wd_filtered_gym)) {
      sp <- wd_filtered_gym$sp[i]
      avg_log_a <- wd_filtered_gym$avg_log_a[i]
      avg_b <- wd_filtered_gym$avg_b[i]
      avg_k <- wd_filtered_gym$avg_k[i]
      avg_wd <- wd_filtered_gym$avg_wd[i]
      predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
      prediction_df_gym <- rbind(prediction_df_gym, temp_df)
    }
    p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p2)
  }

  generate_p3 <- function() {
    beta_ang_h_wb <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
    ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")
    sp_id_mapping_ang <- ang_h_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    beta_ang_h_weibull <- beta_ang_h_wb |>
      mutate(param_type = case_when(
        grepl("\\[.*,1\\]", variable) ~ "log_a",
        grepl("\\[.*,2\\]", variable) ~ "b",
        grepl("\\[.*,3\\]", variable) ~ "k"
      )) |>
      mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))
    angiosperm_params_wide <- beta_ang_h_weibull |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    angiosperm_params_with_sp <- angiosperm_params_wide |>
      left_join(sp_id_mapping_ang, by = "sp_id")
    angiosperm_data_full <- ang_h_df |>
      left_join(angiosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)
    species_summary_ang <- angiosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_h = mean(h, na.rm = TRUE),
                avg_log_a = mean(log_a, na.rm = TRUE), avg_b = mean(b, na.rm = TRUE),
                avg_k = mean(k, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species_ang <- species_summary_ang |>
      filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
    high_wd_species_ang <- species_summary_ang |>
      filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))
    wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)
    dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)
    prediction_df_ang <- data.frame()
    for (i in 1:nrow(wd_filtered_ang)) {
      sp <- wd_filtered_ang$sp[i]
      avg_log_a <- wd_filtered_ang$avg_log_a[i]
      avg_b <- wd_filtered_ang$avg_b[i]
      avg_k <- wd_filtered_ang$avg_k[i]
      avg_wd <- wd_filtered_ang$avg_wd[i]
      predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
      prediction_df_ang <- rbind(prediction_df_ang, temp_df)
    }
    p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Height", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p3)
  }

  # Generate all three plots
  p1 <- generate_p1()
  p2 <- generate_p2()
  p3 <- generate_p3()

  # return(list(p1 = p1, p2 = p2, p3 = p3))
  p <- p1 + p2 + p3 + plot_layout(nrow = 1)

  return(p)
}

# Usage
p <- generate_wd_ef(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr,
                        fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr,
                        fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h)

```

```{r}

```

```{r}

```

### 2025-03-26 docs/wd_para_re.qmd
Initial commit for tree-allo

---
title: "Wood density vs. parameters"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
# Load necessary libraries
library(targets)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(patchwork)
```

```{r}
# 1. Load wood density data (assuming this is the original data)
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
# 2. Filter and clean wood density data
data2 <- tallo_wd_df_200 |>
  filter(!is.na(dbh)) |>     # Filter out NAs in dbh
  filter(!is.na(cr)) |>      # Filter out NAs in crown radius
  filter(!is.na(h)) |>       # Filter out NAs in height
  filter(!is.na(wd_s))       # Filter out NAs in scaled wood density

data2

# 3. Group by species and calculate the mean wood density
wd_df <- data2 |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>  # Calculate mean wood density per species
  mutate(wd_s = scale(wd) |> as.numeric(),  # Scale wood density
         sp_id = 1:n())                     # Assign species ID

# 4. Load Stan model data for height (H-DBH) and fit results
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
wb1_h <- tar_read(fit_nlr_draws_weibull1_h)

# 5. Load Stan model data for crown radius (CR-DBH) and fit results
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
wb1_cr <- tar_read(fit_nlr_draws_weibull1_cr)

# 6. Prepare sequences for wood density values for model fitting
wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)
wd_seq_df <- data.frame(
  wd_s = wd_seq,
  wd = approx(wd_df$wd_s, wd_df$wd, xout = wd_seq)$y
)
```

```{r}
prepare_and_plot_parameter <- function(stan_data, fit_draws, wd_df, param_name, color_point, color_ribbon, line_color, line_type = "solid", show_ribbon = TRUE) {

  # Step 1: Prepare wood density data, ensuring all columns exist
  wd_df <- wd_df |>
    filter(!is.na(wd_s)) |> 
    group_by(sp) |>
    summarize(wd = mean(wd, na.rm = TRUE)) |>
    mutate(wd_s = scale(wd) |> as.numeric(), sp_id = 1:n())

  # Step 2: Extract gamma and beta parameters from fit_draws
  gamma_params <- fit_draws |>
    select(starts_with("gamma"))

  # Step 3: Reshape the beta parameters for analysis
  beta_params <- fit_draws |>
    select(starts_with("beta[")) |>
    pivot_longer(
      cols = everything(),
      names_to = c("species", "parameter"),
      names_pattern = "beta\\[(\\d+),(\\d+)\\]",
      names_transform = list(species = as.integer, parameter = as.integer),
      values_to = "value"
    ) |>
    mutate(parameter = case_when(
      parameter == 1 ~ "log_a",
      parameter == 2 ~ "b",
      parameter == 3 ~ "k"
    ))

  # Step 4: Join wood density data with beta parameters
  plot_data <- beta_params |>
    left_join(wd_df, by = c("species" = "sp_id"))

  # Step 5: Summarize the posterior distributions for plotting
  summary_stats <- plot_data |>
    group_by(species, parameter, wd, wd_s) |>
    summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  # Step 6: Create a sequence for wd_s and merge with corresponding wd values
  wd_seq <- seq(min(wd_df$wd_s), max(wd_df$wd_s), length.out = 100)
  wd_seq_df <- data.frame(
    wd_s = wd_seq,
    wd = approx(wd_df$wd_s, wd_df$wd, xout = wd_seq)$y
  )

  # Step 7: Extract relevant gamma parameters for the chosen model
  gamma_param <- gamma_params |>
    select(paste0("gamma[1,", which(c("log_a", "b", "k") == param_name), "]"), 
           paste0("gamma[2,", which(c("log_a", "b", "k") == param_name), "]"))

  # Step 8: Generate lines and fit data for the selected parameter
  if (param_name == "log_a") {
    transform_value <- function(x) exp(exp(x))
  } else {
    transform_value <- function(x) exp(x)
  }

  lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
    data.frame(
      wd_s = wd_seq_df$wd_s,
      wd = wd_seq_df$wd,
      value = transform_value(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df$wd_s),
      draw = i
    )
  }))

  # Step 9: Compute the fit line and 95% credible intervals
  median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
  median_slope <- median(gamma_param[[2]], na.rm = TRUE)

  fit_line <- data.frame(
    wd_s = wd_seq_df$wd_s,
    wd = wd_seq_df$wd,
    value = transform_value(median_intercept + median_slope * wd_seq_df$wd_s)
  )

  ci_ribbon <- lines_data |>
    group_by(wd) |>
    summarize(
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

  # Apply transformation for log_a in summary_stats
  if (param_name == "log_a") {
    summary_stats <- summary_stats |>
      mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
  }

  # Step 10: Set y-axis label based on parameter
  y_label <- case_when(
    param_name == "log_a" ~ "Amplitude parameter, a",
    param_name == "b" ~ "Scale parameter, b",
    param_name == "k" ~ "Shape parameter, k"
  )

  # Step 11: Create plot
  plot <- ggplot() +
    geom_errorbar(data = summary_stats |> filter(parameter == param_name), 
                  aes(x = wd, ymin = lower, ymax = upper), 
                  width = 0, linewidth = 0.2, color = color_ribbon) +
    geom_point(data = summary_stats |> filter(parameter == param_name), 
               aes(x = wd, y = median), color = color_point, size = 0.5) +
    geom_line(data = fit_line, aes(x = wd, y = value), 
              color = line_color, linewidth = 1, linetype = line_type) +
    theme_minimal() +
    labs(
      x = "Wood Density (g cm⁻³)",
      y = y_label
    ) +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12)
    )

  # Step 12: Conditionally add ribbon if `show_ribbon` is TRUE
  if (show_ribbon) {
    plot <- plot +
      geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), 
                  alpha = 0.5, fill = color_ribbon)
  }

  return(plot)
}
```

```{r}
p1 <- prepare_and_plot_parameter(
  stan_data_nlr_h, wb1_h, wd_df, 
  param_name = "log_a",
  color_point = "#9e9ac8", 
  color_ribbon = "#cbc9e2", 
  line_color = "#6a51a3", 
  show_ribbon = TRUE
)

p2 <- prepare_and_plot_parameter(
  stan_data_nlr_h, wb1_h, wd_df, 
  param_name = "b",
  color_point = "#74c476", 
  color_ribbon = "#bae4b3", 
  line_color = "#238b45", 
  show_ribbon = TRUE
)

p3 <- prepare_and_plot_parameter(
  stan_data_nlr_h, wb1_h, wd_df, 
  param_name = "k",
  color_point = "#6baed6", 
  color_ribbon = "#8fcbee", 
  line_color = "#2171b5", 
  line_type = "dashed", 
  show_ribbon = FALSE
)

p4 <- prepare_and_plot_parameter(
  stan_data_nlr_cr, wb1_cr, wd_df, 
  param_name = "log_a",
  color_point = "#9e9ac8", 
  color_ribbon = "#cbc9e2", 
  line_color = "#6a51a3", 
  show_ribbon = FALSE
)

p5 <- prepare_and_plot_parameter(
  stan_data_nlr_cr, wb1_cr, wd_df, 
  param_name = "b",
  color_point = "#74c476", 
  color_ribbon = "#bae4b3", 
  line_color = "#238b45", 
  show_ribbon = FALSE
)

p6 <- prepare_and_plot_parameter(
  stan_data_nlr_cr, wb1_cr, wd_df, 
  param_name = "k",
  color_point = "#6baed6", 
  color_ribbon = "#8fcbee", 
  line_color = "#2171b5", 
  line_type = "dashed", 
  show_ribbon = FALSE
)
```

```{r}
p <- (p1 + labs(tag = "(a)", x = NULL) + theme(axis.text.x = element_blank()) |
                p4 + labs(tag = "(b)", x = NULL) + theme(axis.title.y = element_blank(), axis.text.x = element_blank()) |
                p2 + labs(tag = "(c)", x = NULL) + theme(axis.text.x = element_blank()) |
                p6 + labs(tag = "(d)", x = NULL) + theme(axis.title.y = element_blank(), axis.text.x = element_blank()) |
                p3 + labs(tag = "(e)", x = expression("Wood Density (g cm"^"-3"~")")) |
                p6 + labs(tag = "(f)", x = expression("Wood Density (g cm"^"-3"~")")) + theme(axis.title.y = element_blank())) +
    plot_layout(ncol = 2, nrow = 3) +
    plot_annotation(
      tag_levels = list(c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"))
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 10),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9),
      plot.margin = margin(10, 10, 10, 10)
    )

p <- ggdraw() +
  draw_plot(p, x = 0, y = 0, width = 1, height = 0.9) +  # Add the original plot
  draw_label("WD vs. H-DBH", x = 0.25, y = 0.91, fontface = 'bold', size = 10) +  # Title for the left column
  draw_label("WD vs. CR-DBH", x = 0.75, y = 0.91, fontface = 'bold', size = 10)   # Title for the right column
p
```

### 2025-03-26 ms/cover_letter.qmd
Initial commit for tree-allo

---
fontsize: 12pt
geometry: margin=1in
link-citations: true
format:
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
    prefer-html: true
  docx:
    toc: false
    include-in-header:
      text: |
        \usepackage{sourcecodepro}
---

February 13, 2025

Dear Editor,

We here submit our manuscript, "Saturating allometric relationships reveal how wood density shapes global tree architecture," for consideration for publication in the *Journal of Forestry Research*.

Selecting an appropriate allometric model is a critical step in estimating forest carbon biomass.
However, developing precise allometric models across regions characterized by diverse tree species composition and varying forest structures remains a persistent challenge.

In this study, we employed Bayesian hierarchical models and three alternative functional forms to characterize how the diameter at breast height, tree height, and crown radius scale with and without wood density as a species-level predictor.
By leveraging a global dataset, we overcome the limitations of previous studies that often relied on site-specific or regional observations with limited samples and species.
By integrating species-specific wood density into a multilevel framework, our methodology avoids the post-hoc correlations used in previous studies, offering a more universal and scalable model and effectively capturing variability from the species to the community level.

Our results emphasize the importance of applying saturating functions for large trees to improve forest biomass estimates and show that wood density, while not always predictive at broad scales, illuminates the biomechanical and ecological constraints underlying diverse tree architectures.

Relationships among tree-size variables are foundational to both ecological theory and practical forestry applications, informing everything from carbon stock estimation to dynamic vegetation modeling.
Our findings offer practical pathways for integrating height- and crown-based metrics into existing carbon monitoring programs worldwide, which we believe would be of significant interest to broad readers.

This manuscript is original, has not been published elsewhere, and is not under consideration by another journal.
All authors have approved the submission and have no conflicts of interest to declare.
The codes and data supporting the findings of this study will be deposited at Zenodo (<https://zenodo.org/>).
Allometry data can be downloaded from <https://doi.org/10.5281/zenodo.6637599>, and wood density data can be requested from <https://www.try-db.org/TryWeb/Home.php>.

Sincerely,

Masatoshi Katabuchi

On behalf of the authors:
Thi Duyen Nguyen,
Masatoshi Katabuchi

Corresponding author and contact information:

Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Mengla, Yunnan 666303, China

Email: katabuchi@xtbg.ac.cn; mattocci27@gmail.com

### 2025-03-26 ms/cover_letter.tex
Initial commit for tree-allo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners, frame hidden, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

February xx, 2025

Dear Editor,

We here submit our manuscript, ``Saturating allometric relationships
reveal how wood density shapes global tree architecture,'' for
consideration for publication in the \emph{Journal of Forestry
Research}.

Selecting an appropriate allometric model is a critical step in
estimating forest carbon biomass. However, developing precise allometric
models across regions characterized by diverse tree species composition
and varying forest structures remains a persistent challenge.

In this study, we employed Bayesian hierarchical models and three
alternative functional forms to characterize how the diameter at breast
height, tree height, and crown radius scale with and without wood
density as a species-level predictor. By leveraging a global dataset, we
overcome the limitations of previous studies that often relied on
site-specific or regional observations with limited samples and species.
By integrating species-specific wood density into a multilevel
framework, our methodology avoids the post-hoc correlations used in
previous studies, offering a more universal and scalable model and
effectively capturing variability from the species to the community
level.

Our results emphasize the importance of applying saturating functions
for large trees to improve forest biomass estimates and show that wood
density, while not always predictive at broad scales, illuminates the
biomechanical and ecological constraints underlying diverse tree
architectures.

Relationships among tree-size variables are foundational to both
ecological theory and practical forestry applications, informing
everything from carbon stock estimation to dynamic vegetation modeling.
Our findings offer practical pathways for integrating height- and
crown-based metrics into existing carbon monitoring programs worldwide,
which we believe would be of significant interest to broad readers.

This manuscript is original, has not been published elsewhere, and is
not under consideration by another journal. All authors have approved
the submission and have no conflicts of interest to declare. The codes
and data supporting the findings of this study will be deposited at
Zenodo (\url{https://zenodo.org/}). Allometry data can be downloaded
from \url{https://doi.org/10.5281/zenodo.6637599}, and wood density data
can be requested from \url{https://www.try-db.org/TryWeb/Home.php}.

Sincerely,

Masatoshi Katabuchi

On behalf of the authors: Thi Duyen Nguyen, Masatoshi Katabuchi

Corresponding author and contact information:

Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical
Botanical Garden, Chinese Academy of Sciences, Mengla, Yunnan 666303,
China

Email: katabuchi@xtbg.ac.cn; mattocci27@gmail.com

\end{document}

### 2025-03-26 ms/ms.qmd
Initial commit for tree-allo

---
fontsize: 12pt
geometry: margin=1in
link-citations: true
bibliography: "../ms/tree-allometry.bib"
csl: journal-of-forestry-research.csl
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
number-sections: true
format:
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \usepackage{setspace}
        \doublespacing
        \linenumbers
    prefer-html: true
  docx:
    prefer-html: true
    toc: false
    highlight-style: github
    html-math-method: katex
    reference-doc: my_template.docx
---

**Saturating allometric relationships reveal how wood density shapes global tree architecture**

$$ $$

Thi Duyen Nguyen^1,2^,
Masatoshi Katabuchi^1^

$$ $$

^1^ CAS Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China

^2^ University of Chinese Academy of Sciences, Beijing 100049, China

$$ $$

**Corresponding Authors**:

Masatoshi Katabuchi

E-mail: katabuchi@xtbg.ac.cn;
mattocci27@gmail.com

$$ $$

Thi Duyen Nguyen:
<https://orcid.org/0009-0003-5296-9569>

Masatoshi Katabuchi:
<https://orcid.org/0000-0001-9900-9029>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r include=FALSE}
library(targets)
library(tidyverse)
library(here)
library(kableExtra)
library(targets)
library(knitr)
library(gsubfn)
library(yaml)
library(flextable)
library(officer)

source("R/figs.R")
source("R/functions.R")
data_params <- read_yaml("data.yaml")
data_posterior <- read_yaml("posterior.yaml")
data_metrics <- read_yaml("agb_metrics_ang.yaml")
```

# Abstract {.unnumbered}

Allometric equations are fundamental tools in ecological research and forestry management, widely used for estimating above-ground biomass and production, serving as the core foundations of dynamic vegetation models.
Using global datasets from Tallo—a tree allometry and crown architecture database encompassing thousands of species—and TRY—a plant traits database—we fit Bayesian hierarchical models with three alternative functional forms (power-law, generalized Michaelis–Menten (gMM), and Weibull) to characterize how diameter at breast height (DBH), tree height (H), and crown radius (CR) scale with and without wood density as a species-level predictor.
Our analysis revealed that the saturating Weibull function best captured the relationship between tree height and DBH in both functional groups, while the CR-DBH relationship was best predicted by a power-law function in angiosperms and the gMM function in gymnosperms.
Although including wood density did not significantly improve predictive performance, it revealed important ecological trade-offs: lighter-wood angiosperms achieve taller mature heights more rapidly, and denser wood promotes wider crown expansion across clades.
We also found that accurately estimating DBH required considering both height and crown size, highlighting how these variables together distinguish trees of similar height but differing trunk diameters.
Our results emphasize the importance of applying saturating functions for large trees to improve forest biomass estimates and show that wood density, while not always predictive at broad scales, illuminates the biomechanical and ecological constraints underlying diverse tree architectures.
These findings offer practical pathways for integrating height- and crown-based metrics into existing carbon monitoring programs worldwide.

**Keywords:**
above ground biomass,
crown radius,
diameter at breast height,
tree allometry model,
tree height,
wood density

# Introduction

Mapping terrestrial carbon stocks is crucial to successfully implementing climate change mitigation programs [@Chave2014].
The accuracy of carbon stock estimates depends on the availability of reliable allometric models to deduce the aboveground biomass (AGB) of trees from census data [@Chave2014].
Selecting an appropriate allometric model is, therefore, a critical step in estimating forest carbon biomass [e.g., @Chave2014; @Molto2014].
Tree allometry describes the correlations between different biometric measurements of trees, such as diameter at breast height (DBH), tree height, and tree crown size [e.g., @Shinozaki1964a; @Niklas1994; @West1999].
These relationships are widely utilized to estimate tree architecture from DBH and to estimate forest biomass and production, serving as fundamental components of dynamic vegetation models [e.g., @Feldpausch2011; @Falster2017; @Song2024].
Despite significant progress, developing precise allometric models across regions characterized by diverse tree species composition and varying forest structures remains a persistent challenge.
This complexity highlights the need for improved global-scale allometric models that capture the diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of allometric relationships is a subject of ongoing debate.
While numerous papers have used various functional forms for allometric equations [e.g., @Thomas1996; @Feldpausch2011; @MartinezCano2019; @Song2024], power-law functions remain the most common choice for describing the allometric scaling of tree size variables, including height and crown size with DBH [e.g., @Sileshi2023; @Jucker2024].
The relationships among tree size variables often involve complex interactions influenced by various factors, such as species diversity, environmental conditions, and individual variability [e.g., @Poorter2012a; @Laurans2024].
Consequently, power-law models often fail to accurately capture the allometric relationships across the full range of tree sizes, leading to underestimation of dimensions for smaller trees and overestimation for larger trees [e.g., @Fayolle2016; @Song2024].
This highlights the need to adopt alternative functional forms—such as saturating relationships—that can more accurately reflect the complexity of allometric scaling in vegetation models [e.g., @Goodman2014; @Song2024].

Earth observation (EO) technologies, including remote sensing and Light Detection and Ranging (LiDAR), have revolutionized the monitoring of forests [e.g., @Jucker2017] by enabling the identification and measurement of individual tree heights and crown dimensions using airborne imagery [e.g., @Shendryk2016; @Jucker2017].
However, because existing allometric models still rely on stem diameter as a crucial input for biomass estimation, this reliance presents a challenge for effectively integrating EO technologies into AGB estimation [@Shendryk2016; @Jucker2017].
While ground-based LiDAR can measure DBH directly and is much more efficient than traditional ground surveys [@Tatsumi2023], it is still less common and convenient than aerial LiDAR.
Alternative approaches focusing on tree height and crown dimensions as model predictors are also necessary for remote sensing applications [@Jucker2017].
One potential solution is to estimate DBH from tree height and crown size and then use established biomass equations, which could be an effective strategy for integrating airborne imagery into carbon monitoring programs [@Jucker2017].

Accurately representing tree allometry relationships also requires considering underlying ecological constraints, particularly the interplay of plant functional traits and their trade-offs influence individual tree architecture and the composition of entire forest communities [e.g., @Iida2012; @LoubotaPanzou2018; @Yang2023].
Under closed-canopy conditions, strong light competition makes the balance between vertical and horizontal space use crucial for species coexistence [e.g., @Aiba2009; @Iida2012].
Among functional traits, wood density is closely linked to both plant life history strategies [@Fajardo2022a] and light demand [e.g., @King2006; @Aiba2009; @Poorter2012a].
Several studies suggest that wood density explains a trade-off between rapid stem expansion and broader crowns, shaping mechanical constraints on species with diverse architecture and light capture strategies [@Aiba2009; @Iida2012; @Yang2023].
However, these types of studies have examined this relationship post hoc, relating species-specific allometric parameters to wood density and risking Type I or Type II errors if the hierarchical data structure is overlooked [@Gelman2006].
One exception is @MartinezCano2019, who incorporated wood density into a hierarchical model of tropical trees and concluded it offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to delineate allometric relationships involving DBH, tree height (H), crown radius (CR), and their interspecific variations.
Using large datasets from the Tallo and TRY databases [@Jucker2022; @Kattge2020], which span a diverse array of tree forms across global forest types and climatic conditions, we overcome the limitations of previous studies that often relied on site-specific or regional observations with limited samples and species.
By integrating species-level wood density (WD) into a multilevel model, our methodology avoids the post-hoc correlations used in previous studies, offering a more universal and scalable model, effectively capturing variability from the species to the community level.
Our main objective is to develop models informed by tree allometric variables (H, CR, DBH, WD) to capture interspecific variation in allometric scaling globally and offer an improved framework for estimating above ground biomass.
Specifically, we ask the following questions:
(1) What is the best function to estimate tree architecture from DBH and *vice versa*?
(2) How does wood density shape interspecific variation in these allometric scaling relationships?
(3) How does the choice of allometric relationships affect subsequent AGB estimation?

# Materials and Methods

## Data

Our analysis combined tree allometry data from Tallo, which includes standardized measurements of DBH, height, and crown radius for 5,163 species across various biomes [@Jucker2022, https://doi.org/10.5281/zenodo.6637599] along with wood density data for 10,995 species from TRY version 6 [@Kattge2020, https://www.try-db.org/TryWeb/Home.php] (see Supplementary Table S1 for further details).
We thoroughly processed the Tallo dataset by removing 998 outliers following @Jucker2022.
To ensure consistency, we selected only species with standardized wood density values in g cm^-3^ from TRY and removed any records with missing data.
Subsequently, we created an overlapping dataset between Tallo and TRY, resulting in a combined dataset encompassing allometric and wood density information for `r data_params$ori_sample_size` individuals across `r data_params$ori_species_number` species.

To reduce computational complexity and ensure the robustness of the models, we further filtered species with fewer than 20 individuals and randomly sampled up to 100 individuals per species.
This step was necessary because the dataset is unbalanced; for example, *Quercus ilex* alone accounts for over 20,000 individuals, representing approximately 10\% of the dataset.
However, *Quercus ilex* is not necessarily the most abundant species globally.
Its prevalence in this dataset is a result of sampling and does not reflect its true global abundance.
By limiting the number of individuals per species, we minimize bias due to data imbalance and prevent the overrepresentation of any single abundant species [@Gelman2006], thereby ensuring that the model remains broadly applicable.
Since not all trees have a complete set of information, we created specific sub-datasets to select the best predictive model for each allometric relationship for two clades (i.e., angiosperms and gymnosperms; see below and Supplementary Table S2 for more details).

## Tree allometry models

We developed hierarchical Bayesian models to estimate the allometric relationships between tree architectural variables, including DBH (cm), tree height (H, m), crown radius (CR, m), and wood density (g cm^-3^).
We assumed a Gaussian likelihood ($N$) for the natural logarithm of the response variable, $y_{(i,j)}$ representing either tree height, crown radius, or DBH for each tree $i$ in species $j$:

$$
\log(y_{i,j}) \sim N(f(X_{i,j}, \pmb{\beta}_{j}), \sigma_y),
$$ {#eq-like}

where $f(\cdot)$ predicts the expected natural log tree height or crown radius from observed DBH (and *vice versa*), $\pmb{\beta}_{j}=\{a_j, b_j, k_j\}$ denotes the vector of species-specific parameters, and $\sigma_y$ represents the standard deviation of the observation error.
We considered three functional forms for the relationship between tree allometric variables [@Sullivan2018; @MartinezCano2019]:

Power-law:
$$
y = aX^b,
$$ {#eq-pl}

Generalized Michaelis-Menten (gMM):
$$
y = \frac{aX^b}{k + X^b},
$$ {#eq-gMM}

Weibull:
$$
y = a\left\{1-\mathrm{exp}(-bX^k)\right\},
$$ {#eq-wb}

The power-law model (@eq-pl) is one of the simplest and most commonly used functions to describe biological scaling relationships [e.g., @West1999; @Brown2004].
We also tested two nonlinear models, the generalized Michaelis–Menten (gMM) (@eq-gMM) and the Weibull function (@eq-wb), to account for a saturating relationship between tree dimensions and DBH [e.g., @MartinezCano2019].

We applied the power-law model to estimate tree height and crown radius from DBH, as well as DBH from tree height and/or crown radius.
Specifically, when predicting DBH from height and/or crown radius, we tested four formulations:
(1) using CR and H as separate factors, $\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \mathrm{H}^{c_1}$, denoted as $CR,H$;
(2) using CR and H as a compound term, $\mathrm{DBH} = a_2 (\mathrm{CR} \times \mathrm{H})^{b_2}$, denoted as $CR \times H$;
(3) using only height, $\mathrm{DBH} = a_3 \mathrm{H}^{b_3}$; and
(4) using only crown radius, $\mathrm{DBH} = a_4 \mathrm{CR}^{b_4}$.
In contrast, the gMM and Weibull models were only used to estimate tree height and crown radius from DBH.
Because these two models specifically describe saturating relationships (@eq-gMM; @eq-wb), their formulation does not allow simply switching the roles of explanatory and response variables.

To compare the performance of different models for the same response variable, we used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2017).
PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), which is a robust method for comparing models with different numbers of parameters [@Vehtari2017].
We utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
A model with lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].

Once the best-performing models for each allometric relationship were identified, we tested whether adding wood density as a species-level predictor would further improve the predictive performance of those models and also examined its influence on tree architecture diversity.
Species-specific parameters ($\beta_j$) were modeled as a function of wood density ($u_j$):
$$
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
$$ {#eq-wood}

where $\gamma_{0}$ and $\gamma_{1}$ are the species-level intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
To facilitate interpretation and visualization, species were then classified into "light" and "dense" wood groups based on the 10th and 90th percentiles of wood density distribution in the observed dataset, respectively.
This classification allowed us to illustrate the relationships between wood density and the parameters of the best-performing models, enabling an evaluation of how wood density contributes to variation in tree architecture.
Additionally, we explored a model that accounts for covariance between species-specific parameters using a multivariate normal distribution.
However, due to the complexity of this approach, the model did not converge for the nonlinear models, and thus, we did not include it in the final analysis.

Posterior distributions of all parameters were estimated using the No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte Carlo (HMC), algorithm implemented in Stan [@Carpenter2017] using weakly-informative priors [@Gelman2008].
The HMC algorithm uses gradient information to propose new states in the Markov chain, leading to a more efficient exploration of the target distribution than traditional Markov chain Monte Carlo (MCMC) methods that rely on random proposals [@Carpenter2017].
This efficiency allows us to achieve convergence with fewer iterations than traditional MCMC methods.
Inferences were based on 1,000 posterior samples following 1,000 warmup iterations for three parallel chains.
Convergence was assessed using the Gelman–Rubin statistic with a convergence threshold of 1.1 [@Gelman2013] and ensuring effective sample sizes greater than 300 [@Vehtari2021].
We also reported the median and 95\% Bayesian credible intervals (CIs) for each parameter (Supplementary Tables S3-S8).
Estimates were considered significant when 95\% CIs did not overlap with zero.

Statistical analyses were performed in R version 4.3.2 [@RCoreTeam2023], using the R packages *targets* version 1.7.0 for workflow management [@Landau2021].
Codes are made available on GitHub at <https://github.com/duyenecology/tree-allometry> (Stan code is provided in Supplementary Section S2).

## Aboveground biomass estimates

For angiosperm species, we used the widely applied equation from @Chave2014 to estimate aboveground biomass (AGB; kg of dry mass):

$$
\mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
$${#eq-agb}

where $\rho$ is wood density (g cm^-3^), $D$ (cm) is DBH, and $H$ (m) is tree height.
We either used both observed DBH and height or substituted one of these variables with an estimate from our allometric models, allowing us to assess how different allometric relationships affect AGB calculations.
Only trees with a DBH ≥ 5 cm were used, as those below this threshold hold a small fraction of forest carbon stocks and were excluded from the calibration of @Chave2014's equation.
Additionally, the sub-datasets used to select the optimal predictive model for each allometric relationship differed as not all trees had a complete information set (see above and Supplementary Table S2).
Consequently, the final dataset for estimating AGB in angiosperms includes `r data_metrics$metrics_agb_ang$num_trees` trees with sufficient allometric information (i.e., DBH, H, CR, WD) across `r data_metrics$metrics_agb_ang$num_species` species.
We did not extend this approach to gymnosperms because no general AGB equation is available for them, and most existing equations are species-specific and do not incorporate wood density [e.g., @Moore2010; @Bazrgar2024].

# Results

## Tree height allometry

The optimal model to predict tree height was a Weibull function for both angiosperms (ang) and gymnosperms (gym) (Table 1).
Across the global dataset, the best model to predict H from DBH of each clade was:
$$
\mathrm{H_{ang}} =`r sprintf("%.1f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$a$Intercept_CI$Median)`
\left\{ 1 - \exp\left( - `r sprintf("%.4f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$b$Intercept_CI$Median)`
\mathrm{DBH}^{`r sprintf("%.3f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$k$Intercept_CI$Median)`}
\right) \right\}
$$ {#eq-h_ang}

$$
\mathrm{H_{gym}} = `r sprintf("%.1f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$a$Intercept_CI$Median)`
\left\{ 1 - \exp\left( - `r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$b$Intercept_CI$Median)`
\mathrm{DBH}^{`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$k$Intercept_CI$Median)`}
\right) \right\}.
$$ {#eq-h_gym}

The posterior medians with their 95\% CIs were as follows:
asymptote $a_{\text{ang}}$ =
`r sprintf("%.1f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$a$Intercept_CI$Median)`
[`r sprintf("%.1f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.1f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$a$Intercept_CI$Upper)`],
$a_{\text{gym}}$ =
`r sprintf("%.1f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$a$Intercept_CI$Median)`
[`r sprintf("%.2f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.1f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$a$Intercept_CI$Upper)`];
exponent $b_{\text{ang}}$ =
`r sprintf("%.4f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$b$Intercept_CI$Median)`
[`r sprintf("%.4f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.4f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$b$Intercept_CI$Upper)`],
$b_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$b$Intercept_CI$Upper)`];
and shape parameter $k_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$k$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$k$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Tree Height'$Angiosperm$Weibull$Without$k$Intercept_CI$Upper)`],
$k_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$k$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$k$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Tree Height'$Gymnosperm$Weibull$Without$k$Intercept_CI$Upper)`].

The second-best model, which used the gMM function, showed poorer predictive performance in angiosperms ($\Delta$LOOIC = 101.01) but showed competitive predictive performance in gymnosperms ($\Delta$LOOIC = 4.87).
While power-law models (@eq-pl) can exhibit saturating behavior when 0 < *b* < 1, they performed the worst overall compared to other saturating models (@eq-gMM; @eq-wb).
Incorporating wood density did not significantly improve predictive performance (Table 1).

**Table 1**: Summary of the results of the model selection procedure.
Models were ranked using the Leave-One-Out Information Criterion (LOOIC) to assess their predictive performance.
For each model, we report the Expected Log Predictive Density (ELPD), the LOOIC, and the $R^2$ values.
The difference in LOOIC relative to the best-performing model, $\Delta$LOOIC, is also included, with a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance [@Sivula2023; @Cochrane2023].
Lower $\Delta$LOOIC values indicate models that perform closer to the best model in terms of predictive accuracy, while higher ELPD values reflect better model fit.
Higher $R^2$ values reflect better explanatory power, indicating how well the model explains the variance in the dependent variable.

\newpage
```{r echo=FALSE, results='asis', eval=TRUE}
loo_r2_tbl <- read_csv("data/loo_r2_tbl.csv")
loo_r2_tbl$Division <- ifelse(loo_r2_tbl$Division == "Angiosperm", "Angiosperm", "Gymnosperm")
loo_r2_tbl <- loo_r2_tbl |> dplyr::select(-pLOO)

dependent_variable_order <- c("Tree Height", "Crown Radius", "DBH")

loo_r2_tbl$`Dependent variable` <- factor(loo_r2_tbl$`Dependent variable`, levels = dependent_variable_order)

loo_r2_tbl <- loo_r2_tbl |> arrange(`Dependent variable`, Division)

dep_var_border_rows <- which(diff(as.numeric(loo_r2_tbl$`Dependent variable`)) != 0)

division_border_rows <- which(diff(as.numeric(factor(paste(loo_r2_tbl$`Dependent variable`, loo_r2_tbl$Division)))) != 0)

ft <- loo_r2_tbl |>
  flextable() |>

  set_header_labels(
    `Dependent variable` = "Dependent\nvariable",
    `Predictor variable` = "Predictor\nvariable",
    Division = "Division",
    `Functional form` = "Functional\nform",
    Wood_density = "Wood\ndensity",
    ELPD = "ELPD",
    LOOIC = "LOOIC",
    ΔLOOIC = "ΔLOOIC",
    R2 = "R²"
  ) |>

  align(j = colnames(loo_r2_tbl), align = "center", part = "all") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

ft <- ft |> border(i = dep_var_border_rows, border.bottom = fp_border(width = 1, color = "black"))

ft <- ft |> border(i = division_border_rows, j = 2:ncol(loo_r2_tbl), border.bottom = fp_border())

ft <- ft |> merge_v(j = 1)

if (knitr::is_latex_output()) {
  loo_r2_tbl |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage

![
Allometric relationships between DBH and tree height or crown radius in angiosperms (a, c) and gymnosperms (b, d), based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
A color gradient from black (low density) to yellow (high density) indicates the density of overlapping data points, calculated via 2D kernel density estimation on a 300 $\times$ 300 grid.
See Supplementary Figure S1 for the log-scale version of these plots.
](../figs/h_cr_dbh_non_log.png){#fig-compare}

## Crown Radius allometry

The power-law function (0 < *b* < 1) provided the best predictive performance for modeling the relationship between crown radius and DBH in angiosperms.
Whereas for gymnosperms, the gMM function offered the best performance for this allometric scaling (Table 1).
Across the global dataset, the best model for predicting CR from DBH of each clade was:
$$
\mathrm{CR_{ang}} = \mathrm{`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Median)`}
\, \mathrm{DBH}^{\mathrm{`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`}}
$$ {#eq-cr_ang}

$$
\mathrm{CR_{gym}} = \frac{\mathrm{`r sprintf("%.1f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$a$Intercept_CI$Median)`}
\, \mathrm{DBH}^{\mathrm{`r sprintf("%.3f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$b$Intercept_CI$Median)`}}}
{\mathrm{`r sprintf("%.0f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$k$Intercept_CI$Median)`}
+ \mathrm{DBH}^{\mathrm{`r sprintf("%.3f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$b$Intercept_CI$Median)`}}}.
$$ {#eq-cr_gym}

The posterior medians with their 95\% CIs were as follows:
scaling factor $a_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Upper)`],
asymptote $a_{\text{gym}}$ =
`r sprintf("%.1f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$a$Intercept_CI$Median)`
[`r sprintf("%.1f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.0f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$a$Intercept_CI$Upper)`];
exponent $b_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`],
$b_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$b$Intercept_CI$Upper)`];
and scaling constant $k_{\text{gym}}$ =
`r sprintf("%.0f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$k$Intercept_CI$Median)`
[`r sprintf("%.1f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$k$Intercept_CI$Lower)`,
`r sprintf("%.0f", data_posterior$'Crown Radius'$Gymnosperm$gMM$Without$k$Intercept_CI$Upper)`].

For angiosperms, the saturating functions (@eq-gMM; @eq-wb) performed worse than the power-law model, with $\Delta$LOOIC values of 124.57 for gMM and 126.35 for the Weibull function.
In gymnosperms, the Weibull function showed a competitive performance compared to the gMM function, while the power-law function (@eq-pl) performed the worst, with a $\Delta$LOOIC of 162.72.
Incorporating wood density did not significantly improve predictive performance (Table 1).

**Table 2**: Posterior estimates of the parameters for the best hierarchical models selected for tree height, crown radius, and DBH allometries in angiosperms and gymnosperms.
Each set includes posterior estimates both without and with the inclusion of wood density, reported as median and 95% Bayesian credible intervals for the community-level parameters of each allometric function.
For tree height, the Weibull model was identified as the best for both clades.
Crown radius was best described by a power-law model in angiosperms and a gMM model for gymnosperms.
The best model for DBH allometry in both clades, which incorporates both tree height and crown radius information ($CR,H$ model).

\newpage
```{r echo=FALSE, results='asis', eval=TRUE}
best_posterior_df <- read_csv("data/best_posterior_df.csv")

best_posterior_df <- best_posterior_df |>
  mutate(
    Intercept_CI = sapply(Intercept_CI, format_ci),
    Slope_CI = sapply(Slope_CI, format_ci),
    Tau_CI = sapply(Tau_CI, format_ci)
  )

dependent_variable_order <- c("Tree Height", "Crown Radius", "DBH")

best_posterior_df$Dependent_variable <- factor(best_posterior_df$Dependent_variable, levels = dependent_variable_order)

best_posterior_df <- best_posterior_df |> arrange(Dependent_variable, Division, Wood_density)

dep_var_border_rows <- which(diff(as.numeric(best_posterior_df$Dependent_variable)) != 0)

division_border_rows <- which(diff(as.numeric(factor(paste(best_posterior_df$Dependent_variable, best_posterior_df$Division)))) != 0)

wood_density_border_rows <- which(diff(as.numeric(factor(paste(best_posterior_df$Dependent_variable, best_posterior_df$Division, best_posterior_df$Wood_density)))) != 0)

ft <- best_posterior_df |>
  flextable() |>

  set_header_labels(
    Dependent_variable = "Dependent\nvariable",
    Division = "Division",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>

  align(j = colnames(best_posterior_df), align = "center", part = "all") |>
  fontsize(size = 5.5, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

ft <- ft |> merge_v(j = c(1, 2))

ft <- ft |> border(i = dep_var_border_rows, border.bottom = fp_border(width = 1, color = "black"))

ft <- ft |> border(i = division_border_rows, j = 2:ncol(best_posterior_df), border.bottom = fp_border())

ft <- ft |> border(i = wood_density_border_rows, j = 3:ncol(best_posterior_df), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  best_posterior_df |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

## DBH allometry

The $CR,H$ model, which treated CR and H as separate factors, exhibited the best predictive performance among all tested models (Table 1).
Across the global dataset, the most effective model for predicting DBH from H and CR for each clade was:
$$
\mathrm{DBH_{ang}} = \mathrm{`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Median)`}
\mathrm{CR}^{\text{`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`}}
\mathrm{H}^{\text{`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$c$Intercept_CI$Median)`}}
$$ {#eq-dbh_ang}

$$
\mathrm{DBH_{gym}} = \mathrm{`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$a$Intercept_CI$Median)`}
\mathrm{CR}^{\text{`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$b$Intercept_CI$Median)`}}
\mathrm{H}^{\text{`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$c$Intercept_CI$Median)`}}.
$$ {#eq-dbh_gym}

The posterior medians with their 95\% CIs were as follows:
scale $a_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$a$Intercept_CI$Upper)`],
$a_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$a$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$a$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$a$Intercept_CI$Upper)`];
exponent for CR $b_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`],
$b_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`];
and exponent for H $c_{\text{ang}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$c$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$c$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Angiosperm$'Power-Law'$Without$c$Intercept_CI$Upper)`],
$c_{\text{gym}}$ =
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$c$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$c$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'DBH'$Gymnosperm$'Power-Law'$Without$c$Intercept_CI$Upper)`].
Incorporating wood density did not significantly improve predictive performance (Table 1).

## Wood density and tree architectural diversity

Although including wood density did not significantly improve the predictive performance of any of the allometric relationships, it can enhance our understanding of the underlying biological mechanisms driving tree architectural diversity.
In H-DBH allometry, the shape parameter $k$ in angiosperms showed a significant negative relationship with wood density, while the asymptote parameter $a$ and the scale parameter $b$ showed no significant associations (@fig-wood1 a, b, c; @fig-wood2 a).
In gymnosperms, none of the parameters significantly correlated with wood density (@fig-wood1 d, e, f).

In CR-DBH allometry, angiosperms exhibited a positive relationship between the scaling factor $a$ and wood density and a negative association between the exponent parameter  $b$ and wood density (@fig-wood1 g, h; @fig-wood2 b).
In gymnosperms, we found a positive association between wood density and the exponent parameter $b$, whereas neither the asymptote parameter $a$ nor the scaling constant $k$ showed significant relationship with this functional trait (@fig-wood1 i, j, k; @fig-wood2 c).

![
Relationships between wood density and species-specific allometry parameters for tree height (a, b, c, d, e, f) and crown radius (g, h, i, j, k) of angiosperms (purple) and gymnosperms (green).
Points represent posterior medians of species-specific parameters, with vertical bars indicating 95% Bayesian credible intervals (CIs).
Thick solid lines represent significant relationships across species, with the shaded area illustrating the 95% CIs, while dashed lines indicate non-significant relationships.
](../figs/wd_com.png){#fig-wood1 fig-pos="H"}

![
Effect of wood density on tree height allometry in angiosperms (a) and on crown radius allometry in angiosperms (b) and gymnosperms (c).
Curves represent the allometric scaling for species with light (orange) and dense (brown) wood density, corresponding to the 10\% and 90\% quantiles of wood density distribution in the observed dataset, respectively.
The shaded areas illustrate the 95% Bayesian credible intervals.
Only statistically significant associations from @fig-wood1 are plotted.
](../figs/wd_ef.png){#fig-wood2 fig-pos="H"}

## Estimating aboveground biomass

AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based both on the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_pl_h)` Mg and bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_pl_h * 100)`% (@fig-agb b).
In contrast, models that used estimated DBH performed substantially worse (@fig-agb c-f).

![
Comparison of individual tree above-ground biomass (AGB) estimated using different approaches with reference AGB ($\text{AGB}_\text{ref}$) derived from observed DBH and tree height in angiosperms.
(a, b) AGB estimates based on observed DBH and tree height predicted by the Weibull function ($\text{AGB}_\text{H-WB}$) or the power-law function ($\text{AGB}_\text{H-PL}$).
(c–f) AGB estimates based on observed tree height and DBH predicted from crown radius and tree height as separate factors ($\text{AGB}_{\text{DBH} \sim \text{CR,H}}$), as a compound term ($\text{AGB}_{\text{DBH} \sim \text{CR × H}}$), from tree height alone ($\text{AGB}_{\text{DBH} \sim \text{H}}$), or from crown radius alone ($\text{AGB}_{\text{DBH} \sim \text{CR}}$).
Dashed lines represent the 1:1 relationships, and the solid lines show linear regression fits to the data, highlighting how predictive accuracy varies with tree size.
RMSE (root mean square error) and bias values for each model are reported.
](../figs/agb_ang.png){#fig-agb fig-pos="H"}

# Discussion

Relationships among tree-size variables are foundational to both ecological theory and practical forestry applications, informing everything from carbon stock estimation to dynamic vegetation modeling.
In this study, we tested multiple functional forms—power-law, generalized Michaelis–Menten (gMM), and Weibull—within a Bayesian hierarchical framework to identify which best describes the linkages between diameter at breast height (DBH), tree height (H), and crown radius (CR) across a global dataset.
Our results reveal that the Weibull function most reliably captures the saturating nature of the H–DBH relationship in both angiosperms and gymnosperms, while the CR–DBH relationship diverges by clade, being well described by a power-law in angiosperms but requiring a gMM form in gymnosperms.
Although incorporating wood density did not significantly enhance predictive performance, it illuminated interspecific architectural constraints and trade-offs, underscoring the complex interplay between mechanical support, crown expansion, and life history strategies.
Below, we discuss these findings in the context of tree allometry models, the role of functional traits, and forest biomass estimation.

## Tree allometry

Our result supported a saturating H-DBH relationship in both angiosperm and gymnosperm species, consistent with theoretical expectations [@Falster2003; @Niklas2007] and empirical observations [e.g., @Feldpausch2011; @MartinezCano2019; @Song2024].
A recent study using Tallo suggested the Weibull function as the best model for six plant functional types (PFTs) [@Song2024].
While PFTs-based approaches can illuminate large-scale patterns, they may obscure species-specific variation in allometric scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns (Supplementary Table S6) despite underestimating the height of individuals with DBH > 30 cm due to substantial interspecific variation in H–DBH allometric scaling (@fig-compare).
This highlights the advantages of a multilevel modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic scaling relationships among different axes of tree size, with an exponent close to 2/3 ($H \propto DBH^{2/3}$, $CR \propto DBH^{2/3}$), based on assumptions regarding mechanical and hydraulic constraints to plant growth [@West1999].
However, our findings revealed that the power-law model exponent of angiosperms (*b* = 0.538) was significantly smaller, while the actual exponent in gymnosperms (*b* = 0.704) was larger than the theoretical value of 2/3 (Supplementary Table S3).
This divergence highlights the importance of separately modeling the allometric relationships of each functional group to accurately capture their distinct scaling patterns.
Moreover, the power-law function performed worse than the saturating functions (Table 1), indicating that the actual H-DBH relationship does not follow the simple scaling laws proposed by MST.

Our analysis also revealed divergent patterns in CR-DBH allometry, highlighting fundamental differences in growth strategies and ecological adaptations between angiosperms and gymnosperms.
In angiosperms, the CR–DBH relationship follows a power-law form, suggesting a capacity for plastic crown expansion in competitive environments [e.g., @Poorter2012a; @Lines2012; @Hulshof2015], a conclusion consistent with a study in tropical moist forests by @MartinezCano2019.
In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent *b* is near the expected 2/3
($b_{\text{ang}}$ = `r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`], Table 2).
In contrast, the CR-DBH allometry in gymnosperms showed an asymptotic relationship best described by the gMM function, with the Weibull function also showing competitive predictive performance (Table 1).
This asymptotic pattern likely arises from structural constraints in gymnosperms growth forms that favour vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].

Finally, we tested models to estimate DBH from biometric variables (i.e., tree height and crown radius) measurable via Earth observation (EO) platforms, such as remote sensing and LiDAR [@Filipescu2012; @Jucker2017].
Our analysis showed that accurately predicting DBH requires both tree height and crown radius [@Jucker2017], which is especially important for distinguishing among trees of similar height but varying DBH [@King2005; @King2011].
Among the tested models, the $CR,H$ model ($\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}$) provided the highest predictive accuracy, a result that contrasts with the previous study [@Jucker2017], which concluded that the $CR \times H$ model ($\mathrm{DBH} = a_2(\mathrm{CR} \times \mathrm{H})^{b_2}$) was the best for predicting DBH.
The $CR,H$ model significantly improves predictive accuracy by accounting for the distinct contributions of crown radius and tree height as independent factors in DBH estimation.
The implications of these differences for biomass estimation are explored in detail in a later section, where we assess how DBH prediction accuracy affects allometric biomass calculations.

## Wood density and tree architecture diversity

For the H-DBH relationship, wood density does not strongly affect the initial growth rate of angiosperm trees but plays a critical role in determining maximum height and the growth trajectory to achieve that height, with lighter wood species tending to grow taller and attain their mature height more rapidly than those with denser wood (@fig-wood1 a-c; @fig-wood2 a).
In contrast, gymnosperms exhibit a height growth strategy that is largely independent of wood density (@fig-wood1 d-f).
These results are consistent with the previous study [@Poorter2012b], which showed that among seven angiosperm species, those with lighter wood exhibited taller maximum heights, while no significant correlations were observed for six gymnosperm species.
A recent study using the Tallo database alongside two additional datasets reported a nonsignificant relationship between tree height and wood density [@Jucker2024].
Their analysis included functional groups (angiosperms vs. gymnosperms) as a fixed effect, providing valuable insights into group-level patterns.
However, it did not incorporate interactions between wood density and functional groups, and thus, the effects of wood density were modeled as uniform across angiosperms and gymnosperms.
In contrast, our approach explicitly modeled these differences, enabling us to capture functional group-specific relationships and provide a more nuanced understanding of height allometry.

In the CR-DBH allometry, our results indicate that both angiosperms and gymnosperms with denser wood tend to have wider crowns across all trunk sizes (@fig-wood1 g-k).
Although the effect of wood density on CR-DBH allometry is significant, the crown expansion patterns do not clearly separate light-wood from dense-wood species (@fig-wood2 b, c).
This may suggest that crown radius is more plastic and responsive to local environmental conditions (e.g., light availability, stand density) than tree height, potentially obscuring the relationship between wood density and crown radius.
Nevertheless, our findings are consistent with previous work in temperate forests [@Aiba2009], tropical forests [@Iida2012], subtropical forests [@Yang2023], and on a global scale [@Jucker2024], all of which similarly report that denser wood promotes more efficient horizontal crown expansion.

In the analysis of DBH allometry, no significant correlation was observed between wood density and DBH in either angiosperms or gymnosperms (Table 2).
This outcome aligns with observations from saplings in an Australian subtropical forest [@Kooyman2009] but contradicts the expectation that species with lower wood density develop thicker stems to ensure structural stability, as suggested by prior studies [@King2006; @Anten2009; @Iida2012].
Although wood density is often regarded as a key functional trait [e.g., @Anten2009; @Jucker2024], variations in tree allometry across individuals are likely influenced by a combination of species identity and environmental factors.
Across species or along broad environmental gradients, wood density systematically varies—tending to be denser in drier, colder climates and in nutrient-poor soils [@Muller-Landau2004; @Chave2009; @Fajardo2022b; @Mo2024]—reflecting species’ adaptive strategies.
Such environmental influences might confound simple correlations between wood density and DBH when examining diverse tree communities.
At the intraspecific level, recent findings suggest that wood density may be more closely related to vessel anatomy than to tree architecture across different environments [@Fajardo2022b].
Furthermore, substantial measurement errors in tree height and crown radius relative to DBH may introduce noise that obscures potential relationships between wood density and stem diameter.
These highlight the importance of adopting a more integrative framework that accounts for additional functional traits and environmental variables to deepen our understanding of the factors driving DBH growth.

Our analysis indicated distinct patterns in which wood density has a greater influence on the interspecific variation in angiosperms than gymnosperms.
Variation in wood density has consequential implications for the mechanical properties of wood and hence influences the mechanical stability of stems and branches [@Sterck1998], ultimately determining the architecture of trees [e.g., @Poorter2006].
Angiosperms have denser wood and a wider wood density range than gymnosperms (see Supplementary Table S2), resulting in greater and more flexible mechanical support [e.g., @Gelder2006; @Swenson2007; @Mo2024].
This allows them to develop a more diverse height and crown to optimize light capture, particularly in environments where light competition is intense, such as tropical regions.
In contrast, gymnosperms predominantly live and often dominate in extreme environments such as boreal and high-elevation regions [@Swenson2007; @Mo2024], prioritising survival and longevity [@Fossdal2024].
Due to their simpler and more homogeneous wood anatomy, these species display lighter wood and a narrower range of wood density [e.g., @Mo2024], which may result in less flexibility in the crown architecture.
These differences underscore how the contrasting evolutionary strategies of angiosperms and gymnosperms shape their architectural diversity, reflecting adaptations to distinct ecological niches.

## Consequences for AGB estimates

Our findings revealed minimal differences in AGB estimates derived from $\text{H}_\text{WB}$ and $\text{H}_\text{PL}$ (@fig-agb a, b).
This result can stem from the fact that over 97\% of the trees used to estimate AGB are small to medium-sized (DBH < 70 cm [@Slik2013; @Murphy2022]), whose H-DBH relationships can still be well captured by the power-law function (@fig-compare a) [e.g., @Molto2014].
In contrast, individuals with DBH ≥ 70 cm, while typically forming a small proportion of the tree population, contribute significantly to biomass storage [e.g., @Fauset2015].
For instance, @Slik2013 reported that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the dataset in tropical forests, yet stored 25-45\% of the total AGB on average.
However, the fundamental assumption of the power-law function does not fully reflect the asymptotic nature of height-diameter allometries in large trees (@fig-compare a), potentially leading to significant inaccuracies in stand-level biomass estimates [e.g., @Feldpausch2011; @Fayolle2016; @Sullivan2018; @MartinezCano2019].
Because large trees contribute disproportionately to overall biomass  [e.g., @Slik2013; @Ligot2018; @Fauset2015], if direct height measurements are unavailable, height estimates should be derived by fitting asymptotic models (e.g., @eq-gMM; @eq-wb) to datasets with sufficient representation of large trees, ensuring that the saturating component is accurately captured [@Sullivan2018].

In the scenarios where DBH is not directly measured, models estimating AGB using estimated DBH consistently perform poorly (@fig-agb c, d, e, f).
Several reasons may account for this underperformance.
Since the gMM and Weibull models (@eq-gMM; @eq-wb) do not allow simply switching the roles of explanatory and response variables.
Consequently, we relied solely on the power-law function to estimate DBH from tree height and crown radius, which may oversimplify the complex interactions between tree-size variables, potentially leading to inaccurate DBH estimates [e.g., @Filipescu2012].
Additionally, most existing AGB estimation equations typically include DBH as a squared term in equations [e.g., @Brown1989; @Chave2014], meaning that even minor errors in DBH predictions can significantly affect the accuracy of AGB estimates [@Jucker2017].
Interestingly, our results revealed that, although the $CR,H$ model provided more accurate direct DBH predictions, using those predictions in the AGB equation resulted in higher RMSE and bias than when DBH estimates from the $CR \times H$ model were used (@fig-agb c, d).
This is likely due to the $CR \times H$ model's ability to capture the inherent interdependence between the crown radius and tree height (*r* = 0.703), offering a more biologically realistic framework [e.g., @Dormann2013].
The observed importance of crown radius in improving AGB estimation aligns with prior studies [e.g., @Henry2010; @Goodman2014], whereas our finding of strong relationships between AGB and crown dimensions is consistent with those reported by @Jucker2017.
While the interdependence of tree height and crown radius requires careful model refinement to address potential collinearity issues [@Dormann2013], their strong relationship with AGB suggests the feasibility of directly estimating AGB from EO data, bypassing DBH prediction and simplifying the estimation process [@Jucker2017].

# Conclusions

Our results underscore that saturating functions—particularly the Weibull model—offer greater accuracy for H–DBH relationships in both angiosperms and gymnosperms than do conventional power-law models, which can underestimate biomass in larger trees.
For CR–DBH allometry, different functional forms emerged as optimal for angiosperms (power-law) and gymnosperms (gMM), reflecting distinct growth strategies.
Although including wood density did not significantly improve predictive performance, it offered insights into interspecific trade-offs in tree architecture.
These findings are especially important for large-scale carbon assessments, where even small misjudgments in tree size can lead to major errors in biomass calculations.
Our results also highlight the importance of incorporating saturating models when forecasting ecosystem services in forests with a broad range of species and structural heterogeneity.
Future studies should investigate additional functional traits and environmental factors to improve both ecological understanding and carbon monitoring practices at global scales.

\newpage
# Acknowledgments {.unnumbered}

We thank the editors and two anonymous reviewers, whose comments greatly improved the manuscript.
We also thank the Alliance of International Science Organizations (ANSO) for providing financial support for TDN’s Master’s studies.

# Funding {.unnumbered}

MK was supported by
the Xingdian Talenj Support Program of Yunnan Province,
the Xishuangbanna State Rainforest Talent Support Program (E4BN041B01), and
the CAS President's International Fellowship Initiative (2020FYB0003).

# Data availability {.unnumbered}

Tallo and TRY data are available from their respective platforms.
Codes are available from Zenodo at xxx

\newpage
# References

### 2025-03-26 ms/ms.tex
Initial commit for tree-allo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{lineno}
\usepackage{setspace}
\doublespacing
\linenumbers
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Fig.}
\else
  \newcommand\figurename{Fig.}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

\textbf{Saturating allometric relationships reveal how wood density
shapes global tree architecture}

\[ \]

Thi Duyen Nguyen\textsuperscript{1,2}, Masatoshi
Katabuchi\textsuperscript{1}

\[ \]

\textsuperscript{1} CAS Key Laboratory of Tropical Forest Ecology,
Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences,
Menglun, Yunnan 666303, China

\textsuperscript{2} University of Chinese Academy of Sciences, Beijing
100049, China

\[ \]

\textbf{Corresponding Authors}:

Masatoshi Katabuchi

E-mail: katabuchi@xtbg.ac.cn; mattocci27@gmail.com

\[ \]

Thi Duyen Nguyen: \url{https://orcid.org/0009-0003-5296-9569}

Masatoshi Katabuchi: \url{https://orcid.org/0000-0001-9900-9029}

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

Allometric equations are fundamental tools in ecological research and
forestry management, widely used for estimating above-ground biomass and
production, serving as the core foundations of dynamic vegetation
models. Using global datasets from Tallo---a tree allometry and crown
architecture database encompassing thousands of species---and TRY---a
plant traits database---we fit Bayesian hierarchical models with three
alternative functional forms (power-law, generalized Michaelis--Menten
(gMM), and Weibull) to characterize how diameter at breast height (DBH),
tree height (H), and crown radius (CR) scale with and without wood
density as a species-level predictor. Our analysis revealed that the
saturating Weibull function best captured the relationship between tree
height and DBH in both functional groups, while the CR-DBH relationship
was best predicted by a power-law function in angiosperms and the gMM
function in gymnosperms. Although including wood density did not
significantly improve predictive performance, it revealed important
ecological trade-offs: lighter-wood angiosperms achieve taller mature
heights more rapidly, and denser wood promotes wider crown expansion
across clades. We also found that accurately estimating DBH required
considering both height and crown size, highlighting how these variables
together distinguish trees of similar height but differing trunk
diameters. Our results emphasize the importance of applying saturating
functions for large trees to improve forest biomass estimates and show
that wood density, while not always predictive at broad scales,
illuminates the biomechanical and ecological constraints underlying
diverse tree architectures. These findings offer practical pathways for
integrating height- and crown-based metrics into existing carbon
monitoring programs worldwide.

\textbf{Keywords:} above ground biomass, crown radius, diameter at
breast height, tree allometry model, tree height, wood density

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Mapping terrestrial carbon stocks is crucial to successfully
implementing climate change mitigation programs
(\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}). The accuracy of
carbon stock estimates depends on the availability of reliable
allometric models to deduce the aboveground biomass (AGB) of trees from
census data (\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}).
Selecting an appropriate allometric model is, therefore, a critical step
in estimating forest carbon biomass (e.g.,
\protect\hyperlink{ref-Chave2014}{Chave et al. 2014};
\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). Tree allometry
describes the correlations between different biometric measurements of
trees, such as diameter at breast height (DBH), tree height, and tree
crown size (e.g., \protect\hyperlink{ref-Shinozaki1964a}{Shinozaki et
al. 1964}; \protect\hyperlink{ref-Niklas1994}{Niklas 1994};
\protect\hyperlink{ref-West1999}{West et al. 1999}). These relationships
are widely utilized to estimate tree architecture from DBH and to
estimate forest biomass and production, serving as fundamental
components of dynamic vegetation models (e.g.,
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-Falster2017}{Falster et al. 2017};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}). Despite significant
progress, developing precise allometric models across regions
characterized by diverse tree species composition and varying forest
structures remains a persistent challenge. This complexity highlights
the need for improved global-scale allometric models that capture the
diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
remain the most common choice for describing the allometric scaling of
tree size variables, including height and crown size with DBH (e.g.,
\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
relationships among tree size variables often involve complex
interactions influenced by various factors, such as species diversity,
environmental conditions, and individual variability (e.g.,
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}). Consequently,
power-law models often fail to accurately capture the allometric
relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and overestimation for
larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
highlights the need to adopt alternative functional forms---such as
saturating relationships---that can more accurately reflect the
complexity of allometric scaling in vegetation models (e.g.,
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized the monitoring of
forests (e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017})
by enabling the identification and measurement of individual tree
heights and crown dimensions using airborne imagery (e.g.,
\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). However,
because existing allometric models still rely on stem diameter as a
crucial input for biomass estimation, this reliance presents a challenge
for effectively integrating EO technologies into AGB estimation
(\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
ground-based LiDAR can measure DBH directly and is much more efficient
than traditional ground surveys
(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
less common and convenient than aerial LiDAR. Alternative approaches
focusing on tree height and crown dimensions as model predictors are
also necessary for remote sensing applications
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
solution is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influence
individual tree architecture and the composition of entire forest
communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
closed-canopy conditions, strong light competition makes the balance
between vertical and horizontal space use crucial for species
coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
functional traits, wood density is closely linked to both plant life
history strategies (\protect\hyperlink{ref-Fajardo2022a}{Fajardo 2022})
and light demand (e.g., \protect\hyperlink{ref-King2006}{King et al.
2006}; \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a}). Several
studies suggest that wood density explains a trade-off between rapid
stem expansion and broader crowns, shaping mechanical constraints on
species with diverse architecture and light capture strategies
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
types of studies have examined this relationship post hoc, relating
species-specific allometric parameters to wood density and risking Type
I or Type II errors if the hierarchical data structure is overlooked
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
exception is Martínez Cano et al.
(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
density into a hierarchical model of tropical trees and concluded it
offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains
insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density (WD) into a
multilevel model, our methodology avoids the post-hoc correlations used
in previous studies, offering a more universal and scalable model,
effectively capturing variability from the species to the community
level. Our main objective is to develop models informed by tree
allometric variables (H, CR, DBH, WD) to capture interspecific variation
in allometric scaling globally and offer an improved framework for
estimating above ground biomass. Specifically, we ask the following
questions: (1) What is the best function to estimate tree architecture
from DBH and \emph{vice versa}? (2) How does wood density shape
interspecific variation in these allometric scaling relationships? (3)
How does the choice of allometric relationships affect subsequent AGB
estimation?

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, and crown radius for 5,163
species across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker
et al. 2022}, https://doi.org/10.5281/zenodo.6637599) along with wood
density data for 10,995 species from TRY version 6
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al.
(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
selected only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

To reduce computational complexity and ensure the robustness of the
models, we further filtered species with fewer than 20 individuals and
randomly sampled up to 100 individuals per species. This step was
necessary because the dataset is unbalanced; for example, \emph{Quercus
ilex} alone accounts for over 20,000 individuals, representing
approximately 10\% of the dataset. However, \emph{Quercus ilex} is not
necessarily the most abundant species globally. Its prevalence in this
dataset is a result of sampling and does not reflect its true global
abundance. By limiting the number of individuals per species, we
minimize bias due to data imbalance and prevent the overrepresentation
of any single abundant species
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}), thereby
ensuring that the model remains broadly applicable. Since not all trees
have a complete set of information, we created specific sub-datasets to
select the best predictive model for each allometric relationship for
two clades (i.e., angiosperms and gymnosperms; see below and
Supplementary Table S2 for more details).

\hypertarget{tree-allometry-models}{%
\subsection{Tree allometry models}\label{tree-allometry-models}}

We developed hierarchical Bayesian models to estimate the allometric
relationships between tree architectural variables, including DBH (cm),
tree height (H, m), crown radius (CR, m), and wood density (g
cm\textsuperscript{-3}). We assumed a Gaussian likelihood (\(N\)) for
the natural logarithm of the response variable, \(y_{(i,j)}\)
representing either tree height, crown radius, or DBH for each tree
\(i\) in species \(j\):

\begin{equation}\protect\hypertarget{eq-like}{}{
\log(y_{i,j}) \sim N(f(X_{i,j}, \pmb{\beta}_{j}), \sigma_y),
}\label{eq-like}\end{equation}

where \(f(\cdot)\) predicts the expected natural log tree height or
crown radius from observed DBH (and \emph{vice versa}),
\(\pmb{\beta}_{j}=\{a_j, b_j, k_j\}\) denotes the vector of
species-specific parameters, and \(\sigma_y\) represents the standard
deviation of the observation error. We considered three functional forms
for the relationship between tree allometric variables
(\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}):

Power-law: \begin{equation}\protect\hypertarget{eq-pl}{}{
y = aX^b,
}\label{eq-pl}\end{equation}

Generalized Michaelis-Menten (gMM):
\begin{equation}\protect\hypertarget{eq-gMM}{}{
y = \frac{aX^b}{k + X^b},
}\label{eq-gMM}\end{equation}

Weibull: \begin{equation}\protect\hypertarget{eq-wb}{}{
y = a\left\{1-\mathrm{exp}(-bX^k)\right\},
}\label{eq-wb}\end{equation}

The power-law model (Eq.~\ref{eq-pl}) is one of the simplest and most
commonly used functions to describe biological scaling relationships
(e.g., \protect\hyperlink{ref-West1999}{West et al. 1999};
\protect\hyperlink{ref-Brown2004}{Brown et al. 2004}). We also tested
two nonlinear models, the generalized Michaelis--Menten (gMM)
(Eq.~\ref{eq-gMM}) and the Weibull function (Eq.~\ref{eq-wb}), to
account for a saturating relationship between tree dimensions and DBH
(e.g., \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
2019}).

We applied the power-law model to estimate tree height and crown radius
from DBH, as well as DBH from tree height and/or crown radius.
Specifically, when predicting DBH from height and/or crown radius, we
tested four formulations: (1) using CR and H as separate factors,
\(\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \mathrm{H}^{c_1}\), denoted as
\(CR,H\); (2) using CR and H as a compound term,
\(\mathrm{DBH} = a_2 (\mathrm{CR} \times \mathrm{H})^{b_2}\), denoted as
\(CR \times H\); (3) using only height,
\(\mathrm{DBH} = a_3 \mathrm{H}^{b_3}\); and (4) using only crown
radius, \(\mathrm{DBH} = a_4 \mathrm{CR}^{b_4}\). In contrast, the gMM
and Weibull models were only used to estimate tree height and crown
radius from DBH. Because these two models specifically describe
saturating relationships (Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}), their
formulation does not allow simply switching the roles of explanatory and
response variables.

To compare the performance of different models for the same response
variable, we used Pareto-smoothed importance sampling leave-one-out
cross-validation (PSIS-LOO; Vehtari et al.
(\protect\hyperlink{ref-Vehtari2017}{2017})). PSIS-LOO is an accurate
and reliable approximation to standard leave-one-out cross-validation
(LOO), which is a robust method for comparing models with different
numbers of parameters (\protect\hyperlink{ref-Vehtari2017}{Vehtari et
al. 2017}). We utilized PSIS-LOO to calculate the LOO Information
Criterion (LOOIC) for each model. A model with lower LOOIC is considered
to have better predictive accuracy
(\protect\hyperlink{ref-Vehtari2017}{Vehtari et al. 2017}).

Once the best-performing models for each allometric relationship were
identified, we tested whether adding wood density as a species-level
predictor would further improve the predictive performance of those
models and also examined its influence on tree architecture diversity.
Species-specific parameters (\(\beta_j\)) were modeled as a function of
wood density (\(u_j\)): \begin{equation}\protect\hypertarget{eq-wood}{}{
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
}\label{eq-wood}\end{equation}

where \(\gamma_{0}\) and \(\gamma_{1}\) are the species-level intercept
and slope, respectively, and \(\sigma_{\beta}\) is the standard
deviation of the species-specific parameters. To facilitate
interpretation and visualization, species were then classified into
``light'' and ``dense'' wood groups based on the 10th and 90th
percentiles of wood density distribution in the observed dataset,
respectively. This classification allowed us to illustrate the
relationships between wood density and the parameters of the
best-performing models, enabling an evaluation of how wood density
contributes to variation in tree architecture. Additionally, we explored
a model that accounts for covariance between species-specific parameters
using a multivariate normal distribution. However, due to the complexity
of this approach, the model did not converge for the nonlinear models,
and thus, we did not include it in the final analysis.

Posterior distributions of all parameters were estimated using the
No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte
Carlo (HMC), algorithm implemented in Stan
(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}) using
weakly-informative priors (\protect\hyperlink{ref-Gelman2008}{Gelman et
al. 2008}). The HMC algorithm uses gradient information to propose new
states in the Markov chain, leading to a more efficient exploration of
the target distribution than traditional Markov chain Monte Carlo (MCMC)
methods that rely on random proposals
(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}). This
efficiency allows us to achieve convergence with fewer iterations than
traditional MCMC methods. Inferences were based on 1,000 posterior
samples following 1,000 warmup iterations for three parallel chains.
Convergence was assessed using the Gelman--Rubin statistic with a
convergence threshold of 1.1 (\protect\hyperlink{ref-Gelman2013}{Gelman
et al. 2015}) and ensuring effective sample sizes greater than 300
(\protect\hyperlink{ref-Vehtari2021}{Vehtari et al. 2021}). We also
reported the median and 95\% Bayesian credible intervals (CIs) for each
parameter (Supplementary Tables S3-S8). Estimates were considered
significant when 95\% CIs did not overlap with zero.

Statistical analyses were performed in R version 4.3.2
(\protect\hyperlink{ref-RCoreTeam2023}{R Core Team 2023}), using the R
packages \emph{targets} version 1.7.0 for workflow management
(\protect\hyperlink{ref-Landau2021}{Landau 2021}). Codes are made
available on GitHub at
\url{https://github.com/duyenecology/tree-allometry} (Stan code is
provided in Supplementary Section S2).

\hypertarget{aboveground-biomass-estimates}{%
\subsection{Aboveground biomass
estimates}\label{aboveground-biomass-estimates}}

For angiosperm species, we used the widely applied equation from Chave
et al. (\protect\hyperlink{ref-Chave2014}{2014}) to estimate aboveground
biomass (AGB; kg of dry mass):

\begin{equation}\protect\hypertarget{eq-agb}{}{
\mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
}\label{eq-agb}\end{equation}

where \(\rho\) is wood density (g cm\textsuperscript{-3}), \(D\) (cm) is
DBH, and \(H\) (m) is tree height. We either used both observed DBH and
height or substituted one of these variables with an estimate from our
allometric models, allowing us to assess how different allometric
relationships affect AGB calculations. Only trees with a DBH ≥ 5 cm were
used, as those below this threshold hold a small fraction of forest
carbon stocks and were excluded from the calibration of Chave et al.
(\protect\hyperlink{ref-Chave2014}{2014})'s equation. Additionally, the
sub-datasets used to select the optimal predictive model for each
allometric relationship differed as not all trees had a complete
information set (see above and Supplementary Table S2). Consequently,
the final dataset for estimating AGB in angiosperms includes 124,206
trees with sufficient allometric information (i.e., DBH, H, CR, WD)
across 624 species. We did not extend this approach to gymnosperms
because no general AGB equation is available for them, and most existing
equations are species-specific and do not incorporate wood density
(e.g., \protect\hyperlink{ref-Moore2010}{Moore 2010};
\protect\hyperlink{ref-Bazrgar2024}{Bazrgar et al. 2024}).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{tree-height-allometry}{%
\subsection{Tree height allometry}\label{tree-height-allometry}}

The optimal model to predict tree height was a Weibull function for both
angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
dataset, the best model to predict H from DBH of each clade was:
\begin{equation}\protect\hypertarget{eq-h_ang}{}{
\mathrm{H_{ang}} =50.0
\left\{ 1 - \exp\left( - 0.0476
\mathrm{DBH}^{0.629}
\right) \right\}
}\label{eq-h_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-h_gym}{}{
\mathrm{H_{gym}} = 117.4
\left\{ 1 - \exp\left( - 0.012
\mathrm{DBH}^{0.750}
\right) \right\}.
}\label{eq-h_gym}\end{equation}

The posterior medians with their 95\% CIs were as follows: asymptote
\(a_{\text{ang}}\) = 50.0 {[}48.0, 52.2{]}, \(a_{\text{gym}}\) = 117.4
{[}89.22, 175.1{]}; exponent \(b_{\text{ang}}\) = 0.0476 {[}0.0458,
0.0495{]}, \(b_{\text{gym}}\) = 0.012 {[}0.008, 0.016{]}; and shape
parameter \(k_{\text{ang}}\) = 0.629 {[}0.618, 0.639{]},
\(k_{\text{gym}}\) = 0.750 {[}0.719, 0.784{]}.

The second-best model, which used the gMM function, showed poorer
predictive performance in angiosperms (\(\Delta\)LOOIC = 101.01) but
showed competitive predictive performance in gymnosperms
(\(\Delta\)LOOIC = 4.87). While power-law models (Eq.~\ref{eq-pl}) can
exhibit saturating behavior when 0 \textless{} \emph{b} \textless{} 1,
they performed the worst overall compared to other saturating models
(Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}). Incorporating wood density did not
significantly improve predictive performance (Table 1).

\textbf{Table 1}: Summary of the results of the model selection
procedure. Models were ranked using the Leave-One-Out Information
Criterion (LOOIC) to assess their predictive performance. For each
model, we report the Expected Log Predictive Density (ELPD), the LOOIC,
and the \(R^2\) values. The difference in LOOIC relative to the
best-performing model, \(\Delta\)LOOIC, is also included, with a
\(\Delta\)LOOIC greater than 4 suggests a meaningful difference in
predictive performance (\protect\hyperlink{ref-Cochrane2023}{Cochrane et
al. 2023}; \protect\hyperlink{ref-Sivula2023}{Sivula et al. 2023}).
Lower \(\Delta\)LOOIC values indicate models that perform closer to the
best model in terms of predictive accuracy, while higher ELPD values
reflect better model fit. Higher \(R^2\) values reflect better
explanatory power, indicating how well the model explains the variance
in the dependent variable.

\newpage

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{ccccccccc}
\toprule
Dependent variable & Predictor variable & Division & Functional form & Wood density & ELPD & LOOIC & ΔLOOIC & R2\\
\midrule
Tree Height & DBH & Angiosperm & Weibull & With & -4317.73 & 8635.46 & 0.00 & 0.823\\
Tree Height & DBH & Angiosperm & Weibull & Without & -4318.55 & 8637.10 & 1.65 & 0.823\\
Tree Height & DBH & Angiosperm & gMM & Without & -4368.23 & 8736.47 & 101.01 & 0.825\\
Tree Height & DBH & Angiosperm & Power-Law & Without & -4720.21 & 9440.42 & 804.96 & 0.813\\
Tree Height & DBH & Gymnosperm & Weibull & Without & -540.02 & 1080.04 & 0.00 & 0.859\\
\addlinespace
Tree Height & DBH & Gymnosperm & Weibull & With & -540.37 & 1080.74 & 0.71 & 0.859\\
Tree Height & DBH & Gymnosperm & gMM & Without & -542.45 & 1084.90 & 4.87 & 0.858\\
Tree Height & DBH & Gymnosperm & Power-Law & Without & -581.04 & 1162.08 & 82.05 & 0.854\\
Crown Radius & DBH & Angiosperm & Power-Law & Without & -20264.58 & 40529.16 & 0.00 & 0.724\\
Crown Radius & DBH & Angiosperm & Power-Law & With & -20265.11 & 40530.22 & 1.07 & 0.724\\
\addlinespace
Crown Radius & DBH & Angiosperm & Weibull & Without & -20326.86 & 40653.73 & 124.57 & 0.722\\
Crown Radius & DBH & Angiosperm & gMM & Without & -20327.75 & 40655.50 & 126.35 & 0.722\\
Crown Radius & DBH & Gymnosperm & gMM & Without & -811.34 & 1622.68 & 0.00 & 0.728\\
Crown Radius & DBH & Gymnosperm & gMM & With & -813.38 & 1626.77 & 4.09 & 0.728\\
Crown Radius & DBH & Gymnosperm & Weibull & Without & -815.13 & 1630.26 & 7.58 & 0.728\\
\addlinespace
Crown Radius & DBH & Gymnosperm & Power-Law & Without & -892.70 & 1785.40 & 162.72 & 0.712\\
DBH & CR, H & Angiosperm & Power-Law & With & -15613.32 & 31226.64 & 0.00 & 0.872\\
DBH & CR, H & Angiosperm & Power-Law & Without & -15615.02 & 31230.05 & 3.41 & 0.803\\
DBH & CR × H & Angiosperm & Power-Law & Without & -17144.13 & 34288.25 & 3061.61 & 0.855\\
DBH & H & Angiosperm & Power-Law & Without & -21457.80 & 42915.59 & 11688.96 & 0.825\\
\addlinespace
DBH & CR & Angiosperm & Power-Law & Without & -25578.55 & 51157.11 & 19930.47 & 0.728\\
DBH & CR, H & Gymnosperm & Power-Law & Without & -961.30 & 1922.59 & 0.00 & 0.854\\
DBH & CR, H & Gymnosperm & Power-Law & With & -961.55 & 1923.10 & 0.51 & 0.897\\
DBH & CR × H & Gymnosperm & Power-Law & Without & -1136.13 & 2272.27 & 349.67 & 0.885\\
DBH & H & Gymnosperm & Power-Law & Without & -2000.82 & 4001.63 & 2079.04 & 0.809\\
\addlinespace
DBH & CR & Gymnosperm & Power-Law & Without & -2392.47 & 4784.95 & 2862.36 & 0.752\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\begin{figure}

{\centering \includegraphics{../figs/h_cr_dbh_non_log.png}

}

\caption{\label{fig-compare}Allometric relationships between DBH and
tree height or crown radius in angiosperms (a, c) and gymnosperms (b,
d), based on three different models: power-law (orange), generalized
Michaelis-Menten (green), and Weibull (blue). A color gradient from
black (low density) to yellow (high density) indicates the density of
overlapping data points, calculated via 2D kernel density estimation on
a 300\,\(\times\)\,300 grid. See Supplementary Figure S1 for the
log-scale version of these plots.}

\end{figure}

\hypertarget{crown-radius-allometry}{%
\subsection{Crown Radius allometry}\label{crown-radius-allometry}}

The power-law function (0 \textless{} \emph{b} \textless{} 1) provided
the best predictive performance for modeling the relationship between
crown radius and DBH in angiosperms. Whereas for gymnosperms, the gMM
function offered the best performance for this allometric scaling (Table
1). Across the global dataset, the best model for predicting CR from DBH
of each clade was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
\mathrm{CR_{ang}} = \mathrm{0.349}
\, \mathrm{DBH}^{\mathrm{0.626}}
}\label{eq-cr_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-cr_gym}{}{
\mathrm{CR_{gym}} = \frac{\mathrm{769.4}
\, \mathrm{DBH}^{\mathrm{0.631}}}
{\mathrm{2657}
+ \mathrm{DBH}^{\mathrm{0.631}}}.
}\label{eq-cr_gym}\end{equation}

The posterior medians with their 95\% CIs were as follows: scaling
factor \(a_{\text{ang}}\) = 0.349 {[}0.340, 0.359{]}, asymptote
\(a_{\text{gym}}\) = 769.4 {[}227.3, 1818{]}; exponent
\(b_{\text{ang}}\) = 0.626 {[}0.617, 0.635{]}, \(b_{\text{gym}}\) =
0.631 {[}0.600, 0.663{]}; and scaling constant \(k_{\text{gym}}\) = 2657
{[}808.7, 6274{]}.

For angiosperms, the saturating functions (Eq.~\ref{eq-gMM};
Eq.~\ref{eq-wb}) performed worse than the power-law model, with
\(\Delta\)LOOIC values of 124.57 for gMM and 126.35 for the Weibull
function. In gymnosperms, the Weibull function showed a competitive
performance compared to the gMM function, while the power-law function
(Eq.~\ref{eq-pl}) performed the worst, with a \(\Delta\)LOOIC of 162.72.
Incorporating wood density did not significantly improve predictive
performance (Table 1).

\textbf{Table 2}: Posterior estimates of the parameters for the best
hierarchical models selected for tree height, crown radius, and DBH
allometries in angiosperms and gymnosperms. Each set includes posterior
estimates both without and with the inclusion of wood density, reported
as median and 95\% Bayesian credible intervals for the community-level
parameters of each allometric function. For tree height, the Weibull
model was identified as the best for both clades. Crown radius was best
described by a power-law model in angiosperms and a gMM model for
gymnosperms. The best model for DBH allometry in both clades, which
incorporates both tree height and crown radius information (\(CR,H\)
model).

\newpage

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{cccccccc}
\toprule
Dependent\_variable & Division & Functional\_form & Wood\_density & Parameter & Intercept\_CI & Slope\_CI & Tau\_CI\\
\midrule
Tree Height & Angiosperm & Weibull & With & a & 50.0 (48.0, 52.3) & 1.00 (0.966, 1.04) & 0.109 (0.0850, 0.134)\\
Tree Height & Angiosperm & Weibull & With & b & 0.0477 (0.0456, 0.0495) & -0.000250 (-0.00189, 0.00142) & 0.0118 (0.0109, 0.0126)\\
Tree Height & Angiosperm & Weibull & With & k & 0.628 (0.618, 0.638) & -0.0155 (-0.0254, -0.00591) & 0.141 (0.134, 0.149)\\
Tree Height & Angiosperm & Weibull & Without & a & 50.0 (48.0, 52.2) & - & 0.109 (0.0847, 0.133)\\
Tree Height & Angiosperm & Weibull & Without & b & 0.0476 (0.0458, 0.0495) & - & 0.0118 (0.0110, 0.0126)\\
\addlinespace
Tree Height & Angiosperm & Weibull & Without & k & 0.629 (0.618, 0.639) & - & 0.142 (0.135, 0.149)\\
Tree Height & Gymnosperm & Weibull & With & a & 114 (85.1, 170) & 0.974 (0.776, 1.21) & 0.0976 (0.00466, 0.257)\\
Tree Height & Gymnosperm & Weibull & With & b & 0.0123 (0.00836, 0.0163) & 0.000633 (-0.00188, 0.00342) & 0.00382 (0.00225, 0.00553)\\
Tree Height & Gymnosperm & Weibull & With & k & 0.751 (0.721, 0.784) & -0.0224 (-0.0538, 0.00762) & 0.113 (0.0905, 0.140)\\
Tree Height & Gymnosperm & Weibull & Without & a & 117 (89.2, 175) & - & 0.0912 (0.00357, 0.229)\\
\addlinespace
Tree Height & Gymnosperm & Weibull & Without & b & 0.0119 (0.00806, 0.0155) & - & 0.00371 (0.00230, 0.00525)\\
Tree Height & Gymnosperm & Weibull & Without & k & 0.750 (0.719, 0.784) & - & 0.114 (0.0919, 0.139)\\
Crown Radius & Angiosperm & Power-Law & With & a & 0.349 (0.340, 0.359) & 0.0680 (0.0395, 0.0970) & 0.331 (0.311, 0.354)\\
Crown Radius & Angiosperm & Power-Law & With & b & 0.626 (0.617, 0.635) & -0.0110 (-0.0199, -0.00253) & 0.0921 (0.0844, 0.100)\\
Crown Radius & Angiosperm & Power-Law & Without & a & 0.349 (0.340, 0.359) & - & 0.336 (0.316, 0.358)\\
\addlinespace
Crown Radius & Angiosperm & Power-Law & Without & b & 0.626 (0.617, 0.635) & - & 0.0929 (0.0852, 0.101)\\
Crown Radius & Gymnosperm & gMM & With & a & 975 (312, 2010) & 0.988 (0.766, 1.27) & 0.367 (0.280, 0.474)\\
Crown Radius & Gymnosperm & gMM & With & b & 0.629 (0.599, 0.659) & 0.0345 (0.00507, 0.0642) & 0.0949 (0.0691, 0.126)\\
Crown Radius & Gymnosperm & gMM & With & k & 3371 (1094, 6978) & 35.7 (-769, 914) & 9.36 (0.100, 52.3)\\
Crown Radius & Gymnosperm & gMM & Without & a & 769 (227, 1818) & - & 0.367 (0.280, 0.465)\\
\addlinespace
Crown Radius & Gymnosperm & gMM & Without & b & 0.631 (0.600, 0.663) & - & 0.101 (0.0755, 0.131)\\
Crown Radius & Gymnosperm & gMM & Without & k & 2657 (809, 6274) & - & 19.0 (0.112, 200)\\
DBH & Angiosperm & Power-Law & With & a & 0.949 (0.928, 0.970) & -0.00254 (-0.0265, 0.0190) & 0.301 (0.284, 0.320)\\
DBH & Angiosperm & Power-Law & With & b & 0.373 (0.361, 0.386) & -0.00327 (-0.0146, 0.00891) & 0.135 (0.125, 0.145)\\
DBH & Angiosperm & Power-Law & With & c & 0.589 (0.575, 0.605) & 0.00511 (-0.00957, 0.0200) & 0.168 (0.156, 0.179)\\
\addlinespace
DBH & Angiosperm & Power-Law & Without & a & 0.948 (0.927, 0.969) & - & 0.302 (0.284, 0.320)\\
DBH & Angiosperm & Power-Law & Without & b & 0.374 (0.361, 0.386) & - & 0.135 (0.125, 0.145)\\
DBH & Angiosperm & Power-Law & Without & c & 0.590 (0.575, 0.605) & - & 0.168 (0.157, 0.180)\\
DBH & Gymnosperm & Power-Law & With & a & 0.971 (0.900, 1.04) & -0.0451 (-0.120, 0.0295) & 0.274 (0.225, 0.339)\\
DBH & Gymnosperm & Power-Law & With & b & 0.408 (0.371, 0.444) & 0.0284 (-0.00773, 0.0616) & 0.112 (0.0845, 0.148)\\
\addlinespace
DBH & Gymnosperm & Power-Law & With & c & 0.588 (0.538, 0.637) & -0.0364 (-0.0841, 0.0140) & 0.165 (0.128, 0.209)\\
DBH & Gymnosperm & Power-Law & Without & a & 0.972 (0.903, 1.05) & - & 0.273 (0.222, 0.336)\\
DBH & Gymnosperm & Power-Law & Without & b & 0.410 (0.373, 0.446) & - & 0.115 (0.0861, 0.150)\\
DBH & Gymnosperm & Power-Law & Without & c & 0.588 (0.540, 0.638) & - & 0.165 (0.129, 0.209)\\
\bottomrule
\end{longtable*}
\endgroup{}

\hypertarget{dbh-allometry}{%
\subsection{DBH allometry}\label{dbh-allometry}}

The \(CR,H\) model, which treated CR and H as separate factors,
exhibited the best predictive performance among all tested models (Table
1). Across the global dataset, the most effective model for predicting
DBH from H and CR for each clade was:
\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
\mathrm{DBH_{ang}} = \mathrm{0.948}
\mathrm{CR}^{\text{0.374}}
\mathrm{H}^{\text{0.590}}
}\label{eq-dbh_ang}\end{equation}

\begin{equation}\protect\hypertarget{eq-dbh_gym}{}{
\mathrm{DBH_{gym}} = \mathrm{0.972}
\mathrm{CR}^{\text{0.410}}
\mathrm{H}^{\text{0.588}}.
}\label{eq-dbh_gym}\end{equation}

The posterior medians with their 95\% CIs were as follows: scale
\(a_{\text{ang}}\) = 0.948 {[}0.927, 0.969{]}, \(a_{\text{gym}}\) =
0.972 {[}0.903, 1.051{]}; exponent for CR \(b_{\text{ang}}\) = 0.374
{[}0.361, 0.386{]}, \(b_{\text{gym}}\) = 0.410 {[}0.373, 0.446{]}; and
exponent for H \(c_{\text{ang}}\) = 0.590 {[}0.575, 0.605{]},
\(c_{\text{gym}}\) = 0.588 {[}0.540, 0.638{]}. Incorporating wood
density did not significantly improve predictive performance (Table 1).

\hypertarget{wood-density-and-tree-architectural-diversity}{%
\subsection{Wood density and tree architectural
diversity}\label{wood-density-and-tree-architectural-diversity}}

Although including wood density did not significantly improve the
predictive performance of any of the allometric relationships, it can
enhance our understanding of the underlying biological mechanisms
driving tree architectural diversity. In H-DBH allometry, the shape
parameter \(k\) in angiosperms showed a significant negative
relationship with wood density, while the asymptote parameter \(a\) and
the scale parameter \(b\) showed no significant associations
(Fig.~\ref{fig-wood1} a, b, c; Fig.~\ref{fig-wood2} a). In gymnosperms,
none of the parameters significantly correlated with wood density
(Fig.~\ref{fig-wood1} d, e, f).

In CR-DBH allometry, angiosperms exhibited a positive relationship
between the scaling factor \(a\) and wood density and a negative
association between the exponent parameter \(b\) and wood density
(Fig.~\ref{fig-wood1} g, h; Fig.~\ref{fig-wood2} b). In gymnosperms, we
found a positive association between wood density and the exponent
parameter \(b\), whereas neither the asymptote parameter \(a\) nor the
scaling constant \(k\) showed significant relationship with this
functional trait (Fig.~\ref{fig-wood1} i, j, k; Fig.~\ref{fig-wood2} c).

\begin{figure}[H]

{\centering \includegraphics{../figs/wd_com.png}

}

\caption{\label{fig-wood1}Relationships between wood density and
species-specific allometry parameters for tree height (a, b, c, d, e, f)
and crown radius (g, h, i, j, k) of angiosperms (purple) and gymnosperms
(green). Points represent posterior medians of species-specific
parameters, with vertical bars indicating 95\% Bayesian credible
intervals (CIs). Thick solid lines represent significant relationships
across species, with the shaded area illustrating the 95\% CIs, while
dashed lines indicate non-significant relationships.}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{../figs/wd_ef.png}

}

\caption{\label{fig-wood2}Effect of wood density on tree height
allometry in angiosperms (a) and on crown radius allometry in
angiosperms (b) and gymnosperms (c). Curves represent the allometric
scaling for species with light (orange) and dense (brown) wood density,
corresponding to the 10\% and 90\% quantiles of wood density
distribution in the observed dataset, respectively. The shaded areas
illustrate the 95\% Bayesian credible intervals. Only statistically
significant associations from Fig.~\ref{fig-wood1} are plotted.}

\end{figure}

\hypertarget{estimating-aboveground-biomass}{%
\subsection{Estimating aboveground
biomass}\label{estimating-aboveground-biomass}}

AGB estimates based on the Weibull model (\(\text{AGB}_\text{H-WB}\))
showed the highest predictive accuracy based both on the lowest root
mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.04\%
(Fig.~\ref{fig-agb} a). Estimates based on the power-law model
(\(\text{AGB}_\text{H-PL}\)) also showed competitive performance but had
higher RMSE and bias compared to the Weibull-based approach, with RMSE
of 0.28 Mg and bias of 2.29\% (Fig.~\ref{fig-agb} b). In contrast,
models that used estimated DBH performed substantially worse
(Fig.~\ref{fig-agb} c-f).

\begin{figure}[H]

{\centering \includegraphics{../figs/agb_ang.png}

}

\caption{\label{fig-agb}Comparison of individual tree above-ground
biomass (AGB) estimated using different approaches with reference AGB
(\(\text{AGB}_\text{ref}\)) derived from observed DBH and tree height in
angiosperms. (a, b) AGB estimates based on observed DBH and tree height
predicted by the Weibull function (\(\text{AGB}_\text{H-WB}\)) or the
power-law function (\(\text{AGB}_\text{H-PL}\)). (c--f) AGB estimates
based on observed tree height and DBH predicted from crown radius and
tree height as separate factors
(\(\text{AGB}_{\text{DBH} \sim \text{CR,H}}\)), as a compound term
(\(\text{AGB}_{\text{DBH} \sim \text{CR × H}}\)), from tree height alone
(\(\text{AGB}_{\text{DBH} \sim \text{H}}\)), or from crown radius alone
(\(\text{AGB}_{\text{DBH} \sim \text{CR}}\)). Dashed lines represent the
1:1 relationships, and the solid lines show linear regression fits to
the data, highlighting how predictive accuracy varies with tree size.
RMSE (root mean square error) and bias values for each model are
reported.}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Relationships among tree-size variables are foundational to both
ecological theory and practical forestry applications, informing
everything from carbon stock estimation to dynamic vegetation modeling.
In this study, we tested multiple functional forms---power-law,
generalized Michaelis--Menten (gMM), and Weibull---within a Bayesian
hierarchical framework to identify which best describes the linkages
between diameter at breast height (DBH), tree height (H), and crown
radius (CR) across a global dataset. Our results reveal that the Weibull
function most reliably captures the saturating nature of the H--DBH
relationship in both angiosperms and gymnosperms, while the CR--DBH
relationship diverges by clade, being well described by a power-law in
angiosperms but requiring a gMM form in gymnosperms. Although
incorporating wood density did not significantly enhance predictive
performance, it illuminated interspecific architectural constraints and
trade-offs, underscoring the complex interplay between mechanical
support, crown expansion, and life history strategies. Below, we discuss
these findings in the context of tree allometry models, the role of
functional traits, and forest biomass estimation.

\hypertarget{tree-allometry}{%
\subsection{Tree allometry}\label{tree-allometry}}

Our result supported a saturating H-DBH relationship in both angiosperm
and gymnosperm species, consistent with theoretical expectations
(\protect\hyperlink{ref-Falster2003}{Falster and Westoby 2003};
\protect\hyperlink{ref-Niklas2007}{Niklas 2007}) and empirical
observations (e.g., \protect\hyperlink{ref-Feldpausch2011}{Feldpausch et
al. 2011}; \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
2019}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). A recent
study using Tallo suggested the Weibull function as the best model for
six plant functional types (PFTs) (\protect\hyperlink{ref-Song2024}{Song
et al. 2024}). While PFTs-based approaches can illuminate large-scale
patterns, they may obscure species-specific variation in allometric
scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns
(Supplementary Table S6) despite underestimating the height of
individuals with DBH \textgreater{} 30 cm due to substantial
interspecific variation in H--DBH allometric scaling
(Fig.~\ref{fig-compare}). This highlights the advantages of a multilevel
modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic
scaling relationships among different axes of tree size, with an
exponent close to 2/3 (\(H \propto DBH^{2/3}\),
\(CR \propto DBH^{2/3}\)), based on assumptions regarding mechanical and
hydraulic constraints to plant growth
(\protect\hyperlink{ref-West1999}{West et al. 1999}). However, our
findings revealed that the power-law model exponent of angiosperms
(\emph{b} = 0.538) was significantly smaller, while the actual exponent
in gymnosperms (\emph{b} = 0.704) was larger than the theoretical value
of 2/3 (Supplementary Table S3). This divergence highlights the
importance of separately modeling the allometric relationships of each
functional group to accurately capture their distinct scaling patterns.
Moreover, the power-law function performed worse than the saturating
functions (Table 1), indicating that the actual H-DBH relationship does
not follow the simple scaling laws proposed by MST.

Our analysis also revealed divergent patterns in CR-DBH allometry,
highlighting fundamental differences in growth strategies and ecological
adaptations between angiosperms and gymnosperms. In angiosperms, the
CR--DBH relationship follows a power-law form, suggesting a capacity for
plastic crown expansion in competitive environments
(\protect\hyperlink{ref-Lines2012}{Lines et al. 2012}; e.g.,
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}), a conclusion
consistent with a study in tropical moist forests by Martínez Cano et
al. (\protect\hyperlink{ref-MartinezCano2019}{2019}). In addition, the
estimated exponent aligns closely with MST predictions for crown radius,
supporting the selection of the power-law model for angiosperms, where
the community-level exponent \emph{b} is near the expected 2/3
(\(b_{\text{ang}}\) = 0.626 {[}0.617, 0.635{]}, Table 2). In contrast,
the CR-DBH allometry in gymnosperms showed an asymptotic relationship
best described by the gMM function, with the Weibull function also
showing competitive predictive performance (Table 1). This asymptotic
pattern likely arises from structural constraints in gymnosperms growth
forms that favour vertical over lateral expansion (e.g.,
\protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), an adaptation
that can mitigate risks such as branch breakage under heavy snow loads
(e.g., \protect\hyperlink{ref-Mikola1938}{Mikola 1938};
\protect\hyperlink{ref-Loehle2016}{Loehle 2016};
\protect\hyperlink{ref-Tavankar2019}{Tavankar et al. 2019}).

Finally, we tested models to estimate DBH from biometric variables
(i.e., tree height and crown radius) measurable via Earth observation
(EO) platforms, such as remote sensing and LiDAR
(\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Our analysis
showed that accurately predicting DBH requires both tree height and
crown radius (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}),
which is especially important for distinguishing among trees of similar
height but varying DBH (\protect\hyperlink{ref-King2005}{King 2005};
\protect\hyperlink{ref-King2011}{King and Clark 2011}). Among the tested
models, the \(CR,H\) model
(\(\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}\))
provided the highest predictive accuracy, a result that contrasts with
the previous study (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
2017}), which concluded that the \(CR \times H\) model
(\(\mathrm{DBH} = a_2(\mathrm{CR} \times \mathrm{H})^{b_2}\)) was the
best for predicting DBH. The \(CR,H\) model significantly improves
predictive accuracy by accounting for the distinct contributions of
crown radius and tree height as independent factors in DBH estimation.
The implications of these differences for biomass estimation are
explored in detail in a later section, where we assess how DBH
prediction accuracy affects allometric biomass calculations.

\hypertarget{wood-density-and-tree-architecture-diversity}{%
\subsection{Wood density and tree architecture
diversity}\label{wood-density-and-tree-architecture-diversity}}

For the H-DBH relationship, wood density does not strongly affect the
initial growth rate of angiosperm trees but plays a critical role in
determining maximum height and the growth trajectory to achieve that
height, with lighter wood species tending to grow taller and attain
their mature height more rapidly than those with denser wood
(Fig.~\ref{fig-wood1} a-c; Fig.~\ref{fig-wood2} a). In contrast,
gymnosperms exhibit a height growth strategy that is largely independent
of wood density (Fig.~\ref{fig-wood1} d-f). These results are consistent
with the previous study (\protect\hyperlink{ref-Poorter2012b}{Poorter et
al. 2012b}), which showed that among seven angiosperm species, those
with lighter wood exhibited taller maximum heights, while no significant
correlations were observed for six gymnosperm species. A recent study
using the Tallo database alongside two additional datasets reported a
nonsignificant relationship between tree height and wood density
(\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). Their analysis
included functional groups (angiosperms vs.~gymnosperms) as a fixed
effect, providing valuable insights into group-level patterns. However,
it did not incorporate interactions between wood density and functional
groups, and thus, the effects of wood density were modeled as uniform
across angiosperms and gymnosperms. In contrast, our approach explicitly
modeled these differences, enabling us to capture functional
group-specific relationships and provide a more nuanced understanding of
height allometry.

In the CR-DBH allometry, our results indicate that both angiosperms and
gymnosperms with denser wood tend to have wider crowns across all trunk
sizes (Fig.~\ref{fig-wood1} g-k). Although the effect of wood density on
CR-DBH allometry is significant, the crown expansion patterns do not
clearly separate light-wood from dense-wood species
(Fig.~\ref{fig-wood2} b, c). This may suggest that crown radius is more
plastic and responsive to local environmental conditions (e.g., light
availability, stand density) than tree height, potentially obscuring the
relationship between wood density and crown radius. Nevertheless, our
findings are consistent with previous work in temperate forests
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009}), tropical
forests (\protect\hyperlink{ref-Iida2012}{Iida et al. 2012}),
subtropical forests (\protect\hyperlink{ref-Yang2023}{Yang and Swenson
2023}), and on a global scale (\protect\hyperlink{ref-Jucker2024}{Jucker
et al. 2024}), all of which similarly report that denser wood promotes
more efficient horizontal crown expansion.

In the analysis of DBH allometry, no significant correlation was
observed between wood density and DBH in either angiosperms or
gymnosperms (Table 2). This outcome aligns with observations from
saplings in an Australian subtropical forest
(\protect\hyperlink{ref-Kooyman2009}{Kooyman and Westoby 2009}) but
contradicts the expectation that species with lower wood density develop
thicker stems to ensure structural stability, as suggested by prior
studies (\protect\hyperlink{ref-King2006}{King et al. 2006};
\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Although wood
density is often regarded as a key functional trait (e.g.,
\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), variations in
tree allometry across individuals are likely influenced by a combination
of species identity and environmental factors. Across species or along
broad environmental gradients, wood density systematically
varies---tending to be denser in drier, colder climates and in
nutrient-poor soils
(\protect\hyperlink{ref-Muller-Landau2004}{Muller-Landau 2004};
\protect\hyperlink{ref-Chave2009}{Chave et al. 2009};
\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022};
\protect\hyperlink{ref-Mo2024}{Mo et al. 2024})---reflecting species'
adaptive strategies. Such environmental influences might confound simple
correlations between wood density and DBH when examining diverse tree
communities. At the intraspecific level, recent findings suggest that
wood density may be more closely related to vessel anatomy than to tree
architecture across different environments
(\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022}).
Furthermore, substantial measurement errors in tree height and crown
radius relative to DBH may introduce noise that obscures potential
relationships between wood density and stem diameter. These highlight
the importance of adopting a more integrative framework that accounts
for additional functional traits and environmental variables to deepen
our understanding of the factors driving DBH growth.

Our analysis indicated distinct patterns in which wood density has a
greater influence on the interspecific variation in angiosperms than
gymnosperms. Variation in wood density has consequential implications
for the mechanical properties of wood and hence influences the
mechanical stability of stems and branches
(\protect\hyperlink{ref-Sterck1998}{Sterck and Bongers 1998}),
ultimately determining the architecture of trees (e.g.,
\protect\hyperlink{ref-Poorter2006}{Poorter et al. 2006}). Angiosperms
have denser wood and a wider wood density range than gymnosperms (see
Supplementary Table S2), resulting in greater and more flexible
mechanical support (e.g., \protect\hyperlink{ref-Gelder2006}{Gelder et
al. 2006}; \protect\hyperlink{ref-Swenson2007}{Swenson and Enquist
2007}; \protect\hyperlink{ref-Mo2024}{Mo et al. 2024}). This allows them
to develop a more diverse height and crown to optimize light capture,
particularly in environments where light competition is intense, such as
tropical regions. In contrast, gymnosperms predominantly live and often
dominate in extreme environments such as boreal and high-elevation
regions (\protect\hyperlink{ref-Swenson2007}{Swenson and Enquist 2007};
\protect\hyperlink{ref-Mo2024}{Mo et al. 2024}), prioritising survival
and longevity (\protect\hyperlink{ref-Fossdal2024}{Fossdal et al.
2024}). Due to their simpler and more homogeneous wood anatomy, these
species display lighter wood and a narrower range of wood density (e.g.,
\protect\hyperlink{ref-Mo2024}{Mo et al. 2024}), which may result in
less flexibility in the crown architecture. These differences underscore
how the contrasting evolutionary strategies of angiosperms and
gymnosperms shape their architectural diversity, reflecting adaptations
to distinct ecological niches.

\hypertarget{consequences-for-agb-estimates}{%
\subsection{Consequences for AGB
estimates}\label{consequences-for-agb-estimates}}

Our findings revealed minimal differences in AGB estimates derived from
\(\text{H}_\text{WB}\) and \(\text{H}_\text{PL}\) (Fig.~\ref{fig-agb} a,
b). This result can stem from the fact that over 97\% of the trees used
to estimate AGB are small to medium-sized (DBH \textless{} 70 cm
(\protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
\protect\hyperlink{ref-Murphy2022}{Murphy and Bradford 2022})), whose
H-DBH relationships can still be well captured by the power-law function
(Fig.~\ref{fig-compare} a) (e.g.,
\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). In contrast,
individuals with DBH ≥ 70 cm, while typically forming a small proportion
of the tree population, contribute significantly to biomass storage
(e.g., \protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015}). For
instance, Slik et al. (\protect\hyperlink{ref-Slik2013}{2013}) reported
that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the dataset in
tropical forests, yet stored 25-45\% of the total AGB on average.
However, the fundamental assumption of the power-law function does not
fully reflect the asymptotic nature of height-diameter allometries in
large trees (Fig.~\ref{fig-compare} a), potentially leading to
significant inaccuracies in stand-level biomass estimates (e.g.,
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-Fayolle2016}{Fayolle et al. 2016};
\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}).
Because large trees contribute disproportionately to overall biomass
(e.g., \protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
\protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015};
\protect\hyperlink{ref-Ligot2018}{Ligot et al. 2018}), if direct height
measurements are unavailable, height estimates should be derived by
fitting asymptotic models (e.g., Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}) to
datasets with sufficient representation of large trees, ensuring that
the saturating component is accurately captured
(\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018}).

In the scenarios where DBH is not directly measured, models estimating
AGB using estimated DBH consistently perform poorly (Fig.~\ref{fig-agb}
c, d, e, f). Several reasons may account for this underperformance.
Since the gMM and Weibull models (Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}) do
not allow simply switching the roles of explanatory and response
variables. Consequently, we relied solely on the power-law function to
estimate DBH from tree height and crown radius, which may oversimplify
the complex interactions between tree-size variables, potentially
leading to inaccurate DBH estimates (e.g.,
\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012}).
Additionally, most existing AGB estimation equations typically include
DBH as a squared term in equations (e.g.,
\protect\hyperlink{ref-Brown1989}{Brown et al. 1989};
\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}), meaning that even
minor errors in DBH predictions can significantly affect the accuracy of
AGB estimates (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
Interestingly, our results revealed that, although the \(CR,H\) model
provided more accurate direct DBH predictions, using those predictions
in the AGB equation resulted in higher RMSE and bias than when DBH
estimates from the \(CR \times H\) model were used (Fig.~\ref{fig-agb}
c, d). This is likely due to the \(CR \times H\) model's ability to
capture the inherent interdependence between the crown radius and tree
height (\emph{r} = 0.703), offering a more biologically realistic
framework (e.g., \protect\hyperlink{ref-Dormann2013}{Dormann et al.
2013}). The observed importance of crown radius in improving AGB
estimation aligns with prior studies (e.g.,
\protect\hyperlink{ref-Henry2010}{Henry et al. 2010};
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014}), whereas our
finding of strong relationships between AGB and crown dimensions is
consistent with those reported by Jucker et al.
(\protect\hyperlink{ref-Jucker2017}{2017}). While the interdependence of
tree height and crown radius requires careful model refinement to
address potential collinearity issues
(\protect\hyperlink{ref-Dormann2013}{Dormann et al. 2013}), their strong
relationship with AGB suggests the feasibility of directly estimating
AGB from EO data, bypassing DBH prediction and simplifying the
estimation process (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
2017}).

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Our results underscore that saturating functions---particularly the
Weibull model---offer greater accuracy for H--DBH relationships in both
angiosperms and gymnosperms than do conventional power-law models, which
can underestimate biomass in larger trees. For CR--DBH allometry,
different functional forms emerged as optimal for angiosperms
(power-law) and gymnosperms (gMM), reflecting distinct growth
strategies. Although including wood density did not significantly
improve predictive performance, it offered insights into interspecific
trade-offs in tree architecture. These findings are especially important
for large-scale carbon assessments, where even small misjudgments in
tree size can lead to major errors in biomass calculations. Our results
also highlight the importance of incorporating saturating models when
forecasting ecosystem services in forests with a broad range of species
and structural heterogeneity. Future studies should investigate
additional functional traits and environmental factors to improve both
ecological understanding and carbon monitoring practices at global
scales.

\newpage

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

We thank the editors and two anonymous reviewers, whose comments greatly
improved the manuscript. We also thank the Alliance of International
Science Organizations (ANSO) for providing financial support for TDN's
Master's studies.

\hypertarget{funding}{%
\section*{Funding}\label{funding}}
\addcontentsline{toc}{section}{Funding}

MK was supported by the Xingdian Talenj Support Program of Yunnan
Province, the Xishuangbanna State Rainforest Talent Support Program
(E4BN041B01), and the CAS President's International Fellowship
Initiative (2020FYB0003).

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Aiba2009}{}}%
Aiba M, Nakashizuka T (2009) Architectural differences associated with
adult stature and wood density in 30 temperate tree species. Functional
Ecology 23:265--273.
\url{https://doi.org/10.1111/j.1365-2435.2008.01500.x}

\leavevmode\vadjust pre{\hypertarget{ref-Anten2009}{}}%
Anten N, Schieving F (2009) The {Role} of {Wood} {Mass} {Density} and
{Mechanical} {Constraints} in the {Economy} of {Tree} {Architecture}.
The American naturalist 175:250--60.
\url{https://doi.org/10.1086/649581}

\leavevmode\vadjust pre{\hypertarget{ref-Bazrgar2024}{}}%
Bazrgar AB, Thevathasan N, Gordon A, Simpson J (2024) Allometric
equations for estimating aboveground biomass carbon in five tree species
grown in an intercropping agroforestry system in southern {Ontario},
{Canada}. Agroforestry Systems 98:739--749.
\url{https://doi.org/10.1007/s10457-023-00942-z}

\leavevmode\vadjust pre{\hypertarget{ref-Brown2004}{}}%
Brown JH, Gillooly JF, Allen AP, et al (2004) Toward a {Metabolic}
{Theory} of {Ecology}. Ecology 85:1771--1789.
\url{https://doi.org/10.1890/03-9000}

\leavevmode\vadjust pre{\hypertarget{ref-Brown1989}{}}%
Brown S, Gillespie AJR, Lugo AE (1989) Biomass {Estimation} {Methods}
for {Tropical} {Forests} with {Applications} to {Forest} {Inventory}
{Data}. Forest Science 35:881--902.
\url{https://doi.org/10.1093/forestscience/35.4.881}

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter B, Gelman A, Hoffman MD, et al (2017) Stan: {A}
{Probabilistic} {Programming} {Language}. Journal of Statistical
Software 76:1--32. \url{https://doi.org/10.18637/jss.v076.i01}

\leavevmode\vadjust pre{\hypertarget{ref-Chave2009}{}}%
Chave J, Coomes D, Jansen S, et al (2009) Towards a worldwide wood
economics spectrum. Ecology Letters 12:351--366.
\url{https://doi.org/10.1111/j.1461-0248.2009.01285.x}

\leavevmode\vadjust pre{\hypertarget{ref-Chave2014}{}}%
Chave J, Réjou-Méchain M, Búrquez A, et al (2014) Improved allometric
models to estimate the aboveground biomass of tropical trees. Global
Change Biology 20:3177--3190. \url{https://doi.org/10.1111/gcb.12629}

\leavevmode\vadjust pre{\hypertarget{ref-Cochrane2023}{}}%
Cochrane A, Sims CR, Bejjanki VR, et al (2023) Multiple timescales of
learning indicated by changes in evidence-accumulation processes during
perceptual decision-making. npj Science of Learning 8:1--10.
\url{https://doi.org/10.1038/s41539-023-00168-9}

\leavevmode\vadjust pre{\hypertarget{ref-Dormann2013}{}}%
Dormann CF, Elith J, Bacher S, et al (2013) Collinearity: A review of
methods to deal with it and a simulation study evaluating their
performance. Ecography 36:27--46.
\url{https://doi.org/10.1111/j.1600-0587.2012.07348.x}

\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022a}{}}%
Fajardo A (2022) Wood density relates negatively to maximum plant height
across major angiosperm and gymnosperm orders. American Journal of
Botany 109:250--258. \url{https://doi.org/10.1002/ajb2.1805}

\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022b}{}}%
Fajardo A, Piper FI, García-Cervigón AI (2022) The intraspecific
relationship between wood density, vessel diameter and other traits
across environmental gradients. Functional Ecology 36:1585--1598.
\url{https://doi.org/10.1111/1365-2435.14069}

\leavevmode\vadjust pre{\hypertarget{ref-Falster2017}{}}%
Falster DS, Brännström Å, Westoby M, Dieckmann U (2017) Multitrait
successional forest dynamics enable diverse competitive coexistence.
Proceedings of the National Academy of Sciences 114:E2719--E2728.
\url{https://doi.org/10.1073/pnas.1610206114}

\leavevmode\vadjust pre{\hypertarget{ref-Falster2003}{}}%
Falster DS, Westoby M (2003) Plant height and evolutionary games. Trends
in Ecology \& Evolution 18:337--343.
\url{https://doi.org/10.1016/S0169-5347(03)00061-2}

\leavevmode\vadjust pre{\hypertarget{ref-Fauset2015}{}}%
Fauset S, Johnson MO, Gloor M, et al (2015) Hyperdominance in
{Amazonian} forest carbon cycling. Nature Communications 6:6857.
\url{https://doi.org/10.1038/ncomms7857}

\leavevmode\vadjust pre{\hypertarget{ref-Fayolle2016}{}}%
Fayolle A, Loubota Panzou GJ, Drouet T, et al (2016) Taller trees,
denser stands and greater biomass in semi-deciduous than in evergreen
lowland central {African} forests. Forest Ecology and Management
374:42--50. \url{https://doi.org/10.1016/j.foreco.2016.04.033}

\leavevmode\vadjust pre{\hypertarget{ref-Feldpausch2011}{}}%
Feldpausch TR, Banin L, Phillips OL, et al (2011) Height-diameter
allometry of tropical forest trees. Biogeosciences 8:1081--1106.
\url{https://doi.org/10.5194/bg-8-1081-2011}

\leavevmode\vadjust pre{\hypertarget{ref-Filipescu2012}{}}%
Filipescu CN, Groot A, MacIsaac DA, et al (2012) Prediction of
{Diameter} {Using} {Height} and {Crown} {Attributes}: {A} {Case}
{Study}. Western Journal of Applied Forestry 27:30--35.
\url{https://doi.org/10.1093/wjaf/27.1.30}

\leavevmode\vadjust pre{\hypertarget{ref-Fossdal2024}{}}%
Fossdal CG, Krokene P, Olsen JE, et al (2024) Epigenetic stress memory
in gymnosperms. Plant Physiology 195:1117--1133.
\url{https://doi.org/10.1093/plphys/kiae051}

\leavevmode\vadjust pre{\hypertarget{ref-Gelder2006}{}}%
Gelder HAV, Poorter L, Sterck FJ (2006) Wood mechanics, allometry, and
life‐history variation in a tropical rain forest tree community.
https://doi.org/\url{https://doi.org/10.1111/j.1469-8137.2006.01757.x}

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
Gelman A, Carlin JB, Stern HS, et al (2015)
\href{https://doi.org/10.1201/b16018}{Bayesian {Data} {Analysis}}, 3rd
edn. Chapman; Hall/CRC, New York

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2006}{}}%
Gelman A, Hill J (2006) Data analysis using regression and
multilevel/hierarchical models. Cambridge University Press

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008}{}}%
Gelman A, Jakulin A, Pittau MG, Su Y-S (2008) A weakly informative
default prior distribution for logistic and other regression models. The
Annals of Applied Statistics 2:1360--1383.
\url{https://doi.org/10.1214/08-AOAS191}

\leavevmode\vadjust pre{\hypertarget{ref-Goodman2014}{}}%
Goodman RC, Phillips OL, Baker TR (2014) The importance of crown
dimensions to improve tropical tree biomass estimates. Ecological
Applications 24:680--698. \url{https://doi.org/10.1890/13-0070.1}

\leavevmode\vadjust pre{\hypertarget{ref-Henry2010}{}}%
Henry M, Besnard A, Asante WA, et al (2010) Wood density, phytomass
variations within and among trees, and allometric equations in a
tropical rainforest of {Africa}. Forest Ecology and Management
260:1375--1388. \url{https://doi.org/10.1016/j.foreco.2010.07.040}

\leavevmode\vadjust pre{\hypertarget{ref-Hulshof2015}{}}%
Hulshof CM, Swenson NG, Weiser MD (2015) Tree height--diameter allometry
across the {United} {States}. Ecology and Evolution 5:1193--1204.
\url{https://doi.org/10.1002/ece3.1328}

\leavevmode\vadjust pre{\hypertarget{ref-Iida2012}{}}%
Iida Y, Poorter L, Sterck FJ, et al (2012) Wood density explains
architectural differentiation across 145 co-occurring tropical tree
species. Functional Ecology 26:274--282.
\url{https://doi.org/10.1111/j.1365-2435.2011.01921.x}

\leavevmode\vadjust pre{\hypertarget{ref-Jucker2017}{}}%
Jucker T, Caspersen J, Chave J, et al (2017) Allometric equations for
integrating remote sensing imagery into forest monitoring programmes.
Global Change Biology 23:177--190.
\url{https://doi.org/10.1111/gcb.13388}

\leavevmode\vadjust pre{\hypertarget{ref-Jucker2022}{}}%
Jucker T, Fischer FJ, Chave J, et al (2022) Tallo: {A} global tree
allometry and crown architecture database. Global Change Biology
28:5254--5268. \url{https://doi.org/10.1111/gcb.16302}

\leavevmode\vadjust pre{\hypertarget{ref-Jucker2024}{}}%
Jucker T, Fischer FJ, Chave J, et al (2024)
\href{https://doi.org/10.1101/2024.09.14.613032}{The global spectrum of
tree crown architecture}

\leavevmode\vadjust pre{\hypertarget{ref-Kattge2020}{}}%
Kattge J, Bönisch G, Díaz S, et al (2020) {TRY} plant trait database --
enhanced coverage and open access. Global Change Biology 26:119--188.
\url{https://doi.org/10.1111/gcb.14904}

\leavevmode\vadjust pre{\hypertarget{ref-King2005}{}}%
King DA (2005) Linking tree form, allocation and growth with an
allometrically explicit model. Ecological Modelling 185:77--91.
\url{https://doi.org/10.1016/j.ecolmodel.2004.11.017}

\leavevmode\vadjust pre{\hypertarget{ref-King2011}{}}%
King DA, Clark DA (2011)
\href{https://www.jstor.org/stable/23012404}{Allometry of emergent tree
species from saplings to above-canopy adults in a {Costa} {Rican} rain
forest}. Journal of Tropical Ecology 27:573--579

\leavevmode\vadjust pre{\hypertarget{ref-King2006}{}}%
King DA, Davies SJ, Tan S, Noor NSMd (2006) The role of wood density and
stem support costs in the growth and mortality of tropical trees.
Journal of Ecology 94:670--680.
\url{https://doi.org/10.1111/j.1365-2745.2006.01112.x}

\leavevmode\vadjust pre{\hypertarget{ref-Kooyman2009}{}}%
Kooyman RM, Westoby M (2009) Costs of height gain in rainforest
saplings: Main-stem scaling, functional traits and strategy variation
across 75 species. Annals of Botany 104:987--993.
\url{https://doi.org/10.1093/aob/mcp185}

\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
Landau WM (2021) \href{https://doi.org/10.21105/joss.02959}{The targets
r package: A dynamic make-like function-oriented pipeline toolkit for
reproducibility and high-performance computing}. Journal of Open Source
Software 6:2959

\leavevmode\vadjust pre{\hypertarget{ref-Laurans2024}{}}%
Laurans M, Munoz F, Charles-Dominique T, et al (2024) Why incorporate
plant architecture into trait-based ecology? Trends in Ecology \&
Evolution. \url{https://doi.org/10.1016/j.tree.2023.11.011}

\leavevmode\vadjust pre{\hypertarget{ref-Ligot2018}{}}%
Ligot G, Gourlet-Fleury S, Ouédraogo D-Y, et al (2018) The limited
contribution of large trees to annual biomass production in an
old-growth tropical forest. Ecological Applications: A Publication of
the Ecological Society of America 28:1273--1281.
\url{https://doi.org/10.1002/eap.1726}

\leavevmode\vadjust pre{\hypertarget{ref-Lines2012}{}}%
Lines ER, Zavala MA, Purves DW, Coomes DA (2012) Predictable changes in
aboveground allometry of trees along gradients of temperature, aridity
and competition. Global Ecology and Biogeography 21:1017--1028.
\url{https://doi.org/10.1111/j.1466-8238.2011.00746.x}

\leavevmode\vadjust pre{\hypertarget{ref-Loehle2016}{}}%
Loehle C (2016) Biomechanical constraints on tree architecture. Trees
30:2061--2070. \url{https://doi.org/10.1007/s00468-016-1433-2}

\leavevmode\vadjust pre{\hypertarget{ref-LoubotaPanzou2018}{}}%
Loubota Panzou GJ, Ligot G, Gourlet-Fleury S, et al (2018) Architectural
differences associated with functional traits among 45 coexisting tree
species in {Central} {Africa}. Functional Ecology 32:2583--2593.
\url{https://doi.org/10.1111/1365-2435.13198}

\leavevmode\vadjust pre{\hypertarget{ref-MartinezCano2019}{}}%
Martínez Cano I, Muller-Landau HC, Wright SJ, et al (2019) Tropical tree
height and crown allometries for the {Barro} {Colorado} {Nature}
{Monument}, {Panama}: A comparison of alternative hierarchical models
incorporating interspecific variation in relation to life history
traits. Biogeosciences 16:847--862.
\url{https://doi.org/10.5194/bg-16-847-2019}

\leavevmode\vadjust pre{\hypertarget{ref-Mikola1938}{}}%
Mikola P (1938) \href{https://www.silvafennica.fi/article/4546}{Crown
and stem form of {Norway} spruce in the snow damage areas of {Maanselkä}
in {Northern} {Finland}}. Silva Fennica

\leavevmode\vadjust pre{\hypertarget{ref-Mo2024}{}}%
Mo L, Crowther TW, Maynard DS, et al (2024) The global distribution and
drivers of wood density and their impact on forest carbon stocks. Nature
Ecology \& Evolution 1--18.
\url{https://doi.org/10.1038/s41559-024-02564-9}

\leavevmode\vadjust pre{\hypertarget{ref-Molto2014}{}}%
Molto Q, Hérault B, Boreux J-J, et al (2014) Predicting tree heights for
biomass estimates in tropical forests \&ndash; a test from {French}
{Guiana}. Biogeosciences 11:3121--3130.
\url{https://doi.org/10.5194/bg-11-3121-2014}

\leavevmode\vadjust pre{\hypertarget{ref-Moore2010}{}}%
Moore JR (2010) Allometric equations to predict the total above-ground
biomass of radiata pine trees. Annals of Forest Science 67:806--806.
\url{https://doi.org/10.1051/forest/2010042}

\leavevmode\vadjust pre{\hypertarget{ref-Muller-Landau2004}{}}%
Muller-Landau HC (2004) Interspecific and {Inter}-site {Variation} in
{Wood} {Specific} {Gravity} of {Tropical} {Trees}. Biotropica 36:20--32.
\url{https://doi.org/10.1111/j.1744-7429.2004.tb00292.x}

\leavevmode\vadjust pre{\hypertarget{ref-Murphy2022}{}}%
Murphy HT, Bradford MG (2022) The role of big trees and abundant species
in driving spatial patterns of species richness in an {Australian}
tropical rainforest. Ecology and Evolution 12:e9324.
\url{https://doi.org/10.1002/ece3.9324}

\leavevmode\vadjust pre{\hypertarget{ref-Niklas2007}{}}%
Niklas KJ (2007) Maximum plant height and the biophysical factors that
limit it. Tree Physiology 27:433--440.
\url{https://doi.org/10.1093/treephys/27.3.433}

\leavevmode\vadjust pre{\hypertarget{ref-Niklas1994}{}}%
Niklas KJ (1994) Plant allometry: The scaling of form and process.
University of Chicago Press, Chicago

\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012a}{}}%
Poorter H, Niklas KJ, Reich PB, et al (2012a) Biomass allocation to
leaves, stems and roots: Meta-analyses of interspecific variation and
environmental control. New Phytologist 193:30--50.
\url{https://doi.org/10.1111/j.1469-8137.2011.03952.x}

\leavevmode\vadjust pre{\hypertarget{ref-Poorter2006}{}}%
Poorter L, Bongers L, Bongers F (2006) Architecture of 54
{Moist}-{Forest} {Tree} {Species}: {Traits}, {Trade}-{Offs}, and
{Functional} {Groups}. Ecology 87:1289--1301.
\url{https://doi.org/10.1890/0012-9658(2006)87\%5B1289:AOMTST\%5D2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012b}{}}%
Poorter L, Lianes E, Moreno-de las Heras M, Zavala MA (2012b)
Architecture of {Iberian} canopy tree species in relation to wood
density, shade tolerance and climate. Plant Ecology 213:707--722.
\url{https://doi.org/10.1007/s11258-012-0032-6}

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2023}{}}%
R Core Team (2023) \href{https://www.R-project.org/}{R: A language and
environment for statistical computing}. R Foundation for Statistical
Computing, Vienna, Austria

\leavevmode\vadjust pre{\hypertarget{ref-Shendryk2016}{}}%
Shendryk I, Broich M, Tulbure MG, Alexandrov SV (2016) Bottom-up
delineation of individual trees from full-waveform airborne laser scans
in a structurally complex eucalypt forest. Remote Sensing of Environment
173:69--83. \url{https://doi.org/10.1016/j.rse.2015.11.008}

\leavevmode\vadjust pre{\hypertarget{ref-Shinozaki1964a}{}}%
Shinozaki K, Yoda K, Hozumi K, Kira T (1964) A quantitative analysis of
plant form-the pipe model theory: I. Basic analyses. Japanese Journal of
ecology 14:97--105

\leavevmode\vadjust pre{\hypertarget{ref-Sileshi2023}{}}%
Sileshi GW, Nath AJ, Kuyah S (2023) Allometric scaling and allocation
patterns: {Implications} for predicting productivity across plant
communities. Frontiers in Forests and Global Change 5:
\url{https://doi.org/10.3389/ffgc.2022.1084480}

\leavevmode\vadjust pre{\hypertarget{ref-Sivula2023}{}}%
Sivula T, Magnusson M, Matamoros AA, Vehtari A (2023)
\href{https://doi.org/10.48550/arXiv.2008.10296}{Uncertainty in
{Bayesian} {Leave}-{One}-{Out} {Cross}-{Validation} {Based} {Model}
{Comparison}}

\leavevmode\vadjust pre{\hypertarget{ref-Slik2013}{}}%
Slik JWF, Paoli G, McGuire K, et al (2013) Large trees drive forest
aboveground biomass variation in moist lowland forests across the
tropics. Global Ecology and Biogeography 22:1261--1271.
\url{https://doi.org/10.1111/geb.12092}

\leavevmode\vadjust pre{\hypertarget{ref-Song2024}{}}%
Song X, Li J, Zeng X (2024) Parameterization of height--diameter and
crown radius--diameter relationships across the globe. Journal of Plant
Ecology 17:rtae005. \url{https://doi.org/10.1093/jpe/rtae005}

\leavevmode\vadjust pre{\hypertarget{ref-Sterck1998}{}}%
Sterck F, Bongers F (1998)
\href{https://www.ncbi.nlm.nih.gov/pubmed/21684910}{Ontogenetic changes
in size, allometry, and mechanical design of tropical rain forest
trees}. American Journal of Botany 85:266

\leavevmode\vadjust pre{\hypertarget{ref-Sullivan2018}{}}%
Sullivan MJP, Lewis SL, Hubau W, et al (2018) Field methods for sampling
tree height for tropical forest biomass estimation. Methods in Ecology
and Evolution 9:1179--1189.
\url{https://doi.org/10.1111/2041-210X.12962}

\leavevmode\vadjust pre{\hypertarget{ref-Swenson2007}{}}%
Swenson N, Enquist B (2007) Ecological and evolutionary determinants of
a key plant functional trait: {Wood} density and its community-wide
variation across latitude and elevation. American journal of botany
94:451--9. \url{https://doi.org/10.3732/ajb.94.3.451}

\leavevmode\vadjust pre{\hypertarget{ref-Tatsumi2023}{}}%
Tatsumi S, Yamaguchi K, Furuya N (2023) {ForestScanner}: {A} mobile
application for measuring and mapping trees with {LiDAR}-equipped
{iPhone} and {iPad}. Methods in Ecology and Evolution 14:1603--1609.
\url{https://doi.org/10.1111/2041-210X.13900}

\leavevmode\vadjust pre{\hypertarget{ref-Tavankar2019}{}}%
Tavankar F, Lo Monaco A, Nikooy M, et al (2019) Snow damages on trees of
an uneven age in mixed broadleaf forests: Effects of topographical
conditions and tree characteristics. Journal of Forestry Research
30:1383--1394. \url{https://doi.org/10.1007/s11676-018-0710-x}

\leavevmode\vadjust pre{\hypertarget{ref-Thomas1996}{}}%
Thomas SC (1996) Asymptotic height as a predictor of growth and
allometric characteristics in malaysian rain forest trees. American
Journal of Botany 83:556--566.
\url{https://doi.org/10.1002/j.1537-2197.1996.tb12739.x}

\leavevmode\vadjust pre{\hypertarget{ref-Vehtari2017}{}}%
Vehtari A, Gelman A, Gabry J (2017) Practical {Bayesian} model
evaluation using leave-one-out cross-validation and {WAIC}. Statistics
and Computing 27:1413--1432.
\url{https://doi.org/10.1007/s11222-016-9696-4}

\leavevmode\vadjust pre{\hypertarget{ref-Vehtari2021}{}}%
Vehtari A, Gelman A, Simpson D, et al (2021) Rank-{Normalization},
{Folding}, and {Localization}: {An} {Improved} {Rˆ} for {Assessing}
{Convergence} of {MCMC} (with {Discussion}). Bayesian Analysis
16:667--718. \url{https://doi.org/10.1214/20-BA1221}

\leavevmode\vadjust pre{\hypertarget{ref-West1999}{}}%
West G, Brown J, Enquist B (1999) A general model for the structure and
allometry of plant vascular systems. Nature 400:664--667.
\url{https://doi.org/10.1038/23251}

\leavevmode\vadjust pre{\hypertarget{ref-Yang2023}{}}%
Yang J, Swenson NG (2023) Height and crown allometries and their
relationship with functional traits: {An} example from a subtropical wet
forest. Ecology and Evolution 13:e9804.
\url{https://doi.org/10.1002/ece3.9804}

\end{CSLReferences}

\end{document}

### 2025-03-26 ms/ms_old.tex
Initial commit for tree-allo

--- 2025-03-26 ms/ms.tex
@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 interactions influenced by various factors, such as species diversity,
-environmental conditions, and individual variability (e.g.,
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
-\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}). Consequently,
-power-law models often fail to accurately capture the allometric
-relationships across the full range of tree sizes, leading to
-underestimation of dimensions for smaller trees and overestimation for
+environmental conditions, and individual variability
+(\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}; e.g.,
+\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}).
+Consequently, power-law models often fail to accurately capture the
+allometric relationships across the full range of tree sizes, leading to
+underestimation of dimensions for smaller trees and underestimation for
 larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
@@
 Earth observation (EO) technologies, including remote sensing and Light
-Detection and Ranging (LiDAR), have revolutionized the monitoring of
-forests (e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017})
-by enabling the identification and measurement of individual tree
-heights and crown dimensions using airborne imagery (e.g.,
+Detection and Ranging (LiDAR), have revolutionized monitoring forests
+(e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}) by
+enabling the identification and measurement of individual tree heights
+and crown dimensions using airborne imagery (e.g.,
 \protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). However,
-because existing allometric models still rely on stem diameter as a
-crucial input for biomass estimation, this reliance presents a challenge
-for effectively integrating EO technologies into AGB estimation
-(\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
+\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Because
+existing allometric models still depend on stem diameter as a crucial
+input for biomass estimation, it also presents the challenge of
+determining the most effective way to utilize EO technologies for AGB
+estimation (\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
 \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
@@
 considering underlying ecological constraints, particularly the
-interplay of plant functional traits and their trade-offs influence
+interplay of plant functional traits and their trade-offs influences
 individual tree architecture and the composition of entire forest
@@
 functional traits, wood density is closely linked to both plant life
-history strategies (\protect\hyperlink{ref-Fajardo2022a}{Fajardo 2022})
-and light demand (e.g., \protect\hyperlink{ref-King2006}{King et al.
-2006}; \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a}). Several
+history strategies
+(\protect\hyperlink{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light
+demand (e.g., \protect\hyperlink{ref-King2006}{King et al. 2006};
+\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}). Several
 studies suggest that wood density explains a trade-off between rapid
@@
 relied on site-specific or regional observations with limited samples
-and species. By integrating species-level wood density (WD) into a
-multilevel model, our methodology avoids the post-hoc correlations used
-in previous studies, offering a more universal and scalable model,
-effectively capturing variability from the species to the community
-level. Our main objective is to develop models informed by tree
-allometric variables (H, CR, DBH, WD) to capture interspecific variation
-in allometric scaling globally and offer an improved framework for
-estimating above ground biomass. Specifically, we ask the following
-questions: (1) What is the best function to estimate tree architecture
-from DBH and \emph{vice versa}? (2) How does wood density shape
-interspecific variation in these allometric scaling relationships? (3)
-How does the choice of allometric relationships affect subsequent AGB
-estimation?
+and species. By integrating species-level wood density into a multilevel
+model, our methodology avoids the post-hoc correlations used in previous
+studies, offering a more universal and scalable model, effectively
+capturing variability from the species to the community level.
+Specifically, we ask the following questions: (1) What is the best
+function to estimate tree architecture from DBH and \emph{vice versa}?
+(2) How does wood density shape interspecific variation in these
+allometric scaling relationships? (3) How does the choice of allometric
+relationships affect subsequent AGB estimation?
 
@@
 Our analysis combined tree allometry data from Tallo, which includes
-standardized measurements of DBH, height, and crown radius for 5,163
-species across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker
-et al. 2022}, https://doi.org/10.5281/zenodo.6637599) along with wood
-density data for 10,995 species from TRY version 6
+standardized measurements of DBH, height, crown radius for 5,163 species
+across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker et al.
+2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
+data for 10,995 species from TRY version 6
 (\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
@@
 
-Once the best-performing models for each allometric relationship were
-identified, we tested whether adding wood density as a species-level
-predictor would further improve the predictive performance of those
-models and also examined its influence on tree architecture diversity.
-Species-specific parameters (\(\beta_j\)) were modeled as a function of
-wood density (\(u_j\)): \begin{equation}\protect\hypertarget{eq-wood}{}{
+Once the best-performing models were identified for each allometric
+relationship, we then tested whether adding a species-level predictor
+(wood density) would further improve that model's predictive
+performance. Species-specific parameters (\(\beta_j\)) were modeled as a
+function of wood density (\(u_j\)):
+\begin{equation}\protect\hypertarget{eq-wood}{}{
 \beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
@@
 and slope, respectively, and \(\sigma_{\beta}\) is the standard
-deviation of the species-specific parameters. To facilitate
-interpretation and visualization, species were then classified into
-``light'' and ``dense'' wood groups based on the 10th and 90th
-percentiles of wood density distribution in the observed dataset,
-respectively. This classification allowed us to illustrate the
-relationships between wood density and the parameters of the
-best-performing models, enabling an evaluation of how wood density
-contributes to variation in tree architecture. Additionally, we explored
+deviation of the species-specific parameters. Additionally, we explored
 a model that accounts for covariance between species-specific parameters
@@
 
-For angiosperm species, we used the widely applied equation from Chave
-et al. (\protect\hyperlink{ref-Chave2014}{2014}) to estimate aboveground
-biomass (AGB; kg of dry mass):
+For angiosperm species, we estimated aboveground biomass (AGB; kg of dry
+mass) using the equation from Chave et al.
+(\protect\hyperlink{ref-Chave2014}{2014}):
 
@@
 angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
-dataset, the best model to predict H from DBH of each clade was:
+dataset, the best model to predict H from DBH of each division was:
 \begin{equation}\protect\hypertarget{eq-h_ang}{}{
@@
 The second-best model, which used the gMM function, showed poorer
-predictive performance in angiosperms (\(\Delta\)LOOIC = 101.01) but
-showed competitive predictive performance in gymnosperms
-(\(\Delta\)LOOIC = 4.87). While power-law models (Eq.~\ref{eq-pl}) can
-exhibit saturating behavior when 0 \textless{} \emph{b} \textless{} 1,
-they performed the worst overall compared to other saturating models
+predictive performance in angiosperms (\(\Delta\) LOOIC = 101.01) but
+showed competitive predictive performance in gymnosperms (\(\Delta\)
+LOOIC = 4.87). While power-law models (Eq.~\ref{eq-pl}) can exhibit
+saturating behavior when 0 \textless{} \emph{b} \textless{} 1, they
+performed the worst overall compared to other saturating models
 (Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}). Incorporating wood density did not
@@
 and the \(R^2\) values. The difference in LOOIC relative to the
-best-performing model, \(\Delta\)LOOIC, is also included, with a
-\(\Delta\)LOOIC greater than 4 suggests a meaningful difference in
-predictive performance (\protect\hyperlink{ref-Cochrane2023}{Cochrane et
-al. 2023}; \protect\hyperlink{ref-Sivula2023}{Sivula et al. 2023}).
-Lower \(\Delta\)LOOIC values indicate models that perform closer to the
-best model in terms of predictive accuracy, while higher ELPD values
-reflect better model fit. Higher \(R^2\) values reflect better
-explanatory power, indicating how well the model explains the variance
-in the dependent variable.
+best-performing model, \(\Delta\)\_LOOIC, is also included. Lower
+\(\Delta\)\_LOOIC values indicate models that perform closer to the best
+model in terms of predictive accuracy, while higher ELPD values reflect
+better model fit. Higher \(R^2\) values reflect better explanatory
+power, indicating how well the model explains the variance in the
+dependent variable.
 
@@
 
-\caption{\label{fig-compare}Allometric relationships between DBH and
-tree height or crown radius in angiosperms (a, c) and gymnosperms (b,
-d), based on three different models: power-law (orange), generalized
-Michaelis-Menten (green), and Weibull (blue). A color gradient from
-black (low density) to yellow (high density) indicates the density of
-overlapping data points, calculated via 2D kernel density estimation on
-a 300\,\(\times\)\,300 grid. See Supplementary Figure S1 for the
-log-scale version of these plots.}
+\caption{\label{fig-compare}Allometric relationships between tree height
+and crown radius with DBH of angiosperm (a, c) and gymnosperm (b, d)
+based on three different models: power-law (orange), generalized
+Michaelis-Menten (green), and Weibull (blue). The density of overlapping
+points is shown using a color gradient ranging from black (low density)
+to yellow (high density). Point density was calculated using a 2D kernel
+density estimation on a 300 \(\times\) 300 grid, estimating the
+concentration of data points in the space of DBH and tree height or
+crown radius (see Supplementary Figure S1 for the log-scale version).}
 
@@
 1). Across the global dataset, the best model for predicting CR from DBH
-of each clade was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
+of each division was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
 \mathrm{CR_{ang}} = \mathrm{0.349}
@@
 Eq.~\ref{eq-wb}) performed worse than the power-law model, with
-\(\Delta\)LOOIC values of 124.57 for gMM and 126.35 for the Weibull
+\(\Delta\) LOOIC values of 124.57 for gMM and 126.35 for the Weibull
 function. In gymnosperms, the Weibull function showed a competitive
 performance compared to the gMM function, while the power-law function
-(Eq.~\ref{eq-pl}) performed the worst, with a \(\Delta\)LOOIC of 162.72.
-Incorporating wood density did not significantly improve predictive
-performance (Table 1).
+(Eq.~\ref{eq-pl}) performed the worst, with a \(\Delta\) LOOIC of
+162.72. Incorporating wood density did not significantly improve
+predictive performance (Table 1).
 
@@
 hierarchical models selected for tree height, crown radius, and DBH
-allometries in angiosperms and gymnosperms. Each set includes posterior
-estimates both without and with the inclusion of wood density, reported
-as median and 95\% Bayesian credible intervals for the community-level
-parameters of each allometric function. For tree height, the Weibull
-model was identified as the best for both clades. Crown radius was best
-described by a power-law model in angiosperms and a gMM model for
-gymnosperms. The best model for DBH allometry in both clades, which
-incorporates both tree height and crown radius information (\(CR,H\)
+allometries. Each set includes the best models for angiosperms and
+gymnosperms, with posterior estimates reported both without and with the
+inclusion of wood density. Table entries represent the median and 95\%
+Bayesian credible intervals for the community-level parameters of each
+allometric function. For tree height, the Weibull model was identified
+as the best for both angiosperm and gymnosperm groups. Crown radius was
+best described by a power-law model for angiosperms and a gMM model for
+gymnosperms. The best model for DBH allometry in both divisions
+incorporated both tree height and crown radius information (\(CR,H\)
 model).
@@
 \midrule
-Tree Height & Angiosperm & Weibull & With & a & 50.0 (48.0, 52.3) & 1.00 (0.966, 1.04) & 0.109 (0.0850, 0.134)\\
-Tree Height & Angiosperm & Weibull & With & b & 0.0477 (0.0456, 0.0495) & -0.000250 (-0.00189, 0.00142) & 0.0118 (0.0109, 0.0126)\\
-Tree Height & Angiosperm & Weibull & With & k & 0.628 (0.618, 0.638) & -0.0155 (-0.0254, -0.00591) & 0.141 (0.134, 0.149)\\
-Tree Height & Angiosperm & Weibull & Without & a & 50.0 (48.0, 52.2) & - & 0.109 (0.0847, 0.133)\\
-Tree Height & Angiosperm & Weibull & Without & b & 0.0476 (0.0458, 0.0495) & - & 0.0118 (0.0110, 0.0126)\\
+Tree Height & Angiosperm & Weibull & With & a & 50.0358625836953 (48.0185910757846, 52.2782021867477) & 1.0030357535955 (0.966279786642521, 1.04204878932111) & 0.1094080515 (0.08504578, 0.133867675)\\
+Tree Height & Angiosperm & Weibull & With & b & 0.0476521 (0.0456004375, 0.04954112) & -0.0002497295 (-0.0018876475, 0.001416113) & 0.0117625713666667 (0.0108990775, 0.01264796)\\
+Tree Height & Angiosperm & Weibull & With & k & 0.627988 (0.6179271, 0.638469025) & -0.0155465 (-0.025364455, -0.005908563) & 0.141325601 (0.134013175, 0.148615)\\
+Tree Height & Angiosperm & Weibull & Without & a & 49.986601524105 (48.0397359146718, 52.1676071716969) & - & 0.108623839 (0.084699285, 0.13259475)\\
+Tree Height & Angiosperm & Weibull & Without & b & 0.04761845 (0.045788395, 0.0495194425) & - & 0.0117794441333333 (0.0109647775, 0.012630005)\\
 \addlinespace
-Tree Height & Angiosperm & Weibull & Without & k & 0.629 (0.618, 0.639) & - & 0.142 (0.135, 0.149)\\
-Tree Height & Gymnosperm & Weibull & With & a & 114 (85.1, 170) & 0.974 (0.776, 1.21) & 0.0976 (0.00466, 0.257)\\
-Tree Height & Gymnosperm & Weibull & With & b & 0.0123 (0.00836, 0.0163) & 0.000633 (-0.00188, 0.00342) & 0.00382 (0.00225, 0.00553)\\
-Tree Height & Gymnosperm & Weibull & With & k & 0.751 (0.721, 0.784) & -0.0224 (-0.0538, 0.00762) & 0.113 (0.0905, 0.140)\\
-Tree Height & Gymnosperm & Weibull & Without & a & 117 (89.2, 175) & - & 0.0912 (0.00357, 0.229)\\
+Tree Height & Angiosperm & Weibull & Without & k & 0.628505 (0.61806375, 0.63871895) & - & 0.142012394 (0.135037575, 0.1493971)\\
+Tree Height & Gymnosperm & Weibull & With & a & 114.392440818366 (85.1394073936058, 170.399810107161) & 0.973669651215167 (0.775709235915639, 1.21272840336665) & 0.0976113393243333 (0.00465741875, 0.256888)\\
+Tree Height & Gymnosperm & Weibull & With & b & 0.01226515 (0.00835557525, 0.01627371) & 0.0006333235 (-0.001878769, 0.003421633) & 0.00382281789033333 (0.0022539225, 0.00552581075)\\
+Tree Height & Gymnosperm & Weibull & With & k & 0.75131 (0.720762275, 0.78392225) & -0.0223577 (-0.053752515, 0.00762075675000001) & 0.11290265 (0.0904752175, 0.139776975)\\
+Tree Height & Gymnosperm & Weibull & Without & a & 117.410929580669 (89.2163884742527, 175.128720630186) & - & 0.0912130060437333 (0.00357252325, 0.229208275)\\
 \addlinespace
-Tree Height & Gymnosperm & Weibull & Without & b & 0.0119 (0.00806, 0.0155) & - & 0.00371 (0.00230, 0.00525)\\
-Tree Height & Gymnosperm & Weibull & Without & k & 0.750 (0.719, 0.784) & - & 0.114 (0.0919, 0.139)\\
-Crown Radius & Angiosperm & Power-Law & With & a & 0.349 (0.340, 0.359) & 0.0680 (0.0395, 0.0970) & 0.331 (0.311, 0.354)\\
-Crown Radius & Angiosperm & Power-Law & With & b & 0.626 (0.617, 0.635) & -0.0110 (-0.0199, -0.00253) & 0.0921 (0.0844, 0.100)\\
-Crown Radius & Angiosperm & Power-Law & Without & a & 0.349 (0.340, 0.359) & - & 0.336 (0.316, 0.358)\\
+Tree Height & Gymnosperm & Weibull & Without & b & 0.01189855 (0.00806481775, 0.0155006175) & - & 0.00370917828 (0.00230250075, 0.005250409)\\
+Tree Height & Gymnosperm & Weibull & Without & k & 0.750261 (0.71890345, 0.7840528) & - & 0.113927871366667 (0.0919335325, 0.139492325)\\
+Crown Radius & Angiosperm & Power-Law & With & a & 0.348870320614843 (0.339558002412552, 0.358617649941404) & 0.067951 (0.0395029625, 0.09696407) & 0.331462941666667 (0.3107583, 0.353737675)\\
+Crown Radius & Angiosperm & Power-Law & With & b & 0.6262655 (0.61724685, 0.6347153) & -0.0110058 (-0.01988032, -0.00253456499999999) & 0.0921479955666667 (0.0844064425, 0.100138025)\\
+Crown Radius & Angiosperm & Power-Law & Without & a & 0.349041308960668 (0.340277352929455, 0.358866706688696) & - & 0.336416977 (0.315543425, 0.35812015)\\
 \addlinespace
-Crown Radius & Angiosperm & Power-Law & Without & b & 0.626 (0.617, 0.635) & - & 0.0929 (0.0852, 0.101)\\
-Crown Radius & Gymnosperm & gMM & With & a & 975 (312, 2010) & 0.988 (0.766, 1.27) & 0.367 (0.280, 0.474)\\
-Crown Radius & Gymnosperm & gMM & With & b & 0.629 (0.599, 0.659) & 0.0345 (0.00507, 0.0642) & 0.0949 (0.0691, 0.126)\\
-Crown Radius & Gymnosperm & gMM & With & k & 3371 (1094, 6978) & 35.7 (-769, 914) & 9.36 (0.100, 52.3)\\
-Crown Radius & Gymnosperm & gMM & Without & a & 769 (227, 1818) & - & 0.367 (0.280, 0.465)\\
+Crown Radius & Angiosperm & Power-Law & Without & b & 0.625775 (0.61686365, 0.634504125) & - & 0.0928698366666667 (0.08518862, 0.101327225)\\
+Crown Radius & Gymnosperm & gMM & With & a & 975.002465780217 (311.898479923573, 2010.23216629045) & 0.98826173732258 (0.765945182590572, 1.26629969928617) & 0.366709977 (0.27977655, 0.473532775)\\
+Crown Radius & Gymnosperm & gMM & With & b & 0.629287 (0.599257875, 0.658672975) & 0.03446325 (0.00506576425, 0.0642449925) & 0.0948956725666667 (0.0691086525, 0.126110225)\\
+Crown Radius & Gymnosperm & gMM & With & k & 3371.455 (1094.23225, 6977.7715) & 35.71465 (-768.70475, 913.79515) & 9.363854693469 (0.10023131, 52.3180125)\\
+Crown Radius & Gymnosperm & gMM & Without & a & 769.406932080409 (227.345355225331, 1818.30585925327) & - & 0.367348107833333 (0.280439875, 0.465157075)\\
 \addlinespace
-Crown Radius & Gymnosperm & gMM & Without & b & 0.631 (0.600, 0.663) & - & 0.101 (0.0755, 0.131)\\
-Crown Radius & Gymnosperm & gMM & Without & k & 2657 (809, 6274) & - & 19.0 (0.112, 200)\\
-DBH & Angiosperm & Power-Law & With & a & 0.949 (0.928, 0.970) & -0.00254 (-0.0265, 0.0190) & 0.301 (0.284, 0.320)\\
-DBH & Angiosperm & Power-Law & With & b & 0.373 (0.361, 0.386) & -0.00327 (-0.0146, 0.00891) & 0.135 (0.125, 0.145)\\
-DBH & Angiosperm & Power-Law & With & c & 0.589 (0.575, 0.605) & 0.00511 (-0.00957, 0.0200) & 0.168 (0.156, 0.179)\\
+Crown Radius & Gymnosperm & gMM & Without & b & 0.630821 (0.60042295, 0.6625657) & - & 0.101279463966667 (0.0755376825, 0.130907775)\\
+Crown Radius & Gymnosperm & gMM & Without & k & 2657.085 (808.712975, 6274.33575) & - & 18.9842685191737 (0.11158275, 200.190725)\\
+DBH & Angiosperm & Power-Law & With & a & 0.948873203444669 (0.92824124982295, 0.970115458681121) & -0.00254114 (-0.026471175, 0.019015775) & 0.301446662333333 (0.28422895, 0.320054325)\\
+DBH & Angiosperm & Power-Law & With & b & 0.3733935 (0.36112825, 0.385692275) & -0.003273 (-0.014598435, 0.008914959) & 0.1350893 (0.12519295, 0.145357375)\\
+DBH & Angiosperm & Power-Law & With & c & 0.589474 (0.574794525, 0.604623) & 0.005107385 (-0.00957317475, 0.020006805) & 0.167959387666667 (0.1564333, 0.1794579)\\
 \addlinespace
-DBH & Angiosperm & Power-Law & Without & a & 0.948 (0.927, 0.969) & - & 0.302 (0.284, 0.320)\\
-DBH & Angiosperm & Power-Law & Without & b & 0.374 (0.361, 0.386) & - & 0.135 (0.125, 0.145)\\
-DBH & Angiosperm & Power-Law & Without & c & 0.590 (0.575, 0.605) & - & 0.168 (0.157, 0.180)\\
-DBH & Gymnosperm & Power-Law & With & a & 0.971 (0.900, 1.04) & -0.0451 (-0.120, 0.0295) & 0.274 (0.225, 0.339)\\
-DBH & Gymnosperm & Power-Law & With & b & 0.408 (0.371, 0.444) & 0.0284 (-0.00773, 0.0616) & 0.112 (0.0845, 0.148)\\
+DBH & Angiosperm & Power-Law & Without & a & 0.948185804143355 (0.927358177080849, 0.969205798294886) & - & 0.302088741 (0.2843181, 0.320217575)\\
+DBH & Angiosperm & Power-Law & Without & b & 0.373565 (0.361136275, 0.38552) & - & 0.134749882666667 (0.125359425, 0.145310525)\\
+DBH & Angiosperm & Power-Law & Without & c & 0.589786 (0.5749327, 0.604729725) & - & 0.168123297666667 (0.156878925, 0.180291175)\\
+DBH & Gymnosperm & Power-Law & With & a & 0.970943839654932 (0.900241719046287, 1.04481714578926) & -0.04506835 (-0.119699575, 0.0295394575) & 0.273574103 (0.22492635, 0.338502075)\\
+DBH & Gymnosperm & Power-Law & With & b & 0.408055 (0.37113765, 0.443653025) & 0.02839555 (-0.007731709, 0.0616446325) & 0.112309440333333 (0.0845160475, 0.14804745)\\
 \addlinespace
-DBH & Gymnosperm & Power-Law & With & c & 0.588 (0.538, 0.637) & -0.0364 (-0.0841, 0.0140) & 0.165 (0.128, 0.209)\\
-DBH & Gymnosperm & Power-Law & Without & a & 0.972 (0.903, 1.05) & - & 0.273 (0.222, 0.336)\\
-DBH & Gymnosperm & Power-Law & Without & b & 0.410 (0.373, 0.446) & - & 0.115 (0.0861, 0.150)\\
-DBH & Gymnosperm & Power-Law & Without & c & 0.588 (0.540, 0.638) & - & 0.165 (0.129, 0.209)\\
+DBH & Gymnosperm & Power-Law & With & c & 0.5880865 (0.538007975, 0.636716225) & -0.0363728 (-0.0841089975, 0.0139803) & 0.164765638 (0.127992925, 0.2091867)\\
+DBH & Gymnosperm & Power-Law & Without & a & 0.972360848599851 (0.903429908240124, 1.05107834246826) & - & 0.272779666666667 (0.22238815, 0.3356855)\\
+DBH & Gymnosperm & Power-Law & Without & b & 0.409878 (0.3725057, 0.446499075) & - & 0.114795101166667 (0.0861228225, 0.15017945)\\
+DBH & Gymnosperm & Power-Law & Without & c & 0.587609 (0.540198275, 0.638262825) & - & 0.165230531666667 (0.129457475, 0.209437675)\\
 \bottomrule
@@
 1). Across the global dataset, the most effective model for predicting
-DBH from H and CR for each clade was:
+DBH from H and CR for each division was:
 \begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
@@
 distribution in the observed dataset, respectively. The shaded areas
-illustrate the 95\% Bayesian credible intervals. Only statistically
-significant associations from Fig.~\ref{fig-wood1} are plotted.}
+illustrate the 95\% CIs. Only statistically significant associations
+from Fig.~\ref{fig-wood1} are plotted.}
 
@@
 showed the highest predictive accuracy based both on the lowest root
-mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.04\%
+mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.043\%
 (Fig.~\ref{fig-agb} a). Estimates based on the power-law model
@@
 \caption{\label{fig-agb}Comparison of individual tree above-ground
-biomass (AGB) estimated using different approaches with reference AGB
+biomass estimated from different approaches with reference AGB
 (\(\text{AGB}_\text{ref}\)) derived from observed DBH and tree height in
@@
 (\(\text{AGB}_{\text{DBH} \sim \text{CR}}\)). Dashed lines represent the
-1:1 relationships, and the solid lines show linear regression fits to
-the data, highlighting how predictive accuracy varies with tree size.
+1:1 relationships, and the solid lines are linear regression fits to the
+data points to highlight how predictive accuracy varies with tree size.
 RMSE (root mean square error) and bias values for each model are
@@
 findings revealed that the power-law model exponent of angiosperms
-(\emph{b} = 0.538) was significantly smaller, while the actual exponent
-in gymnosperms (\emph{b} = 0.704) was larger than the theoretical value
-of 2/3 (Supplementary Table S3). This divergence highlights the
-importance of separately modeling the allometric relationships of each
-functional group to accurately capture their distinct scaling patterns.
-Moreover, the power-law function performed worse than the saturating
-functions (Table 1), indicating that the actual H-DBH relationship does
-not follow the simple scaling laws proposed by MST.
+(\emph{b} = 0.54) was significantly smaller, while the actual exponent
+in gymnosperms (\emph{b} = 0.7) was larger than the theoretical value of
+2/3 (Supplementary Table S3). This divergence highlights the importance
+of separately modeling the allometric relationships of each functional
+group to accurately capture their distinct scaling patterns. Moreover,
+the power-law function performed worse than the saturating functions
+(Table 1), indicating that the actual H-DBH relationship does not follow
+the simple scaling laws proposed by MST.
 
@@
 plastic crown expansion in competitive environments
-(\protect\hyperlink{ref-Lines2012}{Lines et al. 2012}; e.g.,
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}), a conclusion
-consistent with a study in tropical moist forests by Martínez Cano et
-al. (\protect\hyperlink{ref-MartinezCano2019}{2019}). In addition, the
-estimated exponent aligns closely with MST predictions for crown radius,
-supporting the selection of the power-law model for angiosperms, where
-the community-level exponent \emph{b} is near the expected 2/3
-(\(b_{\text{ang}}\) = 0.626 {[}0.617, 0.635{]}, Table 2). In contrast,
-the CR-DBH allometry in gymnosperms showed an asymptotic relationship
-best described by the gMM function, with the Weibull function also
-showing competitive predictive performance (Table 1). This asymptotic
-pattern likely arises from structural constraints in gymnosperms growth
-forms that favour vertical over lateral expansion (e.g.,
-\protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), an adaptation
+(\protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
+\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}; e.g.,
+\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}), a
+conclusion consistent with a study in tropical moist forests by Martínez
+Cano et al. (\protect\hyperlink{ref-MartinezCano2019}{2019}). In
+addition, the estimated exponent aligns closely with MST predictions for
+crown radius, supporting the selection of the power-law model for
+angiosperms, where the community-level exponent \emph{b} is near the
+expected 2/3
+(\(b_{\text{ang}} = 0.625775 \, \left[ 0.6168636, 0.6345041 \right]\),
+Table 2). In contrast, the CR-DBH allometry in gymnosperms showed an
+asymptotic relationship best described by the gMM function, with the
+Weibull function also showing competitive predictive performance (Table
+1). This asymptotic pattern likely arises from structural constraints in
+gymnosperms growth forms that favour vertical over lateral expansion
+(e.g., \protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
+\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}), an adaptation
 that can mitigate risks such as branch breakage under heavy snow loads
 (e.g., \protect\hyperlink{ref-Mikola1938}{Mikola 1938};
-\protect\hyperlink{ref-Loehle2016}{Loehle 2016};
-\protect\hyperlink{ref-Tavankar2019}{Tavankar et al. 2019}).
+\protect\hyperlink{ref-Loehle2016}{Loehle 2016}).
 
@@
 which is especially important for distinguishing among trees of similar
-height but varying DBH (\protect\hyperlink{ref-King2005}{King 2005};
-\protect\hyperlink{ref-King2011}{King and Clark 2011}). Among the tested
-models, the \(CR,H\) model
+height but varying DBH (\protect\hyperlink{ref-King2005}{King 2005}).
+Among the tested models, the \(CR,H\) model
 (\(\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}\))
@@
 of wood density (Fig.~\ref{fig-wood1} d-f). These results are consistent
-with the previous study (\protect\hyperlink{ref-Poorter2012b}{Poorter et
-al. 2012b}), which showed that among seven angiosperm species, those
-with lighter wood exhibited taller maximum heights, while no significant
+with the previous study (\protect\hyperlink{ref-Poorter2012a}{Poorter et
+al. 2012}), which showed that among seven angiosperm species, those with
+lighter wood exhibited taller maximum heights, while no significant
 correlations were observed for six gymnosperm species. A recent study
@@
 tree allometry across individuals are likely influenced by a combination
-of species identity and environmental factors. Across species or along
-broad environmental gradients, wood density systematically
-varies---tending to be denser in drier, colder climates and in
-nutrient-poor soils
-(\protect\hyperlink{ref-Muller-Landau2004}{Muller-Landau 2004};
-\protect\hyperlink{ref-Chave2009}{Chave et al. 2009};
-\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022};
-\protect\hyperlink{ref-Mo2024}{Mo et al. 2024})---reflecting species'
-adaptive strategies. Such environmental influences might confound simple
-correlations between wood density and DBH when examining diverse tree
-communities. At the intraspecific level, recent findings suggest that
-wood density may be more closely related to vessel anatomy than to tree
-architecture across different environments
-(\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022}).
-Furthermore, substantial measurement errors in tree height and crown
-radius relative to DBH may introduce noise that obscures potential
-relationships between wood density and stem diameter. These highlight
-the importance of adopting a more integrative framework that accounts
-for additional functional traits and environmental variables to deepen
-our understanding of the factors driving DBH growth.
+of species identity and environmental factors. Furthermore, substantial
+measurement errors in tree height and crown radius relative to DBH may
+introduce noise that obscures potential relationships between wood
+density and stem diameter. These highlight the importance of adopting a
+more integrative framework that accounts for additional functional
+traits and environmental variables to deepen our understanding of the
+factors driving DBH growth.
 
@@
 to estimate AGB are small to medium-sized (DBH \textless{} 70 cm
-(\protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
-\protect\hyperlink{ref-Murphy2022}{Murphy and Bradford 2022})), whose
-H-DBH relationships can still be well captured by the power-law function
+(\protect\hyperlink{ref-Slik2013}{Slik et al. 2013})), whose H-DBH
+relationships can still be well captured by the power-law function
 (Fig.~\ref{fig-compare} a) (e.g.,
@@
 \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}).
-Because large trees contribute disproportionately to overall biomass
-(e.g., \protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
+Given the disproportionate contribution of large trees in biomass
+dynamics (e.g., \protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
 \protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015};
-\protect\hyperlink{ref-Ligot2018}{Ligot et al. 2018}), if direct height
-measurements are unavailable, height estimates should be derived by
-fitting asymptotic models (e.g., Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}) to
-datasets with sufficient representation of large trees, ensuring that
-the saturating component is accurately captured
+\protect\hyperlink{ref-Ligot2018}{Ligot et al. 2018}), when heights are
+not directly measured, height estimates should be derived by fitting
+asymptotic models (e.g., Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}) to datasets
+with sufficient representation of large trees, ensuring the accurate
+capture of the saturating component
 (\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018}).
@@
 AGB estimates (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
-Interestingly, our results revealed that, although the \(CR,H\) model
-provided more accurate direct DBH predictions, using those predictions
-in the AGB equation resulted in higher RMSE and bias than when DBH
-estimates from the \(CR \times H\) model were used (Fig.~\ref{fig-agb}
-c, d). This is likely due to the \(CR \times H\) model's ability to
-capture the inherent interdependence between the crown radius and tree
-height (\emph{r} = 0.703), offering a more biologically realistic
-framework (e.g., \protect\hyperlink{ref-Dormann2013}{Dormann et al.
-2013}). The observed importance of crown radius in improving AGB
-estimation aligns with prior studies (e.g.,
-\protect\hyperlink{ref-Henry2010}{Henry et al. 2010};
-\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014}), whereas our
-finding of strong relationships between AGB and crown dimensions is
-consistent with those reported by Jucker et al.
+Interestingly, our results revealed that the \(CR,H\) model was more
+effective in predicting DBH; however, when these predicted DBH values
+were used to estimate AGB, the \(CR \times H\) model exhibited
+substantially lower RMSE and bias (Fig.~\ref{fig-agb} c, d). This is
+likely due to the \(CR \times H\) model's ability to capture the
+inherent interdependence between the crown radius and tree height
+(\emph{r} = 0.703), offering a more biologically realistic framework
+(e.g., \protect\hyperlink{ref-Dormann2013}{Dormann et al. 2013}). The
+observed importance of crown radius in improving AGB estimation aligns
+with prior studies (e.g., \protect\hyperlink{ref-Henry2010}{Henry et al.
+2010}; \protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014}),
+whereas our finding of strong relationships between AGB and crown
+dimensions is consistent with those reported by Jucker et al.
 (\protect\hyperlink{ref-Jucker2017}{2017}). While the interdependence of
@@
 ecological understanding and carbon monitoring practices at global
-scales.
+scales monitoring practices at global scales.
 
@@
 
-We thank the editors and two anonymous reviewers, whose comments greatly
-improved the manuscript. We also thank the Alliance of International
-Science Organizations (ANSO) for providing financial support for TDN's
-Master's studies.
+We sincerely thank the Alliance of International Science Organizations
+(ANSO) for providing financial support for TDN's Master's studies.
 
@@
 
-MK was supported by the Xingdian Talenj Support Program of Yunnan
-Province, the Xishuangbanna State Rainforest Talent Support Program
-(E4BN041B01), and the CAS President's International Fellowship
+MK was supported by the Xishuangbanna State Rainforest Talent Support
+Program (E4BN041B01) and the CAS President's International Fellowship
 Initiative (2020FYB0003).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Chave2009}{}}%
-Chave J, Coomes D, Jansen S, et al (2009) Towards a worldwide wood
-economics spectrum. Ecology Letters 12:351--366.
-\url{https://doi.org/10.1111/j.1461-0248.2009.01285.x}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Chave2014}{}}%
@@
 Change Biology 20:3177--3190. \url{https://doi.org/10.1111/gcb.12629}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cochrane2023}{}}%
-Cochrane A, Sims CR, Bejjanki VR, et al (2023) Multiple timescales of
-learning indicated by changes in evidence-accumulation processes during
-perceptual decision-making. npj Science of Learning 8:1--10.
-\url{https://doi.org/10.1038/s41539-023-00168-9}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022a}{}}%
-Fajardo A (2022) Wood density relates negatively to maximum plant height
-across major angiosperm and gymnosperm orders. American Journal of
-Botany 109:250--258. \url{https://doi.org/10.1002/ajb2.1805}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022b}{}}%
-Fajardo A, Piper FI, García-Cervigón AI (2022) The intraspecific
-relationship between wood density, vessel diameter and other traits
-across environmental gradients. Functional Ecology 36:1585--1598.
-\url{https://doi.org/10.1111/1365-2435.14069}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Falster2017}{}}%
@@
 \url{https://doi.org/10.1016/j.ecolmodel.2004.11.017}
-
-\leavevmode\vadjust pre{\hypertarget{ref-King2011}{}}%
-King DA, Clark DA (2011)
-\href{https://www.jstor.org/stable/23012404}{Allometry of emergent tree
-species from saplings to above-canopy adults in a {Costa} {Rican} rain
-forest}. Journal of Tropical Ecology 27:573--579
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Muller-Landau2004}{}}%
-Muller-Landau HC (2004) Interspecific and {Inter}-site {Variation} in
-{Wood} {Specific} {Gravity} of {Tropical} {Trees}. Biotropica 36:20--32.
-\url{https://doi.org/10.1111/j.1744-7429.2004.tb00292.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Murphy2022}{}}%
-Murphy HT, Bradford MG (2022) The role of big trees and abundant species
-in driving spatial patterns of species richness in an {Australian}
-tropical rainforest. Ecology and Evolution 12:e9324.
-\url{https://doi.org/10.1002/ece3.9324}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Niklas2007}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Poorter2012a}{}}%
-Poorter H, Niklas KJ, Reich PB, et al (2012a) Biomass allocation to
+Poorter H, Niklas KJ, Reich PB, et al (2012) Biomass allocation to
 leaves, stems and roots: Meta-analyses of interspecific variation and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012b}{}}%
-Poorter L, Lianes E, Moreno-de las Heras M, Zavala MA (2012b)
-Architecture of {Iberian} canopy tree species in relation to wood
-density, shade tolerance and climate. Plant Ecology 213:707--722.
-\url{https://doi.org/10.1007/s11258-012-0032-6}
-
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2023}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Sivula2023}{}}%
-Sivula T, Magnusson M, Matamoros AA, Vehtari A (2023)
-\href{https://doi.org/10.48550/arXiv.2008.10296}{Uncertainty in
-{Bayesian} {Leave}-{One}-{Out} {Cross}-{Validation} {Based} {Model}
-{Comparison}}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Slik2013}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Tavankar2019}{}}%
-Tavankar F, Lo Monaco A, Nikooy M, et al (2019) Snow damages on trees of
-an uneven age in mixed broadleaf forests: Effects of topographical
-conditions and tree characteristics. Journal of Forestry Research
-30:1383--1394. \url{https://doi.org/10.1007/s11676-018-0710-x}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Thomas1996}{}}%

### 2025-03-26 sup/sup.qmd
Initial commit for tree-allo

---
fontsize: 12pt
geometry: margin=1in
link-citations: true
bibliography: "../ms/tree-allometry.bib"
csl: "../ms/journal-of-forestry-research.csl"
crossref:
  fig-title: Figure
  fig-prefix: Figure
  tbl-title: Table
  eq-prefix: Eq.
format:
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \usepackage{setspace}
        \renewcommand{\thetable}{S\arabic{table}}
        \renewcommand{\thefigure}{S\arabic{figure}}
    prefer-html: true
  docx:
    toc: false
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{lineno}
        \linespread{1.2}

---

*Supplement of*

**Nguyen and Katabuchi, Saturating allometric relationships reveal how wood density shapes global tree architecture**

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r include=FALSE}
library(targets)
library(tidyverse)
library(here)
library(kableExtra)
library(targets)
library(patchwork)
library(MASS)
library(scales)
library(grid)
library(knitr)
library(gsubfn)
library(yaml)
library(flextable)
library(officer)

source("R/figs.R")
source("R/functions.R")
data_params <- read_yaml("data.yaml")
data_posterior <- read_yaml("posterior.yaml")
data_metrics <- read_yaml("agb_metrics_ang.yaml")
```

This supplement includes the following materials:

# Section S1. Extra figures and tables

**Table S1.** List of sources for wood density data used in this study, as released in version 6 of the TRY database.
The table includes authors' names, dataset IDs, and the name of each dataset.

**Table S2.** Characteristics of the sub-datasets used to select the best predictive model for each allometric relationship in angiosperms and gymnosperms.

**Table S3.** Parameter estimates for all hierarchical models for tree height allometry.

**Table S4.** Parameter estimates for all hierarchical models for crown radius allometry.

**Table S5.** Parameter estimates for all hierarchical models for DBH allometry.

**Table S6.** Parameter estimates for predicting tree height (m) from DBH (cm) of 1290 species (see file *TableS6_height_sp_estimates.xlsx*).

**Table S7.** Parameter estimates for predicting crown radius (m) from DBH (cm) of 821 species (see file *TableS7_crown_radius_sp_estimates.xlsx*).

**Table S8.** Parameter estimates for predicting DBH (cm) from crown radius (m) and tree height (m) of 800 species (see file *TableS8_dbh_sp_estimates.xlsx*).

**Figure S1.** Functional curves describing allometric relationships of tree dimensions with DBH on the log scale.

# Section S2. Stan code used to fit alternative allometric models

\newpage

# Section S1. Extra tables and figures

**Table S1**: List of sources for wood density data utilized in this study, as released in version 6 of the TRY database.
The table includes authors' names, dataset IDs, and the name of each dataset.

```{r echo=FALSE, results='asis'}
try_data <- process_try_data("data-raw/30953.txt")
try_data <- try_data |>
  mutate(across(where(is.character), ~iconv(.x, to = "UTF-8", sub = "")))

ft <- try_data |> flextable () |>
  align(j = colnames(try_data), align = "center", part = "all") |>
  fontsize(size = 7, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

if (knitr::is_latex_output()) {
  try_data |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage

**Table S2**: Characteristics of the sub-datasets used to select the best predictive model for each allometric relationship in angiosperms and gymnosperms.
The ranges represent the minimum and maximum values for the respective variables: DBH (cm), tree height (H, m), crown radius (CR, m), and wood density (WD, g cm^-3^).

```{r echo=FALSE, results='asis'}
tbl_subdata <- tar_read(tbl_subdata)

line_rows <- c(5, 10, 15, 20, 25)

ft <- tbl_subdata |>
  flextable() |>
  set_header_labels(
    `Dependent variables` = "Dependent\nvariable",
    `Predictor variable` = "Predictor\nvariable",
    `Characteristics` = "Characteristic",
    `Angiosperm` = "Angiosperm",
    `Gymnosperm` = "Gymnosperm"
  ) |>
  # add_header_row(
  #   values = c("", "", "", "Division"),
  #   colwidths = c(1, 1, 1, 2)
  # ) |>
  align(align = "center", part = "all") |>
  merge_h(i = 1, part = "header") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 1:ncol(tbl_subdata), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  tbl_subdata |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S3**: Posterior estimates of the parameters of the hierarchical models for tree height allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_df <- tar_read(posterior_df)

posterior_df <- posterior_df |>
  mutate(
    Intercept_CI = sapply(Intercept_CI, format_ci),
    Slope_CI = sapply(Slope_CI, format_ci),
    Tau_CI = sapply(Tau_CI, format_ci)
  )

df_h <- posterior_df |>
  filter(Dependent_variable == "Tree Height") |>
  dplyr::select(-Dependent_variable) |>
  mutate(
    Functional_form = factor(Functional_form, levels = c("Power-Law", "gMM", "Weibull")),
    Wood_density = factor(Wood_density, levels = c("Without", "With"))
  ) |>
  arrange(Division, Functional_form, Wood_density)

line_rows <- c(2, 5, 8, 11, 13, 16, 19)
special_line_rows <- c(11)

ft <- df_h |>
  flextable() |>
  set_header_labels(
    Division = "Division",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>
  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 2:ncol(df_h), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_h), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_h |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S4**: Posterior estimates of the parameters of the hierarchical models for crown radius allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_df <- tar_read(posterior_df)

posterior_df <- posterior_df |>
  mutate(
    Intercept_CI = sapply(Intercept_CI, format_ci),
    Slope_CI = sapply(Slope_CI, format_ci),
    Tau_CI = sapply(Tau_CI, format_ci)
  )

df_cr <- posterior_df |>
  filter(Dependent_variable == "Crown Radius") |>
  dplyr::select(-Dependent_variable) |>
  mutate(
    Functional_form = factor(Functional_form, levels = c("Power-Law", "gMM", "Weibull")),
    Wood_density = factor(Wood_density, levels = c("Without", "With"))
  ) |>
  arrange(Division, Functional_form, Wood_density)

line_rows <- c(2, 4, 7, 10, 12, 15, 18)
special_line_rows <- c(10)

ft <- df_cr |>
  flextable() |>
  set_header_labels(
    Division = "Division",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>
  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 2:ncol(df_cr), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_cr), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_cr |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S5**: Posterior estimates of the parameters of the hierarchical models for DBH allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_df <- tar_read(posterior_df)

posterior_df <- posterior_df |>
  mutate(
    Intercept_CI = sapply(Intercept_CI, format_ci),
    Slope_CI = sapply(Slope_CI, format_ci),
    Tau_CI = sapply(Tau_CI, format_ci)
  )
df_dbh <- posterior_df |>
  filter(Dependent_variable %in% c("DBH", "DBH1", "DBH2", "DBH3")) |>
  mutate(
    Predictor = case_when(
      Dependent_variable == "DBH1" ~ "CR × H",
      Dependent_variable == "DBH2" ~ "CR",
      Dependent_variable == "DBH3" ~ "H",
      Dependent_variable == "DBH" ~ "CR, H",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::select(-Dependent_variable)

df_dbh <- df_dbh |>
  mutate(Predictor = factor(Predictor, levels = c("CR, H", "CR × H", "CR", "H"))) |>
  dplyr::select(Division, Predictor, everything()) |>
  dplyr::arrange(Division,
          Predictor,
          factor(Wood_density, levels = c("Without", "With")),
          Parameter)

line_rows <- c(3, 6, 8, 10, 12, 15, 18, 20, 22)
special_line_rows <- c(12)

ft <- df_dbh |>
  flextable() |>

  set_header_labels(
    Division = "Division",
    Predictor = "Predictor\nvariable",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>

  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

ft <- ft |>
  border(i = line_rows, j = 2:ncol(df_dbh), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_dbh), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_dbh |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}
```

\newpage

![
Allometric relationships between tree height and crown radius versus DBH for angiosperms (a, c) and gymnosperms (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
The density of overlapping points is represented using a colour gradient ranging from black (low density) to yellow (high density).
Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, which estimate the concentration of data points in the log-transformed DBH and tree height or crown radius.
All axes are shown on the log scale.
](../figs/h_cr_dbh_log.png){#fig-compare fig-pos="H"}

\newpage

# Section S2. Stan code used to fit alternative allometric models

## Power-law model

```{stan, class.source="small", file="stan/pl.stan", echo=TRUE, eval=FALSE, output.var="hoge", tidy=TRUE}
```

\newpage

## Power-law model with wood density

```{stan, size="small", file="stan/pl_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge", tidy=TRUE}
```

\newpage

## Generalized Michaelis-Menten model

```{stan, size="small", file="stan/gmm.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Generalized Michaelis-Menten model with wood density

```{stan, size="small", file="stan/gmm_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Weibull model

```{stan, size="small", file="stan/weibull.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Weibull model with wood density

```{stan, size="small", file="stan/weibull_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

### 2025-03-26 sup/sup.tex
Initial commit for tree-allo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{lineno}
\usepackage{setspace}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\emph{Supplement of}

\textbf{Nguyen and Katabuchi, Saturating allometric relationships reveal
how wood density shapes global tree architecture}

This supplement includes the following materials:

\section{Section S1. Extra figures and
tables}\label{section-s1.-extra-figures-and-tables}

\textbf{Table S1.} List of sources for wood density data used in this
study, as released in version 6 of the TRY database. The table includes
authors' names, dataset IDs, and the name of each dataset.

\textbf{Table S2.} Characteristics of the sub-datasets used to select
the best predictive model for each allometric relationship in
angiosperms and gymnosperms.

\textbf{Table S3.} Parameter estimates for all hierarchical models for
tree height allometry.

\textbf{Table S4.} Parameter estimates for all hierarchical models for
crown radius allometry.

\textbf{Table S5.} Parameter estimates for all hierarchical models for
DBH allometry.

\textbf{Table S6.} Parameter estimates for predicting tree height (m)
from DBH (cm) of 1290 species (see file
\emph{TableS6\_height\_sp\_estimates.xlsx}).

\textbf{Table S7.} Parameter estimates for predicting crown radius (m)
from DBH (cm) of 821 species (see file
\emph{TableS7\_crown\_radius\_sp\_estimates.xlsx}).

\textbf{Table S8.} Parameter estimates for predicting DBH (cm) from
crown radius (m) and tree height (m) of 800 species (see file
\emph{TableS8\_dbh\_sp\_estimates.xlsx}).

\textbf{Figure S1.} Functional curves describing allometric
relationships of tree dimensions with DBH on the log scale.

\section{Section S2. Stan code used to fit alternative allometric
models}\label{section-s2.-stan-code-used-to-fit-alternative-allometric-models}

\newpage

\section{Section S1. Extra tables and
figures}\label{section-s1.-extra-tables-and-figures}

\textbf{Table S1}: List of sources for wood density data utilized in
this study, as released in version 6 of the TRY database. The table
includes authors' names, dataset IDs, and the name of each dataset.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{ccccc}
\toprule
No. & LastName & FirstName & DatasetID & Dataset\\
\midrule
1 & Cornelissen & Johannes & 37 & Sheffield Database\\
2 & Swaine & Emily & 51 & Tropical Plant Traits From Borneo Database\\
3 & Kleyer & Michael & 25 & The LEDA Traitbase\\
4 & Cornwell & Will & 55 & Jasper Ridge Californian Woody Plants Database\\
5 & Lloyd & Jon & 34 & The RAINFOR Plant Trait Database\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\textbf{Table S2}: Characteristics of the sub-datasets used to select
the best predictive model for each allometric relationship in
angiosperms and gymnosperms. The ranges represent the minimum and
maximum values for the respective variables: DBH (cm), tree height (H,
m), crown radius (CR, m), and wood density (WD, g
cm\textsuperscript{-3}).

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{ccccc}
\toprule
Dependent variables & Predictor variable & Characteristics & Angiosperm & Gymnosperm\\
\midrule
H & DBH & Number of trees & 79,332 & 6,306\\
H & DBH & Number of species & 1,214 & 76\\
H & DBH & Dependent variable range & 1.3 - 94 & 1.3 - 115.8\\
H & DBH & Predictor variable range & 1 - 560 & 1 - 648\\
H & DBH & WD range & 0.131 - 1.17 & 0.32 - 0.689\\
\addlinespace
CR & DBH & Number of trees & 45,350 & 4,632\\
CR & DBH & Number of species & 762 & 59\\
CR & DBH & Dependent variable range & 0.05 - 24.25 & 0.1 - 8.75\\
CR & DBH & Predictor variable range & 1 - 385 & 1 - 266\\
CR & DBH & WD range & 0.131 - 1.15 & 0.32 - 0.689\\
\addlinespace
DBH & CR, H & Number of trees & 41,713 & 3,661\\
DBH & CR, H & Number of species & 744 & 56\\
DBH & CR, H & Dependent variable range & 1 - 385 & 1 - 266\\
DBH & CR, H & Predictor variable range & CR: 0.1 - 24.25 H: 1.3 - 70.6 & CR: 0.1 - 10.65 H: 1.3 - 46.2\\
DBH & CR, H & WD range & 0.131 - 1.15 & 0.32 - 0.651\\
\addlinespace
DBH & CR × H & Number of trees & 41,733 & 3,716\\
DBH & CR × H & Number of species & 743 & 56\\
DBH & CR × H & Dependent variable range & 1 - 385 & 1 - 266\\
DBH & CR × H & Predictor variable range & CR: 0.1 - 24.25 H: 1.3 - 70.6 & CR: 0.15 - 10.65 H: 1.5 - 53.9\\
DBH & CR × H & WD range & 0.131 - 1.15 & 0.32 - 0.651\\
\addlinespace
DBH & CR & Number of trees & 41,733 & 3,702\\
DBH & CR & Number of species & 744 & 56\\
DBH & CR & Dependent variable range & 1 - 385 & 1 - 266\\
DBH & CR & Predictor variable range & 0.1 - 24.25 & 0.15 - 8.75\\
DBH & CR & WD range & 0.131 - 1.15 & 0.32 - 0.651\\
\addlinespace
DBH & H & Number of trees & 41,689 & 3,648\\
DBH & H & Number of species & 746 & 55\\
DBH & H & Dependent variable range & 1 - 385 & 1 - 266\\
DBH & H & Predictor variable range & 1.3 - 70.6 & 1.5 - 46.2\\
DBH & H & WD range & 0.131 - 1.15 & 0.32 - 0.651\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\textbf{Table S3}: Posterior estimates of the parameters of the
hierarchical models for tree height allometry. Table entries correspond
to the median and 95\% CIs for the community-level parameters of each
allometric function.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{ccccccc}
\toprule
Division & Functional\_form & Wood\_density & Parameter & Intercept\_CI & Slope\_CI & Tau\_CI\\
\midrule
Angiosperm & Power-Law & Without & a & 2.55 (2.50, 2.60) & - & 0.311 (0.296, 0.326)\\
Angiosperm & Power-Law & Without & b & 0.538 (0.530, 0.544) & - & 0.108 (0.103, 0.114)\\
Angiosperm & gMM & Without & a & 77.9 (73.8, 82.5) & - & 0.140 (0.115, 0.164)\\
Angiosperm & gMM & Without & b & 0.640 (0.629, 0.651) & - & 0.148 (0.140, 0.156)\\
Angiosperm & gMM & Without & k & 35.0 (33.2, 36.9) & - & 8.95 (8.23, 9.66)\\
\addlinespace
Angiosperm & Weibull & Without & a & 50.0 (48.0, 52.2) & - & 0.109 (0.0847, 0.133)\\
Angiosperm & Weibull & Without & b & 0.0476 (0.0458, 0.0495) & - & 0.0118 (0.0110, 0.0126)\\
Angiosperm & Weibull & Without & k & 0.629 (0.618, 0.639) & - & 0.142 (0.135, 0.149)\\
Angiosperm & Weibull & With & a & 50.0 (48.0, 52.3) & 1.00 (0.966, 1.04) & 0.109 (0.0850, 0.134)\\
Angiosperm & Weibull & With & b & 0.0477 (0.0456, 0.0495) & -0.000250 (-0.00189, 0.00142) & 0.0118 (0.0109, 0.0126)\\
\addlinespace
Angiosperm & Weibull & With & k & 0.628 (0.618, 0.638) & -0.0155 (-0.0254, -0.00591) & 0.141 (0.134, 0.149)\\
Gymnosperm & Power-Law & Without & a & 1.43 (1.30, 1.56) & - & 0.371 (0.303, 0.447)\\
Gymnosperm & Power-Law & Without & b & 0.704 (0.675, 0.732) & - & 0.113 (0.0927, 0.137)\\
Gymnosperm & gMM & Without & a & 2005 (386, 7618) & - & 0.360 (0.298, 0.434)\\
Gymnosperm & gMM & Without & b & 0.716 (0.687, 0.746) & - & 0.112 (0.0908, 0.138)\\
\addlinespace
Gymnosperm & gMM & Without & k & 1427 (287, 5308) & - & 5.90 (0.0814, 32.2)\\
Gymnosperm & Weibull & Without & a & 117 (89.2, 175) & - & 0.0912 (0.00357, 0.229)\\
Gymnosperm & Weibull & Without & b & 0.0119 (0.00806, 0.0155) & - & 0.00371 (0.00230, 0.00525)\\
Gymnosperm & Weibull & Without & k & 0.750 (0.719, 0.784) & - & 0.114 (0.0919, 0.139)\\
Gymnosperm & Weibull & With & a & 114 (85.1, 170) & 0.974 (0.776, 1.21) & 0.0976 (0.00466, 0.257)\\
\addlinespace
Gymnosperm & Weibull & With & b & 0.0123 (0.00836, 0.0163) & 0.000633 (-0.00188, 0.00342) & 0.00382 (0.00225, 0.00553)\\
Gymnosperm & Weibull & With & k & 0.751 (0.721, 0.784) & -0.0224 (-0.0538, 0.00762) & 0.113 (0.0905, 0.140)\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\textbf{Table S4}: Posterior estimates of the parameters of the
hierarchical models for crown radius allometry. Table entries correspond
to the median and 95\% CIs for the community-level parameters of each
allometric function.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{ccccccc}
\toprule
Division & Functional\_form & Wood\_density & Parameter & Intercept\_CI & Slope\_CI & Tau\_CI\\
\midrule
Angiosperm & Power-Law & Without & a & 0.349 (0.340, 0.359) & - & 0.336 (0.316, 0.358)\\
Angiosperm & Power-Law & Without & b & 0.626 (0.617, 0.635) & - & 0.0929 (0.0852, 0.101)\\
Angiosperm & Power-Law & With & a & 0.349 (0.340, 0.359) & 0.0680 (0.0395, 0.0970) & 0.331 (0.311, 0.354)\\
Angiosperm & Power-Law & With & b & 0.626 (0.617, 0.635) & -0.0110 (-0.0199, -0.00253) & 0.0921 (0.0844, 0.100)\\
Angiosperm & gMM & Without & a & 1382 (625, 2564) & - & 0.336 (0.315, 0.357)\\
\addlinespace
Angiosperm & gMM & Without & b & 0.627 (0.618, 0.636) & - & 0.0932 (0.0858, 0.101)\\
Angiosperm & gMM & Without & k & 3975 (1807, 7386) & - & 7.24 (0.0919, 50.7)\\
Angiosperm & Weibull & Without & a & 458 (175, 1761) & - & 0.311 (0.255, 0.348)\\
Angiosperm & Weibull & Without & b & 0.000765 (0.000200, 0.00200) & - & 9.6665278454e-05 (5.092834e-06, 0.00031932395)\\
Angiosperm & Weibull & Without & k & 0.627 (0.619, 0.637) & - & 0.0934 (0.0859, 0.101)\\
\addlinespace
Gymnosperm & Power-Law & Without & a & 0.298 (0.270, 0.329) & - & 0.324 (0.251, 0.415)\\
Gymnosperm & Power-Law & Without & b & 0.619 (0.591, 0.647) & - & 0.0861 (0.0627, 0.114)\\
Gymnosperm & gMM & Without & a & 769 (227, 1818) & - & 0.367 (0.280, 0.465)\\
Gymnosperm & gMM & Without & b & 0.631 (0.600, 0.663) & - & 0.101 (0.0755, 0.131)\\
Gymnosperm & gMM & Without & k & 2657 (809, 6274) & - & 19.0 (0.112, 200)\\
\addlinespace
Gymnosperm & gMM & With & a & 975 (312, 2010) & 0.988 (0.766, 1.27) & 0.367 (0.280, 0.474)\\
Gymnosperm & gMM & With & b & 0.629 (0.599, 0.659) & 0.0345 (0.00507, 0.0642) & 0.0949 (0.0691, 0.126)\\
Gymnosperm & gMM & With & k & 3371 (1094, 6978) & 35.7 (-769, 914) & 9.36 (0.100, 52.3)\\
Gymnosperm & Weibull & Without & a & 61.0 (21.5, 322) & - & 0.202 (0.0145, 0.411)\\
Gymnosperm & Weibull & Without & b & 0.00487 (0.000906, 0.0139) & - & 0.00156209428218 (5.166957e-05, 0.00459459775)\\
\addlinespace
Gymnosperm & Weibull & Without & k & 0.640 (0.606, 0.674) & - & 0.102 (0.0775, 0.133)\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\textbf{Table S5}: Posterior estimates of the parameters of the
hierarchical models for DBH allometry. Table entries correspond to the
median and 95\% CIs for the community-level parameters of each
allometric function.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{cccccccc}
\toprule
Division & Predictor & Functional\_form & Wood\_density & Parameter & Intercept\_CI & Slope\_CI & Tau\_CI\\
\midrule
Angiosperm & CR, H & Power-Law & Without & a & 0.948 (0.927, 0.969) & - & 0.302 (0.284, 0.320)\\
Angiosperm & CR, H & Power-Law & Without & b & 0.374 (0.361, 0.386) & - & 0.135 (0.125, 0.145)\\
Angiosperm & CR, H & Power-Law & Without & c & 0.590 (0.575, 0.605) & - & 0.168 (0.157, 0.180)\\
Angiosperm & CR, H & Power-Law & With & a & 0.949 (0.928, 0.970) & -0.00254 (-0.0265, 0.0190) & 0.301 (0.284, 0.320)\\
Angiosperm & CR, H & Power-Law & With & b & 0.373 (0.361, 0.386) & -0.00327 (-0.0146, 0.00891) & 0.135 (0.125, 0.145)\\
\addlinespace
Angiosperm & CR, H & Power-Law & With & c & 0.589 (0.575, 0.605) & 0.00511 (-0.00957, 0.0200) & 0.168 (0.156, 0.179)\\
Angiosperm & CR × H & Power-Law & Without & a & 0.955 (0.934, 0.978) & - & 0.307 (0.291, 0.325)\\
Angiosperm & CR × H & Power-Law & Without & b & 0.873 (0.859, 0.886) & - & 0.167 (0.156, 0.178)\\
Angiosperm & CR & Power-Law & Without & a & 0.950 (0.917, 0.983) & - & 0.424 (0.403, 0.447)\\
Angiosperm & CR & Power-Law & Without & b & 0.717 (0.699, 0.735) & - & 0.236 (0.223, 0.251)\\
\addlinespace
Angiosperm & H & Power-Law & Without & a & 0.967 (0.941, 0.993) & - & 0.353 (0.334, 0.373)\\
Angiosperm & H & Power-Law & Without & b & 0.866 (0.849, 0.881) & - & 0.200 (0.188, 0.214)\\
Gymnosperm & CR, H & Power-Law & Without & a & 0.972 (0.903, 1.05) & - & 0.273 (0.222, 0.336)\\
Gymnosperm & CR, H & Power-Law & Without & b & 0.410 (0.373, 0.446) & - & 0.115 (0.0861, 0.150)\\
Gymnosperm & CR, H & Power-Law & Without & c & 0.588 (0.540, 0.638) & - & 0.165 (0.129, 0.209)\\
\addlinespace
Gymnosperm & CR, H & Power-Law & With & a & 0.971 (0.900, 1.04) & -0.0451 (-0.120, 0.0295) & 0.274 (0.225, 0.339)\\
Gymnosperm & CR, H & Power-Law & With & b & 0.408 (0.371, 0.444) & 0.0284 (-0.00773, 0.0616) & 0.112 (0.0845, 0.148)\\
Gymnosperm & CR, H & Power-Law & With & c & 0.588 (0.538, 0.637) & -0.0364 (-0.0841, 0.0140) & 0.165 (0.128, 0.209)\\
Gymnosperm & CR × H & Power-Law & Without & a & 0.990 (0.920, 1.07) & - & 0.282 (0.231, 0.347)\\
Gymnosperm & CR × H & Power-Law & Without & b & 0.895 (0.843, 0.942) & - & 0.182 (0.147, 0.225)\\
\addlinespace
Gymnosperm & CR & Power-Law & Without & a & 0.971 (0.887, 1.06) & - & 0.344 (0.280, 0.419)\\
Gymnosperm & CR & Power-Law & Without & b & 0.791 (0.712, 0.870) & - & 0.274 (0.221, 0.336)\\
Gymnosperm & H & Power-Law & Without & a & 0.998 (0.901, 1.09) & - & 0.348 (0.287, 0.427)\\
Gymnosperm & H & Power-Law & Without & b & 0.849 (0.787, 0.915) & - & 0.224 (0.179, 0.279)\\
\bottomrule
\end{longtable*}
\endgroup{}

\newpage

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figs/h_cr_dbh_log.png}}

}

\caption{\label{fig-compare}Allometric relationships between tree height
and crown radius versus DBH for angiosperms (a, c) and gymnosperms (b,
d) based on three different models: power-law (orange), generalized
Michaelis-Menten (green), and Weibull (blue). The density of overlapping
points is represented using a colour gradient ranging from black (low
density) to yellow (high density). Point density was calculated using a
2D kernel density estimation on a 300 \(\times\) 300 grid, which
estimate the concentration of data points in the log-transformed DBH and
tree height or crown radius. All axes are shown on the log scale.}

\end{figure}%

\newpage

\section{Section S2. Stan code used to fit alternative allometric
models}\label{section-s2.-stan-code-used-to-fit-alternative-allometric-models-1}

\subsection{Power-law model}\label{power-law-model}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each trees}
  \DataTypeTok{matrix}\NormalTok{[N, K] log\_x;                }\CommentTok{// tree{-}level predictors}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[K] gamma;}
  \DataTypeTok{matrix}\NormalTok{[K, J] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[K, J] beta;}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[k, ] = gamma[k] + tau[k] * z[k, ];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  sigma \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{);}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  tau \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  gamma \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_mu[n] = log\_x[n, ] * beta[, jj[n]];}
\NormalTok{  \}}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] | log\_x[n, ] * beta[, jj[n]], sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\newpage

\subsection{Power-law model with wood
density}\label{power-law-model-with-wood-density}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} L;                    }\CommentTok{// num sp{-}level predictor}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each trees}
  \DataTypeTok{matrix}\NormalTok{[N, K] log\_x;                }\CommentTok{// tree{-}level predictors}
  \DataTypeTok{matrix}\NormalTok{[L, J] u;                    }\CommentTok{// sp{-}level predictors}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[K, L] gamma;}
  \DataTypeTok{matrix}\NormalTok{[K, J] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[K, J] beta;}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[k, ] = gamma[k, ] * u + tau[k] * z[k, ];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  to\_vector(gamma) \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  tau \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_mu[n] = log\_x[n, ] * beta[, jj[n]];}
\NormalTok{  \}}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] | log\_x[n, ] * beta[, jj[n]], sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\newpage

\subsection{Generalized Michaelis-Menten
model}\label{generalized-michaelis-menten-model}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each tree}
  \DataTypeTok{vector}\NormalTok{[N] x;                       }\CommentTok{// DBH (non{-}log{-}scale)}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] gamma\_hat;}
  \DataTypeTok{matrix}\NormalTok{[J, K] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[J, K] beta;}
  \DataTypeTok{vector}\NormalTok{[K] gamma;}
\NormalTok{  gamma[}\DecValTok{1}\NormalTok{] = gamma\_hat[}\DecValTok{1}\NormalTok{] * }\DecValTok{10}\NormalTok{;}
\NormalTok{  gamma[}\DecValTok{2}\NormalTok{] = gamma\_hat[}\DecValTok{2}\NormalTok{];}
\NormalTok{  gamma[}\DecValTok{3}\NormalTok{] = gamma\_hat[}\DecValTok{3}\NormalTok{] * }\DecValTok{1000}\NormalTok{;}
  \CommentTok{// Corrected matrix multiplication and addition}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[, k] =  gamma[k] + tau[k] * z[, k];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  sigma \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{);}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  tau \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  gamma\_hat \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  log\_mu = beta[jj, }\DecValTok{1}\NormalTok{] + beta[jj, }\DecValTok{2}\NormalTok{] .* log(x) {-}}
\NormalTok{      log(beta[jj, }\DecValTok{3}\NormalTok{] + pow(x, beta[jj, }\DecValTok{2}\NormalTok{]));}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] |}
\NormalTok{      beta[jj[n], }\DecValTok{1}\NormalTok{] +}
\NormalTok{      beta[jj[n], }\DecValTok{2}\NormalTok{] * log(x[n]) {-}}
\NormalTok{      log(beta[jj[n], }\DecValTok{3}\NormalTok{] + pow(x[n], beta[jj[n], }\DecValTok{2}\NormalTok{])),}
\NormalTok{      sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\newpage

\subsection{Generalized Michaelis-Menten model with wood
density}\label{generalized-michaelis-menten-model-with-wood-density}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} L;                    }\CommentTok{// num sp{-}level predictor}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each tree}
  \DataTypeTok{vector}\NormalTok{[N] x;                       }\CommentTok{// DBH (non{-}log{-}scale)}
  \DataTypeTok{matrix}\NormalTok{[J, L] u;                    }\CommentTok{// sp{-}level predictors}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] gamma\_int\_hat;}
  \DataTypeTok{vector}\NormalTok{[K] gamma\_slope\_hat;}
  \DataTypeTok{matrix}\NormalTok{[J, K] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[J, K] beta;}
  \DataTypeTok{row\_vector}\NormalTok{[K] gamma\_int;}
  \DataTypeTok{row\_vector}\NormalTok{[K] gamma\_slope;}
\NormalTok{  gamma\_int[}\DecValTok{1}\NormalTok{] = gamma\_int\_hat[}\DecValTok{1}\NormalTok{] * }\DecValTok{10}\NormalTok{;}
\NormalTok{  gamma\_int[}\DecValTok{2}\NormalTok{] = gamma\_int\_hat[}\DecValTok{2}\NormalTok{];}
\NormalTok{  gamma\_int[}\DecValTok{3}\NormalTok{] = gamma\_int\_hat[}\DecValTok{3}\NormalTok{] * }\DecValTok{1000}\NormalTok{;}
\NormalTok{  gamma\_slope[}\DecValTok{1}\NormalTok{] = gamma\_slope\_hat[}\DecValTok{1}\NormalTok{] * }\DecValTok{10}\NormalTok{;}
\NormalTok{  gamma\_slope[}\DecValTok{2}\NormalTok{] = gamma\_slope\_hat[}\DecValTok{2}\NormalTok{];}
\NormalTok{  gamma\_slope[}\DecValTok{3}\NormalTok{] = gamma\_slope\_hat[}\DecValTok{3}\NormalTok{] * }\DecValTok{1000}\NormalTok{;}
  \DataTypeTok{matrix}\NormalTok{[}\DecValTok{2}\NormalTok{, K] gamma = append\_row(gamma\_int, gamma\_slope);}

  \CommentTok{// Corrected matrix multiplication and addition}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[, k] = u * gamma[, k] + tau[k] * z[, k];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  tau \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  gamma\_int\_hat \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  gamma\_slope\_hat \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{1.25}\NormalTok{);}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_mu[n] = beta[jj[n], }\DecValTok{1}\NormalTok{] + beta[jj[n], }\DecValTok{2}\NormalTok{] * log(x[n]) {-}}
\NormalTok{      log(beta[jj[n], }\DecValTok{3}\NormalTok{] + pow(x[n], beta[jj[n], }\DecValTok{2}\NormalTok{]));}
\NormalTok{  \}}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] |}
\NormalTok{      beta[jj[n], }\DecValTok{1}\NormalTok{] +}
\NormalTok{      beta[jj[n], }\DecValTok{2}\NormalTok{] * log(x[n]) {-}}
\NormalTok{      log(beta[jj[n], }\DecValTok{3}\NormalTok{] + pow(x[n], beta[jj[n], }\DecValTok{2}\NormalTok{])),}
\NormalTok{      sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\newpage

\subsection{Weibull model}\label{weibull-model}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each tree}
  \DataTypeTok{vector}\NormalTok{[N] x;                       }\CommentTok{// DBH (non{-}log{-}scale)}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] gamma;}
  \DataTypeTok{matrix}\NormalTok{[J, K] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[J, K] beta;}
  \CommentTok{// Corrected matrix multiplication and addition}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[, k] =  gamma[k] + tau[k] * z[, k];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  sigma \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{);}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  tau \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  gamma \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  log\_mu = beta[jj, }\DecValTok{1}\NormalTok{] + log1m\_exp({-}beta[jj, }\DecValTok{2}\NormalTok{] .* pow(x, beta[jj, }\DecValTok{3}\NormalTok{]));}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] |}
\NormalTok{      beta[jj[n], }\DecValTok{1}\NormalTok{] + log1m\_exp({-}beta[jj[n], }\DecValTok{2}\NormalTok{] * pow(x[n], beta[jj[n], }\DecValTok{3}\NormalTok{])),}
\NormalTok{      sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\newpage

\subsection{Weibull model with wood
density}\label{weibull-model-with-wood-density}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{ \{}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} N;                    }\CommentTok{// num trees}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} J;                    }\CommentTok{// num sp}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} K;                    }\CommentTok{// num of tree{-}level predictors}
  \DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{\textgreater{} L;                    }\CommentTok{// num sp{-}level predictor}
  \DataTypeTok{array}\NormalTok{[N] }\DataTypeTok{int}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{1}\NormalTok{, }\KeywordTok{upper}\NormalTok{=J\textgreater{} jj; }\CommentTok{// sp indicator for each tree}
  \DataTypeTok{vector}\NormalTok{[N] x;                       }\CommentTok{// DBH (non{-}log{-}scale)}
  \DataTypeTok{matrix}\NormalTok{[J, L] u;                    }\CommentTok{// sp{-}level predictors}
  \DataTypeTok{vector}\NormalTok{[N] log\_y;                   }\CommentTok{// outcomes}
\NormalTok{\}}

\KeywordTok{parameters}\NormalTok{ \{}
  \DataTypeTok{row\_vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] gamma\_int;}
  \DataTypeTok{row\_vector}\NormalTok{[K] gamma\_slope;}
  \DataTypeTok{matrix}\NormalTok{[J, K] z;}
  \DataTypeTok{vector}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{}[K] tau;}
  \DataTypeTok{real}\NormalTok{\textless{}}\KeywordTok{lower}\NormalTok{=}\DecValTok{0}\NormalTok{\textgreater{} sigma;}
\NormalTok{\}}

\KeywordTok{transformed parameters}\NormalTok{ \{}
  \DataTypeTok{matrix}\NormalTok{[J, K] beta;}
  \DataTypeTok{matrix}\NormalTok{[}\DecValTok{2}\NormalTok{, K] gamma = append\_row(gamma\_int, gamma\_slope);}

  \CommentTok{// Corrected matrix multiplication and addition}
  \ControlFlowTok{for}\NormalTok{ (k }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:K) \{}
\NormalTok{    beta[, k] = u * gamma[, k] + tau[k] * z[, k];}
\NormalTok{  \}}
\NormalTok{\}}

\KeywordTok{model}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_mu;}
\NormalTok{  to\_vector(z) \textasciitilde{} std\_normal();}
\NormalTok{  to\_vector(gamma\_int) \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
\NormalTok{  to\_vector(gamma\_slope) \textasciitilde{} normal(}\DecValTok{0}\NormalTok{, }\FloatTok{1.25}\NormalTok{);}
\NormalTok{  to\_vector(tau) \textasciitilde{} cauchy(}\DecValTok{0}\NormalTok{, }\FloatTok{2.5}\NormalTok{);}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_mu[n] = beta[jj[n], }\DecValTok{1}\NormalTok{] + log1m\_exp({-}beta[jj[n], }\DecValTok{2}\NormalTok{] * pow(x[n], beta[jj[n], }\DecValTok{3}\NormalTok{]));}
\NormalTok{  \}}
\NormalTok{  log\_y \textasciitilde{} normal(log\_mu, sigma);}
\NormalTok{\}}

\KeywordTok{generated quantities}\NormalTok{ \{}
  \DataTypeTok{vector}\NormalTok{[N] log\_lik;}
  \ControlFlowTok{for}\NormalTok{ (n }\ControlFlowTok{in} \DecValTok{1}\NormalTok{:N) \{}
\NormalTok{    log\_lik[n] = normal\_lpdf(log\_y[n] |}
\NormalTok{      beta[jj[n], }\DecValTok{1}\NormalTok{] + log1m\_exp({-}beta[jj[n], }\DecValTok{2}\NormalTok{] * pow(x[n], beta[jj[n], }\DecValTok{3}\NormalTok{])),}
\NormalTok{      sigma);}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2025-03-26 R/figs.R

@@ -1,0 +1,3873 @@
+my_ggsave <- function(filename, plot, units = "mm", height = NA, width = NA, dpi = NA, ...) {
+  
+  # # Save the plot as .tiff
+  # ggsave(
+  #   filename = paste0(filename, ".tiff"),
+  #   plot = plot,
+  #   height = height,
+  #   width = width,
+  #   units = units,
+  #   dpi = dpi,
+  #   ...
+  # )
+  
+  # Save the plot as .pdf
+  ggsave(
+    filename = paste0(filename, ".pdf"),
+    plot = plot,
+    height = height,
+    width = width,
+    units = units,
+    dpi = dpi,
+    ...
+  )
+  
+    # Save the plot as .png
+  ggsave(
+    filename = paste0(filename, ".png"),
+    plot = plot,
+    height = height,
+    width = width,
+    units = units,
+    dpi = dpi,
+    ...
+  )
+  # Return the paths to the saved files
+  # return(paste0(filename, c(".tiff", ".pdf", ".png")))
+  return(paste0(filename, c(".pdf", ".png")))
+
+}
+
+
+my_theme <- function(){
+  theme_bw() %+replace%
+  theme(
+    plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"),
+    legend.key.size = unit(0.3, "cm"),
+    legend.spacing.y = unit(0.15, "cm"),
+    legend.text.align = 0,
+    legend.text = element_text(size = 8),
+    legend.title = element_text(size = 9),
+    legend.background = element_blank()
+  )
+}
+
+
+
+#' Plot Latitude and Longitude Data on a World Map
+#' Generates a map displaying locations based on latitude and longitude data. Points are overlaid on a world map to visualize their distribution.
+#' @param tallo_wd_df0 A data frame containing latitude and longitude columns.
+#' @return A ggplot object with the world map and plotted latitude and longitude points.
+#' @export
+
+data_map <- function(tallo_wd_df0) {
+  lat_long <- tallo_wd_df0 |> 
+    dplyr::select(latitude, longitude) |> 
+    na.omit()
+
+  world_map <- map_data("world")
+
+  p <- ggplot() +
[…]

### 2025-03-26 R/functions.R

@@ -1,0 +1,2672 @@
+#' @inheritParams readr::write_csv
+my_write_csv <- function(x, path, append = FALSE, col_names = !append) {
+    write_csv(x, path, append = FALSE, col_names = !append)
+    paste(path)
+}
+
+#=============================
+# CLEANING DATA
+#=============================f
+
+clean_tallo_try <- function(tallo_csv,  tallo_env_csv, try_wd_txt) {
+  tallo <- read_csv(tallo_csv)
+  tallo2 <- tallo |>
+    dplyr::select(tree_id,
+          division, family, genus, latitude, longitude, reference_id,
+          sp = species, dbh = stem_diameter_cm, h = height_m, cr = crown_radius_m,
+          height_outlier, crown_radius_outlier) |>
+          filter(height_outlier != 'Y' & crown_radius_outlier != 'Y' & !is.na(sp) &
+          sp != "unknown" & !is.na(dbh))
+
+  env_df <- read_csv(tallo_env_csv)
+
+  ai_df <- env_df |>
+    group_by(biome) |>
+    summarize(mean_aridity_index= mean(aridity_index, na.rm = TRUE))
+
+  tallo_env <- left_join(tallo2, env_df, by = c("tree_id" = "tree_id"))
+
+  try_data <- read_delim(try_wd_txt, delim = "\t") |>
+    dplyr::select(sp = AccSpeciesName, wd = StdValue, unit = UnitName) |>
+    filter(!is.na(sp) & sp != "unknown" & !is.na(wd) & !is.na(unit) & unit == "g/cm3")
+
+  # Calculate mean wood density for each species
+  mean_density <- try_data |>
+    group_by(sp) |>
+    summarize(wd = mean(as.numeric(wd), na.rm = TRUE))
+
+  # Join selected_data and wd_filtered
+  left_join(tallo_env, mean_density, by = "sp") |>
+    left_join(ai_df) |>
+    mutate(biome_sp = paste(biome, sp, sep = "_")) |>
+    filter(!is.na(wd)) |>
+    mutate(biome_div = paste(biome, division, sep = "_"))
+
+}
+
+
+#=============================
+# STAN DATA
+#=============================
+
+generate_stan_data <- function(tallo_wd_df_200, model = c("lr", "nlr"), div = c("ang", "gym"), variable = c("cr", "h", "dbh", "dbh1", "dbh2", "dbh3")) {
+
+   # Filter out rows with missing values in dbh, cr, h, and wd_s columns
+  if (variable == "cr") {
[…]

### 2025-03-26 _targets.R

@@ -1,0 +1,1741 @@
+# packages for this _targets.R file (not for the pipeline)
+library(targets)
+library(tidyverse)
+library(tarchetypes)
+library(furrr)
+library(here)
+library(ggplot2)
+library(future)
+library(cmdstanr)
+library(stantargets)
+library(clustermq)
+library(lme4)
+
+# Source the user-defined functions from functions.R and figs.R
+source("R/functions.R")
+source("R/figs.R")
+# parallel computing on local or on the same node
+plan(multicore)
+options(clustermq.scheduler = "multicore")
+
+# Set options for the targets pipeline, specifying required packages
+tar_option_set(
+  packages = c(
+    "tidyverse",
+    "lme4",
+    "janitor",
+    "yaml",
+    "e1071",
+    "car",
+    "sjPlot",
+    "gridExtra",
+    "loo",
+    "knitr",
+    "kableExtra",
+    "MASS",
+    "patchwork",
+    "cmdstanr",
+    "cowplot",
+    "MASS",
+    "stringr",
+    "grid",
+    "gsubfn",
+    "scales",
+    "maps",
+    "writexl",
+    "openxlsx"
+  )
+)
+
+# tar_option_set(
+#   garbage_collection = TRUE,
+#   memory = "transient"
+# )
+
+
+# mcmc (model) names -------------------------------------
+lr_values <- expand_grid(
+  lr = "lr",
+  model = c("pl", "pl_biome"),
+  div = c("ang", "gym"),
+  pred = c("cr", "h", "dbh", "dbh1", "dbh2", "dbh3")
+)
+
+nlr_values <- expand_grid(
+  lr = "nlr",
+  model = c("gmm", "weibull", "gmm_biome", "weibull_biome"),
+  div = c("ang", "gym"),
+  pred = c("cr", "h")
+)
+
+values <- bind_rows(lr_values, nlr_values)
+
+data_names <- values |>
+  mutate(data_names = str_c(model, div, pred, sep = "_")) |>
+  pull(data_names)
+
+mcmc_names <- values |>
+  mutate(mcmc_names = str_c("fit_", lr, "_nou_mcmc_",  data_names)) |>
+  filter(mcmc_names != "fit_nlr_nou_mcmc_gmm_biome_ang_h") |>
+  filter(mcmc_names != "fit_nlr_nou_mcmc_gmm_biome_ang_cr") |>
+  pull(mcmc_names)
+
+mcmc_names <- c(mcmc_names,
[…]

### 2025-03-26 docs/gmm_h.r

@@ -1,0 +1,128 @@
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(stringr)
+
+# Load MCMC results and Stan data
+gmm_h <- tar_read(fit_nlr3_mcmc_gmm_h)
+stan_nlr_data_h <- tar_read(stan_data_nlr_h)
+
+# Extract relevant data
+log_y <- stan_nlr_data_h$log_y
+x <- stan_nlr_data_h$x
+sp <- as.factor(stan_nlr_data_h$jj)
+
+# Extract beta parameters
+beta <- gmm_h$summary(variable = "beta")
+
+# Ensure dplyr and stringr are loaded and then extract the mean values of beta parameters
+log_a2 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,1\\]")) %>% pull(mean)
+b2 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,2\\]")) %>% pull(mean)
+k1 <- beta %>% filter(str_detect(variable, "beta\\[\\d+,3\\]")) %>% pull(mean)
+
+# Verify the lengths of the extracted parameters
+print(length(log_a2))
+print(length(b2))
+print(length(k1))
+print(length(sp))
+
+# Compute fitted values for H using the gMM formula
+fitted_log_y_h <- sapply(1:length(log_a2), function(i) {
+  log_a2[i] + b2[i] * log(x) - log(k1[i] + x^b2[i])
+})
+
+# Calculate community-level fitted line by averaging the species-level coefficients
+community_log_a2 <- mean(log_a2)
+community_b2 <- mean(b2)
+community_k1 <- mean(k1)
+community_fitted_log_y_h <- community_log_a2 + community_b2 * log(x) - log(community_k1 + x^community_b2)
+community_fitted_H <- exp(community_fitted_log_y_h)
+
+# Create a data frame for plotting
+data_h <- data.frame(H = exp(log_y), DBH = x, sp = sp)
+
+# Add species-specific fitted values to the data frame
+for (i in 1:length(log_a2)) {
+  data_h[[paste0("fitted_H_sp_", i)]] <- exp(fitted_log_y_h[, i])
+}
+
+# Plot observed H vs DBH with fitted lines for each species and community-level line
+gmm_h_plot <- ggplot(data_h, aes(x = DBH, y = H)) +
+  geom_point(alpha = 0.5, color = "grey") +
+  geom_line(aes(y = community_fitted_H), color = "white", linewidth = 1.2) +
[…]

### 2025-03-26 run.R

@@ -1,0 +1,9 @@
+# Uncomment to run targets sequentially on your local machine.
+#targets::tar_make()
+
+# Uncomment to run targets in parallel
+# on local processes or a Sun Grid Engine cluster.
+# targets::tar_make_clustermq(workers = parallel::detectCores())
+
+arg = commandArgs(T)
+targets::tar_make_clustermq(workers = as.numeric(arg[1]))

### 2025-03-26 run.sh

@@ -1,0 +1,42 @@
+#!/usr/bin/env bash
+set -e
+
+menu() {
+    echo "1) tar_make() on local (or inside docker)"
+    echo "2) tar_make() on Apptainer"
+    echo "3) Enter in the Apptainer container"
+    echo "4) Enter in the Singularity container on HPC"
+    read -rp "Enter number: " menu_num
+    case $menu_num in
+    1|2)
+        # Prompt for the number of CPUs after selecting option 1 or 2
+        read -rp "Enter the number of CPUs to use: " num_cpus
+        if [[ "$menu_num" == "1" ]]; then
+            Rscript R/run_script.R "$num_cpus"
+        elif [[ "$menu_num" == "2" ]]; then
+            apptainer exec \
+            --env RENV_PATHS_CACHE=$HOME/renv \
+            --env INSIDE_CONTAINER=true \
+            apptainer.sif Rscript R/run_script.R "$num_cpus" \
+            > log/script_output.log 2> log/script_error.log
+        fi
+        ;;
+    3)
+        apptainer shell \
+        --env RENV_PATHS_CACHE=$HOME/renv \
+        --env INSIDE_CONTAINER=true \
+        apptainer.sif bash
+        ;;
+    4)
+        singularity shell \
+        --env RENV_PATHS_CACHE=$HOME/renv \
+        --env INSIDE_CONTAINER=true \
+        apptainer.sif bash
+        ;;
+    *)
+        echo "Invalid option. Please type a number from 1-4."
+        ;;
+    esac
+}
+
+menu "$@"

### 2025-03-26 ms/ms.tex
clean

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
+% Set default figure placement to htbp
+\def\fps@figure{htbp}
 \makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
 \makeatletter
-\def\fps@figure{htbp}
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
 \makeatother
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{abstract}{%
-\section*{Abstract}\label{abstract}}
+\section*{Abstract}\label{abstract}
 \addcontentsline{toc}{section}{Abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 implementing climate change mitigation programs
-(\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}). The accuracy of
-carbon stock estimates depends on the availability of reliable
-allometric models to deduce the aboveground biomass (AGB) of trees from
-census data (\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}).
-Selecting an appropriate allometric model is, therefore, a critical step
-in estimating forest carbon biomass (e.g.,
-\protect\hyperlink{ref-Chave2014}{Chave et al. 2014};
-\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). Tree allometry
-describes the correlations between different biometric measurements of
-trees, such as diameter at breast height (DBH), tree height, and tree
-crown size (e.g., \protect\hyperlink{ref-Shinozaki1964a}{Shinozaki et
-al. 1964}; \protect\hyperlink{ref-Niklas1994}{Niklas 1994};
-\protect\hyperlink{ref-West1999}{West et al. 1999}). These relationships
-are widely utilized to estimate tree architecture from DBH and to
-estimate forest biomass and production, serving as fundamental
-components of dynamic vegetation models (e.g.,
-\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
-\protect\hyperlink{ref-Falster2017}{Falster et al. 2017};
-\protect\hyperlink{ref-Song2024}{Song et al. 2024}). Despite significant
+(\citeproc{ref-Chave2014}{Chave et al. 2014}). The accuracy of carbon
+stock estimates depends on the availability of reliable allometric
+models to deduce the aboveground biomass (AGB) of trees from census data
+(\citeproc{ref-Chave2014}{Chave et al. 2014}). Selecting an appropriate
+allometric model is, therefore, a critical step in estimating forest
+carbon biomass (e.g., \citeproc{ref-Chave2014}{Chave et al. 2014};
+\citeproc{ref-Molto2014}{Molto et al. 2014}). Tree allometry describes
+the correlations between different biometric measurements of trees, such
+as diameter at breast height (DBH), tree height, and tree crown size
+(e.g., \citeproc{ref-Shinozaki1964a}{Shinozaki et al. 1964};
+\citeproc{ref-Niklas1994}{Niklas 1994}; \citeproc{ref-West1999}{West et
+al. 1999}). These relationships are widely utilized to estimate tree
+architecture from DBH and to estimate forest biomass and production,
+serving as fundamental components of dynamic vegetation models (e.g.,
+\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-Falster2017}{Falster et al. 2017};
+\citeproc{ref-Song2024}{Song et al. 2024}). Despite significant
 progress, developing precise allometric models across regions
@@
 papers have used various functional forms for allometric equations
-(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
-\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
-\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
-\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
-remain the most common choice for describing the allometric scaling of
-tree size variables, including height and crown size with DBH (e.g.,
-\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
-relationships among tree size variables often involve complex
-interactions influenced by various factors, such as species diversity,
-environmental conditions, and individual variability (e.g.,
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
-\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}). Consequently,
-power-law models often fail to accurately capture the allometric
-relationships across the full range of tree sizes, leading to
+(e.g., \citeproc{ref-Thomas1996}{Thomas 1996};
+\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019};
+\citeproc{ref-Song2024}{Song et al. 2024}), power-law functions remain
+the most common choice for describing the allometric scaling of tree
+size variables, including height and crown size with DBH (e.g.,
+\citeproc{ref-Sileshi2023}{Sileshi et al. 2023};
+\citeproc{ref-Jucker2024}{Jucker et al. 2024}). The relationships among
+tree size variables often involve complex interactions influenced by
+various factors, such as species diversity, environmental conditions,
+and individual variability (e.g., \citeproc{ref-Poorter2012a}{Poorter et
+al. 2012a}; \citeproc{ref-Laurans2024}{Laurans et al. 2024}).
+Consequently, power-law models often fail to accurately capture the
+allometric relationships across the full range of tree sizes, leading to
 underestimation of dimensions for smaller trees and overestimation for
-larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
-2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
-highlights the need to adopt alternative functional forms---such as
-saturating relationships---that can more accurately reflect the
-complexity of allometric scaling in vegetation models (e.g.,
-\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
-\protect\hyperlink{ref-Song2024}{Song et al. 2024}).
+larger trees (e.g., \citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
+\citeproc{ref-Song2024}{Song et al. 2024}). This highlights the need to
+adopt alternative functional forms---such as saturating
+relationships---that can more accurately reflect the complexity of
+allometric scaling in vegetation models (e.g.,
+\citeproc{ref-Goodman2014}{Goodman et al. 2014};
+\citeproc{ref-Song2024}{Song et al. 2024}).
 
@@
 Detection and Ranging (LiDAR), have revolutionized the monitoring of
-forests (e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017})
-by enabling the identification and measurement of individual tree
-heights and crown dimensions using airborne imagery (e.g.,
-\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). However,
-because existing allometric models still rely on stem diameter as a
-crucial input for biomass estimation, this reliance presents a challenge
-for effectively integrating EO technologies into AGB estimation
-(\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
-ground-based LiDAR can measure DBH directly and is much more efficient
-than traditional ground surveys
-(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
-less common and convenient than aerial LiDAR. Alternative approaches
-focusing on tree height and crown dimensions as model predictors are
-also necessary for remote sensing applications
-(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
-solution is to estimate DBH from tree height and crown size and then use
+forests (e.g., \citeproc{ref-Jucker2017}{Jucker et al. 2017}) by
+enabling the identification and measurement of individual tree heights
+and crown dimensions using airborne imagery (e.g.,
+\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). However, because
+existing allometric models still rely on stem diameter as a crucial
+input for biomass estimation, this reliance presents a challenge for
+effectively integrating EO technologies into AGB estimation
+(\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). While ground-based LiDAR
+can measure DBH directly and is much more efficient than traditional
+ground surveys (\citeproc{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is
+still less common and convenient than aerial LiDAR. Alternative
+approaches focusing on tree height and crown dimensions as model
+predictors are also necessary for remote sensing applications
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}). One potential solution
+is to estimate DBH from tree height and crown size and then use
 established biomass equations, which could be an effective strategy for
 integrating airborne imagery into carbon monitoring programs
-(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
 
@@
 individual tree architecture and the composition of entire forest
-communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
-\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
-\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
-closed-canopy conditions, strong light competition makes the balance
-between vertical and horizontal space use crucial for species
-coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
-2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
-functional traits, wood density is closely linked to both plant life
-history strategies (\protect\hyperlink{ref-Fajardo2022a}{Fajardo 2022})
-and light demand (e.g., \protect\hyperlink{ref-King2006}{King et al.
-2006}; \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a}). Several
-studies suggest that wood density explains a trade-off between rapid
-stem expansion and broader crowns, shaping mechanical constraints on
-species with diverse architecture and light capture strategies
-(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
-\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
-\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
-types of studies have examined this relationship post hoc, relating
-species-specific allometric parameters to wood density and risking Type
-I or Type II errors if the hierarchical data structure is overlooked
-(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
-exception is Martínez Cano et al.
-(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
-density into a hierarchical model of tropical trees and concluded it
-offered minimal explanatory power for interspecific variation.
-Therefore, the role of wood density in global allometric scaling remains
-insufficiently explored.
+communities (e.g., \citeproc{ref-Iida2012}{Iida et al. 2012};
+\citeproc{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
+\citeproc{ref-Yang2023}{Yang and Swenson 2023}). Under closed-canopy
+conditions, strong light competition makes the balance between vertical
+and horizontal space use crucial for species coexistence (e.g.,
+\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}). Among functional traits,
+wood density is closely linked to both plant life history strategies
+(\citeproc{ref-Fajardo2022a}{Fajardo 2022}) and light demand (e.g.,
+\citeproc{ref-King2006}{King et al. 2006}; \citeproc{ref-Aiba2009}{Aiba
+and Nakashizuka 2009}; \citeproc{ref-Poorter2012a}{Poorter et al.
+2012a}). Several studies suggest that wood density explains a trade-off
+between rapid stem expansion and broader crowns, shaping mechanical
+constraints on species with diverse architecture and light capture
+strategies (\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}; \citeproc{ref-Yang2023}{Yang
+and Swenson 2023}). However, these types of studies have examined this
+relationship post hoc, relating species-specific allometric parameters
+to wood density and risking Type I or Type II errors if the hierarchical
+data structure is overlooked (\citeproc{ref-Gelman2006}{Gelman and Hill
+2006}). One exception is Martínez Cano et al.
+(\citeproc{ref-MartinezCano2019}{2019}), who incorporated wood density
+into a hierarchical model of tropical trees and concluded it offered
+minimal explanatory power for interspecific variation. Therefore, the
+role of wood density in global allometric scaling remains insufficiently
+explored.
 
@@
 radius (CR), and their interspecific variations. Using large datasets
-from the Tallo and TRY databases
-(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
-\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
+from the Tallo and TRY databases (\citeproc{ref-Kattge2020}{Kattge et
+al. 2020}; \citeproc{ref-Jucker2022}{Jucker et al. 2022}), which span a
 diverse array of tree forms across global forest types and climatic
@@
 
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
+\section{Materials and Methods}\label{materials-and-methods}
+
+\subsection{Data}\label{data}
 
@@
 standardized measurements of DBH, height, and crown radius for 5,163
-species across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker
-et al. 2022}, https://doi.org/10.5281/zenodo.6637599) along with wood
-density data for 10,995 species from TRY version 6
-(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
+species across various biomes (\citeproc{ref-Jucker2022}{Jucker et al.
+2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
+data for 10,995 species from TRY version 6
+(\citeproc{ref-Kattge2020}{Kattge et al. 2020},
 https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
 further details). We thoroughly processed the Tallo dataset by removing
-998 outliers following Jucker et al.
-(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
-selected only species with standardized wood density values in g
-cm\textsuperscript{-3} from TRY and removed any records with missing
-data. Subsequently, we created an overlapping dataset between Tallo and
-TRY, resulting in a combined dataset encompassing allometric and wood
-density information for 355,881 individuals across 2,961 species.
+998 outliers following Jucker et al. (\citeproc{ref-Jucker2022}{2022}).
+To ensure consistency, we selected only species with standardized wood
+density values in g cm\textsuperscript{-3} from TRY and removed any
+records with missing data. Subsequently, we created an overlapping
+dataset between Tallo and TRY, resulting in a combined dataset
+encompassing allometric and wood density information for 355,881
+individuals across 2,961 species.
 
@@
 minimize bias due to data imbalance and prevent the overrepresentation
-of any single abundant species
-(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}), thereby
-ensuring that the model remains broadly applicable. Since not all trees
-have a complete set of information, we created specific sub-datasets to
-select the best predictive model for each allometric relationship for
-two clades (i.e., angiosperms and gymnosperms; see below and
-Supplementary Table S2 for more details).
-
-\hypertarget{tree-allometry-models}{%
-\subsection{Tree allometry models}\label{tree-allometry-models}}
+of any single abundant species (\citeproc{ref-Gelman2006}{Gelman and
+Hill 2006}), thereby ensuring that the model remains broadly applicable.
+Since not all trees have a complete set of information, we created
+specific sub-datasets to select the best predictive model for each
+allometric relationship for two clades (i.e., angiosperms and
+gymnosperms; see below and Supplementary Table S2 for more details).
+
+\subsection{Tree allometry models}\label{tree-allometry-models}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-like}{}{
+\begin{equation}\phantomsection\label{eq-like}{
 \log(y_{i,j}) \sim N(f(X_{i,j}, \pmb{\beta}_{j}), \sigma_y),
-}\label{eq-like}\end{equation}
+}\end{equation}
 
@@
 for the relationship between tree allometric variables
-(\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018};
-\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}):
-
-Power-law: \begin{equation}\protect\hypertarget{eq-pl}{}{
+(\citeproc{ref-Sullivan2018}{Sullivan et al. 2018};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019}):
+
+Power-law: \begin{equation}\phantomsection\label{eq-pl}{
 y = aX^b,
-}\label{eq-pl}\end{equation}
+}\end{equation}
 
 Generalized Michaelis-Menten (gMM):
-\begin{equation}\protect\hypertarget{eq-gMM}{}{
+\begin{equation}\phantomsection\label{eq-gMM}{
 y = \frac{aX^b}{k + X^b},
-}\label{eq-gMM}\end{equation}
-
-Weibull: \begin{equation}\protect\hypertarget{eq-wb}{}{
+}\end{equation}
+
+Weibull: \begin{equation}\phantomsection\label{eq-wb}{
 y = a\left\{1-\mathrm{exp}(-bX^k)\right\},
-}\label{eq-wb}\end{equation}
+}\end{equation}
 
@@
 commonly used functions to describe biological scaling relationships
-(e.g., \protect\hyperlink{ref-West1999}{West et al. 1999};
-\protect\hyperlink{ref-Brown2004}{Brown et al. 2004}). We also tested
-two nonlinear models, the generalized Michaelis--Menten (gMM)
+(e.g., \citeproc{ref-West1999}{West et al. 1999};
+\citeproc{ref-Brown2004}{Brown et al. 2004}). We also tested two
+nonlinear models, the generalized Michaelis--Menten (gMM)
 (Eq.~\ref{eq-gMM}) and the Weibull function (Eq.~\ref{eq-wb}), to
 account for a saturating relationship between tree dimensions and DBH
-(e.g., \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
-2019}).
+(e.g., \citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019}).
 
@@
 cross-validation (PSIS-LOO; Vehtari et al.
-(\protect\hyperlink{ref-Vehtari2017}{2017})). PSIS-LOO is an accurate
-and reliable approximation to standard leave-one-out cross-validation
-(LOO), which is a robust method for comparing models with different
-numbers of parameters (\protect\hyperlink{ref-Vehtari2017}{Vehtari et
-al. 2017}). We utilized PSIS-LOO to calculate the LOO Information
-Criterion (LOOIC) for each model. A model with lower LOOIC is considered
-to have better predictive accuracy
-(\protect\hyperlink{ref-Vehtari2017}{Vehtari et al. 2017}).
+(\citeproc{ref-Vehtari2017}{2017})). PSIS-LOO is an accurate and
+reliable approximation to standard leave-one-out cross-validation (LOO),
+which is a robust method for comparing models with different numbers of
+parameters (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}). We
+utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for
+each model. A model with lower LOOIC is considered to have better
+predictive accuracy (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}).
 
@@
 Species-specific parameters (\(\beta_j\)) were modeled as a function of
-wood density (\(u_j\)): \begin{equation}\protect\hypertarget{eq-wood}{}{
+wood density (\(u_j\)): \begin{equation}\phantomsection\label{eq-wood}{
 \beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
-}\label{eq-wood}\end{equation}
+}\end{equation}
 
@@
 Carlo (HMC), algorithm implemented in Stan
-(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}) using
-weakly-informative priors (\protect\hyperlink{ref-Gelman2008}{Gelman et
-al. 2008}). The HMC algorithm uses gradient information to propose new
+(\citeproc{ref-Carpenter2017}{Carpenter et al. 2017}) using
+weakly-informative priors (\citeproc{ref-Gelman2008}{Gelman et al.
+2008}). The HMC algorithm uses gradient information to propose new
 states in the Markov chain, leading to a more efficient exploration of
@@
 methods that rely on random proposals
-(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}). This
-efficiency allows us to achieve convergence with fewer iterations than
-traditional MCMC methods. Inferences were based on 1,000 posterior
-samples following 1,000 warmup iterations for three parallel chains.
-Convergence was assessed using the Gelman--Rubin statistic with a
-convergence threshold of 1.1 (\protect\hyperlink{ref-Gelman2013}{Gelman
-et al. 2015}) and ensuring effective sample sizes greater than 300
-(\protect\hyperlink{ref-Vehtari2021}{Vehtari et al. 2021}). We also
-reported the median and 95\% Bayesian credible intervals (CIs) for each
-parameter (Supplementary Tables S3-S8). Estimates were considered
-significant when 95\% CIs did not overlap with zero.
+(\citeproc{ref-Carpenter2017}{Carpenter et al. 2017}). This efficiency
+allows us to achieve convergence with fewer iterations than traditional
+MCMC methods. Inferences were based on 1,000 posterior samples following
+1,000 warmup iterations for three parallel chains. Convergence was
+assessed using the Gelman--Rubin statistic with a convergence threshold
+of 1.1 (\citeproc{ref-Gelman2013}{Gelman et al. 2015}) and ensuring
+effective sample sizes greater than 300
+(\citeproc{ref-Vehtari2021}{Vehtari et al. 2021}). We also reported the
+median and 95\% Bayesian credible intervals (CIs) for each parameter
+(Supplementary Tables S3-S8). Estimates were considered significant when
+95\% CIs did not overlap with zero.
 
 Statistical analyses were performed in R version 4.3.2
-(\protect\hyperlink{ref-RCoreTeam2023}{R Core Team 2023}), using the R
-packages \emph{targets} version 1.7.0 for workflow management
-(\protect\hyperlink{ref-Landau2021}{Landau 2021}). Codes are made
-available on GitHub at
-\url{https://github.com/duyenecology/tree-allometry} (Stan code is
-provided in Supplementary Section S2).
-
-\hypertarget{aboveground-biomass-estimates}{%
+(\citeproc{ref-RCoreTeam2023}{R Core Team 2023}), using the R packages
+\emph{targets} version 1.7.0 for workflow management
+(\citeproc{ref-Landau2021}{Landau 2021}). Codes are made available on
+GitHub at \url{https://github.com/duyenecology/tree-allometry} (Stan
+code is provided in Supplementary Section S2).
+
 \subsection{Aboveground biomass
-estimates}\label{aboveground-biomass-estimates}}
+estimates}\label{aboveground-biomass-estimates}
 
 For angiosperm species, we used the widely applied equation from Chave
-et al. (\protect\hyperlink{ref-Chave2014}{2014}) to estimate aboveground
-biomass (AGB; kg of dry mass):
-
-\begin{equation}\protect\hypertarget{eq-agb}{}{
+et al. (\citeproc{ref-Chave2014}{2014}) to estimate aboveground biomass
+(AGB; kg of dry mass):
+
+\begin{equation}\phantomsection\label{eq-agb}{
 \mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
-}\label{eq-agb}\end{equation}
+}\end{equation}
 
@@
 carbon stocks and were excluded from the calibration of Chave et al.
-(\protect\hyperlink{ref-Chave2014}{2014})'s equation. Additionally, the
+(\citeproc{ref-Chave2014}{2014})'s equation. Additionally, the
 sub-datasets used to select the optimal predictive model for each
@@
 equations are species-specific and do not incorporate wood density
-(e.g., \protect\hyperlink{ref-Moore2010}{Moore 2010};
-\protect\hyperlink{ref-Bazrgar2024}{Bazrgar et al. 2024}).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{tree-height-allometry}{%
-\subsection{Tree height allometry}\label{tree-height-allometry}}
+(e.g., \citeproc{ref-Moore2010}{Moore 2010};
+\citeproc{ref-Bazrgar2024}{Bazrgar et al. 2024}).
+
+\section{Results}\label{results}
+
+\subsection{Tree height allometry}\label{tree-height-allometry}
 
@@
 dataset, the best model to predict H from DBH of each clade was:
-\begin{equation}\protect\hypertarget{eq-h_ang}{}{
+\begin{equation}\phantomsection\label{eq-h_ang}{
 \mathrm{H_{ang}} =50.0
@@
 \right) \right\}
-}\label{eq-h_ang}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-h_gym}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-h_gym}{
 \mathrm{H_{gym}} = 117.4
@@
 \right) \right\}.
-}\label{eq-h_gym}\end{equation}
+}\end{equation}
 
@@
 \(\Delta\)LOOIC greater than 4 suggests a meaningful difference in
-predictive performance (\protect\hyperlink{ref-Cochrane2023}{Cochrane et
-al. 2023}; \protect\hyperlink{ref-Sivula2023}{Sivula et al. 2023}).
-Lower \(\Delta\)LOOIC values indicate models that perform closer to the
-best model in terms of predictive accuracy, while higher ELPD values
-reflect better model fit. Higher \(R^2\) values reflect better
-explanatory power, indicating how well the model explains the variance
-in the dependent variable.
+predictive performance (\citeproc{ref-Cochrane2023}{Cochrane et al.
+2023}; \citeproc{ref-Sivula2023}{Sivula et al. 2023}). Lower
+\(\Delta\)LOOIC values indicate models that perform closer to the best
+model in terms of predictive accuracy, while higher ELPD values reflect
+better model fit. Higher \(R^2\) values reflect better explanatory
+power, indicating how well the model explains the variance in the
+dependent variable.
 
@@
 
-{\centering \includegraphics{../figs/h_cr_dbh_non_log.png}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{../figs/h_cr_dbh_non_log.png}}
 
@@
 
-\end{figure}
-
-\hypertarget{crown-radius-allometry}{%
-\subsection{Crown Radius allometry}\label{crown-radius-allometry}}
+\end{figure}%
+
+\subsection{Crown Radius allometry}\label{crown-radius-allometry}
 
@@
 1). Across the global dataset, the best model for predicting CR from DBH
-of each clade was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
+of each clade was: \begin{equation}\phantomsection\label{eq-cr_ang}{
 \mathrm{CR_{ang}} = \mathrm{0.349}
 \, \mathrm{DBH}^{\mathrm{0.626}}
-}\label{eq-cr_ang}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-cr_gym}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-cr_gym}{
 \mathrm{CR_{gym}} = \frac{\mathrm{769.4}
@@
 + \mathrm{DBH}^{\mathrm{0.631}}}.
-}\label{eq-cr_gym}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{dbh-allometry}{%
-\subsection{DBH allometry}\label{dbh-allometry}}
+\subsection{DBH allometry}\label{dbh-allometry}
 
@@
 DBH from H and CR for each clade was:
-\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
+\begin{equation}\phantomsection\label{eq-dbh_ang}{
 \mathrm{DBH_{ang}} = \mathrm{0.948}
@@
 \mathrm{H}^{\text{0.590}}
-}\label{eq-dbh_ang}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-dbh_gym}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-dbh_gym}{
 \mathrm{DBH_{gym}} = \mathrm{0.972}
@@
 \mathrm{H}^{\text{0.588}}.
-}\label{eq-dbh_gym}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{wood-density-and-tree-architectural-diversity}{%
 \subsection{Wood density and tree architectural
-diversity}\label{wood-density-and-tree-architectural-diversity}}
+diversity}\label{wood-density-and-tree-architectural-diversity}
 
@@
 
-{\centering \includegraphics{../figs/wd_com.png}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{../figs/wd_com.png}}
 
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-{\centering \includegraphics{../figs/wd_ef.png}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{../figs/wd_ef.png}}
 
@@
 
-\end{figure}
-
-\hypertarget{estimating-aboveground-biomass}{%
+\end{figure}%
+
 \subsection{Estimating aboveground
-biomass}\label{estimating-aboveground-biomass}}
+biomass}\label{estimating-aboveground-biomass}
 
@@
 
-{\centering \includegraphics{../figs/agb_ang.png}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{../figs/agb_ang.png}}
 
@@
 
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\end{figure}%
+
+\section{Discussion}\label{discussion}
 
@@
 
-\hypertarget{tree-allometry}{%
-\subsection{Tree allometry}\label{tree-allometry}}
+\subsection{Tree allometry}\label{tree-allometry}
 
@@
 and gymnosperm species, consistent with theoretical expectations
-(\protect\hyperlink{ref-Falster2003}{Falster and Westoby 2003};
-\protect\hyperlink{ref-Niklas2007}{Niklas 2007}) and empirical
-observations (e.g., \protect\hyperlink{ref-Feldpausch2011}{Feldpausch et
-al. 2011}; \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
-2019}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). A recent
-study using Tallo suggested the Weibull function as the best model for
-six plant functional types (PFTs) (\protect\hyperlink{ref-Song2024}{Song
-et al. 2024}). While PFTs-based approaches can illuminate large-scale
-patterns, they may obscure species-specific variation in allometric
-scaling, and locally abundant species may bias the overall estimates.
-Our best predictive model effectively captures species-specific patterns
+(\citeproc{ref-Falster2003}{Falster and Westoby 2003};
+\citeproc{ref-Niklas2007}{Niklas 2007}) and empirical observations
+(e.g., \citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019};
+\citeproc{ref-Song2024}{Song et al. 2024}). A recent study using Tallo
+suggested the Weibull function as the best model for six plant
+functional types (PFTs) (\citeproc{ref-Song2024}{Song et al. 2024}).
+While PFTs-based approaches can illuminate large-scale patterns, they
+may obscure species-specific variation in allometric scaling, and
+locally abundant species may bias the overall estimates. Our best
+predictive model effectively captures species-specific patterns
 (Supplementary Table S6) despite underestimating the height of
@@
 \(CR \propto DBH^{2/3}\)), based on assumptions regarding mechanical and
-hydraulic constraints to plant growth
-(\protect\hyperlink{ref-West1999}{West et al. 1999}). However, our
-findings revealed that the power-law model exponent of angiosperms
-(\emph{b} = 0.538) was significantly smaller, while the actual exponent
-in gymnosperms (\emph{b} = 0.704) was larger than the theoretical value
-of 2/3 (Supplementary Table S3). This divergence highlights the
-importance of separately modeling the allometric relationships of each
-functional group to accurately capture their distinct scaling patterns.
-Moreover, the power-law function performed worse than the saturating
-functions (Table 1), indicating that the actual H-DBH relationship does
-not follow the simple scaling laws proposed by MST.
+hydraulic constraints to plant growth (\citeproc{ref-West1999}{West et
+al. 1999}). However, our findings revealed that the power-law model
+exponent of angiosperms (\emph{b} = 0.538) was significantly smaller,
+while the actual exponent in gymnosperms (\emph{b} = 0.704) was larger
+than the theoretical value of 2/3 (Supplementary Table S3). This
+divergence highlights the importance of separately modeling the
+allometric relationships of each functional group to accurately capture
+their distinct scaling patterns. Moreover, the power-law function
+performed worse than the saturating functions (Table 1), indicating that
+the actual H-DBH relationship does not follow the simple scaling laws
+proposed by MST.
 
@@
 plastic crown expansion in competitive environments
-(\protect\hyperlink{ref-Lines2012}{Lines et al. 2012}; e.g.,
-\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}), a conclusion
+(\citeproc{ref-Lines2012}{Lines et al. 2012}; e.g.,
+\citeproc{ref-Poorter2012a}{Poorter et al. 2012a};
+\citeproc{ref-Hulshof2015}{Hulshof et al. 2015}), a conclusion
 consistent with a study in tropical moist forests by Martínez Cano et
-al. (\protect\hyperlink{ref-MartinezCano2019}{2019}). In addition, the
-estimated exponent aligns closely with MST predictions for crown radius,
+al. (\citeproc{ref-MartinezCano2019}{2019}). In addition, the estimated
+exponent aligns closely with MST predictions for crown radius,
 supporting the selection of the power-law model for angiosperms, where
@@
 forms that favour vertical over lateral expansion (e.g.,
-\protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), an adaptation
-that can mitigate risks such as branch breakage under heavy snow loads
-(e.g., \protect\hyperlink{ref-Mikola1938}{Mikola 1938};
-\protect\hyperlink{ref-Loehle2016}{Loehle 2016};
-\protect\hyperlink{ref-Tavankar2019}{Tavankar et al. 2019}).
+\citeproc{ref-Lines2012}{Lines et al. 2012};
+\citeproc{ref-Hulshof2015}{Hulshof et al. 2015};
+\citeproc{ref-Jucker2024}{Jucker et al. 2024}), an adaptation that can
+mitigate risks such as branch breakage under heavy snow loads (e.g.,
+\citeproc{ref-Mikola1938}{Mikola 1938}; \citeproc{ref-Loehle2016}{Loehle
+2016}; \citeproc{ref-Tavankar2019}{Tavankar et al. 2019}).
 
@@
 (EO) platforms, such as remote sensing and LiDAR
-(\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Our analysis
-showed that accurately predicting DBH requires both tree height and
-crown radius (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}),
-which is especially important for distinguishing among trees of similar
-height but varying DBH (\protect\hyperlink{ref-King2005}{King 2005};
-\protect\hyperlink{ref-King2011}{King and Clark 2011}). Among the tested
-models, the \(CR,H\) model
+(\citeproc{ref-Filipescu2012}{Filipescu et al. 2012};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). Our analysis showed that
+accurately predicting DBH requires both tree height and crown radius
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}), which is especially
+important for distinguishing among trees of similar height but varying
+DBH (\citeproc{ref-King2005}{King 2005}; \citeproc{ref-King2011}{King
+and Clark 2011}). Among the tested models, the \(CR,H\) model
 (\(\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}\))
 provided the highest predictive accuracy, a result that contrasts with
-the previous study (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
-2017}), which concluded that the \(CR \times H\) model
+the previous study (\citeproc{ref-Jucker2017}{Jucker et al. 2017}),
+which concluded that the \(CR \times H\) model
 (\(\mathrm{DBH} = a_2(\mathrm{CR} \times \mathrm{H})^{b_2}\)) was the
@@
 
-\hypertarget{wood-density-and-tree-architecture-diversity}{%
 \subsection{Wood density and tree architecture
-diversity}\label{wood-density-and-tree-architecture-diversity}}
+diversity}\label{wood-density-and-tree-architecture-diversity}
 
@@
 of wood density (Fig.~\ref{fig-wood1} d-f). These results are consistent
-with the previous study (\protect\hyperlink{ref-Poorter2012b}{Poorter et
-al. 2012b}), which showed that among seven angiosperm species, those
-with lighter wood exhibited taller maximum heights, while no significant
+with the previous study (\citeproc{ref-Poorter2012b}{Poorter et al.
+2012b}), which showed that among seven angiosperm species, those with
+lighter wood exhibited taller maximum heights, while no significant
 correlations were observed for six gymnosperm species. A recent study
@@
 nonsignificant relationship between tree height and wood density
-(\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). Their analysis
-included functional groups (angiosperms vs.~gymnosperms) as a fixed
-effect, providing valuable insights into group-level patterns. However,
-it did not incorporate interactions between wood density and functional
-groups, and thus, the effects of wood density were modeled as uniform
-across angiosperms and gymnosperms. In contrast, our approach explicitly
+(\citeproc{ref-Jucker2024}{Jucker et al. 2024}). Their analysis included
+functional groups (angiosperms vs.~gymnosperms) as a fixed effect,
+providing valuable insights into group-level patterns. However, it did
+not incorporate interactions between wood density and functional groups,
+and thus, the effects of wood density were modeled as uniform across
+angiosperms and gymnosperms. In contrast, our approach explicitly
 modeled these differences, enabling us to capture functional
@@
 findings are consistent with previous work in temperate forests
-(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009}), tropical
-forests (\protect\hyperlink{ref-Iida2012}{Iida et al. 2012}),
-subtropical forests (\protect\hyperlink{ref-Yang2023}{Yang and Swenson
-2023}), and on a global scale (\protect\hyperlink{ref-Jucker2024}{Jucker
-et al. 2024}), all of which similarly report that denser wood promotes
-more efficient horizontal crown expansion.
+(\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009}), tropical forests
+(\citeproc{ref-Iida2012}{Iida et al. 2012}), subtropical forests
+(\citeproc{ref-Yang2023}{Yang and Swenson 2023}), and on a global scale
+(\citeproc{ref-Jucker2024}{Jucker et al. 2024}), all of which similarly
+report that denser wood promotes more efficient horizontal crown
+expansion.
 
@@
 saplings in an Australian subtropical forest
-(\protect\hyperlink{ref-Kooyman2009}{Kooyman and Westoby 2009}) but
-contradicts the expectation that species with lower wood density develop
-thicker stems to ensure structural stability, as suggested by prior
-studies (\protect\hyperlink{ref-King2006}{King et al. 2006};
-\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
-\protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Although wood
-density is often regarded as a key functional trait (e.g.,
-\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), variations in
-tree allometry across individuals are likely influenced by a combination
-of species identity and environmental factors. Across species or along
+(\citeproc{ref-Kooyman2009}{Kooyman and Westoby 2009}) but contradicts
+the expectation that species with lower wood density develop thicker
+stems to ensure structural stability, as suggested by prior studies
+(\citeproc{ref-King2006}{King et al. 2006};
+\citeproc{ref-Anten2009}{Anten and Schieving 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}). Although wood density is
+often regarded as a key functional trait (e.g.,
+\citeproc{ref-Anten2009}{Anten and Schieving 2009};
+\citeproc{ref-Jucker2024}{Jucker et al. 2024}), variations in tree
+allometry across individuals are likely influenced by a combination of
+species identity and environmental factors. Across species or along
 broad environmental gradients, wood density systematically
 varies---tending to be denser in drier, colder climates and in
-nutrient-poor soils
-(\protect\hyperlink{ref-Muller-Landau2004}{Muller-Landau 2004};
-\protect\hyperlink{ref-Chave2009}{Chave et al. 2009};
-\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022};
-\protect\hyperlink{ref-Mo2024}{Mo et al. 2024})---reflecting species'
-adaptive strategies. Such environmental influences might confound simple
+nutrient-poor soils (\citeproc{ref-Muller-Landau2004}{Muller-Landau
+2004}; \citeproc{ref-Chave2009}{Chave et al. 2009};
+\citeproc{ref-Fajardo2022b}{Fajardo et al. 2022};
+\citeproc{ref-Mo2024}{Mo et al. 2024})---reflecting species' adaptive
+strategies. Such environmental influences might confound simple
 correlations between wood density and DBH when examining diverse tree
@@
 architecture across different environments
-(\protect\hyperlink{ref-Fajardo2022b}{Fajardo et al. 2022}).
-Furthermore, substantial measurement errors in tree height and crown
-radius relative to DBH may introduce noise that obscures potential
-relationships between wood density and stem diameter. These highlight
-the importance of adopting a more integrative framework that accounts
-for additional functional traits and environmental variables to deepen
-our understanding of the factors driving DBH growth.
+(\citeproc{ref-Fajardo2022b}{Fajardo et al. 2022}). Furthermore,
+substantial measurement errors in tree height and crown radius relative
+to DBH may introduce noise that obscures potential relationships between
+wood density and stem diameter. These highlight the importance of
+adopting a more integrative framework that accounts for additional
+functional traits and environmental variables to deepen our
+understanding of the factors driving DBH growth.
 
@@
 mechanical stability of stems and branches
-(\protect\hyperlink{ref-Sterck1998}{Sterck and Bongers 1998}),
-ultimately determining the architecture of trees (e.g.,
-\protect\hyperlink{ref-Poorter2006}{Poorter et al. 2006}). Angiosperms
-have denser wood and a wider wood density range than gymnosperms (see
+(\citeproc{ref-Sterck1998}{Sterck and Bongers 1998}), ultimately
+determining the architecture of trees (e.g.,
+\citeproc{ref-Poorter2006}{Poorter et al. 2006}). Angiosperms have
+denser wood and a wider wood density range than gymnosperms (see
 Supplementary Table S2), resulting in greater and more flexible
-mechanical support (e.g., \protect\hyperlink{ref-Gelder2006}{Gelder et
-al. 2006}; \protect\hyperlink{ref-Swenson2007}{Swenson and Enquist
-2007}; \protect\hyperlink{ref-Mo2024}{Mo et al. 2024}). This allows them
-to develop a more diverse height and crown to optimize light capture,
-particularly in environments where light competition is intense, such as
-tropical regions. In contrast, gymnosperms predominantly live and often
-dominate in extreme environments such as boreal and high-elevation
-regions (\protect\hyperlink{ref-Swenson2007}{Swenson and Enquist 2007};
-\protect\hyperlink{ref-Mo2024}{Mo et al. 2024}), prioritising survival
-and longevity (\protect\hyperlink{ref-Fossdal2024}{Fossdal et al.
-2024}). Due to their simpler and more homogeneous wood anatomy, these
-species display lighter wood and a narrower range of wood density (e.g.,
-\protect\hyperlink{ref-Mo2024}{Mo et al. 2024}), which may result in
-less flexibility in the crown architecture. These differences underscore
-how the contrasting evolutionary strategies of angiosperms and
-gymnosperms shape their architectural diversity, reflecting adaptations
-to distinct ecological niches.
-
-\hypertarget{consequences-for-agb-estimates}{%
+mechanical support (e.g., \citeproc{ref-Gelder2006}{Gelder et al. 2006};
+\citeproc{ref-Swenson2007}{Swenson and Enquist 2007};
+\citeproc{ref-Mo2024}{Mo et al. 2024}). This allows them to develop a
+more diverse height and crown to optimize light capture, particularly in
+environments where light competition is intense, such as tropical
+regions. In contrast, gymnosperms predominantly live and often dominate
+in extreme environments such as boreal and high-elevation regions
+(\citeproc{ref-Swenson2007}{Swenson and Enquist 2007};
+\citeproc{ref-Mo2024}{Mo et al. 2024}), prioritising survival and
+longevity (\citeproc{ref-Fossdal2024}{Fossdal et al. 2024}). Due to
+their simpler and more homogeneous wood anatomy, these species display
+lighter wood and a narrower range of wood density (e.g.,
+\citeproc{ref-Mo2024}{Mo et al. 2024}), which may result in less
+flexibility in the crown architecture. These differences underscore how
+the contrasting evolutionary strategies of angiosperms and gymnosperms
+shape their architectural diversity, reflecting adaptations to distinct
+ecological niches.
+
 \subsection{Consequences for AGB
-estimates}\label{consequences-for-agb-estimates}}
+estimates}\label{consequences-for-agb-estimates}
 
@@
 to estimate AGB are small to medium-sized (DBH \textless{} 70 cm
-(\protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
-\protect\hyperlink{ref-Murphy2022}{Murphy and Bradford 2022})), whose
-H-DBH relationships can still be well captured by the power-law function
-(Fig.~\ref{fig-compare} a) (e.g.,
-\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). In contrast,
-individuals with DBH ≥ 70 cm, while typically forming a small proportion
-of the tree population, contribute significantly to biomass storage
-(e.g., \protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015}). For
-instance, Slik et al. (\protect\hyperlink{ref-Slik2013}{2013}) reported
-that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the dataset in
-tropical forests, yet stored 25-45\% of the total AGB on average.
-However, the fundamental assumption of the power-law function does not
-fully reflect the asymptotic nature of height-diameter allometries in
-large trees (Fig.~\ref{fig-compare} a), potentially leading to
-significant inaccuracies in stand-level biomass estimates (e.g.,
-\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
-\protect\hyperlink{ref-Fayolle2016}{Fayolle et al. 2016};
-\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018};
-\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}).
-Because large trees contribute disproportionately to overall biomass
-(e.g., \protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
-\protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015};
-\protect\hyperlink{ref-Ligot2018}{Ligot et al. 2018}), if direct height
+(\citeproc{ref-Slik2013}{Slik et al. 2013};
+\citeproc{ref-Murphy2022}{Murphy and Bradford 2022})), whose H-DBH
+relationships can still be well captured by the power-law function
+(Fig.~\ref{fig-compare} a) (e.g., \citeproc{ref-Molto2014}{Molto et al.
+2014}). In contrast, individuals with DBH ≥ 70 cm, while typically
+forming a small proportion of the tree population, contribute
+significantly to biomass storage (e.g., \citeproc{ref-Fauset2015}{Fauset
+et al. 2015}). For instance, Slik et al. (\citeproc{ref-Slik2013}{2013})
+reported that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the
+dataset in tropical forests, yet stored 25-45\% of the total AGB on
+average. However, the fundamental assumption of the power-law function
+does not fully reflect the asymptotic nature of height-diameter
+allometries in large trees (Fig.~\ref{fig-compare} a), potentially
+leading to significant inaccuracies in stand-level biomass estimates
+(e.g., \citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
+\citeproc{ref-Sullivan2018}{Sullivan et al. 2018};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019}). Because
+large trees contribute disproportionately to overall biomass (e.g.,
+\citeproc{ref-Slik2013}{Slik et al. 2013};
+\citeproc{ref-Fauset2015}{Fauset et al. 2015};
+\citeproc{ref-Ligot2018}{Ligot et al. 2018}), if direct height
 measurements are unavailable, height estimates should be derived by
@@
 the saturating component is accurately captured
-(\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018}).
+(\citeproc{ref-Sullivan2018}{Sullivan et al. 2018}).
 
@@
 leading to inaccurate DBH estimates (e.g.,
-\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012}).
-Additionally, most existing AGB estimation equations typically include
-DBH as a squared term in equations (e.g.,
-\protect\hyperlink{ref-Brown1989}{Brown et al. 1989};
-\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}), meaning that even
-minor errors in DBH predictions can significantly affect the accuracy of
-AGB estimates (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
+\citeproc{ref-Filipescu2012}{Filipescu et al. 2012}). Additionally, most
+existing AGB estimation equations typically include DBH as a squared
+term in equations (e.g., \citeproc{ref-Brown1989}{Brown et al. 1989};
+\citeproc{ref-Chave2014}{Chave et al. 2014}), meaning that even minor
+errors in DBH predictions can significantly affect the accuracy of AGB
+estimates (\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
 Interestingly, our results revealed that, although the \(CR,H\) model
@@
 height (\emph{r} = 0.703), offering a more biologically realistic
-framework (e.g., \protect\hyperlink{ref-Dormann2013}{Dormann et al.
-2013}). The observed importance of crown radius in improving AGB
-estimation aligns with prior studies (e.g.,
-\protect\hyperlink{ref-Henry2010}{Henry et al. 2010};
-\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014}), whereas our
-finding of strong relationships between AGB and crown dimensions is
-consistent with those reported by Jucker et al.
-(\protect\hyperlink{ref-Jucker2017}{2017}). While the interdependence of
-tree height and crown radius requires careful model refinement to
-address potential collinearity issues
-(\protect\hyperlink{ref-Dormann2013}{Dormann et al. 2013}), their strong
+framework (e.g., \citeproc{ref-Dormann2013}{Dormann et al. 2013}). The
+observed importance of crown radius in improving AGB estimation aligns
+with prior studies (e.g., \citeproc{ref-Henry2010}{Henry et al. 2010};
+\citeproc{ref-Goodman2014}{Goodman et al. 2014}), whereas our finding of
+strong relationships between AGB and crown dimensions is consistent with
+those reported by Jucker et al. (\citeproc{ref-Jucker2017}{2017}). While
+the interdependence of tree height and crown radius requires careful
+model refinement to address potential collinearity issues
+(\citeproc{ref-Dormann2013}{Dormann et al. 2013}), their strong
 relationship with AGB suggests the feasibility of directly estimating
 AGB from EO data, bypassing DBH prediction and simplifying the
-estimation process (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
-2017}).
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+estimation process (\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
+
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
+\section*{Acknowledgments}\label{acknowledgments}
 \addcontentsline{toc}{section}{Acknowledgments}
@@
 
-\hypertarget{funding}{%
-\section*{Funding}\label{funding}}
+\section*{Funding}\label{funding}
 \addcontentsline{toc}{section}{Funding}
@@
 
+\section*{Data availability}\label{data-availability}
+\addcontentsline{toc}{section}{Data availability}
+
+Tallo and TRY data are available from their respective platforms. Codes
+are available from Zenodo at xxx
+
 \newpage
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Aiba2009}{}}%
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{1}
+\bibitem[\citeproctext]{ref-Aiba2009}
 Aiba M, Nakashizuka T (2009) Architectural differences associated with
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Anten2009}{}}%
+\bibitem[\citeproctext]{ref-Anten2009}
 Anten N, Schieving F (2009) The {Role} of {Wood} {Mass} {Density} and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bazrgar2024}{}}%
+\bibitem[\citeproctext]{ref-Bazrgar2024}
 Bazrgar AB, Thevathasan N, Gordon A, Simpson J (2024) Allometric
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Brown2004}{}}%
+\bibitem[\citeproctext]{ref-Brown2004}
 Brown JH, Gillooly JF, Allen AP, et al (2004) Toward a {Metabolic}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Brown1989}{}}%
+\bibitem[\citeproctext]{ref-Brown1989}
 Brown S, Gillespie AJR, Lugo AE (1989) Biomass {Estimation} {Methods}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
+\bibitem[\citeproctext]{ref-Carpenter2017}
 Carpenter B, Gelman A, Hoffman MD, et al (2017) Stan: {A}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Chave2009}{}}%
+\bibitem[\citeproctext]{ref-Chave2009}
 Chave J, Coomes D, Jansen S, et al (2009) Towards a worldwide wood
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Chave2014}{}}%
+\bibitem[\citeproctext]{ref-Chave2014}
 Chave J, Réjou-Méchain M, Búrquez A, et al (2014) Improved allometric
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Cochrane2023}{}}%
+\bibitem[\citeproctext]{ref-Cochrane2023}
 Cochrane A, Sims CR, Bejjanki VR, et al (2023) Multiple timescales of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Dormann2013}{}}%
+\bibitem[\citeproctext]{ref-Dormann2013}
 Dormann CF, Elith J, Bacher S, et al (2013) Collinearity: A review of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022a}{}}%
+\bibitem[\citeproctext]{ref-Fajardo2022a}
 Fajardo A (2022) Wood density relates negatively to maximum plant height
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fajardo2022b}{}}%
+\bibitem[\citeproctext]{ref-Fajardo2022b}
 Fajardo A, Piper FI, García-Cervigón AI (2022) The intraspecific
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Falster2017}{}}%
+\bibitem[\citeproctext]{ref-Falster2017}
 Falster DS, Brännström Å, Westoby M, Dieckmann U (2017) Multitrait
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Falster2003}{}}%
+\bibitem[\citeproctext]{ref-Falster2003}
 Falster DS, Westoby M (2003) Plant height and evolutionary games. Trends
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fauset2015}{}}%
+\bibitem[\citeproctext]{ref-Fauset2015}
 Fauset S, Johnson MO, Gloor M, et al (2015) Hyperdominance in
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fayolle2016}{}}%
+\bibitem[\citeproctext]{ref-Fayolle2016}
 Fayolle A, Loubota Panzou GJ, Drouet T, et al (2016) Taller trees,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Feldpausch2011}{}}%
+\bibitem[\citeproctext]{ref-Feldpausch2011}
 Feldpausch TR, Banin L, Phillips OL, et al (2011) Height-diameter
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Filipescu2012}{}}%
+\bibitem[\citeproctext]{ref-Filipescu2012}
 Filipescu CN, Groot A, MacIsaac DA, et al (2012) Prediction of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fossdal2024}{}}%
+\bibitem[\citeproctext]{ref-Fossdal2024}
 Fossdal CG, Krokene P, Olsen JE, et al (2024) Epigenetic stress memory
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelder2006}{}}%
+\bibitem[\citeproctext]{ref-Gelder2006}
 Gelder HAV, Poorter L, Sterck FJ (2006) Wood mechanics, allometry, and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
+\bibitem[\citeproctext]{ref-Gelman2013}
 Gelman A, Carlin JB, Stern HS, et al (2015)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2006}{}}%
+\bibitem[\citeproctext]{ref-Gelman2006}
 Gelman A, Hill J (2006) Data analysis using regression and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008}{}}%
+\bibitem[\citeproctext]{ref-Gelman2008}
 Gelman A, Jakulin A, Pittau MG, Su Y-S (2008) A weakly informative
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Goodman2014}{}}%
+\bibitem[\citeproctext]{ref-Goodman2014}
 Goodman RC, Phillips OL, Baker TR (2014) The importance of crown
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Henry2010}{}}%
+\bibitem[\citeproctext]{ref-Henry2010}
 Henry M, Besnard A, Asante WA, et al (2010) Wood density, phytomass
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hulshof2015}{}}%
+\bibitem[\citeproctext]{ref-Hulshof2015}
 Hulshof CM, Swenson NG, Weiser MD (2015) Tree height--diameter allometry
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Iida2012}{}}%
+\bibitem[\citeproctext]{ref-Iida2012}
 Iida Y, Poorter L, Sterck FJ, et al (2012) Wood density explains
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jucker2017}{}}%
+\bibitem[\citeproctext]{ref-Jucker2017}
 Jucker T, Caspersen J, Chave J, et al (2017) Allometric equations for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jucker2022}{}}%
+\bibitem[\citeproctext]{ref-Jucker2022}
 Jucker T, Fischer FJ, Chave J, et al (2022) Tallo: {A} global tree
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jucker2024}{}}%
+\bibitem[\citeproctext]{ref-Jucker2024}
 Jucker T, Fischer FJ, Chave J, et al (2024)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Kattge2020}{}}%
+\bibitem[\citeproctext]{ref-Kattge2020}
 Kattge J, Bönisch G, Díaz S, et al (2020) {TRY} plant trait database --
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-King2005}{}}%
+\bibitem[\citeproctext]{ref-King2005}
 King DA (2005) Linking tree form, allocation and growth with an
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-King2011}{}}%
+\bibitem[\citeproctext]{ref-King2011}
 King DA, Clark DA (2011)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-King2006}{}}%
+\bibitem[\citeproctext]{ref-King2006}
 King DA, Davies SJ, Tan S, Noor NSMd (2006) The role of wood density and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Kooyman2009}{}}%
+\bibitem[\citeproctext]{ref-Kooyman2009}
 Kooyman RM, Westoby M (2009) Costs of height gain in rainforest
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
+\bibitem[\citeproctext]{ref-Landau2021}
 Landau WM (2021) \href{https://doi.org/10.21105/joss.02959}{The targets
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Laurans2024}{}}%
+\bibitem[\citeproctext]{ref-Laurans2024}
 Laurans M, Munoz F, Charles-Dominique T, et al (2024) Why incorporate
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Ligot2018}{}}%
+\bibitem[\citeproctext]{ref-Ligot2018}
 Ligot G, Gourlet-Fleury S, Ouédraogo D-Y, et al (2018) The limited
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Lines2012}{}}%
+\bibitem[\citeproctext]{ref-Lines2012}
 Lines ER, Zavala MA, Purves DW, Coomes DA (2012) Predictable changes in
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Loehle2016}{}}%
+\bibitem[\citeproctext]{ref-Loehle2016}
 Loehle C (2016) Biomechanical constraints on tree architecture. Trees
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-LoubotaPanzou2018}{}}%
+\bibitem[\citeproctext]{ref-LoubotaPanzou2018}
 Loubota Panzou GJ, Ligot G, Gourlet-Fleury S, et al (2018) Architectural
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-MartinezCano2019}{}}%
+\bibitem[\citeproctext]{ref-MartinezCano2019}
 Martínez Cano I, Muller-Landau HC, Wright SJ, et al (2019) Tropical tree
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mikola1938}{}}%
+\bibitem[\citeproctext]{ref-Mikola1938}
 Mikola P (1938) \href{https://www.silvafennica.fi/article/4546}{Crown
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mo2024}{}}%
+\bibitem[\citeproctext]{ref-Mo2024}
 Mo L, Crowther TW, Maynard DS, et al (2024) The global distribution and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Molto2014}{}}%
+\bibitem[\citeproctext]{ref-Molto2014}
 Molto Q, Hérault B, Boreux J-J, et al (2014) Predicting tree heights for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Moore2010}{}}%
+\bibitem[\citeproctext]{ref-Moore2010}
 Moore JR (2010) Allometric equations to predict the total above-ground
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Muller-Landau2004}{}}%
+\bibitem[\citeproctext]{ref-Muller-Landau2004}
 Muller-Landau HC (2004) Interspecific and {Inter}-site {Variation} in
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Murphy2022}{}}%
+\bibitem[\citeproctext]{ref-Murphy2022}
 Murphy HT, Bradford MG (2022) The role of big trees and abundant species
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Niklas2007}{}}%
+\bibitem[\citeproctext]{ref-Niklas2007}
 Niklas KJ (2007) Maximum plant height and the biophysical factors that
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Niklas1994}{}}%
+\bibitem[\citeproctext]{ref-Niklas1994}
 Niklas KJ (1994) Plant allometry: The scaling of form and process.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012a}{}}%
+\bibitem[\citeproctext]{ref-Poorter2012a}
 Poorter H, Niklas KJ, Reich PB, et al (2012a) Biomass allocation to
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Poorter2006}{}}%
+\bibitem[\citeproctext]{ref-Poorter2006}
 Poorter L, Bongers L, Bongers F (2006) Architecture of 54
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012b}{}}%
+\bibitem[\citeproctext]{ref-Poorter2012b}
 Poorter L, Lianes E, Moreno-de las Heras M, Zavala MA (2012b)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2023}{}}%
+\bibitem[\citeproctext]{ref-RCoreTeam2023}
 R Core Team (2023) \href{https://www.R-project.org/}{R: A language and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Shendryk2016}{}}%
+\bibitem[\citeproctext]{ref-Shendryk2016}
 Shendryk I, Broich M, Tulbure MG, Alexandrov SV (2016) Bottom-up
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Shinozaki1964a}{}}%
+\bibitem[\citeproctext]{ref-Shinozaki1964a}
 Shinozaki K, Yoda K, Hozumi K, Kira T (1964) A quantitative analysis of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Sileshi2023}{}}%
+\bibitem[\citeproctext]{ref-Sileshi2023}
 Sileshi GW, Nath AJ, Kuyah S (2023) Allometric scaling and allocation
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Sivula2023}{}}%
+\bibitem[\citeproctext]{ref-Sivula2023}
 Sivula T, Magnusson M, Matamoros AA, Vehtari A (2023)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Slik2013}{}}%
+\bibitem[\citeproctext]{ref-Slik2013}
 Slik JWF, Paoli G, McGuire K, et al (2013) Large trees drive forest
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Song2024}{}}%
+\bibitem[\citeproctext]{ref-Song2024}
 Song X, Li J, Zeng X (2024) Parameterization of height--diameter and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Sterck1998}{}}%
+\bibitem[\citeproctext]{ref-Sterck1998}
 Sterck F, Bongers F (1998)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Sullivan2018}{}}%
+\bibitem[\citeproctext]{ref-Sullivan2018}
 Sullivan MJP, Lewis SL, Hubau W, et al (2018) Field methods for sampling
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Swenson2007}{}}%
+\bibitem[\citeproctext]{ref-Swenson2007}
 Swenson N, Enquist B (2007) Ecological and evolutionary determinants of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Tatsumi2023}{}}%
+\bibitem[\citeproctext]{ref-Tatsumi2023}
 Tatsumi S, Yamaguchi K, Furuya N (2023) {ForestScanner}: {A} mobile
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Tavankar2019}{}}%
+\bibitem[\citeproctext]{ref-Tavankar2019}
 Tavankar F, Lo Monaco A, Nikooy M, et al (2019) Snow damages on trees of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Thomas1996}{}}%
+\bibitem[\citeproctext]{ref-Thomas1996}
 Thomas SC (1996) Asymptotic height as a predictor of growth and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Vehtari2017}{}}%
+\bibitem[\citeproctext]{ref-Vehtari2017}
 Vehtari A, Gelman A, Gabry J (2017) Practical {Bayesian} model
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Vehtari2021}{}}%
+\bibitem[\citeproctext]{ref-Vehtari2021}
 Vehtari A, Gelman A, Simpson D, et al (2021) Rank-{Normalization},
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-West1999}{}}%
+\bibitem[\citeproctext]{ref-West1999}
 West G, Brown J, Enquist B (1999) A general model for the structure and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Yang2023}{}}%
+\bibitem[\citeproctext]{ref-Yang2023}
 Yang J, Swenson NG (2023) Height and crown allometries and their

### 2025-03-26 sup/sup.tex
clean

@@
 5 & Lloyd & Jon & 34 & The RAINFOR Plant Trait Database\\
+\addlinespace
+6 & Zanne & Amy & 61 & Global Wood Density Database\\
+7 & Weiher & Evan & 66 & Midwestern and Southern US Herbaceous Species Trait Database\\
+8 & Cornelissen & Johannes & 72 & Sheffield \& Spain Woody Database\\
+9 & Bond & William & 47 & South African Woody Plants Database (ZLTP)\\
+10 & Ordonez & Jenny & 88 & The Netherlands Plant Traits Database\\
+\addlinespace
+11 & Peuelas & Josep & 91 & Catalonian Mediterranean Forest Trait Database\\
+12 & Reich & Peter & 94 & Global A, N, P, SLA Database\\
+13 & Sosinski & Enio & 77 & FAPESP Brazil Rainforest Database\\
+14 & van Bodegom & Peter & 90 & Ukraine Wetlands Plant Traits Database\\
+15 & Wright & S. Joseph & 112 & Panama Plant Traits Database\\
+\addlinespace
+16 & Anand & Madhur & 100 & Plant Traits in Pollution Gradients Database\\
+17 & Wright & Ian & 64 & Neotropical Plant Traits Database\\
+18 & Poorter & Lourens & 33 & Tropical Rainforest Traits Database\\
+19 & Finegan & Bryan & 74 & Costa Rica Rainforest Trees Database\\
+20 & Laughlin & Daniel & 193 & Plant Traits for Pinus and Juniperus Forests in Arizona\\
+\addlinespace
+21 & Kramer & Koen & 191 & Baccara - Plant Traits of European Forests\\
+22 & Jansen & Steven & 241 & Xylem Functional Traits (XFT) Database\\
+23 & Craven & Dylan & 230 & Panama Tree Traits\\
+24 & Gutierrez & Alvaro G. & 123 & VirtualForests Trait Database\\
+25 & Powers & Jennifer & 263 & Costa Rican Tropical Dry Forest Trees\\
+\addlinespace
+26 & Gross & Nicolas & 275 & Plant Traits From Spanish Mediteranean shrublands\\
+27 & Wirth & Christian & 68 & The Functional Ecology of Trees (FET) Database  - Jena\\
+28 & Adler & Peter & 285 & Functional traits explaining variation in plant life history strategies\\
+29 & Higuchi & Pedro & 305 & Araucaria Forest Database\\
+30 & Holl & Karen & 306 & Plant traits from Costa Rica\\
+\addlinespace
+31 & Mazzochini & Guilherme & 357 & Functional traits of woody species in the Brazilian semi-arid region\\
+32 & van Bodegom & Peter & 330 & Traits of Ukraine native and invasive plant species\\
+33 & Swenson & Nathan & 288 & CTFS Luquillo Forest Dynamics Plot\\
+34 & Weigelt & Alexandra & 347 & Traits of 59 grassland species\\
+35 & Hietz & Peter & 384 & Panama wood anatomy\\
+\addlinespace
+36 & Chacon & Eduardo & 377 & Functional Traits of Trees in Golfo Dulce, Costa Rica\\
+37 & Rolo Romero & Victor & 396 & Rehabilitating Coastal dune forest\\
+38 & Li & Yuanzhi & 419 & Sherbrooke\\
+39 & Boonman & Coline & 420 & Chinese savanna trees - aboveground trait data\\
+40 & Jacobsen & Kim & 398 & Yangambi (DR Congo) tropical forest tree traits\\
+\addlinespace
+41 & Paine & C. E. Timothy & 439 & Functional Traits of Trees\\
+42 & Lopez-Gonzalez & NA & 413 & Zanne et al. Global wood density database.\\
+43 & Silva & Ana Carolina & 393 & LABDENDRO Brazilian Subtropical Forest Traits Database [Dataset II]\\
+44 & Falster & Daniel & 431 & BAAD: a biomass and allometry database for woody plants\\
+45 & Bjorkman & Anne & 468 & Tundra Trait Team\\
+\addlinespace
+46 & Thomas & Evert & 473 & Functional trait data Colombian dry Forest trees\\
+47 & Pausas & Juli & 474 & BROT 2.0\\
+48 & Torres-Ruiz & Jose M. & 492 & Torres-Ruiz\\
+49 & Schmitt & Sylvain & 532 & UppangalaTraits\\
+50 & Laughlin & Daniel & 538 & NZ tree traits\\
+\addlinespace
+51 & Michelaki & Chrysanthi & 539 & Thermo-Mediterranean species along Greece\\
+52 & Richardson & Sarah & 572 & New Zealand Nothofagus leaf and stem traits\\
+53 & Schrader & Julian & 587 & Raja Ampat tree dataset\\
+54 & Apgaua & Deborah & 622 & Daintree Rainforest Functional Traits Data\\
+55 & Tng & David & 625 & Tng et al 2013 Traits\\
+\addlinespace
+56 & Vasquez & Maribel & 627 & functional traits native and invasive species in tropical dry forest\\
+57 & Aubin & Isabelle & 629 & Traits of Plants in Canada (TOPIC)\\
+58 & Watkins & Harry & 630 & Watkins, Sjoman and Hitchmough CSR ordination of trees\\
+59 & Bruy & David & 632 & Atractocarpus from new Caledonia\\
+60 & Olson & Mark E. & 635 & Olson PNAS 2018\\
+\addlinespace
+61 & Boukili & Vanessa & 636 & La Selva FT Data\\
+62 & Olson & Mark E. & 638 & Olson et al AnnBot 2018 Corners Rules\\
+63 & VILLACIS & JAIME & 653 & RASGOS FUNCIONALES ESPECIES ARBOREAS CUENCA AMAZONICA\\
+64 & Staples & Timothy & 674 & Staples et al Australian Reforestation Tree Database\\
+65 & Helsen & Kenny & 706 & riparian forb traits W-Europe\\
+\addlinespace
+66 & Olson & Mark E. & 708 & Olson et al. 2020 Ecological Monographs vessel diameter-plant height scaling\\
+67 & Thom & Dominik & 662 & Thom 2019\\
+68 & McCarthy & James & 475 & Woody plant traits from southeast Queensland, Australia\\
+69 & Fernndez-Mndez & Fernando & 668 & Traits of urban species from Ibagu Colombia\\
+70 & Jansen & Steven & 673 & Wagenfhr Woodatlas\\
+\addlinespace
+71 & Scalon & Marina & 681 & Scalon bark traits\\
+72 & Li & Yuanzhi & 744 & Heishiding\_Performance\_Trait\_Environment\_Data\\
 \bottomrule

### 2025-03-26 _targets.R

@@ -101,17 +101,6 @@ biomes <- c(
       "Dryland",
       "Mangrove")
 
-biome_mcmc_names <- expand_grid(
-  model = c("gmm", "weibull", "pl"),
-  div = c("ang", "gym"),
-  variable = c("cr", "h"),
-  biome = biomes
-) |>
-  mutate(mcmc_names = case_when(
-    model == "pl" ~ str_c("fit_lr_nou_each_mcmc_", model, "_", div, "_", variable, "_", biome),
-    TRUE ~ str_c("fit_nlr_nou_each_mcmc_", model, "_", div, "_", variable, "_", biome)
-  )) |>
-  pull(mcmc_names)
 
 
 # Define the tar_map for LOOIC calculation
@@ -119,14 +108,6 @@ loo_map <- tar_map(
   values = list(mcmc = rlang::syms(mcmc_names)),
   tar_target(
     loo,
-    my_loo(mcmc)
-  )
-)
-
-loo_each_map <- tar_map(
-  values = list(mcmc = rlang::syms(biome_mcmc_names)),
-  tar_target(
-    loo_each,
     my_loo(mcmc)
   )
 )
@@ -348,128 +329,6 @@ tar_nlr_nou2 <-
       )
    )
 
-tar_map_lr_nou_each <-
-  tar_map(
-    values = expand_grid(
-      div = c("ang", "gym"),
-      variable = c("cr", "h"),
-      biome = c(
-        "Boreal/montane forest",
-        "Tropical rain forest",
-        "Temperate broadleaf forest",
-        "Tropical savanna",
-        "Temperate grassland",
-        "Temperate conifer forest",
-        "Mediterranean woodland",
-        "Tropical dry forest",
-        "Dryland",
-        "Mangrove")),
-    tar_target(
-      tallo_reduced_lr_df_each, {
-        if (variable == "cr") {
-          tmp <- tallo_wd_df0 |>
-            filter(!is.na(cr))
-        } else if (variable == "h") {
-          tmp <- tallo_wd_df0 |>
-            filter(!is.na(h))
-        }
-        reduce_trees_simple(tmp, 100)
-      }
-    ),
-    tar_target(
-      stan_data_lr_each, {
-        tallo_reduced_lr_df_each |>
-          filter(biome == biome) |>
-          generate_stan_data(
-            model = "lr", div = div, variable = variable)
-      }
[…]

### 2025-03-26 ms/ms_old.qmd
update data availability section

--- 2025-03-26 ms/ms.qmd
@@
 While numerous papers have used various functional forms for allometric equations [e.g., @Thomas1996; @Feldpausch2011; @MartinezCano2019; @Song2024], power-law functions remain the most common choice for describing the allometric scaling of tree size variables, including height and crown size with DBH [e.g., @Sileshi2023; @Jucker2024].
-The relationships among tree size variables often involve complex interactions influenced by various factors, such as species diversity, environmental conditions, and individual variability [e.g., @Poorter2012a; @Laurans2024].
-Consequently, power-law models often fail to accurately capture the allometric relationships across the full range of tree sizes, leading to underestimation of dimensions for smaller trees and overestimation for larger trees [e.g., @Fayolle2016; @Song2024].
+The relationships among tree size variables often involve complex interactions influenced by various factors, such as species diversity, environmental conditions, and individual variability [e.g., @Poorter2012; @Laurans2024].
+Consequently, power-law models often fail to accurately capture the allometric relationships across the full range of tree sizes, leading to underestimation of dimensions for smaller trees and underestimation for larger trees [e.g., @Fayolle2016; @Song2024].
 This highlights the need to adopt alternative functional forms—such as saturating relationships—that can more accurately reflect the complexity of allometric scaling in vegetation models [e.g., @Goodman2014; @Song2024].
 
-Earth observation (EO) technologies, including remote sensing and Light Detection and Ranging (LiDAR), have revolutionized the monitoring of forests [e.g., @Jucker2017] by enabling the identification and measurement of individual tree heights and crown dimensions using airborne imagery [e.g., @Shendryk2016; @Jucker2017].
-However, because existing allometric models still rely on stem diameter as a crucial input for biomass estimation, this reliance presents a challenge for effectively integrating EO technologies into AGB estimation [@Shendryk2016; @Jucker2017].
+Earth observation (EO) technologies, including remote sensing and Light Detection and Ranging (LiDAR), have revolutionized monitoring forests [e.g., @Jucker2017] by enabling the identification and measurement of individual tree heights and crown dimensions using airborne imagery [e.g., @Shendryk2016; @Jucker2017].
+Because existing allometric models still depend on stem diameter as a crucial input for biomass estimation, it also presents the challenge of determining the most effective way to utilize EO technologies for AGB estimation [@Shendryk2016; @Jucker2017].
 While ground-based LiDAR can measure DBH directly and is much more efficient than traditional ground surveys [@Tatsumi2023], it is still less common and convenient than aerial LiDAR.
@@
 
-Accurately representing tree allometry relationships also requires considering underlying ecological constraints, particularly the interplay of plant functional traits and their trade-offs influence individual tree architecture and the composition of entire forest communities [e.g., @Iida2012; @LoubotaPanzou2018; @Yang2023].
+Accurately representing tree allometry relationships also requires considering underlying ecological constraints, particularly the interplay of plant functional traits and their trade-offs influences individual tree architecture and the composition of entire forest communities [e.g., @Iida2012; @LoubotaPanzou2018; @Yang2023].
 Under closed-canopy conditions, strong light competition makes the balance between vertical and horizontal space use crucial for species coexistence [e.g., @Aiba2009; @Iida2012].
-Among functional traits, wood density is closely linked to both plant life history strategies [@Fajardo2022a] and light demand [e.g., @King2006; @Aiba2009; @Poorter2012a].
+Among functional traits, wood density is closely linked to both plant life history strategies [@Fajardo2022] and light demand [e.g., @King2006; @Aiba2009; @Poorter2012].
 Several studies suggest that wood density explains a trade-off between rapid stem expansion and broader crowns, shaping mechanical constraints on species with diverse architecture and light capture strategies [@Aiba2009; @Iida2012; @Yang2023].
@@
 Using large datasets from the Tallo and TRY databases [@Jucker2022; @Kattge2020], which span a diverse array of tree forms across global forest types and climatic conditions, we overcome the limitations of previous studies that often relied on site-specific or regional observations with limited samples and species.
-By integrating species-level wood density (WD) into a multilevel model, our methodology avoids the post-hoc correlations used in previous studies, offering a more universal and scalable model, effectively capturing variability from the species to the community level.
-Our main objective is to develop models informed by tree allometric variables (H, CR, DBH, WD) to capture interspecific variation in allometric scaling globally and offer an improved framework for estimating above ground biomass.
+By integrating species-level wood density into a multilevel model, our methodology avoids the post-hoc correlations used in previous studies, offering a more universal and scalable model, effectively capturing variability from the species to the community level.
 Specifically, we ask the following questions:
@@
 
-Our analysis combined tree allometry data from Tallo, which includes standardized measurements of DBH, height, and crown radius for 5,163 species across various biomes [@Jucker2022, https://doi.org/10.5281/zenodo.6637599] along with wood density data for 10,995 species from TRY version 6 [@Kattge2020, https://www.try-db.org/TryWeb/Home.php] (see Supplementary Table S1 for further details).
+Our analysis combined tree allometry data from Tallo, which includes standardized measurements of DBH, height, crown radius for 5,163 species across various biomes [@Jucker2022, https://doi.org/10.5281/zenodo.6637599] along with wood density data for 10,995 species from TRY version 6 [@Kattge2020, https://www.try-db.org/TryWeb/Home.php] (see Supplementary Table S1 for further details).
 We thoroughly processed the Tallo dataset by removing 998 outliers following @Jucker2022.
@@
 
-Once the best-performing models for each allometric relationship were identified, we tested whether adding wood density as a species-level predictor would further improve the predictive performance of those models and also examined its influence on tree architecture diversity.
+Once the best-performing models were identified for each allometric relationship, we then tested whether adding a species-level predictor (wood density) would further improve that model’s predictive performance.
 Species-specific parameters ($\beta_j$) were modeled as a function of wood density ($u_j$):
@@
 where $\gamma_{0}$ and $\gamma_{1}$ are the species-level intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
-To facilitate interpretation and visualization, species were then classified into "light" and "dense" wood groups based on the 10th and 90th percentiles of wood density distribution in the observed dataset, respectively.
-This classification allowed us to illustrate the relationships between wood density and the parameters of the best-performing models, enabling an evaluation of how wood density contributes to variation in tree architecture.
 Additionally, we explored a model that accounts for covariance between species-specific parameters using a multivariate normal distribution.
@@
 
-For angiosperm species, we used the widely applied equation from @Chave2014 to estimate aboveground biomass (AGB; kg of dry mass):
+For angiosperm species, we estimated aboveground biomass (AGB; kg of dry mass) using the equation from @Chave2014:
 
@@
 The optimal model to predict tree height was a Weibull function for both angiosperms (ang) and gymnosperms (gym) (Table 1).
-Across the global dataset, the best model to predict H from DBH of each clade was:
+Across the global dataset, the best model to predict H from DBH of each division was:
 $$
@@
 
-The second-best model, which used the gMM function, showed poorer predictive performance in angiosperms ($\Delta$LOOIC = 101.01) but showed competitive predictive performance in gymnosperms ($\Delta$LOOIC = 4.87).
+The second-best model, which used the gMM function, showed poorer predictive performance in angiosperms ($\Delta$ LOOIC = 101.01) but showed competitive predictive performance in gymnosperms ($\Delta$ LOOIC = 4.87).
 While power-law models (@eq-pl) can exhibit saturating behavior when 0 < *b* < 1, they performed the worst overall compared to other saturating models (@eq-gMM; @eq-wb).
@@
 For each model, we report the Expected Log Predictive Density (ELPD), the LOOIC, and the $R^2$ values.
-The difference in LOOIC relative to the best-performing model, $\Delta$LOOIC, is also included, with a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance [@Sivula2023; @Cochrane2023].
-Lower $\Delta$LOOIC values indicate models that perform closer to the best model in terms of predictive accuracy, while higher ELPD values reflect better model fit.
+The difference in LOOIC relative to the best-performing model, $\Delta$\_LOOIC, is also included.
+Lower $\Delta$\_LOOIC values indicate models that perform closer to the best model in terms of predictive accuracy, while higher ELPD values reflect better model fit.
 Higher $R^2$ values reflect better explanatory power, indicating how well the model explains the variance in the dependent variable.
@@
 ![
-Allometric relationships between DBH and tree height or crown radius in angiosperms (a, c) and gymnosperms (b, d), based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
-A color gradient from black (low density) to yellow (high density) indicates the density of overlapping data points, calculated via 2D kernel density estimation on a 300 $\times$ 300 grid.
-See Supplementary Figure S1 for the log-scale version of these plots.
+Allometric relationships between tree height and crown radius with DBH of angiosperm (a, c) and gymnosperm (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
+The density of overlapping points is shown using a color gradient ranging from black (low density) to yellow (high density).
+Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, estimating the concentration of data points in the space of DBH and tree height or crown radius (see Supplementary Figure S1 for the log-scale version).
 ](../figs/h_cr_dbh_non_log.png){#fig-compare}
@@
 Whereas for gymnosperms, the gMM function offered the best performance for this allometric scaling (Table 1).
-Across the global dataset, the best model for predicting CR from DBH of each clade was:
+Across the global dataset, the best model for predicting CR from DBH of each division was:
 $$
@@
 
-For angiosperms, the saturating functions (@eq-gMM; @eq-wb) performed worse than the power-law model, with $\Delta$LOOIC values of 124.57 for gMM and 126.35 for the Weibull function.
-In gymnosperms, the Weibull function showed a competitive performance compared to the gMM function, while the power-law function (@eq-pl) performed the worst, with a $\Delta$LOOIC of 162.72.
+For angiosperms, the saturating functions (@eq-gMM; @eq-wb) performed worse than the power-law model, with $\Delta$ LOOIC values of 124.57 for gMM and 126.35 for the Weibull function.
+In gymnosperms, the Weibull function showed a competitive performance compared to the gMM function, while the power-law function (@eq-pl) performed the worst, with a $\Delta$ LOOIC of 162.72.
 Incorporating wood density did not significantly improve predictive performance (Table 1).
 
-**Table 2**: Posterior estimates of the parameters for the best hierarchical models selected for tree height, crown radius, and DBH allometries in angiosperms and gymnosperms.
-Each set includes posterior estimates both without and with the inclusion of wood density, reported as median and 95% Bayesian credible intervals for the community-level parameters of each allometric function.
-For tree height, the Weibull model was identified as the best for both clades.
-Crown radius was best described by a power-law model in angiosperms and a gMM model for gymnosperms.
-The best model for DBH allometry in both clades, which incorporates both tree height and crown radius information ($CR,H$ model).
+**Table 2**: Posterior estimates of the parameters for the best hierarchical models selected for tree height, crown radius, and DBH allometries.
+Each set includes the best models for angiosperms and gymnosperms, with posterior estimates reported both without and with the inclusion of wood density.
+Table entries represent the median and 95\% Bayesian credible intervals for the community-level parameters of each allometric function.
+For tree height, the Weibull model was identified as the best for both angiosperm and gymnosperm groups.
+Crown radius was best described by a power-law model for angiosperms and a gMM model for gymnosperms.
+The best model for DBH allometry in both divisions incorporated both tree height and crown radius information ($CR,H$ model).
 
@@
 best_posterior_df <- read_csv("data/best_posterior_df.csv")
-
-best_posterior_df <- best_posterior_df |>
-  mutate(
-    Intercept_CI = sapply(Intercept_CI, format_ci),
-    Slope_CI = sapply(Slope_CI, format_ci),
-    Tau_CI = sapply(Tau_CI, format_ci)
-  )
 
@@
 The $CR,H$ model, which treated CR and H as separate factors, exhibited the best predictive performance among all tested models (Table 1).
-Across the global dataset, the most effective model for predicting DBH from H and CR for each clade was:
+Across the global dataset, the most effective model for predicting DBH from H and CR for each division was:
 $$
@@
 Curves represent the allometric scaling for species with light (orange) and dense (brown) wood density, corresponding to the 10\% and 90\% quantiles of wood density distribution in the observed dataset, respectively.
-The shaded areas illustrate the 95% Bayesian credible intervals.
+The shaded areas illustrate the 95% CIs.
 Only statistically significant associations from @fig-wood1 are plotted.
@@
 
-AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based both on the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
-Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_pl_h)` Mg and bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_pl_h * 100)`% (@fig-agb b).
+AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based both on the lowest root mean square error (RMSE) of `r round(data_metrics$metrics_agb_ang$rmse_wb_h, 3)` Mg and the smallest bias of `r round(data_metrics$metrics_agb_ang$bias_wb_h * 100, 3)`% (@fig-agb a).
+Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r round(data_metrics$metrics_agb_ang$rmse_pl_h, 3)` Mg and bias of `r round(data_metrics$metrics_agb_ang$bias_pl_h * 100, 3)`% (@fig-agb b).
 In contrast, models that used estimated DBH performed substantially worse (@fig-agb c-f).
@@
 ![
-Comparison of individual tree above-ground biomass (AGB) estimated using different approaches with reference AGB ($\text{AGB}_\text{ref}$) derived from observed DBH and tree height in angiosperms.
+Comparison of individual tree above-ground biomass estimated from different approaches with reference AGB ($\text{AGB}_\text{ref}$) derived from observed DBH and tree height in angiosperms.
 (a, b) AGB estimates based on observed DBH and tree height predicted by the Weibull function ($\text{AGB}_\text{H-WB}$) or the power-law function ($\text{AGB}_\text{H-PL}$).
 (c–f) AGB estimates based on observed tree height and DBH predicted from crown radius and tree height as separate factors ($\text{AGB}_{\text{DBH} \sim \text{CR,H}}$), as a compound term ($\text{AGB}_{\text{DBH} \sim \text{CR × H}}$), from tree height alone ($\text{AGB}_{\text{DBH} \sim \text{H}}$), or from crown radius alone ($\text{AGB}_{\text{DBH} \sim \text{CR}}$).
-Dashed lines represent the 1:1 relationships, and the solid lines show linear regression fits to the data, highlighting how predictive accuracy varies with tree size.
+Dashed lines represent the 1:1 relationships, and the solid lines are linear regression fits to the data points to highlight how predictive accuracy varies with tree size.
 RMSE (root mean square error) and bias values for each model are reported.
@@
 Additionally, the metabolic scaling theory (MST) posits logarithmic scaling relationships among different axes of tree size, with an exponent close to 2/3 ($H \propto DBH^{2/3}$, $CR \propto DBH^{2/3}$), based on assumptions regarding mechanical and hydraulic constraints to plant growth [@West1999].
-However, our findings revealed that the power-law model exponent of angiosperms (*b* = 0.538) was significantly smaller, while the actual exponent in gymnosperms (*b* = 0.704) was larger than the theoretical value of 2/3 (Supplementary Table S3).
+However, our findings revealed that the power-law model exponent of angiosperms (*b* = 0.54) was significantly smaller, while the actual exponent in gymnosperms (*b* = 0.7) was larger than the theoretical value of 2/3 (Supplementary Table S3).
 This divergence highlights the importance of separately modeling the allometric relationships of each functional group to accurately capture their distinct scaling patterns.
@@
 Our analysis also revealed divergent patterns in CR-DBH allometry, highlighting fundamental differences in growth strategies and ecological adaptations between angiosperms and gymnosperms.
-In angiosperms, the CR–DBH relationship follows a power-law form, suggesting a capacity for plastic crown expansion in competitive environments [e.g., @Poorter2012a; @Lines2012; @Hulshof2015], a conclusion consistent with a study in tropical moist forests by @MartinezCano2019.
+In angiosperms, the CR–DBH relationship follows a power-law form, suggesting a capacity for plastic crown expansion in competitive environments [e.g., @Poorter2012; @Lines2012; @Hulshof2015], a conclusion consistent with a study in tropical moist forests by @MartinezCano2019.
 In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent *b* is near the expected 2/3
-($b_{\text{ang}}$ = `r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
-[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
-`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`], Table 2).
+($b_{\text{ang}} = `r data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median` \,
+\left[ `r data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower`, `r data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper` \right]$, Table 2).
 In contrast, the CR-DBH allometry in gymnosperms showed an asymptotic relationship best described by the gMM function, with the Weibull function also showing competitive predictive performance (Table 1).
-This asymptotic pattern likely arises from structural constraints in gymnosperms growth forms that favour vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].
+This asymptotic pattern likely arises from structural constraints in gymnosperms growth forms that favour vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016].
 
 Finally, we tested models to estimate DBH from biometric variables (i.e., tree height and crown radius) measurable via Earth observation (EO) platforms, such as remote sensing and LiDAR [@Filipescu2012; @Jucker2017].
-Our analysis showed that accurately predicting DBH requires both tree height and crown radius [@Jucker2017], which is especially important for distinguishing among trees of similar height but varying DBH [@King2005; @King2011].
+Our analysis showed that accurately predicting DBH requires both tree height and crown radius [@Jucker2017], which is especially important for distinguishing among trees of similar height but varying DBH [@King2005].
 Among the tested models, the $CR,H$ model ($\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}$) provided the highest predictive accuracy, a result that contrasts with the previous study [@Jucker2017], which concluded that the $CR \times H$ model ($\mathrm{DBH} = a_2(\mathrm{CR} \times \mathrm{H})^{b_2}$) was the best for predicting DBH.
@@
 In contrast, gymnosperms exhibit a height growth strategy that is largely independent of wood density (@fig-wood1 d-f).
-These results are consistent with the previous study [@Poorter2012b], which showed that among seven angiosperm species, those with lighter wood exhibited taller maximum heights, while no significant correlations were observed for six gymnosperm species.
+These results are consistent with the previous study [@Poorter2012a], which showed that among seven angiosperm species, those with lighter wood exhibited taller maximum heights, while no significant correlations were observed for six gymnosperm species.
 A recent study using the Tallo database alongside two additional datasets reported a nonsignificant relationship between tree height and wood density [@Jucker2024].
@@
 Although wood density is often regarded as a key functional trait [e.g., @Anten2009; @Jucker2024], variations in tree allometry across individuals are likely influenced by a combination of species identity and environmental factors.
-Across species or along broad environmental gradients, wood density systematically varies—tending to be denser in drier, colder climates and in nutrient-poor soils [@Muller-Landau2004; @Chave2009; @Fajardo2022b; @Mo2024]—reflecting species’ adaptive strategies.
-Such environmental influences might confound simple correlations between wood density and DBH when examining diverse tree communities.
-At the intraspecific level, recent findings suggest that wood density may be more closely related to vessel anatomy than to tree architecture across different environments [@Fajardo2022b].
 Furthermore, substantial measurement errors in tree height and crown radius relative to DBH may introduce noise that obscures potential relationships between wood density and stem diameter.
@@
 Our findings revealed minimal differences in AGB estimates derived from $\text{H}_\text{WB}$ and $\text{H}_\text{PL}$ (@fig-agb a, b).
-This result can stem from the fact that over 97\% of the trees used to estimate AGB are small to medium-sized (DBH < 70 cm [@Slik2013; @Murphy2022]), whose H-DBH relationships can still be well captured by the power-law function (@fig-compare a) [e.g., @Molto2014].
+This result can stem from the fact that over 97\% of the trees used to estimate AGB are small to medium-sized (DBH < 70 cm [@Slik2013]), whose H-DBH relationships can still be well captured by the power-law function (@fig-compare a) [e.g., @Molto2014].
 In contrast, individuals with DBH ≥ 70 cm, while typically forming a small proportion of the tree population, contribute significantly to biomass storage [e.g., @Fauset2015].
@@
 However, the fundamental assumption of the power-law function does not fully reflect the asymptotic nature of height-diameter allometries in large trees (@fig-compare a), potentially leading to significant inaccuracies in stand-level biomass estimates [e.g., @Feldpausch2011; @Fayolle2016; @Sullivan2018; @MartinezCano2019].
-Because large trees contribute disproportionately to overall biomass  [e.g., @Slik2013; @Ligot2018; @Fauset2015], if direct height measurements are unavailable, height estimates should be derived by fitting asymptotic models (e.g., @eq-gMM; @eq-wb) to datasets with sufficient representation of large trees, ensuring that the saturating component is accurately captured [@Sullivan2018].
+Given the disproportionate contribution of large trees in biomass dynamics [e.g., @Slik2013; @Ligot2018; @Fauset2015], when heights are not directly measured, height estimates should be derived by fitting asymptotic models (e.g., @eq-gMM; @eq-wb) to datasets with sufficient representation of large trees, ensuring the accurate capture of the saturating component [@Sullivan2018].
 
@@
 Additionally, most existing AGB estimation equations typically include DBH as a squared term in equations [e.g., @Brown1989; @Chave2014], meaning that even minor errors in DBH predictions can significantly affect the accuracy of AGB estimates [@Jucker2017].
-Interestingly, our results revealed that, although the $CR,H$ model provided more accurate direct DBH predictions, using those predictions in the AGB equation resulted in higher RMSE and bias than when DBH estimates from the $CR \times H$ model were used (@fig-agb c, d).
+Interestingly, our results revealed that the $CR,H$ model was more effective in predicting DBH; however, when these predicted DBH values were used to estimate AGB, the $CR \times H$ model exhibited substantially lower RMSE and bias (@fig-agb c, d).
 This is likely due to the $CR \times H$ model's ability to capture the inherent interdependence between the crown radius and tree height (*r* = 0.703), offering a more biologically realistic framework [e.g., @Dormann2013].
@@
 Our results also highlight the importance of incorporating saturating models when forecasting ecosystem services in forests with a broad range of species and structural heterogeneity.
-Future studies should investigate additional functional traits and environmental factors to improve both ecological understanding and carbon monitoring practices at global scales.
+Future studies should investigate additional functional traits and environmental factors to improve both ecological understanding and carbon monitoring practices at global scales monitoring practices at global scales.
 
@@
 
-We thank the editors and two anonymous reviewers, whose comments greatly improved the manuscript.
-We also thank the Alliance of International Science Organizations (ANSO) for providing financial support for TDN’s Master’s studies.
+We sincerely thank the Alliance of International Science Organizations (ANSO) for providing financial support for TDN’s Master’s studies.
 
@@
 
-MK was supported by
-the Xingdian Talenj Support Program of Yunnan Province,
-the Xishuangbanna State Rainforest Talent Support Program (E4BN041B01), and
-the CAS President's International Fellowship Initiative (2020FYB0003).
-
-# Data availability {.unnumbered}
-
-Tallo and TRY data are available from their respective platforms.
-Codes are available from Zenodo at xxx
+MK was supported by the Xishuangbanna State Rainforest Talent Support Program (E4BN041B01) and the CAS President's International Fellowship Initiative (2020FYB0003).
 

### 2025-03-26 ms/ms_old.tex
update data availability section

@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 papers have used various functional forms for allometric equations
-(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
-\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
-\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
-\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
-remain the most common choice for describing the allometric scaling of
-tree size variables, including height and crown size with DBH (e.g.,
-\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
-relationships among tree size variables often involve complex
-interactions influenced by various factors, such as species diversity,
-environmental conditions, and individual variability
-(\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}; e.g.,
-\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}).
+(e.g., \citeproc{ref-Thomas1996}{Thomas 1996};
+\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019};
+\citeproc{ref-Song2024}{Song et al. 2024}), power-law functions remain
+the most common choice for describing the allometric scaling of tree
+size variables, including height and crown size with DBH (e.g.,
+\citeproc{ref-Sileshi2023}{Sileshi et al. 2023};
+\citeproc{ref-Jucker2024}{Jucker et al. 2024}). The relationships among
+tree size variables often involve complex interactions influenced by
+various factors, such as species diversity, environmental conditions,
+and individual variability (\citeproc{ref-Laurans2024}{Laurans et al.
+2024}; e.g., \citeproc{ref-Poorter2012}{\textbf{Poorter2012?}}).
 Consequently, power-law models often fail to accurately capture the
@@
 underestimation of dimensions for smaller trees and underestimation for
-larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
-2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
-highlights the need to adopt alternative functional forms---such as
-saturating relationships---that can more accurately reflect the
-complexity of allometric scaling in vegetation models (e.g.,
-\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
-\protect\hyperlink{ref-Song2024}{Song et al. 2024}).
+larger trees (e.g., \citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
+\citeproc{ref-Song2024}{Song et al. 2024}). This highlights the need to
+adopt alternative functional forms---such as saturating
+relationships---that can more accurately reflect the complexity of
+allometric scaling in vegetation models (e.g.,
+\citeproc{ref-Goodman2014}{Goodman et al. 2014};
+\citeproc{ref-Song2024}{Song et al. 2024}).
 
@@
 Detection and Ranging (LiDAR), have revolutionized monitoring forests
-(e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}) by
-enabling the identification and measurement of individual tree heights
-and crown dimensions using airborne imagery (e.g.,
-\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Because
-existing allometric models still depend on stem diameter as a crucial
-input for biomass estimation, it also presents the challenge of
-determining the most effective way to utilize EO technologies for AGB
-estimation (\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
-ground-based LiDAR can measure DBH directly and is much more efficient
-than traditional ground surveys
-(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
-less common and convenient than aerial LiDAR. Alternative approaches
-focusing on tree height and crown dimensions as model predictors are
-also necessary for remote sensing applications
-(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
-solution is to estimate DBH from tree height and crown size and then use
+(e.g., \citeproc{ref-Jucker2017}{Jucker et al. 2017}) by enabling the
+identification and measurement of individual tree heights and crown
+dimensions using airborne imagery (e.g.,
+\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). Because existing
+allometric models still depend on stem diameter as a crucial input for
+biomass estimation, it also presents the challenge of determining the
+most effective way to utilize EO technologies for AGB estimation
+(\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). While ground-based LiDAR
+can measure DBH directly and is much more efficient than traditional
+ground surveys (\citeproc{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is
+still less common and convenient than aerial LiDAR. Alternative
+approaches focusing on tree height and crown dimensions as model
+predictors are also necessary for remote sensing applications
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}). One potential solution
+is to estimate DBH from tree height and crown size and then use
 established biomass equations, which could be an effective strategy for
 integrating airborne imagery into carbon monitoring programs
-(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
 
@@
 individual tree architecture and the composition of entire forest
-communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
-\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
-\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
-closed-canopy conditions, strong light competition makes the balance
-between vertical and horizontal space use crucial for species
-coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
-2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
-functional traits, wood density is closely linked to both plant life
-history strategies
-(\protect\hyperlink{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light
-demand (e.g., \protect\hyperlink{ref-King2006}{King et al. 2006};
-\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
-\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}). Several
-studies suggest that wood density explains a trade-off between rapid
-stem expansion and broader crowns, shaping mechanical constraints on
-species with diverse architecture and light capture strategies
-(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
-\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
-\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
-types of studies have examined this relationship post hoc, relating
-species-specific allometric parameters to wood density and risking Type
-I or Type II errors if the hierarchical data structure is overlooked
-(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
-exception is Martínez Cano et al.
-(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
-density into a hierarchical model of tropical trees and concluded it
-offered minimal explanatory power for interspecific variation.
-Therefore, the role of wood density in global allometric scaling remains
-insufficiently explored.
+communities (e.g., \citeproc{ref-Iida2012}{Iida et al. 2012};
+\citeproc{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
+\citeproc{ref-Yang2023}{Yang and Swenson 2023}). Under closed-canopy
+conditions, strong light competition makes the balance between vertical
+and horizontal space use crucial for species coexistence (e.g.,
+\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}). Among functional traits,
+wood density is closely linked to both plant life history strategies
+(\citeproc{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light demand
+(e.g., \citeproc{ref-King2006}{King et al. 2006};
+\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\citeproc{ref-Poorter2012}{\textbf{Poorter2012?}}). Several studies
+suggest that wood density explains a trade-off between rapid stem
+expansion and broader crowns, shaping mechanical constraints on species
+with diverse architecture and light capture strategies
+(\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}; \citeproc{ref-Yang2023}{Yang
+and Swenson 2023}). However, these types of studies have examined this
+relationship post hoc, relating species-specific allometric parameters
+to wood density and risking Type I or Type II errors if the hierarchical
+data structure is overlooked (\citeproc{ref-Gelman2006}{Gelman and Hill
+2006}). One exception is Martínez Cano et al.
+(\citeproc{ref-MartinezCano2019}{2019}), who incorporated wood density
+into a hierarchical model of tropical trees and concluded it offered
+minimal explanatory power for interspecific variation. Therefore, the
+role of wood density in global allometric scaling remains insufficiently
+explored.
 
@@
 standardized measurements of DBH, height, crown radius for 5,163 species
-across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker et al.
-2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
-data for 10,995 species from TRY version 6
-(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
-https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
-further details). We thoroughly processed the Tallo dataset by removing
-998 outliers following Jucker et al.
-(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
-selected only species with standardized wood density values in g
+across various biomes (\citeproc{ref-Jucker2022}{Jucker et al. 2022},
+https://doi.org/10.5281/zenodo.6637599) along with wood density data for
+10,995 species from TRY version 6 (\citeproc{ref-Kattge2020}{Kattge et
+al. 2020}, https://www.try-db.org/TryWeb/Home.php) (see Supplementary
+Table S1 for further details). We thoroughly processed the Tallo dataset
+by removing 998 outliers following Jucker et al.
+(\citeproc{ref-Jucker2022}{2022}). To ensure consistency, we selected
+only species with standardized wood density values in g
 cm\textsuperscript{-3} from TRY and removed any records with missing
@@
 function of wood density (\(u_j\)):
-\begin{equation}\protect\hypertarget{eq-wood}{}{
+\begin{equation}\phantomsection\label{eq-wood}{
 \beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
-}\label{eq-wood}\end{equation}
+}\end{equation}
 
@@
 methods that rely on random proposals
-(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al. 2017}). This
-efficiency allows us to achieve convergence with fewer iterations than
-traditional MCMC methods. Inferences were based on 1,000 posterior
-samples following 1,000 warmup iterations for three parallel chains.
-Convergence was assessed using the Gelman--Rubin statistic with a
-convergence threshold of 1.1 (\protect\hyperlink{ref-Gelman2013}{Gelman
-et al. 2015}) and ensuring effective sample sizes greater than 300
-(\protect\hyperlink{ref-Vehtari2021}{Vehtari et al. 2021}). We also
-reported the median and 95\% Bayesian credible intervals (CIs) for each
-parameter (Supplementary Tables S3-S8). Estimates were considered
-significant when 95\% CIs did not overlap with zero.
+(\citeproc{ref-Carpenter2017}{Carpenter et al. 2017}). This efficiency
+allows us to achieve convergence with fewer iterations than traditional
+MCMC methods. Inferences were based on 1,000 posterior samples following
+1,000 warmup iterations for three parallel chains. Convergence was
+assessed using the Gelman--Rubin statistic with a convergence threshold
+of 1.1 (\citeproc{ref-Gelman2013}{Gelman et al. 2015}) and ensuring
+effective sample sizes greater than 300
+(\citeproc{ref-Vehtari2021}{Vehtari et al. 2021}). We also reported the
+median and 95\% Bayesian credible intervals (CIs) for each parameter
+(Supplementary Tables S3-S8). Estimates were considered significant when
+95\% CIs did not overlap with zero.
 
 Statistical analyses were performed in R version 4.3.2
-(\protect\hyperlink{ref-RCoreTeam2023}{R Core Team 2023}), using the R
-packages \emph{targets} version 1.7.0 for workflow management
-(\protect\hyperlink{ref-Landau2021}{Landau 2021}). Codes are made
-available on GitHub at
-\url{https://github.com/duyenecology/tree-allometry} (Stan code is
-provided in Supplementary Section S2).
-
-\hypertarget{aboveground-biomass-estimates}{%
+(\citeproc{ref-RCoreTeam2023}{R Core Team 2023}), using the R packages
+\emph{targets} version 1.7.0 for workflow management
+(\citeproc{ref-Landau2021}{Landau 2021}). Codes are made available on
+GitHub at \url{https://github.com/duyenecology/tree-allometry} (Stan
+code is provided in Supplementary Section S2).
+
 \subsection{Aboveground biomass
-estimates}\label{aboveground-biomass-estimates}}
+estimates}\label{aboveground-biomass-estimates}
 
@@
 mass) using the equation from Chave et al.
-(\protect\hyperlink{ref-Chave2014}{2014}):
-
-\begin{equation}\protect\hypertarget{eq-agb}{}{
+(\citeproc{ref-Chave2014}{2014}):
+
+\begin{equation}\phantomsection\label{eq-agb}{
 \mathrm{AGB} = 0.0673 \times (\rho D^2H)^{0.976}
-}\label{eq-agb}\end{equation}
+}\end{equation}
 
@@
 dataset, the best model to predict H from DBH of each division was:
-\begin{equation}\protect\hypertarget{eq-h_ang}{}{
+\begin{equation}\phantomsection\label{eq-h_ang}{
 \mathrm{H_{ang}} =50.0
@@
 1). Across the global dataset, the best model for predicting CR from DBH
-of each division was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
+of each division was: \begin{equation}\phantomsection\label{eq-cr_ang}{
 \mathrm{CR_{ang}} = \mathrm{0.349}
 \, \mathrm{DBH}^{\mathrm{0.626}}
-}\label{eq-cr_ang}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-cr_gym}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-cr_gym}{
 \mathrm{CR_{gym}} = \frac{\mathrm{769.4}
@@
 DBH from H and CR for each division was:
-\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
+\begin{equation}\phantomsection\label{eq-dbh_ang}{
 \mathrm{DBH_{ang}} = \mathrm{0.948}
@@
 \(CR \propto DBH^{2/3}\)), based on assumptions regarding mechanical and
-hydraulic constraints to plant growth
-(\protect\hyperlink{ref-West1999}{West et al. 1999}). However, our
-findings revealed that the power-law model exponent of angiosperms
-(\emph{b} = 0.54) was significantly smaller, while the actual exponent
-in gymnosperms (\emph{b} = 0.7) was larger than the theoretical value of
-2/3 (Supplementary Table S3). This divergence highlights the importance
-of separately modeling the allometric relationships of each functional
-group to accurately capture their distinct scaling patterns. Moreover,
-the power-law function performed worse than the saturating functions
-(Table 1), indicating that the actual H-DBH relationship does not follow
-the simple scaling laws proposed by MST.
+hydraulic constraints to plant growth (\citeproc{ref-West1999}{West et
+al. 1999}). However, our findings revealed that the power-law model
+exponent of angiosperms (\emph{b} = 0.54) was significantly smaller,
+while the actual exponent in gymnosperms (\emph{b} = 0.7) was larger
+than the theoretical value of 2/3 (Supplementary Table S3). This
+divergence highlights the importance of separately modeling the
+allometric relationships of each functional group to accurately capture
+their distinct scaling patterns. Moreover, the power-law function
+performed worse than the saturating functions (Table 1), indicating that
+the actual H-DBH relationship does not follow the simple scaling laws
+proposed by MST.
 
@@
 plastic crown expansion in competitive environments
-(\protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}; e.g.,
-\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}), a
-conclusion consistent with a study in tropical moist forests by Martínez
-Cano et al. (\protect\hyperlink{ref-MartinezCano2019}{2019}). In
-addition, the estimated exponent aligns closely with MST predictions for
-crown radius, supporting the selection of the power-law model for
-angiosperms, where the community-level exponent \emph{b} is near the
-expected 2/3
-(\(b_{\text{ang}} = 0.625775 \, \left[ 0.6168636, 0.6345041 \right]\),
-Table 2). In contrast, the CR-DBH allometry in gymnosperms showed an
-asymptotic relationship best described by the gMM function, with the
-Weibull function also showing competitive predictive performance (Table
-1). This asymptotic pattern likely arises from structural constraints in
-gymnosperms growth forms that favour vertical over lateral expansion
-(e.g., \protect\hyperlink{ref-Lines2012}{Lines et al. 2012};
-\protect\hyperlink{ref-Hulshof2015}{Hulshof et al. 2015}), an adaptation
-that can mitigate risks such as branch breakage under heavy snow loads
-(e.g., \protect\hyperlink{ref-Mikola1938}{Mikola 1938};
-\protect\hyperlink{ref-Loehle2016}{Loehle 2016}).
+(\citeproc{ref-Lines2012}{Lines et al. 2012};
+\citeproc{ref-Hulshof2015}{Hulshof et al. 2015}; e.g.,
+\citeproc{ref-Poorter2012}{\textbf{Poorter2012?}}), a conclusion
+consistent with a study in tropical moist forests by Martínez Cano et
+al. (\citeproc{ref-MartinezCano2019}{2019}). In addition, the estimated
+exponent aligns closely with MST predictions for crown radius,
+supporting the selection of the power-law model for angiosperms, where
+the community-level exponent \emph{b} is near the expected 2/3
+(\(b_{\text{ang}} = 0.625775 \,
+\left[ 0.6168636, 0.6345041 \right]\), Table 2). In contrast, the CR-DBH
+allometry in gymnosperms showed an asymptotic relationship best
+described by the gMM function, with the Weibull function also showing
+competitive predictive performance (Table 1). This asymptotic pattern
+likely arises from structural constraints in gymnosperms growth forms
+that favour vertical over lateral expansion (e.g.,
+\citeproc{ref-Lines2012}{Lines et al. 2012};
+\citeproc{ref-Hulshof2015}{Hulshof et al. 2015}), an adaptation that can
+mitigate risks such as branch breakage under heavy snow loads (e.g.,
+\citeproc{ref-Mikola1938}{Mikola 1938}; \citeproc{ref-Loehle2016}{Loehle
+2016}).
 
@@
 (EO) platforms, such as remote sensing and LiDAR
-(\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012};
-\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Our analysis
-showed that accurately predicting DBH requires both tree height and
-crown radius (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}),
-which is especially important for distinguishing among trees of similar
-height but varying DBH (\protect\hyperlink{ref-King2005}{King 2005}).
-Among the tested models, the \(CR,H\) model
+(\citeproc{ref-Filipescu2012}{Filipescu et al. 2012};
+\citeproc{ref-Jucker2017}{Jucker et al. 2017}). Our analysis showed that
+accurately predicting DBH requires both tree height and crown radius
+(\citeproc{ref-Jucker2017}{Jucker et al. 2017}), which is especially
+important for distinguishing among trees of similar height but varying
+DBH (\citeproc{ref-King2005}{King 2005}). Among the tested models, the
+\(CR,H\) model
 (\(\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \times \mathrm{H}^{c_1}\))
 provided the highest predictive accuracy, a result that contrasts with
-the previous study (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
-2017}), which concluded that the \(CR \times H\) model
+the previous study (\citeproc{ref-Jucker2017}{Jucker et al. 2017}),
+which concluded that the \(CR \times H\) model
 (\(\mathrm{DBH} = a_2(\mathrm{CR} \times \mathrm{H})^{b_2}\)) was the
@@
 of wood density (Fig.~\ref{fig-wood1} d-f). These results are consistent
-with the previous study (\protect\hyperlink{ref-Poorter2012a}{Poorter et
-al. 2012}), which showed that among seven angiosperm species, those with
+with the previous study (\citeproc{ref-Poorter2012a}{Poorter et al.
+2012}), which showed that among seven angiosperm species, those with
 lighter wood exhibited taller maximum heights, while no significant
@@
 saplings in an Australian subtropical forest
-(\protect\hyperlink{ref-Kooyman2009}{Kooyman and Westoby 2009}) but
-contradicts the expectation that species with lower wood density develop
-thicker stems to ensure structural stability, as suggested by prior
-studies (\protect\hyperlink{ref-King2006}{King et al. 2006};
-\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
-\protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Although wood
-density is often regarded as a key functional trait (e.g.,
-\protect\hyperlink{ref-Anten2009}{Anten and Schieving 2009};
-\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}), variations in
-tree allometry across individuals are likely influenced by a combination
-of species identity and environmental factors. Furthermore, substantial
+(\citeproc{ref-Kooyman2009}{Kooyman and Westoby 2009}) but contradicts
+the expectation that species with lower wood density develop thicker
+stems to ensure structural stability, as suggested by prior studies
+(\citeproc{ref-King2006}{King et al. 2006};
+\citeproc{ref-Anten2009}{Anten and Schieving 2009};
+\citeproc{ref-Iida2012}{Iida et al. 2012}). Although wood density is
+often regarded as a key functional trait (e.g.,
+\citeproc{ref-Anten2009}{Anten and Schieving 2009};
+\citeproc{ref-Jucker2024}{Jucker et al. 2024}), variations in tree
+allometry across individuals are likely influenced by a combination of
+species identity and environmental factors. Furthermore, substantial
 measurement errors in tree height and crown radius relative to DBH may
@@
 to estimate AGB are small to medium-sized (DBH \textless{} 70 cm
-(\protect\hyperlink{ref-Slik2013}{Slik et al. 2013})), whose H-DBH
-relationships can still be well captured by the power-law function
-(Fig.~\ref{fig-compare} a) (e.g.,
-\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). In contrast,
-individuals with DBH ≥ 70 cm, while typically forming a small proportion
-of the tree population, contribute significantly to biomass storage
-(e.g., \protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015}). For
-instance, Slik et al. (\protect\hyperlink{ref-Slik2013}{2013}) reported
-that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the dataset in
-tropical forests, yet stored 25-45\% of the total AGB on average.
-However, the fundamental assumption of the power-law function does not
-fully reflect the asymptotic nature of height-diameter allometries in
-large trees (Fig.~\ref{fig-compare} a), potentially leading to
-significant inaccuracies in stand-level biomass estimates (e.g.,
-\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
-\protect\hyperlink{ref-Fayolle2016}{Fayolle et al. 2016};
-\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018};
-\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019}).
-Given the disproportionate contribution of large trees in biomass
-dynamics (e.g., \protect\hyperlink{ref-Slik2013}{Slik et al. 2013};
-\protect\hyperlink{ref-Fauset2015}{Fauset et al. 2015};
-\protect\hyperlink{ref-Ligot2018}{Ligot et al. 2018}), when heights are
-not directly measured, height estimates should be derived by fitting
+(\citeproc{ref-Slik2013}{Slik et al. 2013})), whose H-DBH relationships
+can still be well captured by the power-law function
+(Fig.~\ref{fig-compare} a) (e.g., \citeproc{ref-Molto2014}{Molto et al.
+2014}). In contrast, individuals with DBH ≥ 70 cm, while typically
+forming a small proportion of the tree population, contribute
+significantly to biomass storage (e.g., \citeproc{ref-Fauset2015}{Fauset
+et al. 2015}). For instance, Slik et al. (\citeproc{ref-Slik2013}{2013})
+reported that large trees (DBH ≥ 70 cm) comprised only 2-4\% of the
+dataset in tropical forests, yet stored 25-45\% of the total AGB on
+average. However, the fundamental assumption of the power-law function
+does not fully reflect the asymptotic nature of height-diameter
+allometries in large trees (Fig.~\ref{fig-compare} a), potentially
+leading to significant inaccuracies in stand-level biomass estimates
+(e.g., \citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
+\citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
+\citeproc{ref-Sullivan2018}{Sullivan et al. 2018};
+\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019}). Given the
+disproportionate contribution of large trees in biomass dynamics (e.g.,
+\citeproc{ref-Slik2013}{Slik et al. 2013};
+\citeproc{ref-Fauset2015}{Fauset et al. 2015};
+\citeproc{ref-Ligot2018}{Ligot et al. 2018}), when heights are not
+directly measured, height estimates should be derived by fitting
 asymptotic models (e.g., Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}) to datasets
@@
 capture of the saturating component
-(\protect\hyperlink{ref-Sullivan2018}{Sullivan et al. 2018}).
+(\citeproc{ref-Sullivan2018}{Sullivan et al. 2018}).
 
@@
 leading to inaccurate DBH estimates (e.g.,
-\protect\hyperlink{ref-Filipescu2012}{Filipescu et al. 2012}).
-Additionally, most existing AGB estimation equations typically include
-DBH as a squared term in equations (e.g.,
-\protect\hyperlink{ref-Brown1989}{Brown et al. 1989};
-\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}), meaning that even
-minor errors in DBH predictions can significantly affect the accuracy of
-AGB estimates (\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).
+\citeproc{ref-Filipescu2012}{Filipescu et al. 2012}). Additionally, most
+existing AGB estimation equations typically include DBH as a squared
+term in equations (e.g., \citeproc{ref-Brown1989}{Brown et al. 1989};
+\citeproc{ref-Chave2014}{Chave et al. 2014}), meaning that even minor
+errors in DBH predictions can significantly affect the accuracy of AGB
+estimates (\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
 Interestingly, our results revealed that the \(CR,H\) model was more
@@
 (\emph{r} = 0.703), offering a more biologically realistic framework
-(e.g., \protect\hyperlink{ref-Dormann2013}{Dormann et al. 2013}). The
-observed importance of crown radius in improving AGB estimation aligns
-with prior studies (e.g., \protect\hyperlink{ref-Henry2010}{Henry et al.
-2010}; \protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014}),
-whereas our finding of strong relationships between AGB and crown
-dimensions is consistent with those reported by Jucker et al.
-(\protect\hyperlink{ref-Jucker2017}{2017}). While the interdependence of
-tree height and crown radius requires careful model refinement to
-address potential collinearity issues
-(\protect\hyperlink{ref-Dormann2013}{Dormann et al. 2013}), their strong
+(e.g., \citeproc{ref-Dormann2013}{Dormann et al. 2013}). The observed
+importance of crown radius in improving AGB estimation aligns with prior
+studies (e.g., \citeproc{ref-Henry2010}{Henry et al. 2010};
+\citeproc{ref-Goodman2014}{Goodman et al. 2014}), whereas our finding of
+strong relationships between AGB and crown dimensions is consistent with
+those reported by Jucker et al. (\citeproc{ref-Jucker2017}{2017}). While
+the interdependence of tree height and crown radius requires careful
+model refinement to address potential collinearity issues
+(\citeproc{ref-Dormann2013}{Dormann et al. 2013}), their strong
 relationship with AGB suggests the feasibility of directly estimating
 AGB from EO data, bypassing DBH prediction and simplifying the
-estimation process (\protect\hyperlink{ref-Jucker2017}{Jucker et al.
-2017}).
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+estimation process (\citeproc{ref-Jucker2017}{Jucker et al. 2017}).
+
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Aiba2009}{}}%
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{1}
+\bibitem[\citeproctext]{ref-Aiba2009}
 Aiba M, Nakashizuka T (2009) Architectural differences associated with
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Poorter2012a}{}}%
+\bibitem[\citeproctext]{ref-Poorter2012a}
 Poorter H, Niklas KJ, Reich PB, et al (2012) Biomass allocation to

### 2025-03-26 _targets.R
clean targets

@@ -128,29 +128,6 @@ data_ <- list(
     format = "file"
   ),
   tar_target(
-    bb_allometry_csv,
-    "data-raw/bb_allometry.csv",
-    format = "file"
-  ),
-  tar_target(
-    bb_sp_list_csv,
-    "data-raw/bb_sp_list.csv",
-    format = "file"
-  ),
-  tar_target(
-    abund_xtbg_csv,
-    "data-raw/abund_xtbg.csv",
-    format = "file"
-  ),
-  tar_target(
-    bb_allometry_re_csv,
-    sp_name_match(
-      bb_sp_list_csv, abund_xtbg_csv, bb_allometry_csv,
-      "data/bb_allometry_re.csv"
-    ),
-    format = "file"
-  ),
-  tar_target(
     try_wd_txt,
     "data-raw/30953.txt",
     format = "file"

### 2025-03-27 ms/response_letter.tex

@@
 \author{}
-\date{2025-03-25}
+\date{2025-03-26}
 
@@
 significance of our findings.
+
+We have also updated the data availability statement. The programming
+code of this study will be deposited at Zenodo (https://zenodo.org) and
+made available on Github
+(\url{https://github.com/duyenecology/tree-allo}).
 
@@
 We have revised the manuscript for language, grammar, and sentence
-structure to enhance clarity and improve overall readability.
+structure to enhance clarity and improve overall readability. For
+example, we revised lines XXX--YYY in the Discussion section to ensure
+sentence brevity and clarity. Additionally, we updated the figure
+legends and table captions for greater clarity and carefully
+cross-checked all technical terms to maintain consistency throughout the
+text.
 

### 2025-03-27 ms/ms.qmd

@@
 
-Our analysis combined tree allometry data from Tallo, which includes standardized measurements of DBH, height, and crown radius for 5,163 species across various biomes [@Jucker2022, https://doi.org/10.5281/zenodo.6637599] along with wood density data for 10,995 species from TRY version 6 [@Kattge2020, https://www.try-db.org/TryWeb/Home.php] (see Supplementary Table S1 for further details).
+Our analysis combined tree allometry data from Tallo, which includes standardized measurements of DBH, height, and crown radius for 5,163 species across various biomes [@Jucker2022], with wood density data for 10,995 species from TRY version 6 [@Kattge2020] (see Supplementary Table S1 for further details).
+Detailed information on data access and availability can be found in the Data availability section.
 We thoroughly processed the Tallo dataset by removing 998 outliers following @Jucker2022.
@@
 Statistical analyses were performed in R version 4.3.2 [@RCoreTeam2023], using the R packages *targets* version 1.7.0 for workflow management [@Landau2021].
-Codes are made available on GitHub at <https://github.com/duyenecology/tree-allometry> (Stan code is provided in Supplementary Section S2).
+Codes are made available on Zenodo (https://zenodo.org) and GitHub (<https://github.com/duyenecology/tree-allo>).
+Stan code is provided in Supplementary Section S2.
 
@@
 \newpage
+
 # Acknowledgments {.unnumbered}
@@
 
-Tallo and TRY data are available from their respective platforms.
-Codes are available from Zenodo at xxx
+The tree allometry data is avaiable from tallo [@Jucker2022, https://doi.org/10.5281/zenodo.6637599], and
+the wood density data are available on request from TRY version 6  [@Kattge2020, <https://www.try-db.org/TryWeb/Home.php>].
+All codes and scripts used for analysis are available on Zenodo (https://zenodo.org) and GitHub (<https://github.com/duyenecology/tree-allo>).
 

### 2025-03-27 ms/ms.tex

@@
 species across various biomes (\citeproc{ref-Jucker2022}{Jucker et al.
-2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
-data for 10,995 species from TRY version 6
-(\citeproc{ref-Kattge2020}{Kattge et al. 2020},
-https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
-further details). We thoroughly processed the Tallo dataset by removing
-998 outliers following Jucker et al. (\citeproc{ref-Jucker2022}{2022}).
-To ensure consistency, we selected only species with standardized wood
-density values in g cm\textsuperscript{-3} from TRY and removed any
-records with missing data. Subsequently, we created an overlapping
-dataset between Tallo and TRY, resulting in a combined dataset
-encompassing allometric and wood density information for 355,881
-individuals across 2,961 species.
+2022}), with wood density data for 10,995 species from TRY version 6
+(\citeproc{ref-Kattge2020}{Kattge et al. 2020}) (see Supplementary Table
+S1 for further details). Detailed information on data access and
+availability can be found in the Data availability section. We
+thoroughly processed the Tallo dataset by removing 998 outliers
+following Jucker et al. (\citeproc{ref-Jucker2022}{2022}). To ensure
+consistency, we selected only species with standardized wood density
+values in g cm\textsuperscript{-3} from TRY and removed any records with
+missing data. Subsequently, we created an overlapping dataset between
+Tallo and TRY, resulting in a combined dataset encompassing allometric
+and wood density information for 355,881 individuals across 2,961
+species.
 
@@
 (\citeproc{ref-Landau2021}{Landau 2021}). Codes are made available on
-GitHub at \url{https://github.com/duyenecology/tree-allometry} (Stan
-code is provided in Supplementary Section S2).
+Zenodo (https://zenodo.org) and GitHub
+(\url{https://github.com/duyenecology/tree-allo}). Stan code is provided
+in Supplementary Section S2.
 
@@
 
-Tallo and TRY data are available from their respective platforms. Codes
-are available from Zenodo at xxx
+The tree allometry data is avaiable from tallo
+(\citeproc{ref-Jucker2022}{Jucker et al. 2022},
+https://doi.org/10.5281/zenodo.6637599), and the wood density data are
+available on request from TRY version 6
+(\citeproc{ref-Kattge2020}{Kattge et al. 2020},
+\url{https://www.try-db.org/TryWeb/Home.php}). All codes and scripts
+used for analysis are available on Zenodo (https://zenodo.org) and
+GitHub (\url{https://github.com/duyenecology/tree-allo}).
 

### 2025-03-27 sup/sup_old.qmd

--- 2025-03-26 sup/sup.qmd
@@
         \usepackage{setspace}
+        \doublespacing
+        \linenumbers
         \renewcommand{\thetable}{S\arabic{table}}
@@
 
-**Nguyen and Katabuchi, Saturating allometric relationships reveal how wood density shapes global tree architecture**
+**Saturating allometric relationships reveal how wood density shapes global tree architecture**
 
@@
 
+\newpage
+
 This supplement includes the following materials:
@@
 
-**Table S2.** Characteristics of the sub-datasets used to select the best predictive model for each allometric relationship in angiosperms and gymnosperms.
+**Table S2.** Characteristics of sub-datasets were used to select the best predictive model of each allometric relationship for angiosperms and gymnosperms.
 
@@
 
-**Table S6.** Parameter estimates for predicting tree height (m) from DBH (cm) of 1290 species (see file *TableS6_height_sp_estimates.xlsx*).
+**Table S6.** Parameter estimates for predicting tree height (m) from DBH (cm) of 1290 species (see file TableS6_height_sp_estimates.xlsx).
 
-**Table S7.** Parameter estimates for predicting crown radius (m) from DBH (cm) of 821 species (see file *TableS7_crown_radius_sp_estimates.xlsx*).
+**Table S7.** Parameter estimates for predicting crown radius (m) from DBH (cm) of 821 species (see file TableS7_crown_radius_sp_estimates.xlsx).
 
-**Table S8.** Parameter estimates for predicting DBH (cm) from crown radius (m) and tree height (m) of 800 species (see file *TableS8_dbh_sp_estimates.xlsx*).
+**Table S8.** Parameter estimates for predicting DBH (cm) from crown radius (m) and tree height (m) of 800 species (see file TableS8_dbh_sp_estimates.xlsx).
 
-**Figure S1.** Functional curves describing allometric relationships of tree dimensions with DBH on the log scale.
+**Figure S1.** Function curves are used to describe allometric relationships of tree dimensions with DBH in log scale.
 
@@
 
-**Table S2**: Characteristics of the sub-datasets used to select the best predictive model for each allometric relationship in angiosperms and gymnosperms.
-The ranges represent the minimum and maximum values for the respective variables: DBH (cm), tree height (H, m), crown radius (CR, m), and wood density (WD, g cm^-3^).
+**Table S2**: Characteristics of sub-datasets used to select the best predictive model of each allometric relationship for angiosperms and gymnosperms.
+The ranges represent the minimum and maximum values for the respective variables: DBH (cm), tree height (H, m), crown radius (CR, m), and wood density (WD, g cm$^{-3}$).
 
@@
 ```{r echo=FALSE, results='asis'}
-posterior_df <- tar_read(posterior_df)
-
-posterior_df <- posterior_df |>
-  mutate(
-    Intercept_CI = sapply(Intercept_CI, format_ci),
-    Slope_CI = sapply(Slope_CI, format_ci),
-    Tau_CI = sapply(Tau_CI, format_ci)
-  )
-
-df_h <- posterior_df |>
+posterior_rounded_df <- tar_read(posterior_rounded_df)
+df_h <- posterior_rounded_df |>
   filter(Dependent_variable == "Tree Height") |>
@@
 ```{r echo=FALSE, results='asis'}
-posterior_df <- tar_read(posterior_df)
-
-posterior_df <- posterior_df |>
-  mutate(
-    Intercept_CI = sapply(Intercept_CI, format_ci),
-    Slope_CI = sapply(Slope_CI, format_ci),
-    Tau_CI = sapply(Tau_CI, format_ci)
-  )
-
-df_cr <- posterior_df |>
+posterior_rounded_df <- tar_read(posterior_rounded_df)
+df_cr <- posterior_rounded_df |>
   filter(Dependent_variable == "Crown Radius") |>
@@
 ```{r echo=FALSE, results='asis'}
-posterior_df <- tar_read(posterior_df)
+posterior_rounded_df <- tar_read(posterior_rounded_df)
 
-posterior_df <- posterior_df |>
-  mutate(
-    Intercept_CI = sapply(Intercept_CI, format_ci),
-    Slope_CI = sapply(Slope_CI, format_ci),
-    Tau_CI = sapply(Tau_CI, format_ci)
-  )
-df_dbh <- posterior_df |>
+df_dbh <- posterior_rounded_df |>
   filter(Dependent_variable %in% c("DBH", "DBH1", "DBH2", "DBH3")) |>
@@
 ![
-Allometric relationships between tree height and crown radius versus DBH for angiosperms (a, c) and gymnosperms (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
-The density of overlapping points is represented using a colour gradient ranging from black (low density) to yellow (high density).
-Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, which estimate the concentration of data points in the log-transformed DBH and tree height or crown radius.
-All axes are shown on the log scale.
+Allometric relationships between tree height and crown radius with DBH of angiosperm (a, c) and gymnosperm (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
+The density of overlapping points is shown using a colour gradient ranging from black (low density) to yellow (high density).
+Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, estimating the concentration of data points in the log-transformed space of DBH and tree height or crown radius.
+All axes are on the log scale.
 ](../figs/h_cr_dbh_log.png){#fig-compare fig-pos="H"}

### 2025-03-27 ms/response_letter.qmd
updated response letter

@@
 author: ""
-date: "`r format(Sys.time(), '%B %d, %Y')`"
 fontsize: 12pt
@@
 ```
+
+March xxx, 2025
 
@@
 Thank you for your suggestion.
-We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line XXX).
+We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line 117-119).
 
@@
 
-We have standardized the presentation of all parameter estimates throughout the manuscript, including Table 2 and Table S6-8 and corresponding text.
+We have standardized the presentation of all parameter estimates throughout the manuscript, including Table 2 and Tables S6-8 and corresponding text.
 
@@
 
-We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line XXX).
+We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 262-264).
 
@@
 
-We have updated the references with recent publications, including xxx, and formatted them according to the journal guidelines.
+We have updated the references with recent publications, including Tavankar et al.
+(2019), Murphy and Bradford (2022), Fajardo et al.
+(2022), Cochrane et al.
+(2023), Sivula et al.
+(2023), and formatted them according to the journal guidelines.
 
@@
 We have revised the manuscript for language, grammar, and sentence structure to enhance clarity and improve overall readability.
-For example, we revised lines XXX–YYY in the Discussion section to ensure sentence brevity and clarity.
+For example, we revised lines 481–486 in the Discussion section to ensure sentence brevity and clarity.
 Additionally, we updated the figure legends and table captions for greater clarity and carefully cross-checked all technical terms to maintain consistency throughout the text.
@@
 Thank you for pointing this out.
-We have added a description in the Methods section clarifying the classification criteria (line XXX).
+We have added a description in the Methods section clarifying the classification criteria (line 197-202).
 
@@
 
-We have corrected the formatting issue in Figure 4.
+We were unable to see the absence of the letters "a" and "d" in Figure 4 of the submitted manuscript, so we are unsure how to address this issue.
+Nevertheless, we slightly increased the left margin of the panel labels in case clipping was the cause.
 
@@
 We agree with your point and have expanded the Discussion section to address potential environmental confounding effects.
-We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line XXX).
+We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line 437-444).
 
@@
 
-We have resolved the formatting issue that caused the blank pages in the supplementary file.
+We have resolved the formatting issue of Tables S6–S8 that may have caused the blank pages in the supplementary file.
+These tables, which were submitted as Excel files, were included in the PDF due to a system-related submission setting.

### 2025-03-27 ms/response_letter.tex
updated response letter

--- 2025-03-26 ms/response_letter.tex
@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatletter
-\newsavebox\pandoc@box
-\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
-  \sbox\pandoc@box{#1}%
-  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
-  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
-  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
-  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
-  \else\usebox{\pandoc@box}%
-  \fi%
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 % Set default figure placement to htbp
+\makeatletter
 \def\fps@figure{htbp}
@@
 \KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-
-\usepackage{bookmark}
-
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \author{}
-\date{2025-03-25}
+\date{}
 
 \begin{document}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+
+March xxx, 2025
 
@@
 
-\section{Responses}\label{responses}
+\hypertarget{responses}{%
+\section{Responses}\label{responses}}
 
@@
 clearly state the main objective of the study before presenting the
-specific questions (line XXX).
+specific questions (line 117-119).
 
@@
 We have standardized the presentation of all parameter estimates
-throughout the manuscript, including Table 2 and Table S6-8 and
+throughout the manuscript, including Table 2 and Tables S6-8 and
 corresponding text.
@@
 greater than 4 suggests a meaningful difference in predictive
-performance (line XXX).
+performance (line 262-264).
 
@@
 
-We have updated the references with recent publications, including xxx,
-and formatted them according to the journal guidelines.
+We have updated the references with recent publications, including
+Tavankar et al.~(2019), Murphy and Bradford (2022), Fajardo et
+al.~(2022), Cochrane et al.~(2023), Sivula et al.~(2023), and formatted
+them according to the journal guidelines.
 
@@
 We have revised the manuscript for language, grammar, and sentence
-structure to enhance clarity and improve overall readability.
+structure to enhance clarity and improve overall readability. For
+example, we revised lines 481--486 in the Discussion section to ensure
+sentence brevity and clarity. Additionally, we updated the figure
+legends and table captions for greater clarity and carefully
+cross-checked all technical terms to maintain consistency throughout the
+text.
 
@@
 Thank you for pointing this out. We have added a description in the
-Methods section clarifying the classification criteria (line XXX).
+Methods section clarifying the classification criteria (line 197-202).
 
@@
 
-We have corrected the formatting issue in Figure 4.
+We were unable to see the absence of the letters ``a'' and ``d'' in
+Figure 4 of the submitted manuscript, so we are unsure how to address
+this issue. Nevertheless, we slightly increased the left margin of the
+panel labels in case clipping was the cause.
 
@@
 Fajardo et al.~(2022) to highlight less clear linkages between tree
-architecture and intraspecific variation in wood density (line XXX).
+architecture and intraspecific variation in wood density (line 437-444).
 
@@
 
-We have resolved the formatting issue that caused the blank pages in the
-supplementary file.
+We have resolved the formatting issue of Tables S6--S8 that may have
+caused the blank pages in the supplementary file. These tables, which
+were submitted as Excel files, were included in the PDF due to a
+system-related submission setting.
 

### 2025-03-27 ms/ms.qmd
typo

--- 2025-03-26 ms/ms.qmd
@@
 MK was supported by
-the Xingdian Talenj Support Program of Yunnan Province,
+the Xingdian Talent Support Program of Yunnan Province,
 the Xishuangbanna State Rainforest Talent Support Program (E4BN041B01), and

### 2025-03-27 ms/ms.tex
typo

--- 2025-03-26 ms/ms.tex
@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-MK was supported by the Xingdian Talenj Support Program of Yunnan
+MK was supported by the Xingdian Talent Support Program of Yunnan
 Province, the Xishuangbanna State Rainforest Talent Support Program
@@
 Initiative (2020FYB0003).
+
+\hypertarget{data-availability}{%
+\section*{Data availability}\label{data-availability}}
+\addcontentsline{toc}{section}{Data availability}
+
+Tallo and TRY data are available from their respective platforms. Codes
+are available from Zenodo at xxx
 

### 2025-03-27 ms/response_letter.tex

@@
 \date{}
+\date{2025-03-26}
 
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 

### 2025-03-27 ms/response_letter.qmd
updated response letter

@@
 
+We have also updated the data availability statement.
+The programming code of this study will be deposited at Zenodo (https://zenodo.org) and made available on Github (<https://github.com/duyenecology/tree-allo>).
+
 Below we provide detailed responses.
@@
 
-We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 262-264).
+We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 263-265).
 
@@
 We have revised the manuscript for language, grammar, and sentence structure to enhance clarity and improve overall readability.
-For example, we revised lines 481–486 in the Discussion section to ensure sentence brevity and clarity.
+For example, we revised lines 482–487 in the Discussion section to ensure sentence brevity and clarity.
 Additionally, we updated the figure legends and table captions for greater clarity and carefully cross-checked all technical terms to maintain consistency throughout the text.

### 2025-03-27 ms/response_letter.tex
updated response letter

@@
 \date{}
-\date{2025-03-26}
 
 \begin{document}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 greater than 4 suggests a meaningful difference in predictive
-performance (line 262-264).
+performance (line 263-265).
 
@@
 structure to enhance clarity and improve overall readability. For
-example, we revised lines 481--486 in the Discussion section to ensure
+example, we revised lines 482--487 in the Discussion section to ensure
 sentence brevity and clarity. Additionally, we updated the figure

### 2025-03-27 ms/response_letter.qmd

@@
 
-Please note that line numbers refer to ms_diff.pdf that contains editing history.
+Please note that line numbers refer to **ms_diff.pdf** that contains editing history.
 The word file has slightly different line numbers.
@@
 
-We were unable to see the absence of the letters "a" and "d" in Figure 4 of the submitted manuscript, so we are unsure how to address this issue.
+We were unable to see the absence of the letters "a" and "d" in Figure 4 of the generated PDF file.
 Nevertheless, we slightly increased the left margin of the panel labels in case clipping was the cause.
+We hope this adjustment addresses the concern and would be happy to make further modifications if needed.
 
@@
 
-We have resolved the formatting issue of Tables S6–S8 that may have caused the blank pages in the supplementary file.
-These tables, which were submitted as Excel files, were included in the PDF due to a system-related submission setting.
+We have reduced the number of significant digits in the tables to prevent column overflow and ensure a complete display of cell information, which should resolve the formatting issue.
+These tables were submitted as Excel files, but due to a system-related submission setting, they are still included in the PDF.

### 2025-03-27 ms/response_letter.tex

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \begin{document}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{responses}{%
-\section{Responses}\label{responses}}
+\section{Responses}\label{responses}
 
-Please note that line numbers refer to ms\_diff.pdf that contains
-editing history. The word file has slightly different line numbers.
+Please note that line numbers refer to \textbf{ms\_diff.pdf} that
+contains editing history. The word file has slightly different line
+numbers.
 
@@
 We were unable to see the absence of the letters ``a'' and ``d'' in
-Figure 4 of the submitted manuscript, so we are unsure how to address
-this issue. Nevertheless, we slightly increased the left margin of the
-panel labels in case clipping was the cause.
+Figure 4 of the generated PDF file. Nevertheless, we slightly increased
+the left margin of the panel labels in case clipping was the cause. We
+hope this adjustment addresses the concern and would be happy to make
+further modifications if needed.
 
@@
 
-We have resolved the formatting issue of Tables S6--S8 that may have
-caused the blank pages in the supplementary file. These tables, which
-were submitted as Excel files, were included in the PDF due to a
-system-related submission setting.
+We have reduced the number of significant digits in the tables to
+prevent column overflow and ensure a complete display of cell
+information, which should resolve the formatting issue. These tables
+were submitted as Excel files, but due to a system-related submission
+setting, they are still included in the PDF.
 

### 2025-03-27 sup/sup.qmd
Update sup.qmd and response letter

@@
 ![
-Allometric relationships between tree height and crown radius versus DBH for angiosperms (a, c) and gymnosperms (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
-The density of overlapping points is represented using a colour gradient ranging from black (low density) to yellow (high density).
-Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, which estimate the concentration of data points in the log-transformed DBH and tree height or crown radius.
+Allometric relationships between DBH and tree height or crown radius in angiosperms (a, c) and gymnosperms (b, d), based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
+A color gradient from black (low density) to yellow (high density) indicates the density of overlapping data points, calculated via 2D kernel density estimation on a 300 $\times$ 300 grid.
 All axes are shown on the log scale.

### 2025-03-28 sup/sup.tex

--- 2025-03-26 sup/sup.tex
@@
 
-\caption{\label{fig-compare}Allometric relationships between tree height
-and crown radius versus DBH for angiosperms (a, c) and gymnosperms (b,
-d) based on three different models: power-law (orange), generalized
-Michaelis-Menten (green), and Weibull (blue). The density of overlapping
-points is represented using a colour gradient ranging from black (low
-density) to yellow (high density). Point density was calculated using a
-2D kernel density estimation on a 300 \(\times\) 300 grid, which
-estimate the concentration of data points in the log-transformed DBH and
-tree height or crown radius. All axes are shown on the log scale.}
+\caption{\label{fig-compare}Allometric relationships between DBH and
+tree height or crown radius in angiosperms (a, c) and gymnosperms (b,
+d), based on three different models: power-law (orange), generalized
+Michaelis-Menten (green), and Weibull (blue). A color gradient from
+black (low density) to yellow (high density) indicates the density of
+overlapping data points, calculated via 2D kernel density estimation on
+a 300\,\(\times\)\,300 grid. All axes are shown on the log scale.}
 

### 2025-03-28 ms/response_letter.qmd
updated response letter

@@
 
-March xxx, 2025
+March 28, 2025
 
@@
 Thank you for your suggestion.
-We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line 117-119).
+We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line 105-107).
 
@@
 
-We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 263-265).
+We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 238-239).
 
@@
 We have revised the manuscript for language, grammar, and sentence structure to enhance clarity and improve overall readability.
-For example, we revised lines 482–487 in the Discussion section to ensure sentence brevity and clarity.
+For example, we revised lines 432-438, 448-452 in the Discussion section to ensure sentence brevity and clarity.
 Additionally, we updated the figure legends and table captions for greater clarity and carefully cross-checked all technical terms to maintain consistency throughout the text.
@@
 Thank you for pointing this out.
-We have added a description in the Methods section clarifying the classification criteria (line 197-202).
+We have added a description in the Methods section clarifying the classification criteria (line 178-183).
 
@@
 We agree with your point and have expanded the Discussion section to address potential environmental confounding effects.
-We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line 437-444).
+We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line 392-398).
 

### 2025-03-28 ms/response_letter.tex
Update sup.qmd and response letter
updated response letter

@@
 clearly state the main objective of the study before presenting the
-specific questions (line 117-119).
+specific questions (line 105-107).
 
@@
 greater than 4 suggests a meaningful difference in predictive
-performance (line 263-265).
+performance (line 238-239).
 
@@
 structure to enhance clarity and improve overall readability. For
-example, we revised lines 482--487 in the Discussion section to ensure
-sentence brevity and clarity. Additionally, we updated the figure
+example, we revised lines 432-438, 448-452 in the Discussion section to
+ensure sentence brevity and clarity. Additionally, we updated the figure
 legends and table captions for greater clarity and carefully
@@
 Thank you for pointing this out. We have added a description in the
-Methods section clarifying the classification criteria (line 197-202).
+Methods section clarifying the classification criteria (line 178-183).
 
@@
 Fajardo et al.~(2022) to highlight less clear linkages between tree
-architecture and intraspecific variation in wood density (line 437-444).
+architecture and intraspecific variation in wood density (line 392-398).
 

### 2025-07-03 R/figs.R
updated figures

@@ -32,27 +32,53 @@ my_ggsave <- function(filename, plot, units = "mm", height = NA, width = NA, dpi
     dpi = dpi,
     ...
   )
+
+    # Save as JPG
+  ggsave(
+    filename = paste0(filename, ".jpg"),
+    plot = plot,
+    height = height,
+    width = width,
+    units = units,
+    dpi = dpi,
+    ...
+  )
   # Return the paths to the saved files
   # return(paste0(filename, c(".tiff", ".pdf", ".png")))
-  return(paste0(filename, c(".pdf", ".png")))
+  return(paste0(filename, c(".pdf", ".png", ".jpg")))
 
 }
 
+
+# my_theme <- function(){
+#   theme_bw() %+replace%
+#   theme(
+#     plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"),
+#     legend.key.size = unit(0.3, "cm"),
+#     legend.spacing.y = unit(0.15, "cm"),
+#     legend.text.align = 0,
+#     legend.text = element_text(size = 8),
+#     legend.title = element_text(size = 9),
+#     legend.background = element_blank()
+#   )
+# }
 
 my_theme <- function(){
-  theme_bw() %+replace%
-  theme(
-    plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"),
-    legend.key.size = unit(0.3, "cm"),
-    legend.spacing.y = unit(0.15, "cm"),
-    legend.text.align = 0,
-    legend.text = element_text(size = 8),
-    legend.title = element_text(size = 9),
-    legend.background = element_blank()
-  )
+  theme_bw(base_size = 8, base_family = "serif") %+replace%
+    theme(
+      plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"),
+      legend.key.size = unit(0.3, "cm"),
+      legend.spacing.y = unit(0.15, "cm"),
+      legend.text.align = 0,
+      legend.text = element_text(size = 8, family = "serif"),
+      legend.title = element_text(size = 8, family = "serif"),
+      axis.title = element_text(size = 8, family = "serif"),
+      axis.text = element_text(size = 8, family = "serif"),
+      strip.text = element_text(size = 8, family = "serif"),
+      plot.title = element_text(size = 8, face = "bold", family = "serif"),
+      legend.background = element_blank()
+    )
 }
-
-
 
[…]

### 2025-07-03 R/functions.R

@@ -2410,13 +2410,22 @@ subset_charac <- function(tallo_reduced_lr_df_ang_h,
     group_by(sp) |> filter(n() >= 20) |>
     ungroup()
 
+  # ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
+  #   filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
+  #   group_by(sp) |> filter(n() >= 20) |>
+  #   ungroup()
   ang_data_dbh1 <- tallo_reduced_lr_df_ang_dbh1 |>
-    filter(division == "Angiosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
+    filter(division == "Angiosperm", !is.na(dbh), !is.na(cr*h)) |>
     group_by(sp) |> filter(n() >= 20) |>
     ungroup()
 
+  # gym_data_dbh1 <- tallo_reduced_lr_df_gym_dbh1 |>
+  #   filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
+  #   group_by(sp) |> filter(n() >= 20) |>
+  #   ungroup()
+
   gym_data_dbh1 <- tallo_reduced_lr_df_gym_dbh1 |>
-    filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr), !is.na(h)) |>
+    filter(division == "Gymnosperm", !is.na(dbh), !is.na(cr*h)) |>
     group_by(sp) |> filter(n() >= 20) |>
     ungroup()
 

### 2025-07-03 _targets.R

@@ -1329,12 +1329,48 @@ main_ <- list(
         sp_posterior_h_df
       )
       my_ggsave(
-        plot = p,
+        plot     = p,
         filename = "figs/h_sp",
-        dpi = 600,
-        width = 200,
-        height = 300,
-        units = "mm"
+        dpi      = 600,
+        width    = 200,
+        height   = 300,
+        units    = "mm"
+      )
+    },
+    format = "file"
+  ),
+# ANG
+  tar_target(
+    h_sp_ang_plot, {
+      p <- generate_h_sp_ang_plot(
+        tallo_reduced_lr_df_ang_h,
+        sp_posterior_h_df
+      )
+      my_ggsave(
+        plot     = p,
+        filename = "figs/h_sp_ang",
+        dpi      = 600,
+        width    = 200,
+        height   = 300,
+        units    = "mm"
+      )
+    },
+    format = "file"
+  ),
+#GYM
+  tar_target(
+    h_sp_gym_plot, {
+      p <- generate_h_sp_gym_plot(
+        tallo_reduced_nlr_df_gym_h,
+        sp_posterior_h_df
+      )
+      my_ggsave(
+        plot     = p,
+        filename = "figs/h_sp_gym",
+        dpi      = 600,
+        width    = 200,
+        height   = 300,
+        units    = "mm"
       )
     },
     format = "file"
@@ -1357,6 +1393,43 @@ main_ <- list(
     },
     format = "file"
   ),
+# SPECIES-LEVEL CR-ANG
+  tar_target(
+    cr_sp_ang_plot, {
+      p <- generate_cr_sp_ang_plot(
+        tallo_reduced_lr_df_ang_cr,
+        sp_posterior_cr_df
+      )
+      my_ggsave(
+        plot     = p,
+        filename = "figs/cr_sp_ang",
+        dpi      = 600,
+        width    = 200,
+        height   = 300,
+        units    = "mm"
+      )
+    },
+    format = "file"
+  ),
+# SPECIES-LEVEL CR-GYM
+  tar_target(
+    cr_sp_gym_plot, {
+      p <- generate_cr_sp_gym_plot(
+        tallo_reduced_nlr_df_gym_cr,
+        sp_posterior_cr_df
+      )
+      my_ggsave(
+        plot     = p,
+        filename = "figs/cr_sp_gym",
+        dpi      = 600,
+        width    = 200,
+        height   = 300,
+        units    = "mm"
+      )
+    },
+    format = "file"
+  ),
+
 NULL
 )
 

### 2025-07-03 ms/ms.qmd
updated ms

@@
 
-^1^ CAS Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China
+^1^ Yunnan Key Laboratory of Forest Ecosystem Stability and Global Change Response, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China
 
@@
 
-where $f(\cdot)$ predicts the expected natural log tree height or crown radius from observed DBH (and *vice versa*), $\pmb{\beta}_{j}=\{a_j, b_j, k_j\}$ denotes the vector of species-specific parameters, and $\sigma_y$ represents the standard deviation of the observation error.
+where $f(\cdot)$ predicts the expected natural log tree height or crown radius from observed DBH (and *vice versa*), $\pmb{\beta}_{j}=\{a_j, b_j, c_j, k_j\}$ denotes the vector of species-specific parameters, and $\sigma_y$ represents the standard deviation of the observation error.
 We considered three functional forms for the relationship between tree allometric variables [@Sullivan2018; @MartinezCano2019]:
@@
 In contrast, the gMM and Weibull models were only used to estimate tree height and crown radius from DBH.
-Because these two models specifically describe saturating relationships (@eq-gMM; @eq-wb), their formulation does not allow simply switching the roles of explanatory and response variables.
+Because these two models specifically describe saturating relationships (@eq-gMM; @eq-wb), their formulation does not allow for simply switching the roles of explanatory and response variables.
 
 To compare the performance of different models for the same response variable, we used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2017).
-PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), which is a robust method for comparing models with different numbers of parameters [@Vehtari2017].
+PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), a robust method for comparing models with different numbers of parameters [@Vehtari2017].
 We utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
-A model with lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].
+A model with a lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].
 
@@
 
-where $\gamma_{0}$ and $\gamma_{1}$ are the species-level intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
+where $\gamma_{0}$ and $\gamma_{1}$ are the global-level hyperparameters for intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
 To facilitate interpretation and visualization, species were then classified into "light" and "dense" wood groups based on the 10th and 90th percentiles of wood density distribution in the observed dataset, respectively.
@@
 We also reported the median and 95\% Bayesian credible intervals (CIs) for each parameter (Supplementary Tables S3-S8).
-Estimates were considered significant when 95\% CIs did not overlap with zero.
+Estimates were considered significant when their 95\% CIs did not overlap with zero.
 
@@
 In CR-DBH allometry, angiosperms exhibited a positive relationship between the scaling factor $a$ and wood density and a negative association between the exponent parameter  $b$ and wood density (@fig-wood1 g, h; @fig-wood2 b).
-In gymnosperms, we found a positive association between wood density and the exponent parameter $b$, whereas neither the asymptote parameter $a$ nor the scaling constant $k$ showed significant relationship with this functional trait (@fig-wood1 i, j, k; @fig-wood2 c).
+In gymnosperms, we found a positive association between wood density and the exponent parameter $b$, whereas neither the asymptote parameter $a$ nor the scaling constant $k$ showed a significant relationship with this functional trait (@fig-wood1 i, j, k; @fig-wood2 c).
 
@@
 
-AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based both on the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
+AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based on both the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
 Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_pl_h)` Mg and bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_pl_h * 100)`% (@fig-agb b).
@@
 
-Our result supported a saturating H-DBH relationship in both angiosperm and gymnosperm species, consistent with theoretical expectations [@Falster2003; @Niklas2007] and empirical observations [e.g., @Feldpausch2011; @MartinezCano2019; @Song2024].
+Our results supported a saturating H-DBH relationship in both angiosperm and gymnosperm species, consistent with theoretical expectations [@Falster2003; @Niklas2007] and empirical observations [e.g., @Feldpausch2011; @MartinezCano2019; @Song2024].
 A recent study using Tallo suggested the Weibull function as the best model for six plant functional types (PFTs) [@Song2024].
@@
 In contrast, the CR-DBH allometry in gymnosperms showed an asymptotic relationship best described by the gMM function, with the Weibull function also showing competitive predictive performance (Table 1).
-This asymptotic pattern likely arises from structural constraints in gymnosperms growth forms that favour vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].
+This asymptotic pattern likely arises from structural constraints in gymnosperms' growth forms that favor vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].
 
@@
 
-Our analysis indicated distinct patterns in which wood density has a greater influence on the interspecific variation in angiosperms than gymnosperms.
+Our analysis indicated distinct patterns in which wood density has a greater influence on interspecific variation in angiosperms than in gymnosperms.
 Variation in wood density has consequential implications for the mechanical properties of wood and hence influences the mechanical stability of stems and branches [@Sterck1998], ultimately determining the architecture of trees [e.g., @Poorter2006].

### 2025-07-03 ms/ms.tex
updated ms

@@
 
-\textsuperscript{1} CAS Key Laboratory of Tropical Forest Ecology,
-Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences,
-Menglun, Yunnan 666303, China
+\textsuperscript{1} Yunnan Key Laboratory of Forest Ecosystem Stability
+and Global Change Response, Xishuangbanna Tropical Botanical Garden,
+Chinese Academy of Sciences, Menglun, Yunnan 666303, China
 
@@
 crown radius from observed DBH (and \emph{vice versa}),
-\(\pmb{\beta}_{j}=\{a_j, b_j, k_j\}\) denotes the vector of
+\(\pmb{\beta}_{j}=\{a_j, b_j, c_j, k_j\}\) denotes the vector of
 species-specific parameters, and \(\sigma_y\) represents the standard
@@
 saturating relationships (Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}), their
-formulation does not allow simply switching the roles of explanatory and
-response variables.
+formulation does not allow for simply switching the roles of explanatory
+and response variables.
 
@@
 reliable approximation to standard leave-one-out cross-validation (LOO),
-which is a robust method for comparing models with different numbers of
+a robust method for comparing models with different numbers of
 parameters (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}). We
 utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for
-each model. A model with lower LOOIC is considered to have better
+each model. A model with a lower LOOIC is considered to have better
 predictive accuracy (\citeproc{ref-Vehtari2017}{Vehtari et al. 2017}).
@@
 
-where \(\gamma_{0}\) and \(\gamma_{1}\) are the species-level intercept
-and slope, respectively, and \(\sigma_{\beta}\) is the standard
-deviation of the species-specific parameters. To facilitate
-interpretation and visualization, species were then classified into
-``light'' and ``dense'' wood groups based on the 10th and 90th
-percentiles of wood density distribution in the observed dataset,
-respectively. This classification allowed us to illustrate the
+where \(\gamma_{0}\) and \(\gamma_{1}\) are the global-level
+hyperparameters for intercept and slope, respectively, and
+\(\sigma_{\beta}\) is the standard deviation of the species-specific
+parameters. To facilitate interpretation and visualization, species were
+then classified into ``light'' and ``dense'' wood groups based on the
+10th and 90th percentiles of wood density distribution in the observed
+dataset, respectively. This classification allowed us to illustrate the
 relationships between wood density and the parameters of the
@@
 (Supplementary Tables S3-S8). Estimates were considered significant when
-95\% CIs did not overlap with zero.
+their 95\% CIs did not overlap with zero.
 
@@
 parameter \(b\), whereas neither the asymptote parameter \(a\) nor the
-scaling constant \(k\) showed significant relationship with this
+scaling constant \(k\) showed a significant relationship with this
 functional trait (Fig.~\ref{fig-wood1} i, j, k; Fig.~\ref{fig-wood2} c).
@@
 AGB estimates based on the Weibull model (\(\text{AGB}_\text{H-WB}\))
-showed the highest predictive accuracy based both on the lowest root
+showed the highest predictive accuracy based on both the lowest root
 mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.04\%
@@
 
-Our result supported a saturating H-DBH relationship in both angiosperm
+Our results supported a saturating H-DBH relationship in both angiosperm
 and gymnosperm species, consistent with theoretical expectations
@@
 showing competitive predictive performance (Table 1). This asymptotic
-pattern likely arises from structural constraints in gymnosperms growth
-forms that favour vertical over lateral expansion (e.g.,
+pattern likely arises from structural constraints in gymnosperms' growth
+forms that favor vertical over lateral expansion (e.g.,
 \citeproc{ref-Lines2012}{Lines et al. 2012};
@@
 Our analysis indicated distinct patterns in which wood density has a
-greater influence on the interspecific variation in angiosperms than
+greater influence on interspecific variation in angiosperms than in
 gymnosperms. Variation in wood density has consequential implications
